### 2016-05-04 Semenov_SCT_statya.tex
Наработки по состоянию на 4 мая 2016 года

\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel} %используем русский и английский языки с переносами
\usepackage{amssymb,amsfonts,amsmath,mathtext,cite,enumerate,float,amsthm} %подключаем нужные пакеты расширений
\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{indentfirst} % включить отступ у первого абзаца
\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
\graphicspath{{illustr/}}%путь к рисункам

\makeatletter
\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"

\usepackage{geometry} % Меняем поля страницы. 
\geometry{left=1cm}% левое поле
\geometry{right=1cm}% правое поле
\geometry{top=1cm}% верхнее поле
\geometry{bottom=2cm}% нижнее поле

\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра

\begin{document}\large

УДК ???

О множествах точек на плоскости с целочисленными расстояниями

Н.Н. Авдеев, Е.М. Семёнов.

Известна следующая

\paragraph{Теорема 1.} Пусть $\{M_1,M_2\}$ - счётное множество точек на плоскости и расстояние $|M_i,M_j|$ - натуральные для всех $1\leq i < j <\infty$, где $\mathbb{N} - $множество натуральных чисел. Тогда найдется такая прямая на плоскости $l$, что $M_i\in l$ для всех $i \in \mathbb{N}$.

Формулировка теоремы и идея её доказательства приведены в \cite{Newman}. Полное доказательство можно найти в \cite{angem1kurs}. Там же показан, что для любого $N\in \mathbb{N}$ существует такое множество $\{M_1,M_2,...,M_n\}\subset \mathbb{R}^2$, что $|M_i,M_j|$ для всех $1\leq i < j  \leq n$ и $M_1, M_2, ..., M_n$ не лежат на прямой. Изучению таких подмножеств посвящена настоящая работа.

Для заданного $n\in \mathbb{N}$ обозначим через $C_n$ множество таких последовательностей $M_1,M_2,...,M_n \in \mathbb{R}^2$, что $|M_i,M_j|\in\mathbb{N}$ ля всех $1\leq i < j  \leq n$ и  $M_1,M_2,...,M_n$ не принадлежат никакой прямой. Положим
$$
F(n)=\min\limits_{A\in C_n} d(A),
$$
где $d(A)$ --- диаметр $A$, т. е.
$$
d(A)=\max\limits_{x,y\in A}|x,y|
$$

Точную асимптотику последовательности  $F(n)$ найти не удалось, получены лишь верхняя и нижняя оценки.

Число элементов множества $A$ обозначим через $|A|$.
В \cite{angem1kurs} была доказана

\paragraph{Лемма 2.}

Пусть $A=(M_1, M_2, ..., M_n) \in C_n$ для некоторого $n\in N$ и $M_1, M_2, M_3$ не принадлежат прямой.
Тогда $n\leq (a+1)(b+1)+3$, где $a=|M_1,M_2|$, $|M_2,M_3|$.

Аналогичное утверждение справедливо, когда $M_1,M_2,M_3$ принадлежат некоторой прямой и $M_2$ лежит между $M_1$ и $M_3$:
$n\leq (a+1)(b+1)+3+d(A)$.

(Евгений Михайлович! Здесь вылезло $+d(A)$ в правой части.
В следующей лемме из-за этого может испортиться константа перед $m^2$.
Чтобы её сберечь, повысил $m$ в условии с 2 до 4.
Надо всё пересчитывать $=($ ~~ )

\paragraph{Лемма 3.}
Пусть $m\in\mathbb{N}, m\geq 4$, последовательность $(M_1,M_2,...,M_{2m^2+1})$ принадлежит $C_{2m^2+1}$ и содержится в квадрате со стороной $d$.
Тогда $d>\frac{1}{2}m^2$.

\paragraph{Доказательство.} Разобьём квадрат со стороной $d$ на $m^2$ квадратов со стороной $\frac{d}{m}$.
Тогда по крайней мере один из маленьких квадратов содержит некоторые три точки исходной последовательности.
Без ограничения общности $M_1,M_2,M_3$ содержатся в квадрате со стороной $\frac{d}{m}$.
Поэтому $|M_1,M_2|,|M_2,M_3| \leq \frac{d}{m}\sqrt{2}$.
В силу леммы 2
$$
2m^2+1 \leq\left(\frac{d}{m}\sqrt{2}+1\right)^2+3+d\sqrt{2}
$$

Отсюда

$$
d\geq\frac{1}{4} (-\sqrt(2) m^2+\sqrt(2) \sqrt(m^2 (9 m^2+4 m-8))-2 \sqrt(2) m)
$$

Так как для $m\geq 4$

$$
\frac{1}{4} (-\sqrt(2) m^2+\sqrt(2) \sqrt(m^2 (9 m^2+4 m-8))-2 \sqrt(2) m) \geq \frac{m^2}{2}
$$

то

$$
d>\frac{1}{2}m^2
$$

Лемма доказана.

(Здесь вставляю кусок с примориальной конструкцией)

Обозначим через $p_i, i\in\mathbb{N}$ простые числа, начиная с 3.
По теореме Чебышева (а не надо ли здесь ссылку?) $p_i \leq bi\ln(i+1)$ для некоторого $b>0$ и всех $i\in\mathbb{N}$.
Обозначим
$$
A_n=\prod_{i=1}^{n} p_i, n\in\mathbb{N}
$$

Тогда
$$
a_n\leq b^n n! \prod_{i=1}^{n} \ln(i+1)
$$

и по формуле Стирлинга
\begin{multline*}
A_n\leq \left(\frac{bn\ln n}{e}\right)
\end{multline*}

\paragraph{Теорема 4.}
$F(2^n)<A_n$ для всех $n\in\mathbb{N}$.

\paragraph{Доказательство.} Через $S$ обозначим множество подмножеств $\{1,2,..,n\}$ и каждому $I\in S$ поставим в соответствие числа $c_I=\prod\limits_{c\in I}p_i, b_I=\frac{1}{2}\left(c_I-\frac{A_n}{c_I}\right)$.
Так как $c_I$ и $\frac{A_n}{c_I}$ нечётны, то $b_I \in \mathbb{Z}$.

Рассмотрим подмножество точек на плоскости
\begin{equation}\label{konstrukcia_primorialy}
M_I=\{(b_I,0), I\in S\}, N=(0, \sqrt{A_n})
\end{equation}

Если $I,J \in S, I \neq J$, то $M_I \neq M_J$.
Поэтоу множество (\ref{konstrukcia_primorialy}) содержит $2^n+2$ элементов.
Так как
\begin{multline*}
	|M_I,N|=\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)=\\
	\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)=\frac{c_I+\frac{A_n}{C_I}}{2} \in \mathbb{N},
\end{multline*}

то все расстояния между точками множества (\ref{konstrukcia_primorialy}) есть целые числа.
Диаметр множества (\ref{konstrukcia_primorialy}) достигается на паре точек $M_{(1,2,...,n)}$ и $M_{\varnothing}$, для которых
$$
|M_{(1,2,...,n)},M_{\varnothing}|=\frac{1}{2}(A_n-1)-\frac{1}{2}(1-A_n) = A_n-1 < A_n
$$

Поэтому 
$$
F(2^n+2) < A_n
$$

Отсюда, из \ref{stirling_primorial} и монотонности $F(n)$ вытекает, что
$$
F(n) \leq \left( \frac{b \log_2(n+1) \ln \log_2 (n+1)}{e}\right)^{\log_2(n+1)}
$$
для всех $n\in \mathbb{N}$.

Заметим, что подобная конструкция не позволяет получить оценку сверху в качестве полинома.
Это вызвано тем, что количество точек в множестве подобного типа (подмножество прямой и точка, на ней не лежащая) ограничено числом  делителей числа $2\mu^2+2$ (эту оценку обосновать?), где $\mu$ --- расстояние от точки до прямой.
С другой стороны, в \cite{Ramanujan} (формулы 198-200) показано, что
$$
p(n)=\max\limits_{1\leq k \leq n} d(k),
$$
где $d(k)$ --- количество делителей числа $k$ (дивизор-функция Рамануджана?),
растёт медленнее любого полинома.

(Осталось свести оба ограничения в кучку)

В ходе численного расчёта были вычислены значения $F(n)$ для $4\leq n \leq 41$.
Время, требуемое для вычисления $F(n)$ разработанными алгоритмами, растёт, как установлено эмпирически, не медленнее, чем $n^4$.
Так, расчёт для $n=5$ занимает меньше секунды, а для вычисления $F(41)$ при известном (вычисленном ранее) $F(40)$ ушло больше суток.

\addcontentsline{toc}{chapter}{Литература}
\begin{thebibliography}{99}

\bibitem{angem1kurs} Аналитическая геометрия на плоскости / Е.М. Семенов, С.Н. Уксусов. – Воронеж : Воронежский государственный университет, 2013. – 100с.

\bibitem{Newman} Donald J. Newman. A Problem Seminar. Springer - Verlag.1982.

\bibitem{Ramanujan} Highly composite numbers. Proceedings of the London Mathematical Society, 2, XIV, 1915, 347 – 409 (http://ramanujan.sirinudi.org/Volumes/published/ram15.pdf)

\end{thebibliography}

\end{document}

### 2016-05-04 Semenov_SCT_statya.tex
Вставлена ссылка на Бухштаба
Внесены исправления из черновика статьи

@@
 
-\paragraph{Теорема 1.} Пусть $\{M_1,M_2\}$ - счётное множество точек на плоскости и расстояние $|M_i,M_j|$ - натуральные для всех $1\leq i < j <\infty$, где $\mathbb{N} - $множество натуральных чисел.
+\paragraph{Теорема 1.} Пусть $\{M_1,M_2\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N} - $множество натуральных чисел.
 Тогда найдется такая прямая на плоскости $l$, что $M_i\in l$ для всех $i \in \mathbb{N}$.
@@
 
-Аналогичное утверждение справедливо, когда $M_1,M_2,M_3$ принадлежат некоторой прямой и $M_2$ лежит между $M_1$ и $M_3$:
+Аналогичное утверждение справедливо, когда $M_1,M_2,M_3$ принадлежат некоторой прямой и $M_2$ лежит между $M_1$ и $M_3$.
+В этом случае
 $n\leq (a+1)(b+1)+3+d(A)$.
@@
 
-Отсюда
-
+Положим $d=\lambda m^2$.
+Тогда
 $$
-d\geq\frac{1}{4} (-\sqrt(2) m^2+\sqrt(2) \sqrt(m^2 (9 m^2+4 m-8))-2 \sqrt(2) m)
+2m^2+1 \leq \left(\lambda m \sqrt2 + 1 \right)^2 + 3 + \lambda m^2 \sqrt2
+$$
+и
+$$
+0\leq \left( 2\lambda^2 + \sqrt2 \lambda - 2 \right)m^2 + 2\sqrt2 \lambda m + 3
 $$
 
-Так как для $m\geq 4$
-
-$$
-\frac{1}{4} (-\sqrt(2) m^2+\sqrt(2) \sqrt(m^2 (9 m^2+4 m-8))-2 \sqrt(2) m) \geq \frac{m^2}{2}
-$$
-
-то
-
-$$
-d>\frac{1}{2}m^2
-$$
+Для $m\geq 4$ это неравенство не выполнено, если $\lambda \leq \frac{1}{2}$.
+Поэтому $\lambda > \frac{1}{2}$ и $d>\frac{m^2}{2}$.
 
@@
 Обозначим через $p_i, i\in\mathbb{N}$ простые числа, начиная с 3.
-По теореме Чебышева (а не надо ли здесь ссылку?) $p_i \leq bi\ln(i+1)$ для некоторого $b>0$ и всех $i\in\mathbb{N}$.
+По теореме Чебышева (\cite{Buhshtab}) $p_i \leq bi\ln(i+1)$ для некоторого $b>0$ и всех $i\in\mathbb{N}$.
 Обозначим
@@
 $$
-a_n\leq b^n n! \prod_{i=1}^{n} \ln(i+1)
+A_n\leq b^n n! \prod_{i=1}^{n} \ln(i+1)
 $$
@@
 и по формуле Стирлинга
-\begin{multline*}
-A_n\leq \left(\frac{bn\ln n}{e}\right)
-\end{multline*}
+\begin{equation*}\label{ocenka_Stirling}
+A_n <
+b^n \sqrt{2\pi n} \left(\frac{n}{e}\right)^n \left(1+\frac{1}{n}\right) \prod_{i=1}^{n} \ln(i+1)
+\leq \left(\frac{bn\ln (n+1)}{e}\right)^n
+\end{equation*}
 
@@
 \paragraph{Доказательство.} Через $S$ обозначим множество подмножеств $\{1,2,..,n\}$ и каждому $I\in S$ поставим в соответствие числа $c_I=\prod\limits_{c\in I}p_i, b_I=\frac{1}{2}\left(c_I-\frac{A_n}{c_I}\right)$.
-Так как $c_I$ и $\frac{A_n}{c_I}$ нечётны, то $b_I \in \mathbb{Z}$.
+Так как $c_I$ и $\frac{A_n}{c_I}$ нечётны, то $b_I$ --- целые числа.
 
@@
 Если $I,J \in S, I \neq J$, то $M_I \neq M_J$.
-Поэтоу множество (\ref{konstrukcia_primorialy}) содержит $2^n+2$ элементов.
+Поэтому множество (\ref{konstrukcia_primorialy}) содержит $2^n+2$ элементов.
 Так как
 \begin{multline*}
-	|M_I,N|=\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)=\\
-	\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)=\frac{c_I+\frac{A_n}{C_I}}{2} \in \mathbb{N},
+	|M_I,N|=\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)^\frac{1}{2}=\\
+	\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)^\frac{1}{2}=
+	\frac{c_I+\frac{A_n}{C_I}}{2} \in \mathbb{N},
 \end{multline*}
@@
 
-Отсюда, из \ref{stirling_primorial} и монотонности $F(n)$ вытекает, что
+Отсюда, из \ref{ocenka_Stirling} и монотонности $F(n)$ вытекает, что
 $$
@@
 
-\bibitem{angem1kurs} Аналитическая геометрия на плоскости / Е.М. Семенов, С.Н. Уксусов. – Воронеж : Воронежский государственный университет, 2013. – 100с.
+\bibitem{Newman} Donald J. Newman. A Problem Seminar. Springer - Verlag.1982. Problem 29.
 
-\bibitem{Newman} Donald J. Newman. A Problem Seminar. Springer - Verlag.1982.
+\bibitem{angem1kurs} Аналитическая геометрия на плоскости / Е.М. Семенов, С.Н. Уксусов. – Воронеж : Воронежский государственный университет, 2016. – 100с.
 
 \bibitem{Ramanujan} Highly composite numbers. Proceedings of the London Mathematical Society, 2, XIV, 1915, 347 – 409 (http://ramanujan.sirinudi.org/Volumes/published/ram15.pdf)
+
+\bibitem{Buhshtab} Теория чисел / А.А. Бухштаб. --- М., 1966 (теорема 325)
 

### 2016-05-04 Semenov_SCT_statya.tex
Косметически правки

@@
 
-\paragraph{Теорема 1.} Пусть $\{M_1,M_2\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N} - $множество натуральных чисел.
+\paragraph{Теорема 1.} Пусть $\{M_1,M_2\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N}$ --- множество натуральных чисел.
 Тогда найдется такая прямая на плоскости $l$, что $M_i\in l$ для всех $i \in \mathbb{N}$.
@@
 Полное доказательство можно найти в \cite{angem1kurs}.
-Там же показан, что для любого $N\in \mathbb{N}$ существует такое множество $\{M_1,M_2,...,M_n\}\subset \mathbb{R}^2$, что $|M_i,M_j|$ для всех $1\leq i < j  \leq n$ и $M_1, M_2, ..., M_n$ не лежат на прямой.
+Там же показано, что для любого $n\in \mathbb{N}$ существует такое множество $\{M_1,M_2,...,M_n\}\subset \mathbb{R}^2$, что $|M_i,M_j|$ для всех $1\leq i < j  \leq n$ и $M_1, M_2, ..., M_n$ не лежат на прямой.
 Изучению таких подмножеств посвящена настоящая работа.
 
-Для заданного $n\in \mathbb{N}$ обозначим через $C_n$ множество таких последовательностей $M_1,M_2,...,M_n \in \mathbb{R}^2$, что $|M_i,M_j|\in\mathbb{N}$ ля всех $1\leq i < j  \leq n$ и  $M_1,M_2,...,M_n$ не принадлежат никакой прямой.
+Для заданного $n\in \mathbb{N}$ обозначим через $C_n$ множество таких последовательностей $M_1,M_2,...,M_n \in \mathbb{R}^2$, что $|M_i,M_j|\in\mathbb{N}$ для всех $1\leq i < j  \leq n$ и  $M_1,M_2,...,M_n$ не принадлежат никакой прямой.
 Положим

### 2016-05-04 Semenov_SCT_statya.tex
Внесена благодарность проф. Ю. А. Брудному и упоминание о гранте
Исправлены мелкие косметические недочёты

@@
 
-(Здесь вставляю кусок с примориальной конструкцией)
+(Здесь вставляю кусок с примориальной конструкцией, т.е. он уже набран ниже)
 
@@
 Отсюда, из \ref{ocenka_Stirling} и монотонности $F(n)$ вытекает, что
-$$
+\begin{equation}\label{ocenka_primoryaly}
 F(n) \leq \left( \frac{b \log_2(n+1) \ln \log_2 (n+1)}{e}\right)^{\log_2(n+1)}
-$$
+\end{equation}
 для всех $n\in \mathbb{N}$.
 
-Заметим, что подобная конструкция не позволяет получить оценку сверху в качестве полинома.
+Заметим, что подобная конструкция не позволяет получить оценку сверху в виде полинома.
 Это вызвано тем, что количество точек в множестве подобного типа (подмножество прямой и точка, на ней не лежащая) ограничено числом  делителей числа $2\mu^2+2$ (эту оценку обосновать?), где $\mu$ --- расстояние от точки до прямой.
@@
 
-(Осталось свести оба ограничения в кучку)
+\paragraph{Теорема 5.}
 
-В ходе численного расчёта были вычислены значения $F(n)$ для $4\leq n \leq 41$.
+$$
+ \max\left( \frac{n-5}{8} \right) \geq F(n) \leq \left( \frac{b \log_2(n+1) \ln \log_2 (n+1)}{e}\right)^{\log_2(n+1)}
+$$
+для всех $n\in\mathbb{N}$.
+
+\paragraph{Доказательство.}
+Верхняя оценка непосредственно следует из \ref{ocenka_primoryaly}.
+
+Пусть $\left(M_1, M_2, ..., M_n\right) \in C_n$.
+Найдём такое $m\in \mathbb{N}$, что
+$$
+2m^2+1 \leq n < 2(m+1)^2 +1
+$$
+
+Пусть $d$ --- сторона минимального квадрата, содержащего все точки $M_i, 1\leq i \leq m^2+1$.
+Используя монотонность последовательности $F(n)$ и лемму 3, получаем для $n\geq 17$
+$$
+F(n) \leq F\left(2m^2+1\right)\geq d \geq \frac{1}{2}m^2 > \frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} - \right)
+$$
+
+Простые оценки показывают, что
+$$
+\frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} -1 \right) > \frac{n-5}{8}
+$$
+
+Отсюда вытекает, что
+$$
+\frac{n-5}{8}<F(n)
+$$
+
+Для $3 \leq n < 15$ значения $F(n)$ были вычислены на ЭВМ (см. далее).
+
+Теорема доказана.
+
+В ходе численного расчёта были вычислены значения $F(n)$ для $4\leq n \leq 42$.
+
 Время, требуемое для вычисления $F(n)$ разработанными алгоритмами, растёт, как установлено эмпирически, не медленнее, чем $n^4$.
 Так, расчёт для $n=5$ занимает меньше секунды, а для вычисления $F(41)$ при известном (вычисленном ранее) $F(40)$ ушло больше суток.
+
+Авторы благодарят проф. Ю. А. Брудного за информацию о работе \cite{Newman}.
+
+Работа выполнена при поддержке РФФИ, грант 14-01-00141а.
 

### 2016-05-04 Semenov_SCT_statya.tex
Добавлена таблица с вычисленными значениями F(n)
Замечание о целочисленной решётке
Очепятка

@@
 $$
-F(n) \leq F\left(2m^2+1\right)\geq d \geq \frac{1}{2}m^2 > \frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} - \right)
+F(n) \leq F\left(2m^2+1\right)\geq d \geq \frac{1}{2}m^2 > \frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} - 1 \right)
 $$
@@
 В ходе численного расчёта были вычислены значения $F(n)$ для $4\leq n \leq 42$.
+
+\begin{table}[H]
+\caption{Результаты численного расчёта}
+\label{tabular:pc_counted}
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
+\hline
+\textbf{n}    &  3 & 4 & 5 & 6 &  7 &  8 &  9 & 10 & 11 & 12 & 13 \\
+\hline
+\textbf{F(n)} &  1 & 4 & 7 & 8 & 17 & 21 & 29 & 40 & 51 & 63 & 74  \\
+\hline
+\hline
+\textbf{n}    &  14 &  15 &  16 &  17 &  18 &  19 &  20 &  21 &  22 &  23 &  24 \\
+\hline
+\textbf{F(n)} &  91 & 104 & 121 & 134 & 153 & 164 & 196 & 212 & 234 & 256 & 286  \\
+\hline
+\hline
+\textbf{n}    &  25 &  26 &  27 &  28 &  29 &  30 &  31 &  32 &  33 &  34 &  35 \\
+\hline
+\textbf{F(n)} & 304 & 338 & 370 & 384 & 414 & 448 & 464 & 494 & 524 & 553 & 578  \\
+\hline
+\end{tabular}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\hline
+\textbf{n}    &  36 &  37 &  38 &  39 &  40 &  41 &  42 &  43  \\
+\hline
+\textbf{F(n)} & 608 & 642 & 667 & 692 & 754 & 816 & 897 & >955  \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Заметим, что соответствующие множества нельзя расположить на целочисленной решётке, но можно расположить на решётке $\left\{\pm\frac{\sqrt{p}}{q}\right\}$, где $p,q \in \mathbb{N}$.
 

### 2016-05-04 Semenov_SCT_statya.tex
Дальнейшие заключительные замечания

@@
 
-Заметим, что соответствующие множества нельзя расположить на целочисленной решётке, но можно расположить на решётке $\left\{\pm\frac{\sqrt{p}}{q}\right\}$, где $p,q \in \mathbb{N}$.
+Заметим, что соответствующие множества нельзя расположить на целочисленной решётке, но можно расположить на решётке 
+$$
+	\left(
+		\left\{\pm\frac{\sqrt{p_1}}{q_1}\right\};
+		\left\{\pm\frac{\sqrt{p_2}}{q_2}\right\}
+	\right),$$
+где $p_1,q_1,p_2,q_2 \in \mathbb{N} \cup{0}$.
+
+(Надо ли это доказывать?).
+
+В большинстве обсчитанных случаев оптимальное расположение точек определяется однозначно с точностью до симметрий.
+Тем не менее, например, для $n=18$ имеем два набора:
+
+\begin{multline*}
+\{( 0 ; 0 ); ( 0 ; 153 ); ( 0 ; 144 ); ( 0 ; 130 ); ( 0 ; 115 ); ( 0 ; 111 ); ( 0 ; 104 ); ( 0 ; 98 ); ( 0 ; 88 ); ( 0 ; 76 );\\
+( 0 ; 66 ); ( 0 ; 60 ); ( 0 ; 53 ); ( 0 ; 49 ); ( 0 ; 34 ); ( 0 ; 20 ); ( 0 ; 11 ); ( \sqrt{2880} ; 82 )\}
+\end{multline*}
+и
+\begin{multline*}
+\{( 0 ; 0 ); ( 0 ; 153 ); ( 0 ; 134 ); ( 0 ; 121 ); ( 0 ; 104 ); ( 0 ; 98 ); ( 0 ; 93 ); ( 0 ; 85 ); ( 0 ; 76 ); ( 0 ; 69 );\\
+( 0 ; 65 ); ( 0 ; 58 ); ( 0 ; 49 ); ( 0 ; 41 ); ( 0 ; 36 ); ( 0 ; 30 ); ( 0 ; 13 ); ( \sqrt{1440} ; 67 )\}
+\end{multline*}
+
+Также примечательно, что, начиная с $n=9$, для всех обсчитанных $n$ оптимальное множество, кроме одной точки, лежит на прямой (т. е. $n-1$ точек на прямой и одна точка вне её).
+
+(Я готов выложить текст программы, которой считал. Нужно ли дать ссылку?)
 

### 2016-05-05 Semenov_SCT_statya.tex
Убрана часть технических пометок
Инкремент F(43)
Изменён порядок библиографии
УДК и прочие реверансы
Увеличен шрифт

@@
 
-УДК ???
+УДК 511.95 + 514.112
 
-О множествах точек на плоскости с целочисленными расстояниями
+\begin{center}
+	\textbf{\LARGE О множествах точек на плоскости с целочисленными расстояниями}
+\end{center}
 
-Н.Н. Авдеев, Е.М. Семёнов.
+{\Large{Н.Н. Авдеев, Е.М. Семёнов.}}
 
@@
 
-(Евгений Михайлович! Здесь вылезло $+d(A)$ в правой части.
-В следующей лемме из-за этого может испортиться константа перед $m^2$.
-Чтобы её сберечь, повысил $m$ в условии с 2 до 4.
-Надо всё пересчитывать $=($ ~~ )
-
 \paragraph{Лемма 3.}
@@
 Лемма доказана.
-
-(Здесь вставляю кусок с примориальной конструкцией, т.е. он уже набран ниже)
 
@@
 \hline
-\textbf{F(n)} & 608 & 642 & 667 & 692 & 754 & 816 & 897 & >955  \\
+\textbf{F(n)} & 608 & 642 & 667 & 692 & 754 & 816 & 897 & >956  \\
 \hline
@@
 
-Авторы благодарят проф. Ю. А. Брудного за информацию о работе \cite{Newman}.
+Авторы благодарят проф. Ю. А. Брудного за информацию о работе \cite{Newman} и ценные замечания.
 
-Работа выполнена при поддержке РФФИ, грант 14-01-00141а.
+Работа выполнена при поддержке РНФ, грант ???.
 
@@
 
+\bibitem{Buhshtab} Теория чисел / А.А. Бухштаб. --- М., 1966 (теорема 325)
+
 \bibitem{Ramanujan} Highly composite numbers. Proceedings of the London Mathematical Society, 2, XIV, 1915, 347 – 409 (http://ramanujan.sirinudi.org/Volumes/published/ram15.pdf)
-
-\bibitem{Buhshtab} Теория чисел / А.А. Бухштаб. --- М., 1966 (теорема 325)
 

### 2016-05-05 Semenov_SCT_statya.tex
Расстановка значков параграфа и конца доказательства
Внесены исправления

@@
 
+\paragraph{\S 1} % Махровейшее вордописчество =(
 Известна следующая
@@
 
-Формулировка теоремы и идея её доказательства приведены в \cite{Newman}.
-Полное доказательство можно найти в \cite{angem1kurs}.
+Формулировка теоремы и идея её доказательства приведены в \cite{Newman}, problem 29. Полное доказательство можно найти в \cite{angem1kurs}.
 Там же показано, что для любого $n\in \mathbb{N}$ существует такое множество $\{M_1,M_2,...,M_n\}\subset \mathbb{R}^2$, что $|M_i,M_j|$ для всех $1\leq i < j  \leq n$ и $M_1, M_2, ..., M_n$ не лежат на прямой.
@@
 
+\paragraph{\S 2}
 Число элементов множества $A$ обозначим через $|A|$.
@@
 Поэтому $\lambda > \frac{1}{2}$ и $d>\frac{m^2}{2}$.
-
-Лемма доказана.
+$\Box$
 
 Обозначим через $p_i, i\in\mathbb{N}$ простые числа, начиная с 3.
-По теореме Чебышева (\cite{Buhshtab}) $p_i \leq bi\ln(i+1)$ для некоторого $b>0$ и всех $i\in\mathbb{N}$.
+По теореме Чебышева (\cite{Buhshtab}, теорема 325) $p_i \leq bi\ln(i+1)$ для некоторого $b>0$ и всех $i\in\mathbb{N}$.
 Обозначим
@@
 и по формуле Стирлинга
-\begin{equation*}\label{ocenka_Stirling}
+\begin{equation}\label{ocenka_Stirling}
 A_n <
@@
 \leq \left(\frac{bn\ln (n+1)}{e}\right)^n
-\end{equation*}
+\end{equation}
 
@@
 
-\paragraph{Доказательство.} Через $S$ обозначим множество подмножеств $\{1,2,..,n\}$ и каждому $I\in S$ поставим в соответствие числа $c_I=\prod\limits_{c\in I}p_i, b_I=\frac{1}{2}\left(c_I-\frac{A_n}{c_I}\right)$.
+\paragraph{Доказательство.}
+Через $S$ обозначим множество подмножеств $\{1,2,..,n\}$ и каждому $I\in S$ поставим в соответствие числа $c_I=\prod\limits_{c\in I}p_i, b_I=\frac{1}{2}\left(c_I-\frac{A_n}{c_I}\right)$.
 Так как $c_I$ и $\frac{A_n}{c_I}$ нечётны, то $b_I$ --- целые числа.
@@
 Так как
-\begin{multline*}
-	|M_I,N|=\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)^\frac{1}{2}=\\
+\begin{equation*}
+	|M_I,N|=\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)^\frac{1}{2}=
 	\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)^\frac{1}{2}=
 	\frac{c_I+\frac{A_n}{C_I}}{2} \in \mathbb{N},
-\end{multline*}
-
+\end{equation*}
 то все расстояния между точками множества (\ref{konstrukcia_primorialy}) есть целые числа.
@@
 $$
- \max\left( \frac{n-5}{8} \right) \geq F(n) \leq \left( \frac{b \log_2(n+1) \ln \log_2 (n+1)}{e}\right)^{\log_2(n+1)}
+ \max\left( \frac{n-5}{8} \right) \leq F(n) \leq \left( \frac{b \log_2(n+1) \ln \log_2 (n+1)}{e}\right)^{\log_2(n+1)}
 $$
@@
 Для $3 \leq n < 15$ значения $F(n)$ были вычислены на ЭВМ (см. далее).
-
-Теорема доказана.
+$\Box$
 
@@
 
-\bibitem{Newman} Donald J. Newman. A Problem Seminar. Springer - Verlag.1982. Problem 29.
+\bibitem{Newman} Donald J. Newman. A Problem Seminar. Springer - Verlag.1982.
 
@@
 
-\bibitem{Buhshtab} Теория чисел / А.А. Бухштаб. --- М., 1966 (теорема 325)
+\bibitem{Buhshtab} Теория чисел / А.А. Бухштаб. --- М., 1966.
 

### 2016-05-05 Semenov_SCT_statya.tex
Изменена формула верхней оценки
Две теорему слиты в одну

@@
 \paragraph{Теорема 4.}
+$$
+ \max\left( \frac{n-5}{8}, 7  \right) \leq \left( \frac{b (1+\log_2 n )\ln (1 + \log_2 n)}{e}\right)^{1+\log_2 n}
+$$
+для всех $n\in\mathbb{N}$.
+
+\paragraph{Доказательство.}
+Для доказательства верхней оценки докажем сначала, что
 $F(2^n)<A_n$ для всех $n\in\mathbb{N}$.
 
-\paragraph{Доказательство.}
 Через $S$ обозначим множество подмножеств $\{1,2,..,n\}$ и каждому $I\in S$ поставим в соответствие числа $c_I=\prod\limits_{c\in I}p_i, b_I=\frac{1}{2}\left(c_I-\frac{A_n}{c_I}\right)$.
@@
 \begin{equation}\label{ocenka_primoryaly}
-F(n) \leq \left( \frac{b \log_2(n+1) \ln \log_2 (n+1)}{e}\right)^{\log_2(n+1)}
+	F(n) \leq \left( \frac{b (1+\log_2 n )\ln (1 + \log_2 n)}{e}\right)^{1+\log_2 n}
 \end{equation}
@@
 
-\paragraph{Теорема 5.}
-
-$$
- \max\left( \frac{n-5}{8} \right) \leq F(n) \leq \left( \frac{b \log_2(n+1) \ln \log_2 (n+1)}{e}\right)^{\log_2(n+1)}
-$$
-для всех $n\in\mathbb{N}$.
-
-\paragraph{Доказательство.}
-Верхняя оценка непосредственно следует из \ref{ocenka_primoryaly}.
-
+Докажем теперь нижнюю оценку.
 Пусть $\left(M_1, M_2, ..., M_n\right) \in C_n$.
@@
 $$
-F(n) \leq F\left(2m^2+1\right)\geq d \geq \frac{1}{2}m^2 > \frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} - 1 \right)
+F(n) \geq F\left(2m^2+1\right)\geq d \geq \frac{1}{2}m^2 > \frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} - 1 \right)
 $$
@@
 
-Для $3 \leq n < 15$ значения $F(n)$ были вычислены на ЭВМ (см. далее).
+Для $3 \leq n < 17$ значения $F(n)$ были вычислены на ЭВМ (см. далее).
 $\Box$

### 2016-05-05 Semenov_SCT_statya.tex
Рокировка координат в примере для n=18
Замечание о n<9
Донабран кусок про численный эксперимент

@@
 
-В ходе численного расчёта были вычислены значения $F(n)$ для $4\leq n \leq 42$.
+\paragraph{\S 3} Множество точек $A=\{M_1, M_2, ..., M_n\}$ называется оптимальным, если $A \in C_n$ и $d(A) = F(n)$.
+Например, правильный треугольник со стороной 1 показывает, что $F(3)=1$.
+Для нахождения оптимальных множеств и вычисления $F(n)$ была создана программа, которую удалось реализовать на ЭВМ.
+
+Результаты численного эксперимента.
+
+\begin{enumerate}
+
+\item.
+Были вычислены значения $F(n)$ для $4\leq n \leq 42$.
 
 \begin{table}[H]
-\caption{Результаты численного расчёта}
+\caption{Результаты численного эксперимента, значения $F(n)$}
 \label{tabular:pc_counted}
@@
 
-Заметим, что соответствующие множества нельзя расположить на целочисленной решётке, но можно расположить на решётке 
+\item.
+Оптимальные множества не принадлежат целочисленной решётке.
+Однако координаты любого (не обязательно оптимального) множества с точностью до движения имеют вид
 $$
 	\left(
-		\left\{\pm\frac{\sqrt{p_1}}{q_1}\right\};
-		\left\{\pm\frac{\sqrt{p_2}}{q_2}\right\}
+		\left\{\pm\frac{\sqrt{p}}{q}\right\};
+		\left\{\pm\frac{\sqrt{r}}{s}\right\}
 	\right),$$
-где $p_1,q_1,p_2,q_2 \in \mathbb{N} \cup{0}$.
+где $p,q,r,s \in \mathbb{N} \cup \{0\}$.
 
-(Надо ли это доказывать?).
-
-В большинстве обсчитанных случаев оптимальное расположение точек определяется однозначно с точностью до симметрий.
+\item.
+В большинстве обсчитанных случаев оптимальное расположение точек определяется однозначно с точностью до движения.
 Тем не менее, например, для $n=18$ имеем два набора:
@@
 \begin{multline*}
-\{( 0 ; 0 ); ( 0 ; 153 ); ( 0 ; 144 ); ( 0 ; 130 ); ( 0 ; 115 ); ( 0 ; 111 ); ( 0 ; 104 ); ( 0 ; 98 ); ( 0 ; 88 ); ( 0 ; 76 );\\
-( 0 ; 66 ); ( 0 ; 60 ); ( 0 ; 53 ); ( 0 ; 49 ); ( 0 ; 34 ); ( 0 ; 20 ); ( 0 ; 11 ); ( \sqrt{2880} ; 82 )\}
+\{( 0 ; 0 ); ( 153 ; 0 ); ( 144 ; 0 ); ( 130 ; 0 ); ( 115 ; 0 ); ( 111 ; 0 ); ( 104 ; 0 ); ( 98 ; 0 ); ( 88 ; 0 ); ( 76 ; 0 );\\
+( 66 ; 0 ); ( 60 ; 0 ); ( 53 ; 0 ); ( 49 ; 0 ); ( 34 ; 0 ); ( 20 ; 0 ); ( 11 ; 0 ); ( 82 ; \sqrt{2880})\}
 \end{multline*}
@@
 \begin{multline*}
-\{( 0 ; 0 ); ( 0 ; 153 ); ( 0 ; 134 ); ( 0 ; 121 ); ( 0 ; 104 ); ( 0 ; 98 ); ( 0 ; 93 ); ( 0 ; 85 ); ( 0 ; 76 ); ( 0 ; 69 );\\
-( 0 ; 65 ); ( 0 ; 58 ); ( 0 ; 49 ); ( 0 ; 41 ); ( 0 ; 36 ); ( 0 ; 30 ); ( 0 ; 13 ); ( \sqrt{1440} ; 67 )\}
+\{( 0 ; 0 ); ( 153 ; 0 ); ( 134 ; 0 ); ( 121 ; 0 ); ( 104 ; 0 ); ( 98 ; 0 ); ( 93 ; 0 ); ( 85 ; 0 ); ( 76 ; 0 ); ( 69 ; 0 );\\
+( 65 ; 0 ); ( 58 ; 0 ); ( 49 ; 0 ); ( 41 ; 0 ); ( 36 ; 0 ); ( 30 ; 0 ); ( 13 ; 0 ); ( 67 ; \sqrt{1440} )\}
 \end{multline*}
 
-Также примечательно, что, начиная с $n=9$, для всех обсчитанных $n$ оптимальное множество, кроме одной точки, лежит на прямой (т. е. $n-1$ точек на прямой и одна точка вне её).
+\item.
+Для $N \geq 9$ оптимальное множество имеет вид, описанный в теореме 4:
+$n-1$ точек лежат на оси иксов с целыми координатами и одна точка не принадлежит оси иксов, её вторая координата иррациональна.
+Для меньших $n$ эта закономерность не выполняется.
 
-(Я готов выложить текст программы, которой считал. Нужно ли дать ссылку?)
+\item.
+Время, требуемое для вычисления $F(n)$ разработанными алгоритмами, растёт, как установлено эмпирически, не медленнее, чем $n^4$.
+Так, расчёт для $n \leq 5$ занимает меньше секунды, а для вычисления $F(41)$ при известном (вычисленном ранее) $F(40)$ ушло больше суток.
 
-Время, требуемое для вычисления $F(n)$ разработанными алгоритмами, растёт, как установлено эмпирически, не медленнее, чем $n^4$.
-Так, расчёт для $n=5$ занимает меньше секунды, а для вычисления $F(41)$ при известном (вычисленном ранее) $F(40)$ ушло больше суток.
+\end{enumerate}
 

### 2016-05-14 Semenov_SCT_statya.tex
Очередная порция правок

@@
 	\textbf{\LARGE О множествах точек на плоскости с целочисленными расстояниями}
-\end{center}
+
+	\vspace{15mm}
 
 	{\Large{Н.Н. Авдеев, Е.М. Семёнов.}}
+
+\end{center}
 
@@
 Формулировка теоремы и идея её доказательства приведены в \cite{Newman}, problem 29. Полное доказательство можно найти в \cite{angem1kurs}.
-Там же показано, что для любого $n\in \mathbb{N}$ существует такое множество $\{M_1,M_2,...,M_n\}\subset \mathbb{R}^2$, что $|M_i,M_j|$ для всех $1\leq i < j  \leq n$ и $M_1, M_2, ..., M_n$ не лежат на прямой.
+Там же показано, что для любого $n\in \mathbb{N}$ существует такое множество $\{M_1,M_2,...,M_n\}\subset \mathbb{R}^2$, что $|M_i,M_j|\in\mathbb{N}$ для всех $1\leq i < j  \leq n$ и $M_1, M_2, ..., M_n$ не лежат на прямой.
 Изучению таких подмножеств посвящена настоящая работа.
 
-Для заданного $n\in \mathbb{N}$ обозначим через $C_n$ множество таких последовательностей $M_1,M_2,...,M_n \in \mathbb{R}^2$, что $|M_i,M_j|\in\mathbb{N}$ для всех $1\leq i < j  \leq n$ и  $M_1,M_2,...,M_n$ не принадлежат никакой прямой.
+Для заданного $n\in \mathbb{N}, n\geq 3$ обозначим через $C_n$ множество таких последовательностей $M_1,M_2,...,M_n \in \mathbb{R}^2$, что $|M_i,M_j|\in\mathbb{N}$ для всех $1\leq i < j  \leq n$ и  $M_1,M_2,...,M_n$ не принадлежат никакой прямой.
 Положим
@@
 
-Точную асимптотику последовательности  $F(n)$ найти не удалось, получены лишь верхняя и нижняя оценки.
+Точную асимптотику последовательности  $F(n)$ найти не удалось, получены лишь верхняя и нижняя оценки и найдены $F(n)$ для $3\leq n \leq 43$.
 
@@
 В этом случае
-$n\leq (a+1)(b+1)+3+d(A)$.
+\begin{equation}\label{ocenka_iz_lemmy_2}
+	n\leq (a+1)(b+1)+3+d(A).
+\end{equation}
 
@@
 Поэтому $|M_1,M_2|,|M_2,M_3| \leq \frac{d}{m}\sqrt{2}$.
-В силу леммы 2
+В силу (\ref{ocenka_iz_lemmy_2})
 $$
@@
 \paragraph{Теорема 4.}
+Неравенства
 $$
- \max\left( \frac{n-5}{8}, 7  \right) \leq \left( \frac{b (1+\log_2 n )\ln (1 + \log_2 n)}{e}\right)^{1+\log_2 n}
+	\max\left( \frac{n-5}{8}, 1 \right) \leq F(n) \leq \left( \frac{b (1+\log_2 n )\ln (1 + \log_2 n)}{e}\right)^{1+\log_2 n}
 $$
@@
 Для доказательства верхней оценки докажем сначала, что
-$F(2^n)<A_n$ для всех $n\in\mathbb{N}$.
+$F(2^n)<A_n$ для всех $n\in\mathbb{N}, n \geq 3$, где $b$ - константа из неравенств Чебышева.
 
@@
 Если $I,J \in S, I \neq J$, то $M_I \neq M_J$.
-Поэтому множество (\ref{konstrukcia_primorialy}) содержит $2^n+2$ элементов.
+Поэтому множество (\ref{konstrukcia_primorialy}) содержит $2^n+1$ элементов.
 Так как
@@
 $$
-F(2^n+2) < A_n
+F(2^n+1) < A_n
 $$
@@
 для всех $n\in \mathbb{N}$.
-
-Заметим, что подобная конструкция не позволяет получить оценку сверху в виде полинома.
-Это вызвано тем, что количество точек в множестве подобного типа (подмножество прямой и точка, на ней не лежащая) ограничено числом  делителей числа $2\mu^2+2$ (эту оценку обосновать?), где $\mu$ --- расстояние от точки до прямой.
-С другой стороны, в \cite{Ramanujan} (формулы 198-200) показано, что
-$$
-p(n)=\max\limits_{1\leq k \leq n} d(k),
-$$
-где $d(k)$ --- количество делителей числа $k$ (дивизор-функция Рамануджана?),
-растёт медленнее любого полинома.
 
@@
 
+Заметим, что подобная конструкция не позволяет получить оценку сверху в виде полинома.
+Это вызвано тем, что количество точек в множестве подобного типа (подмножество прямой и точка, на ней не лежащая) ограничено числом  делителей числа $2\mu^2+2$, где $\mu$ --- расстояние от точки до прямой.
+С другой стороны, в \cite{Ramanujan} (формулы 198-200) показано, что
+$$
+p(n)=\max\limits_{1\leq k \leq n} D(k),
+$$
+где $D(k)$ --- количество делителей числа $k$ (дивизор-функция Рамануджана),
+растёт медленнее степенной функции.
+
 \paragraph{\S 3} Множество точек $A=\{M_1, M_2, ..., M_n\}$ называется оптимальным, если $A \in C_n$ и $d(A) = F(n)$.
-Например, правильный треугольник со стороной 1 показывает, что $F(3)=1$.
+Пример правильного треугольника со стороной 1 показывает, что $F(3)=1$.
 Для нахождения оптимальных множеств и вычисления $F(n)$ была создана программа, которую удалось реализовать на ЭВМ.
@@
 \end{tabular}
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
 \hline
-\textbf{n}    &  36 &  37 &  38 &  39 &  40 &  41 &  42 &  43  \\
+\textbf{n}    &  36 &  37 &  38 &  39 &  40 &  41 &  42 &  43 & 44 \\
 \hline
-\textbf{F(n)} & 608 & 642 & 667 & 692 & 754 & 816 & 897 & >956  \\
+\textbf{F(n)} & 608 & 642 & 667 & 692 & 754 & 816 & 897 & 959 & >963 \\
 \hline
@@
 
+Например,
+$$
+A_4=\left\{( 0 ; 0 ); ( 0 ; 1 ); \left( \frac{\sqrt{15}}{2} ; \frac{1}{2} \right); ( 0 ; 4 ); \right\}
+$$
+
 \item.
 Оптимальные множества не принадлежат целочисленной решётке.
-Однако координаты любого (не обязательно оптимального) множества с точностью до движения имеют вид
+Однако координаты любого множества из $C_n$ с точностью до движения имеют вид
 $$
@@
 \item.
-В большинстве обсчитанных случаев оптимальное расположение точек определяется однозначно с точностью до движения.
-Тем не менее, например, для $n=18$ имеем два набора:
+В большинстве случаев оптимальное множество с точностью до движения определяется однозначно.
+Тем не менее, например, для $n=18$ имеем два оптимальных набора:
 
@@
 
+\item.
+Оптимальные множества для $ 25 \leq n \leq 42$ являются подмножествами оптимального множества для $n=43$ и получаются по конструкции, приведённой в теореме 4, если вместо $A_n$ (перегруз обозначения?) взять 20160 (это не просто так, это HCN - hoghly composite number).
+
 \end{enumerate}
@@
 
-\bibitem{Newman} Donald J. Newman. A Problem Seminar. Springer - Verlag.1982.
+\bibitem{Newman} D.J. Newman. A Problem Seminar. Springer - Verlag.1982.
 
-\bibitem{angem1kurs} Аналитическая геометрия на плоскости / Е.М. Семенов, С.Н. Уксусов. – Воронеж : Воронежский государственный университет, 2016. – 100с.
+\bibitem{angem1kurs} Е.М. Семенов, С.Н. Уксусов / Аналитическая геометрия на плоскости. – Воронеж : Воронежский государственный университет, 2016. – 130с.
 
-\bibitem{Buhshtab} Теория чисел / А.А. Бухштаб. --- М., 1966.
+\bibitem{Buhshtab} А.А. Бухштаб / Теория чисел. --- М., 1966.
 
-\bibitem{Ramanujan} Highly composite numbers. Proceedings of the London Mathematical Society, 2, XIV, 1915, 347 – 409 (http://ramanujan.sirinudi.org/Volumes/published/ram15.pdf)
+\bibitem{Ramanujan} Ramanujan S. / Highly composite numbers. Proceedings of the London Mathematical Society, 2, XIV, 1915.
 

### 2016-05-18 Semenov_SCT_statya.tex
Внесён номер гранта
Недопущение вылезание строк на поля справа
Список литературы вынесен на отдельную страницу
Текущие правки

@@
 \renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра
+
+\sloppy
 
@@
 $$
-	\max\left( \frac{n-5}{8}, 1 \right) \leq F(n) \leq \left( \frac{b (1+\log_2 n )\ln (1 + \log_2 n)}{e}\right)^{1+\log_2 n}
+	\max\left( \frac{n-5}{8}, 1 \right) \leq F(n) \leq \left( \frac{b (1+\log_2 n )\ln (1 + \log_2 n)}{e}\right)^{1+\log_2 n},
 $$
-для всех $n\in\mathbb{N}$.
+где $b$ --- константа из неравенств Чебышева, справедливы для всех $n\in\mathbb{N}$, $n \geq 3$.
 
 \paragraph{Доказательство.}
-Для доказательства верхней оценки докажем сначала, что
-$F(2^n)<A_n$ для всех $n\in\mathbb{N}, n \geq 3$, где $b$ - константа из неравенств Чебышева.
+Докажем сначала, что
+$F(2^n)<A_n$ для всех $n\in\mathbb{N}, n \geq 3$.
 
-Через $S$ обозначим множество подмножеств $\{1,2,..,n\}$ и каждому $I\in S$ поставим в соответствие числа $c_I=\prod\limits_{c\in I}p_i, b_I=\frac{1}{2}\left(c_I-\frac{A_n}{c_I}\right)$.
+Через $S$ обозначим множество подмножеств $\{1,2,..,n\}$ и каждому $I\in S$ поставим в соответствие числа
+$c_I=\prod\limits_{c\in I}p_i, b_I=\frac{1}{2}\left(c_I-\frac{A_n}{c_I}\right)$,
+полагая $c_\varnothing = 1$.
 Так как $c_I$ и $\frac{A_n}{c_I}$ нечётны, то $b_I$ --- целые числа.
@@
 Пусть $d$ --- сторона минимального квадрата, содержащего все точки $M_i, 1\leq i \leq m^2+1$.
-Используя монотонность последовательности $F(n)$ и лемму 3, получаем для $n\geq 17$
+Используя монотонность последовательности $F(n)$ и лемму 3, получаем
 $$
@@
 
-Отсюда вытекает, что
+Для $n\geq 17$ отсюда вытекает, что
 $$
@@
 
-Для $3 \leq n < 17$ значения $F(n)$ были вычислены на ЭВМ (см. далее).
+Для $4 \leq n < 17$ значения $F(n)$ были вычислены на ЭВМ (см. далее).
 $\Box$
 
-Заметим, что подобная конструкция не позволяет получить оценку сверху в виде полинома.
-Это вызвано тем, что количество точек в множестве подобного типа (подмножество прямой и точка, на ней не лежащая) ограничено числом  делителей числа $2\mu^2+2$, где $\mu$ --- расстояние от точки до прямой.
-С другой стороны, в \cite{Ramanujan} (формулы 198-200) показано, что
+Заметим, что подобная конструкция не позволяет получить верхнюю оценку $F(n)$ в виде полинома.
+Это вызвано тем, что количество точек в множестве подобного типа (подмножество прямой и точка, на ней не лежащая) ограничено числом $p(2\mu^2)+2$, где $\mu$ --- расстояние от точки до прямой,
 $$
@@
 $$
-где $D(k)$ --- количество делителей числа $k$ (дивизор-функция Рамануджана),
+$D(k)$ --- количество делителей числа $k$ (дивизор-функция Рамануджана),
+С другой стороны, в \cite{Ramanujan} (формулы 198-200) показано, что $p(n)$
 растёт медленнее степенной функции.
 
-\paragraph{\S 3} Множество точек $A=\{M_1, M_2, ..., M_n\}$ называется оптимальным, если $A \in C_n$ и $d(A) = F(n)$.
+\paragraph{\S 3}
+Множество точек $B=\{M_1, M_2, ..., M_n\}$ называется оптимальным, если $B \in C_n$ и $d(B) = F(n)$.
+Ясно, что любое оптимальное множество определяется с точностью до движения.
+Пусть $B_n \in C_n$ и оптимально.
 Пример правильного треугольника со стороной 1 показывает, что $F(3)=1$.
@@
 \begin{table}[H]
-\caption{Результаты численного эксперимента, значения $F(n)$}
+\caption{Значения $F(n)$}
 \label{tabular:pc_counted}
@@
 $$
-A_4=\left\{( 0 ; 0 ); ( 0 ; 1 ); \left( \frac{\sqrt{15}}{2} ; \frac{1}{2} \right); ( 0 ; 4 ); \right\}
+B_4=\left\{( 0 ; 0 ); ( 0 ; 1 ); \left( \frac{\sqrt{15}}{2} ; \frac{1}{2} \right); ( 0 ; 4 ); \right\}
 $$
@@
 	\right),$$
-где $p,q,r,s \in \mathbb{N} \cup \{0\}$.
+где $p,q,r,s \in \mathbb{N}$.
 
@@
 \item.
-Время, требуемое для вычисления $F(n)$ разработанными алгоритмами, растёт, как установлено эмпирически, не медленнее, чем $n^4$.
-Так, расчёт для $n \leq 5$ занимает меньше секунды, а для вычисления $F(41)$ при известном (вычисленном ранее) $F(40)$ ушло больше суток.
+Если $ 25 \leq n \leq 42$, то $B_n \subset B_43$.
 
 \item.
-Оптимальные множества для $ 25 \leq n \leq 42$ являются подмножествами оптимального множества для $n=43$ и получаются по конструкции, приведённой в теореме 4, если вместо $A_n$ (перегруз обозначения?) взять 20160 (это не просто так, это HCN - hoghly composite number).
+Время, требуемое для вычисления $F(n)$ разработанными алгоритмами, растёт, как установлено эмпирически, не медленнее, чем $n^4$.
+Так, нахождение для $n \leq 5$ занимает меньше секунды, а для вычисления $F(41)$ при известном (вычисленном ранее) $F(40)$ ушло больше суток.
 
@@
 
-Работа выполнена при поддержке РНФ, грант ???.
+Работа выполнена при поддержке РНФ, грант 16-11-101-25.
+
+\newpage
 

### 2016-05-25 Semenov_SCT_statya.tex
Увеличены поля
Ещё поля
Комментарий по поводу содержательного ТеХ-кода
Косметика
Уменьшено правое поле
Ссылки сделаны чёрными
Увеличен шрифт до 14pt
Правки по оформление под требования МЗ

@@
 \usepackage{amssymb,amsfonts,amsmath,mathtext,cite,enumerate,float,amsthm} %подключаем нужные пакеты расширений
-\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
+\usepackage[pdftex,unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
+%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
 \usepackage{indentfirst} % включить отступ у первого абзаца
@@
 \usepackage{geometry} % Меняем поля страницы. 
-\geometry{left=1cm}% левое поле
+\geometry{left=4cm}% левое поле
 \geometry{right=1cm}% правое поле
-\geometry{top=1cm}% верхнее поле
+\geometry{top=2cm}% верхнее поле
 \geometry{bottom=2cm}% нижнее поле
@@
 
-\begin{document}\large
+\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}
+
+\begin{document}
+
+% !!!
+% Здесь начинается реальный ТеХ-код
+% Всё, что выше - беллетристика
 
@@
 
-Пусть $A=(M_1, M_2, ..., M_n) \in C_n$ для некоторого $n\in N$ и $M_1, M_2, M_3$ не принадлежат прямой.
+Пусть $A=(M_1, M_2, ..., M_n) \in C_n$ для некоторого $n\in \mathbb{N}$ и $M_1, M_2, M_3$ не принадлежат прямой.
 Тогда $n\leq (a+1)(b+1)+3$, где $a=|M_1,M_2|$, $|M_2,M_3|$.
@@
 
-Отсюда, из \ref{ocenka_Stirling} и монотонности $F(n)$ вытекает, что
+Отсюда, из (\ref{ocenka_Stirling}) и монотонности $F(n)$ вытекает, что
 \begin{equation}\label{ocenka_primoryaly}
@@
 $$
-$D(k)$ --- количество делителей числа $k$ (дивизор-функция Рамануджана),
-С другой стороны, в \cite{Ramanujan} (формулы 198-200) показано, что $p(n)$
+$D(k)$ --- количество делителей числа $k$ (дивизор-функция Рамануджана).
+В \cite{Ramanujan} (формулы 198-200) показано, что $p(n)$
 растёт медленнее степенной функции.
@@
 $$
-B_4=\left\{( 0 ; 0 ); ( 0 ; 1 ); \left( \frac{\sqrt{15}}{2} ; \frac{1}{2} \right); ( 0 ; 4 ); \right\}
+B_4=\left\{( 0 ; 0 ); ( 1 ; 0 ); ( 4 ; 0 ); \left(  \frac{1}{2}  ; \frac{\sqrt{15}}{2} \right) \right\}
 $$
@@
 \begin{multline*}
-\{( 0 ; 0 ); ( 153 ; 0 ); ( 144 ; 0 ); ( 130 ; 0 ); ( 115 ; 0 ); ( 111 ; 0 ); ( 104 ; 0 ); ( 98 ; 0 ); ( 88 ; 0 ); ( 76 ; 0 );\\
-( 66 ; 0 ); ( 60 ; 0 ); ( 53 ; 0 ); ( 49 ; 0 ); ( 34 ; 0 ); ( 20 ; 0 ); ( 11 ; 0 ); ( 82 ; \sqrt{2880})\}
+\{( 0 ; 0 ); ( 153 ; 0 ); ( 144 ; 0 ); ( 130 ; 0 ); ( 115 ; 0 ); ( 111 ; 0 );\\
+( 104 ; 0 ); ( 98 ; 0 ); ( 88 ; 0 ); ( 76 ; 0 ); ( 66 ; 0 ); ( 60 ; 0 ); ( 53 ; 0 );\\
+( 49 ; 0 ); ( 34 ; 0 ); ( 20 ; 0 ); ( 11 ; 0 ); ( 82 ; \sqrt{2880})\}
 \end{multline*}
@@
 \begin{multline*}
-\{( 0 ; 0 ); ( 153 ; 0 ); ( 134 ; 0 ); ( 121 ; 0 ); ( 104 ; 0 ); ( 98 ; 0 ); ( 93 ; 0 ); ( 85 ; 0 ); ( 76 ; 0 ); ( 69 ; 0 );\\
-( 65 ; 0 ); ( 58 ; 0 ); ( 49 ; 0 ); ( 41 ; 0 ); ( 36 ; 0 ); ( 30 ; 0 ); ( 13 ; 0 ); ( 67 ; \sqrt{1440} )\}
+\{( 0 ; 0 ); ( 153 ; 0 ); ( 134 ; 0 ); ( 121 ; 0 ); ( 104 ; 0 ); ( 98 ; 0 );\\
+( 93 ; 0 ); ( 85 ; 0 ); ( 76 ; 0 ); ( 69 ; 0 ); ( 65 ; 0 ); ( 58 ; 0 ); ( 49 ; 0 );\\
+( 41 ; 0 ); ( 36 ; 0 ); ( 30 ; 0 ); ( 13 ; 0 ); ( 67 ; \sqrt{1440} )\}
 \end{multline*}
@@
 \item.
-Если $ 25 \leq n \leq 42$, то $B_n \subset B_43$.
+Если $ 25 \leq n \leq 42$, то $B_n \subset B_{43}$.
 
@@
 Время, требуемое для вычисления $F(n)$ разработанными алгоритмами, растёт, как установлено эмпирически, не медленнее, чем $n^4$.
-Так, нахождение для $n \leq 5$ занимает меньше секунды, а для вычисления $F(41)$ при известном (вычисленном ранее) $F(40)$ ушло больше суток.
+Так, нахождение $F(n)$ для $n \leq 5$ занимает менее секунды, а для вычисления $F(41)$ при известном (вычисленном ранее) $F(40)$ потребовалось больше суток.
 
@@
 
-Работа выполнена при поддержке РНФ, грант 16-11-101-25.
+Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-101-25.
 

### 2016-05-25 Semenov_Avdeev_SCT.tex
Версия по образцу МЗ

--- 2016-05-25 Semenov_SCT_statya.tex
@@
-\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
+\documentclass{article}
+
 \usepackage[T2A]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english,russian]{babel} %используем русский и английский языки с переносами
-\usepackage{amssymb,amsfonts,amsmath,mathtext,cite,enumerate,float,amsthm} %подключаем нужные пакеты расширений
-\usepackage[pdftex,unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
-%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
-\usepackage{indentfirst} % включить отступ у первого абзаца
-\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
-\graphicspath{{illustr/}}%путь к рисункам
+\usepackage[cp1251]{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage[tbtags]{amsmath}
+\usepackage{amsfonts,amssymb}
 
-\makeatletter
-\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
-\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"
+\hoffset -10mm
+\voffset -7mm
 
-\usepackage{geometry} % Меняем поля страницы. 
-\geometry{left=4cm}% левое поле
-\geometry{right=1cm}% правое поле
-\geometry{top=2cm}% верхнее поле
-\geometry{bottom=2cm}% нижнее поле
+\usepackage{mathrsfs}
+\usepackage[short]{mz2ewin}
+\usepackage{graphicx}
 
-\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра
+\usepackage{float} % Иначе ругается на таблицу
 
-\sloppy
+\let\rom=\textup
 
-\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}
+\theoremstyle{plain}
 
+\newtheorem{condition}{Условие}
+\newtheorem{lemma}{Лемма}
+\newtheorem{theorem}{Теорема}
+\newtheorem{propos}{Предложение}
+\newtheorem{corol}{Следствие}
+\newtheorem{notation}{Обозначения}
+\newtheorem{definit}{Определение}
+
+\theoremstyle{definition}
+\newtheorem{proof}{Доказательство}\def\theproof{}
+\newtheorem{remark}{Замечание}
+
+\overfullrule5pt
+\mag1300
 \begin{document}
 
-% !!!
-% Здесь начинается реальный ТеХ-код
-% Всё, что выше - беллетристика
+\udk{511.95 + 514.112}
 
-УДК 511.95 + 514.112
+\date{\\
+Исправленный вариант\\ }
 
-\begin{center}
-	\textbf{\LARGE О множествах точек на плоскости с целочисленными расстояниями}
+\author{Н.\,Н.~Авдеев}
+\address{Воронежский госуниверситет}
+\email{avdeev@math.vsu.ru}
 
-	\vspace{15mm}
+\author{Е.\,М.~Семёнов}
+\address{Воронежский госуниверситет}
+%\email{ivanov@gmail.com}
 
-	{\Large{Н.Н. Авдеев, Е.М. Семёнов.}}
+\title{О множествах точек на плоскости с целочисленными расстояниями}
 
-\end{center}
+\maketitle
+\begin{fulltext}
+
+%\begin{abstract}
+
+%\end{abstract}
+
+\footnotetext{Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-101-25.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\paragraph{Теорема 1.} Пусть $\{M_1,M_2\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N}$ --- множество натуральных чисел.
+\begin{theorem}
+	Пусть $\{M_1,M_2\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N}$ --- множество натуральных чисел.
 Тогда найдется такая прямая на плоскости $l$, что $M_i\in l$ для всех $i \in \mathbb{N}$.
+\end{theorem}
 
@@
 
-\paragraph{Лемма 2.}
-
+\begin{lemma}
 	Пусть $A=(M_1, M_2, ..., M_n) \in C_n$ для некоторого $n\in \mathbb{N}$ и $M_1, M_2, M_3$ не принадлежат прямой.
-Тогда $n\leq (a+1)(b+1)+3$, где $a=|M_1,M_2|$, $|M_2,M_3|$.
+	Тогда $n\leq (a+1)(b+1)+3$, где $a=|M_1,M_2|$, $b=|M_2,M_3|$.
+\end{lemma}
 
@@
 
-\paragraph{Лемма 3.}
-Пусть $m\in\mathbb{N}, m\geq 4$, последовательность $(M_1,M_2,...,M_{2m^2+1})$ принадлежит $C_{2m^2+1}$ и содержится в квадрате со стороной $d$.
+\begin{lemma}
+	Пусть $m\in\mathbb{N}, m\geq 4$, последовательность $(M_1,M_2,...,M_{2m^2+1})$ принадлежит $C_{2m^2+1}$ и
+	содержится в квадрате со стороной $d$.
 	Тогда $d>\frac{1}{2}m^2$.
+\end{lemma}
 
-\paragraph{Доказательство.} Разобьём квадрат со стороной $d$ на $m^2$ квадратов со стороной $\frac{d}{m}$.
+\begin{proof}
+	Разобьём квадрат со стороной $d$ на $m^2$ квадратов со стороной $\frac{d}{m}$.
 	Тогда по крайней мере один из маленьких квадратов содержит некоторые три точки исходной последовательности.
@@
 	Поэтому $\lambda > \frac{1}{2}$ и $d>\frac{m^2}{2}$.
-$\Box$
+\end{proof}
 
@@
 
-\paragraph{Теорема 4.}
+\begin{theorem}
 	Неравенства
@@
 	где $b$ --- константа из неравенств Чебышева, справедливы для всех $n\in\mathbb{N}$, $n \geq 3$.
+\end{theorem}
 
-\paragraph{Доказательство.}
+\begin{proof}
 	Докажем сначала, что
@@
 	Так как
-\begin{equation*}
+	\begin{multline*}
 		|M_I,N|=\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)^\frac{1}{2}=
+\\=
 		\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)^\frac{1}{2}=
 		\frac{c_I+\frac{A_n}{C_I}}{2} \in \mathbb{N},
-\end{equation*}
+	\end{multline*}
 	то все расстояния между точками множества (\ref{konstrukcia_primorialy}) есть целые числа.
@@
 	Для $4 \leq n < 17$ значения $F(n)$ были вычислены на ЭВМ (см. далее).
-$\Box$
+\end{proof}
 
 Заметим, что подобная конструкция не позволяет получить верхнюю оценку $F(n)$ в виде полинома.
-Это вызвано тем, что количество точек в множестве подобного типа (подмножество прямой и точка, на ней не лежащая) ограничено числом $p(2\mu^2)+2$, где $\mu$ --- расстояние от точки до прямой,
+Это вызвано тем, что количество точек в множестве подобного типа (подмножество прямой и точка, на ней не лежащая)
+ограничено числом $p(2\mu^2)+2$, где $\mu$ --- расстояние от точки до прямой,
 $$
@@
 \paragraph{\S 3}
+%\begin{definit}
 	Множество точек $B=\{M_1, M_2, ..., M_n\}$ называется оптимальным, если $B \in C_n$ и $d(B) = F(n)$.
+%\end{definit}
 Ясно, что любое оптимальное множество определяется с точностью до движения.
@@
 		\left\{\pm\frac{\sqrt{r}}{s}\right\}
-	\right),$$
+	\right),
+$$
 где $p,q,r,s \in \mathbb{N}$.
@@
 
-Авторы благодарят проф. Ю. А. Брудного за информацию о работе \cite{Newman} и ценные замечания.
+Авторы благодарят проф. Ю.\,А.~Брудного за информацию о работе \cite{Newman} и ценные замечания.
 
-Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-101-25.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\newpage
+%Замечание
+%\begin{remark} Отметим, что доказательство теоремы \ref{theo1} 
+%может быть распространено на случай ...
+%\end{remark}
 
-\addcontentsline{toc}{chapter}{Литература}
+%Определение
+%\begin{definit} Будем называть ...
+%\end{definit}
+
+%Предложение
+%\begin{propos} Пусть выполняются следующие неравенства...
+%\end{propos}
+
+%\begin{figure}[ht]
+%\centering
+%\includegraphics{1kriv80}
+%\caption{Многообразие точек смены кратности.}
+%\end{figure}
+
+%===============Список литературы==================
+\end{fulltext}
+
 \begin{thebibliography}{99}
 
-\bibitem{Newman} D.J. Newman. A Problem Seminar. Springer - Verlag.1982.
+\Bibitem{Newman}
+\by D.J. Newman
+\book A Problem Seminar
+\publ Springer -- Verlag
+\yr 1982
 
-\bibitem{angem1kurs} Е.М. Семенов, С.Н. Уксусов / Аналитическая геометрия на плоскости. – Воронеж : Воронежский государственный университет, 2016. – 130с.
+\RBibitem{angem1kurs}
+\by Е.М. Семенов, С.Н. Уксусов
+\book Аналитическая геометрия на плоскости
+\publaddr Воронеж
+\publ Воронежский государственный университет
+\yr 2016
+%\totalpages 130
 
-\bibitem{Buhshtab} А.А. Бухштаб / Теория чисел. --- М., 1966.
+\RBibitem{Buhshtab}
+\by А.А. Бухштаб 
+\book Теория чисел
+\publaddr М
+\yr 1966
 
-\bibitem{Ramanujan} Ramanujan S. / Highly composite numbers. Proceedings of the London Mathematical Society, 2, XIV, 1915.
+\Bibitem{Ramanujan}
+\by Ramanujan S.
+\paper Highly composite numbers
+\jour Proceedings of the London Mathematical Society
+\vol {2,~XIV}
+\yr 1915
+
+%\RBibitem{1}
+%\by И.\,Я.~Арефьева, И.\,В.~Волович, Е.\,В.~Писковский
+%\paper Скатывание в модели Хиггса и эллиптические функции
+%\jour Теоретическая и математическая физика
+%\vol 172
+%\issue 1 
+%\yr 2012 
+%\pages 138--154
+
+%\RBibitem{2}
+%\by В.\,П.~Маслов
+%\book Теория возмущений и асимптотические методы
+%\publaddr М.
+%\publ Изд-во Московского университета
+%\yr 1965
+
+%\Bibitem{3}
+%\by B.~Hellfer
+%\paper Theorie spectrale pour des operateurs elliptiques
+%\jour Asterisque
+%\vol 90
+%\pages 1--15
+%\yr 1984
+
+%\RBibitem{4} 
+%\by С.\,Ю.~Доброхотов,
+%А.\,И.~Шафаревич 
+%\paper Квазиклассическое квантование инвариантных изотропных многообразий гамильтоновых систем 
+%\inbook Топологические методы в теории гамильтоновых систем 
+%\publaddr М.
+%\publ Факториал 
+%\yr 1998
+
+%Электронный препринт в arXiv
+%\RBibitem{5} 
+%\by Ph.~Barbe
+%\eprint Approximation of integrals over asymptotic sets with applications 
+%to probability and statistics \eprintinfo 
+%\href{http://arxiv.org/abs/math/0312132}{http://arxiv.org/abs/math/0312132}
+%\yr 2003
+
+%Статья в трудах конференции
+%\RBibitem{6} 
+%\by J.\,R.~Isbell
+%\paper Epimorphisms and dominions
+%\inbook Proceedings of the Conference on Categorical Algebra
+%\publ Springer 
+%\publaddr New York 
+%\yr 1966 
+%\pages 232--246
+
+%Препринт
+%\RBibitem{7} 
+%\by В.\,В.~Кузенов, С.\,В.~Рыжков
+%\preprint Математическая модель взаимодействия лазерных пучков высокой энергии с плазменной мишенью в магнитном поле
+%\preprintinfo Препринт ИПМех РАН, \No 942
+%\publaddr М.
+%\yr 2010
+%\totalpages 57
+
+%\RBibitem{8} 
+%\by K.\,O.~Friedrichs
+%\book Perturbation of spectra in Hilbert space
+%\publaddr Providence
+%\publ American Mathematical Society
+%\yr 1965
 

### 2016-05-29 Semenov_Avdeev_SCT.tex
Какие-то правки (косметические), и да будет СР1251 проклята во веки веков

@@
 
-\footnotetext{Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-101-25.}
+\footnotetext{Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-10125.}
 
@@
 $$
-d(A)=\max\limits_{x,y\in A}|x,y|
+	d(A)=\max\limits_{x,y\in A}|x,y|.
 $$
@@
 	$$
-	2m^2+1 \leq\left(\frac{d}{m}\sqrt{2}+1\right)^2+3+d\sqrt{2}
+		2m^2+1 \leq\left(\frac{d}{m}\sqrt{2}+1\right)^2+3+d\sqrt{2}.
 	$$
@@
 	$$
-	0\leq \left( 2\lambda^2 + \sqrt2 \lambda - 2 \right)m^2 + 2\sqrt2 \lambda m + 3
+		0\leq \left( 2\lambda^2 + \sqrt2 \lambda - 2 \right)m^2 + 2\sqrt2 \lambda m + 3.
 	$$
@@
 Обозначим через $p_i, i\in\mathbb{N}$ простые числа, начиная с 3.
-По теореме Чебышева (\cite{Buhshtab}, теорема 325) $p_i \leq bi\ln(i+1)$ для некоторого $b>0$ и всех $i\in\mathbb{N}$.
+По теореме Чебышева (\cite{Buhshtab}, теорема 325)
+$$
+	p_i \leq bi\ln(i+1)
+$$
+для некоторого $b>0$ и всех $i\in\mathbb{N}$.
 Обозначим
 $$
-A_n=\prod_{i=1}^{n} p_i, n\in\mathbb{N}
+	A_n=\prod_{i=1}^{n} p_i, n\in\mathbb{N}.
 $$
@@
 	b^n \sqrt{2\pi n} \left(\frac{n}{e}\right)^n \left(1+\frac{1}{n}\right) \prod_{i=1}^{n} \ln(i+1)
-	\leq \left(\frac{bn\ln (n+1)}{e}\right)^n
+	\leq \left(\frac{bn\ln (n+1)}{e}\right)^n.
 \end{equation}
 
-\begin{theorem}
+\begin{theorem}\label{sct_main_theorem}
 	Неравенства
@@
 	$$
-		|M_{(1,2,...,n)},M_{\varnothing}|=\frac{1}{2}(A_n-1)-\frac{1}{2}(1-A_n) = A_n-1 < A_n
+		|M_{(1,2,...,n)},M_{\varnothing}|=\frac{1}{2}(A_n-1)-\frac{1}{2}(1-A_n) = A_n-1 < A_n.
 	$$
@@
 	$$
-		F(2^n+1) < A_n
+		F(2^n+1) < A_n.
 	$$
@@
 	$$
-		2m^2+1 \leq n < 2(m+1)^2 +1
+		2m^2+1 \leq n < 2(m+1)^2 +1.
 	$$
@@
 	$$
-		F(n) \geq F\left(2m^2+1\right)\geq d \geq \frac{1}{2}m^2 > \frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} - 1 \right)
+		F(n) \geq F\left(2m^2+1\right)\geq d \geq \frac{1}{2}m^2 > \frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} - 1 \right).
 	$$
@@
 	$$
-		\frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} -1 \right) > \frac{n-5}{8}
+		\frac{1}{2}\left(\left(\frac{n-1}{2}\right)^\frac{1}{2} -1 \right) > \frac{n-5}{8}.
 	$$
@@
 	$$
-		\frac{n-5}{8}<F(n)
+		\frac{n-5}{8}<F(n).
 	$$
@@
 
-\item.
-Были вычислены значения $F(n)$ для $4\leq n \leq 42$.
+\item
+Были вычислены значения $F(n)$ для $4\leq n \leq 43$.
 
@@
 
-\item.
+\item
 Оптимальные множества не принадлежат целочисленной решётке.
@@
 
-\item.
+\item
 В большинстве случаев оптимальное множество с точностью до движения определяется однозначно.
@@
 
-\item.
-Для $N \geq 9$ оптимальное множество имеет вид, описанный в теореме 4:
-$n-1$ точек лежат на оси иксов с целыми координатами и одна точка не принадлежит оси иксов, её вторая координата иррациональна.
+\item
+Для $n \geq 9$ оптимальное множество имеет вид, описанный в теореме \ref{sct_main_theorem}:
+$n-1$ точек лежат на оси абсцисс с целыми координатами и одна точка не принадлежит оси абсцисс, её вторая координата иррациональна.
 Для меньших $n$ эта закономерность не выполняется.
 
-\item.
+\item
 Если $ 25 \leq n \leq 42$, то $B_n \subset B_{43}$.
 
-\item.
+\item
 Время, требуемое для вычисления $F(n)$ разработанными алгоритмами, растёт, как установлено эмпирически, не медленнее, чем $n^4$.

### 2016-09-27 Semenov_Avdeev_SCT.tex
Правки в таблицу
Исправлен диапазон вложенных СЦТ
Добавлен e-mail Е. М. Семёнова
Исправлен повтор в формуле
Исправлена последовательность точек

@@
 \address{Воронежский госуниверситет}
-%\email{ivanov@gmail.com}
+\email{nadezhka_ssm@geophys.vsu.ru}
 
@@
 \begin{theorem}
-	Пусть $\{M_1,M_2\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N}$ --- множество натуральных чисел.
+	Пусть $\{M_1,M_2,...,M_n\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N}$ --- множество натуральных чисел.
 Тогда найдется такая прямая на плоскости $l$, что $M_i\in l$ для всех $i \in \mathbb{N}$.
@@
 	\\=
-		\left(\frac{1}{4}\left( c_I^2 - 2 A_n + \frac{A_n^2}{c_I^2} \right)+A_n\right)^\frac{1}{2}=
+		\left(\frac{1}{4}\left( c_I^2 + 2 A_n + \frac{A_n^2}{c_I^2} \right)\right)^\frac{1}{2}=
 		\frac{c_I+\frac{A_n}{C_I}}{2} \in \mathbb{N},
@@
 \hline
-\textbf{F(n)} &  91 & 104 & 121 & 134 & 153 & 164 & 196 & 212 & 234 & 256 & 286  \\
+\textbf{F(n)} &  91 & 104 & 121 & 134 & 153 & 164 & 196 & 212 & 228 & 244 & 272  \\
 \hline
@@
 \hline
-\textbf{F(n)} & 304 & 338 & 370 & 384 & 414 & 448 & 464 & 494 & 524 & 553 & 578  \\
+\textbf{F(n)} & 288 & 319 & 332 & 364 & 396 & 437 & 464 & 494 & 524 & 553 & 578  \\
 \hline
@@
 \item
-Если $ 25 \leq n \leq 42$, то $B_n \subset B_{43}$.
+Если $ 31 \leq n \leq 42$, то $B_n \subset B_{43}$.
 

### 2016-09-27 Semenov_Avdeev_SCT.tex
Исправлена большая C на маленькую c
Экранировано _ в адресе
Исправлено счётное множество

@@
 \address{Воронежский госуниверситет}
-\email{nadezhka_ssm@geophys.vsu.ru}
+\email{nadezhka\_ssm@geophys.vsu.ru}
 
@@
 \begin{theorem}
-	Пусть $\{M_1,M_2,...,M_n\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N}$ --- множество натуральных чисел.
+	Пусть $\{M_1,M_2,...\}$ --- счётное множество точек на плоскости и расстояние $|M_i,M_j|\in \mathbb{N}$ для всех $1\leq i < j <\infty$, где $\mathbb{N}$ --- множество натуральных чисел.
 Тогда найдется такая прямая на плоскости $l$, что $M_i\in l$ для всех $i \in \mathbb{N}$.
@@
 		\left(\frac{1}{4}\left( c_I^2 + 2 A_n + \frac{A_n^2}{c_I^2} \right)\right)^\frac{1}{2}=
-		\frac{c_I+\frac{A_n}{C_I}}{2} \in \mathbb{N},
+		\frac{c_I+\frac{A_n}{c_I}}{2} \in \mathbb{N},
 	\end{multline*}

### 2018-05-02 vvmsh-2018/ErdoshExamples.tex
Theses to VVMSH-2018

\documentclass{article}
\usepackage[russian]{babel}
\usepackage[cp1251]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\pagestyle{empty} \textwidth=108mm \textheight=165mm
\usepackage{calc,ifthen}

\input{style.tex}

\begin{document}

\tezis{НЕКОТОРЫЕ ПРИМЕРЫ СИСТЕМ ЭРДЁША}
{Работа выполнена за счёт гранта РНФ, проект 16-11-10125}%
{Н.Н. Авдеев}
{\me{Авдеев Н.Н.}}
{(Воронеж, ВГУ)}
{nickkolok@mail.ru}
{Некоторые примеры систем Эрдёша}
{Авдеев Н.Н.}

%\keywords
Ключевые слова:
{системы Эрдёша, комбинаторная геометрия}

Название на английском:
Some examples of Erdos systems
%%%%%% Текст тезисов  %%%%%%%

Системой Эрдёша будем называть множество $\{M_1, ..., M_n\}\subset \mathbb{R}^2$, такое,
что для любых $1 \leq i < j \leq n$ расстояние $|M_i M_j|\in\mathbb{Z}$.
%
В [1] даны оценки на диаметр для таких систем;
для небольших диаметров возможно компьютерное моделирование [2--3].
%
%
При моделировании найдено несколько сотен различных систем Эрдёша.
Ниже приведены наиболее интересные примеры, не имеющие отношения к оценкам на диаметр.

Система содержится в двух параллельных прямых (2 точки на одной, 6 на другой):
$
\left\{
(\pm 21/2, 0);
(\pm 4, 3\sqrt{35}/2);
(\pm 5/2, 0);
(\pm 11/2, 0);
\right\}
$

Система содержится в двух параллельных прямых (3 точки на одной, 4 на другой):
\\
$
\left\{
\left( \pm{17/2} ; 0\right);
\left( {44/17} ; {39\sqrt{35}/34}\right);
\left( -{72/17} ;
{15\sqrt{35}/34}\right);
\right.$ \\ $\left.
\left( -{175/34} ; -{24\sqrt{35}/17}\right);
\left( -{57/34} ; {12\sqrt{35}/17}\right);
%\right.$ \\ $\left.
\left( {57/34} ; -{12\sqrt{35}/17}\right);
\right\}
$

Система содержится в объединении окружности с её центром (7~точек):
$
\left\{
(0,0);(\pm 50, 0); (\pm 14, \pm 48)
\right\}
$

Система содержится в окружности (12 точек):
\\
$
\left\{
\left( \pm{153/2} ; 0\right);
%\left( {153/2} ; 0\right);
%\left( 60 ; {63\sqrt{3}/2}\right);
\left( \pm60 ; {63\sqrt{3}/2}\right);
\left( \pm{65/2} ; {44\sqrt{3}}\right);
%\left( {65/2} ; {44\sqrt{3}}\right);
\right.$ \\ $\left.
\left( \pm 44 ; -{65\sqrt{3}/2}\right);
%\left( 44 ; -{65\sqrt{3}/2}\right);
%\left( {143/2} ; {20\sqrt{3}}\right);
\left( \pm{143/2} ; {20\sqrt{3}}\right);
%\left( {23/2} ; -{40\sqrt{3}}\right);
\left( \pm{23/2} ; -{40\sqrt{3}}\right);
\right\}
$

Система содержится в двух прямых, пересекающихся под острым углом (7~точек):
\\
$
\left\{
\left( \pm30 ; 0\right);
%\left( 30 ; 0\right);
\left( \pm{15/2} ; {25\sqrt{7}/2}\right);
%\left( {15/2} ; {25\sqrt{7}/2}\right);
%\left( {33/2} ; {9\sqrt{7}/2}\right);
\left( 0 ; {10\sqrt{7}/1}\right);
\left( \pm{33/2} ; {9\sqrt{7}/2}\right);
\right\}
$

По результатам моделирования может быть выдвинута

\textbf{Гипотеза.} {\it
	Любая система Эрдёша содержится в одном из следующих множеств:

	а) Две прямые (параллельные или пересекающиеся)

	б) Объединение окружности с её центром
}

Планируется проведение более масштабного компьютерного моделирования.

%%%%  ОФОРМЛЕНИЕ СПИСКА ЛИТЕРАТУРЫ %%%
\liter

1.
Авдеев Н. Н., Семёнов Е. М. О множествах точек на плоскости с целочисленными расстояниями
//Математические заметки. – 2016. – Т. 100. – №. 5. – С. 757-761.

2.
Авдеев Н.Н.
Об отыскании множеств точек на плоскости с целочисленными расстояниями
//В сб.: МКМИТУ - 2016. С. 15-19.

3.
Авдеев Н.Н.
О множествах точек на плоскости с целочисленными расстояниями и алгоритмах их отыскания
//Некоторые вопросы анализа, алгебры, геометрии и математического образования. 2017. № 7-1. С. 9-10.

\end{document}

### 2018-05-02 grid/grid_theorem.tex
Drafts on grid theorem

\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel} %используем русский и английский языки с переносами
\usepackage{amssymb,amsfonts,amsmath,mathtext,cite,enumerate,float,amsthm} %подключаем нужные пакеты расширений
\usepackage[pdftex,unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{indentfirst} % включить отступ у первого абзаца
\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
\graphicspath{{illustr/}}%путь к рисункам

\makeatletter
\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"

\usepackage{geometry} % Меняем поля страницы.
\geometry{left=4cm}% левое поле
\geometry{right=1cm}% правое поле
\geometry{top=2cm}% верхнее поле
\geometry{bottom=2cm}% нижнее поле

\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра

\sloppy

\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}

\begin{document}
\renewcommand{\bibname}{Список цитированной литературы}
\renewcommand\refname{\bibname}
% !!!
% Здесь начинается реальный ТеХ-код
% Всё, что выше - беллетристика

\paragraph{Определение.}
Системой Эрдёша называется множество точек $\{M_1, M_2, ..., M_n\}$ на плоскости, не содержащееся ни в какой прямой,
такое, что для любых $i\neq j$ расстояние $|M_i M_j| \in \mathbb{N}$,
т.е. является натуральным числом.

\paragraph{Лемма 1.}

Для любой системы Эрдёша $S=\{M_1, M_2, ..., M_n\}$ существует такое целое число $m$, не превосходящее её диаметра $diam(S)$,
что можно выбрать систему координат способом, при котором координаты каждой точки системы будут иметь вид
\begin{equation}
	M_i = \left(
		\frac{p_i}{2m}
		;
		\frac{\pm\sqrt{q_i}}{2m}
	\right),
\end{equation}
где $p_i \in \mathbb{Z}, q_i \in \mathbb{N}$.

\paragraph{Доказательство.}
Пусть $|M_1 M_2| = m$, тогда, очевидно, $m \leq diam(S)$.
Введём систему координат так, что $M_1=(-m/2, 0)$, $M_2=(m/2, 0)$
Рассмотрим некоторую другую точку $M_i=(x, y)\in S$.
Обозначим $M_i M_1 = a$, $M_i M_2 = b$.
Тогда $M_i$ лежит на пересечении двух окружностей,
задаваемых уравнениями
\begin{gather}
	\left(-\frac{m}{2} - x\right)^2 + y ^2 = a^2,
\\
	\left( \frac{m}{2} - x\right)^2 + y ^2 = b^2.
\end{gather}

Решив её, имеем
\begin{gather}
	x = \frac{a^2 - b^2}{2 m} = \frac{p_i}{2m},
\\
%	y = \frac{\pm \sqrt{-(a^2 - b^2)^2 + m^2 (2 a^2 + 2 b^2  - m^2)} }{2m}.
	y = \pm\sqrt{a^2 - \left(\frac{m}{2}+x\right)^2} =
	\pm\sqrt{a^2 - \left(\frac{m}{2}+\frac{p_i}{2m}\right)^2} =
	\frac{\pm\sqrt{q_i}}{2m},
\end{gather}
что и требовалось доказать.

\paragraph{Лемма 2.}

Для любой системы Эрдёша $S=\{M_1, M_2, ..., M_n\}$ существует такое целое число $m$, не превосходящее её диаметра $diam(S)$,
и такое натуральное число $q$, свободное от квадратов (ссылка!!),
что можно выбрать систему координат способом, при котором координаты каждой точки системы будут иметь вид
\begin{equation}
	M_i = \left(
		\frac{p_i}{2m}
		;
		\frac{s_i\sqrt{q}}{2m}
	\right),
\end{equation}
где $p_i, s_i \in \mathbb{Z}$.

\paragraph{Доказательство.}
Рассмотрим две точки $M_1, M_2 \in S$.
По лемме 1 их координаты (при введении соответствующей системы координат) выражаются в виде
\begin{equation}
	M_1 = \left(
		\frac{p_1}{2m}
		;
		\frac{\pm\sqrt{q_1}}{2m}
	\right),
	M_2 = \left(
		\frac{p_2}{2m}
		;
		\frac{\pm\sqrt{q_2}}{2m}
	\right),
\end{equation}
где $p_i \in \mathbb{Z}, q_i \in \mathbb{N}$, $i=1,2$.

Тогда
\begin{equation}
	|M_1 M_2| = \frac{1}{2m} \sqrt{p_1^2 + p_2^2 - 2p_1p_2 + q_1 + q_2 - 2\sqrt{q_1 q_2}}
\end{equation}
Для того, чтобы расстояние $|M_1 M_2| \in \mathbb{N}$, необходимо,
чтобы $|M_1 M_2|^2 \in \mathbb{N}$, для чего, в свою очередь, необходимо, чтобы
$p_1^2 + p_2^2 - 2p_1p_2 + q_1 + q_2 - 2\sqrt{q_1 q_2} \in \mathbb{N}$,
что влечёт $\sqrt{q_1 q_2} \in \mathbb{N}$,
что, в свою очередь, требует соотношения
$q_1 = s_1^2 \sqrt{q}$, $q_2 = s_2^2 \sqrt{q}$.
Лемма доказана.

\paragraph{Лемма 3.}
Число $q$ в лемме 2 не зависит от выбора <<опорных>> точек для системы координат в лемме 1.

\paragraph{Доказательство.}
Покажем для простоты, что при замене <<опорной>> точки $M_2$ на $M_3$ число $q$ сохраняется.
Пусть в системе координат с <<опорными>> точками $M_1$ и $M_2$ имеем
$M_1=(x_1, 0)$, $M_2=(x_2, 0)$, $M_3=(x_3, y_3)$,
где $x_1, x_2, x_3$~--- рациональные, $y_3 = \frac{s_3\sqrt{q}}{2m}$ по лемме 2;
и пусть в системе координат с <<опорными>> точками $M_1$ и $M_3$ имеем
$M_1=(\tilde{x}_1, 0)$, $M_3=(\tilde{x}_3, 0)$, $M_2=(\tilde{x}_2, \tilde{y}_2)$,
где $\tilde{x}_1, \tilde{x}_2, \tilde{x}_3$~--- рациональные, $\tilde{y}_2 = \frac{s_2\sqrt{\tilde{q}}}{2\tilde{m}}$ по лемме 2.
Тогда площадь треугольника $M_1 M_2 M_3$ двумя способами выражается как полупроизведение основания на высоту:
\begin{equation}
	\frac{1}{2} |y_3| \cdot |x_1 - x_2|
	=
	\frac{1}{2} |\tilde{y}_2| \cdot |\tilde{x}_1 - \tilde{x}_3|,
\end{equation}
откуда имеем
\begin{equation}
	|y_3|
	=
	|\tilde{y}_2| \frac{ |\tilde{x}_1 - \tilde{x}_3|}{|x_1 - x_2|},
\end{equation}
или, подставив выражения для $\tilde{y}_2$ и $y_3$,
\begin{equation}
	\left|\frac{s_3\sqrt{q}}{2m}\right|
	=
	\left|\frac{s_2\sqrt{\tilde{q}}}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|}{|x_1 - x_2|},
\end{equation}

Сгруппировав иррациональные множители в левой части, а рациональные~--- в правой, получим
\begin{equation}
	\left|\frac{\sqrt{q}}{\sqrt{\tilde{q}}}\right|
	=
	\left|\frac{s_2}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|\cdot|2m|}{|s_3|\cdot|x_1 - x_2|},
\end{equation}
или, что то же самое,
\begin{equation}
	\sqrt{\frac{q}{\tilde{q}}}
	=
	\left|\frac{s_2}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|\cdot|2m|}{|s_3|\cdot|x_1 - x_2|},
\end{equation}

Так как $q$ b $\tilde{q}$ свободны от квадратов, то $\frac{q}{\tilde{q}} = 1$, т.е. $q = \tilde{q}$,
что и требовалось доказать.

\end{document}

### 2018-05-02 mz/sqr-test.js
MZ paper moved to separate folder

@@ -1,0 +1,68 @@
+/*var i2;
+var s;
+var t=Date.now();
+for(var i=1,j=0; i<10000000; i++){
+  i2=i;
+  while(!(i2&3)){
+    i2 = i2 >> 2;
+  }
+  if(i2 & 2 || i2 & 7 === 5){
+    continue;
+  }
+  s = Math.sqrt(i2);
+  j+=( s == Math.floor(s));
+}
+console.log(Date.now()-t,' ',j);
+*/
+var i2;
+var s;
+var t=Date.now();
+for(var i=1,j=0; i<10000000; i++){
+  i2=i;
+  while(!(i2&3)){
+    i2 = i2 >> 2;
+  }
+  if(i2 & 2 || i2 & 7 === 5 || ((i2 & 15) != 1 && (i2 & 15) != 4 && (i2 & 15) != 9)){
+    continue;
+  }
+  s = Math.sqrt(i2);
+  j+=( s == Math.floor(s));
+}
+console.log(Date.now()-t,' ',j);
+
+
+var i2;
+var s;
+var t=Date.now();
+for(var i=1,j=0; i<10000000; i++){
+  i2=i;
+  while(!(i2&3)){
+    i2 = i2 >> 2;
+  }
+  if((i2 & 7) != 1){
+    continue;
+  }
+  s = Math.sqrt(i2);
+  j+=( s == Math.floor(s));
+}
+console.log(Date.now()-t,' ',j);
+
+
+
+/*
+var i2;
+var s;
+var t=Date.now();
+for(var i=1,j=0; i<10000000; i++){
+  i2=i;
+  while(!(i2&3)){
+    i2 = i2 >> 2;
+  }
+  if(i2 & 2 || i2 & 7 === 5){
+    continue;
+  }
+  s = Math.sqrt(i2);
+  j+=( s == Math.floor(s));
+}
+console.log(Date.now()-t);
+*/

### 2018-05-02 grid/grid_theorem.tex
Biblatex
Aux definitions and designations

@@
 \usepackage[english,russian]{babel} %используем русский и английский языки с переносами
-\usepackage{amssymb,amsfonts,amsmath,mathtext,cite,enumerate,float,amsthm} %подключаем нужные пакеты расширений
+\usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
 \usepackage[pdftex,unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
@@
 
+\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+\addbibresource{../common/notmy.bib}
+
 \begin{document}
@@
 т.е. является натуральным числом.
+
+Для дальнейшего описания свойств систем Эрдёша нам потребуются следующие определения и обозначения.
+
+\paragraph{Обозначение.}
+$(n,r)$--решёткой или $\L{}(n,r)$ будем называть решётку \cite{polygons-on-lattices}, построенную на векторах $(1/n, 0)$ и $(0, \sqrt{r}/n)$,
+т.е. множество
+\begin{equation}
+	\L{}(n,r) = \left\{
+		\left( \frac{a}{n}, \frac{b\sqrt{r}}{n} \right):
+		a,b \in \mathbb{Z}
+	\right\}.
+\end{equation}
+
+\paragraph{Обозначение.}
+Площадь треугольника $ABC$ будем обозначать символом $S\triangle ABC$.
+
+\paragraph{Определение.}
+Натуральное число $r$ называется свободным от квадратов, если оно не допускает представления в виде $r = r_1^2 \cdot r_2$,
+где $r_1, r_2$~--- натуральные числа и $r_1 > 1$.
+TODO: ссылка
+
+\paragraph{Обозначение.}
+Через $\mathbb{Q}\sqrt{r}$ будем обозначать множество $\{q\sqrt{r} : q \in \mathbb{Q}, r\in\mathbb{N}\}$
+и подразумевать, что $r$ свободно от квадратов.
 
@@
 
+\printbibliography
+
 \end{document}

### 2018-05-02 grid/grid_theorem.tex
Ref to Bukhstab in def of square-free number

@@
 \paragraph{Определение.}
-Натуральное число $r$ называется свободным от квадратов, если оно не допускает представления в виде $r = r_1^2 \cdot r_2$,
+Натуральное число $r$ называется свободным от квадратов [\cite{Bukhstab-number-theory}, гл. 34, п. 3], если оно не допускает представления в виде $r = r_1^2 \cdot r_2$,
 где $r_1, r_2$~--- натуральные числа и $r_1 > 1$.
-TODO: ссылка
 

### 2018-05-02 grid/grid_theorem.tex
File with my bibliography
Small restructurization

@@
 \addbibresource{../common/notmy.bib}
+\addbibresource{../common/my.bib}
 
@@
 
+\paragraph{Определение.}
+Натуральное число $r$ называется свободным от квадратов [\cite{Bukhstab-number-theory}, гл. 34, п. 3], если оно не допускает представления в виде $r = r_1^2 \cdot r_2$,
+где $r_1, r_2$~--- натуральные числа и $r_1 > 1$.
+
 \paragraph{Обозначение.}
-$(n,r)$--решёткой или $\L{}(n,r)$ будем называть решётку \cite{polygons-on-lattices}, построенную на векторах $(1/n, 0)$ и $(0, \sqrt{r}/n)$,
+$(n,r)$--решёткой или $\L{}(n,r)$ будем называть решётку \cite{polygons-on-lattices},
+построенную на векторах $(1/n, 0)$ и $(0, \sqrt{r}/n)$,
+где $n\in\mathbb{N}$, и $r$ свободно от квадратов,
 т.е. множество
@@
 \paragraph{Обозначение.}
+Через $\mathbb{Q}\sqrt{r}$ будем обозначать множество $\{q\sqrt{r} : q \in \mathbb{Q}, r\in\mathbb{N}\}$
+и подразумевать, что $r$ свободно от квадратов.
+
+\paragraph{Обозначение.}
 Площадь треугольника $ABC$ будем обозначать символом $S\triangle ABC$.
 
-\paragraph{Определение.}
-Натуральное число $r$ называется свободным от квадратов [\cite{Bukhstab-number-theory}, гл. 34, п. 3], если оно не допускает представления в виде $r = r_1^2 \cdot r_2$,
-где $r_1, r_2$~--- натуральные числа и $r_1 > 1$.
+Также нам потребуются следующие утверждения.
 
-\paragraph{Обозначение.}
-Через $\mathbb{Q}\sqrt{r}$ будем обозначать множество $\{q\sqrt{r} : q \in \mathbb{Q}, r\in\mathbb{N}\}$
-и подразумевать, что $r$ свободно от квадратов.
+\paragraph{Утверждение 1.}
+Если $a\in\mathbb{Q}\sqrt{r}$ и $a\in\mathbb{Q}\sqrt{p}$, то $p=r$.
+
+Надо ли это доказывать?
+
+\paragraph{Утверждение 2.}
+Если $A,B,C\in\L{}(n,r)$, то $S\triangle ABC \in \mathbb{Q}\sqrt{r}$.
+
+Это утверждение непосредственно следует из [\cite{polygons-on-lattices}, теорема 3.1].
 

### 2018-05-05 new-linear-lower-bound/linear-lower-bound.tex
Drafts for new linear lower bound

--- 2018-05-02 grid/grid_theorem.tex
@@
 % !!!
-% Здесь начинается реальный ТеХ-код
+% Здесь начинается значащий текст
 % Всё, что выше - беллетристика
@@
 
-Для дальнейшего описания свойств систем Эрдёша нам потребуются следующие определения и обозначения.
-
-\paragraph{Обозначение.}
-$(n,r)$--решёткой или $\L{}(n,r)$ будем называть решётку \cite{polygons-on-lattices}, построенную на векторах $(1/n, 0)$ и $(0, \sqrt{r}/n)$,
-т.е. множество
-\begin{equation}
-	\L{}(n,r) = \left\{
-		\left( \frac{a}{n}, \frac{b\sqrt{r}}{n} \right):
-		a,b \in \mathbb{Z}
-	\right\}.
-\end{equation}
-
-\paragraph{Обозначение.}
-Площадь треугольника $ABC$ будем обозначать символом $S\triangle ABC$.
-
-\paragraph{Определение.}
-Натуральное число $r$ называется свободным от квадратов, если оно не допускает представления в виде $r = r_1^2 \cdot r_2$,
-где $r_1, r_2$~--- натуральные числа и $r_1 > 1$.
-TODO: ссылка
-
-\paragraph{Обозначение.}
-Через $\mathbb{Q}\sqrt{r}$ будем обозначать множество $\{q\sqrt{r} : q \in \mathbb{Q}, r\in\mathbb{N}\}$
-и подразумевать, что $r$ свободно от квадратов.
-
 \paragraph{Лемма 1.}
-
-Для любой системы Эрдёша $S=\{M_1, M_2, ..., M_n\}$ существует такое целое число $m$, не превосходящее её диаметра $diam(S)$,
-что можно выбрать систему координат способом, при котором координаты каждой точки системы будут иметь вид
-\begin{equation}
-	M_i = \left(
-		\frac{p_i}{2m}
-		;
-		\frac{\pm\sqrt{q_i}}{2m}
-	\right),
-\end{equation}
-где $p_i \in \mathbb{Z}, q_i \in \mathbb{N}$.
+В системе Эрдёша $S$ не может быть точек $M_1$, $M_2$, $M_3$, $M_4$,
+лежащих на одной прямой $m$ и таких, что $|M_1 M_2| = |M_3 M_4| = 1$.
 
 \paragraph{Доказательство.}
-Пусть $|M_1 M_2| = m$, тогда, очевидно, $m \leq diam(S)$.
-Введём систему координат так, что $M_1=(-m/2, 0)$, $M_2=(m/2, 0)$
-Рассмотрим некоторую другую точку $M_i=(x, y)\in S$.
-Обозначим $M_i M_1 = a$, $M_i M_2 = b$.
-Тогда $M_i$ лежит на пересечении двух окружностей,
-задаваемых уравнениями
-\begin{gather}
-	\left(-\frac{m}{2} - x\right)^2 + y ^2 = a^2,
-\\
-	\left( \frac{m}{2} - x\right)^2 + y ^2 = b^2.
-\end{gather}
+Обозначим через $m_{12}$ и $m_{34}$ серединные перпендикуляры к отрезкам $M_1 M_2$ и $m_3 M_4$ соответственно.
 
-Решив её, имеем
-\begin{gather}
-	x = \frac{a^2 - b^2}{2 m} = \frac{p_i}{2m},
-\\
-%	y = \frac{\pm \sqrt{-(a^2 - b^2)^2 + m^2 (2 a^2 + 2 b^2  - m^2)} }{2m}.
-	y = \pm\sqrt{a^2 - \left(\frac{m}{2}+x\right)^2} =
-	\pm\sqrt{a^2 - \left(\frac{m}{2}+\frac{p_i}{2m}\right)^2} =
-	\frac{\pm\sqrt{q_i}}{2m},
-\end{gather}
-что и требовалось доказать.
+Пусть точка $M\in S$.
+Тогда либо $|MM_1| - |MM_2| = 0$ и $M\in m_{12}$, либо $\left||MM_1| - |MM_2|\right| = 1$ и $M\in m$,
+т.е. в любом случае $M\in m \cup m_{12}$.
+Аналогично $M\in m \cup m_{34}$.
+Следовательно, $M\in (m \cup m_{12}) \cap (m \cup m_{34}) = m \cup (m_{12} \cap m_{34}) = m$,
+т.к. $m_{12} \cap m_{34} = \varnothing$ (перпендикуляры к одной прямой, проведённые в разных точках, не пересекаются между собой).
 
-\paragraph{Лемма 2.}
+В силу произвольности выбора $M \in S$ получаем $S \subset m$, чего быть не должно.
 
-Для любой системы Эрдёша $S=\{M_1, M_2, ..., M_n\}$ существует такое целое число $m$, не превосходящее её диаметра $diam(S)$,
-и такое натуральное число $q$, свободное от квадратов (ссылка!!),
-что можно выбрать систему координат способом, при котором координаты каждой точки системы будут иметь вид
-\begin{equation}
-	M_i = \left(
-		\frac{p_i}{2m}
-		;
-		\frac{s_i\sqrt{q}}{2m}
-	\right),
-\end{equation}
-где $p_i, s_i \in \mathbb{Z}$.
+\paragraph{Следствие 1.1.}
+Пусть $S$~--- система Эрдёша, $\mathop{diam} S = d$.
+Тогда ни на какой прямой не лежит более $(d+3)/2$ точек из $S$.
 
 \paragraph{Доказательство.}
-Рассмотрим две точки $M_1, M_2 \in S$.
-По лемме 1 их координаты (при введении соответствующей системы координат) выражаются в виде
-\begin{equation}
-	M_1 = \left(
-		\frac{p_1}{2m}
-		;
-		\frac{\pm\sqrt{q_1}}{2m}
-	\right),
-	M_2 = \left(
-		\frac{p_2}{2m}
-		;
-		\frac{\pm\sqrt{q_2}}{2m}
-	\right),
-\end{equation}
-где $p_i \in \mathbb{Z}, q_i \in \mathbb{N}$, $i=1,2$.
+Предположим проивное.
+Очевидно, что на прямой $m$ может быть не более $d+1$ точек системы $S$.
 
-Тогда
-\begin{equation}
-	|M_1 M_2| = \frac{1}{2m} \sqrt{p_1^2 + p_2^2 - 2p_1p_2 + q_1 + q_2 - 2\sqrt{q_1 q_2}}
-\end{equation}
-Для того, чтобы расстояние $|M_1 M_2| \in \mathbb{N}$, необходимо,
-чтобы $|M_1 M_2|^2 \in \mathbb{N}$, для чего, в свою очередь, необходимо, чтобы
-$p_1^2 + p_2^2 - 2p_1p_2 + q_1 + q_2 - 2\sqrt{q_1 q_2} \in \mathbb{N}$,
-что влечёт $\sqrt{q_1 q_2} \in \mathbb{N}$,
-что, в свою очередь, требует соотношения
-$q_1 = s_1^2 \sqrt{q}$, $q_2 = s_2^2 \sqrt{q}$.
-Лемма доказана.
+Пусть сначала $d$ чётно.
+Тогда из каждой пары соседних точек, кроме, быть может, одной пары,
+нужно выбросить хотя бы одну точку в силу леммы 1.
+Итого мы выбросим не менее $d/2$ точек,
+останется не более $(d+2)/2$ точек.
 
-\paragraph{Лемма 3.}
-Число $q$ в лемме 2 не зависит от выбора <<опорных>> точек для системы координат в лемме 1.
+Пусть теперь $d$ нечётно.
+Тогда нужно выбросить не менее $(d-1)/2$ точек.
+Останется не более $(d+3)/2$ точек.
 
-\paragraph{Доказательство.}
-Покажем для простоты, что при замене <<опорной>> точки $M_2$ на $M_3$ число $q$ сохраняется.
-Пусть в системе координат с <<опорными>> точками $M_1$ и $M_2$ имеем
-$M_1=(x_1, 0)$, $M_2=(x_2, 0)$, $M_3=(x_3, y_3)$,
-где $x_1, x_2, x_3$~--- рациональные, $y_3 = \frac{s_3\sqrt{q}}{2m}$ по лемме 2;
-и пусть в системе координат с <<опорными>> точками $M_1$ и $M_3$ имеем
-$M_1=(\tilde{x}_1, 0)$, $M_3=(\tilde{x}_3, 0)$, $M_2=(\tilde{x}_2, \tilde{y}_2)$,
-где $\tilde{x}_1, \tilde{x}_2, \tilde{x}_3$~--- рациональные, $\tilde{y}_2 = \frac{s_2\sqrt{\tilde{q}}}{2\tilde{m}}$ по лемме 2.
-Тогда площадь треугольника $M_1 M_2 M_3$ двумя способами выражается как полупроизведение основания на высоту:
-\begin{equation}
-	\frac{1}{2} |y_3| \cdot |x_1 - x_2|
-	=
-	\frac{1}{2} |\tilde{y}_2| \cdot |\tilde{x}_1 - \tilde{x}_3|,
-\end{equation}
-откуда имеем
-\begin{equation}
-	|y_3|
-	=
-	|\tilde{y}_2| \frac{ |\tilde{x}_1 - \tilde{x}_3|}{|x_1 - x_2|},
-\end{equation}
-или, подставив выражения для $\tilde{y}_2$ и $y_3$,
-\begin{equation}
-	\left|\frac{s_3\sqrt{q}}{2m}\right|
-	=
-	\left|\frac{s_2\sqrt{\tilde{q}}}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|}{|x_1 - x_2|},
-\end{equation}
-
-Сгруппировав иррациональные множители в левой части, а рациональные~--- в правой, получим
-\begin{equation}
-	\left|\frac{\sqrt{q}}{\sqrt{\tilde{q}}}\right|
-	=
-	\left|\frac{s_2}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|\cdot|2m|}{|s_3|\cdot|x_1 - x_2|},
-\end{equation}
-или, что то же самое,
-\begin{equation}
-	\sqrt{\frac{q}{\tilde{q}}}
-	=
-	\left|\frac{s_2}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|\cdot|2m|}{|s_3|\cdot|x_1 - x_2|},
-\end{equation}
-
-Так как $q$ b $\tilde{q}$ свободны от квадратов, то $\frac{q}{\tilde{q}} = 1$, т.е. $q = \tilde{q}$,
-что и требовалось доказать.
-
+\cite{our-mz-rus}
 \printbibliography

### 2018-05-06 grid/grid_theorem.tex
Last drafts of grid theorem (not useful anymore)

@@
 
+В дальнейшем, говоря о системе координат,
+мы будем подразумевать прямоугольную декартову систему координат,
+в которой расстояния сохраняются.
+
+(TODO: как эту мысль выразить по-человечеки?!)
+
 \paragraph{Лемма 1.}
@@
 \\
-	\left( \frac{m}{2} - x\right)^2 + y ^2 = b^2.
+	\left( \frac{m}{2} - x\right)^2 + y ^2 = b^2,
 \end{gather}
+где $a+b\geq m$.
 
@@
 Для любой системы Эрдёша $S=\{M_1, M_2, ..., M_n\}$ существует такое целое число $m$, не превосходящее её диаметра $diam(S)$,
-и такое натуральное число $q$, свободное от квадратов (ссылка!!),
+и такое натуральное число $q$, свободное от квадратов,
 что можно выбрать систему координат способом, при котором координаты каждой точки системы будут иметь вид
-\begin{equation}
+\begin{equation}\label{grid_for_Erdosh_system}
 	M_i = \left(
@@
 
-\paragraph{Лемма 3.}
-Число $q$ в лемме 2 не зависит от выбора <<опорных>> точек для системы координат в лемме 1.
+\paragraph{Теорема 1.}
+Любая система Эрдёша $S$ может быть помещена на $\L{}(n,q)$ в некоторой системе координат,
+причём $q$ не зависит от выбора системы координат.
 
 \paragraph{Доказательство.}
-Покажем для простоты, что при замене <<опорной>> точки $M_2$ на $M_3$ число $q$ сохраняется.
-Пусть в системе координат с <<опорными>> точками $M_1$ и $M_2$ имеем
-$M_1=(x_1, 0)$, $M_2=(x_2, 0)$, $M_3=(x_3, y_3)$,
-где $x_1, x_2, x_3$~--- рациональные, $y_3 = \frac{s_3\sqrt{q}}{2m}$ по лемме 2;
-и пусть в системе координат с <<опорными>> точками $M_1$ и $M_3$ имеем
-$M_1=(\tilde{x}_1, 0)$, $M_3=(\tilde{x}_3, 0)$, $M_2=(\tilde{x}_2, \tilde{y}_2)$,
-где $\tilde{x}_1, \tilde{x}_2, \tilde{x}_3$~--- рациональные, $\tilde{y}_2 = \frac{s_2\sqrt{\tilde{q}}}{2\tilde{m}}$ по лемме 2.
-Тогда площадь треугольника $M_1 M_2 M_3$ двумя способами выражается как полупроизведение основания на высоту:
-\begin{equation}
-	\frac{1}{2} |y_3| \cdot |x_1 - x_2|
-	=
-	\frac{1}{2} |\tilde{y}_2| \cdot |\tilde{x}_1 - \tilde{x}_3|,
-\end{equation}
-откуда имеем
-\begin{equation}
-	|y_3|
-	=
-	|\tilde{y}_2| \frac{ |\tilde{x}_1 - \tilde{x}_3|}{|x_1 - x_2|},
-\end{equation}
-или, подставив выражения для $\tilde{y}_2$ и $y_3$,
-\begin{equation}
-	\left|\frac{s_3\sqrt{q}}{2m}\right|
-	=
-	\left|\frac{s_2\sqrt{\tilde{q}}}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|}{|x_1 - x_2|},
-\end{equation}
+Существование $n$ и $q$ следует из леммы 2.
+Покажем, что такое $q$ единственно.
+Пусть нашлась система координат, в которой $S \subset \L{}(m,p)$.
+Пусть $M_1, M_2, M_3 \in S$ и не лежат на одной прямой.
+Тогда по утверждению 2 $S\triangle M_1 M_2 M_3 \in \mathbb{Q}\sqrt{q}$
+и $S\triangle M_1 M_2 M_3 \in \mathbb{Q}\sqrt{p}$,
+откуда в силу утверждения 1 $p=q$,
+что и требовалось доказать.
 
-Сгруппировав иррациональные множители в левой части, а рациональные~--- в правой, получим
-\begin{equation}
-	\left|\frac{\sqrt{q}}{\sqrt{\tilde{q}}}\right|
-	=
-	\left|\frac{s_2}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|\cdot|2m|}{|s_3|\cdot|x_1 - x_2|},
-\end{equation}
-или, что то же самое,
-\begin{equation}
-	\sqrt{\frac{q}{\tilde{q}}}
-	=
-	\left|\frac{s_2}{2\tilde{m}}\right| \frac{ |\tilde{x}_1 - \tilde{x}_3|\cdot|2m|}{|s_3|\cdot|x_1 - x_2|},
-\end{equation}
+\paragraph{Замечание.}
+$n$, вообще говоря, определяется не единственным образом.
 
-Так как $q$ b $\tilde{q}$ свободны от квадратов, то $\frac{q}{\tilde{q}} = 1$, т.е. $q = \tilde{q}$,
-что и требовалось доказать.
+\paragraph{Определение.}
+Число $q$ из предыдущей теоремы будем называть орнаментом системы Эрдёша $S$.
+
+TODO: хороший ли это термин? На английский переводится легко (ornament).
+
+Таким образом, у системы Эрдёша, наряду с двумя наиболее очевидными характеристиками~--- диаметром и мощностью~\cite{our-mz-rus}~---
+%TODO: english mz paper
+существует ещё одна характеристика, инвариантная относительно выбора системы координат~--- орнамент.
+Следует отметить, что в отличие от диаметра и мощности,
+орнамент есть локальная характеристика в том смысле, что при рассмотрении подсистемы
+(т.е. системы Эрдёша $S'$, являющейся подмножеством данной системы $S$)
+орнамент сохраняется.
+С точки зрения геометрического смысла, орнамент~---
+это минимальное натуральное число, с корнем из которого соизмерима
+площадь любого треугольника, вершины которого принадлежат данной системе
+(говоря точнее, из (\ref{grid_for_Erdosh_system}) следует,
+что площадь такого треугольника кратна $\sqrt{q}/4$).
+
+Примеры систем Эрдёша с различным орнаментом можно найти, например,
+в \cite{our-vvmsh-2018,our-mkmitu-2016}
+
+Возникает закономерный вопрос о соотношении орнамента, диаметра и мощности.
+
+Из леммы 1 непосредственно вытекает
+
+\paragraph{Следствие 1.}
+Если диаметр системы Эрдёша равен $m$,
+то её орнамент принадлежит множеству
+\begin{multline}
+	\mathbb{N} \cap \{
+		-(m+a+b)(m-a-b)(m+a-b)(m-a+b)/c
+	\\
+	\mbox{где }
+		a,b,c\in\mathbb{N}; a \leq m; b\leq m; a+b > m
+	\},
+	%\right\}
+\end{multline}
+
+\paragraph{Следствие 2.}
+Для диаметра $m$ существует не более $(m^2 + m)/2$ возможных орнаментов.
+
+TODO: доказательство.
+
+TODO: улучшить оценку из соображений симметрии предположительно до $(m^2+3m)/4 - 1$.
+
+\paragraph{Следствие 3.}
+Если диаметр системы Эрдёша равен $m$, то её орнамент строго меньше $12m^4$.
+
+TODO: доказательство.
+
+Обратной оценки, т.е. оценки на диаметр, зависящей от орнамента, нет.
+(Достаточно заметить, что при растяжении системы в целое количество раз её орнамент не изменяется.)
+
+Нет и подобной взаимосвязи между орнаментом и мощностью, а именно верна следующая
+\paragraph{Теорема 2.}
+Для любых наперёд заданных натурального $p\leq 3$ и свободного от квадратов $q$
+существует система Эрдёша с мощностью $p$ и орнаментом $q$.
+
+TODO: доказательство (конструктивное).
+
+Ещё сюда можно добавить табличку как в \cite{our-mz-rus},
+но с орнаментами оптимальных систем.
 

### 2018-05-06 new-linear-lower-bound/linear-lower-bound.tex
Finishing the draft of new linear lower bound

@@
 
+Avdeev N.N, Semenov E.M.
+
+Better linear lower bound for integral point sets
+
+Это НЕ статья, это просто доказательство.
+Имеются вольности языка.
+
 \paragraph{Определение.}
@@
 
-\cite{our-mz-rus}
+\paragraph{Лемма 2.}
+Систему Эрдёша $S$ диаметром $d$ можно вместить в квадрат со стороной $d$.
+
+\paragraph{Доказательство.}
+Пусть $M_1, M_2 \in S$, $|M_1 M_2| = d$,
+т.е. диаметр системы $S$ достигается на $M_1$  $M_2$.
+Введём прямоугольную декартову систему координат на плоскости таким образом, что
+$M_1 = (-d/2; 0)$, $M_1 = (d/2; 0)$.
+
+Тогда прочие точки $M_i \in S$ имеют координаты $M_i=(x_i, y_i)$.
+Очевидно, что для $i>2$ имеем $|x_i| < d/2$
+(иначе $|M_i M_1| > d$ или $|M_i M_2| > d$).
+Более того, $|y_i| < d$ (иначе $|M_i M_1| > d$ или $|M_i M_2| > d$).
+
+Пусть $y_+ = \max_{i} y_i$, $y_- = \min_{i} y_i$, $M_+=(x_+, y_+)$, $M_-=(x_-, y_-)$
+(если максимум досигается на нескольких точках, возьмём любые).
+Тогда
+\begin{multline}
+	d \geq |M_+ M_-| = \sqrt{(x_+ - x_-)^2 + (y_+ - y_-)^2}
+	\geq \sqrt{(y_+ - y_-)^2} = y_+ - y_-
+\end{multline}
+Отсуюда следует, что $S$ можно вместить в квадрат со сторонами $x=\pm d/2$,
+$y=y_+$, $y=y_+ - d$.
+
+\paragraph{Примечание.}
+Скорее всего, улучшить оценку на сторону квадрата, сделав её меньше диаметра,
+не получится: часто встречаются конструкции систем Эрдёша, расположенных на окружности
+(Ссылка!!!).
+
+\paragraph{Ремарка.}
+Терминология местами диковата,
+но я на неё не особо заморачиваюсь,
+потому что всё равно потом переводить на английский.
+
+\paragraph{Определение.}
+Крестом точек $M_1$ и $M_2$ будем называть объединение прямой,
+проходящей через эти точки,
+и серединного перпендикуляра к отрезку $M_1 M_2$
+и обозначать $cr(M_1,M_2)$.
+
+\paragraph{Ремарка.}
+Возможно, в финальной версии мы крест уберём, но пока пусть будет.
+Без него формулировать утверждение 3 и лемму 4 вообще как-то грустно.
+
+\paragraph{Обозначение.}
+Количесво точек в множестве $P$ будем обозначать как $\#P$
+и считать его равным $\infty$, если $P$ не менее чем счётно.
+
+Вроде как в англоязычной литературе это стандартное обозначение.
+
+\paragraph{Утверждение 3.}
+Если интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются,
+$M_1 \neq M_2$, $M_3 \neq M_4$,
+то $cr(M_1,M_2) \cap cr(M_3,M_4)$~--- либо не более 4 точек, либо прямая.
+
+\paragraph{Доказательство.}
+Пересечением $cr(M_1,M_2) \cap cr(M_3,M_4)$ может быть либо от 2 до 4 точек, либо прямая,
+либо, если эти кресты совпадают, их пересечение совпадает с каждым из эих крестов.
+Но $cr(M_1,M_2) \neq cr(M_3,M_4)$, потому что середины отрезков $M_1 M_2$ и $M_3 M_4$
+совпасть не могут, т.к. интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются по условию.
+
+\paragraph{Лемма 4.}
+Пусть $S$~--- система Эрдёша,
+$M_1, M_2, M_3, M_4 \in S$,
+$M_1 \neq M_2$, $M_3 \neq M_4$,
+интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются,
+$d = \mathop{diam} S > 5$.
+Тогда $\#S \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2}$.
+
+\paragraph{Доказательство.}
+Если $\#(cr(M_1, M_2) \cap cr(M_3, M_4)) < \infty$,
+то (ссылка на Солимоси!)
+$\#S \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
+Иначе (в силу утверждения 3) $cr(M_1, M_2) \cap cr(M_3, M_4) = m$,
+где $m$~--- прямая.
+В силу следствия 1.1 на этой прямой лежит не более $(d+3)/2$ точек;
+кроме того, из общего количества точек нужно вычесть 4,
+которые бы дали эти 2 креста в случае дискретного переечения.
+Получаем верхнюю оценку
+\begin{equation}
+	\#S \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| - 4 + \frac{d+3}{2}
+	=
+	4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2},
+\end{equation}
+что и требовалось доказать.
+
+\paragraph{Обозначение.}
+Через $\varphi_k$ будем обозначать максимально возможный минимум попарных расстояний между $k$ точками,
+упакованными в единичный квадрат (ссылка!!).
+
+На английском это есть в статье и звучит как-то адекватно.
+
+\paragraph{Лемма 5.}
+Пусть $S$~--- система Эрдёша,
+$d = \mathop{diam} S > 5$,
+$k \geq 2$,
+$m \geq 2$,
+$ \#S = (k-1)m^2 + 2$.
+
+Тогда
+\begin{equation}
+	d \geq \mu (\#S - 2),
+\end{equation}
+где
+\begin{equation}
+	\mu = \frac{\sqrt{64\varphi_k^2 (k-1)+1}-1}{16\varphi_k^2 (k-1)}
+\end{equation}
+
+\paragraph{Доказательство.}
+Впишем $S$ в квадрат со стороной $d$ (см. лемму 2)
+и разобьём этот квадрат на $m^2$ маленьких квадратов со стороной $d/m$.
+Тогда по принципу Дирихле либо:
+
+а) хотя бы в два маленьких квадрата $Q_1$ и $Q_2$ попало не менее чем по $k$ точек;
+
+б) хотя бы в один маленький квадрат $Q_1$ попало не менее чем $k+1$ точка.
+\\
+В случае (а) в $Q_1$ выберем точки $M_1$ и $M_2$,
+а в $Q_2$ выберем точки $M_3$ и $M_4$
+так, что $|M_1 M_2| \leq \varphi_k d /m$, $|M_3 M_4| \leq \varphi_k d/m$
+(это всегда можно сделать в силу определения $\varphi_k$).
+
+В случае (б) в $Q_1$ выберем точки $M_1$ и $M_2$ так, что
+$|M_1 M_2| \leq \varphi_{k+1} d /m \leq \varphi_k d /m$.
+Из всех точек, кроме $M_1$, выберем $M_3$ и $M_4$ так, что
+$|M_3 M_4| \leq \varphi_k d/m$.
+Возможное совпадение $M_2$ и $M_3$ не помешает дальнейшему доказательству.
+
+По лемме 4
+\begin{equation}
+	\#S \leq 4 \left( \frac{d}{m} \varphi_k \right)^2 + \frac{d}{2} - \frac{5}{2}
+\end{equation}
+или, с учётом того, что $ \#S = (k-1)m^2 + 2$,
+\begin{equation}
+	 4 \left( \frac{d}{m} \varphi_k \right)^2 + \frac{d}{2} - \left( (k-1)m^2 + \frac{9}{2}\right) \geq 0
+\end{equation}
+
+Положим $d = \lambda m$:
+\begin{equation}
+	 4 \lambda^2 \varphi_k^2 + \frac{m}{2} \lambda - \left( (k-1)m^2 + \frac{9}{2}\right) \geq 0
+\end{equation}
+
+Решим это квадратное неравенство относительно $\lambda$,
+зная, что $\lambda > 0$.
+
+Честно считаем дискриминант:
+\begin{multline}
+	D =
+	\frac{m^2}{4} + 4 \cdot 4 \varphi_k^2 \cdot \left( (k-1)m^2 + \frac{9}{2}\right)
+	=
+	\frac{m^2}{4} + 16 \varphi_k^2 \cdot \left( (k-1)m^2 + \frac{9}{2}\right)
+	>\\>
+	\frac{m^2}{4} + 16 \varphi_k^2 \cdot (k-1)m^2
+	=
+	\frac{m^2}{4} \cdot (1 + 64 \varphi_k^2 \cdot (k-1))
+	> 0
+\end{multline}
+
+Итак,
+\begin{multline}
+	\frac{d}{m} = \lambda >
+	\frac{-\frac{m}{2} + \sqrt{\frac{m^2}{4} \cdot (1 + 64 \varphi_k^2 \cdot (k-1))} }{8 \varphi_k^2}
+	=\\=
+	\frac{-m + \sqrt{m^2  (1 + 64 \varphi_k^2 \cdot (k-1))} }{16 \varphi_k^2}
+	=
+	m\frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2}
+\end{multline}
+Разделим обе части неравенства на положительное число $m(k-1)$:
+\begin{equation}
+	\frac{d}{(k-1)m^2}
+	>
+	\frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2 (k-1)}
+\end{equation}
+Положим
+\begin{equation}
+	\mu = \frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2 (k-1)},
+\end{equation}
+тогда
+\begin{equation}
+	\frac{d}{(k-1)m^2}
+	>
+	\mu
+\end{equation}
+откуда незамедлительно
+\begin{equation}
+	d > \mu \cdot (\#S-2)
+\end{equation}
+что и требовалось доказать.
+
+Какие-то умные и добрые люди вычислили $\varphi_k$ до $k=10$ включительно (ссылка!) с точностью 4 знака после запятой.
+(Умные, потому что вычислили, а добрые, потому что выложили.)
+
+По их расчётам получаем $\varphi_{10} = 0.4214$,
+что даёт оценку $d>0.3583(\#S-2)$.
+
+Это, конечно, для $\#S$ специального вида, но множитель $(m+1)^2 / m^2$ съестся на бесконечности,
+а итоговые коэфициенты и так будем брать с запасом (кто их там знает, как они округляли).
+
+Это для $\#S \geq 41$, но для меньших мощностей посчитано (и не один раз, в том числе нами).
+
+Кстати, для $\#S = 3$ не проходит (и не должно).
+
+А особенная приятность в том, что вычисление новых $\phi_k$ может давать автоматическое улучшение оценки
+(может, правда, и не давать).
+Однако, как ни печально, константу не получится сделать даже 0.42 (есть нижняя оценка на $\varphi_k$).
+
+Следующий прикол:
+де-факто нам нужна не какая попало расстановка точек в единичном квадрате,
+а расстановка с рациональными расстояниями (они потенциально могут превратиться в целые при умножении на $d/m$).
+А возможность аппроксимировать произвольную расстановку расстановкой с рациональными расстояниями
+уже имеет прямое отношение к проблеме Улама-Эрдёша (ссылка!!).
+
+Просто чтобы список литературы был непуст: \cite{our-mz-rus}
 \printbibliography

### 2018-08-17 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Better liear bound included as is

--- 2018-05-06 new-linear-lower-bound/linear-lower-bound.tex
@@
-\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
-\usepackage[T2A]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[english,russian]{babel} %используем русский и английский языки с переносами
-\usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
-\usepackage[pdftex,unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
-%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
-\usepackage{indentfirst} % включить отступ у первого абзаца
-\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
-\graphicspath{{illustr/}}%путь к рисункам
-
-\makeatletter
-\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
-\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"
-
-\usepackage{geometry} % Меняем поля страницы.
-\geometry{left=4cm}% левое поле
-\geometry{right=1cm}% правое поле
-\geometry{top=2cm}% верхнее поле
-\geometry{bottom=2cm}% нижнее поле
-
-\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра
-
-\sloppy
-
-\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}
-
-\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
-\addbibresource{../common/notmy.bib}
-\addbibresource{../common/my.bib}
-
-\begin{document}
-\renewcommand{\bibname}{Список цитированной литературы}
-\renewcommand\refname{\bibname}
-% !!!
-% Здесь начинается значащий текст
-% Всё, что выше - беллетристика
-
-Avdeev N.N, Semenov E.M.
-
-Better linear lower bound for integral point sets
-
-Это НЕ статья, это просто доказательство.
-Имеются вольности языка.
-
 \paragraph{Определение.}
@@
 уже имеет прямое отношение к проблеме Улама-Эрдёша (ссылка!!).
-
-Просто чтобы список литературы был непуст: \cite{our-mz-rus}
-\printbibliography
-
-\end{document}

### 2018-08-19 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Linear bound beautified a bit: amsthm, refs

@@
-\paragraph{Определение.}
-Системой Эрдёша называется множество точек $\{M_1, M_2, ..., M_n\}$ на плоскости, не содержащееся ни в какой прямой,
-такое, что для любых $i\neq j$ расстояние $|M_i M_j| \in \mathbb{N}$,
-т.е. является натуральным числом.
+\begin{lemma}
+	\label{lemma_points_on_line}
+	Пусть $S\in\mathfrak{M}_n$ для некоторого $n \geqslant 3$,
+	$\{M_1, M_2, M_3, M_4\} \subset S$, как минимум три из этих точек различны,
+	$|M_1 M_2| = |M_3 M_4| = 1$.
+	Тогда никакая прямая $m$ не содержит точки $M_1$, $M_2$, $M_3$ и $M_4$ одновременно.
+\end{lemma}
 
-\paragraph{Лемма 1.}
-В системе Эрдёша $S$ не может быть точек $M_1$, $M_2$, $M_3$, $M_4$,
-лежащих на одной прямой $m$ и таких, что $|M_1 M_2| = |M_3 M_4| = 1$.
-
-\paragraph{Доказательство.}
-Обозначим через $m_{12}$ и $m_{34}$ серединные перпендикуляры к отрезкам $M_1 M_2$ и $m_3 M_4$ соответственно.
+\begin{proof}
+	Предположим противное.
+	Обозначим через $m_{12}$ и $m_{34}$ серединные перпендикуляры к отрезкам $M_1 M_2$ и $M_3 M_4$ соответственно.
 
 	Пусть точка $M\in S$.
-Тогда либо $|MM_1| - |MM_2| = 0$ и $M\in m_{12}$, либо $\left||MM_1| - |MM_2|\right| = 1$ и $M\in m$,
+	Тогда либо $|MM_1| - |MM_2| = 0$ и $M\in m_{12}$, либо $\Bigl||MM_1| - |MM_2|\Bigr| = 1$ и $M\in m$,
 	т.е. в любом случае $M\in m \cup m_{12}$.
@@
 
-В силу произвольности выбора $M \in S$ получаем $S \subset m$, чего быть не должно.
+	В силу произвольности выбора $M \in S$ получаем $S \subset m$, что противоречит $S\in\mathfrak{M}_n$.
+\end{proof}
 
-\paragraph{Следствие 1.1.}
+\begin{corollary}
+	\label{corollary:max_points_on_line}
 	Пусть $S$~--- система Эрдёша, $\mathop{diam} S = d$.
 	Тогда ни на какой прямой не лежит более $(d+3)/2$ точек из $S$.
+\end{corollary}
 
-\paragraph{Доказательство.}
-Предположим проивное.
+\begin{proof}
+	Предположим противное.
 	Очевидно, что на прямой $m$ может быть не более $d+1$ точек системы $S$.
@@
 	Тогда из каждой пары соседних точек, кроме, быть может, одной пары,
-нужно выбросить хотя бы одну точку в силу леммы 1.
+	нужно выбросить хотя бы одну точку в силу леммы~\ref{lemma_points_on_line}.
 	Итого мы выбросим не менее $d/2$ точек,
@@
 	Останется не более $(d+3)/2$ точек.
+\end{proof}
 
-\paragraph{Лемма 2.}
+\begin{lemma}
+	\label{lemma:square_container}
 	Систему Эрдёша $S$ диаметром $d$ можно вместить в квадрат со стороной $d$.
+\end{lemma}
 
-\paragraph{Доказательство.}
+\begin{proof}
 	Пусть $M_1, M_2 \in S$, $|M_1 M_2| = d$,
@@
 	\end{multline}
-Отсуюда следует, что $S$ можно вместить в квадрат со сторонами $x=\pm d/2$,
+	Следовательно, $S$ можно вписать в квадрат со сторонами $x=\pm d/2$,
 	$y=y_+$, $y=y_+ - d$.
+\end{proof}
 
-\paragraph{Примечание.}
+\begin{remark}
 	Скорее всего, улучшить оценку на сторону квадрата, сделав её меньше диаметра,
@@
 	(Ссылка!!!).
+\end{remark}
 
-\paragraph{Ремарка.}
-Терминология местами диковата,
-но я на неё не особо заморачиваюсь,
-потому что всё равно потом переводить на английский.
-
-\paragraph{Определение.}
+\begin{definition}
 	Крестом точек $M_1$ и $M_2$ будем называть объединение прямой,
@@
 	и обозначать $cr(M_1,M_2)$.
+\end{definition}
 
-\paragraph{Ремарка.}
-Возможно, в финальной версии мы крест уберём, но пока пусть будет.
-Без него формулировать утверждение 3 и лемму 4 вообще как-то грустно.
-
-\paragraph{Обозначение.}
-Количесво точек в множестве $P$ будем обозначать как $\#P$
-и считать его равным $\infty$, если $P$ не менее чем счётно.
-
-Вроде как в англоязычной литературе это стандартное обозначение.
-
-\paragraph{Утверждение 3.}
+\begin{proposition}
+	\label{proposition:intervals_cross}
 	Если интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются,
@@
 	то $cr(M_1,M_2) \cap cr(M_3,M_4)$~--- либо не более 4 точек, либо прямая.
+\end{proposition}
 
-\paragraph{Доказательство.}
+\begin{proof}
 	Пересечением $cr(M_1,M_2) \cap cr(M_3,M_4)$ может быть либо от 2 до 4 точек, либо прямая,
-либо, если эти кресты совпадают, их пересечение совпадает с каждым из эих крестов.
+	либо, если эти кресты совпадают, их пересечение совпадает с каждым из этих крестов.
 	Но $cr(M_1,M_2) \neq cr(M_3,M_4)$, потому что середины отрезков $M_1 M_2$ и $M_3 M_4$
 	совпасть не могут, т.к. интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются по условию.
+\end{proof}
 
-\paragraph{Лемма 4.}
+\begin{lemma}
+	\label{lemma_preliminary_size}
 	Пусть $S$~--- система Эрдёша,
@@
 	Тогда $\#S \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2}$.
+\end{lemma}
 
-\paragraph{Доказательство.}
+\begin{proof}
 	Если $\#(cr(M_1, M_2) \cap cr(M_3, M_4)) < \infty$,
@@
 	$\#S \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
-Иначе (в силу утверждения 3) $cr(M_1, M_2) \cap cr(M_3, M_4) = m$,
+	Иначе (в силу утверждения~\ref{proposition:intervals_cross}) $cr(M_1, M_2) \cap cr(M_3, M_4) = m$,
 	где $m$~--- прямая.
-В силу следствия 1.1 на этой прямой лежит не более $(d+3)/2$ точек;
+	В силу следствия~\ref{corollary:max_points_on_line} на этой прямой лежит не более $(d+3)/2$ точек;
 	кроме того, из общего количества точек нужно вычесть 4,
@@
 		=
-	4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2},
+		4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2}
+		.
 	\end{equation}
-что и требовалось доказать.
+\end{proof}
 
-\paragraph{Обозначение.}
+\begin{designation}
 	Через $\varphi_k$ будем обозначать максимально возможный минимум попарных расстояний между $k$ точками,
 	упакованными в единичный квадрат (ссылка!!).
+\end{designation}
 
-На английском это есть в статье и звучит как-то адекватно.
-
-\paragraph{Лемма 5.}
+\begin{lemma}
 	Пусть $S$~--- система Эрдёша,
@@
 	\end{equation}
+\end{lemma}
 
-\paragraph{Доказательство.}
-Впишем $S$ в квадрат со стороной $d$ (см. лемму 2)
+\begin{proof}
+	Впишем $S$ в квадрат со стороной $d$ (см. лемму~\ref{lemma:square_container})
 	и разобьём этот квадрат на $m^2$ маленьких квадратов со стороной $d/m$.
@@
 
-По лемме 4
+	TODO: реверанс про выпуклый четырёхугольник.
+
+	По лемме~\ref{lemma_preliminary_size}
 	\begin{equation}
@@
 
-Честно считаем дискриминант:
+	Непосредственно выпишем дискриминант:
 	\begin{multline}
@@
 		d > \mu \cdot (\#S-2)
+		.
 	\end{equation}
-что и требовалось доказать.
+\end{proof}
 

### 2018-08-19 vmmsh-2018/BetterLinearBound.tex
[vmmsh] References

@@
 	не получится: часто встречаются конструкции систем Эрдёша, расположенных на окружности
-	(Ссылка!!!).
+	~\cite{anning1915discussions,harborth1993upper,piepmeyer1996maximum,kurz2008bounds,our-vvmsh-2018}.
 \end{remark}
@@
 \begin{proof}
-	Если $\#(cr(M_1, M_2) \cap cr(M_3, M_4)) < \infty$,
-	то (ссылка на Солимоси!)
-	$\#S \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
+	Если $|(cr(M_1, M_2) \cap cr(M_3, M_4))| < \infty$,
+	то (см. доказательство теоремы~\ref{thmErdos} или~\cite[часть 2, неравенство (1)]{solymosi2003note})
+	$\#S \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4|$.
 	Иначе (в силу утверждения~\ref{proposition:intervals_cross}) $cr(M_1, M_2) \cap cr(M_3, M_4) = m$,
@@
 	Через $\varphi_k$ будем обозначать максимально возможный минимум попарных расстояний между $k$ точками,
-	упакованными в единичный квадрат (ссылка!!).
+	упакованными в единичный квадрат~\cite{locatelli2002packing}.
 \end{designation}
@@
 
-Какие-то умные и добрые люди вычислили $\varphi_k$ до $k=10$ включительно (ссылка!) с точностью 4 знака после запятой.
+Какие-то умные и добрые люди вычислили $\varphi_k$ до $k=10$ включительно~\cite{locatelli2002packing} с точностью 4 знака после запятой.
 (Умные, потому что вычислили, а добрые, потому что выложили.)
@@
 
-А особенная приятность в том, что вычисление новых $\phi_k$ может давать автоматическое улучшение оценки
+А особенная приятность в том, что вычисление новых $\varphi_k$ может давать автоматическое улучшение оценки
 (может, правда, и не давать).

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] #S to |S|

@@
 	$d = \mathop{diam} S > 5$.
-	Тогда $\#S \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2}$.
+	Тогда $|S| \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2}$.
 \end{lemma}
@@
 	то (см. доказательство теоремы~\ref{thmErdos} или~\cite[часть 2, неравенство (1)]{solymosi2003note})
-	$\#S \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4|$.
+	$|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4|$.
 	Иначе (в силу утверждения~\ref{proposition:intervals_cross}) $cr(M_1, M_2) \cap cr(M_3, M_4) = m$,
@@
 	\begin{equation}
-		\#S \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| - 4 + \frac{d+3}{2}
+		|S| \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| - 4 + \frac{d+3}{2}
 		=
@@
 	$m \geq 2$,
-	$ \#S = (k-1)m^2 + 2$.
+	$ |S| = (k-1)m^2 + 2$.
 
@@
 	\begin{equation}
-		d \geq \mu (\#S - 2),
+		d \geq \mu (|S| - 2),
 	\end{equation}
@@
 	\begin{equation}
-		\#S \leq 4 \left( \frac{d}{m} \varphi_k \right)^2 + \frac{d}{2} - \frac{5}{2}
+		|S| \leq 4 \left( \frac{d}{m} \varphi_k \right)^2 + \frac{d}{2} - \frac{5}{2}
 	\end{equation}
-	или, с учётом того, что $ \#S = (k-1)m^2 + 2$,
+	или, с учётом того, что $ |S| = (k-1)m^2 + 2$,
 	\begin{equation}
@@
 	\begin{equation}
-		d > \mu \cdot (\#S-2)
+		d > \mu \cdot (|S|-2)
 		.
@@
 По их расчётам получаем $\varphi_{10} = 0.4214$,
-что даёт оценку $d>0.3583(\#S-2)$.
+что даёт оценку $d>0.3583(|S|-2)$.
 
-Это, конечно, для $\#S$ специального вида, но множитель $(m+1)^2 / m^2$ съестся на бесконечности,
+Это, конечно, для $|S|$ специального вида, но множитель $(m+1)^2 / m^2$ съестся на бесконечности,
 а итоговые коэфициенты и так будем брать с запасом (кто их там знает, как они округляли).
 
-Это для $\#S \geq 41$, но для меньших мощностей посчитано (и не один раз, в том числе нами).
+Это для $|S| \geq 41$, но для меньших мощностей посчитано (и не один раз, в том числе нами).
 
-Кстати, для $\#S = 3$ не проходит (и не должно).
+Кстати, для $|S| = 3$ не проходит (и не должно).
 

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Designation of length unified

@@
 	$\{M_1, M_2, M_3, M_4\} \subset S$, как минимум три из этих точек различны,
-	$|M_1 M_2| = |M_3 M_4| = 1$.
+	$|M_1 - M_2| = |M_3 - M_4| = 1$.
 	Тогда никакая прямая $m$ не содержит точки $M_1$, $M_2$, $M_3$ и $M_4$ одновременно.
@@
 	Пусть точка $M\in S$.
-	Тогда либо $|MM_1| - |MM_2| = 0$ и $M\in m_{12}$, либо $\Bigl||MM_1| - |MM_2|\Bigr| = 1$ и $M\in m$,
+	Тогда либо $|M - M_1| - |M - M_2| = 0$ и $M\in m_{12}$, либо $\Bigl||M - M_1| - |M - M_2|\Bigr| = 1$ и $M\in m$,
 	т.е. в любом случае $M\in m \cup m_{12}$.
@@
 \begin{proof}
-	Пусть $M_1, M_2 \in S$, $|M_1 M_2| = d$,
+	Пусть $M_1, M_2 \in S$, $|M_1 - M_2| = d$,
 	т.е. диаметр системы $S$ достигается на $M_1$  $M_2$.
@@
 	Очевидно, что для $i>2$ имеем $|x_i| < d/2$
-	(иначе $|M_i M_1| > d$ или $|M_i M_2| > d$).
-	Более того, $|y_i| < d$ (иначе $|M_i M_1| > d$ или $|M_i M_2| > d$).
+	(иначе $|M_i - M_1| > d$ или $|M_i - M_2| > d$).
+	Более того, $|y_i| < d$ (иначе $|M_i - M_1| > d$ или $|M_i - M_2| > d$).
 
@@
 	$d = \mathop{diam} S > 5$.
-	Тогда $|S| \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2}$.
+	Тогда $|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| + \frac{d-5}{2}$.
 \end{lemma}
@@
 	\begin{equation}
-		|S| \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| - 4 + \frac{d+3}{2}
+		|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| - 4 + \frac{d+3}{2}
 		=
-		4 \cdot |M_1 M_2| \cdot |M_3 M_4| + \frac{d-5}{2}
+		4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| + \frac{d-5}{2}
 		.
@@
 	а в $Q_2$ выберем точки $M_3$ и $M_4$
-	так, что $|M_1 M_2| \leq \varphi_k d /m$, $|M_3 M_4| \leq \varphi_k d/m$
+	так, что $|M_1 - M_2| \leq \varphi_k d /m$, $|M_3 - M_4| \leq \varphi_k d/m$
 	(это всегда можно сделать в силу определения $\varphi_k$).
@@
 	В случае (б) в $Q_1$ выберем точки $M_1$ и $M_2$ так, что
-	$|M_1 M_2| \leq \varphi_{k+1} d /m \leq \varphi_k d /m$.
+	$|M_1 - M_2| \leq \varphi_{k+1} d /m \leq \varphi_k d /m$.
 	Из всех точек, кроме $M_1$, выберем $M_3$ и $M_4$ так, что
-	$|M_3 M_4| \leq \varphi_k d/m$.
+	$|M_3 - M_4| \leq \varphi_k d/m$.
 	Возможное совпадение $M_2$ и $M_3$ не помешает дальнейшему доказательству.

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Cosmetics

@@
 	\label{corollary:max_points_on_line}
-	Пусть $S$~--- система Эрдёша, $\mathop{diam} S = d$.
+	Пусть $S\in\mathfrak{M}_n$, $\operatorname{diam} S = d$.
 	Тогда ни на какой прямой не лежит более $(d+3)/2$ точек из $S$.
@@
 	интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются,
-	$d = \mathop{diam} S > 5$.
+	$d = \operatorname{diam} S > 5$.
 	Тогда $|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| + \frac{d-5}{2}$.
@@
 	Пусть $S$~--- система Эрдёша,
-	$d = \mathop{diam} S > 5$,
+	$d = \operatorname{diam} S > 5$,
 	$k \geq 2$,

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Cosmetics

@@
 	кроме того, из общего количества точек нужно вычесть 4,
-	которые бы дали эти 2 креста в случае дискретного переечения.
+	которые бы дали эти 2 креста в случае дискретного переcечения.
 	Получаем верхнюю оценку
-	\begin{equation}
+	\begin{multline}
 		|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| - 4 + \frac{d+3}{2}
 		=
+		\\=
 		4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| + \frac{d-5}{2}
 		.
-	\end{equation}
+	\end{multline}
 \end{proof}

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Note about packing points in the square

@@
 
-\begin{designation}
-	Через $\varphi_k$ будем обозначать максимально возможный минимум попарных расстояний между $k$ точками,
-	упакованными в единичный квадрат~\cite{locatelli2002packing}.
-\end{designation}
+Определим теперь коэффициенты упаковки точек в квадрат $\varphi_k$.
+Пусть
+\begin{equation*}
+	\Phi_k = \{ P \subset [0;1]^2 : |P|=k\}
+	,
+\end{equation*}
+где $[0;1]^2$~--- замкнутый единичный квадрат на плоскости.
+Тогда
+\begin{equation*}
+	\varphi_k = \max_{P \in \Phi_k} \min_{A,B \in P} |A - B|
+	.
+\end{equation*}
+Иначе говоря, через $\varphi_k$ будем обозначать наибольшее число, такое,
+что в единичном квадрате нельзя разместить $k$ точек так,
+чтобы расстояние между любыми двумя точками было не менее $\varphi_k$.
+Проблема отыскания $\varphi_k$ носит название проблемы упаковки точек в квадрат~\cite{locatelli2002packing,costa2013valid}.
 
@@
 
-Какие-то умные и добрые люди вычислили $\varphi_k$ до $k=10$ включительно~\cite{locatelli2002packing} с точностью 4 знака после запятой.
-(Умные, потому что вычислили, а добрые, потому что выложили.)
+В~\cite{costa2013valid} приведены значения $\varphi_k$ до $k=10$ включительно
+
+TODO: разобраться, насколько эти значения точны и откуда они взяты =/
 

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] One definition made more clear

@@
 
+\begin{definition}
+	Будем говорить, что подмножество плоскости $S\subset\mathbb{R}^2$
+	вписано в квадрат со стороной $d$, если
+	можно построить такой квадрат $Q$ со стороной $d$,
+	что каждая точка $S$ окажется либо внутри $Q$, либо на одной из его сторон.
+\end{definition}
+
 \begin{lemma}
 	\label{lemma:square_container}
-	Систему Эрдёша $S$ диаметром $d$ можно вместить в квадрат со стороной $d$.
+	Пусть $S\in\mathfrak{M}_n$, $\operatorname{diam} S = d$.
+	Тогда $S$ можно вписать в квадрат со стороной $d$.
 \end{lemma}

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Typo
[vmmsh] Preambula for linear bound

@@
+Для конечного дискретного подмножества плоскости $M\subset \mathbb{R}^2$
+естественным образом определяется диаметр:
+\begin{equation*}
+	\operatorname{diam} M = \max_{A,B\in M} |A-B|
+	.
+\end{equation*}
+
+Солимоси~\cite{solymosi2003note} показал существование такой константы $c$,
+что для любого $S\in\mathfrak{M}_n$ выполнено $\operatorname{diam} M \geq c n$.
+Солимоси не привёл значения этой константы, однако из его доказательства можно вывести,
+что эта константа равна $1/24$.
+
+Мы сейчас дадим оригинальное доказательство, конструкция которого отличается от конструкции,
+предложенной Солимоси, дающее большее значение константы $c$.
+
+TODO: надо ли ссылаться на стаью в матзаметках и рассказывать про независимость тамошнего доказательства?
+
 \begin{lemma}
@@
 	Введём прямоугольную декартову систему координат на плоскости таким образом, что
-	$M_1 = (-d/2; 0)$, $M_1 = (d/2; 0)$.
+	$M_1 = (-d/2; 0)$, $M_2 = (d/2; 0)$.
 

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Cosmetics

@@
 	Пусть $y_+ = \max_{i} y_i$, $y_- = \min_{i} y_i$, $M_+=(x_+, y_+)$, $M_-=(x_-, y_-)$
-	(если максимум досигается на нескольких точках, возьмём любые).
+	(если максимум досигается на нескольких точках, возьмём любые из них).
 	Тогда
@@
 		d \geq |M_+ M_-| = \sqrt{(x_+ - x_-)^2 + (y_+ - y_-)^2}
-		\geq \sqrt{(y_+ - y_-)^2} = y_+ - y_-
+		\geq
+		\\\geq
+		\sqrt{(y_+ - y_-)^2} =
+		y_+ - y_-
 	\end{multline}
@@
 \begin{remark}
-	Скорее всего, улучшить оценку на сторону квадрата, сделав её меньше диаметра,
-	не получится: часто встречаются конструкции систем Эрдёша, расположенных на окружности
+	Вероятно, улучшить оценку на сторону квадрата, сделав её меньше диаметра,
+	невозможно: часто встречаются конструкции множеств $S\in\mathfrak{M}_n$, расположенных на окружности
 	~\cite{anning1915discussions,harborth1993upper,piepmeyer1996maximum,kurz2008bounds,our-vvmsh-2018}.

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Terminology

@@
 	\label{lemma_preliminary_size}
-	Пусть $S$~--- система Эрдёша,
+	Пусть $S\in\mathfrak{M}_n$,
 	$M_1, M_2, M_3, M_4 \in S$,
@@
 \begin{lemma}
-	Пусть $S$~--- система Эрдёша,
+	Пусть $S\in\mthfrak{M}_n$,
 	$d = \operatorname{diam} S > 5$,

### 2018-08-22 vmmsh-2018/Review.tex
[vmmsh] Review - char M

Итак, каждому множеству $M\in\mathfrak{M}_n$ сопоставлены
два натуральных числа: мощность и диаметр.
Кемнитц~\cite{kemnitz1988punktmengen} обнаружил и третье,
которое назвал характеристикой.

\begin{definition}
	\label{def_char_classic}
	Характеристикой множества $M\in\mathfrak{M}_n$ называется свободное от квадратов~\cite[гл. 34, п. 3]{Bukhstab-number-theory}
	число $p$, такое, что площадь любого треугольника $ABC$, где $A,B,C\in M$,
	соизмерима с $\sqrt{p}$.
\end{definition}

Опираясь на работы~\cite{our-mkmitu-2016,our-ped-2017} и используя~\cite[теорема 3.1]{polygons-on-lattices},
можно доказать, что определение~\ref{def_char_classic} равносильно следующему определению.
\begin{definition}
	\label{def_char_classic}
	Характеристикой множества $M\in\mathfrak{M}_n$ называется свободное от квадратов
	число $p$, такое, что множество $M$ может быть размещена на решётке
\begin{equation}\label{grid_for_Erdosh_system}
	\left\{\left(
		\frac{a_i}{2m}
		;
		\frac{b_i\sqrt{p}}{2m}
	\right)\right\},
\end{equation}
где $a_i, b_i \in \mathbb{Z}$, в качестве $m$ можно взять длину любого ребра $M$.
\end{definition}

TODO: надо ли писать доказательство?

TODO: можно ли использовать термин "ребро $M$" или его надо определить? Где?

Характеристика для данного множества $M\in\mathfrak{M}_n$ определяется единственным образом и обозначается $\operatorname{char}M$.
Заметим, что понятие характеристики можно обобщить и на множества $M\in\mathfrak{M}$,
содержащиеся в некоторой прямой.
В таком случае логично полагать $\operatorname{char} M = 0$.

\begin{theorem}
	Для любого $n\geq 3$ и любого свободного от квадратов числа $p$
	существует множество $M\in\mathfrak{M}_n$ такое, что $\operatorname{char} M = p$.
\end{theorem}

\begin{proof}
	Не теряя общности положим, что $n$ нечётно.
	Пусть
	\begin{equation*}
		M = \{M_i, i =0,1,...,n+1\},
	\end{equation*}
	где
	$M_i = (2^{n-i}-p\cdot 2^{i-1},0)$, $i=0,1,...,n$;
	$M_{n+1} = (0,\sqrt{p \cdot 2^{n+1}})$.
	Очевидно, для $0\leqslant i \leqslant j \leqslant n$ выполнено
	$|M_i- M_j|\in\mathbb{N}$.
	Заметим, что для $i=1,...,n$
	\begin{multline*}
		|M_{n+1} - M_i| =
		\sqrt{p\cdot 2^{n+1} + (2^{n-i}-p\cdot2^{i-1})^2}
		=
		\\=
		\sqrt{4p\cdot 2^{n-1} + (2^{n-i})^2 - 2 \cdot 2^{n-i} \cdot p \cdot 2^{i-1} + (p\cdot 2^{i-1})^2}
		=
		\\=
		\sqrt{4p\cdot 2^{i-1} \cdot 2^{n-i} + (2^{n-i})^2 - 2 \cdot 2^{n-i} \cdot p \cdot 2^{i-1} + (p\cdot 2^{i-1})^2}
		=
		\\=
		\sqrt{(2^{n-i})^2 + 2 \cdot 2^{n-i} \cdot p \cdot 2^{i-1} + (p\cdot 2^{i-1})^2}
		=
		%\\=
		%\sqrt{(2^{n-i})^2 + 2 \cdot 2^{n-i} \cdot p \cdot 2^{i-1} + (p\cdot 2^{i-1})^2}
		%=
		2^{n-i} + p\cdot 2^{i-1}
		\in\mathbb{N}
		.
	\end{multline*}
\end{proof}

TODO: не знаю, доказывалось ли это ранее. Скорее всего, да, но где - неизвестно.

### 2018-08-22 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Typo fix
[vmmsh] New proof

@@
 \begin{proof}
-	Предположим противное.
-	Очевидно, что на прямой $m$ может быть не более $d+1$ точек системы $S$.
-
-	Пусть сначала $d$ чётно.
-	Тогда из каждой пары соседних точек, кроме, быть может, одной пары,
-	нужно выбросить хотя бы одну точку в силу леммы~\ref{lemma_points_on_line}.
-	Итого мы выбросим не менее $d/2$ точек,
-	останется не более $(d+2)/2$ точек.
-
-	Пусть теперь $d$ нечётно.
-	Тогда нужно выбросить не менее $(d-1)/2$ точек.
-	Останется не более $(d+3)/2$ точек.
+	Предположим противное, т.е на прямой $m$ лежит $k$ точек из $S$ и $k > (d+3)/2$.
+	Пусть $\rho_i$, $i = 1,...,k-1$~--- расстояния между соседними точками,
+	пронумерованные в порядке возрастания.
+	Из леммы~\ref{lemma_points_on_line} получаем $\rho_{k-1}\geq...\geq\rho_2\geq 2$.
+	Значит,
+	\begin{equation*}
+		\sum_{i=1}^{k-1} \rho_i
+		=
+		\rho_1 + \sum_{i=2}^{k-1} \rho_i
+		\geq
+		1 + \sum_{i=2}^{k-1} \rho_i
+		\geq
+		1 + 2*k-3
+		>
+		1 + d + 3 - 3
+		=
+		d + 1
+		,
+	\end{equation*}
+	т.е. расстояние между крайними точками превосходит диаметр системы,
+	чего быть не может.
 \end{proof}
@@
 \begin{lemma}
-	Пусть $S\in\mthfrak{M}_n$,
+	Пусть $S\in\mathfrak{M}_n$,
 	$d = \operatorname{diam} S > 5$,

### 2018-08-22 vmmsh-2018/ReviewChar.tex
[vmmsh] Review on characteristics

--- 2018-08-22 vmmsh-2018/Review.tex
@@
 TODO: не знаю, доказывалось ли это ранее. Скорее всего, да, но где - неизвестно.
+
+В отличие от диаметра и мощности, характеристика в некотором смысле локальна:
+если $S\in\mathfrak{M}_n$, $H \subset S$, $H\in\mathfrak{M}_{p}$, $p<n$,
+то
+\begin{equation*}
+	|H| < |S|, \quad \operatorname{diam} H \leq \operatorname{diam} S, \quad \operatorname{char} H = \operatorname{char} S
+	.
+\end{equation*}
+
+Характеристика (наряду с диаметром и мощностью) обобщается на случай, когда множество точек с целочисленными расстояниями
+рассмативается в $\mathbb{R}^k$~\cite{kurz2005characteristic}.

### 2018-08-23 vmmsh-2018/ReviewMultidim.tex
[vmmsh] Multidimensional review - drafts

В пространствах более высокой размерности можно построить множества, аналогичные $\mathfrak{M}$.

\begin{definition}
	Пусть $m \geq 2$, $n \geq m + 1$.
	Через $\mathfrak{M}(m,n)$ будем обозначать множество таких подмножеств $m$--мерного евклидова пространства
	$M\subset\mathbb{R}^m$, что $|M| = n$, для любых $M_1,M_2 \in M$ выполнено $|M_1 - M_2| \in\mathbb{N}$
	и $M$ не содержится ни в какой $(m-1)$--мерной гиперплоскости.
\end{definition}

Характеристика (наряду с диаметром и мощностью) обобщается на случай, когда множество точек с целочисленными расстояниями
рассмативается в $\mathbb{R}^k$~\cite{kurz2005characteristic}.

При изучении множеств $\mathfrak{M}(m,n)$ возникает закономерный вопрос
о минимальной диаметре для заданных размерности и мощности.
Говоря формально, определим функцию
\begin{equation*}
	d(m,n) = \min_{M\in\mathfrak{M}(m,n)} \operatorname{diam} M
	.
\end{equation*}

Имеют место быть нижеследующие оценки на $d(m,n)$
(за основу взят список из~\cite{kurz2008bounds}):

%, дополненный результатом работы~\cite{nozaki2013lower}
%и результатом, полученным в данной статье.

\begin{align}
	& d(m, n - 1) \leq d(m,n) \\
	& d(n, n + 1) = 1 \\
	& d(m, n) \leq \begin{cases}
		2^{n-m+1} -2 & \mbox{для} n-m \equiv 0 \mod 2
		\\
		3(2^{n-m} -1) & \mbox{для} n-m \equiv 1 \mod 2
	\end{cases} \qquad \mbox{\cite{harborth1985diameters}} \\
	& d(2,n) \geq cn \qquad \mbox{TODO: вставить константу} \\
	& d(m,n) \leq (n-m)^{c \log \log (n-m)} \mbox{~для некоторого $c$} \qquad \mbox{\cite{harborth1993upper}} \\
	& d(m,n) > \sqrt{\frac{3}{2m}} n^{1/m} \mbox{\cite{kanold1981punktmengen},в частных случаях улучшена в \cite{nozaki2013lower}} \\
	& d(3,n) > \frac{1}{\sqrt{14}}n^{1/2} \mbox{~для~} n \geq 5 \quad \mbox{\cite{kanold1981punktmengen}} \\
	& d(m, 2m + 1) \leq 8 \qquad \mbox{\cite{piepmeyer1988raumliche}} \\
	& d(m, 2m + 2) \leq 13 \qquad \mbox{\cite{piepmeyer1988raumliche}} \\
	& d(m, 3m    ) \leq 109 \qquad \mbox{\cite{kemnitz1988punktmengen}} \\
	& d(m, m^2 + m) \leq 17 \qquad \mbox{\cite{kurz2008bounds}} \\
	& 3 \leq d(m, n) \leq 4 \mbox{~для~} m + 2 \leq n \leq 2m \mbox{~и~} d(m, 2m) = 4 ~ \mbox{\cite{piepmeyer1988raumliche,harborth1991point}}\\
\end{align}

Кроме того, Курц~\cite{kurz2008bounds} выдвинул предположение, что $d(m - 1, n) \geq d(m, n)$.

Существуют алгоритмы вычисления $d(m,n)$ на ЭВМ, однако они требуют большого количества машинного времени,
причём требуемое время быстро растёт с ростом $m$ и $n$.
Первые несколько значений $d(2,n)$ были получены в~\cite{harborth1998integral},
следующие известные значения~--- в \cite{kurz2005characteristic,kurz2006konstruktion,kurz2008minimum,kurz2008bounds,our-mz-rus}.
Курцу удалось существенно оптимизировать алгоритм, основанный на переборе, применяя характеристику~\cite{kurz2005characteristic,kreisel2008heptagon}.
Однако сложность вычисления максимальной возможной мощности для заданного диаметра $d$, по оценке Курца~\cite{kreisel2008heptagon},
составлет $O(d^3)$ (в двумерном случае).
Курц (TODO: ссылка!) отмечает, что важно перепроверять вычисленные значения независимо составленными программами для ЭВМ.
Приведём некоторые известные значения:
\\
$
(d(2, n))_{n=3,...,122} = 1, 4, 7, 8, 17, 21, 29, 40, 51, 63, 74, 91, 104, 121,
134, $\\$153, 164,
196, 212, 228, 244, 272, 288, 319, 332, 364, 396, 437, 464, 494, 524,$\\$ 553, 578, 608,
642, 667, 692, 754, 816, 897, 959, 1026, 1066, 1139, 1190, $\\$ 1248, 1306, 1363, 1410,
1460, 1514, 1564, 1614, 1675, 1727, 1770, 1817, $\\$1887, 1906, 2060, 2140, 2169,
2231, 2299, 2432, 2494, 2556, 2624, 2692, $\\$2827, 2895, 2993, 3098, 3196, 3294,
3465, 3575, 3658, 3749, 3885, 3922, $\\$4223, 4380, 4437, 4559, 4693, 4883,
5018, 5109, 5264, 5332, 5480, 5603, $\\$5738, 5938, 5995, 6052,
6324, 6432, 6630, 6738, 6939, 7061, 7245, 7384, $\\$7568, 7752, 7935, 8119, 8321,
8406, 8648, 8729, 8927, 9052, 9211, 9423, $\\$9534, 9794, 9905
$.

Примечательно, что для $n=4$ и $n=18$ оптимальные множества не единственны.
Курц выяснил, что для $9 \leq n \leq 122$ опимальное множество содержится
в объединении прямой и точки (т.е. все точки, кроме одной, лежат на некоторой прямой).

План

2. Несколько слов про вычисление. Обзор результатов Харборса, Курца и наших.
Независимые эксперименты - это хорошо. Курц и характеристика. Что-нибудь про сложность вычислений.
Неединственные оптимальные системы. Про веерность со ссылкой на Курца.

1. Особые случаи

1.1. Нечётные расстояния (3 работы)

1.2. Круговые системы (1 работа) - оценка сверху

1.3. Веерные системы (1 работа) - и оценка на них снизу

1.4. Системы с миниметром 2 (Солимоси)

1.5. Системы общего положения (гипотеза Эрдёша, работа Курца)

1.6. 7-кластеры (Курц)

3. Связанные проблемы: гипотеза Эрдёша о числе расстояний, гипотеза Улама--Эрдёша.
Ссылка на 2 работы Гарибальди. Ссылка на доказательство гипотезы Эрдёша о числе расстояний.

### 2018-08-23 vmmsh-2018/ReviewMultidim.tex
[vmmsh] Review - WIP

@@
 	& d(2,n) \geq cn \qquad \mbox{TODO: вставить константу} \\
-	& d(m,n) \leq (n-m)^{c \log \log (n-m)} \mbox{~для некоторого $c$} \qquad \mbox{\cite{harborth1993upper}} \\
+	& d(m,n) \leq (n-m)^{c \log \log (n-m)} \mbox{~для некоторого $c$} \qquad \mbox{\cite{harborth1993upper}} \label{upper_bound_d_m_n_log_log}\\
 	& d(m,n) > \sqrt{\frac{3}{2m}} n^{1/m} \mbox{\cite{kanold1981punktmengen},в частных случаях улучшена в \cite{nozaki2013lower}} \\
@@
 
-Примечательно, что для $n=4$ и $n=18$ оптимальные множества не единственны.
-Курц выяснил, что для $9 \leq n \leq 122$ опимальное множество содержится
+Примечательно, что для $m=2$, $n=4$ и $n=18$ оптимальные множества не единственны.
+Курц выяснил, что для $9 \leq n \leq 122$ оптимальное множество содержится
 в объединении прямой и точки (т.е. все точки, кроме одной, лежат на некоторой прямой).
 
-План
+Для многомерного случая вычисленные значения можно найти в~\cite{kurz2008bounds}.
 
-2. Несколько слов про вычисление. Обзор результатов Харборса, Курца и наших.
-Независимые эксперименты - это хорошо. Курц и характеристика. Что-нибудь про сложность вычислений.
-Неединственные оптимальные системы. Про веерность со ссылкой на Курца.
+При изучении множеств из $\mathfrak{M}_n$ часто возникают множества специальных видов.
+Наиболее простым примером являются веерные множества, т.е. множества, содержащиеся в объединении прямой и точки.
+Курц и Вассерман~\cite{kurz2008minimum} показали, что диаметр веерного множества мощности $n$
+не менее, чем $n^{c \log \log n}$, где $c$~--- некоторая константа.
+В этой же работе высказано предположение, что подобная оценка верна для всех множеств.
+Заметим, что эта оценка совпадает с верхней оценкой~\eqref{upper_bound_d_m_n_log_log} на $d(2,n)$
+с точностью до константы.
+Это вкупе с тем фактом, что большинство известных оптимальных множеств являются веерными,
+позволяет предположить, что все оптимальные множества являются веерными.
+
+Возникает закономерный вопрос: а какие бывают множества из $\mathfrak{M}_n$,
+кроме веерных?
+
+\begin{definition}
+	\cite{kurz2008minimum} Множество $M\in\mathfrak{M}_n$ называется множеством полуобщего положения,
+	если ни через какие три точки из $M$ нельзя провести прямую.
+	Множество всех множеств полуобщего положения мощности $n$ будем обозначать $\overline{\mathfrak{M}}_n$.
+\end{definition}
+
+В \cite{harborth1993upper} показано, что $\overline{\mathfrak{M}}_n\neq\varnothing$ для любого $n \geq 3$
+и приведены примеры множеств из $\overline{\mathfrak{M}}_n$, содержащихся в окружности.
+На эти множествах достигается верхняя оценка~\eqref{upper_bound_d_m_n_log_log} на $d(2,n)$
+с некоторой константой.
+
+\begin{definition}
+	\cite{kurz2008minimum} Множество $M\in\overline{\mathfrak{M}}_n$, $n\geq 4$ называется множеством общего положения,
+	если ни через какие четыре точки из $M$ нельзя провести окружность.
+	Множество всех множеств общего положения мощности $n$ будем обозначать $\dot{\mathfrak{M}}_n$.
+\end{definition}
 
@@
 1.5. Системы общего положения (гипотеза Эрдёша, работа Курца)
+\cite{kreisel2008heptagon}
 
@@
 
+Интересны, на взгляд авторов, и ещё несколько результатов для множеств из $\mathfrak{M}_n$ специального вида.
+Так, Солимоси построил~\cite{solymosi2003note} для любого $n\geq 3$ конструкцию множества $M\in\mathfrak{M}_n$
+такого, что для некоторых $M_1, M_2 \in M$ выполнено $|M_1 - M_2| = 2$.
+Полученное им множество также является веерным.
+
+Для всех известных значений d(2,n) < d(2,n+1).
+Для больших размерностей это неверно (ссылка, пример).
+
 3. Связанные проблемы: гипотеза Эрдёша о числе расстояний, гипотеза Улама--Эрдёша.
 Ссылка на 2 работы Гарибальди. Ссылка на доказательство гипотезы Эрдёша о числе расстояний.
+
+4. Примеры множеств.
+
+Перспективными направлениями исследований являются, кроме непосредственно отыскания как можно более точных оценок на $d(m,n)$,
+например, поиск систем из $\mathfrak{M}_n$, содержащихся в двух прямых (параллельных или пересекающихся).

### 2018-08-24 vmmsh-2018/ReviewMultidim.tex
[vmmsh] On before iteration

@@
 При изучении множеств $\mathfrak{M}(m,n)$ возникает закономерный вопрос
-о минимальной диаметре для заданных размерности и мощности.
+о минимальном диаметре для заданных размерности и мощности.
 Говоря формально, определим функцию
@@
 Приведём некоторые известные значения:
+
+TODO: cite
 \\
@@
 не менее, чем $n^{c \log \log n}$, где $c$~--- некоторая константа.
-В этой же работе высказано предположение, что подобная оценка верна для всех множеств.
+
+TODO: про лемму
+
+В этой же работе высказано предположение, что подобная оценка верна для всех множеств из $\mathfrak{M}_n$.
 Заметим, что эта оценка совпадает с верхней оценкой~\eqref{upper_bound_d_m_n_log_log} на $d(2,n)$
@@
 
+Определим функцию~\cite{kurz2008minimum}
+\begin{equation*}
+	\overline{d}(m,n) = \min_{M\in\overline{\mathfrak{M}}(m,n)} \operatorname{diam} M
+	.
+\end{equation*}
+
+В работе~\cite{kurz2008minimum} вычислено, что
+\\
+$\bigl(\overline{d}(2, n)\bigr)_{n=3,...,36} = 1,$
+4, 8, 8, 33, 56, 56, 105, 105, 105, 532, 532, 735, 735, 735, 735,
+1995, 1995, 1995, 1995, 1995, 1995, 9555, 9555, 9555, 10672,
+13975, 13975, 13975, 13975, 13975, 13975, 13975, 13975.
+
 \begin{definition}
@@
 
+В отличие от $\mathfrak{M}_n$ и $M\in\overline{\mathfrak{M}}_n$,
+существование множеств из $\dot{\mathfrak{M}}_n$ является достаточно сложным вопросом.
+Множество из $\dot{\mathfrak{M}}_6$ предъявлено в~\cite{noll1989nclusters},
+множество из $\dot{\mathfrak{M}}_7$~--- в~\cite{kurz2008minimum}.
+Определим функцию~\cite{kurz2008minimum}
+\begin{equation*}
+	\dot{d}(m,n) = \min_{M\in\dot{\mathfrak{M}}(m,n)} \operatorname{diam} M
+	.
+\end{equation*}
+Согласно~\cite{kurz2008minimum},
+$\bigl(\dot{d}(2,n)\bigr)_{n=3,...,7}
+= 1,$ 8, 73, 174, 22270.
+
+Вопрос о существовании множеств из $\dot{\mathfrak{M}}_8$
+остаётся открытым.
+
+\begin{definition}
+	\cite{noll1989nclusters,kurz2013constructing}
+	$n$--кластером называется множество $M\in\dot{\mathfrak{M}}_n$,
+	содержащееся в целочисленной решётке.
+\end{definition}
+Существование $n$--кластера эквивалентно существованию множества $M\in\dot{\mathfrak{M}}_n$
+характеристики 1.
+В~\cite{noll1989nclusters} показано существование 6--кластеров,
+в~\cite{kurz2013constructing}~--- 7--кластеров.
+
 1. Особые случаи
-
-1.1. Нечётные расстояния (3 работы)
-
-1.2. Круговые системы (1 работа) - оценка сверху
-
-1.3. Веерные системы (1 работа) - и оценка на них снизу
-
-1.4. Системы с миниметром 2 (Солимоси)
 
@@
 Интересны, на взгляд авторов, и ещё несколько результатов для множеств из $\mathfrak{M}_n$ специального вида.
+
 Так, Солимоси построил~\cite{solymosi2003note} для любого $n\geq 3$ конструкцию множества $M\in\mathfrak{M}_n$
@@
 Полученное им множество также является веерным.
+
+TODO: про сущ-е множеств с миниметром 1.
+
+Если $M\in\mathfrak{M}(m,n)$ и все рёбра $M$ имею чётную длину,
+то понятно, что множество $H$, полученное из $M$ сжатием в два раза,
+также будет принадлежать $\mathfrak{M}(m,n)$.
+Выполняя такую операцию, мы рано или поздно получим множество из $\mathfrak{M}(m,n)$ такое,
+что для некоторых $M_1,M_2\in M$ расстояние $|M_1 - M_2|$ нечётно.
+Возникает закономерный вопрос: сколько может быть нечётных расстояний?
+
+В~\cite{graham1974there} доказано, что
+для существования множества $M\in\mathfrak{M}(m,m+2)$ такого, что для любых
+$M_i,M_j\in M$, $i\neq j$ расстояние $|M_i-M_j|$ нечётно,
+необходимо и достаточно, чтобы
+$m+2\equiv 0 (\operatorname{mod} 16)$.
+В частности, не существует такого множества $M\in\mathfrak{M}_4$,
+что все его рёбра имеют нечётную длину.
+На основе этого факта в работе~\cite{piepmeyer1996maximum} с использованием результатов теории графов доказано,
+что количество рёбер нечётной длины в множестве $M\in\mathfrak{M}_n$ не превосходит
+\begin{equation}
+	\frac{n^2}{3} + \frac{r(r - 3)}{6}, \mbox{~где~} r = 1, 2, 3 \mbox{~и~} n \equiv r (\operatorname{mod} 3)
+\end{equation}
+и конструктивно показана точность этой оценки.
+
+Отдельные оценки количества рёбер нечётной длины для множеств из $\overline{\mathfrak{M}}_n$ и $\dot{\mathfrak{M}}_n$
+авторам неизвестны.
 

### 2018-08-24 vmmsh-2018/ReviewChar.tex
[vmmsh] Multidimensional review - drafts
[vmmsh] Iteration - intro

@@
 \begin{theorem}
+	\label{thm:power_char_exist}
 	Для любого $n\geq 3$ и любого свободного от квадратов числа $p$
@@
 \end{equation*}
-
-Характеристика (наряду с диаметром и мощностью) обобщается на случай, когда множество точек с целочисленными расстояниями
-рассмативается в $\mathbb{R}^k$~\cite{kurz2005characteristic}.

### 2018-08-25 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Cosmetics
[vmmsh] Iteration

@@
 
-Мы сейчас дадим оригинальное доказательство, конструкция которого отличается от конструкции,
-предложенной Солимоси, дающее большее значение константы $c$.
-
-TODO: надо ли ссылаться на стаью в матзаметках и рассказывать про независимость тамошнего доказательства?
+Мы сейчас дадим другое доказательство, дающее лучшее значение константы $c$.
 
@@
 		\geq
-		1 + 2*k-3
+		1 + 2(k-2)
+		=
+		2k-3
 		>
-		1 + d + 3 - 3
+		d+3-3
 		=
-		d + 1
+		d
 		,
@@
 
-\begin{definition}
-	Будем говорить, что подмножество плоскости $S\subset\mathbb{R}^2$
-	вписано в квадрат со стороной $d$, если
-	можно построить такой квадрат $Q$ со стороной $d$,
-	что каждая точка $S$ окажется либо внутри $Q$, либо на одной из его сторон.
-\end{definition}
-
 \begin{lemma}
@@
 	Пусть $S\in\mathfrak{M}_n$, $\operatorname{diam} S = d$.
-	Тогда $S$ можно вписать в квадрат со стороной $d$.
+	Тогда $S$ содержится в квадрате со стороной $d$.
 \end{lemma}
@@
 	Пусть $M_1, M_2 \in S$, $|M_1 - M_2| = d$,
-	т.е. диаметр системы $S$ достигается на $M_1$  $M_2$.
+	т.е. диаметр системы $S$ достигается на $M_1 M_2$.
 	Введём прямоугольную декартову систему координат на плоскости таким образом, что
@@
 
-	Пусть $y_+ = \max_{i} y_i$, $y_- = \min_{i} y_i$, $M_+=(x_+, y_+)$, $M_-=(x_-, y_-)$
-	(если максимум досигается на нескольких точках, возьмём любые из них).
+	Пусть $y_{\max} = \max_{i} y_i$, $y_{\min} = \min_{i} y_i$, $M_{\max}=(x_{\max}, y_{\max})$, $M_{\min}=(x_{\min}, y_{\min})$
+	(если максимум достигается на нескольких точках, возьмём любые из них).
 	Тогда
 	\begin{multline}
-		d \geq |M_+ M_-| = \sqrt{(x_+ - x_-)^2 + (y_+ - y_-)^2}
+		d \geq |M_{\max} - M_{\min}| = \sqrt{(x_{\max} - x_{\min})^2 + (y_{\max} - y_{\min})^2}
 		\geq
 		\\\geq
-		\sqrt{(y_+ - y_-)^2} =
-		y_+ - y_-
+		\sqrt{(y_{\max} - y_{\min})^2} =
+		y_{\max} - y_{\min}
 	\end{multline}
-	Следовательно, $S$ можно вписать в квадрат со сторонами $x=\pm d/2$,
-	$y=y_+$, $y=y_+ - d$.
+	Следовательно, $S$ содержится в квадрате со сторонами $x=\pm d/2$,
+	$y=y_{\max}$, $y=y_{\max} - d$.
 \end{proof}
 
-\begin{remark}
-	Вероятно, улучшить оценку на сторону квадрата, сделав её меньше диаметра,
-	невозможно: часто встречаются конструкции множеств $S\in\mathfrak{M}_n$, расположенных на окружности
-	~\cite{anning1915discussions,harborth1993upper,piepmeyer1996maximum,kurz2008bounds,our-vvmsh-2018}.
-\end{remark}
+\begin{hypothesis}
+	Не существует такого числа $\gamma < 1$,
+	что любое множество из $M\in \mathfrak{M}_n$, $\geq 3$,
+	содержится в квадрате со стороной $\gamma \cdot \operatorname{diam} M$.
+\end{hypothesis}
+
+В пользу этой гипотезы говорят часто встречающиеся конструкции множеств $M\in\mathfrak{M}_n$,
+расположенных на окружности~\cite{anning1915discussions,harborth1993upper,piepmeyer1996maximum,kurz2008bounds,our-vvmsh-2018}.
 
@@
 \end{proposition}
+
+TODO: напротив этого утверждения стояла галочка, но мы её не обсудили.
 
@@
 
-	а) хотя бы в два маленьких квадрата $Q_1$ и $Q_2$ попало не менее чем по $k$ точек;
+	а) хотя бы в два маленьких квадрата $Q_1$ и $Q_2$ попало не менее, чем по $k$ точек;
 
-	б) хотя бы в один маленький квадрат $Q_1$ попало не менее чем $k+1$ точка.
+	б) хотя бы в один маленький квадрат $Q_1$ попало не менее, чем $k+1$ точка.
 	\\
@@
 	зная, что $\lambda > 0$.
-
 	Непосредственно выпишем дискриминант:

### 2018-08-25 vmmsh-2018/ReviewChar.tex
[vmmsh] Iteration

@@
 \begin{definition}
+	\cite[гл. 34, п. 3]{Bukhstab-number-theory}
+	Натуральное число называется свободным от квдратов,
+	если оно не делится ни на один квадрат простого числа.
+\end{definition}
+
+\begin{definition}
 	\label{def_char_classic}
-	Характеристикой множества $M\in\mathfrak{M}_n$ называется свободное от квадратов~\cite[гл. 34, п. 3]{Bukhstab-number-theory}
+	Характеристикой множества $M\in\mathfrak{M}_n$ называется свободное от квадратов
 	число $p$, такое, что площадь любого треугольника $ABC$, где $A,B,C\in M$,
@@
 \end{definition}
+
+Пусть $\mathbb{Z}$~--- множество целых чисел;
+ребром множества $M\in\mathfrak{M}$
+будем называть любой отрезок $M_1 M_2$,
+где $M_1, M_2 \in M$.
 
@@
 \end{equation}
-где $a_i, b_i \in \mathbb{Z}$, в качестве $m$ можно взять длину любого ребра $M$.
+где $a_i, b_i \in \mathbb{Z}$,
+в качестве $m$ можно взять длину любого ребра $M$.
 \end{definition}
@@
 TODO: надо ли писать доказательство?
-
-TODO: можно ли использовать термин "ребро $M$" или его надо определить? Где?
 
@@
 
-TODO: не знаю, доказывалось ли это ранее. Скорее всего, да, но где - неизвестно.
-
 В отличие от диаметра и мощности, характеристика в некотором смысле локальна:

### 2018-08-25 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Better linear bound - iteration, from draft to not so sloppy text

@@
 
+\begin{theorem}[\cite{costa2013valid}]
+	\label{thm:varphi_k_bounds}
+	\cite{costa2013valid}
+	Для $k\geq 2$ выполнено
+	\begin{equation*}
+		\sqrt{\frac{2}{k\sqrt{3}}}
+		\leq
+		\varphi_k
+		\leq
+		\frac{1}{k-1} +
+		\sqrt{
+			\frac{1}{(k-1)^2}
+			+
+			\frac{2}{(k-1)\sqrt{3}}
+		}
+	\end{equation*}
+\end{theorem}
+
 \begin{lemma}
@@
 	\begin{equation}
-		d \geq \mu (|S| - 2),
+		d \geq \mu_k (|S| - 2),
 	\end{equation}
@@
 	\begin{equation}
-		\mu = \frac{\sqrt{64\varphi_k^2 (k-1)+1}-1}{16\varphi_k^2 (k-1)}
+		\mu_k = \frac{\sqrt{64\varphi_k^2 (k-1)+1}-1}{16\varphi_k^2 (k-1)}
 	\end{equation}
@@
 \begin{proof}
-	Впишем $S$ в квадрат со стороной $d$ (см. лемму~\ref{lemma:square_container})
-	и разобьём этот квадрат на $m^2$ маленьких квадратов со стороной $d/m$.
+	Множество $S$ содержится в квадрате со стороной $d$ (см. лемму~\ref{lemma:square_container}).
+	Разобьём этот квадрат на $m^2$ маленьких квадратов со стороной $d/m$.
 	Тогда по принципу Дирихле либо:
@@
 	\begin{equation}
-		\mu = \frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2 (k-1)},
+		\mu_k = \frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2 (k-1)},
 	\end{equation}
@@
 		>
-		\mu
+		\mu_k
 	\end{equation}
@@
 	\begin{equation}
-		d > \mu \cdot (|S|-2)
+		d > \mu_k \cdot (|S|-2)
 		.
@@
 
-В~\cite{costa2013valid} приведены значения $\varphi_k$ до $k=10$ включительно
+У только что полученной оценки есть две существенных особенности:
+во-первых, она справедлива лишь для мощности специального вида,
+во-вторых, зависит от $\varphi_k$.
 
-TODO: разобраться, насколько эти значения точны и откуда они взяты =/
+\begin{corollary}
+	\label{corollary:common_linear_bound}
+	Пусть $S\in\mathfrak{M}_n$, $ (k-1)m^2 + 2 \leq n \leq (k-1)(m+1)^2 + 1$.
+	Тогда $\operatorname{diam} S \geq \frac{m^2}{(m+1)^2} \mu_k \cdot |S|$.
+\end{corollary}
 
-По их расчётам получаем $\varphi_{10} = 0.4214$,
-что даёт оценку $d>0.3583(|S|-2)$.
+Выбор $k$ во многом произволен.
+Так, например, для $k=29$ имеем $\varphi_k=0.226882900744209089183505212941$
 
-Это, конечно, для $|S|$ специального вида, но множитель $(m+1)^2 / m^2$ съестся на бесконечности,
-а итоговые коэфициенты и так будем брать с запасом (кто их там знает, как они округляли).
+TODO: ссылка с http://hydra.nat.uni-magdeburg.de/packing/csq/
 
-Это для $|S| \geq 41$, но для меньших мощностей посчитано (и не один раз, в том числе нами).
+(округление здесь и далее осуществляется в сторону уменьшения $\mu_k$, чтобы не испортить неравенство)
+и $\mu_k > 0.3753$.
+Следовательно, верно
+\begin{proposition}
+	Для достаточно больших $n$ диаметр любого множества $M\in\mathfrak{M}_n$ не менее $\frac{3}{8}n$.
+\end{proposition}
+Заметим, что для $n=3$ эта оценка, очевидно, неверна.
 
-Кстати, для $|S| = 3$ не проходит (и не должно).
+Выбрав другое $k$, можно получить оценку, применимую, начиная с некоторого известного $n$.
+Например, $\varphi_{10} = 0.42127954398390343276882176065...$
+(это число не выражается в радикалах TODO:ссылка) и
+$\mu_k > 0.3584$.
+Для $4 \leq n \leq 122$ вычислены точные нижние оценки на диаметр (см. утверждение~\ref{proposition:d(2,n)}) и верно, что
+\begin{equation}
+	\label{eq:linear_bound_10}
+	\operatorname{diam} M \geq 0.3584 \cdot |M|
+\end{equation}
+Кроме того, для $M \in \mathfrak{M}_n$, $n \geq 122$ выполнено
+$\operatorname{diam} M \geq 9905$.
+Следовательно, для $4 \leq n \leq 27636$ выполнено~\eqref{eq:linear_bound_10}.
+Тогда с учётом следствия~\ref{corollary:common_linear_bound} получаем
+\begin{proposition}
+	\label{proposition:linear_bound_29}
+	Для всех $n \geq 4$, $M\in\mathfrak{M}$
+	\begin{equation*}
+		\operatorname{diam} M \geq \frac{55^2}{56^2} \cdot 0.3584 \cdot |M|
+		>
+		0.3457 \cdot |M|
+		.
+	\end{equation*}
+\end{proposition}
 
-А особенная приятность в том, что вычисление новых $\varphi_k$ может давать автоматическое улучшение оценки
-(может, правда, и не давать).
-Однако, как ни печально, константу не получится сделать даже 0.42 (есть нижняя оценка на $\varphi_k$).
+\begin{remark}
+	Нахождение $\varphi_k$ для $k > 30$ может позволить улучшить оценку в утверждении~\ref{proposition:linear_bound_29},
+	равно как и вычисление точных нижних оценок на диаметр для конкретных $n>122$,
+	но лишь до определённого предела, поскольку $\mu_k$ зависит от $\varphi_k$ и $k$
+	только в составе выражения $\varphi_k^2 (k-1)$,
+	которое ограничено в силу теоремы~\ref{thm:varphi_k_bounds}.
+\end{remark}
 
-Следующий прикол:
-де-факто нам нужна не какая попало расстановка точек в единичном квадрате,
-а расстановка с рациональными расстояниями (они потенциально могут превратиться в целые при умножении на $d/m$).
-А возможность аппроксимировать произвольную расстановку расстановкой с рациональными расстояниями
-уже имеет прямое отношение к проблеме Улама-Эрдёша (ссылка!!).
+\begin{remark}
+	Де-факто нам требуется не произвольная расстановка точек в единичном квадрате,
+	а такая расстановка, при которой все расстояния между точками~--- рациональные числа.
+	Это требование уже имеет отношение к проблеме Улама-Эрдёша (TODO:ссылка!).
+\end{remark}

### 2018-08-26 vmmsh-2018/ReviewChar.tex
[vmmsh] Proof that 2 charact. defs are equivalent

@@
 
-Опираясь на работы~\cite{our-mkmitu-2016,our-ped-2017} и используя~\cite[теорема 3.1]{polygons-on-lattices},
+Опираясь на работы~\cite{our-mkmitu-2016,our-ped-2017} и используя,
 можно доказать, что определение~\ref{def_char_classic} равносильно следующему определению.
 \begin{definition}
-	\label{def_char_classic}
+	\label{def_char_lattice}
 	Характеристикой множества $M\in\mathfrak{M}_n$ называется свободное от квадратов
@@
 
-TODO: надо ли писать доказательство?
+\begin{theorem}
+	Определения~\ref{def_char_classic} и~\ref{def_char_lattice} эквивалентны.
+\end{theorem}
+\begin{proof}
+	Пусть $M\in\mathfrak{M}_n$, $M_1,M_2,M_3 \in M$ и не лежат на одной прямой.
+	Пусть сначала $M_1,M_2,M_3$ лежат на решётке~\eqref{grid_for_Erdosh_system}.
+	Тогда по формуле Пика~\cite[теорема 3.1]{polygons-on-lattices} площадь треугольника $M_1 M_2 M_3$
+	соизмерима с $\sqrt{p}$.
+	Пусть теперь площадь треугольника $M_1 M_2 M_3$ соизмерима с $\sqrt{p}$ и $|M_1 - M_2|=m$.
+	Введём систему координат так, что $M_1 = (0,0)$,
+	$M_2 = (|M_1 - M_2|, 0)$,
+	$M_3 = (x_3, y_3)$.
+	Тогда площадь треугольника $M_1 M_2 M_3$ равна $\frac{1}{2} |M_1 - M_2| y_3$,
+	откуда $y_3$ соизмерим с $\sqrt{p}$.
+	Более того, выполнены равенства
+	\begin{gather*}
+		x_3^2 + y_3^2 = |M_1 - M_3|^2
+		\\
+		(m^2 - x_3^2)^2 = |M_2 - M_3|^2
+		.
+	\end{gather*}
+	Вычитая одно из другого, получаем $2mx_3 = a$, где $a$~--- целое число.
+\end{proof}
 

### 2018-08-26 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Lemma about edges and diagonals

@@
 
+Нам потребуется вспомогательная
+\begin{lemma}
+	\label{lemma:quadr_diag_edges}
+	Пусть $ABCD$~--- выпуклый четырёхугольник на плоскости.
+	Тогда $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
+	т.е. хотя бы одна диагональ больше хотя бы одной стороны.
+\end{lemma}
+\begin{proof}
+	Предположим, что любая диагональ меньше любой стороны.
+	Тогда в каждом треугольнике, образованном диагональю и двумя сторонами,
+	меньший угол лежит против меньшей стороны,
+	т.е. угол между сторонами, лежащий против диагонали, острый.
+	Таким образом, получаем, что все 4 угла четырёхугольника~---
+	острые, чего быть не может.
+\end{proof}
+
+Докажем теперь линейную оценку снизу на диаметр множеств из $\mathfrak{M}_n$,
+где $n$ имеет специальный вид.
 \begin{lemma}
@@
 	$|M_3 - M_4| \leq \varphi_k d/m$.
+	В силу леммы~\ref{lemma:quadr_diag_edges} это всегда можно сделать так,
+	чтобы интервалы $M_1 M_2$ и $M_3 M_4$ не пересекались.
 	Возможное совпадение $M_2$ и $M_3$ не помешает дальнейшему доказательству.
-
-	TODO: реверанс про выпуклый четырёхугольник.
 

### 2018-09-17 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Minor strikes

@@
 	\label{corollary:common_linear_bound}
-	Пусть $S\in\mathfrak{M}_n$, $ (k-1)m^2 + 2 \leq n \leq (k-1)(m+1)^2 + 1$.
+	Пусть $S\in\mathfrak{M}_n$, $$ (k-1)m^2 + 2 \leq n \leq (k-1)(m+1)^2 + 1.$$
 	Тогда $\operatorname{diam} S \geq \frac{m^2}{(m+1)^2} \mu_k \cdot |S|$.
@@
 Выбор $k$ во многом произволен.
-Так, например, для $k=29$ имеем $\varphi_k=0.226882900744209089183505212941$
-
-TODO: ссылка с http://hydra.nat.uni-magdeburg.de/packing/csq/
-
-(округление здесь и далее осуществляется в сторону уменьшения $\mu_k$, чтобы не испортить неравенство)
+Так, например, для $k=29$ имеем $\varphi_k=0.226882900744209089183505212941...$
+~\cite{markot2005newverified}
+%(округление здесь и далее осуществляется в сторону уменьшения $\mu_k$, чтобы не испортить неравенство)
 и $\mu_k > 0.3753$.
@@
 \begin{proposition}
+	\label{proposition:linear_bound_29}
 	Для достаточно больших $n$ диаметр любого множества $M\in\mathfrak{M}_n$ не менее $\frac{3}{8}n$.
@@
 Например, $\varphi_{10} = 0.42127954398390343276882176065...$
-(это число не выражается в радикалах TODO:ссылка) и
+(для этого числа неизвестно выражение в радикалах)~\cite{degroot1990optimal10points} и
 $\mu_k > 0.3584$.
@@
 \begin{proposition}
-	\label{proposition:linear_bound_29}
+	\label{proposition:linear_bound_10}
 	Для всех $n \geq 4$, $M\in\mathfrak{M}$
@@
 \begin{remark}
-	Нахождение $\varphi_k$ для $k > 30$ может позволить улучшить оценку в утверждении~\ref{proposition:linear_bound_29},
-	равно как и вычисление точных нижних оценок на диаметр для конкретных $n>122$,
+	Нахождение $\varphi_k$ для $k > 30$ может позволить улучшить оценку
+	в утверждениях~\ref{proposition:linear_bound_29} и~\ref{proposition:linear_bound_10},
+	равно как и вычисление точных нижних оценок на диаметр для конкретных $n>122$ (в утверждении~\ref{proposition:linear_bound_10}),
 	но лишь до определённого предела, поскольку $\mu_k$ зависит от $\varphi_k$ и $k$
@@
 	а такая расстановка, при которой все расстояния между точками~--- рациональные числа.
-	Это требование уже имеет отношение к проблеме Улама-Эрдёша (TODO:ссылка!).
+	Это требование уже имеет отношение к проблеме Улама-Эрдёша (см. гипотезу~\ref{hypot:Ulam}).
 \end{remark}

### 2018-09-17 vmmsh-2018/ReviewMultidim.tex
[vmmsh] Iteration
[vmmsh] Cosmetics
[vmmsh] Minor strikes

@@
 
+Более того, аналогичные множества были построены в $\mathbb{Z}^m_n$~\cite{Kohnert2006IntegralPS}.
+
 Характеристика (наряду с диаметром и мощностью) обобщается на случай, когда множество точек с целочисленными расстояниями
-рассмативается в $\mathbb{R}^k$~\cite{kurz2005characteristic}.
+рассмативается в $\mathbb{R}^m$~\cite{kurz2005characteristic}.
 
@@
 
-Имеют место быть нижеследующие оценки на $d(m,n)$
+\begin{definition}
+	Множество $M\in\mathfrak{M}(m,n)$ называется оптимальным,
+	если $\operatorname{diam} M = d(m,n)$.
+\end{definition}
+
+Приведём известные оценки на $d(m,n)$
 (за основу взят список из~\cite{kurz2008bounds}):
@@
 	& d(m, n) \leq \begin{cases}
-		2^{n-m+1} -2 & \mbox{для} n-m \equiv 0 \mod 2
+		2^{n-m+1} -2 & \mbox{~для~} n-m \equiv 0 \mod 2
 		\\
-		3(2^{n-m} -1) & \mbox{для} n-m \equiv 1 \mod 2
+		3(2^{n-m} -1) & \mbox{~для~} n-m \equiv 1 \mod 2
 	\end{cases} \qquad \mbox{\cite{harborth1985diameters}} \\
-	& d(2,n) \geq cn \qquad \mbox{TODO: вставить константу} \\
+	& d(2,n) \geq \frac{3}{8}n \qquad \mbox{для достаточно больших $n$ (утверждение~\ref{proposition:linear_bound_29})} \\
+	& d(2,n) > 0.3457 n \qquad \mbox{для $n\geq 4$ (утверждение~\ref{proposition:linear_bound_10})} \\
 	& d(m,n) \leq (n-m)^{c \log \log (n-m)} \mbox{~для некоторого $c$} \qquad \mbox{\cite{harborth1993upper}} \label{upper_bound_d_m_n_log_log}\\
-	& d(m,n) > \sqrt{\frac{3}{2m}} n^{1/m} \mbox{\cite{kanold1981punktmengen},в частных случаях улучшена в \cite{nozaki2013lower}} \\
+	& d(m,n) > \sqrt{\frac{3}{2m}} n^{1/m} \mbox{~\cite{kanold1981punktmengen}, в частных случаях улучшена в \cite{nozaki2013lower}} \\
 	& d(3,n) > \frac{1}{\sqrt{14}}n^{1/2} \mbox{~для~} n \geq 5 \quad \mbox{\cite{kanold1981punktmengen}} \\
@@
 
-Кроме того, Курц~\cite{kurz2008bounds} выдвинул предположение, что $d(m - 1, n) \geq d(m, n)$.
+Кроме того, в~\cite{kurz2008bounds} выдвинуто предположение, что $d(m - 1, n) \geq d(m, n)$.
 
@@
 следующие известные значения~--- в \cite{kurz2005characteristic,kurz2006konstruktion,kurz2008minimum,kurz2008bounds,our-mz-rus}.
-Курцу удалось существенно оптимизировать алгоритм, основанный на переборе, применяя характеристику~\cite{kurz2005characteristic,kreisel2008heptagon}.
-Однако сложность вычисления максимальной возможной мощности для заданного диаметра $d$, по оценке Курца~\cite{kreisel2008heptagon},
+Алгоритм, основанный на переборе, может быть существенно опимизирван с применением характеристики~\cite{kurz2005characteristic,kreisel2008heptagon}.
+Однако сложность вычисления максимальной возможной мощности для заданного диаметра $d$, по оценке~\cite{kreisel2008heptagon},
 составлет $O(d^3)$ (в двумерном случае).
-Курц (TODO: ссылка!) отмечает, что важно перепроверять вычисленные значения независимо составленными программами для ЭВМ.
+%В (TODO: ссылка!) отмечено, что важно перепроверять вычисленные значения независимо составленными программами для ЭВМ.
 Приведём некоторые известные значения:
 
-TODO: cite
-\\
-$
-(d(2, n))_{n=3,...,122} = 1, 4, 7, 8, 17, 21, 29, 40, 51, 63, 74, 91, 104, 121,
-134, $\\$153, 164,
-196, 212, 228, 244, 272, 288, 319, 332, 364, 396, 437, 464, 494, 524,$\\$ 553, 578, 608,
-642, 667, 692, 754, 816, 897, 959, 1026, 1066, 1139, 1190, $\\$ 1248, 1306, 1363, 1410,
-1460, 1514, 1564, 1614, 1675, 1727, 1770, 1817, $\\$1887, 1906, 2060, 2140, 2169,
-2231, 2299, 2432, 2494, 2556, 2624, 2692, $\\$2827, 2895, 2993, 3098, 3196, 3294,
-3465, 3575, 3658, 3749, 3885, 3922, $\\$4223, 4380, 4437, 4559, 4693, 4883,
-5018, 5109, 5264, 5332, 5480, 5603, $\\$5738, 5938, 5995, 6052,
-6324, 6432, 6630, 6738, 6939, 7061, 7245, 7384, $\\$7568, 7752, 7935, 8119, 8321,
-8406, 8648, 8729, 8927, 9052, 9211, 9423, $\\$9534, 9794, 9905
-$.
+\begin{proposition}
+\label{proposition:d(2,n)}
+\end{proposition}
+$(d(2, n))_{n=3,...,122} = 1$, 4, 7, 8, 17, 21, 29, 40, 51, 63, 74, 91, 104, 121,
+134, 153, 164,
+196, 212, 228, 244, 272, 288, 319, 332, 364, 396, 437, 464, 494, 524, 553, 578, 608,
+642, 667, 692, 754, 816, 897, 959, 1026, 1066, 1139, 1190,  1248, 1306, 1363, 1410,
+1460, 1514, 1564, 1614, 1675, 1727, 1770, 1817, 1887, 1906, 2060, 2140, 2169,
+2231, 2299, 2432, 2494, 2556, 2624, 2692, 2827, 2895, 2993, 3098, 3196, 3294,
+3465, 3575, 3658, 3749, 3885, 3922, 4223, 4380, 4437, 4559, 4693, 4883,
+5018, 5109, 5264, 5332, 5480, 5603, 5738, 5938, 5995, 6052,
+6324, 6432, 6630, 6738, 6939, 7061, 7245, 7384, 7568, 7752, 7935, 8119, 8321,
+8406, 8648, 8729, 8927, 9052, 9211, 9423, 9534, 9794, 9905
+.
+
+Для многомерного случая были вычислены~\cite{piepmeyer1988raumliche,harborth1998integral,kurz2005characteristic,kurz2006konstruktion,kurz2008bounds} следующие значения.
+\begin{proposition}~
+\end{proposition}
+$(d(3, n))_{n=4,...,24} = 1$, 3, 4, 8, 13, 16, 17, 17, 17, 56, 65, 77, 86, 99, 112, 133, 154,
+195, 212, 228, 244.
+
+\begin{hypothesis}
+	Для любого $n\geq 3$ выполнено $d(2,n) < d(2,n+1)$.
+\end{hypothesis}
+В пользу этой гипотезы говорят известные значения $d(2,n)$.
+Для больших размерностей утверждение гипотезы неверно: так, $d(3,10) = d(3,11) = 17$.
+
+\begin{hypothesis}
+	Никакое оптимальное множество $M\in\mathfrak{M}$ не лежит на целочисленной решётке.
+\end{hypothesis}
+В пользу этой гипотезы говорят реузльтаты вычислений для $m=2$, $3 \leq n \leq 43$.
 
 Примечательно, что для $m=2$, $n=4$ и $n=18$ оптимальные множества не единственны.
-Курц выяснил, что для $9 \leq n \leq 122$ оптимальное множество содержится
-в объединении прямой и точки (т.е. все точки, кроме одной, лежат на некоторой прямой).
+В~\cite{kurz2008bounds} установлено, что для $9 \leq n \leq 122$ оптимальное множество является веерным, т.е. содержится
+в объединении прямой и точки (все точки, кроме одной, лежат на некоторой прямой).
 
-Для многомерного случая вычисленные значения можно найти в~\cite{kurz2008bounds}.
+\begin{hypothesis}
+	Все оптимальные множества из $n$ точек на плоскости, $n\geq 9$, являются веерными.
+\end{hypothesis}
 
 При изучении множеств из $\mathfrak{M}_n$ часто возникают множества специальных видов.
-Наиболее простым примером являются веерные множества, т.е. множества, содержащиеся в объединении прямой и точки.
+Наиболее простым примером являются веерные множества.
 Курц и Вассерман~\cite{kurz2008minimum} показали, что диаметр веерного множества мощности $n$
 не менее, чем $n^{c \log \log n}$, где $c$~--- некоторая константа.
+Это результат указывает на возможность улучшения оценки в следствии~\ref{corollary:max_points_on_line}.
 
-TODO: про лемму
+В работе~\cite{kurz2008minimum} высказана
+\begin{hypothesis}
+	\label{hypot:log_lower_bound}
+	Существует такая константа $c$,
+	что для любого $n\geq 3$ и $M\in\mathfrak{M}_n$ выполнено
+	$\operatorname{diam} M \geq n^{c \log \log n}$.
+\end{hypothesis}
 
-В этой же работе высказано предположение, что подобная оценка верна для всех множеств из $\mathfrak{M}_n$.
-Заметим, что эта оценка совпадает с верхней оценкой~\eqref{upper_bound_d_m_n_log_log} на $d(2,n)$
+Заметим, что оценка гипотезы~\ref{hypot:log_lower_bound}
+совпадает с верхней оценкой~\eqref{upper_bound_d_m_n_log_log} на $d(2,n)$
 с точностью до константы.
-Это вкупе с тем фактом, что большинство известных оптимальных множеств являются веерными,
-позволяет предположить, что все оптимальные множества являются веерными.
 
@@
 
+Положив $p=1$ в доказательстве теоремы~\ref{thm:power_char_exist} и добавив к построеннному множеству
+точку $M_{n+2} = (0, -\sqrt{2^{n+1}})$, где $n$ нечётно,
+получим множество из $n+2$ точек, из которых $n$ лежат на некоторой прямой и 2 вне её.
+
+В~\cite{our-vvmsh-2018} приведены примеры различных множеств из $\mathfrak{M}_n$:
+\begin{itemize}
+	\item
+		$
+		\left\{
+		(\pm 21/2, 0);
+		(\pm 4, 3\sqrt{35}/2);
+		(\pm 5/2, 0);
+		(\pm 11/2, 0);
+		\right\}
+		$~---
+		содержится в двух параллельных прямых (2 точки на одной, 6 на другой);
+
+	\item
+		$
+		\left\{
+		\left( {57/34} ; -{12\sqrt{35}/17}\right);
+		\left( {44/17} ; {39\sqrt{35}/34}\right);
+		\left( -{72/17} ;
+		{15\sqrt{35}/34}\right);
+		\right.$ \\ $\left.
+		\left( -{175/34} ; -{24\sqrt{35}/17}\right);
+		\left( -{57/34} ; {12\sqrt{35}/17}\right);
+		%\right.$ \\ $\left.
+		\left( \pm{17/2} ; 0\right);
+		\right\}
+		$~---
+		содержится в двух параллельных прямых (3 точки на одной, 4 на другой);
+
+	\item
+		$
+		\left\{
+		(0,0);(\pm 50, 0); (\pm 14, \pm 48)
+		\right\}
+		$~---
+		содержится в объединении окружности с её центром (7~точек);
+
+	\item
+		$
+		\left\{
+		\left( \pm30 ; 0\right);
+		%\left( 30 ; 0\right);
+		\left( \pm{15/2} ; {25\sqrt{7}/2}\right);
+		%\left( {15/2} ; {25\sqrt{7}/2}\right);
+		%\left( {33/2} ; {9\sqrt{7}/2}\right);
+		\left( 0 ; {10\sqrt{7}/1}\right);
+		\left( \pm{33/2} ; {9\sqrt{7}/2}\right);
+		\right\}
+		$~---
+		содержится в двух прямых, пересекающихся под острым углом (7~точек).
+\end{itemize}
+
+Информация о некоторых других конструкциях может быть найдена в~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved}.
+
+TODO: спрямить ссылки!
+
 \begin{definition}
-	\cite{kurz2008minimum} Множество $M\in\mathfrak{M}_n$ называется множеством полуобщего положения,
-	если ни через какие три точки из $M$ нельзя провести прямую.
-	Множество всех множеств полуобщего положения мощности $n$ будем обозначать $\overline{\mathfrak{M}}_n$.
+	\cite{kurz2008minimum} Множество $M\in\mathfrak{M}(m,n)$ называется множеством полуобщего положения,
+	если ни через какие $m+1$ точек из $M$ нельзя провести $(m-1)$--мерную гиперплоскость.
+	Множество всех множеств полуобщего положения мощности $n$ в $\mathbb{R}^m$ будем обозначать $\overline{\mathfrak{M}}(m,n)$.
+	Для краткости обозначим $\overline{\mathfrak{M}}_n=\overline{\mathfrak{M}}(2,n)$.
 \end{definition}
@@
 \begin{definition}
-	\cite{kurz2008minimum} Множество $M\in\overline{\mathfrak{M}}_n$, $n\geq 4$ называется множеством общего положения,
-	если ни через какие четыре точки из $M$ нельзя провести окружность.
-	Множество всех множеств общего положения мощности $n$ будем обозначать $\dot{\mathfrak{M}}_n$.
+	\cite{kurz2008minimum} Множество $M\in\overline{\mathfrak{M}}(m,n)$, $n\geq 4$ называется множеством общего положения,
+	если ни через какие $m+2$ точек из $M$ нельзя провести сферу.
+	Множество всех множеств общего положения мощности $n$ будем обозначать $\dot{\mathfrak{M}}(m,n)$.
+	Для краткости обозначим $\dot{\mathfrak{M}}_n=\dot{\mathfrak{M}}(2,n)$.
 \end{definition}
@@
 
+\begin{hypothesis}
+	\label{hypot:general_position_nonempty}
+	Для любого $n\geq 4$ непусто $\dot{\mathfrak{M}}_n$.
+\end{hypothesis}
+
 \begin{definition}
@@
 
-1. Особые случаи
-
-1.5. Системы общего положения (гипотеза Эрдёша, работа Курца)
-\cite{kreisel2008heptagon}
-
-1.6. 7-кластеры (Курц)
-
 Интересны, на взгляд авторов, и ещё несколько результатов для множеств из $\mathfrak{M}_n$ специального вида.
@@
 
-TODO: про сущ-е множеств с миниметром 1.
+Известно, что для $3\leq n \leq 5$ существуют множества из $\mathfrak{M}_n$
+c минимальной длиной ребра 1.
+Все известные примеры являются веерными~\cite{harborth1993upper}.
+Существуют ли множества произвольной мощности с длиной ребра 1,
+неизвестно, однако легко доказать следующее
+\begin{proposition}
+	Для $n > 4$ не существует таких $M\in\overline{\mathfrak{M}}_n$,
+	что для некоторых $M_1, M_2$ выполнено $|M_1 - M_2| = 1$.
+\end{proposition}
+\begin{proof}
+	Пусть $M\in\overline{\mathfrak{M}}_n$ и для некоторых $M_1, M_2$ выполнено $|M_1 - M_2| = 1$.
+	Тогда все точки множества $M$ лежат на $cr(M_1,M_2)$.
+	$cr(M_1,M_2)$ есть объединение двух прямых, на каждой из которых может быть не более 2 точек в силу того, что
+	$M\in\overline{\mathfrak{M}}_n$.
+\end{proof}
+Для $n=3$ утверждение неверно.
+Существует ли множество $M\in\mathfrak{M}_4$ c ребром единичной длины,
+авторам незвестно, но если такое $M$ существует, то $\operatorname{char} M = 1$.
 
-Если $M\in\mathfrak{M}(m,n)$ и все рёбра $M$ имею чётную длину,
+Если $M\in\mathfrak{M}(m,n)$ и все рёбра $M$ имеют чётную длину,
 то понятно, что множество $H$, полученное из $M$ сжатием в два раза,
@@
 и конструктивно показана точность этой оценки.
+Множество, на котром эта оценка достигается, лежит на окружности.
 
-Отдельные оценки количества рёбер нечётной длины для множеств из $\overline{\mathfrak{M}}_n$ и $\dot{\mathfrak{M}}_n$
+Отдельные оценки количества рёбер нечётной длины для множеств из $\dot{\mathfrak{M}}_n$
 авторам неизвестны.
 
-Для всех известных значений d(2,n) < d(2,n+1).
-Для больших размерностей это неверно (ссылка, пример).
-
-3. Связанные проблемы: гипотеза Эрдёша о числе расстояний, гипотеза Улама--Эрдёша.
-Ссылка на 2 работы Гарибальди. Ссылка на доказательство гипотезы Эрдёша о числе расстояний.
-
-4. Примеры множеств.
+TODO: выкинуть следующий абзац?
 
@@
 например, поиск систем из $\mathfrak{M}_n$, содержащихся в двух прямых (параллельных или пересекающихся).
+
+С исследованием множеств $\mathfrak{M}_n$ тесно связаны ещё две гипотезы о расстояниях на плоскости.
+\begin{hypothesis}[Улам, \cite{ulam1960collection}]
+	\label{hypot:Ulam}
+	Существует всюду плотное подмножество $U\subset\mathbb{R}^2$, такое,
+	что для любых $U_1, U_2 \in U$ выполнено $|U_1 - U_2|\in\mathbb{Q}$,
+	где $\mathbb{Q}$~--- множество рациональных чисел.
+\end{hypothesis}
+Если гипотеза~\ref{hypot:Ulam} справедлива, то для любого $n\geq 4$ множество $\dot{\mathfrak{M}}_n$ непусто
+(гипотеза~\ref{hypot:general_position_nonempty}).
+Однако авторам неизвестны доказательства того, что из справедливости
+гипотезы~\ref{hypot:general_position_nonempty} следует справедливость гипотезы~\ref{hypot:Ulam}.
+
+В работах~\cite{garibaldi2005erdos,garibaldi2011erdos} исследуется связь множеств $\mathfrak{M}_n$ и
+проблемы Эрдёша о числе различных расстояний~\cite{erdos1946sets}.
+В работе~\cite{guth2015erdos} эта проблема в своей изначальной постановке была решена и доказана следующая
+\begin{theorem}
+	Пусть $G\subset{R}^2$, $|G| = n$.
+	Тогда количество различных расстояний $|G_i - G_j|, G_i,G_j \in G$
+	не менее $cn/\log n $.
+\end{theorem}
+
+TODO: Примеры множеств - рисунки (если будет время и место).

### 2018-09-18 vmmsh-2018/BetterLinearBound.tex
[vmmsh] Minnor strikes

@@
 	Не существует такого числа $\gamma < 1$,
-	что любое множество из $M\in \mathfrak{M}_n$, $\geq 3$,
+	что любое множество из $M\in \mathfrak{M}_n$, $n\geq 3$,
 	содержится в квадрате со стороной $\gamma \cdot \operatorname{diam} M$.
@@
 	$M_1 \neq M_2$, $M_3 \neq M_4$,
-	то $cr(M_1,M_2) \cap cr(M_3,M_4)$~--- либо не более 4 точек, либо прямая.
+	то множество $cr(M_1,M_2) \cap cr(M_3,M_4)$~--- либо прямая, либо содержит не более 4 точек.
 \end{proposition}
-
-TODO: напротив этого утверждения стояла галочка, но мы её не обсудили.
 
@@
 
-\begin{theorem}[\cite{costa2013valid}]
+\begin{theorem}%[\cite{costa2013valid}]
 	\label{thm:varphi_k_bounds}
@@
 Кроме того, для $M \in \mathfrak{M}_n$, $n \geq 122$ выполнено
-$\operatorname{diam} M \geq 9905$.
+$\operatorname{diam} M \geq 9905$ (см. утверждение~\ref{proposition:d(2,n)}).
 Следовательно, для $4 \leq n \leq 27636$ выполнено~\eqref{eq:linear_bound_10}.

### 2018-09-18 vmmsh-2018/ReviewChar.tex
[vmmsh] Minnor strikes

@@
 	\cite[гл. 34, п. 3]{Bukhstab-number-theory}
-	Натуральное число называется свободным от квдратов,
+	Натуральное число называется свободным от квадратов,
 	если оно не делится ни на один квадрат простого числа.
@@
 
+Корректность определения~\ref{def_char_classic} показана в~\cite{kemnitz1988punktmengen}.
+
 Пусть $\mathbb{Z}$~--- множество целых чисел;
@@
 
-Опираясь на работы~\cite{our-mkmitu-2016,our-ped-2017} и используя,
+Опираясь на работы~\cite{our-mkmitu-2016,our-ped-2017},
 можно доказать, что определение~\ref{def_char_classic} равносильно следующему определению.

### 2018-09-18 vmmsh-2018/ReviewMultidim.tex
[vmmsh] Minnor strikes

@@
-В пространствах более высокой размерности можно построить множества, аналогичные $\mathfrak{M}$.
+Конструкцию $\mathfrak{M}_n$ можно распространить на пространства более высокой размерности.
+
+%В пространствах более высокой размерности можно построить множества, аналогичные $\mathfrak{M}$.
 
@@
 
-Более того, аналогичные множества были построены в $\mathbb{Z}^m_n$~\cite{Kohnert2006IntegralPS}.
+Более того, конструкцию $\mathfrak{M}_n$ можно распространить и на $\mathbb{Z}^m_n$~\cite{Kohnert2006IntegralPS}.
+%Более того, аналогичные множества были построены в $\mathbb{Z}^m_n$~\cite{Kohnert2006IntegralPS}.
 
 Характеристика (наряду с диаметром и мощностью) обобщается на случай, когда множество точек с целочисленными расстояниями
-рассмативается в $\mathbb{R}^m$~\cite{kurz2005characteristic}.
+рассматривается в $\mathbb{R}^m$~\cite{kurz2005characteristic}.
 
@@
 следующие известные значения~--- в \cite{kurz2005characteristic,kurz2006konstruktion,kurz2008minimum,kurz2008bounds,our-mz-rus}.
-Алгоритм, основанный на переборе, может быть существенно опимизирван с применением характеристики~\cite{kurz2005characteristic,kreisel2008heptagon}.
+Алгоритм, основанный на переборе,
+может быть существенно оптимизирован с применением характеристики~\cite{kurz2005characteristic,kreisel2008heptagon}.
 Однако сложность вычисления максимальной возможной мощности для заданного диаметра $d$, по оценке~\cite{kreisel2008heptagon},
@@
 
-Заметим, что оценка гипотезы~\ref{hypot:log_lower_bound}
+Заметим, что неравенство в гипотезе~\ref{hypot:log_lower_bound}
 совпадает с верхней оценкой~\eqref{upper_bound_d_m_n_log_log} на $d(2,n)$
@@
 
-Интересны, на взгляд авторов, и ещё несколько результатов для множеств из $\mathfrak{M}_n$ специального вида.
-
-Так, Солимоси построил~\cite{solymosi2003note} для любого $n\geq 3$ конструкцию множества $M\in\mathfrak{M}_n$
+Солимоси построил~\cite{solymosi2003note} для любого $n\geq 3$ конструкцию множества $M\in\mathfrak{M}_n$
 такого, что для некоторых $M_1, M_2 \in M$ выполнено $|M_1 - M_2| = 2$.
@@
 \begin{proposition}
-	Для $n > 4$ не существует таких $M\in\overline{\mathfrak{M}}_n$,
+	\label{prop:edge_one}
+	Для $n \geq 4$ не существует таких $M\in\overline{\mathfrak{M}}_n$,
 	что для некоторых $M_1, M_2$ выполнено $|M_1 - M_2| = 1$.
@@
 	$cr(M_1,M_2)$ есть объединение двух прямых, на каждой из которых может быть не более 2 точек в силу того, что
-	$M\in\overline{\mathfrak{M}}_n$.
+	$M\in\overline{\mathfrak{M}}_n$,
+	откуда непосредственно следует требуемое утверждение для $n\geq 5$.
+
+	Пусть теперь $n=4$.
+	Предположим противное.
+	Пусть $M=\{M_1, M_2, M_3, M_4\} \in \overline{\mathfrak{M}}_4$,
+	$M_1 = (-1/2, 0)$, $M_2 = (1/2, 0)$, $O=(0,0)$.
+	Тогда $M_3$ и $M_4$ лежат на оси ординат (см. выше).
+	Так как $|M_3 - M_4| \in \mathbb{Z}$,
+	то площадь треугольника $M_1 M_3 M_4$ есть рациональное число.
+	По определению~\ref{def_char_classic} это означает, что $\operatorname{char} M = 1$.
+	Тогда по определению~\ref{def_char_lattice} имеем $M_3 = (0, a/2)$.
+	Понятно, что $a \neq 0$.
+	Не теряя общности, положим $a\in\mathbb{N}$.
+	Пусть $|M_1 - M_3| = b$.
+	Применяя теорему Пифагора к треугольнику $O M_1 M_3$, получим уравнение
+	\begin{equation*}
+		\frac{1}{4} + \frac{a^2}{4} = b^2
+		,
+	\end{equation*}
+	или, что то же самое,
+	\begin{equation*}
+		(2b)^2 - a^2 = 1
+		.
+	\end{equation*}
+	Это уравнение не имеет решений в натуральных числах.
+	Полученное противоречие завершает доказательство.
 \end{proof}
-Для $n=3$ утверждение неверно.
-Существует ли множество $M\in\mathfrak{M}_4$ c ребром единичной длины,
-авторам незвестно, но если такое $M$ существует, то $\operatorname{char} M = 1$.
+Условие $n \geq 4$ в утверждении~\ref{prop:edge_one} существенно.
+Для $n=3$ существует контрпример: равносторонний треугольник со стороной 1.
 
@@
 
-TODO: выкинуть следующий абзац?
-
 Перспективными направлениями исследований являются, кроме непосредственно отыскания как можно более точных оценок на $d(m,n)$,
@@
 (гипотеза~\ref{hypot:general_position_nonempty}).
-Однако авторам неизвестны доказательства того, что из справедливости
-гипотезы~\ref{hypot:general_position_nonempty} следует справедливость гипотезы~\ref{hypot:Ulam}.
+Неизвестно, следует ли из
+гипотезы~\ref{hypot:general_position_nonempty} следует гипотеза~\ref{hypot:Ulam}.
 

### 2018-09-18 vmmsh-2018/style09/Primer.tex
[vmmsh] The original style

\input style_konf.tex

\setcounter{footnote}{0}
 \setcounter{equation}{0}

\UDC{???}

\Title{НАЗВАНИЕ СТАТЬИ}

 \Author{ФИО}

\Abstract{Текст аннотации. \par}

{\hangindent17pt\hangafter=0\noindent\footnotesize{\bf Ключевые
 слова:} ключевое слово 1, ключевое слово 2, \ldots\par}

\section*{1. Название параграфа}

\Teorema{ 1}Текст теоремы. \Endproc

\beginproof~Текст доказательства.~\endproof

\Lema{ 1}Текст леммы.\Endproc

\section*{Литература}

\begin{enumerate}
 {\footnotesize
 \itemsep=0pt \parskip=0pt
 \leftskip=-10pt

%+
 \bib{Полянин А.~Д.}{Справочник по линейным уравнениям математической фи\-зи\-ки.\,---\,М.: Физматлит, 2001.\,---\,576~c.}

%+
 \bib{Киприянов И.~А., Катрахов В.~В., Ляпин В.~М.} {О краевых задачах в области общего вида для сингулярных
 параболических систем уравнений~/\!/ Докл. АН СССР.\,---\,1976.\,---\,Т.~230, \No~6.\,---\,С.~1271--1274.}

%+
 \bib{Ильин В.~А., Позняк Э.~Г.}{Основы математического анализа.  Ч.~1: Учеб. для вузов~/ 7-е изд.\,---\,М.:
 Физматлит, 2005.\,---\,648~с.}

}
 \end{enumerate}

 \Adress{\textsc{ФИО}\\
Место работы\\
 %старший научный сотрудник (докторант),\\
 Адрес (страна, индекс, город, улица, номер дома)\\
 E-mail}

 \bigskip

 \begin{center}
 НАЗВАНИЕ СТАТЬИ НА АНГЛИЙСКОМ ЯЗЫКЕ
 \end{center}
 \begin{center}
  ФИО на английском языке
 \end{center}

\Abstract{Аннотация на английском языке. \par}

\vspace*{-4pt}

{\hangindent17pt\hangafter=0\noindent\footnotesize {\bf Key
 words:}~Ключевые слова на английском языке.\par}

\end{document}

### 2018-09-23 vmmsh-2018/blg.tex
[vmmsh] Generating the bibliography

\begin{thebibliography}{99}
\bibitem{erdos1945integral}
\emph{Erdös} \emph{P.} Integral distances // Bulletin of the
American Mathematical Society. — 1945. — Т. 51, No 12. — С. 996.
\bibitem{anning1945integral}
\emph{Anning} \emph{N. H.}, \emph{Erdös} \emph{P.}
Integral distances // Bulletin of the American Mathematical
Society. — 1945. — Т. 51, No 8. — С. 598—600.
\bibitem{solymosi2003note}
\emph{Solymosi} \emph{J.} Note on integral distances //
Discrete \& Computational Geometry. — 2003. — Т. 30, No 2. —
С. 337—342.
\bibitem{anning1915discussions}
\emph{Anning} \emph{N.} Discussions: Relating to a Geometric
Representation of Integral Solutions of Certain Quadratic
Equations // The American Mathematical Monthly. — 1915. —
Т. 22, No 9. — С. 321—321.
\bibitem{harborth1993upper}
\emph{Harborth} \emph{H.},
\emph{Möller} \emph{M.} An
diameter of integral point sets
Geometry. — 1993. — Т. 9, No 4.
\emph{Kemnitz} \emph{A.},
upper bound for the minimum
// Discrete \& Computational
— С. 427—432.\bibitem{piepmeyer1996maximum}
\emph{Piepmeyer} \emph{L.} The maximum number of odd
integral distances between points in the plane // Discrete \&
Computational Geometry. — 1996. — Т. 16, No 1. — С. 113—115.
\bibitem{kurz2008bounds}
\emph{Kurz} \emph{S.}, \emph{Laue} \emph{R.} Bounds for
the minimum diameter of integral point sets // arXiv preprint
arXiv:0804.1296. — 2008.
\bibitem{our-vvmsh-2018}
\emph{Авдеев} \emph{Н. Н.} Некоторые примеры систем
Эрдёша // Современные методы теории краевых задач :
материалы международной конференции «Понтрягинские
чтения — XXIX», посвященной 90-летию Владимира
Александровича Ильина (2–6 мая 2018 г.) — 2018. — С. 30—31.
\bibitem{locatelli2002packing}
\emph{Locatelli} \emph{M.}, \emph{Raber} \emph{U.} Packing
equal circles in a square: a deterministic global optimization
approach // Discrete Applied Mathematics. — 2002. — Т. 122,
No 1—3. — С. 139—166.
\bibitem{costa2013valid}
\emph{Costa} \emph{A.} Valid constraints for the point packing
in a square problem // Discrete Applied Mathematics. — 2013. —
Т. 161, No 18. — С. 2901—2909.
\bibitem{markot2005newverified}
\emph{Markót} \emph{M. C.}, \emph{Csendes} \emph{T.} A
new verified optimization technique for the "packing circles in
a unit square" problems // SIAM Journal on Optimization. —
2005. — Т. 16, No 1. — С. 193—219.
\bibitem{degroot1990optimal10points}
\emph{DeGroot} \emph{C.}, \emph{Peikert} \emph{R.},
\emph{Würtz} \emph{D.} The optimal packing of ten equal
circles in a square. — Eidgenössische Technische Hochschule
Zürich, Interdisziplinäres Projektzentrum für Supercomputing,
1990.\bibitem{kemnitz1988punktmengen}
\emph{Kemnitz} \emph{A.} Punktmengen mit ganzzahligen
Abständen. — 1988.
\bibitem{Bukhstab-number-theory}
\emph{Бухштаб} \emph{А.} Теория чисел. — 1966. — С. 384.
\bibitem{our-mkmitu-2016}
\emph{Авдеев} \emph{Н. Н.} Об отыскании множеств точек на
плоскости с целочисленными расстояниями // Математическое
и компьютерное моделирование, информационные технологии
управления: сб. тр. Школы для студентов, аспирантов и
молодых ученых «МКМИТУ-2016». — 2016. — С. 15—19.
\bibitem{our-ped-2017}
\emph{Авдеев} \emph{Н. Н.} О множествах точек на
плоскости с целочисленными расстояниями и алгоритмах их
отыскания // Некоторые вопросы анализа, алгебры, геометрии
и математического образования: материалы международной
молодежной научной школы «Актуальные направления
математического анализа и смежные вопросы». — 2017. — Т. 7,
No I. — С. 9—10.
\bibitem{polygons-on-lattices}
\emph{Вавилов} \emph{В.}, \emph{Устинов}
Многоугольники на решетках. — 2006.
\emph{А.}
\bibitem{Kohnert2006IntegralPS}
\emph{Kohnert} \emph{A.}, \emph{Kurz} \emph{S.} Integral
point sets over $Z^m_n$
// Discrete Applied Mathematics. — 2006. —
Т. 157 (2009). — С. 2105—2117.
\bibitem{kurz2005characteristic}
\emph{Kurz} \emph{S.} On the characteristic of integral point
sets in // arXiv preprint math/0511704. — 2005.
\bibitem{harborth1985diameters}
\emph{Harborth} \emph{H.}, \emph{Kemnitz} \emph{A.}
Diameters of integral point sets // Colloquia Mathematica
Societatis Jano Bolyai. Т. 48. — 1985.
\bibitem{kanold1981punktmengen}
\emph{Kanold}
\emph{H.-J.}
Über
Punktmengen
im
$k$-dimensionalen euklidischen Raum // Abh. Braunschw. Wiss.
Ges. — 1981. — Т. 32, No 55—65. — С. 252.
\bibitem{nozaki2013lower}
\emph{Nozaki} \emph{H.} Lower bounds for the minimum
diameter of integral point sets // Australasian Journal of
Combinatorics. — 2013. — Т. 56. — С. 139—143.
\bibitem{piepmeyer1988raumliche}
\emph{Piepmeyer}
\emph{L.}
Räumliche
ganzzahlige
Punktmengen: дис. . . . канд. / Piepmeyer L. — Master’s
thesis, TU Braunschweig, 1988.
\bibitem{harborth1991point}
\emph{Harborth} \emph{H.}, \emph{Piepmeyer} \emph{L.}
Point sets with small integral distances // Applied geometry and
discrete mathematics: the Victor Klee Festschrift. Т. 4. — American
Mathematical Soc., 1991.
\bibitem{harborth1998integral}
\emph{Harborth} \emph{H.} Integral distances in point sets //
Karl der Grosse und sein Nachwirken. 1200 Jahre Kultur und
Wissenschaft in Europa: Band II, Mathematisches Wissen. —
1998. — С. 213—224.
\bibitem{kurz2006konstruktion}
\emph{Kurz} \emph{S.} Konstruktion und Eigenschaften
ganzzahliger Punktmengen: дис. . . . канд. / Kurz Sascha. — 2006.
\bibitem{kurz2008minimum}
\emph{Kurz} \emph{S.}, \emph{Wassermann} \emph{A.} On
the minimum diameter of plane integral point sets // arXiv
preprint arXiv:0804.1307. — 2008.
\bibitem{our-mz-rus}
\emph{Авдеев} \emph{Н. Н.}, \emph{Семёнов} \emph{Е. М.}
О множествах точек на плоскости с целочисленными
расстояниями // Математические заметки. — 2016. — Т. 100,
No 5. — С. 757—761.\bibitem{kreisel2008heptagon}
\emph{Kreisel} \emph{T.}, \emph{Kurz} \emph{S.} There are
integral heptagons, no three points on a line, no four on a circle //
Discrete \& Computational Geometry. — 2008. — Т. 39, No 4. —
С. 786—790.
\bibitem{brass2006research}
\emph{Brass} \emph{P.}, \emph{Moser} \emph{W. O.},
\emph{Pach} \emph{J.} Research problems in discrete
geometry. — Springer Science \& Business Media, 2006.
\bibitem{guy2013unsolved}
\emph{Guy} \emph{R.} Unsolved problems in number theory. Т.
1. — Springer Science \& Business Media, 2013.
\bibitem{noll1989nclusters}
\emph{Noll} \emph{L. C.}, \emph{Bell} \emph{D. I.} n-clusters
for $1 < n < 7$ // Mathematics of Computation. — 1989. — Т. 53,
No 187. — С. 439—444.
\bibitem{kurz2013constructing}
Constructing 7-clusters / S. Kurz [и др.] // arXiv preprint
arXiv:1312.2318. — 2013.
\bibitem{graham1974there}
\emph{Graham} \emph{R.}, \emph{Rothschild} \emph{B.},
\emph{Straus} \emph{E.} Are there $n + 2$ points in $E^n$ with odd
integral distances? // The American Mathematical Monthly. —
1974. — Т. 81, No 1. — С. 21—25.
\bibitem{ulam1960collection}
\emph{Ulam} \emph{S. M.} A collection of mathematical
problems. Т. 8. — Interscience Publishers, 1960.
\bibitem{garibaldi2005erdos}
\emph{Garibaldi} \emph{J.}, \emph{Iosevich} \emph{A.} The
Erdos distance problem: lecture notes. — 2005.
\bibitem{garibaldi2011erdos}
\emph{Garibaldi} \emph{J.}, \emph{Iosevich} \emph{A.},
\emph{Senger} \emph{S.} The Erdös Distance Problem. Т. 56. —
American Mathematical Soc., 2011.\bibitem{erdos1946sets}
\emph{Erdös} \emph{P.} On sets of distances of n points //
The American Mathematical Monthly. — 1946. — Т. 53, No 5. —
С. 248—250.
\bibitem{guth2015erdos}
\emph{Guth} \emph{L.}, \emph{Katz} \emph{N. H.} On the
Erdos distinct distances problem in the plane // Annals of
Mathematics. — 2015. — Т. 181. — С. 155—190.
\end{thebibliography}

### 2018-09-23 vmmsh-2018/blg.tex
[vmmsh] Bibiliography- WIP

@@
 \bibitem{erdos1945integral}
-\emph{Erdös} \emph{P.} Integral distances // Bulletin of the
+\textsl{Erdös} \textsl{P.} Integral distances // Bulletin of the
 American Mathematical Society. — 1945. — Т. 51, No 12. — С. 996.
 \bibitem{anning1945integral}
-\emph{Anning} \emph{N. H.}, \emph{Erdös} \emph{P.}
+\textsl{Anning} \textsl{N. H.}, \textsl{Erdös} \textsl{P.}
 Integral distances // Bulletin of the American Mathematical
@@
 \bibitem{solymosi2003note}
-\emph{Solymosi} \emph{J.} Note on integral distances //
+\textsl{Solymosi} \textsl{J.} Note on integral distances //
 Discrete \& Computational Geometry. — 2003. — Т. 30, No 2. —
@@
 \bibitem{anning1915discussions}
-\emph{Anning} \emph{N.} Discussions: Relating to a Geometric
+\textsl{Anning} \textsl{N.} Discussions: Relating to a Geometric
 Representation of Integral Solutions of Certain Quadratic
@@
 \bibitem{harborth1993upper}
-\emph{Harborth} \emph{H.},
-\emph{Möller} \emph{M.} An
+\textsl{Harborth} \textsl{H.},
+\textsl{Möller} \textsl{M.} An
 diameter of integral point sets
 Geometry. — 1993. — Т. 9, No 4.
-\emph{Kemnitz} \emph{A.},
+\textsl{Kemnitz} \textsl{A.},
 upper bound for the minimum
@@
 — С. 427—432.\bibitem{piepmeyer1996maximum}
-\emph{Piepmeyer} \emph{L.} The maximum number of odd
+\textsl{Piepmeyer} \textsl{L.} The maximum number of odd
 integral distances between points in the plane // Discrete \&
@@
 \bibitem{kurz2008bounds}
-\emph{Kurz} \emph{S.}, \emph{Laue} \emph{R.} Bounds for
+\textsl{Kurz} \textsl{S.}, \textsl{Laue} \textsl{R.} Bounds for
 the minimum diameter of integral point sets // arXiv preprint
@@
 \bibitem{our-vvmsh-2018}
-\emph{Авдеев} \emph{Н. Н.} Некоторые примеры систем
+\textsl{Авдеев} \textsl{Н. Н.} Некоторые примеры систем
 Эрдёша // Современные методы теории краевых задач :
@@
 \bibitem{locatelli2002packing}
-\emph{Locatelli} \emph{M.}, \emph{Raber} \emph{U.} Packing
+\textsl{Locatelli} \textsl{M.}, \textsl{Raber} \textsl{U.} Packing
 equal circles in a square: a deterministic global optimization
@@
 \bibitem{costa2013valid}
-\emph{Costa} \emph{A.} Valid constraints for the point packing
+\textsl{Costa} \textsl{A.} Valid constraints for the point packing
 in a square problem // Discrete Applied Mathematics. — 2013. —
@@
 \bibitem{markot2005newverified}
-\emph{Markót} \emph{M. C.}, \emph{Csendes} \emph{T.} A
+\textsl{Markót} \textsl{M. C.}, \textsl{Csendes} \textsl{T.} A
 new verified optimization technique for the "packing circles in
@@
 \bibitem{degroot1990optimal10points}
-\emph{DeGroot} \emph{C.}, \emph{Peikert} \emph{R.},
-\emph{Würtz} \emph{D.} The optimal packing of ten equal
+\textsl{DeGroot} \textsl{C.}, \textsl{Peikert} \textsl{R.},
+\textsl{Würtz} \textsl{D.} The optimal packing of ten equal
 circles in a square. — Eidgenössische Technische Hochschule
@@
 1990.\bibitem{kemnitz1988punktmengen}
-\emph{Kemnitz} \emph{A.} Punktmengen mit ganzzahligen
+\textsl{Kemnitz} \textsl{A.} Punktmengen mit ganzzahligen
 Abständen. — 1988.
 \bibitem{Bukhstab-number-theory}
-\emph{Бухштаб} \emph{А.} Теория чисел. — 1966. — С. 384.
+\textsl{Бухштаб} \textsl{А.} Теория чисел. — 1966. — С. 384.
 \bibitem{our-mkmitu-2016}
-\emph{Авдеев} \emph{Н. Н.} Об отыскании множеств точек на
+\textsl{Авдеев} \textsl{Н. Н.} Об отыскании множеств точек на
 плоскости с целочисленными расстояниями // Математическое
@@
 \bibitem{our-ped-2017}
-\emph{Авдеев} \emph{Н. Н.} О множествах точек на
+\textsl{Авдеев} \textsl{Н. Н.} О множествах точек на
 плоскости с целочисленными расстояниями и алгоритмах их
@@
 \bibitem{polygons-on-lattices}
-\emph{Вавилов} \emph{В.}, \emph{Устинов}
+\textsl{Вавилов} \textsl{В.}, \textsl{Устинов}
 Многоугольники на решетках. — 2006.
-\emph{А.}
+\textsl{А.}
 \bibitem{Kohnert2006IntegralPS}
-\emph{Kohnert} \emph{A.}, \emph{Kurz} \emph{S.} Integral
+\textsl{Kohnert} \textsl{A.}, \textsl{Kurz} \textsl{S.} Integral
 point sets over $Z^m_n$
@@
 \bibitem{kurz2005characteristic}
-\emph{Kurz} \emph{S.} On the characteristic of integral point
+\textsl{Kurz} \textsl{S.} On the characteristic of integral point
 sets in // arXiv preprint math/0511704. — 2005.
 \bibitem{harborth1985diameters}
-\emph{Harborth} \emph{H.}, \emph{Kemnitz} \emph{A.}
+\textsl{Harborth} \textsl{H.}, \textsl{Kemnitz} \textsl{A.}
 Diameters of integral point sets // Colloquia Mathematica
@@
 \bibitem{kanold1981punktmengen}
-\emph{Kanold}
-\emph{H.-J.}
+\textsl{Kanold}
+\textsl{H.-J.}
 Über
@@
 \bibitem{nozaki2013lower}
-\emph{Nozaki} \emph{H.} Lower bounds for the minimum
+\textsl{Nozaki} \textsl{H.} Lower bounds for the minimum
 diameter of integral point sets // Australasian Journal of
@@
 \bibitem{piepmeyer1988raumliche}
-\emph{Piepmeyer}
-\emph{L.}
+\textsl{Piepmeyer}
+\textsl{L.}
 Räumliche
@@
 \bibitem{harborth1991point}
-\emph{Harborth} \emph{H.}, \emph{Piepmeyer} \emph{L.}
+\textsl{Harborth} \textsl{H.}, \textsl{Piepmeyer} \textsl{L.}
 Point sets with small integral distances // Applied geometry and
@@
 \bibitem{harborth1998integral}
-\emph{Harborth} \emph{H.} Integral distances in point sets //
+\textsl{Harborth} \textsl{H.} Integral distances in point sets //
 Karl der Grosse und sein Nachwirken. 1200 Jahre Kultur und
@@
 \bibitem{kurz2006konstruktion}
-\emph{Kurz} \emph{S.} Konstruktion und Eigenschaften
+\textsl{Kurz} \textsl{S.} Konstruktion und Eigenschaften
 ganzzahliger Punktmengen: дис. . . . канд. / Kurz Sascha. — 2006.
 \bibitem{kurz2008minimum}
-\emph{Kurz} \emph{S.}, \emph{Wassermann} \emph{A.} On
+\textsl{Kurz} \textsl{S.}, \textsl{Wassermann} \textsl{A.} On
 the minimum diameter of plane integral point sets // arXiv
@@
 \bibitem{our-mz-rus}
-\emph{Авдеев} \emph{Н. Н.}, \emph{Семёнов} \emph{Е. М.}
+\textsl{Авдеев} \textsl{Н. Н.}, \textsl{Семёнов} \textsl{Е. М.}
 О множествах точек на плоскости с целочисленными
@@
 No 5. — С. 757—761.\bibitem{kreisel2008heptagon}
-\emph{Kreisel} \emph{T.}, \emph{Kurz} \emph{S.} There are
+\textsl{Kreisel} \textsl{T.}, \textsl{Kurz} \textsl{S.} There are
 integral heptagons, no three points on a line, no four on a circle //
@@
 \bibitem{brass2006research}
-\emph{Brass} \emph{P.}, \emph{Moser} \emph{W. O.},
-\emph{Pach} \emph{J.} Research problems in discrete
+\textsl{Brass} \textsl{P.}, \textsl{Moser} \textsl{W. O.},
+\textsl{Pach} \textsl{J.} Research problems in discrete
 geometry. — Springer Science \& Business Media, 2006.
 \bibitem{guy2013unsolved}
-\emph{Guy} \emph{R.} Unsolved problems in number theory. Т.
+\textsl{Guy} \textsl{R.} Unsolved problems in number theory. Т.
 1. — Springer Science \& Business Media, 2013.
 \bibitem{noll1989nclusters}
-\emph{Noll} \emph{L. C.}, \emph{Bell} \emph{D. I.} n-clusters
+\textsl{Noll} \textsl{L. C.}, \textsl{Bell} \textsl{D. I.} n-clusters
 for $1 < n < 7$ // Mathematics of Computation. — 1989. — Т. 53,
@@
 \bibitem{graham1974there}
-\emph{Graham} \emph{R.}, \emph{Rothschild} \emph{B.},
-\emph{Straus} \emph{E.} Are there $n + 2$ points in $E^n$ with odd
+\textsl{Graham} \textsl{R.}, \textsl{Rothschild} \textsl{B.},
+\textsl{Straus} \textsl{E.} Are there $n + 2$ points in $E^n$ with odd
 integral distances? // The American Mathematical Monthly. —
@@
 \bibitem{ulam1960collection}
-\emph{Ulam} \emph{S. M.} A collection of mathematical
+\textsl{Ulam} \textsl{S. M.} A collection of mathematical
 problems. Т. 8. — Interscience Publishers, 1960.
 \bibitem{garibaldi2005erdos}
-\emph{Garibaldi} \emph{J.}, \emph{Iosevich} \emph{A.} The
+\textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.} The
 Erdos distance problem: lecture notes. — 2005.
 \bibitem{garibaldi2011erdos}
-\emph{Garibaldi} \emph{J.}, \emph{Iosevich} \emph{A.},
-\emph{Senger} \emph{S.} The Erdös Distance Problem. Т. 56. —
+\textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.},
+\textsl{Senger} \textsl{S.} The Erdös Distance Problem. Т. 56. —
 American Mathematical Soc., 2011.\bibitem{erdos1946sets}
-\emph{Erdös} \emph{P.} On sets of distances of n points //
+\textsl{Erdös} \textsl{P.} On sets of distances of n points //
 The American Mathematical Monthly. — 1946. — Т. 53, No 5. —
@@
 \bibitem{guth2015erdos}
-\emph{Guth} \emph{L.}, \emph{Katz} \emph{N. H.} On the
+\textsl{Guth} \textsl{L.}, \textsl{Katz} \textsl{N. H.} On the
 Erdos distinct distances problem in the plane // Annals of

### 2018-09-23 vmmsh-2018/blg.tex
[vmmsh] Bibiliography - rather working

@@
 \bibitem{erdos1945integral}
-\textsl{Erdös} \textsl{P.} Integral distances // Bulletin of the
+\textsl{Erdös} \textsl{P.} Integral distances /\!/ Bulletin of the
 American Mathematical Society. — 1945. — Т. 51, No 12. — С. 996.
@@
 \textsl{Anning} \textsl{N. H.}, \textsl{Erdös} \textsl{P.}
-Integral distances // Bulletin of the American Mathematical
+Integral distances /\!/ Bulletin of the American Mathematical
 Society. — 1945. — Т. 51, No 8. — С. 598—600.
 \bibitem{solymosi2003note}
-\textsl{Solymosi} \textsl{J.} Note on integral distances //
+\textsl{Solymosi} \textsl{J.} Note on integral distances /\!/
 Discrete \& Computational Geometry. — 2003. — Т. 30, No 2. —
@@
 Representation of Integral Solutions of Certain Quadratic
-Equations // The American Mathematical Monthly. — 1915. —
+Equations /\!/ The American Mathematical Monthly. — 1915. —
 Т. 22, No 9. — С. 321—321.
@@
 upper bound for the minimum
-// Discrete \& Computational
+/\!/ Discrete \& Computational Geometry
 — С. 427—432.\bibitem{piepmeyer1996maximum}
 \textsl{Piepmeyer} \textsl{L.} The maximum number of odd
-integral distances between points in the plane // Discrete \&
+integral distances between points in the plane /\!/ Discrete \&
 Computational Geometry. — 1996. — Т. 16, No 1. — С. 113—115.
@@
 \textsl{Kurz} \textsl{S.}, \textsl{Laue} \textsl{R.} Bounds for
-the minimum diameter of integral point sets // arXiv preprint
+the minimum diameter of integral point sets /\!/ arXiv preprint
 arXiv:0804.1296. — 2008.
@@
 \textsl{Авдеев} \textsl{Н. Н.} Некоторые примеры систем
-Эрдёша // Современные методы теории краевых задач :
+Эрдёша /\!/ Современные методы теории краевых задач :
 материалы международной конференции «Понтрягинские
@@
 equal circles in a square: a deterministic global optimization
-approach // Discrete Applied Mathematics. — 2002. — Т. 122,
+approach /\!/ Discrete Applied Mathematics. — 2002. — Т. 122,
 No 1—3. — С. 139—166.
@@
 \textsl{Costa} \textsl{A.} Valid constraints for the point packing
-in a square problem // Discrete Applied Mathematics. — 2013. —
+in a square problem /\!/ Discrete Applied Mathematics. — 2013. —
 Т. 161, No 18. — С. 2901—2909.
@@
 new verified optimization technique for the "packing circles in
-a unit square" problems // SIAM Journal on Optimization. —
+a unit square" problems /\!/ SIAM Journal on Optimization. —
 2005. — Т. 16, No 1. — С. 193—219.
@@
 \textsl{Авдеев} \textsl{Н. Н.} Об отыскании множеств точек на
-плоскости с целочисленными расстояниями // Математическое
+плоскости с целочисленными расстояниями /\!/ Математическое
 и компьютерное моделирование, информационные технологии
@@
 плоскости с целочисленными расстояниями и алгоритмах их
-отыскания // Некоторые вопросы анализа, алгебры, геометрии
+отыскания /\!/ Некоторые вопросы анализа, алгебры, геометрии
 и математического образования: материалы международной
@@
 point sets over $Z^m_n$
-// Discrete Applied Mathematics. — 2006. —
+/\!/ Discrete Applied Mathematics. — 2006. —
 Т. 157 (2009). — С. 2105—2117.
@@
 \textsl{Kurz} \textsl{S.} On the characteristic of integral point
-sets in // arXiv preprint math/0511704. — 2005.
+sets in /\!/ arXiv preprint math/0511704. — 2005.
 \bibitem{harborth1985diameters}
 \textsl{Harborth} \textsl{H.}, \textsl{Kemnitz} \textsl{A.}
-Diameters of integral point sets // Colloquia Mathematica
+Diameters of integral point sets /\!/ Colloquia Mathematica
 Societatis Jano Bolyai. Т. 48. — 1985.
@@
 im
-$k$-dimensionalen euklidischen Raum // Abh. Braunschw. Wiss.
+$k$-dimensionalen euklidischen Raum /\!/ Abh. Braunschw. Wiss.
 Ges. — 1981. — Т. 32, No 55—65. — С. 252.
@@
 \textsl{Nozaki} \textsl{H.} Lower bounds for the minimum
-diameter of integral point sets // Australasian Journal of
+diameter of integral point sets /\!/ Australasian Journal of
 Combinatorics. — 2013. — Т. 56. — С. 139—143.
@@
 \textsl{Harborth} \textsl{H.}, \textsl{Piepmeyer} \textsl{L.}
-Point sets with small integral distances // Applied geometry and
+Point sets with small integral distances /\!/ Applied geometry and
 discrete mathematics: the Victor Klee Festschrift. Т. 4. — American
@@
 \bibitem{harborth1998integral}
-\textsl{Harborth} \textsl{H.} Integral distances in point sets //
+\textsl{Harborth} \textsl{H.} Integral distances in point sets /\!/
 Karl der Grosse und sein Nachwirken. 1200 Jahre Kultur und
@@
 \textsl{Kurz} \textsl{S.}, \textsl{Wassermann} \textsl{A.} On
-the minimum diameter of plane integral point sets // arXiv
+the minimum diameter of plane integral point sets /\!/ arXiv
 preprint arXiv:0804.1307. — 2008.
@@
 О множествах точек на плоскости с целочисленными
-расстояниями // Математические заметки. — 2016. — Т. 100,
+расстояниями /\!/ Математические заметки. — 2016. — Т. 100,
 No 5. — С. 757—761.\bibitem{kreisel2008heptagon}
 \textsl{Kreisel} \textsl{T.}, \textsl{Kurz} \textsl{S.} There are
-integral heptagons, no three points on a line, no four on a circle //
+integral heptagons, no three points on a line, no four on a circle /\!/
 Discrete \& Computational Geometry. — 2008. — Т. 39, No 4. —
@@
 \textsl{Noll} \textsl{L. C.}, \textsl{Bell} \textsl{D. I.} n-clusters
-for $1 < n < 7$ // Mathematics of Computation. — 1989. — Т. 53,
+for $1 < n < 7$ /\!/ Mathematics of Computation. — 1989. — Т. 53,
 No 187. — С. 439—444.
 \bibitem{kurz2013constructing}
-Constructing 7-clusters / S. Kurz [и др.] // arXiv preprint
+Constructing 7-clusters / S. Kurz [и др.] /\!/ arXiv preprint
 arXiv:1312.2318. — 2013.
@@
 \textsl{Straus} \textsl{E.} Are there $n + 2$ points in $E^n$ with odd
-integral distances? // The American Mathematical Monthly. —
+integral distances? /\!/ The American Mathematical Monthly. —
 1974. — Т. 81, No 1. — С. 21—25.
@@
 American Mathematical Soc., 2011.\bibitem{erdos1946sets}
-\textsl{Erdös} \textsl{P.} On sets of distances of n points //
+\textsl{Erdös} \textsl{P.} On sets of distances of n points /\!/
 The American Mathematical Monthly. — 1946. — Т. 53, No 5. —
@@
 \textsl{Guth} \textsl{L.}, \textsl{Katz} \textsl{N. H.} On the
-Erdos distinct distances problem in the plane // Annals of
+Erdos distinct distances problem in the plane /\!/ Annals of
 Mathematics. — 2015. — Т. 181. — С. 155—190.

### 2018-09-23 vmmsh-2018/style09/Primer.utf8.tex
[vmmsh-style] UTF-8 versions
[vmmsh-style] UTF-8 versions - full
[vmmsh-style] overloading amsthm proof environment
[vmmsh-style] resetting the counters when \Title is called -- example
[vmmsh-style] Bibliography with examples

--- 2018-09-18 vmmsh-2018/style09/Primer.tex
@@
-\input style_konf.tex
-
-\setcounter{footnote}{0}
- \setcounter{equation}{0}
+\input style_konf.utf8.tex
 
@@
 
+\begin{proof}
+Очевидно.
+\end{proof}
+
+\begin{equation}
+	\label{eq:main}
+	\sin^2 t + \cos^2 t = 1
+\end{equation}
+
+\begin{equation}
+	\label{eq:aux}
+	\sin^2 t = 1 - \cos^2 t
+\end{equation}
+
+Уравнение \eqref{eq:main} написано раньше уравнения (\ref{eq:aux}).
+
 \section*{Литература}
@@
  \bib{Киприянов И.~А., Катрахов В.~В., Ляпин В.~М.} {О краевых задачах в области общего вида для сингулярных
- параболических систем уравнений~/\!/ Докл. АН СССР.\,---\,1976.\,---\,Т.~230, \No~6.\,---\,С.~1271--1274.}
+ параболических систем уравнений~/\!/ Докл. АН СССР.\,---\,1976.\,---\,Т.~230, No~6.\,---\,С.~1271--1274.}
 
@@
 
+\Title{НАЗВАНИЕ СТАТЬИ}
+
+ \Author{ФИО}
+
+\Abstract{Текст аннотации. \par}
+
+{\hangindent17pt\hangafter=0\noindent\footnotesize{\bf Ключевые
+ слова:} ключевое слово 1, ключевое слово 2, \ldots\par}
+
+\section*{1. Название параграфа}
+
+\Teorema{ 1}Текст теоремы. \Endproc
+
+\beginproof~Текст доказательства.~\endproof
+
+\Lema{ 1}Текст леммы.\Endproc
+
+\begin{proof}
+Очевидно.
+\end{proof}
+
+\begin{equation}
+	\label{eq:aux}
+	\sin^2 t = 1 - \cos^2 t
+\end{equation}
+
+\begin{equation}
+	\label{eq:main}
+	\sin^2 t + \cos^2 t = 1
+\end{equation}
+
+Уравнение \eqref{eq:aux} написано раньше уравнения \eqref{eq:main}.
+
+Ссылки на работы \cite{Polyanin2001,Kipryanov1976} и \cite{Ilin2005}
+
+\begin{thebibliography}{99}
+  \bibitem{Polyanin2001}
+    \textsl{Полянин А.~Д.}
+	Справочник по линейным уравнениям математической фи\-зи\-ки.\,---\,М.: Физматлит, 2001.\,---\,576~c.
+
+  \bibitem{Kipryanov1976}
+    \textsl{Киприянов И.~А., Катрахов В.~В., Ляпин В.~М.}
+    О краевых задачах в области общего вида для сингулярных
+    параболических систем уравнений~/\!/ Докл. АН СССР.\,---\,1976.\,---\,Т.~230, No~6.\,---\,С.~1271--1274.
+
+  \bibitem{Ilin2005}
+    \textsl{Ильин В.~А., Позняк Э.~Г.}
+    Основы математического анализа.  Ч.~1: Учеб. для вузов~/ 7-е изд.\,---\,М.:
+    Физматлит, 2005.\,---\,648~с.
+
+\end{thebibliography}
+
 \end{document}

### 2018-09-23 vmmsh-2018/AvdeevNN_09.cp1251.tex
[vmmsh] Recoding to 1251 - WIP

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Тезисы доклада на XIV «Владикавказская молодежная математическая школа»
%%  Россия, РСО-А, с. Цей, 16-21 июля 2018 г.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Загрузка файла style.tex со служебными командами
\input style09/style_konf.tex

\mag 1468

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Пожалуйста, не меняйте порядок команд!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Название доклада

\UDC{511.95 + 514.112}

\Title{МНОЖЕСТВА ТОЧЕК С ЦЕЛОЧИСЛЕННЫМИ РАССТОЯНИЯМИ НА ПЛОСКОСТИ И В ЕВКЛИДОВОМ ПРОСТРАНСТВЕ} % обязательное поле!

 \Author{ФИО}

\Abstract{Текст аннотации. \par}

{\hangindent17pt\hangafter=0\noindent\footnotesize{\bf Ключевые
 слова:} ключевое слово 1, ключевое слово 2, \ldots\par}

\input{IntroBySemenov.cp1251.tex}

\input{BetterLinearBound.cp1251.tex}

\input{ReviewChar.cp1251.tex}

\input{ReviewMultidim.cp1251.tex}

{\small Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-10125.}

\input{blg.cp1251.tex}

\end{document}

### 2018-09-23 vmmsh-2018/IntroBySemenov.cp1251.tex
[vmmsh] Recoding to 1251 - WIP

Oбозначим через $\mathfrak{M}$ множество таких подмножеств
$
	M = \{M_1, M_2, ... \}\subset \mathbb{R}^2
$,
что
$
	|M_i - M_j|\in\mathbb{N}
$
для всех $i$, $j$, где $\mathbb{N}$~--- множество натуральных чисел
и $|M_i - M_j|$~--- расстояние между точками $M_i$ и $M_j$.
П. Эрдёш доказал~\cite{erdos1945integral},
что справедлива следующая

\begin{theorem}
	\label{thmErdos}
	Всякое бесконечное подмножество $M\in\mathfrak{M}$
	содержится в некоторой прямой $L\in\mathbb{R}^2$.
\end{theorem}

Теорему~\ref{thmErdos} можно доказать, основываясь на двух элементарных леммах.
Интересно, что работа~\cite{erdos1945integral} была опубликована через несколько
месяцев после совместной работы Эннинга и Эрдёша~\cite{anning1945integral},
в которой было дано другое, более сложное доказательство.

\begin{lemma}
	\label{thmQuadCurves}
	Пусть $A$ и $B$~--- две кривые второго порядка на плоскости и $A$ не является парой прямых.
	Тогда $A\cap B$ содержит не более 4 точек.
\end{lemma}

\begin{proof}
	Без ограничения общности кривая $A$ определяется уравнением
	\begin{equation*}
		x^2 = ay^2 + by +c
		,
	\end{equation*}
	а кривая $B$ определяется уравнением второго порядка общего вида.
	Система этих двух уравнений легко сводится к уравнению с одной неизвестной 4-го порядка.
	Если это уравнение не является тривиальным, то оно имеет не более 4 корней.
\end{proof}

Если $A$ и $B$~--- пары прямых $xy=0$ и $xy=x$,
то $A\cap B$ содержит ось игреков.
Поэтому предположение в лемме~\ref{thmQuadCurves} о том,
что одна из кривых не является парой прямых, существенно.

\begin{lemma}
	\label{thm:quadCurveFamily}
	Пусть $M_1, M_2 \in \mathbb{R}^2$ и
	$|M_1 - M_2| = n \in \mathbb{N}$.
	Тогда
	\begin{equation}
		\label{eq:quadCurveFamily}
		\left\{
			M: |M-M_1|\in\mathbb{N}, |M-M_2|\in\mathbb{N}
		\right\}
		=\bigcup_{i=1}^n Q_i
		,
	\end{equation}
	где $Q_1$ есть пара взаимно перпендикулярных прямых (крест)
	и $Q_2, Q_3,...,Q_n$~--- гиперболы.
\end{lemma}

\begin{proof}
	По неравенству треугольника
	\begin{equation*}
		\bigl| |M-M_1| - |M-M_2| \bigr|
		\leqslant
		|M_1 - M_2| = n
		.
	\end{equation*}
	Так как $\bigl| |M-M_1| - |M-M_2| \bigr|$
	есть целое неотрицательное число, то
	\begin{equation*}
		\bigl| |M-M_1| - |M-M_2| \bigr|
		\in\{0,1,2,...,n\}
		.
	\end{equation*}
	Ясно, что множество
	\begin{equation}
		\label{eq:centrPerp}
		\left\{ M: \bigl| |M-M_1| - |M-M_2| \bigr| = 0 \right\}
	\end{equation}
	принадлежит прямой, проходящей через середину отрезка
	$[M_1, M_2]$ перпендикулярно этому отрезку,
	\begin{equation}
		\label{eq:tails}
		\left\{ M: \bigl| |M-M_1| - |M-M_2| \bigr| = n \right\}
	\end{equation}
	содержится в прямой, проходящей через точки $M_1, M_2$,
	\begin{equation}
		\label{eq:hyperbola}
		\left\{ M: \bigl| |M-M_1| - |M-M_2| \bigr| = k \right\}
	\end{equation}
	есть гипербола для любого $k=1,2,...,n-1$.
	Обозначая объединение множеств \eqref{eq:centrPerp} и \eqref {eq:tails}
	через $Q_1$,
	а множества \eqref{eq:hyperbola} через $Q_i$,
	мы получаем требуемое представление.
	Таким образом, \eqref{eq:quadCurveFamily}
	есть объединение $n$ кривых второго порядка.
\end{proof}

\begin{proof}[Доказательство теоремы \ref{thmErdos}]
	Предположим противное,
	т.е. $\{M_1, M_2, ... \}\in\mathfrak{M}$
	и точки $M_1$, $M_2$, $M_3$ не лежат на одной прямой.
	Применяя лемму~\ref{thm:quadCurveFamily}
	к паре точек $M_1, M_2$ и к паре точек $M_1, M_3$,
	найдём такие кривые 2-го порядка
	$Q_1, Q_2,...,.Q_n$ и $R_1, R_2,...,.R_m$, что
	\begin{equation*}
		\left\{
			M: |M-M_1|\in\mathbb{N}, |M-M_2|\in\mathbb{N}
		\right\}
		=\bigcup_{i=1}^n Q_i
		,
	\end{equation*}
	\begin{equation*}
		\left\{
			M: |M-M_1|\in\mathbb{N}, |M-M_3|\in\mathbb{N}
		\right\}
		=\bigcup_{j=1}^m R_j
		.
	\end{equation*}
	Отсюда
	\begin{multline*}
		\left\{
			M:
			|M-M_1|\in\mathbb{N},
			|M-M_2|\in\mathbb{N},
			|M-M_3|\in\mathbb{N}
		\right\}
		\subset
		\\ \subset
		\left(\bigcup_{i=1}^n Q_i\right)
		\cap
		\left(\bigcup_{j=1}^m R_j\right)
		=
		\bigcup_{i=1}^n \bigcup_{j=1}^m \left( Q_i \cap R_i \right)
		.
	\end{multline*}
	Из предположения о том, что $M_1$, $M_2$, $M_3$
	не лежат на одной прямой, вытекает,
	что $Q_1\cap R_1$ не содержит прямую.
	Поэтому $|Q_1\cap R_1|\leqslant 4$,
	и, следовательно, в силу леммы~\ref{thmQuadCurves}
	$|Q_i\cap R_j|\leqslant 4$
	для всех $i=1,2,...,n$, $j=1,2,...,m$.
	Отсюда
	\begin{equation*}
		|M| \leqslant 4nm
		,
	\end{equation*}
	т.е. $M$ конечно,
	что противоречит предположению о бесконечности $M$.
	Полученное проиворечие доказывает, что все точки $M$
	принадлежат некоторой прямой $L$.
	В качестве $L$ можно взять прямую,
	проходящую через пару любых точек из $M$.
\end{proof}

Предположение о бесконечности $M$ в теореме~\ref{thmErdos} существенно.
Для заданного $n\in\mathbb{N}$ обозначим через $\mathfrak{M}_n$
множество таких $M\in\mathfrak{M}$, что
$|M|=n$ и $M \not\subset L$ для любой прямой $L \subset\mathbb{R}^2$.

\begin{theorem}
	\label{thm:power_exist}
	Для любого $n\in\mathbb{N}$ выполнено $\mathfrak{M}_n\neq\varnothing$.
\end{theorem}

\begin{proof}
	Из теоремы Пифагора следует, что треугольник со сторонами
	$2k+1$, $2k^2+2k$, $2k^2+2k+1$
	является прямоугольным для любого $k\in\mathbb{N}$.
	Выпишем стороны первых трёх треугольников,
	а затем стороны подобных им треугольников с совпадающей первой стороной:
	\begin{equation}
		\label{eq:PifagTriangles}
		\begin{array}{ll}
			3,  4,  5 & \quad 105, 140, 175 \\
			5, 12, 13 & \quad 105, 252, 273 \\
			7, 24, 25 & \quad 105, 360, 375.
		\end{array}
	\end{equation}
	Ясно, точки $(0,0)$, $(0,105)$, $(140,0)$, $(252,0)$, $(360,0)$
	образуют множество из $\mathfrak{M}_5$.
	Очевидно, если взять не 3, а $n$ точек, и затем найти подобные
	прямоугольные треугольники с совпадающей первой стороной,
	то можно построить множество из $\mathfrak{M}_{n+2}$.
\end{proof}

Другой пример множества из $\mathfrak{M}_{n}$ для заданного $n$
можно найти в доказательстве теоремы~\ref{thm:power_char_exist},
уверждение которой усиливает утверждение теоремы~\ref{thm:power_exist}.

### 2018-09-23 vmmsh-2018/blg.cp1251.tex
[vmmsh] Recoding to 1251 - WIP

--- 2018-09-23 vmmsh-2018/blg.tex
@@
 \bibitem{erdos1945integral}
-\textsl{Erdös} \textsl{P.} Integral distances /\!/ Bulletin of the
-American Mathematical Society. — 1945. — Т. 51, No 12. — С. 996.
+\textsl{Erd\"{o}s} \textsl{P.} Integral distances /\!/ Bulletin of the
+American Mathematical Society. --- 1945. --- Т. 51, No 12. --- С. 996.
 \bibitem{anning1945integral}
-\textsl{Anning} \textsl{N. H.}, \textsl{Erdös} \textsl{P.}
+\textsl{Anning} \textsl{N. H.}, \textsl{Erd\"{o}s} \textsl{P.}
 Integral distances /\!/ Bulletin of the American Mathematical
-Society. — 1945. — Т. 51, No 8. — С. 598—600.
+Society. --- 1945. --- Т. 51, No 8. --- С. 598---600.
 \bibitem{solymosi2003note}
 \textsl{Solymosi} \textsl{J.} Note on integral distances /\!/
-Discrete \& Computational Geometry. — 2003. — Т. 30, No 2. —
-С. 337—342.
+Discrete \& Computational Geometry. --- 2003. --- Т. 30, No 2. ---
+С. 337---342.
 \bibitem{anning1915discussions}
@@
 Representation of Integral Solutions of Certain Quadratic
-Equations /\!/ The American Mathematical Monthly. — 1915. —
-Т. 22, No 9. — С. 321—321.
+Equations /\!/ The American Mathematical Monthly. --- 1915. ---
+Т. 22, No 9. --- С. 321---321.
 \bibitem{harborth1993upper}
 \textsl{Harborth} \textsl{H.},
-\textsl{Möller} \textsl{M.} An
+\textsl{M\"{o}ller} \textsl{M.} An
 diameter of integral point sets
-Geometry. — 1993. — Т. 9, No 4.
+Geometry. --- 1993. --- Т. 9, No 4.
 \textsl{Kemnitz} \textsl{A.},
@@
 /\!/ Discrete \& Computational Geometry
-— С. 427—432.\bibitem{piepmeyer1996maximum}
+--- С. 427---432.\bibitem{piepmeyer1996maximum}
 \textsl{Piepmeyer} \textsl{L.} The maximum number of odd
 integral distances between points in the plane /\!/ Discrete \&
-Computational Geometry. — 1996. — Т. 16, No 1. — С. 113—115.
+Computational Geometry. --- 1996. --- Т. 16, No 1. --- С. 113---115.
 \bibitem{kurz2008bounds}
@@
 the minimum diameter of integral point sets /\!/ arXiv preprint
-arXiv:0804.1296. — 2008.
+arXiv:0804.1296. --- 2008.
 \bibitem{our-vvmsh-2018}
@@
 Эрдёша /\!/ Современные методы теории краевых задач :
-материалы международной конференции «Понтрягинские
-чтения — XXIX», посвященной 90-летию Владимира
-Александровича Ильина (2–6 мая 2018 г.) — 2018. — С. 30—31.
+материалы международной конференции <<Понтрягинские
+чтения --- XXIX>>, посвященной 90-летию Владимира
+Александровича Ильина (2--6 мая 2018 г.) --- 2018. --- С. 30---31.
 \bibitem{locatelli2002packing}
@@
 equal circles in a square: a deterministic global optimization
-approach /\!/ Discrete Applied Mathematics. — 2002. — Т. 122,
-No 1—3. — С. 139—166.
+approach /\!/ Discrete Applied Mathematics. --- 2002. --- Т. 122,
+No 1---3. --- С. 139---166.
 \bibitem{costa2013valid}
 \textsl{Costa} \textsl{A.} Valid constraints for the point packing
-in a square problem /\!/ Discrete Applied Mathematics. — 2013. —
-Т. 161, No 18. — С. 2901—2909.
+in a square problem /\!/ Discrete Applied Mathematics. --- 2013. ---
+Т. 161, No 18. --- С. 2901---2909.
 \bibitem{markot2005newverified}
-\textsl{Markót} \textsl{M. C.}, \textsl{Csendes} \textsl{T.} A
+\textsl{Mark\'{o}t} \textsl{M. C.}, \textsl{Csendes} \textsl{T.} A
 new verified optimization technique for the "packing circles in
-a unit square" problems /\!/ SIAM Journal on Optimization. —
-2005. — Т. 16, No 1. — С. 193—219.
+a unit square" problems /\!/ SIAM Journal on Optimization. ---
+2005. --- Т. 16, No 1. --- С. 193---219.
 \bibitem{degroot1990optimal10points}
 \textsl{DeGroot} \textsl{C.}, \textsl{Peikert} \textsl{R.},
-\textsl{Würtz} \textsl{D.} The optimal packing of ten equal
-circles in a square. — Eidgenössische Technische Hochschule
-Zürich, Interdisziplinäres Projektzentrum für Supercomputing,
+\textsl{W\"{u}rtz} \textsl{D.} The optimal packing of ten equal
+circles in a square. --- Eidgen\"{o}ssische Technische Hochschule
+Z\"{u}rich, Interdisziplin\"{a}res Projektzentrum f\"{u}r Supercomputing,
 1990.\bibitem{kemnitz1988punktmengen}
 \textsl{Kemnitz} \textsl{A.} Punktmengen mit ganzzahligen
-Abständen. — 1988.
+Abst\"{a}nden. --- 1988.
 \bibitem{Bukhstab-number-theory}
-\textsl{Бухштаб} \textsl{А.} Теория чисел. — 1966. — С. 384.
+\textsl{Бухштаб} \textsl{А.} Теория чисел. --- 1966. --- С. 384.
 \bibitem{our-mkmitu-2016}
@@
 управления: сб. тр. Школы для студентов, аспирантов и
-молодых ученых «МКМИТУ-2016». — 2016. — С. 15—19.
+молодых ученых <<МКМИТУ-2016>>. --- 2016. --- С. 15---19.
 \bibitem{our-ped-2017}
@@
 и математического образования: материалы международной
-молодежной научной школы «Актуальные направления
-математического анализа и смежные вопросы». — 2017. — Т. 7,
-No I. — С. 9—10.
+молодежной научной школы <<Актуальные направления
+математического анализа и смежные вопросы>>. --- 2017. --- Т. 7,
+No I. --- С. 9---10.
 \bibitem{polygons-on-lattices}
 \textsl{Вавилов} \textsl{В.}, \textsl{Устинов}
-Многоугольники на решетках. — 2006.
+Многоугольники на решетках. --- 2006.
 \textsl{А.}
@@
 point sets over $Z^m_n$
-/\!/ Discrete Applied Mathematics. — 2006. —
-Т. 157 (2009). — С. 2105—2117.
+/\!/ Discrete Applied Mathematics. --- 2006. ---
+Т. 157 (2009). --- С. 2105---2117.
 \bibitem{kurz2005characteristic}
 \textsl{Kurz} \textsl{S.} On the characteristic of integral point
-sets in /\!/ arXiv preprint math/0511704. — 2005.
+sets in /\!/ arXiv preprint math/0511704. --- 2005.
 \bibitem{harborth1985diameters}
@@
 Diameters of integral point sets /\!/ Colloquia Mathematica
-Societatis Jano Bolyai. Т. 48. — 1985.
+Societatis Jano Bolyai. Т. 48. --- 1985.
 \bibitem{kanold1981punktmengen}
@@
 \textsl{H.-J.}
-Über
+\"{U}ber
 Punktmengen
@@
 $k$-dimensionalen euklidischen Raum /\!/ Abh. Braunschw. Wiss.
-Ges. — 1981. — Т. 32, No 55—65. — С. 252.
+Ges. --- 1981. --- Т. 32, No 55---65. --- С. 252.
 \bibitem{nozaki2013lower}
@@
 diameter of integral point sets /\!/ Australasian Journal of
-Combinatorics. — 2013. — Т. 56. — С. 139—143.
+Combinatorics. --- 2013. --- Т. 56. --- С. 139---143.
 \bibitem{piepmeyer1988raumliche}
@@
 \textsl{L.}
-Räumliche
+R\"{a}umliche
 ganzzahlige
-Punktmengen: дис. . . . канд. / Piepmeyer L. — Master’s
+Punktmengen: дис. . . . канд. / Piepmeyer L. --- Master’s
 thesis, TU Braunschweig, 1988.
@@
 Point sets with small integral distances /\!/ Applied geometry and
-discrete mathematics: the Victor Klee Festschrift. Т. 4. — American
+discrete mathematics: the Victor Klee Festschrift. Т. 4. --- American
 Mathematical Soc., 1991.
@@
 Karl der Grosse und sein Nachwirken. 1200 Jahre Kultur und
-Wissenschaft in Europa: Band II, Mathematisches Wissen. —
-1998. — С. 213—224.
+Wissenschaft in Europa: Band II, Mathematisches Wissen. ---
+1998. --- С. 213---224.
 \bibitem{kurz2006konstruktion}
 \textsl{Kurz} \textsl{S.} Konstruktion und Eigenschaften
-ganzzahliger Punktmengen: дис. . . . канд. / Kurz Sascha. — 2006.
+ganzzahliger Punktmengen: дис. . . . канд. / Kurz Sascha. --- 2006.
 \bibitem{kurz2008minimum}
@@
 the minimum diameter of plane integral point sets /\!/ arXiv
-preprint arXiv:0804.1307. — 2008.
+preprint arXiv:0804.1307. --- 2008.
 \bibitem{our-mz-rus}
@@
 О множествах точек на плоскости с целочисленными
-расстояниями /\!/ Математические заметки. — 2016. — Т. 100,
-No 5. — С. 757—761.\bibitem{kreisel2008heptagon}
+расстояниями /\!/ Математические заметки. --- 2016. --- Т. 100,
+No 5. --- С. 757---761.\bibitem{kreisel2008heptagon}
 \textsl{Kreisel} \textsl{T.}, \textsl{Kurz} \textsl{S.} There are
 integral heptagons, no three points on a line, no four on a circle /\!/
-Discrete \& Computational Geometry. — 2008. — Т. 39, No 4. —
-С. 786—790.
+Discrete \& Computational Geometry. --- 2008. --- Т. 39, No 4. ---
+С. 786---790.
 \bibitem{brass2006research}
@@
 \textsl{Pach} \textsl{J.} Research problems in discrete
-geometry. — Springer Science \& Business Media, 2006.
+geometry. --- Springer Science \& Business Media, 2006.
 \bibitem{guy2013unsolved}
 \textsl{Guy} \textsl{R.} Unsolved problems in number theory. Т.
-1. — Springer Science \& Business Media, 2013.
+1. --- Springer Science \& Business Media, 2013.
 \bibitem{noll1989nclusters}
 \textsl{Noll} \textsl{L. C.}, \textsl{Bell} \textsl{D. I.} n-clusters
-for $1 < n < 7$ /\!/ Mathematics of Computation. — 1989. — Т. 53,
-No 187. — С. 439—444.
+for $1 < n < 7$ /\!/ Mathematics of Computation. --- 1989. --- Т. 53,
+No 187. --- С. 439---444.
 \bibitem{kurz2013constructing}
 Constructing 7-clusters / S. Kurz [и др.] /\!/ arXiv preprint
-arXiv:1312.2318. — 2013.
+arXiv:1312.2318. --- 2013.
 \bibitem{graham1974there}
@@
 \textsl{Straus} \textsl{E.} Are there $n + 2$ points in $E^n$ with odd
-integral distances? /\!/ The American Mathematical Monthly. —
-1974. — Т. 81, No 1. — С. 21—25.
+integral distances? /\!/ The American Mathematical Monthly. ---
+1974. --- Т. 81, No 1. --- С. 21---25.
 \bibitem{ulam1960collection}
 \textsl{Ulam} \textsl{S. M.} A collection of mathematical
-problems. Т. 8. — Interscience Publishers, 1960.
+problems. Т. 8. --- Interscience Publishers, 1960.
 \bibitem{garibaldi2005erdos}
 \textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.} The
-Erdos distance problem: lecture notes. — 2005.
+Erdos distance problem: lecture notes. --- 2005.
 \bibitem{garibaldi2011erdos}
 \textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.},
-\textsl{Senger} \textsl{S.} The Erdös Distance Problem. Т. 56. —
+\textsl{Senger} \textsl{S.} The Erd\"{o}s Distance Problem. Т. 56. ---
 American Mathematical Soc., 2011.\bibitem{erdos1946sets}
-\textsl{Erdös} \textsl{P.} On sets of distances of n points /\!/
-The American Mathematical Monthly. — 1946. — Т. 53, No 5. —
-С. 248—250.
+\textsl{Erd\"{o}s} \textsl{P.} On sets of distances of n points /\!/
+The American Mathematical Monthly. --- 1946. --- Т. 53, No 5. ---
+С. 248---250.
 \bibitem{guth2015erdos}
@@
 Erdos distinct distances problem in the plane /\!/ Annals of
-Mathematics. — 2015. — Т. 181. — С. 155—190.
+Mathematics. --- 2015. --- Т. 181. --- С. 155---190.
 \end{thebibliography}

### 2018-09-23 vmmsh-2018/blg.cp1251.tex.tex
[vmmsh] Recoding to 1251 - WIP

\begin{thebibliography}{99}
\bibitem{erdos1945integral}
\textsl{Erd\"{o}s} \textsl{P.} Integral distances /\!/ Bulletin of the
American Mathematical Society. Ч 1945. Ч “. 51, No 12. Ч —. 996.
\bibitem{anning1945integral}
\textsl{Anning} \textsl{N. H.}, \textsl{Erd\"{o}s} \textsl{P.}
Integral distances /\!/ Bulletin of the American Mathematical
Society. Ч 1945. Ч “. 51, No 8. Ч —. 598Ч600.
\bibitem{solymosi2003note}
\textsl{Solymosi} \textsl{J.} Note on integral distances /\!/
Discrete \& Computational Geometry. Ч 2003. Ч “. 30, No 2. Ч
—. 337Ч342.
\bibitem{anning1915discussions}
\textsl{Anning} \textsl{N.} Discussions: Relating to a Geometric
Representation of Integral Solutions of Certain Quadratic
Equations /\!/ The American Mathematical Monthly. Ч 1915. Ч
“. 22, No 9. Ч —. 321Ч321.
\bibitem{harborth1993upper}
\textsl{Harborth} \textsl{H.},
\textsl{M\"{o}ller} \textsl{M.} An
diameter of integral point sets
Geometry. Ч 1993. Ч “. 9, No 4.
\textsl{Kemnitz} \textsl{A.},
upper bound for the minimum
/\!/ Discrete \& Computational Geometry
Ч —. 427Ч432.\bibitem{piepmeyer1996maximum}
\textsl{Piepmeyer} \textsl{L.} The maximum number of odd
integral distances between points in the plane /\!/ Discrete \&
Computational Geometry. Ч 1996. Ч “. 16, No 1. Ч —. 113Ч115.
\bibitem{kurz2008bounds}
\textsl{Kurz} \textsl{S.}, \textsl{Laue} \textsl{R.} Bounds for
the minimum diameter of integral point sets /\!/ arXiv preprint
arXiv:0804.1296. Ч 2008.
\bibitem{our-vvmsh-2018}
\textsl{јвдеев} \textsl{Ќ. Ќ.} Ќекоторые примеры систем
ЁрдЄша /\!/ —овременные методы теории краевых задач :
материалы международной конференции <<ѕонтр€гинские
чтени€ Ч XXIX>>, посв€щенной 90-летию ¬ладимира
јлександровича »льина (2Ц6 ма€ 2018 г.) Ч 2018. Ч —. 30Ч31.
\bibitem{locatelli2002packing}
\textsl{Locatelli} \textsl{M.}, \textsl{Raber} \textsl{U.} Packing
equal circles in a square: a deterministic global optimization
approach /\!/ Discrete Applied Mathematics. Ч 2002. Ч “. 122,
No 1Ч3. Ч —. 139Ч166.
\bibitem{costa2013valid}
\textsl{Costa} \textsl{A.} Valid constraints for the point packing
in a square problem /\!/ Discrete Applied Mathematics. Ч 2013. Ч
“. 161, No 18. Ч —. 2901Ч2909.
\bibitem{markot2005newverified}
\textsl{Mark

### 2018-09-23 vmmsh-2018/style09/Primer.tex
[vmmsh-style] \No temporary disabled
[vmmsh] Recoding to 1251 - WIP

@@
 \input style_konf.tex
-
-\setcounter{footnote}{0}
- \setcounter{equation}{0}
 
@@
 
+\Title{НАЗВАНИЕ СТАТЬИ}
+
+ \Author{ФИО}
+
+\Abstract{Текст аннотации. \par}
+
+{\hangindent17pt\hangafter=0\noindent\footnotesize{\bf Ключевые
+ слова:} ключевое слово 1, ключевое слово 2, \ldots\par}
+
+\section*{1. Название параграфа}
+
+\begin{aTeorema}
+	\label{thm1}
+	Текст теоремы.
+\end{aTeorema}
+
+\begin{proof}
+	Текст доказательства.
+\end{proof}
+
+\begin{aLema}
+	\label{lem2}
+	Текст леммы.
+\end{aLema}
+
+\begin{proof}
+	Очевидно.
+\end{proof}
+
+Из теоремы~\ref{thm1} и леммы~\ref{lem2} cледует
+\begin{aTeorema*}
+	Текст теоремы.
+\end{aTeorema*}
+
+\begin{equation}
+	\label{eq:aux}
+	\sin^2 t = 1 - \cos^2 t
+\end{equation}
+
+\begin{equation}
+	\label{eq:main}
+	\sin^2 t + \cos^2 t = 1
+\end{equation}
+
+Уравнение \eqref{eq:aux} написано раньше уравнения \eqref{eq:main}.
+
+Ссылки на работы \cite{Polyanin2001,Kipryanov1976} и \cite{Ilin2005}
+
+\begin{thebibliography}{99}
+  \bibitem{Polyanin2001}
+    \textsl{Полянин А.~Д.}
+	Справочник по линейным уравнениям математической фи\-зи\-ки.\,---\,М.: Физматлит, 2001.\,---\,576~c.
+
+  \bibitem{Kipryanov1976}
+    \textsl{Киприянов И.~А., Катрахов В.~В., Ляпин В.~М.}
+    О краевых задачах в области общего вида для сингулярных
+    параболических систем уравнений~/\!/ Докл. АН СССР.\,---\,1976.\,---\,Т.~230, No~6.\,---\,С.~1271--1274.
+
+  \bibitem{Ilin2005}
+    \textsl{Ильин В.~А., Позняк Э.~Г.}
+    Основы математического анализа.  Ч.~1: Учеб. для вузов~/ 7-е изд.\,---\,М.:
+    Физматлит, 2005.\,---\,648~с.
+
+\end{thebibliography}
+
 \end{document}

### 2018-09-23 vmmsh-2018/BetterLinearBound.cp1251.tex
[vmmsh] Recoding to 1251 - WIP
[vmmsh] It is rather built than not

--- 2018-09-18 vmmsh-2018/BetterLinearBound.tex
@@
 
-\begin{lemma}
+\begin{aLema}
 	\label{lemma_points_on_line}
@@
 	Тогда никакая прямая $m$ не содержит точки $M_1$, $M_2$, $M_3$ и $M_4$ одновременно.
-\end{lemma}
+\end{aLema}
 
@@
 
-\begin{corollary}
+\begin{aSledstvie}
 	\label{corollary:max_points_on_line}
@@
 	Тогда ни на какой прямой не лежит более $(d+3)/2$ точек из $S$.
-\end{corollary}
+\end{aSledstvie}
 
@@
 
-\begin{lemma}
+\begin{aLema}
 	\label{lemma:square_container}
@@
 	Тогда $S$ содержится в квадрате со стороной $d$.
-\end{lemma}
+\end{aLema}
 
@@
 
-\begin{hypothesis}
+\begin{aGipot}
 	Не существует такого числа $\gamma < 1$,
@@
 	содержится в квадрате со стороной $\gamma \cdot \operatorname{diam} M$.
-\end{hypothesis}
+\end{aGipot}
 
@@
 
-\begin{definition}
+\begin{aDef}
 	Крестом точек $M_1$ и $M_2$ будем называть объединение прямой,
@@
 	и обозначать $cr(M_1,M_2)$.
-\end{definition}
+\end{aDef}
 
-\begin{proposition}
+\begin{aUtv}
 	\label{proposition:intervals_cross}
@@
 	то множество $cr(M_1,M_2) \cap cr(M_3,M_4)$~--- либо прямая, либо содержит не более 4 точек.
-\end{proposition}
+\end{aUtv}
 
@@
 
-\begin{lemma}
+\begin{aLema}
 	\label{lemma_preliminary_size}
@@
 	Тогда $|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| + \frac{d-5}{2}$.
-\end{lemma}
+\end{aLema}
 
@@
 
-\begin{theorem}%[\cite{costa2013valid}]
+\begin{aTeorema}%[\cite{costa2013valid}]
 	\label{thm:varphi_k_bounds}
@@
 	\end{equation*}
-\end{theorem}
+\end{aTeorema}
 
 Нам потребуется вспомогательная
-\begin{lemma}
+\begin{aLema}
 	\label{lemma:quadr_diag_edges}
@@
 	т.е. хотя бы одна диагональ больше хотя бы одной стороны.
-\end{lemma}
+\end{aLema}
 \begin{proof}
@@
 где $n$ имеет специальный вид.
-\begin{lemma}
+\begin{aLema}
 	Пусть $S\in\mathfrak{M}_n$,
@@
 	\end{equation}
-\end{lemma}
+\end{aLema}
 
@@
 
-\begin{corollary}
+\begin{aSledstvie}
 	\label{corollary:common_linear_bound}
@@
 	Тогда $\operatorname{diam} S \geq \frac{m^2}{(m+1)^2} \mu_k \cdot |S|$.
-\end{corollary}
+\end{aSledstvie}
 
@@
 Следовательно, верно
-\begin{proposition}
+\begin{aUtv}
 	\label{proposition:linear_bound_29}
 	Для достаточно больших $n$ диаметр любого множества $M\in\mathfrak{M}_n$ не менее $\frac{3}{8}n$.
-\end{proposition}
+\end{aUtv}
 Заметим, что для $n=3$ эта оценка, очевидно, неверна.
@@
 $\operatorname{diam} M \geq 9905$ (см. утверждение~\ref{proposition:d(2,n)}).
+Разделив 9905 на 0.3584  и округлив результат до меньшего целого, получаем 27636.
 Следовательно, для $4 \leq n \leq 27636$ выполнено~\eqref{eq:linear_bound_10}.
+Если $n \geq 27636$, то $ (k-1)m^2 + 2 \leq n \leq (k-1)(m+1)^2 + 1$ для $k=10$ и некоторого $m \geq 55$.
 Тогда с учётом следствия~\ref{corollary:common_linear_bound} получаем
-\begin{proposition}
+\begin{aUtv}
 	\label{proposition:linear_bound_10}
@@
 	\end{equation*}
-\end{proposition}
+\end{aUtv}
 
-\begin{remark}
+\begin{aZam}
 	Нахождение $\varphi_k$ для $k > 30$ может позволить улучшить оценку
@@
 	которое ограничено в силу теоремы~\ref{thm:varphi_k_bounds}.
-\end{remark}
+\end{aZam}
 
-\begin{remark}
+\begin{aZam}
 	Де-факто нам требуется не произвольная расстановка точек в единичном квадрате,
@@
 	Это требование уже имеет отношение к проблеме Улама-Эрдёша (см. гипотезу~\ref{hypot:Ulam}).
-\end{remark}
+\end{aZam}

### 2018-09-23 vmmsh-2018/ReviewMultidim.cp1251.tex
[vmmsh] Recoding to 1251 - WIP
[vmmsh] It is rather built than not

--- 2018-09-18 vmmsh-2018/ReviewMultidim.tex
@@
 
-\begin{definition}
+\begin{aDef}
 	Пусть $m \geq 2$, $n \geq m + 1$.
@@
 	и $M$ не содержится ни в какой $(m-1)$--мерной гиперплоскости.
-\end{definition}
+\end{aDef}
 
@@
 
-\begin{definition}
+\begin{aDef}
 	Множество $M\in\mathfrak{M}(m,n)$ называется оптимальным,
 	если $\operatorname{diam} M = d(m,n)$.
-\end{definition}
+\end{aDef}
 
@@
 
-\begin{proposition}
+\begin{aUtv}
 \label{proposition:d(2,n)}
-\end{proposition}
+\end{aUtv}
 $(d(2, n))_{n=3,...,122} = 1$, 4, 7, 8, 17, 21, 29, 40, 51, 63, 74, 91, 104, 121,
@@
 Для многомерного случая были вычислены~\cite{piepmeyer1988raumliche,harborth1998integral,kurz2005characteristic,kurz2006konstruktion,kurz2008bounds} следующие значения.
-\begin{proposition}~
-\end{proposition}
+\begin{aUtv}~
+\end{aUtv}
 $(d(3, n))_{n=4,...,24} = 1$, 3, 4, 8, 13, 16, 17, 17, 17, 56, 65, 77, 86, 99, 112, 133, 154,
@@
 
-\begin{hypothesis}
+\begin{aGipot}
 	Для любого $n\geq 3$ выполнено $d(2,n) < d(2,n+1)$.
-\end{hypothesis}
+\end{aGipot}
 В пользу этой гипотезы говорят известные значения $d(2,n)$.
@@
 
-\begin{hypothesis}
+\begin{aGipot}
 	Никакое оптимальное множество $M\in\mathfrak{M}$ не лежит на целочисленной решётке.
-\end{hypothesis}
+\end{aGipot}
 В пользу этой гипотезы говорят реузльтаты вычислений для $m=2$, $3 \leq n \leq 43$.
@@
 
-\begin{hypothesis}
+\begin{aGipot}
 	Все оптимальные множества из $n$ точек на плоскости, $n\geq 9$, являются веерными.
-\end{hypothesis}
+\end{aGipot}
 
@@
 В работе~\cite{kurz2008minimum} высказана
-\begin{hypothesis}
+\begin{aGipot}
 	\label{hypot:log_lower_bound}
@@
 	$\operatorname{diam} M \geq n^{c \log \log n}$.
-\end{hypothesis}
+\end{aGipot}
 
@@
 
-\begin{definition}
+\begin{aDef}
 	\cite{kurz2008minimum} Множество $M\in\mathfrak{M}(m,n)$ называется множеством полуобщего положения,
@@
 	Для краткости обозначим $\overline{\mathfrak{M}}_n=\overline{\mathfrak{M}}(2,n)$.
-\end{definition}
+\end{aDef}
 
@@
 
-\begin{definition}
+\begin{aDef}
 	\cite{kurz2008minimum} Множество $M\in\overline{\mathfrak{M}}(m,n)$, $n\geq 4$ называется множеством общего положения,
@@
 	Для краткости обозначим $\dot{\mathfrak{M}}_n=\dot{\mathfrak{M}}(2,n)$.
-\end{definition}
+\end{aDef}
 
@@
 
-\begin{hypothesis}
+\begin{aGipot}
 	\label{hypot:general_position_nonempty}
 	Для любого $n\geq 4$ непусто $\dot{\mathfrak{M}}_n$.
-\end{hypothesis}
+\end{aGipot}
 
-\begin{definition}
+\begin{aDef}
 	\cite{noll1989nclusters,kurz2013constructing}
@@
 	содержащееся в целочисленной решётке.
-\end{definition}
+\end{aDef}
 Существование $n$--кластера эквивалентно существованию множества $M\in\dot{\mathfrak{M}}_n$
@@
 неизвестно, однако легко доказать следующее
-\begin{proposition}
+\begin{aUtv}
 	\label{prop:edge_one}
@@
 	что для некоторых $M_1, M_2$ выполнено $|M_1 - M_2| = 1$.
-\end{proposition}
+\end{aUtv}
 \begin{proof}
@@
 С исследованием множеств $\mathfrak{M}_n$ тесно связаны ещё две гипотезы о расстояниях на плоскости.
-\begin{hypothesis}[Улам, \cite{ulam1960collection}]
+\begin{aGipot}[Улам, \cite{ulam1960collection}]
 	\label{hypot:Ulam}
@@
 	где $\mathbb{Q}$~--- множество рациональных чисел.
-\end{hypothesis}
+\end{aGipot}
 Если гипотеза~\ref{hypot:Ulam} справедлива, то для любого $n\geq 4$ множество $\dot{\mathfrak{M}}_n$ непусто
@@
 Неизвестно, следует ли из
-гипотезы~\ref{hypot:general_position_nonempty} следует гипотеза~\ref{hypot:Ulam}.
+гипотезы~\ref{hypot:general_position_nonempty} гипотеза~\ref{hypot:Ulam}.
 
@@
 В работе~\cite{guth2015erdos} эта проблема в своей изначальной постановке была решена и доказана следующая
-\begin{theorem}
+\begin{aTeorema}
 	Пусть $G\subset{R}^2$, $|G| = n$.
@@
 	не менее $cn/\log n $.
-\end{theorem}
+\end{aTeorema}
 

### 2018-09-23 vmmsh-2018/AvdeevNN_09.cp1251.tex
[vmmsh] Meta

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Тезисы доклада на XIV «Владикавказская молодежная математическая школа»
@@
 
- \Author{ФИО}
+ \Author{Авдеев Н.Н., Семёнов Е.М.}
 
-\Abstract{Текст аннотации. \par}
+\Abstract{
+	Рассмотрены такие множества точек на плоскости,
+	что расстояние между любыми двумя точками множества есть целое число
+	и множество не содержится ни в какой прямой.
+	Приведены оценки на минимальный диаметр таких множеств.
+	Дано обобщение этих результатов на n-мерный случай.
+	Также рассмотрены некоторые частные случаи таких множеств
+	и дан обзор известных результатов.
+\par}
 
 {\hangindent17pt\hangafter=0\noindent\footnotesize{\bf Ключевые
- слова:} ключевое слово 1, ключевое слово 2, \ldots\par}
+ слова:}
+	комбинаторная геометрия,
+	целочисленные расстояния,
+	диаметр множества,
+	экстремальные подмножества точек евклидова пространства,
+	проблема Эрдёша
+\par},
 
@@
 
+ \Adress{\textsc{Авдеев Николай Николаевич}\\
+ Воронежский госуниверситет\\
+ Россия, Воронеж, Университетская пл., 1\\
+ avdeev@math.vsu.ru}
+
+ \Adress{\textsc{Семенов Евгений Михайлович}\\
+ Воронежский госуниверситет\\
+ Россия, Воронеж, Университетская пл., 1\\
+ E-mail}
+
+ \bigskip
+
+ \begin{center}
+ НАЗВАНИЕ СТАТЬИ НА АНГЛИЙСКОМ ЯЗЫКЕ
+ \end{center}
+ \begin{center}
+  Avdeev N.N., Semenov E.M.
+ \end{center}
+
+\Abstract{
+	We considered such a sets of points on the plane
+	that the distance between any two points is integral
+	and the set isn't contained in any straight line.
+	Minimal diameter of such sets is estimated.
+	The results are generalized to n-dimensional euclidean space.
+	Some special cases of such sets are considered
+	and known results are reviewed.
+\par}
+
+\vspace*{-4pt}
+
+{\hangindent17pt\hangafter=0\noindent\footnotesize {\bf Key
+ words:}~
+	combinatorial geometry,
+	integral distances,
+	diameter of point set,
+	extremal subsets of Euclidean space,
+	integral point sets,
+	Erd\"{o}s problem,
+	minimum diameter,
+	characteristic of an integral point set
+\par}
+
 \end{document}

### 2018-09-23 vmmsh-2018/AvdeevNN_09.cp1251.tex
[vmmsh] Meta

@@
  Воронежский госуниверситет\\
- Россия, Воронеж, Университетская пл., 1\\
+ Россия, 394036, Воронеж, Университетская пл., 1\\
  avdeev@math.vsu.ru}
@@
  Воронежский госуниверситет\\
- Россия, Воронеж, Университетская пл., 1\\
- E-mail}
+ Россия, 394036, Воронеж, Университетская пл., 1\\
+ nadezhka\_ssm@geophys.vsu.ru}
 

### 2018-09-23 vmmsh-2018/AvdeevNN_09.cp1251.tex
[vmmsh] Joining of files

@@
 
-\input{IntroBySemenov.cp1251.tex}
+Oбозначим через $\mathfrak{M}$ множество таких подмножеств
+$
+	M = \{M_1, M_2, ... \}\subset \mathbb{R}^2
+$,
+что
+$
+	|M_i - M_j|\in\mathbb{N}
+$
+для всех $i$, $j$, где $\mathbb{N}$~--- множество натуральных чисел
+и $|M_i - M_j|$~--- расстояние между точками $M_i$ и $M_j$.
+П. Эрдёш доказал~\cite{erdos1945integral},
+что справедлива следующая
 
-\input{BetterLinearBound.cp1251.tex}
+\begin{aTeorema}
+	\label{thmErdos}
+	Всякое бесконечное подмножество $M\in\mathfrak{M}$
+	содержится в некоторой прямой $L\in\mathbb{R}^2$.
+\end{aTeorema}
 
-\input{ReviewChar.cp1251.tex}
+Теорему~\ref{thmErdos} можно доказать, основываясь на двух элементарных леммах.
+Интересно, что работа~\cite{erdos1945integral} была опубликована через несколько
+месяцев после совместной работы Эннинга и Эрдёша~\cite{anning1945integral},
+в которой было дано другое, более сложное доказательство.
 
-\input{ReviewMultidim.cp1251.tex}
+\begin{aLema}
+	\label{thmQuadCurves}
+	Пусть $A$ и $B$~--- две кривые второго порядка на плоскости и $A$ не является парой прямых.
+	Тогда $A\cap B$ содержит не более 4 точек.
+\end{aLema}
+
+\begin{proof}
+	Без ограничения общности кривая $A$ определяется уравнением
+	\begin{equation*}
+		x^2 = ay^2 + by +c
+		,
+	\end{equation*}
+	а кривая $B$ определяется уравнением второго порядка общего вида.
+	Система этих двух уравнений легко сводится к уравнению с одной неизвестной 4-го порядка.
+	Если это уравнение не является тривиальным, то оно имеет не более 4 корней.
+\end{proof}
+
+Если $A$ и $B$~--- пары прямых $xy=0$ и $xy=x$,
+то $A\cap B$ содержит ось игреков.
+Поэтому предположение в лемме~\ref{thmQuadCurves} о том,
+что одна из кривых не является парой прямых, существенно.
+
+\begin{aLema}
+	\label{thm:quadCurveFamily}
+	Пусть $M_1, M_2 \in \mathbb{R}^2$ и
+	$|M_1 - M_2| = n \in \mathbb{N}$.
+	Тогда
+	\begin{equation}
+		\label{eq:quadCurveFamily}
+		\left\{
+			M: |M-M_1|\in\mathbb{N}, |M-M_2|\in\mathbb{N}
+		\right\}
+		=\bigcup_{i=1}^n Q_i
+		,
+	\end{equation}
+	где $Q_1$ есть пара взаимно перпендикулярных прямых (крест)
+	и $Q_2, Q_3,...,Q_n$~--- гиперболы.
+\end{aLema}
+
+\begin{proof}
+	По неравенству треугольника
+	\begin{equation*}
+		\bigl| |M-M_1| - |M-M_2| \bigr|
+		\leqslant
+		|M_1 - M_2| = n
+		.
+	\end{equation*}
+	Так как $\bigl| |M-M_1| - |M-M_2| \bigr|$
+	есть целое неотрицательное число, то
+	\begin{equation*}
+		\bigl| |M-M_1| - |M-M_2| \bigr|
+		\in\{0,1,2,...,n\}
+		.
+	\end{equation*}
+	Ясно, что множество
+	\begin{equation}
+		\label{eq:centrPerp}
+		\left\{ M: \bigl| |M-M_1| - |M-M_2| \bigr| = 0 \right\}
+	\end{equation}
+	принадлежит прямой, проходящей через середину отрезка
+	$[M_1, M_2]$ перпендикулярно этому отрезку,
+	\begin{equation}
+		\label{eq:tails}
+		\left\{ M: \bigl| |M-M_1| - |M-M_2| \bigr| = n \right\}
+	\end{equation}
+	содержится в прямой, проходящей через точки $M_1, M_2$,
+	\begin{equation}
+		\label{eq:hyperbola}
+		\left\{ M: \bigl| |M-M_1| - |M-M_2| \bigr| = k \right\}
+	\end{equation}
+	есть гипербола для любого $k=1,2,...,n-1$.
+	Обозначая объединение множеств \eqref{eq:centrPerp} и \eqref {eq:tails}
+	через $Q_1$,
+	а множества \eqref{eq:hyperbola} через $Q_i$,
+	мы получаем требуемое представление.
+	Таким образом, \eqref{eq:quadCurveFamily}
+	есть объединение $n$ кривых второго порядка.
+\end{proof}
+
+\begin{proof}[Доказательство теоремы \ref{thmErdos}]
+	Предположим противное,
+	т.е. $\{M_1, M_2, ... \}\in\mathfrak{M}$
+	и точки $M_1$, $M_2$, $M_3$ не лежат на одной прямой.
+	Применяя лемму~\ref{thm:quadCurveFamily}
+	к паре точек $M_1, M_2$ и к паре точек $M_1, M_3$,
+	найдём такие кривые 2-го порядка
+	$Q_1, Q_2,...,.Q_n$ и $R_1, R_2,...,.R_m$, что
+	\begin{equation*}
+		\left\{
+			M: |M-M_1|\in\mathbb{N}, |M-M_2|\in\mathbb{N}
+		\right\}
+		=\bigcup_{i=1}^n Q_i
+		,
+	\end{equation*}
+	\begin{equation*}
+		\left\{
+			M: |M-M_1|\in\mathbb{N}, |M-M_3|\in\mathbb{N}
+		\right\}
+		=\bigcup_{j=1}^m R_j
+		.
+	\end{equation*}
+	Отсюда
+	\begin{multline*}
+		\left\{
+			M:
+			|M-M_1|\in\mathbb{N},
+			|M-M_2|\in\mathbb{N},
+			|M-M_3|\in\mathbb{N}
+		\right\}
+		\subset
+		\\ \subset
+		\left(\bigcup_{i=1}^n Q_i\right)
+		\cap
+		\left(\bigcup_{j=1}^m R_j\right)
+		=
+		\bigcup_{i=1}^n \bigcup_{j=1}^m \left( Q_i \cap R_i \right)
+		.
+	\end{multline*}
+	Из предположения о том, что $M_1$, $M_2$, $M_3$
+	не лежат на одной прямой, вытекает,
+	что $Q_1\cap R_1$ не содержит прямую.
+	Поэтому $|Q_1\cap R_1|\leqslant 4$,
+	и, следовательно, в силу леммы~\ref{thmQuadCurves}
+	$|Q_i\cap R_j|\leqslant 4$
+	для всех $i=1,2,...,n$, $j=1,2,...,m$.
+	Отсюда
+	\begin{equation*}
+		|M| \leqslant 4nm
+		,
+	\end{equation*}
+	т.е. $M$ конечно,
+	что противоречит предположению о бесконечности $M$.
+	Полученное проиворечие доказывает, что все точки $M$
+	принадлежат некоторой прямой $L$.
+	В качестве $L$ можно взять прямую,
+	проходящую через пару любых точек из $M$.
+\end{proof}
+
+Предположение о бесконечности $M$ в теореме~\ref{thmErdos} существенно.
+Для заданного $n\in\mathbb{N}$ обозначим через $\mathfrak{M}_n$
+множество таких $M\in\mathfrak{M}$, что
+$|M|=n$ и $M \not\subset L$ для любой прямой $L \subset\mathbb{R}^2$.
+
+\begin{aTeorema}
+	\label{thm:power_exist}
+	Для любого $n\in\mathbb{N}$ выполнено $\mathfrak{M}_n\neq\varnothing$.
+\end{aTeorema}
+
+\begin{proof}
+	Из теоремы Пифагора следует, что треугольник со сторонами
+	$2k+1$, $2k^2+2k$, $2k^2+2k+1$
+	является прямоугольным для любого $k\in\mathbb{N}$.
+	Выпишем стороны первых трёх треугольников,
+	а затем стороны подобных им треугольников с совпадающей первой стороной:
+	\begin{equation}
+		\label{eq:PifagTriangles}
+		\begin{array}{ll}
+			3,  4,  5 & \quad 105, 140, 175 \\
+			5, 12, 13 & \quad 105, 252, 273 \\
+			7, 24, 25 & \quad 105, 360, 375.
+		\end{array}
+	\end{equation}
+	Ясно, точки $(0,0)$, $(0,105)$, $(140,0)$, $(252,0)$, $(360,0)$
+	образуют множество из $\mathfrak{M}_5$.
+	Очевидно, если взять не 3, а $n$ точек, и затем найти подобные
+	прямоугольные треугольники с совпадающей первой стороной,
+	то можно построить множество из $\mathfrak{M}_{n+2}$.
+\end{proof}
+
+Другой пример множества из $\mathfrak{M}_{n}$ для заданного $n$
+можно найти в доказательстве теоремы~\ref{thm:power_char_exist},
+уверждение которой усиливает утверждение теоремы~\ref{thm:power_exist}.
+
+Для конечного дискретного подмножества плоскости $M\subset \mathbb{R}^2$
+естественным образом определяется диаметр:
+\begin{equation*}
+	\operatorname{diam} M = \max_{A,B\in M} |A-B|
+	.
+\end{equation*}
+
+Солимоси~\cite{solymosi2003note} показал существование такой константы $c$,
+что для любого $S\in\mathfrak{M}_n$ выполнено $\operatorname{diam} M \geq c n$.
+Солимоси не привёл значения этой константы, однако из его доказательства можно вывести,
+что эта константа равна $1/24$.
+
+Мы сейчас дадим другое доказательство, дающее лучшее значение константы $c$.
+
+\begin{aLema}
+	\label{lemma_points_on_line}
+	Пусть $S\in\mathfrak{M}_n$ для некоторого $n \geqslant 3$,
+	$\{M_1, M_2, M_3, M_4\} \subset S$, как минимум три из этих точек различны,
+	$|M_1 - M_2| = |M_3 - M_4| = 1$.
+	Тогда никакая прямая $m$ не содержит точки $M_1$, $M_2$, $M_3$ и $M_4$ одновременно.
+\end{aLema}
+
+\begin{proof}
+	Предположим противное.
+	Обозначим через $m_{12}$ и $m_{34}$ серединные перпендикуляры к отрезкам $M_1 M_2$ и $M_3 M_4$ соответственно.
+
+	Пусть точка $M\in S$.
+	Тогда либо $|M - M_1| - |M - M_2| = 0$ и $M\in m_{12}$, либо $\Bigl||M - M_1| - |M - M_2|\Bigr| = 1$ и $M\in m$,
+	т.е. в любом случае $M\in m \cup m_{12}$.
+	Аналогично $M\in m \cup m_{34}$.
+	Следовательно, $M\in (m \cup m_{12}) \cap (m \cup m_{34}) = m \cup (m_{12} \cap m_{34}) = m$,
+	т.к. $m_{12} \cap m_{34} = \varnothing$ (перпендикуляры к одной прямой, проведённые в разных точках, не пересекаются между собой).
+
+	В силу произвольности выбора $M \in S$ получаем $S \subset m$, что противоречит $S\in\mathfrak{M}_n$.
+\end{proof}
+
+\begin{aSledstvie}
+	\label{corollary:max_points_on_line}
+	Пусть $S\in\mathfrak{M}_n$, $\operatorname{diam} S = d$.
+	Тогда ни на какой прямой не лежит более $(d+3)/2$ точек из $S$.
+\end{aSledstvie}
+
+\begin{proof}
+	Предположим противное, т.е на прямой $m$ лежит $k$ точек из $S$ и $k > (d+3)/2$.
+	Пусть $\rho_i$, $i = 1,...,k-1$~--- расстояния между соседними точками,
+	пронумерованные в порядке возрастания.
+	Из леммы~\ref{lemma_points_on_line} получаем $\rho_{k-1}\geq...\geq\rho_2\geq 2$.
+	Значит,
+	\begin{equation*}
+		\sum_{i=1}^{k-1} \rho_i
+		=
+		\rho_1 + \sum_{i=2}^{k-1} \rho_i
+		\geq
+		1 + \sum_{i=2}^{k-1} \rho_i
+		\geq
+		1 + 2(k-2)
+		=
+		2k-3
+		>
+		d+3-3
+		=
+		d
+		,
+	\end{equation*}
+	т.е. расстояние между крайними точками превосходит диаметр системы,
+	чего быть не может.
+\end{proof}
+
+\begin{aLema}
+	\label{lemma:square_container}
+	Пусть $S\in\mathfrak{M}_n$, $\operatorname{diam} S = d$.
+	Тогда $S$ содержится в квадрате со стороной $d$.
+\end{aLema}
+
+\begin{proof}
+	Пусть $M_1, M_2 \in S$, $|M_1 - M_2| = d$,
+	т.е. диаметр системы $S$ достигается на $M_1 M_2$.
+	Введём прямоугольную декартову систему координат на плоскости таким образом, что
+	$M_1 = (-d/2; 0)$, $M_2 = (d/2; 0)$.
+
+	Тогда прочие точки $M_i \in S$ имеют координаты $M_i=(x_i, y_i)$.
+	Очевидно, что для $i>2$ имеем $|x_i| < d/2$
+	(иначе $|M_i - M_1| > d$ или $|M_i - M_2| > d$).
+	Более того, $|y_i| < d$ (иначе $|M_i - M_1| > d$ или $|M_i - M_2| > d$).
+
+	Пусть $y_{\max} = \max_{i} y_i$, $y_{\min} = \min_{i} y_i$, $M_{\max}=(x_{\max}, y_{\max})$, $M_{\min}=(x_{\min}, y_{\min})$
+	(если максимум достигается на нескольких точках, возьмём любые из них).
+	Тогда
+	\begin{multline}
+		d \geq |M_{\max} - M_{\min}| = \sqrt{(x_{\max} - x_{\min})^2 + (y_{\max} - y_{\min})^2}
+		\geq
+		\\\geq
+		\sqrt{(y_{\max} - y_{\min})^2} =
+		y_{\max} - y_{\min}
+	\end{multline}
+	Следовательно, $S$ содержится в квадрате со сторонами $x=\pm d/2$,
+	$y=y_{\max}$, $y=y_{\max} - d$.
+\end{proof}
+
+\begin{aGipot}
+	Не существует такого числа $\gamma < 1$,
+	что любое множество из $M\in \mathfrak{M}_n$, $n\geq 3$,
+	содержится в квадрате со стороной $\gamma \cdot \operatorname{diam} M$.
+\end{aGipot}
+
+В пользу этой гипотезы говорят часто встречающиеся конструкции множеств $M\in\mathfrak{M}_n$,
+расположенных на окружности~\cite{anning1915discussions,harborth1993upper,piepmeyer1996maximum,kurz2008bounds,our-vvmsh-2018}.
+
+\begin{aDef}
+	Крестом точек $M_1$ и $M_2$ будем называть объединение прямой,
+	проходящей через эти точки,
+	и серединного перпендикуляра к отрезку $M_1 M_2$
+	и обозначать $cr(M_1,M_2)$.
+\end{aDef}
+
+\begin{aUtv}
+	\label{proposition:intervals_cross}
+	Если интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются,
+	$M_1 \neq M_2$, $M_3 \neq M_4$,
+	то множество $cr(M_1,M_2) \cap cr(M_3,M_4)$~--- либо прямая, либо содержит не более 4 точек.
+\end{aUtv}
+
+\begin{proof}
+	Пересечением $cr(M_1,M_2) \cap cr(M_3,M_4)$ может быть либо от 2 до 4 точек, либо прямая,
+	либо, если эти кресты совпадают, их пересечение совпадает с каждым из этих крестов.
+	Но $cr(M_1,M_2) \neq cr(M_3,M_4)$, потому что середины отрезков $M_1 M_2$ и $M_3 M_4$
+	совпасть не могут, т.к. интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются по условию.
+\end{proof}
+
+\begin{aLema}
+	\label{lemma_preliminary_size}
+	Пусть $S\in\mathfrak{M}_n$,
+	$M_1, M_2, M_3, M_4 \in S$,
+	$M_1 \neq M_2$, $M_3 \neq M_4$,
+	интервалы $M_1 M_2$ и $M_3 M_4$ не пересекаются,
+	$d = \operatorname{diam} S > 5$.
+	Тогда $|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| + \frac{d-5}{2}$.
+\end{aLema}
+
+\begin{proof}
+	Если $|(cr(M_1, M_2) \cap cr(M_3, M_4))| < \infty$,
+	то (см. доказательство теоремы~\ref{thmErdos} или~\cite[часть 2, неравенство (1)]{solymosi2003note})
+	$|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4|$.
+	Иначе (в силу утверждения~\ref{proposition:intervals_cross}) $cr(M_1, M_2) \cap cr(M_3, M_4) = m$,
+	где $m$~--- прямая.
+	В силу следствия~\ref{corollary:max_points_on_line} на этой прямой лежит не более $(d+3)/2$ точек;
+	кроме того, из общего количества точек нужно вычесть 4,
+	которые бы дали эти 2 креста в случае дискретного переcечения.
+	Получаем верхнюю оценку
+	\begin{multline}
+		|S| \leq 4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| - 4 + \frac{d+3}{2}
+		=
+		\\=
+		4 \cdot |M_1 - M_2| \cdot |M_3 - M_4| + \frac{d-5}{2}
+		.
+	\end{multline}
+\end{proof}
+
+Определим теперь коэффициенты упаковки точек в квадрат $\varphi_k$.
+Пусть
+\begin{equation*}
+	\Phi_k = \{ P \subset [0;1]^2 : |P|=k\}
+	,
+\end{equation*}
+где $[0;1]^2$~--- замкнутый единичный квадрат на плоскости.
+Тогда
+\begin{equation*}
+	\varphi_k = \max_{P \in \Phi_k} \min_{A,B \in P} |A - B|
+	.
+\end{equation*}
+Иначе говоря, через $\varphi_k$ будем обозначать наибольшее число, такое,
+что в единичном квадрате нельзя разместить $k$ точек так,
+чтобы расстояние между любыми двумя точками было не менее $\varphi_k$.
+Проблема отыскания $\varphi_k$ носит название проблемы упаковки точек в квадрат~\cite{locatelli2002packing,costa2013valid}.
+
+\begin{aTeorema}%[\cite{costa2013valid}]
+	\label{thm:varphi_k_bounds}
+	\cite{costa2013valid}
+	Для $k\geq 2$ выполнено
+	\begin{equation*}
+		\sqrt{\frac{2}{k\sqrt{3}}}
+		\leq
+		\varphi_k
+		\leq
+		\frac{1}{k-1} +
+		\sqrt{
+			\frac{1}{(k-1)^2}
+			+
+			\frac{2}{(k-1)\sqrt{3}}
+		}
+	\end{equation*}
+\end{aTeorema}
+
+Нам потребуется вспомогательная
+\begin{aLema}
+	\label{lemma:quadr_diag_edges}
+	Пусть $ABCD$~--- выпуклый четырёхугольник на плоскости.
+	Тогда $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
+	т.е. хотя бы одна диагональ больше хотя бы одной стороны.
+\end{aLema}
+\begin{proof}
+	Предположим, что любая диагональ меньше любой стороны.
+	Тогда в каждом треугольнике, образованном диагональю и двумя сторонами,
+	меньший угол лежит против меньшей стороны,
+	т.е. угол между сторонами, лежащий против диагонали, острый.
+	Таким образом, получаем, что все 4 угла четырёхугольника~---
+	острые, чего быть не может.
+\end{proof}
+
+Докажем теперь линейную оценку снизу на диаметр множеств из $\mathfrak{M}_n$,
+где $n$ имеет специальный вид.
+\begin{aLema}
+	Пусть $S\in\mathfrak{M}_n$,
+	$d = \operatorname{diam} S > 5$,
+	$k \geq 2$,
+	$m \geq 2$,
+	$ |S| = (k-1)m^2 + 2$.
+
+	Тогда
+	\begin{equation}
+		d \geq \mu_k (|S| - 2),
+	\end{equation}
+	где
+	\begin{equation}
+		\mu_k = \frac{\sqrt{64\varphi_k^2 (k-1)+1}-1}{16\varphi_k^2 (k-1)}
+	\end{equation}
+\end{aLema}
+
+\begin{proof}
+	Множество $S$ содержится в квадрате со стороной $d$ (см. лемму~\ref{lemma:square_container}).
+	Разобьём этот квадрат на $m^2$ маленьких квадратов со стороной $d/m$.
+	Тогда по принципу Дирихле либо:
+
+	а) хотя бы в два маленьких квадрата $Q_1$ и $Q_2$ попало не менее, чем по $k$ точек;
+
+	б) хотя бы в один маленький квадрат $Q_1$ попало не менее, чем $k+1$ точка.
+	\\
+	В случае (а) в $Q_1$ выберем точки $M_1$ и $M_2$,
+	а в $Q_2$ выберем точки $M_3$ и $M_4$
+	так, что $|M_1 - M_2| \leq \varphi_k d /m$, $|M_3 - M_4| \leq \varphi_k d/m$
+	(это всегда можно сделать в силу определения $\varphi_k$).
+
+	В случае (б) в $Q_1$ выберем точки $M_1$ и $M_2$ так, что
+	$|M_1 - M_2| \leq \varphi_{k+1} d /m \leq \varphi_k d /m$.
+	Из всех точек, кроме $M_1$, выберем $M_3$ и $M_4$ так, что
+	$|M_3 - M_4| \leq \varphi_k d/m$.
+	В силу леммы~\ref{lemma:quadr_diag_edges} это всегда можно сделать так,
+	чтобы интервалы $M_1 M_2$ и $M_3 M_4$ не пересекались.
+	Возможное совпадение $M_2$ и $M_3$ не помешает дальнейшему доказательству.
+
+	По лемме~\ref{lemma_preliminary_size}
+	\begin{equation}
+		|S| \leq 4 \left( \frac{d}{m} \varphi_k \right)^2 + \frac{d}{2} - \frac{5}{2}
+	\end{equation}
+	или, с учётом того, что $ |S| = (k-1)m^2 + 2$,
+	\begin{equation}
+		 4 \left( \frac{d}{m} \varphi_k \right)^2 + \frac{d}{2} - \left( (k-1)m^2 + \frac{9}{2}\right) \geq 0
+	\end{equation}
+
+	Положим $d = \lambda m$:
+	\begin{equation}
+		 4 \lambda^2 \varphi_k^2 + \frac{m}{2} \lambda - \left( (k-1)m^2 + \frac{9}{2}\right) \geq 0
+	\end{equation}
+
+	Решим это квадратное неравенство относительно $\lambda$,
+	зная, что $\lambda > 0$.
+	Непосредственно выпишем дискриминант:
+	\begin{multline}
+		D =
+		\frac{m^2}{4} + 4 \cdot 4 \varphi_k^2 \cdot \left( (k-1)m^2 + \frac{9}{2}\right)
+		=
+		\frac{m^2}{4} + 16 \varphi_k^2 \cdot \left( (k-1)m^2 + \frac{9}{2}\right)
+		>\\>
+		\frac{m^2}{4} + 16 \varphi_k^2 \cdot (k-1)m^2
+		=
+		\frac{m^2}{4} \cdot (1 + 64 \varphi_k^2 \cdot (k-1))
+		> 0
+	\end{multline}
+
+	Итак,
+	\begin{multline}
+		\frac{d}{m} = \lambda >
+		\frac{-\frac{m}{2} + \sqrt{\frac{m^2}{4} \cdot (1 + 64 \varphi_k^2 \cdot (k-1))} }{8 \varphi_k^2}
+		=\\=
+		\frac{-m + \sqrt{m^2  (1 + 64 \varphi_k^2 \cdot (k-1))} }{16 \varphi_k^2}
+		=
+		m\frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2}
+	\end{multline}
+	Разделим обе части неравенства на положительное число $m(k-1)$:
+	\begin{equation}
+		\frac{d}{(k-1)m^2}
+		>
+		\frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2 (k-1)}
+	\end{equation}
+	Положим
+	\begin{equation}
+		\mu_k = \frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2 (k-1)},
+	\end{equation}
+	тогда
+	\begin{equation}
+		\frac{d}{(k-1)m^2}
+		>
+		\mu_k
+	\end{equation}
+	откуда незамедлительно
+	\begin{equation}
+		d > \mu_k \cdot (|S|-2)
+		.
+	\end{equation}
+\end{proof}
+
+У только что полученной оценки есть две существенных особенности:
+во-первых, она справедлива лишь для мощности специального вида,
+во-вторых, зависит от $\varphi_k$.
+
+\begin{aSledstvie}
+	\label{corollary:common_linear_bound}
+	Пусть $S\in\mathfrak{M}_n$, $$ (k-1)m^2 + 2 \leq n \leq (k-1)(m+1)^2 + 1.$$
+	Тогда $\operatorname{diam} S \geq \frac{m^2}{(m+1)^2} \mu_k \cdot |S|$.
+\end{aSledstvie}
+
+Выбор $k$ во многом произволен.
+Так, например, для $k=29$ имеем $\varphi_k=0.226882900744209089183505212941...$
+~\cite{markot2005newverified}
+%(округление здесь и далее осуществляется в сторону уменьшения $\mu_k$, чтобы не испортить неравенство)
+и $\mu_k > 0.3753$.
+Следовательно, верно
+\begin{aUtv}
+	\label{proposition:linear_bound_29}
+	Для достаточно больших $n$ диаметр любого множества $M\in\mathfrak{M}_n$ не менее $\frac{3}{8}n$.
+\end{aUtv}
+Заметим, что для $n=3$ эта оценка, очевидно, неверна.
+
+Выбрав другое $k$, можно получить оценку, применимую, начиная с некоторого известного $n$.
+Например, $\varphi_{10} = 0.42127954398390343276882176065...$
+(для этого числа неизвестно выражение в радикалах)~\cite{degroot1990optimal10points} и
+$\mu_k > 0.3584$.
+Для $4 \leq n \leq 122$ вычислены точные нижние оценки на диаметр (см. утверждение~\ref{proposition:d(2,n)}) и верно, что
+\begin{equation}
+	\label{eq:linear_bound_10}
+	\operatorname{diam} M \geq 0.3584 \cdot |M|
+\end{equation}
+Кроме того, для $M \in \mathfrak{M}_n$, $n \geq 122$ выполнено
+$\operatorname{diam} M \geq 9905$ (см. утверждение~\ref{proposition:d(2,n)}).
+Разделив 9905 на 0.3584  и округлив результат до меньшего целого, получаем 27636.
+Следовательно, для $4 \leq n \leq 27636$ выполнено~\eqref{eq:linear_bound_10}.
+Если $n \geq 27636$, то $ (k-1)m^2 + 2 \leq n \leq (k-1)(m+1)^2 + 1$ для $k=10$ и некоторого $m \geq 55$.
+Тогда с учётом следствия~\ref{corollary:common_linear_bound} получаем
+\begin{aUtv}
+	\label{proposition:linear_bound_10}
+	Для всех $n \geq 4$, $M\in\mathfrak{M}$
+	\begin{equation*}
+		\operatorname{diam} M \geq \frac{55^2}{56^2} \cdot 0.3584 \cdot |M|
+		>
+		0.3457 \cdot |M|
+		.
+	\end{equation*}
+\end{aUtv}
+
+\begin{aZam}
+	Нахождение $\varphi_k$ для $k > 30$ может позволить улучшить оценку
+	в утверждениях~\ref{proposition:linear_bound_29} и~\ref{proposition:linear_bound_10},
+	равно как и вычисление точных нижних оценок на диаметр для конкретных $n>122$ (в утверждении~\ref{proposition:linear_bound_10}),
+	но лишь до определённого предела, поскольку $\mu_k$ зависит от $\varphi_k$ и $k$
+	только в составе выражения $\varphi_k^2 (k-1)$,
+	которое ограничено в силу теоремы~\ref{thm:varphi_k_bounds}.
+\end{aZam}
+
+\begin{aZam}
+	Де-факто нам требуется не произвольная расстановка точек в единичном квадрате,
+	а такая расстановка, при которой все расстояния между точками~--- рациональные числа.
+	Это требование уже имеет отношение к проблеме Улама-Эрдёша (см. гипотезу~\ref{hypot:Ulam}).
+\end{aZam}
+
+Конструкцию $\mathfrak{M}_n$ можно распространить на пространства более высокой размерности.
+
+%В пространствах более высокой размерности можно построить множества, аналогичные $\mathfrak{M}$.
+
+\begin{aDef}
+	Пусть $m \geq 2$, $n \geq m + 1$.
+	Через $\mathfrak{M}(m,n)$ будем обозначать множество таких подмножеств $m$--мерного евклидова пространства
+	$M\subset\mathbb{R}^m$, что $|M| = n$, для любых $M_1,M_2 \in M$ выполнено $|M_1 - M_2| \in\mathbb{N}$
+	и $M$ не содержится ни в какой $(m-1)$--мерной гиперплоскости.
+\end{aDef}
+
+Более того, конструкцию $\mathfrak{M}_n$ можно распространить и на $\mathbb{Z}^m_n$~\cite{Kohnert2006IntegralPS}.
+%Более того, аналогичные множества были построены в $\mathbb{Z}^m_n$~\cite{Kohnert2006IntegralPS}.
+
+Характеристика (наряду с диаметром и мощностью) обобщается на случай, когда множество точек с целочисленными расстояниями
+рассматривается в $\mathbb{R}^m$~\cite{kurz2005characteristic}.
+
+При изучении множеств $\mathfrak{M}(m,n)$ возникает закономерный вопрос
+о минимальном диаметре для заданных размерности и мощности.
+Говоря формально, определим функцию
+\begin{equation*}
+	d(m,n) = \min_{M\in\mathfrak{M}(m,n)} \operatorname{diam} M
+	.
+\end{equation*}
+
+\begin{aDef}
+	Множество $M\in\mathfrak{M}(m,n)$ называется оптимальным,
+	если $\operatorname{diam} M = d(m,n)$.
+\end{aDef}
+
+Приведём известные оценки на $d(m,n)$
+(за основу взят список из~\cite{kurz2008bounds}):
+
+%, дополненный результатом работы~\cite{nozaki2013lower}
+%и результатом, полученным в данной статье.
+
+\begin{align}
+	& d(m, n - 1) \leq d(m,n) \\
+	& d(n, n + 1) = 1 \\
+	& d(m, n) \leq \begin{cases}
+		2^{n-m+1} -2 & \mbox{~для~} n-m \equiv 0 \mod 2
+		\\
+		3(2^{n-m} -1) & \mbox{~для~} n-m \equiv 1 \mod 2
+	\end{cases} \qquad \mbox{\cite{harborth1985diameters}} \\
+	& d(2,n) \geq \frac{3}{8}n \qquad \mbox{для достаточно больших $n$ (утверждение~\ref{proposition:linear_bound_29})} \\
+	& d(2,n) > 0.3457 n \qquad \mbox{для $n\geq 4$ (утверждение~\ref{proposition:linear_bound_10})} \\
+	& d(m,n) \leq (n-m)^{c \log \log (n-m)} \mbox{~для некоторого $c$} \qquad \mbox{\cite{harborth1993upper}} \label{upper_bound_d_m_n_log_log}\\
+	& d(m,n) > \sqrt{\frac{3}{2m}} n^{1/m} \mbox{~\cite{kanold1981punktmengen}, в частных случаях улучшена в \cite{nozaki2013lower}} \\
+	& d(3,n) > \frac{1}{\sqrt{14}}n^{1/2} \mbox{~для~} n \geq 5 \quad \mbox{\cite{kanold1981punktmengen}} \\
+	& d(m, 2m + 1) \leq 8 \qquad \mbox{\cite{piepmeyer1988raumliche}} \\
+	& d(m, 2m + 2) \leq 13 \qquad \mbox{\cite{piepmeyer1988raumliche}} \\
+	& d(m, 3m    ) \leq 109 \qquad \mbox{\cite{kemnitz1988punktmengen}} \\
+	& d(m, m^2 + m) \leq 17 \qquad \mbox{\cite{kurz2008bounds}} \\
+	& 3 \leq d(m, n) \leq 4 \mbox{~для~} m + 2 \leq n \leq 2m \mbox{~и~} d(m, 2m) = 4 ~ \mbox{\cite{piepmeyer1988raumliche,harborth1991point}}\\
+\end{align}
+
+Кроме того, в~\cite{kurz2008bounds} выдвинуто предположение, что $d(m - 1, n) \geq d(m, n)$.
+
+Существуют алгоритмы вычисления $d(m,n)$ на ЭВМ, однако они требуют большого количества машинного времени,
+причём требуемое время быстро растёт с ростом $m$ и $n$.
+Первые несколько значений $d(2,n)$ были получены в~\cite{harborth1998integral},
+следующие известные значения~--- в \cite{kurz2005characteristic,kurz2006konstruktion,kurz2008minimum,kurz2008bounds,our-mz-rus}.
+Алгоритм, основанный на переборе,
+может быть существенно оптимизирован с применением характеристики~\cite{kurz2005characteristic,kreisel2008heptagon}.
+Однако сложность вычисления максимальной возможной мощности для заданного диаметра $d$, по оценке~\cite{kreisel2008heptagon},
+составлет $O(d^3)$ (в двумерном случае).
+%В (TODO: ссылка!) отмечено, что важно перепроверять вычисленные значения независимо составленными программами для ЭВМ.
+Приведём некоторые известные значения:
+
+\begin{aUtv}
+\label{proposition:d(2,n)}
+\end{aUtv}
+$(d(2, n))_{n=3,...,122} = 1$, 4, 7, 8, 17, 21, 29, 40, 51, 63, 74, 91, 104, 121,
+134, 153, 164,
+196, 212, 228, 244, 272, 288, 319, 332, 364, 396, 437, 464, 494, 524, 553, 578, 608,
+642, 667, 692, 754, 816, 897, 959, 1026, 1066, 1139, 1190,  1248, 1306, 1363, 1410,
+1460, 1514, 1564, 1614, 1675, 1727, 1770, 1817, 1887, 1906, 2060, 2140, 2169,
+2231, 2299, 2432, 2494, 2556, 2624, 2692, 2827, 2895, 2993, 3098, 3196, 3294,
+3465, 3575, 3658, 3749, 3885, 3922, 4223, 4380, 4437, 4559, 4693, 4883,
+5018, 5109, 5264, 5332, 5480, 5603, 5738, 5938, 5995, 6052,
+6324, 6432, 6630, 6738, 6939, 7061, 7245, 7384, 7568, 7752, 7935, 8119, 8321,
+8406, 8648, 8729, 8927, 9052, 9211, 9423, 9534, 9794, 9905
+.
+
+Для многомерного случая были вычислены~\cite{piepmeyer1988raumliche,harborth1998integral,kurz2005characteristic,kurz2006konstruktion,kurz2008bounds} следующие значения.
+\begin{aUtv}~
+\end{aUtv}
+$(d(3, n))_{n=4,...,24} = 1$, 3, 4, 8, 13, 16, 17, 17, 17, 56, 65, 77, 86, 99, 112, 133, 154,
+195, 212, 228, 244.
+
+\begin{aGipot}
+	Для любого $n\geq 3$ выполнено $d(2,n) < d(2,n+1)$.
+\end{aGipot}
+В пользу этой гипотезы говорят известные значения $d(2,n)$.
+Для больших размерностей утверждение гипотезы неверно: так, $d(3,10) = d(3,11) = 17$.
+
+\begin{aGipot}
+	Никакое оптимальное множество $M\in\mathfrak{M}$ не лежит на целочисленной решётке.
+\end{aGipot}
+В пользу этой гипотезы говорят реузльтаты вычислений для $m=2$, $3 \leq n \leq 43$.
+
+Примечательно, что для $m=2$, $n=4$ и $n=18$ оптимальные множества не единственны.
+В~\cite{kurz2008bounds} установлено, что для $9 \leq n \leq 122$ оптимальное множество является веерным, т.е. содержится
+в объединении прямой и точки (все точки, кроме одной, лежат на некоторой прямой).
+
+\begin{aGipot}
+	Все оптимальные множества из $n$ точек на плоскости, $n\geq 9$, являются веерными.
+\end{aGipot}
+
+При изучении множеств из $\mathfrak{M}_n$ часто возникают множества специальных видов.
+Наиболее простым примером являются веерные множества.
+Курц и Вассерман~\cite{kurz2008minimum} показали, что диаметр веерного множества мощности $n$
+не менее, чем $n^{c \log \log n}$, где $c$~--- некоторая константа.
+Это результат указывает на возможность улучшения оценки в следствии~\ref{corollary:max_points_on_line}.
+
+В работе~\cite{kurz2008minimum} высказана
+\begin{aGipot}
+	\label{hypot:log_lower_bound}
+	Существует такая константа $c$,
+	что для любого $n\geq 3$ и $M\in\mathfrak{M}_n$ выполнено
+	$\operatorname{diam} M \geq n^{c \log \log n}$.
+\end{aGipot}
+
+Заметим, что неравенство в гипотезе~\ref{hypot:log_lower_bound}
+совпадает с верхней оценкой~\eqref{upper_bound_d_m_n_log_log} на $d(2,n)$
+с точностью до константы.
+
+Возникает закономерный вопрос: а какие бывают множества из $\mathfrak{M}_n$,
+кроме веерных?
+
+Положив $p=1$ в доказательстве теоремы~\ref{thm:power_char_exist} и добавив к построеннному множеству
+точку $M_{n+2} = (0, -\sqrt{2^{n+1}})$, где $n$ нечётно,
+получим множество из $n+2$ точек, из которых $n$ лежат на некоторой прямой и 2 вне её.
+
+В~\cite{our-vvmsh-2018} приведены примеры различных множеств из $\mathfrak{M}_n$:
+\begin{itemize}
+	\item
+		$
+		\left\{
+		(\pm 21/2, 0);
+		(\pm 4, 3\sqrt{35}/2);
+		(\pm 5/2, 0);
+		(\pm 11/2, 0);
+		\right\}
+		$~---
+		содержится в двух параллельных прямых (2 точки на одной, 6 на другой);
+
+	\item
+		$
+		\left\{
+		\left( {57/34} ; -{12\sqrt{35}/17}\right);
+		\left( {44/17} ; {39\sqrt{35}/34}\right);
+		\left( -{72/17} ;
+		{15\sqrt{35}/34}\right);
+		\right.$ \\ $\left.
+		\left( -{175/34} ; -{24\sqrt{35}/17}\right);
+		\left( -{57/34} ; {12\sqrt{35}/17}\right);
+		%\right.$ \\ $\left.
+		\left( \pm{17/2} ; 0\right);
+		\right\}
+		$~---
+		содержится в двух параллельных прямых (3 точки на одной, 4 на другой);
+
+	\item
+		$
+		\left\{
+		(0,0);(\pm 50, 0); (\pm 14, \pm 48)
+		\right\}
+		$~---
+		содержится в объединении окружности с её центром (7~точек);
+
+	\item
+		$
+		\left\{
+		\left( \pm30 ; 0\right);
+		%\left( 30 ; 0\right);
+		\left( \pm{15/2} ; {25\sqrt{7}/2}\right);
+		%\left( {15/2} ; {25\sqrt{7}/2}\right);
+		%\left( {33/2} ; {9\sqrt{7}/2}\right);
+		\left( 0 ; {10\sqrt{7}/1}\right);
+		\left( \pm{33/2} ; {9\sqrt{7}/2}\right);
+		\right\}
+		$~---
+		содержится в двух прямых, пересекающихся под острым углом (7~точек).
+\end{itemize}
+
+Информация о некоторых других конструкциях может быть найдена в~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved}.
+
+TODO: спрямить ссылки!
+
+\begin{aDef}
+	\cite{kurz2008minimum} Множество $M\in\mathfrak{M}(m,n)$ называется множеством полуобщего положения,
+	если ни через какие $m+1$ точек из $M$ нельзя провести $(m-1)$--мерную гиперплоскость.
+	Множество всех множеств полуобщего положения мощности $n$ в $\mathbb{R}^m$ будем обозначать $\overline{\mathfrak{M}}(m,n)$.
+	Для краткости обозначим $\overline{\mathfrak{M}}_n=\overline{\mathfrak{M}}(2,n)$.
+\end{aDef}
+
+В \cite{harborth1993upper} показано, что $\overline{\mathfrak{M}}_n\neq\varnothing$ для любого $n \geq 3$
+и приведены примеры множеств из $\overline{\mathfrak{M}}_n$, содержащихся в окружности.
+На эти множествах достигается верхняя оценка~\eqref{upper_bound_d_m_n_log_log} на $d(2,n)$
+с некоторой константой.
+
+Определим функцию~\cite{kurz2008minimum}
+\begin{equation*}
+	\overline{d}(m,n) = \min_{M\in\overline{\mathfrak{M}}(m,n)} \operatorname{diam} M
+	.
+\end{equation*}
+
+В работе~\cite{kurz2008minimum} вычислено, что
+\\
+$\bigl(\overline{d}(2, n)\bigr)_{n=3,...,36} = 1,$
+4, 8, 8, 33, 56, 56, 105, 105, 105, 532, 532, 735, 735, 735, 735,
+1995, 1995, 1995, 1995, 1995, 1995, 9555, 9555, 9555, 10672,
+13975, 13975, 13975, 13975, 13975, 13975, 13975, 13975.
+
+\begin{aDef}
+	\cite{kurz2008minimum} Множество $M\in\overline{\mathfrak{M}}(m,n)$, $n\geq 4$ называется множеством общего положения,
+	если ни через какие $m+2$ точек из $M$ нельзя провести сферу.
+	Множество всех множеств общего положения мощности $n$ будем обозначать $\dot{\mathfrak{M}}(m,n)$.
+	Для краткости обозначим $\dot{\mathfrak{M}}_n=\dot{\mathfrak{M}}(2,n)$.
+\end{aDef}
+
+В отличие от $\mathfrak{M}_n$ и $M\in\overline{\mathfrak{M}}_n$,
+существование множеств из $\dot{\mathfrak{M}}_n$ является достаточно сложным вопросом.
+Множество из $\dot{\mathfrak{M}}_6$ предъявлено в~\cite{noll1989nclusters},
+множество из $\dot{\mathfrak{M}}_7$~--- в~\cite{kurz2008minimum}.
+Определим функцию~\cite{kurz2008minimum}
+\begin{equation*}
+	\dot{d}(m,n) = \min_{M\in\dot{\mathfrak{M}}(m,n)} \operatorname{diam} M
+	.
+\end{equation*}
+Согласно~\cite{kurz2008minimum},
+$\bigl(\dot{d}(2,n)\bigr)_{n=3,...,7}
+= 1,$ 8, 73, 174, 22270.
+
+Вопрос о существовании множеств из $\dot{\mathfrak{M}}_8$
+остаётся открытым.
+
+\begin{aGipot}
+	\label{hypot:general_position_nonempty}
+	Для любого $n\geq 4$ непусто $\dot{\mathfrak{M}}_n$.
+\end{aGipot}
+
+\begin{aDef}
+	\cite{noll1989nclusters,kurz2013constructing}
+	$n$--кластером называется множество $M\in\dot{\mathfrak{M}}_n$,
+	содержащееся в целочисленной решётке.
+\end{aDef}
+Существование $n$--кластера эквивалентно существованию множества $M\in\dot{\mathfrak{M}}_n$
+характеристики 1.
+В~\cite{noll1989nclusters} показано существование 6--кластеров,
+в~\cite{kurz2013constructing}~--- 7--кластеров.
+
+Солимоси построил~\cite{solymosi2003note} для любого $n\geq 3$ конструкцию множества $M\in\mathfrak{M}_n$
+такого, что для некоторых $M_1, M_2 \in M$ выполнено $|M_1 - M_2| = 2$.
+Полученное им множество также является веерным.
+
+Известно, что для $3\leq n \leq 5$ существуют множества из $\mathfrak{M}_n$
+c минимальной длиной ребра 1.
+Все известные примеры являются веерными~\cite{harborth1993upper}.
+Существуют ли множества произвольной мощности с длиной ребра 1,
+неизвестно, однако легко доказать следующее
+\begin{aUtv}
+	\label{prop:edge_one}
+	Для $n \geq 4$ не существует таких $M\in\overline{\mathfrak{M}}_n$,
+	что для некоторых $M_1, M_2$ выполнено $|M_1 - M_2| = 1$.
+\end{aUtv}
+\begin{proof}
+	Пусть $M\in\overline{\mathfrak{M}}_n$ и для некоторых $M_1, M_2$ выполнено $|M_1 - M_2| = 1$.
+	Тогда все точки множества $M$ лежат на $cr(M_1,M_2)$.
+	$cr(M_1,M_2)$ есть объединение двух прямых, на каждой из которых может быть не более 2 точек в силу того, что
+	$M\in\overline{\mathfrak{M}}_n$,
+	откуда непосредственно следует требуемое утверждение для $n\geq 5$.
+
+	Пусть теперь $n=4$.
+	Предположим противное.
+	Пусть $M=\{M_1, M_2, M_3, M_4\} \in \overline{\mathfrak{M}}_4$,
+	$M_1 = (-1/2, 0)$, $M_2 = (1/2, 0)$, $O=(0,0)$.
+	Тогда $M_3$ и $M_4$ лежат на оси ординат (см. выше).
+	Так как $|M_3 - M_4| \in \mathbb{Z}$,
+	то площадь треугольника $M_1 M_3 M_4$ есть рациональное число.
+	По определению~\ref{def_char_classic} это означает, что $\operatorname{char} M = 1$.
+	Тогда по определению~\ref{def_char_lattice} имеем $M_3 = (0, a/2)$.
+	Понятно, что $a \neq 0$.
+	Не теряя общности, положим $a\in\mathbb{N}$.
+	Пусть $|M_1 - M_3| = b$.
+	Применяя теорему Пифагора к треугольнику $O M_1 M_3$, получим уравнение
+	\begin{equation*}
+		\frac{1}{4} + \frac{a^2}{4} = b^2
+		,
+	\end{equation*}
+	или, что то же самое,
+	\begin{equation*}
+		(2b)^2 - a^2 = 1
+		.
+	\end{equation*}
+	Это уравнение не имеет решений в натуральных числах.
+	Полученное противоречие завершает доказательство.
+\end{proof}
+Условие $n \geq 4$ в утверждении~\ref{prop:edge_one} существенно.
+Для $n=3$ существует контрпример: равносторонний треугольник со стороной 1.
+
+Если $M\in\mathfrak{M}(m,n)$ и все рёбра $M$ имеют чётную длину,
+то понятно, что множество $H$, полученное из $M$ сжатием в два раза,
+также будет принадлежать $\mathfrak{M}(m,n)$.
+Выполняя такую операцию, мы рано или поздно получим множество из $\mathfrak{M}(m,n)$ такое,
+что для некоторых $M_1,M_2\in M$ расстояние $|M_1 - M_2|$ нечётно.
+Возникает закономерный вопрос: сколько может быть нечётных расстояний?
+
+В~\cite{graham1974there} доказано, что
+для существования множества $M\in\mathfrak{M}(m,m+2)$ такого, что для любых
+$M_i,M_j\in M$, $i\neq j$ расстояние $|M_i-M_j|$ нечётно,
+необходимо и достаточно, чтобы
+$m+2\equiv 0 (\operatorname{mod} 16)$.
+В частности, не существует такого множества $M\in\mathfrak{M}_4$,
+что все его рёбра имеют нечётную длину.
+На основе этого факта в работе~\cite{piepmeyer1996maximum} с использованием результатов теории графов доказано,
+что количество рёбер нечётной длины в множестве $M\in\mathfrak{M}_n$ не превосходит
+\begin{equation}
+	\frac{n^2}{3} + \frac{r(r - 3)}{6}, \mbox{~где~} r = 1, 2, 3 \mbox{~и~} n \equiv r (\operatorname{mod} 3)
+\end{equation}
+и конструктивно показана точность этой оценки.
+Множество, на котром эта оценка достигается, лежит на окружности.
+
+Отдельные оценки количества рёбер нечётной длины для множеств из $\dot{\mathfrak{M}}_n$
+авторам неизвестны.
+
+Перспективными направлениями исследований являются, кроме непосредственно отыскания как можно более точных оценок на $d(m,n)$,
+например, поиск систем из $\mathfrak{M}_n$, содержащихся в двух прямых (параллельных или пересекающихся).
+
+С исследованием множеств $\mathfrak{M}_n$ тесно связаны ещё две гипотезы о расстояниях на плоскости.
+\begin{aGipot}[Улам, \cite{ulam1960collection}]
+	\label{hypot:Ulam}
+	Существует всюду плотное подмножество $U\subset\mathbb{R}^2$, такое,
+	что для любых $U_1, U_2 \in U$ выполнено $|U_1 - U_2|\in\mathbb{Q}$,
+	где $\mathbb{Q}$~--- множество рациональных чисел.
+\end{aGipot}
+Если гипотеза~\ref{hypot:Ulam} справедлива, то для любого $n\geq 4$ множество $\dot{\mathfrak{M}}_n$ непусто
+(гипотеза~\ref{hypot:general_position_nonempty}).
+Неизвестно, следует ли из
+гипотезы~\ref{hypot:general_position_nonempty} гипотеза~\ref{hypot:Ulam}.
+
+В работах~\cite{garibaldi2005erdos,garibaldi2011erdos} исследуется связь множеств $\mathfrak{M}_n$ и
+проблемы Эрдёша о числе различных расстояний~\cite{erdos1946sets}.
+В работе~\cite{guth2015erdos} эта проблема в своей изначальной постановке была решена и доказана следующая
+\begin{aTeorema}
+	Пусть $G\subset{R}^2$, $|G| = n$.
+	Тогда количество различных расстояний $|G_i - G_j|, G_i,G_j \in G$
+	не менее $cn/\log n $.
+\end{aTeorema}
+
+TODO: Примеры множеств - рисунки (если будет время и место).
 
@@
 
-\input{blg.cp1251.tex}
+\begin{thebibliography}{99}
+	\bibitem{erdos1945integral}
+	\textsl{Erd\"{o}s} \textsl{P.} Integral distances /\!/ Bulletin of the
+	American Mathematical Society. --- 1945. --- Т. 51, No 12. --- С. 996.
+	\bibitem{anning1945integral}
+	\textsl{Anning} \textsl{N. H.}, \textsl{Erd\"{o}s} \textsl{P.}
+	Integral distances /\!/ Bulletin of the American Mathematical
+	Society. --- 1945. --- Т. 51, No 8. --- С. 598---600.
+	\bibitem{solymosi2003note}
+	\textsl{Solymosi} \textsl{J.} Note on integral distances /\!/
+	Discrete \& Computational Geometry. --- 2003. --- Т. 30, No 2. ---
+	С. 337---342.
+	\bibitem{anning1915discussions}
+	\textsl{Anning} \textsl{N.} Discussions: Relating to a Geometric
+	Representation of Integral Solutions of Certain Quadratic
+	Equations /\!/ The American Mathematical Monthly. --- 1915. ---
+	Т. 22, No 9. --- С. 321---321.
+	\bibitem{harborth1993upper}
+	\textsl{Harborth} \textsl{H.},
+	\textsl{M\"{o}ller} \textsl{M.} An
+	diameter of integral point sets
+	Geometry. --- 1993. --- Т. 9, No 4.
+	\textsl{Kemnitz} \textsl{A.},
+	upper bound for the minimum
+	/\!/ Discrete \& Computational Geometry
+	--- С. 427---432.\bibitem{piepmeyer1996maximum}
+	\textsl{Piepmeyer} \textsl{L.} The maximum number of odd
+	integral distances between points in the plane /\!/ Discrete \&
+	Computational Geometry. --- 1996. --- Т. 16, No 1. --- С. 113---115.
+	\bibitem{kurz2008bounds}
+	\textsl{Kurz} \textsl{S.}, \textsl{Laue} \textsl{R.} Bounds for
+	the minimum diameter of integral point sets /\!/ arXiv preprint
+	arXiv:0804.1296. --- 2008.
+	\bibitem{our-vvmsh-2018}
+	\textsl{Авдеев} \textsl{Н. Н.} Некоторые примеры систем
+	Эрдёша /\!/ Современные методы теории краевых задач :
+	материалы международной конференции <<Понтрягинские
+	чтения --- XXIX>>, посвященной 90-летию Владимира
+	Александровича Ильина (2--6 мая 2018 г.) --- 2018. --- С. 30---31.
+	\bibitem{locatelli2002packing}
+	\textsl{Locatelli} \textsl{M.}, \textsl{Raber} \textsl{U.} Packing
+	equal circles in a square: a deterministic global optimization
+	approach /\!/ Discrete Applied Mathematics. --- 2002. --- Т. 122,
+	No 1---3. --- С. 139---166.
+	\bibitem{costa2013valid}
+	\textsl{Costa} \textsl{A.} Valid constraints for the point packing
+	in a square problem /\!/ Discrete Applied Mathematics. --- 2013. ---
+	Т. 161, No 18. --- С. 2901---2909.
+	\bibitem{markot2005newverified}
+	\textsl{Mark\'{o}t} \textsl{M. C.}, \textsl{Csendes} \textsl{T.} A
+	new verified optimization technique for the "packing circles in
+	a unit square" problems /\!/ SIAM Journal on Optimization. ---
+	2005. --- Т. 16, No 1. --- С. 193---219.
+	\bibitem{degroot1990optimal10points}
+	\textsl{DeGroot} \textsl{C.}, \textsl{Peikert} \textsl{R.},
+	\textsl{W\"{u}rtz} \textsl{D.} The optimal packing of ten equal
+	circles in a square. --- Eidgen\"{o}ssische Technische Hochschule
+	Z\"{u}rich, Interdisziplin\"{a}res Projektzentrum f\"{u}r Supercomputing,
+	1990.\bibitem{kemnitz1988punktmengen}
+	\textsl{Kemnitz} \textsl{A.} Punktmengen mit ganzzahligen
+	Abst\"{a}nden. --- 1988.
+	\bibitem{Bukhstab-number-theory}
+	\textsl{Бухштаб} \textsl{А.} Теория чисел. --- 1966. --- С. 384.
+	\bibitem{our-mkmitu-2016}
+	\textsl{Авдеев} \textsl{Н. Н.} Об отыскании множеств точек на
+	плоскости с целочисленными расстояниями /\!/ Математическое
+	и компьютерное моделирование, информационные технологии
+	управления: сб. тр. Школы для студентов, аспирантов и
+	молодых ученых <<МКМИТУ-2016>>. --- 2016. --- С. 15---19.
+	\bibitem{our-ped-2017}
+	\textsl{Авдеев} \textsl{Н. Н.} О множествах точек на
+	плоскости с целочисленными расстояниями и алгоритмах их
+	отыскания /\!/ Некоторые вопросы анализа, алгебры, геометрии
+	и математического образования: материалы международной
+	молодежной научной школы <<Актуальные направления
+	математического анализа и смежные вопросы>>. --- 2017. --- Т. 7,
+	No I. --- С. 9---10.
+	\bibitem{polygons-on-lattices}
+	\textsl{Вавилов} \textsl{В.}, \textsl{Устинов}
+	Многоугольники на решетках. --- 2006.
+	\textsl{А.}
+	\bibitem{Kohnert2006IntegralPS}
+	\textsl{Kohnert} \textsl{A.}, \textsl{Kurz} \textsl{S.} Integral
+	point sets over $Z^m_n$
+	/\!/ Discrete Applied Mathematics. --- 2006. ---
+	Т. 157 (2009). --- С. 2105---2117.
+	\bibitem{kurz2005characteristic}
+	\textsl{Kurz} \textsl{S.} On the characteristic of integral point
+	sets in /\!/ arXiv preprint math/0511704. --- 2005.
+	\bibitem{harborth1985diameters}
+	\textsl{Harborth} \textsl{H.}, \textsl{Kemnitz} \textsl{A.}
+	Diameters of integral point sets /\!/ Colloquia Mathematica
+	Societatis Jano Bolyai. Т. 48. --- 1985.
+	\bibitem{kanold1981punktmengen}
+	\textsl{Kanold}
+	\textsl{H.-J.}
+	\"{U}ber
+	Punktmengen
+	im
+	$k$-dimensionalen euklidischen Raum /\!/ Abh. Braunschw. Wiss.
+	Ges. --- 1981. --- Т. 32, No 55---65. --- С. 252.
+	\bibitem{nozaki2013lower}
+	\textsl{Nozaki} \textsl{H.} Lower bounds for the minimum
+	diameter of integral point sets /\!/ Australasian Journal of
+	Combinatorics. --- 2013. --- Т. 56. --- С. 139---143.
+	\bibitem{piepmeyer1988raumliche}
+	\textsl{Piepmeyer}
+	\textsl{L.}
+	R\"{a}umliche
+	ganzzahlige
+	Punktmengen: дис. . . . канд. / Piepmeyer L. --- Master’s
+	thesis, TU Braunschweig, 1988.
+	\bibitem{harborth1991point}
+	\textsl{Harborth} \textsl{H.}, \textsl{Piepmeyer} \textsl{L.}
+	Point sets with small integral distances /\!/ Applied geometry and
+	discrete mathematics: the Victor Klee Festschrift. Т. 4. --- American
+	Mathematical Soc., 1991.
+	\bibitem{harborth1998integral}
+	\textsl{Harborth} \textsl{H.} Integral distances in point sets /\!/
+	Karl der Grosse und sein Nachwirken. 1200 Jahre Kultur und
+	Wissenschaft in Europa: Band II, Mathematisches Wissen. ---
+	1998. --- С. 213---224.
+	\bibitem{kurz2006konstruktion}
+	\textsl{Kurz} \textsl{S.} Konstruktion und Eigenschaften
+	ganzzahliger Punktmengen: дис. . . . канд. / Kurz Sascha. --- 2006.
+	\bibitem{kurz2008minimum}
+	\textsl{Kurz} \textsl{S.}, \textsl{Wassermann} \textsl{A.} On
+	the minimum diameter of plane integral point sets /\!/ arXiv
+	preprint arXiv:0804.1307. --- 2008.
+	\bibitem{our-mz-rus}
+	\textsl{Авдеев} \textsl{Н. Н.}, \textsl{Семёнов} \textsl{Е. М.}
+	О множествах точек на плоскости с целочисленными
+	расстояниями /\!/ Математические заметки. --- 2016. --- Т. 100,
+	No 5. --- С. 757---761.\bibitem{kreisel2008heptagon}
+	\textsl{Kreisel} \textsl{T.}, \textsl{Kurz} \textsl{S.} There are
+	integral heptagons, no three points on a line, no four on a circle /\!/
+	Discrete \& Computational Geometry. --- 2008. --- Т. 39, No 4. ---
+	С. 786---790.
+	\bibitem{brass2006research}
+	\textsl{Brass} \textsl{P.}, \textsl{Moser} \textsl{W. O.},
+	\textsl{Pach} \textsl{J.} Research problems in discrete
+	geometry. --- Springer Science \& Business Media, 2006.
+	\bibitem{guy2013unsolved}
+	\textsl{Guy} \textsl{R.} Unsolved problems in number theory. Т.
+	1. --- Springer Science \& Business Media, 2013.
+	\bibitem{noll1989nclusters}
+	\textsl{Noll} \textsl{L. C.}, \textsl{Bell} \textsl{D. I.} n-clusters
+	for $1 < n < 7$ /\!/ Mathematics of Computation. --- 1989. --- Т. 53,
+	No 187. --- С. 439---444.
+	\bibitem{kurz2013constructing}
+	Constructing 7-clusters / S. Kurz [и др.] /\!/ arXiv preprint
+	arXiv:1312.2318. --- 2013.
+	\bibitem{graham1974there}
+	\textsl{Graham} \textsl{R.}, \textsl{Rothschild} \textsl{B.},
+	\textsl{Straus} \textsl{E.} Are there $n + 2$ points in $E^n$ with odd
+	integral distances? /\!/ The American Mathematical Monthly. ---
+	1974. --- Т. 81, No 1. --- С. 21---25.
+	\bibitem{ulam1960collection}
+	\textsl{Ulam} \textsl{S. M.} A collection of mathematical
+	problems. Т. 8. --- Interscience Publishers, 1960.
+	\bibitem{garibaldi2005erdos}
+	\textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.} The
+	Erdos distance problem: lecture notes. --- 2005.
+	\bibitem{garibaldi2011erdos}
+	\textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.},
+	\textsl{Senger} \textsl{S.} The Erd\"{o}s Distance Problem. Т. 56. ---
+	American Mathematical Soc., 2011.\bibitem{erdos1946sets}
+	\textsl{Erd\"{o}s} \textsl{P.} On sets of distances of n points /\!/
+	The American Mathematical Monthly. --- 1946. --- Т. 53, No 5. ---
+	С. 248---250.
+	\bibitem{guth2015erdos}
+	\textsl{Guth} \textsl{L.}, \textsl{Katz} \textsl{N. H.} On the
+	Erdos distinct distances problem in the plane /\!/ Annals of
+	Mathematics. --- 2015. --- Т. 181. --- С. 155---190.
+\end{thebibliography}
 

### 2018-09-24 vmmsh-2018/AvdeevNN_09.cp1251.tex
[vmmsh] Iteration

@@
 
- \Author{Авдеев Н.Н., Семёнов Е.М.}
-
-\Abstract{
-	Рассмотрены такие множества точек на плоскости,
+ \Author{Авдеев Н.Н., Семeнов Е.М.}
+
+\Abstract{Рассмотрены такие множества точек на плоскости,
 	что расстояние между любыми двумя точками множества есть целое число
@@
 	экстремальные подмножества точек евклидова пространства,
-	проблема Эрдёша
-\par},
+	проблема Эрд\"{e}ша
+\par}
 
@@
 и $|M_i - M_j|$~--- расстояние между точками $M_i$ и $M_j$.
-П. Эрдёш доказал~\cite{erdos1945integral},
+П. Эрд\"{e}ш доказал~\cite{erdos1945integral},
 что справедлива следующая
@@
 Интересно, что работа~\cite{erdos1945integral} была опубликована через несколько
-месяцев после совместной работы Эннинга и Эрдёша~\cite{anning1945integral},
+месяцев после совместной работы Эннинга и Эрд\"{e}ша~\cite{anning1945integral},
 в которой было дано другое, более сложное доказательство.
@@
 	к паре точек $M_1, M_2$ и к паре точек $M_1, M_3$,
-	найдём такие кривые 2-го порядка
+	найд\"{e}м такие кривые 2-го порядка
 	$Q_1, Q_2,...,.Q_n$ и $R_1, R_2,...,.R_m$, что
@@
 	является прямоугольным для любого $k\in\mathbb{N}$.
-	Выпишем стороны первых трёх треугольников,
+	Выпишем стороны первых тр\"{e}х треугольников,
 	а затем стороны подобных им треугольников с совпадающей первой стороной:
@@
 что для любого $S\in\mathfrak{M}_n$ выполнено $\operatorname{diam} M \geq c n$.
-Солимоси не привёл значения этой константы, однако из его доказательства можно вывести,
+Солимоси не прив\"{e}л значения этой константы, однако из его доказательства можно вывести,
 что эта константа равна $1/24$.
@@
 	Следовательно, $M\in (m \cup m_{12}) \cap (m \cup m_{34}) = m \cup (m_{12} \cap m_{34}) = m$,
-	т.к. $m_{12} \cap m_{34} = \varnothing$ (перпендикуляры к одной прямой, проведённые в разных точках, не пересекаются между собой).
+	т.к. $m_{12} \cap m_{34} = \varnothing$ (перпендикуляры к одной прямой, провед\"{e}нные в разных точках, не пересекаются между собой).
 
@@
 	т.е. диаметр системы $S$ достигается на $M_1 M_2$.
-	Введём прямоугольную декартову систему координат на плоскости таким образом, что
+	Введ\"{e}м прямоугольную декартову систему координат на плоскости таким образом, что
 	$M_1 = (-d/2; 0)$, $M_2 = (d/2; 0)$.
@@
 	Пусть $y_{\max} = \max_{i} y_i$, $y_{\min} = \min_{i} y_i$, $M_{\max}=(x_{\max}, y_{\max})$, $M_{\min}=(x_{\min}, y_{\min})$
-	(если максимум достигается на нескольких точках, возьмём любые из них).
+	(если максимум достигается на нескольких точках, возьм\"{e}м любые из них).
 	Тогда
@@
 	\label{lemma:quadr_diag_edges}
-	Пусть $ABCD$~--- выпуклый четырёхугольник на плоскости.
+	Пусть $ABCD$~--- выпуклый четыр\"{e}хугольник на плоскости.
 	Тогда $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
@@
 	т.е. угол между сторонами, лежащий против диагонали, острый.
-	Таким образом, получаем, что все 4 угла четырёхугольника~---
+	Таким образом, получаем, что все 4 угла четыр\"{e}хугольника~---
 	острые, чего быть не может.
@@
 	Множество $S$ содержится в квадрате со стороной $d$ (см. лемму~\ref{lemma:square_container}).
-	Разобьём этот квадрат на $m^2$ маленьких квадратов со стороной $d/m$.
+	Разобь\"{e}м этот квадрат на $m^2$ маленьких квадратов со стороной $d/m$.
 	Тогда по принципу Дирихле либо:
@@
 	\end{equation}
-	или, с учётом того, что $ |S| = (k-1)m^2 + 2$,
+	или, с уч\"{e}том того, что $ |S| = (k-1)m^2 + 2$,
 	\begin{equation}
@@
 Если $n \geq 27636$, то $ (k-1)m^2 + 2 \leq n \leq (k-1)(m+1)^2 + 1$ для $k=10$ и некоторого $m \geq 55$.
-Тогда с учётом следствия~\ref{corollary:common_linear_bound} получаем
+Тогда с уч\"{e}том следствия~\ref{corollary:common_linear_bound} получаем
 \begin{aUtv}
@@
 	равно как и вычисление точных нижних оценок на диаметр для конкретных $n>122$ (в утверждении~\ref{proposition:linear_bound_10}),
-	но лишь до определённого предела, поскольку $\mu_k$ зависит от $\varphi_k$ и $k$
+	но лишь до определ\"{e}нного предела, поскольку $\mu_k$ зависит от $\varphi_k$ и $k$
 	только в составе выражения $\varphi_k^2 (k-1)$,
@@
 	а такая расстановка, при которой все расстояния между точками~--- рациональные числа.
-	Это требование уже имеет отношение к проблеме Улама-Эрдёша (см. гипотезу~\ref{hypot:Ulam}).
+	Это требование уже имеет отношение к проблеме Улама-Эрд\"{e}ша (см. гипотезу~\ref{hypot:Ulam}).
 \end{aZam}
+
+Итак, каждому множеству $M\in\mathfrak{M}_n$ сопоставлены
+два натуральных числа: мощность и диаметр.
+Кемнитц~\cite{kemnitz1988punktmengen} обнаружил и третье,
+которое назвал характеристикой.
+
+\begin{aDef}
+	\cite[гл. 34, п. 3]{Bukhstab-number-theory}
+	Натуральное число называется свободным от квадратов,
+	если оно не делится ни на один квадрат простого числа.
+\end{aDef}
+
+\begin{aDef}
+	\label{def_char_classic}
+	Характеристикой множества $M\in\mathfrak{M}_n$ называется свободное от квадратов
+	число $p$, такое, что площадь любого треугольника $ABC$, где $A,B,C\in M$,
+	соизмерима с $\sqrt{p}$.
+\end{aDef}
+
+Корректность определения~\ref{def_char_classic} показана в~\cite{kemnitz1988punktmengen}.
+
+Пусть $\mathbb{Z}$~--- множество целых чисел;
+ребром множества $M\in\mathfrak{M}$
+будем называть любой отрезок $M_1 M_2$,
+где $M_1, M_2 \in M$.
+
+Опираясь на работы~\cite{our-mkmitu-2016,our-ped-2017},
+можно доказать, что определение~\ref{def_char_classic} равносильно следующему определению.
+\begin{aDef}
+	\label{def_char_lattice}
+	Характеристикой множества $M\in\mathfrak{M}_n$ называется свободное от квадратов
+	число $p$, такое, что множество $M$ может быть размещена на решётке
+\begin{equation}\label{grid_for_Erdosh_system}
+	\left\{\left(
+		\frac{a_i}{2m}
+		;
+		\frac{b_i\sqrt{p}}{2m}
+	\right)\right\},
+\end{equation}
+где $a_i, b_i \in \mathbb{Z}$,
+в качестве $m$ можно взять длину любого ребра $M$.
+\end{aDef}
+
+\begin{aTeorema}
+	Определения~\ref{def_char_classic} и~\ref{def_char_lattice} эквивалентны.
+\end{aTeorema}
+\begin{proof}
+	Пусть $M\in\mathfrak{M}_n$, $M_1,M_2,M_3 \in M$ и не лежат на одной прямой.
+	Пусть сначала $M_1,M_2,M_3$ лежат на решётке~\eqref{grid_for_Erdosh_system}.
+	Тогда по формуле Пика~\cite[теорема 3.1]{polygons-on-lattices} площадь треугольника $M_1 M_2 M_3$
+	соизмерима с $\sqrt{p}$.
+	Пусть теперь площадь треугольника $M_1 M_2 M_3$ соизмерима с $\sqrt{p}$ и $|M_1 - M_2|=m$.
+	Введём систему координат так, что $M_1 = (0,0)$,
+	$M_2 = (|M_1 - M_2|, 0)$,
+	$M_3 = (x_3, y_3)$.
+	Тогда площадь треугольника $M_1 M_2 M_3$ равна $\frac{1}{2} |M_1 - M_2| y_3$,
+	откуда $y_3$ соизмерим с $\sqrt{p}$.
+	Более того, выполнены равенства
+	\begin{gather*}
+		x_3^2 + y_3^2 = |M_1 - M_3|^2
+		\\
+		(m^2 - x_3^2)^2 = |M_2 - M_3|^2
+		.
+	\end{gather*}
+	Вычитая одно из другого, получаем $2mx_3 = a$, где $a$~--- целое число.
+\end{proof}
+
+Характеристика для данного множества $M\in\mathfrak{M}_n$ определяется единственным образом и обозначается $\operatorname{char}M$.
+Заметим, что понятие характеристики можно обобщить и на множества $M\in\mathfrak{M}$,
+содержащиеся в некоторой прямой.
+В таком случае логично полагать $\operatorname{char} M = 0$.
+
+\begin{aTeorema}
+	\label{thm:power_char_exist}
+	Для любого $n\geq 3$ и любого свободного от квадратов числа $p$
+	существует множество $M\in\mathfrak{M}_n$ такое, что $\operatorname{char} M = p$.
+\end{aTeorema}
+
+\begin{proof}
+	Не теряя общности положим, что $n$ нечётно.
+	Пусть
+	\begin{equation*}
+		M = \{M_i, i =0,1,...,n+1\},
+	\end{equation*}
+	где
+	$M_i = (2^{n-i}-p\cdot 2^{i-1},0)$, $i=0,1,...,n$;
+	$M_{n+1} = (0,\sqrt{p \cdot 2^{n+1}})$.
+	Очевидно, для $0\leqslant i \leqslant j \leqslant n$ выполнено
+	$|M_i- M_j|\in\mathbb{N}$.
+	Заметим, что для $i=1,...,n$
+	\begin{multline*}
+		|M_{n+1} - M_i| =
+		\sqrt{p\cdot 2^{n+1} + (2^{n-i}-p\cdot2^{i-1})^2}
+		=
+		\\=
+		\sqrt{4p\cdot 2^{n-1} + (2^{n-i})^2 - 2 \cdot 2^{n-i} \cdot p \cdot 2^{i-1} + (p\cdot 2^{i-1})^2}
+		=
+		\\=
+		\sqrt{4p\cdot 2^{i-1} \cdot 2^{n-i} + (2^{n-i})^2 - 2 \cdot 2^{n-i} \cdot p \cdot 2^{i-1} + (p\cdot 2^{i-1})^2}
+		=
+		\\=
+		\sqrt{(2^{n-i})^2 + 2 \cdot 2^{n-i} \cdot p \cdot 2^{i-1} + (p\cdot 2^{i-1})^2}
+		=
+		%\\=
+		%\sqrt{(2^{n-i})^2 + 2 \cdot 2^{n-i} \cdot p \cdot 2^{i-1} + (p\cdot 2^{i-1})^2}
+		%=
+		2^{n-i} + p\cdot 2^{i-1}
+		\in\mathbb{N}
+		.
+	\end{multline*}
+\end{proof}
+
+В отличие от диаметра и мощности, характеристика в некотором смысле локальна:
+если $S\in\mathfrak{M}_n$, $H \subset S$, $H\in\mathfrak{M}_{p}$, $p<n$,
+то
+\begin{equation*}
+	|H| < |S|, \quad \operatorname{diam} H \leq \operatorname{diam} S, \quad \operatorname{char} H = \operatorname{char} S
+	.
+\end{equation*}
 
@@
 
-Приведём известные оценки на $d(m,n)$
+Привед\"{e}м известные оценки на $d(m,n)$
 (за основу взят список из~\cite{kurz2008bounds}):
@@
 Существуют алгоритмы вычисления $d(m,n)$ на ЭВМ, однако они требуют большого количества машинного времени,
-причём требуемое время быстро растёт с ростом $m$ и $n$.
+прич\"{e}м требуемое время быстро раст\"{e}т с ростом $m$ и $n$.
 Первые несколько значений $d(2,n)$ были получены в~\cite{harborth1998integral},
@@
 составлет $O(d^3)$ (в двумерном случае).
-%В (TODO: ссылка!) отмечено, что важно перепроверять вычисленные значения независимо составленными программами для ЭВМ.
-Приведём некоторые известные значения:
+Привед\"{e}м некоторые известные значения:
 
@@
 \begin{aGipot}
-	Никакое оптимальное множество $M\in\mathfrak{M}$ не лежит на целочисленной решётке.
+	Никакое оптимальное множество $M\in\mathfrak{M}$ не лежит на целочисленной реш\"{e}тке.
 \end{aGipot}
@@
 Положив $p=1$ в доказательстве теоремы~\ref{thm:power_char_exist} и добавив к построеннному множеству
-точку $M_{n+2} = (0, -\sqrt{2^{n+1}})$, где $n$ нечётно,
-получим множество из $n+2$ точек, из которых $n$ лежат на некоторой прямой и 2 вне её.
+точку $M_{n+2} = (0, -\sqrt{2^{n+1}})$, где $n$ неч\"{e}тно,
+получим множество из $n+2$ точек, из которых $n$ лежат на некоторой прямой и 2 вне е\"{e}.
 
@@
 		$~---
-		содержится в объединении окружности с её центром (7~точек);
+		содержится в объединении окружности с е\"{e} центром (7~точек);
 
@@
 
-TODO: спрямить ссылки!
-
 \begin{aDef}
@@
 Вопрос о существовании множеств из $\dot{\mathfrak{M}}_8$
-остаётся открытым.
+оста\"{e}тся открытым.
 
@@
 	$n$--кластером называется множество $M\in\dot{\mathfrak{M}}_n$,
-	содержащееся в целочисленной решётке.
+	содержащееся в целочисленной реш\"{e}тке.
 \end{aDef}
@@
 
-Если $M\in\mathfrak{M}(m,n)$ и все рёбра $M$ имеют чётную длину,
+Если $M\in\mathfrak{M}(m,n)$ и все р\"{e}бра $M$ имеют ч\"{e}тную длину,
 то понятно, что множество $H$, полученное из $M$ сжатием в два раза,
@@
 Выполняя такую операцию, мы рано или поздно получим множество из $\mathfrak{M}(m,n)$ такое,
-что для некоторых $M_1,M_2\in M$ расстояние $|M_1 - M_2|$ нечётно.
-Возникает закономерный вопрос: сколько может быть нечётных расстояний?
+что для некоторых $M_1,M_2\in M$ расстояние $|M_1 - M_2|$ неч\"{e}тно.
+Возникает закономерный вопрос: сколько может быть неч\"{e}тных расстояний?
 
@@
 для существования множества $M\in\mathfrak{M}(m,m+2)$ такого, что для любых
-$M_i,M_j\in M$, $i\neq j$ расстояние $|M_i-M_j|$ нечётно,
+$M_i,M_j\in M$, $i\neq j$ расстояние $|M_i-M_j|$ неч\"{e}тно,
 необходимо и достаточно, чтобы
@@
 В частности, не существует такого множества $M\in\mathfrak{M}_4$,
-что все его рёбра имеют нечётную длину.
+что все его р\"{e}бра имеют неч\"{e}тную длину.
 На основе этого факта в работе~\cite{piepmeyer1996maximum} с использованием результатов теории графов доказано,
-что количество рёбер нечётной длины в множестве $M\in\mathfrak{M}_n$ не превосходит
+что количество р\"{e}бер неч\"{e}тной длины в множестве $M\in\mathfrak{M}_n$ не превосходит
 \begin{equation}
@@
 
-Отдельные оценки количества рёбер нечётной длины для множеств из $\dot{\mathfrak{M}}_n$
+Отдельные оценки количества р\"{e}бер неч\"{e}тной длины для множеств из $\dot{\mathfrak{M}}_n$
 авторам неизвестны.
@@
 
-С исследованием множеств $\mathfrak{M}_n$ тесно связаны ещё две гипотезы о расстояниях на плоскости.
+С исследованием множеств $\mathfrak{M}_n$ тесно связаны ещ\"{e} две гипотезы о расстояниях на плоскости.
 \begin{aGipot}[Улам, \cite{ulam1960collection}]
@@
 В работах~\cite{garibaldi2005erdos,garibaldi2011erdos} исследуется связь множеств $\mathfrak{M}_n$ и
-проблемы Эрдёша о числе различных расстояний~\cite{erdos1946sets}.
+проблемы Эрд\"{e}ша о числе различных расстояний~\cite{erdos1946sets}.
 В работе~\cite{guth2015erdos} эта проблема в своей изначальной постановке была решена и доказана следующая
@@
 \end{aTeorema}
-
-TODO: Примеры множеств - рисунки (если будет время и место).
 
@@
 	\textsl{Авдеев} \textsl{Н. Н.} Некоторые примеры систем
-	Эрдёша /\!/ Современные методы теории краевых задач :
+	Эрд\"{e}ша /\!/ Современные методы теории краевых задач :
 	материалы международной конференции <<Понтрягинские
@@
 	\bibitem{our-mz-rus}
-	\textsl{Авдеев} \textsl{Н. Н.}, \textsl{Семёнов} \textsl{Е. М.}
+	\textsl{Авдеев} \textsl{Н. Н.}, \textsl{Сем\"{e}нов} \textsl{Е. М.}
 	О множествах точек на плоскости с целочисленными
@@
  \begin{center}
- НАЗВАНИЕ СТАТЬИ НА АНГЛИЙСКОМ ЯЗЫКЕ
+ INTEGRAL POINT SETS ON THE PLANE AND IN EUCLIDEAN SPACE
  \end{center}
@@
 
-\Abstract{
-	We considered such a sets of points on the plane
+\Abstract{We considered such a sets of points on the plane
 	that the distance between any two points is integral
@@
 	Minimal diameter of such sets is estimated.
-	The results are generalized to n-dimensional euclidean space.
+	The results are generalized to n-dimensional Euclidean space.
 	Some special cases of such sets are considered

### 2018-10-24 amm2018/Avdeev_IPS_amm2018.utf8.tex
[amm2018] Drafts

\documentclass[draft, 12pt]{article}
\usepackage[utf8]{inputenc} % следующие две строки используются для
\usepackage[english,russian]{babel}   % руссификации AmSLaTeX
\usepackage{amsmath,amsfonts,amssymb,euscript,graphicx,wrapfig,multirow}
\usepackage{dsfont}
\inputencoding{utf8}
\textheight=240mm \textwidth=170mm
\hoffset=-17mm % сдвиг влево
\voffset=-17mm % сдвиг вверх

\begin{document}

\noindent УДК 511.95 + 514.112

\begin{center}
\textbf{Об отыскании целоудалённых множеств специального вида} \\[3mm]
\textbf{Н.~Н.~Авдеев}\\[2mm]
\emph{ВГУ}
\end{center}

Целоудалённым множеством (ЦМ) будем называть такое подмножество $M$ плоскости $\mathbb{R}^2$,
не содержащееся ни в какой прямой,
что расстояние между любыми двумя точками $M$ есть целое число.

Эрдёш [TODO: ссылки!] доказал, что любое ЦМ конечно.
(Доказательство на русском языке можно найти в [TODO: ссылки!].)

Для ЦМ естественным образом определяется диаметр:
$$
	diam(M) = \max_{A,B\in M} |A-B|,
$$
где $|A-B|$~--- обычное расстояние между $A$ и $B$.

Задача об отыскании ЦМ заданного диаметра и мощности первоначально возникла
в ходе определения минимального диаметра,
которым может обладать множество заданной мощности [TODO: ссылка на первые работы].
Затем, в ходе развития классификации ЦМ и выделения множеств полуобщего и общего положения,
встал вопрос от отыскании ЦМ, обладающих специальными свойствами.

В данной работе описывается отыскание <<необычных>>, в некотором смысле, ЦМ.

Выделяют веерные ЦМ, т.е. ЦМ, лежащие на объединении прямой и некоторой точки.
Именно веерные ЦМ имеют наименьший диаметр среди всех ЦМ мощности от 9 до 122 [TODO: Kurz];
однако с помощью веерных ЦМ невозможно построить конструктивную оценку на минимальный диаметр,
которая была бы существенно лучше, чем существующая [TODO:ссылка], основанная на содержащихся в окружности ЦМ.
С помощью инверсии и растяжения плоскости можно из веерного ЦМ сконструировать ЦМ,
содержащееся в объединении окружности с её центром [TODO: ссылка].
Таким образом, относительно хорошо известны ЦМ, лежащие в объединении точки с прямой или окружности с её центром.
Отыскание таких ЦМ выходит за пределы целей данной работы.

На основе веерного ЦМ характеристики 1 легко строится ЦМ, содержащееся в двух перпендикулярных прямых,
при этом на одной из прямых лежит 2 или 3 точки, на другой~--- все остальные
(точка пересечения прямых может как принадлежать ЦМ, так и не принадлежать ему).
Отыскание ЦМ, содержащихся в паре перпендикулярных прямых и распределённых по этим прямым
<<более равномерно>>, де-факто оказывается чисто алгебраической [TODO:ссылка];
наиболее интересным, на взгляд автора, является следующий пример из [TODO:ссылка]:
$$
TODO:
.
$$

Таким образом, нас будут интересовать все ЦМ, кроме ЦМ, содержащихся:
а) В объединении прямой с точкой, не лежащей на ней (веерные ЦМ);
б) В объединении окружности с её центром;
в) В паре перпендикулярных прямых.

Для отыскания ЦМ используется алгоритм, описанный в [TODO: строяк],
со следующими изменениями.
Во-первых, нахождение ЦМ с заданным диаметром и мощностью не приводит к прекращению процесса поиска.
Во-вторых, в модифицированном алгоритме Брона-Кербоша за начальную клику
принимается не произвольная точка, лежащая за пределами прямой, проходящей через точки,
на которых достигается диаметр, а пара таких точек (если расстояние между точкми пары есть число целое).
Это отсекает значительное количество (но не все) веерных ЦМ.

Все найденные таким алгоритмом ЦМ подвергаются проверке на выполнение одного из условий (а)--(в);
если ни одно из условий не выполнено, информация о ЦМ записывается в файл и строится визуализация найденного ЦМ с помощью gnuplot
[TODO: ссылочку бы].

Ниже приводятся иллюстрированные примеры найденных ЦМ из числа наиболее интересных.

Чаще всего из интересующих нас ЦМ встречаются ЦМ, содержащиеся в паре параллельных прямых,
причём на одной из прямых лежит две или три точки.

$ TODO $ (рис. 1).
ЦМ мощности 7, диаметр 17.
На этом ЦМ достигается минимальное значение диаметра для мощности 7,
однако в [TODO: ссылка, fig. ] это ЦМ приведено не было:
примером ЦМ мощности 7 и диаметра 17 послужило другое ЦМ.

%\left\{\left( -79 ; 0\right); \\
%\left( 79 ; 0\right); \\
%\left( -\frac{873}{79} ; \frac{528\sqrt{70}}{79}\right); \\
%\left( \frac{2093}{79} ; -\frac{408\sqrt{70}}{79}\right); \\
%\left( \frac{4838}{79} ; -\frac{138\sqrt{70}}{79}\right); \\
%\left( -\frac{3275}{79} ; -\frac{936\sqrt{70}}{79}\right); \\
%\left( -\frac{1872}{79} ; -\frac{798\sqrt{70}}{79}\right); \\
%\left( \frac{5265}{79} ; -\frac{96\sqrt{70}}{79}\right); \\
%\left( -\frac{2299}{79} ; -\frac{840\sqrt{70}}{79}\right); \\
%\left( \frac{873}{79} ; -\frac{528\sqrt{70}}{79}\right); \\
%\right\}

\begin{figure}[h!]
\begin{center}
%\includegraphics[width=0.75\textwidth]{Ivanov_Petrov_pic_1.eps}\\
\emph{Ðèñ. 1. }
\end{center}
\end{figure}

\begin{wrapfigure}{R}{5.0cm}
\begin{center}
\includegraphics[width=4.5cm, height=4.5cm]{Ivanov_Petrov_pic_1.eps}\\
\emph{ 1.}\end{center}
\end{wrapfigure}

\bigskip\centerline{\bf Литература}

1.	\emph{Инфантьева А.~С., Катышев П.~А.} Тематическая организация Интернет-эгоистории в ранговом распределении Ципфа / А.~С.~Инфантьева, П.~А.~Катышев; Вестник Кемеровского государственного университета. – 2012. – №. 1. (49). С. 160-163.

2.	\emph{Авдеев Н.~Н.} Программа анализа грамотности интер\-нет\--СМИ // Культура общения и её формирование. – 2016. – С. 81-83.

3.	\emph{ Авдеев Н.~Н., Шевелева К.~В.} Анализ орфографической грамотности
математических образовательных ресурсов в сети «Интернет» // Некоторые вопросы анализа, алгебры, геометрии и математического
образования – Воронеж: Издательско-полиграфический центр «Научная книга», 2017. – Вып. 7, Часть I – 244 с.

4.	\emph{Авдеев Н.~Н., Шевелева К.~В.} Математическое моделирование распространённых орфографических ошибок с помощью регулярных выражений на материале математических образовательных ресурсов в сети «Интернет» // Материалы международной конференции  «Воронежская зимняя математическая школа С. Г. Крейна – 2018» [Текст] / под ред. В. А. Костина. – Воронеж : Издательско-полиграфический центр «Научная книга», 2018. –  С.110-112

5.	\emph{Sheveleva K.~V.} Elementary generalization of Zipf's law for spelling mistakes (based on internet resources devoted to mathematics)  / K.~V.~Sheveleva// «Вестник современных исследований»  Выпуск № 8-3 (23) (август, 2018). --  Научный центр «Орка»,  Омск, 2018 -- C.362--363

6.	\emph{Маслов В. П., Маслова Т. В.} О законе Ципфа и ранговых распределениях в лингвистике и семиотике //Математические заметки. – 2006. – Т. 80. – №. 5. – С. 718-732.

7.  \emph{Ляпунов С.~М., Алымов А.~С.}  Методология анализа текста с помощью составления частотных словарей по частям речи //Научный альманах. – 2017. – №. 10-3. – С. 203-206.

8.	\emph{Кочеткова Н.~А., Клышинский Э.~С., Ермаков П.~Д.} Подчиняются ли составные конструкции закону Ципфа? //Системный администратор. – 2016. – №. 11. – С. 89-95.

9.  \emph{Чичкарёв Е.~А.} Компьютерная математика с Maxima: Руководство для школьников и студентов //М.: ALT Linux. – 2012. – Т. 384.

10. \emph{Алексеев Е.~Р., Чеснокова О.~В.} Введение в Octave для инженеров и математиков //М.: ALT Linux. – 2012.

11. \emph{Алексеев Е.~Р.} Использование свободных программ в научных исследованиях // Прикладная информатика. – 2009. – №. 6.

{\small Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-10125.}

{\bf Авдеев Николай Николаевич},
бакалавр математики, магистрант кафедры функционального анализа и операторных уравнений математического факультета
ФГБОУ ВО <<Воронежский государственный университет>>, г. Воронеж.

E-mail: avdeev@math.vsu.ru, nickkolok@mail.ru

Научный руководитель: Семенов Евгений Михайлович.

\end{document}

### 2018-10-24 vmmsh-2018/AvdeevNN_09.cp1251.tex
[vmmsh] Final edits

@@
 
- \Author{Авдеев Н.Н., Семeнов Е.М.}
+ \Author{Н.Н. Авдеев, Е.М. Семeнов}
 
@@
 
-\begin{proof}[Доказательство теоремы \ref{thmErdos}]
+\begin{proof}%[Доказательство теоремы \ref{thmErdos}]
 	Предположим противное,
@@
 	и точки $M_1$, $M_2$, $M_3$ не лежат на одной прямой.
+	Обозначим $|M_1 - M_2| = n$ и $|M_1 - M_3| = m$.
 	Применяя лемму~\ref{thm:quadCurveFamily}
@@
 	Разделим обе части неравенства на положительное число $m(k-1)$:
-	\begin{equation}
+	\begin{equation*}
 		\frac{d}{(k-1)m^2}
@@
 		\frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2 (k-1)}
-	\end{equation}
+	\end{equation*}
 	Положим
-	\begin{equation}
+	\begin{equation*}
 		\mu_k = \frac{ \sqrt{ 64 \varphi_k^2 (k-1) + 1} -1 }{16 \varphi_k^2 (k-1)},
-	\end{equation}
+	\end{equation*}
 	тогда
-	\begin{equation}
+	\begin{equation*}
 		\frac{d}{(k-1)m^2}
@@
 		\mu_k
-	\end{equation}
+	\end{equation*}
 	откуда незамедлительно
-	\begin{equation}
+	\begin{equation*}
 		d > \mu_k \cdot (|S|-2)
 		.
-	\end{equation}
+	\end{equation*}
 \end{proof}
@@
 	Нахождение $\varphi_k$ для $k > 30$ может позволить улучшить оценку
-	в утверждениях~\ref{proposition:linear_bound_29} и~\ref{proposition:linear_bound_10},
-	равно как и вычисление точных нижних оценок на диаметр для конкретных $n>122$ (в утверждении~\ref{proposition:linear_bound_10}),
-	но лишь до определ\"{e}нного предела, поскольку $\mu_k$ зависит от $\varphi_k$ и $k$
+	в утверждениях~\ref{proposition:linear_bound_29} и~\ref{proposition:linear_bound_10}.
+	Вычисление точных нижних оценок на диаметр для конкретных $n>122$~--- в утверждении~\ref{proposition:linear_bound_10}.
+	Однако оценки таким образом можно улучшить
+	лишь до определ\"{e}нного предела, поскольку $\mu_k$ зависит от $\varphi_k$ и $k$
 	только в составе выражения $\varphi_k^2 (k-1)$,
-	которое ограничено в силу теоремы~\ref{thm:varphi_k_bounds}.
+	ограниченого в силу теоремы~\ref{thm:varphi_k_bounds}.
 \end{aZam}
@@
 	Характеристикой множества $M\in\mathfrak{M}_n$ называется свободное от квадратов
-	число $p$, такое, что множество $M$ может быть размещена на решётке
+	число $p$, такое, что множество $M$ может быть размещено на реш\"{e}тке
 \begin{equation}\label{grid_for_Erdosh_system}
@@
 	Пусть $M\in\mathfrak{M}_n$, $M_1,M_2,M_3 \in M$ и не лежат на одной прямой.
-	Пусть сначала $M_1,M_2,M_3$ лежат на решётке~\eqref{grid_for_Erdosh_system}.
+	Пусть сначала $M_1,M_2,M_3$ лежат на реш\"{e}тке~\eqref{grid_for_Erdosh_system}.
 	Тогда по формуле Пика~\cite[теорема 3.1]{polygons-on-lattices} площадь треугольника $M_1 M_2 M_3$
@@
 	Пусть теперь площадь треугольника $M_1 M_2 M_3$ соизмерима с $\sqrt{p}$ и $|M_1 - M_2|=m$.
-	Введём систему координат так, что $M_1 = (0,0)$,
+	Введ\"{e}м систему координат так, что $M_1 = (0,0)$,
 	$M_2 = (|M_1 - M_2|, 0)$,
@@
 \begin{proof}
-	Не теряя общности положим, что $n$ нечётно.
+	Не теряя общности положим, что $n$ неч\"{e}тно.
 	Пусть
@@
 
-Более того, конструкцию $\mathfrak{M}_n$ можно распространить и на $\mathbb{Z}^m_n$~\cite{Kohnert2006IntegralPS}.
+Конструкцию $\mathfrak{M}_n$ можно распространить и на более широкие классы пространств~\cite{Kohnert2006IntegralPS}.
 %Более того, аналогичные множества были построены в $\mathbb{Z}^m_n$~\cite{Kohnert2006IntegralPS}.
@@
 	& d(m, m^2 + m) \leq 17 \qquad \mbox{\cite{kurz2008bounds}} \\
-	& 3 \leq d(m, n) \leq 4 \mbox{~для~} m + 2 \leq n \leq 2m \mbox{~и~} d(m, 2m) = 4 ~ \mbox{\cite{piepmeyer1988raumliche,harborth1991point}}\\
+	& 3 \leq d(m, n) \leq 4 \mbox{~для~} m + 2 \leq n \leq 2m \mbox{~и~} d(m, 2m) = 4 ~ \mbox{\cite{piepmeyer1988raumliche,harborth1991point}}
 \end{align}
@@
 
-Для многомерного случая были вычислены~\cite{piepmeyer1988raumliche,harborth1998integral,kurz2005characteristic,kurz2006konstruktion,kurz2008bounds} следующие значения.
+Для многомерного случая были вычислены в~\cite{piepmeyer1988raumliche,harborth1998integral,kurz2005characteristic,kurz2006konstruktion,kurz2008bounds} следующие значения.
 \begin{aUtv}~
@@
 \end{aGipot}
-В пользу этой гипотезы говорят реузльтаты вычислений для $m=2$, $3 \leq n \leq 43$.
+В пользу этой гипотезы говорят результаты вычислений для $m=2$, $3 \leq n \leq 43$.
 
@@
 
-Солимоси построил~\cite{solymosi2003note} для любого $n\geq 3$ конструкцию множества $M\in\mathfrak{M}_n$
-такого, что для некоторых $M_1, M_2 \in M$ выполнено $|M_1 - M_2| = 2$.
+Солимоси построил~\cite{solymosi2003note} для любого $n\geq 3$ множество $M\in\mathfrak{M}_n$
+такое, что для некоторых $M_1, M_2 \in M$ выполнено $|M_1 - M_2| = 2$.
 Полученное им множество также является веерным.
@@
 	Тогда все точки множества $M$ лежат на $cr(M_1,M_2)$.
-	$cr(M_1,M_2)$ есть объединение двух прямых, на каждой из которых может быть не более 2 точек в силу того, что
+	Крест есть объединение двух прямых, на каждой из которых может быть не более 2 точек в силу того, что
 	$M\in\overline{\mathfrak{M}}_n$,
@@
 	\textsl{Kohnert} \textsl{A.}, \textsl{Kurz} \textsl{S.} Integral
-	point sets over $Z^m_n$
+	point sets over $\mathbb{Z}^m_n$
 	/\!/ Discrete Applied Mathematics. --- 2006. ---
@@
 
-\Abstract{We considered such a sets of points on the plane
+\Abstract{We considered such sets of points on the plane
 	that the distance between any two points is integral

### 2018-10-24 drafts/minimeter1.js
[code] Script to find IPS with minimeter 1

@@ -1,0 +1,27 @@
+
+
+var timeBefore = Date.now();
+
+var triades=[];
+
+for(var a = 2; a < 40000; a++){
+	console.error(a);
+	var a2a = a*(a+1);
+	for(var t = 1; 4*(t+1) <= a2a; t++){
+		var max_s = Math.sqrt(t*t+a2a);
+		var s = Math.floor(max_s);
+//		for(var s = t + 2; s <= max_s; s +=2){
+			if( a2a == (s-t)*(s+t) ) {
+//				console.log(a,s,t);
+				triades.push([a,s,t]);
+			}
+//		}
+	}
+}
+
+triades.sort((x,y)=>(x[2]-y[2]));
+//console.log(triades);
+
+triades.map((arr)=>console.log(arr));
+
+console.error((Date.now() - timeBefore)/1000, 'c');

### 2018-10-25 amm2018/Avdeev_IPS_amm2018.utf8.tex
[amm2018] Asterisk notation
[amm2018] Illustrations

@@
-\documentclass[draft, 12pt]{article}
+%\documentclass[draft, 12pt]{article}
+\documentclass[12pt]{article}
 \usepackage[utf8]{inputenc} % следующие две строки используются для
@@
 Ниже приводятся иллюстрированные примеры найденных ЦМ из числа наиболее интересных.
+Для облегчения восприятия иногда применяется запись вида
+$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
+означающая, что каждую абсциссу нужно умножить на $1/q$,
+а каждую ординату~--- умножить на $\sqrt{p}/q$, т.е.
+$$
+	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
+	=
+	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},\frac{y_n\sqrt{p}}{q}\right)  \right\}
+	.
+$$
 
@@
 
-$ TODO $ (рис. 1).
-ЦМ мощности 7, диаметр 17.
+\begin{itemize}
+\setlength{\itemsep}{-1mm}
+
+\item
+$\sqrt{35}/{34} * \{ (\pm 289, 0),
+%(289 , 0),
+(88 , -39),
+(-144 , -15),
+(-175 , 48),
+(-57 , -24),
+(57 , 24)\} $
+(рис.~\ref{Avdeev_7_17_1538484835387.eps}).
+%ЦМ мощности 7, диаметр 17.
 На этом ЦМ достигается минимальное значение диаметра для мощности 7,
@@
 
-%\left\{\left( -79 ; 0\right); \\
-%\left( 79 ; 0\right); \\
-%\left( -\frac{873}{79} ; \frac{528\sqrt{70}}{79}\right); \\
-%\left( \frac{2093}{79} ; -\frac{408\sqrt{70}}{79}\right); \\
-%\left( \frac{4838}{79} ; -\frac{138\sqrt{70}}{79}\right); \\
-%\left( -\frac{3275}{79} ; -\frac{936\sqrt{70}}{79}\right); \\
-%\left( -\frac{1872}{79} ; -\frac{798\sqrt{70}}{79}\right); \\
-%\left( \frac{5265}{79} ; -\frac{96\sqrt{70}}{79}\right); \\
-%\left( -\frac{2299}{79} ; -\frac{840\sqrt{70}}{79}\right); \\
-%\left( \frac{873}{79} ; -\frac{528\sqrt{70}}{79}\right); \\
-%\right\}
+\item
+$\sqrt{3}/2 * \{
+( \pm 56 , 0),
+( 14 , 0),
+( -34 , 0),
+( -10 , 24),
+( -21 , 35),
+( 35 , -21)
+\}
+$
+(рис.~\ref{Avdeev_7_56_1538484851696.eps}).
+ЦМ лежит на двух пересекающихся под острым углом прямых и имеет ось симметрии
+~--- биссектрису угла между прямыми.
 
-\begin{figure}[h!]
-\begin{center}
-%\includegraphics[width=0.75\textwidth]{Ivanov_Petrov_pic_1.eps}\\
-\emph{Ðèñ. 1. }
-\end{center}
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_7_17_1538484835387.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_7_56_1538484851696.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 17}\label{Avdeev_7_17_1538484835387.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 56}\label{Avdeev_7_56_1538484851696.eps}}
 \end{figure}
 
-\begin{wrapfigure}{R}{5.0cm}
-\begin{center}
-\includegraphics[width=4.5cm, height=4.5cm]{Ivanov_Petrov_pic_1.eps}\\
-\emph{ 1.}\end{center}
-\end{wrapfigure}
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_7_60_1538484853808.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_7_60_1538484854748.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_7_60_1538484853808.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_7_60_1538484854748.eps}}
+\end{figure}
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_7_64_1538484861024.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_7_75_1538484880810.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_7_64_1538484861024.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_7_75_1538484880810.eps}}
+\end{figure}
+
+\begin{figure}[htbp]
+%	\includegraphics[width=.48\linewidth]{}
+	\hfill
+%	\includegraphics[width=.48\linewidth]{}
+	\\
+	\parbox{.48\linewidth}{\caption{}\label{}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{}\label{}}
+\end{figure}
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_8_123_1538485378899.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_8_123_1538485378698.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_123_1538485378899.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_123_1538485378698.eps}}
+\end{figure}
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_8_146_1538486493354.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_8_162_1538488466583.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_146_1538486493354.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_162_1538488466583.eps}}
+\end{figure}
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_8_168_1538490487450.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_8_171_1538491142044.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_168_1538490487450.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_171_1538491142044.eps}}
+\end{figure}
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_9_123_1538485102263.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_9_214_1538495187150.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_9_123_1538485102263.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{}\label{Avdeev_9_214_1538495187150.eps}}
+\end{figure}
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_10_158_1538485325776.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_10_222_1538488624864.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{}\label{F1}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{}\label{F2}}
+\end{figure}
+
+\end{itemize}
 

### 2018-10-25 amm2018/Avdeev_IPS_amm2018.utf8.tex
[amm2018] References - WIP

@@
 
-Эрдёш [TODO: ссылки!] доказал, что любое ЦМ конечно.
-(Доказательство на русском языке можно найти в [TODO: ссылки!].)
+Эрдёш [1,2] доказал, что любое ЦМ конечно.
+(Доказательство на русском языке можно найти в [3].)
 
@@
 в ходе определения минимального диаметра,
-которым может обладать множество заданной мощности [TODO: ссылка на первые работы].
+которым может обладать множество заданной мощности [4].
 Затем, в ходе развития классификации ЦМ и выделения множеств полуобщего и общего положения,
@@
 Выделяют веерные ЦМ, т.е. ЦМ, лежащие на объединении прямой и некоторой точки.
-Именно веерные ЦМ имеют наименьший диаметр среди всех ЦМ мощности от 9 до 122 [TODO: Kurz];
+Именно веерные ЦМ имеют наименьший диаметр среди всех ЦМ мощности от 9 до 122 [5];
 однако с помощью веерных ЦМ невозможно построить конструктивную оценку на минимальный диаметр,
-которая была бы существенно лучше, чем существующая [TODO:ссылка], основанная на содержащихся в окружности ЦМ.
-С помощью инверсии и растяжения плоскости можно из веерного ЦМ сконструировать ЦМ,
-содержащееся в объединении окружности с её центром [TODO: ссылка].
+которая была бы существенно лучше, чем существующая [4], основанная на содержащихся в окружности ЦМ.
 Таким образом, относительно хорошо известны ЦМ, лежащие в объединении точки с прямой или окружности с её центром.
@@
 Отыскание ЦМ, содержащихся в паре перпендикулярных прямых и распределённых по этим прямым
-<<более равномерно>>, де-факто оказывается чисто алгебраической [TODO:ссылка];
-наиболее интересным, на взгляд автора, является следующий пример из [TODO:ссылка]:
+<<более равномерно>>, де-факто оказывается чисто алгебраической [6];
+наиболее интересным, на взгляд автора, является следующий пример из [6]:
 $$
-TODO:
+\{
+(0,0);
+(0,\pm 9282000),
+(0,\pm ),
+(0,\pm ),
+(0,),
+(0,),
+(0,),
+(0,),
+(0,),
+(0,),
+\}
 .
@@
 На этом ЦМ достигается минимальное значение диаметра для мощности 7,
-однако в [TODO: ссылка, fig. ] это ЦМ приведено не было:
+однако в [4, fig. 1] это ЦМ приведено не было:
 примером ЦМ мощности 7 и диаметра 17 послужило другое ЦМ.
@@
 
-1.	\emph{Инфантьева А.~С., Катышев П.~А.} Тематическая организация Интернет-эгоистории в ранговом распределении Ципфа / А.~С.~Инфантьева, П.~А.~Катышев; Вестник Кемеровского государственного университета.
-– 2012.
-– №.
-1.
-(49).
-С.
-160-163.
+1.	\emph{Anning~N. H., Erd\"{o}s~P.}
+	Integral distances // Bulletin of the American Mathematical
+	Society. --- 1945. --- Т. 51, No 8. --- С. 598---600.
 
-2.	\emph{Авдеев Н.~Н.} Программа анализа грамотности интер\-нет\--СМИ // Культура общения и её формирование. – 2016. – С. 81-83.
+2.	\emph{Erd\"{o}s~P.} Integral distances // Bulletin of the
+	American Mathematical Society. --- 1945. --- Т. 51, No 12. --- С. 996.
 
-3.	\emph{ Авдеев Н.~Н., Шевелева К.~В.} Анализ орфографической грамотности
-математических образовательных ресурсов в сети «Интернет» // Некоторые вопросы анализа, алгебры, геометрии и математического
-образования – Воронеж: Издательско-полиграфический центр «Научная книга», 2017. – Вып. 7, Часть I – 244 с.
+3.	\emph{Авдеев Н. Н., Семенов Е. М.}
+	Множества точек с целочисленными расстояниями на плоскости и в евклидовом пространстве
+	//
+	Сборник материалов XIV Владикавказской молодежной математической Школы-конференции.
+	--- 2018 --- (в печати)
 
-4.	\emph{Авдеев Н.~Н., Шевелева К.~В.} Математическое моделирование распространённых орфографических ошибок с помощью регулярных выражений на материале математических образовательных ресурсов в сети «Интернет» // Материалы международной конференции  «Воронежская зимняя математическая школа С.
-Г.
-Крейна – 2018» [Текст] / под ред.
-В.
-А.
-Костина.
-– Воронеж : Издательско-полиграфический центр «Научная книга», 2018.
-–  С.110-112
+4.	\emph{Harborth~H.,
+	Kemnitz~A.,
+	M\"{o}ller~M.}
+	An upper bound for the minimum diameter of integral point sets
+	/\!/ Discrete \& Computational Geometry. --- 1993. --- Т. 9, No 4.
+	--- С. 427---432.
 
-5.	\emph{Sheveleva K.~V.} Elementary generalization of Zipf's law for spelling mistakes (based on internet resources devoted to mathematics)  / K.~V.~Sheveleva// «Вестник современных исследований»  Выпуск № 8-3 (23) (август, 2018).
---  Научный центр «Орка»,  Омск, 2018 -- C.362--363
+5.	\emph{Kurz~S., Wassermann~A.}
+	On the minimum diameter of plane integral point sets /\!/ arXiv
+	preprint arXiv:0804.1307. --- 2008.
 
-6.	\emph{Маслов В. П., Маслова Т. В.} О законе Ципфа и ранговых распределениях в лингвистике и семиотике //Математические заметки. – 2006. – Т. 80. – №. 5. – С. 718-732.
-
-7.  \emph{Ляпунов С.~М., Алымов А.~С.}  Методология анализа текста с помощью составления частотных словарей по частям речи //Научный альманах. – 2017. – №. 10-3. – С. 203-206.
-
-8.	\emph{Кочеткова Н.~А., Клышинский Э.~С., Ермаков П.~Д.} Подчиняются ли составные конструкции закону Ципфа? //Системный администратор. – 2016. – №. 11. – С. 89-95.
-
-9.  \emph{Чичкарёв Е.~А.} Компьютерная математика с Maxima: Руководство для школьников и студентов //М.: ALT Linux. – 2012. – Т. 384.
-
-10. \emph{Алексеев Е.~Р., Чеснокова О.~В.} Введение в Octave для инженеров и математиков //М.: ALT Linux. – 2012.
-
-11. \emph{Алексеев Е.~Р.} Использование свободных программ в научных исследованиях // Прикладная информатика. – 2009. – №. 6.
+6.	\emph{Guy~R.} Unsolved problems in number theory. Т.
+	1. --- Springer Science \& Business Media, 2013.
 

### 2018-10-25 amm2018/Avdeev_IPS_amm2018.utf8.tex
[amm2018] Example of 17 points on a cross
[amm2018] Cosmetics

@@
 Отыскание ЦМ, содержащихся в паре перпендикулярных прямых и распределённых по этим прямым
-<<более равномерно>>, де-факто оказывается чисто алгебраической [6];
+<<более равномерно>>, де-факто оказывается чисто алгебраической задачей [6];
 наиболее интересным, на взгляд автора, является следующий пример из [6]:
@@
 (0,\pm 9282000),
-(0,\pm ),
-(0,\pm ),
-(0,),
-(0,),
-(0,),
-(0,),
-(0,),
-(0,),
+(0,\pm 26822600),
+(0,\pm 60386040);
+$$
+$$
+(\pm 7422030,0),
+(\pm 8947575, 0),
+(\pm 22276800,0),
+(\pm 44142336,0)
 \}
@@
 Таким образом, нас будут интересовать все ЦМ, кроме ЦМ, содержащихся:
-а) В объединении прямой с точкой, не лежащей на ней (веерные ЦМ);
-б) В объединении окружности с её центром;
-в) В паре перпендикулярных прямых.
+а) в объединении прямой с точкой, не лежащей на ней (веерные ЦМ);
+б) в объединении окружности с её центром;
+в) в паре перпендикулярных прямых.
 

### 2019-05-08 amm2018/Avdeev_IPS_amm2018.utf8.tex
Thesis for AMM2018 - full version

@@
 \begin{center}
-\textbf{Об отыскании целоудалённых множеств специального вида} \\[3mm]
+\textbf{ОБ ОТЫСКАНИИ ЦЕЛОУДАЛЁННЫХ МНОЖЕСТВ СПЕЦИАЛЬНОГО ВИДА} \\[3mm]
 \textbf{Н.~Н.~Авдеев}\\[2mm]
@@
 
-Для отыскания ЦМ используется алгоритм, описанный в [TODO: строяк],
+Для отыскания ЦМ используется алгоритм, описанный в [7],
 со следующими изменениями.
@@
 если ни одно из условий не выполнено, информация о ЦМ записывается в файл и строится визуализация найденного ЦМ с помощью gnuplot
-[TODO: ссылочку бы].
+[8,9].
 
 Ниже приводятся иллюстрированные примеры найденных ЦМ из числа наиболее интересных.
-Для облегчения восприятия иногда применяется запись вида
+Для облегчения восприятия применяется запись вида
 $\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
@@
 
+\item
+$
+\sqrt{39}/8 *
+\{
+( \pm240 , 0),
+( -91 , -15),
+( -35 , -55),
+( -140 , 20),
+( -112 , 0),
+( -16 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_60_1538484853808.eps}).
+Ещё один пример с двумя пересекающимися прямыми.
+
+\item
+$
+\sqrt{7}/2 *
+\{
+( \pm60 , 0),
+( \pm15 , -25),
+( \pm33 , -9),
+( 0 , -20)
+\}
+$
+(рис.~\ref{Avdeev_7_60_1538484854748.eps}).
+Прямые не перпендикулярны, как может показаться.
+
 \begin{figure}[htbp]
@@
 	\\
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_7_60_1538484853808.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 60}\label{Avdeev_7_60_1538484853808.eps}}
 	\hfill
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_7_60_1538484854748.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 60}\label{Avdeev_7_60_1538484854748.eps}}
 \end{figure}
+
+\item
+$
+\sqrt{7}/2 *
+\{
+( \pm64 , 0),
+( \pm45 , 15),
+( \pm19 , 15),
+( 0 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_64_1538484861024.eps}).
+
+\item
+$
+\sqrt{3}/2 *
+\{
+( \pm75 , 0),
+( 29 , -24),
+( -35 , 40),
+( 40 , -35),
+( 5 , 0),
+( 53 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_75_1538484880810.eps}).
 
@@
 	\\
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_7_64_1538484861024.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 64}\label{Avdeev_7_64_1538484861024.eps}}
 	\hfill
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_7_75_1538484880810.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 75}\label{Avdeev_7_75_1538484880810.eps}}
 \end{figure}
 
+\item
+$
+\sqrt{15}/2 *
+\{
+( \pm77 , 0),
+( 0 , 13),
+( \pm28 , -15),
+( \pm11 , 24)
+\}
+$
+(рис.~\ref{Avdeev_7_77_1538484885363.eps}).
+
+\item
+$
+\sqrt{15}/4 *
+\{
+( \pm156 , 0),
+( -9 , -45),
+( -145 , 11),
+( 112 , -12),
+( 128 , -28),
+( 68 , 32)
+\}
+$
+(рис.~\ref{Avdeev_7_78_1538484891588.eps}).
+
 \begin{figure}[htbp]
-%	\includegraphics[width=.48\linewidth]{}
+	\includegraphics[width=.48\linewidth]{Avdeev_7_77_1538484885363.eps}
 	\hfill
-%	\includegraphics[width=.48\linewidth]{}
+	\includegraphics[width=.48\linewidth]{Avdeev_7_78_1538484891588.eps}
 	\\
-	\parbox{.48\linewidth}{\caption{}\label{}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 77}\label{Avdeev_7_77_1538484885363.eps}}
 	\hfill
-	\parbox{.48\linewidth}{\caption{}\label{}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 78}\label{Avdeev_7_78_1538484891588.eps}}
 \end{figure}
+
+\item
+$
+\sqrt{455}/2 *
+\{
+( \pm87 , 0),
+( \pm64 , -3),
+( 0 , -3),
+( \pm23 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_87_1538484975645.eps}).
+
+\item
+$
+\sqrt{15}/4 *
+\{
+( \pm180 , 0),
+( 135 , 45),
+( -33 , 21),
+( 44 , 32),
+( 76 , 0),
+( -12 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_90_1538485007759.eps}).
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_7_87_1538484975645.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_7_90_1538485007759.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 87}\label{Avdeev_7_87_1538484975645.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 90}\label{Avdeev_7_90_1538485007759.eps}}
+\end{figure}
+
+\item
+$
+\sqrt{35}/1 *
+\{
+( \pm27 , 8),
+( 0 , 8),
+( \pm46 , 0),
+( \pm19 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_92_1538485051452.eps}).
+
+\item
+$
+\sqrt{15}/2 *
+\{
+( -95 , 0),
+( 95 , 0),
+( -84 , -11),
+( -52 , 21),
+( -7 , -24),
+( -73 , 0),
+( -31 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_95_1538485087074.eps}).
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_7_92_1538485051452.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_7_95_1538485087074.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 92}\label{Avdeev_7_92_1538485051452.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 95}\label{Avdeev_7_95_1538485087074.eps}}
+\end{figure}
+
+\item
+$
+\sqrt{15}/8 *
+\{
+( -384 , 0),
+( 384 , 0),
+( -100 , -28),
+( 155 , 77),
+( -253 , -91),
+( -32 , 0),
+( 96 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_96_1538485207044.eps}).
+
+\item
+$
+\sqrt{91}/2 *
+\{
+( \pm99 , 0),
+( 0 , 17),
+( \pm75 , -8),
+( \pm24 , 9)
+\}
+$
+(рис.~\ref{Avdeev_7_99_1538485281307.eps}).
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_7_96_1538485207044.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_7_99_1538485281307.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 96}\label{Avdeev_7_96_1538485207044.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 99}\label{Avdeev_7_99_1538485281307.eps}}
+\end{figure}
+
+\item
+$
+\sqrt{91}/2 *
+\{
+( \pm99 , 0),
+( 0 , 17),
+( \pm24 , 9),
+( 75 , -8),
+( 51 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_99_1538485281335.eps}).
+
+\item
+$
+\sqrt{91}/2 *
+\{
+( \pm99 , 0),
+( 0 , 17),
+( \pm24 , 9),
+( \pm51 , 0)
+\}
+$
+(рис.~\ref{Avdeev_7_99_1538485281337.eps}).
+
+\begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{Avdeev_7_99_1538485281335.eps}
+	\hfill
+	\includegraphics[width=.48\linewidth]{Avdeev_7_99_1538485281337.eps}
+	\\
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 99}\label{Avdeev_7_99_1538485281335.eps}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 7, диаметр 99}\label{Avdeev_7_99_1538485281337.eps}}
+\end{figure}
+
+\item
+$
+\sqrt{3003}/82 *
+\{
+( \pm5043 , 0),
+( 3640 , 23),
+( 1200 , 63),
+( 2277 , -120),
+( -163 , -80),
+( 2603 , 40),
+$ \\ $
+( -2603 , -40)
+\}
+$
+(рис.~\ref{Avdeev_8_123_1538485378899.eps}).
+ЦМ лежит на двух прямых, по 4 точки на каждой.
+
+\item
+$
+\sqrt{3003}/82 *
+\{
+( \pm5043 , 0),
+( -1240 , -103),
+( 3640 , -23),
+( 1200 , -63),
+( -163 , 80),
+( 163 , -80),
+$ \\ $
+( -2603 , 40)
+\}
+$
+(рис.~\ref{Avdeev_8_123_1538485378698.eps}).
+ЦМ лежит на двух прямых, 3 точки на одной и 5 на другой.
 
@@
 	\\
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_123_1538485378899.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 8, диаметр 123}\label{Avdeev_8_123_1538485378899.eps}}
 	\hfill
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_123_1538485378698.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 8, диаметр 123}\label{Avdeev_8_123_1538485378698.eps}}
 \end{figure}
+
+\item
+$
+\sqrt{210}/73 *
+\{
+( \pm5329 , 0),
+( -3783 , -272),
+( 3922 , -42),
+( 1309 , -120),
+( -2376 , -230),
+$ \\ $
+( 237 , -152),
+( 3515 , 264)
+\}
+$
+(рис.~\ref{Avdeev_8_146_1538486493354.eps}).
+
+\item
+$
+\sqrt{2}/27 *
+\{
+( \pm2187 , 0),
+( 1309 , 1520),
+( 899 , -560),
+( 623 , -680),
+( -665 , -1240),
+( 1428 , -330),
+$ \\ $
+( 94 , -910)
+\}
+$
+(рис.~\ref{Avdeev_8_162_1538488466583.eps}).
 
@@
 	\\
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_146_1538486493354.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 8, диаметр 146}\label{Avdeev_8_146_1538486493354.eps}}
 	\hfill
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_162_1538488466583.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 8, диаметр 162}\label{Avdeev_8_162_1538488466583.eps}}
 \end{figure}
+
+\item
+$
+\sqrt{1}/1 *
+\{
+( \pm84 , 0),
+( \pm24 , 45),
+( 0 , 63),
+( 0 , 13),
+( 0 , 35),
+( 0 , 0)
+\}
+$
+(рис.~\ref{Avdeev_8_168_1538490487450.eps}).
+ЦМ лежит на целочисленной решётке.
+
+\item
+$
+\sqrt{231}/2 *
+\{
+( \pm171 , 0),
+( \pm96 , 9),
+( \pm125 , -8),
+( \pm21 , 0)
+\}
+$
+(рис.~\ref{Avdeev_8_171_1538491142044.eps}).
 
@@
 	\\
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_168_1538490487450.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 8, диаметр 168}\label{Avdeev_8_168_1538490487450.eps}}
 	\hfill
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_8_171_1538491142044.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 8, диаметр 171}\label{Avdeev_8_171_1538491142044.eps}}
 \end{figure}
+
+\item
+$
+\sqrt{3003}/82 *
+\{
+( \pm5043 , 0),
+( -163 , -80),
+( 163 , 80),
+( 2603 , 40),
+( -2603 , -40),
+$ \\ $
+( 1240 , -103),
+( -1200 , -63),
+( -3640 , -23)
+\}
+$
+(рис.~\ref{Avdeev_9_123_1538485102263.eps}).
+
+\item
+$
+\sqrt{210}/107 *
+\{
+( \pm11449 , 0),
+( -3795 , -296),
+( -7812 , -374),
+( 4033 , -144),
+( -9975 , -416),
+$ \\ $
+( 1870 , -186),
+( -2147 , -264),
+( 2147 , 264)
+\}
+$
+(рис.~\ref{Avdeev_9_214_1538495187150.eps}).
 
@@
 	\\
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_9_123_1538485102263.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 9, диаметр 123}\label{Avdeev_9_123_1538485102263.eps}}
 	\hfill
-	\parbox{.48\linewidth}{\caption{}\label{Avdeev_9_214_1538495187150.eps}}
+	\parbox{.48\linewidth}{\caption{ЦМ мощности 9, диаметр 214}\label{Avdeev_9_214_1538495187150.eps}}
 \end{figure}
 
-\begin{figure}[htbp]
-	\includegraphics[width=.48\linewidth]{Avdeev_10_158_1538485325776.eps}
-	\hfill
-	\includegraphics[width=.48\linewidth]{Avdeev_10_222_1538488624864.eps}
-	\\
-	\parbox{.48\linewidth}{\caption{}\label{F1}}
-	\hfill
-	\parbox{.48\linewidth}{\caption{}\label{F2}}
-\end{figure}
+%\begin{figure}[htbp]
+%	\includegraphics[width=.48\linewidth]{Avdeev_10_158_1538485325776.eps}
+%	\hfill
+%	\includegraphics[width=.48\linewidth]{Avdeev_10_222_1538488624864.eps}
+%	\\
+%	\parbox{.48\linewidth}{\caption{ЦМ мощности 10, диаметр 158}\label{Avdeev_10_158_1538485325776.eps}}
+%	\hfill
+%	\parbox{.48\linewidth}{\caption{ЦМ мощности 10, диаметр 222}\label{Avdeev_10_222_1538488624864.eps}}
+%\end{figure}
 
 \end{itemize}
+
+{\small Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-10125.}
 
@@
 
-{\small Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-10125.}
+7.	\emph{Авдеев~Н. Н.} Об отыскании множеств точек на
+	плоскости с целочисленными расстояниями /\!/ Математическое
+	и компьютерное моделирование, информационные технологии
+	управления: сб. тр. Школы для студентов, аспирантов и
+	молодых ученых <<МКМИТУ-2016>>. --- 2016. --- С. 15---19.
+
+8.	\emph{Балдин Е.} Gnuplot. Графики заказывали? //Системный администратор. – 2007. – №. 4. – С. 72-77.
+
+9.	\emph{Алексеев Е. Р., Демин П. А., Костюк Д. А.} Возможности графического вывода результатов в последовательных и
+	параллельных кроссплатформенных вычислительных приложениях на Фортране и С(С++) // Advanced science, 2017, №3.
 

### 2019-05-08 ped2018/AvdeevNN_3_geom.tex
Theses for PED2018

\documentclass[11pt]{article}
\usepackage[left=20mm, right=20mm, top=20mm, bottom=20mm, includefoot, a5paper]{geometry}
\usepackage[cp1251]{inputenc}
\usepackage[english,russian]{babel} %ѕоправлено. ѕозволит нормально переносить слова в английском тексте.
\usepackage[intlimits]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{latexsym}
\pagestyle{empty}

\usepackage{graphicx} % ј как иначе вставл€ть картинки??

\begin{document}

\addcontentsline{toc}{section}{
	Avdeev N.\,N.
	On integral point sets in special position
}

\small UDC 511.95 + 514.112
\begin{center}
\bf
ON INTEGRAL POINT SETS IN SPECIAL POSITION%
\footnote{This work was supported by the Russian Science
Foundation under grant 16-11-10125.}\\
\small Avdeev N.\,N.\\
\small Voronezh State University\\
\small avdeev@math.vsu.ru, nickkolok@mail.ru\\
\end{center}

% ќбнул€ю счЄтчик на вс€кий случай
\setcounter{equation}{0}

% Switching to English
\selectlanguage{english}
\setcounter{figure}{0}

\textbf{Definition 1.}
A planar integral point set (PIPS) is a set $M$ of noncollinear points in plane $\mathbb{R}^2$
such that for any pair of points $M_1, M_2\in M$ the distance between $M_1$ and $M_2$ is integral.
Every PIPS is finite [1].

Some general constructions of PIPS are known,
e.g. a concyclic PIPS of an arbitrary power [2]
and a PIPS of an arbitrary power $n$ such that $(n-1)$ points are collinear [3].
However, there is no known algorithm to construct a PIPS of arbitrary power $n$ which is contained in 2 parallel straight lines,
no more than $n-2$ points on each line.

Examples of such PIPS for $n=7,8,9$ can be found in [4,5].
Below an example for $n=10$ is given.
We use notation
\begin{equation*}
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	:=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},\frac{y_n\sqrt{p}}{q}\right)  \right\}
	.
\end{equation*}
\\
Figure~\ref{Avdeev_10_158_1538485325776.crop.eps}: $
\sqrt{70}/79 *
\{
( \pm6241 , 0),
( -873 , -528),
( 2093 , 408),
( 4838 , 138),
$\\$
( -3275 , 936),
( -1872 , 798),
( 5265 , 96),
( -2299 , 840),
( 873 , 528)
\}
$.
%\\
%Figure~\ref{Avdeev_10_222_1538488624864.crop.eps}
%
%$
%\sqrt{70}/37 *
%\{
%( -4107 , 0),
%( 4107 , 0),
%( 840 , 198)
%$\\$
%( -2622 , -90),
%( -1599 , -152),
%( -246 , -234),
%$\\$
%( 381 , -272),
%( 1206 , -322),
%( 1833 , -360),
%( 3186 , -442),
%\}
%$

\begin{figure}[h]
\center
	\includegraphics[width=0.7\linewidth]{Avdeev_10_158_1538485325776.eps}
	\caption{PIPS of power 10 and diameter 158}\label{Avdeev_10_158_1538485325776.crop.eps}%-converted-to-crop.ps}
\end{figure}

%\begin{figure}[h!]
%	\center
%	\includegraphics[width=0.7\linewidth]{Avdeev_10_222_1538488624864.eps}
%	\caption{PIPS of power 10 and diameter 222}\label{Avdeev_10_222_1538488624864.crop.eps}%-converted-to-crop.ps}
%\end{figure}

 \centerline{\bf References}

1.
Erd\"{o}s P. Integral distances //Bulletin of the American Mathematical Society. Ц 1945. Ц V. 51. Ц No. 12. Ц P. 996.

2.
Harborth H., Kemnitz A., M\"{o}ller M. An upper bound for the minimum diameter of integral point sets
//Discrete \& Computational Geometry. Ц 1993. Ц V. 9. Ц No. 4. Ц P. 427-432.

3.
Solymosi J. Note on integral distances //Discrete \& Computational Geometry. Ц 2003. Ц V. 30. Ц No. 2. Ц P. 337-342.

\selectlanguage{russian}
4.
јвдеев Ќ.\,Ќ.
ќб отыскании целоудалЄнных множеств специального вида
//
—б. тр. ћеждународной конференции Ђјктуальные проблемы прикладной математики, информатики и механикиї
Ц ¬оронеж, 2018. Ц в печати. (in Russian)

\selectlanguage{english}

5.
Guy~R. Unsolved problems in number theory. V.
	1. --- Springer Science \& Business Media, 2013.

{\bf Avdeev Nikolai Nikolaevich},
master student of department of functional analysis and operator equations, Voronezh State University.
%магистрант каф. функционального анализа и операторных уравнений математического факультета ‘√Ѕќ” ¬ќ <<¬√”>>

% ќбнул€ю счЄтчик за собой
\setcounter{equation}{0}
\setcounter{figure}{0}

\end{document}

### 2019-05-08 grant201810/AvdeevNN_Erdos_sets.tex
For RSF grant

\documentclass[a4paper,12pt]{article}

\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
\addbibresource{../common/notmy.bib}
\addbibresource{../common/my.bib}

%\usepackage{cite}

\usepackage{amsthm}
\newtheoremstyle{VMMSHtheorem}% name of the style to be used
	{\topsep}% measure of space to leave above the theorem. E.g.: 3pt
	{\topsep}% measure of space to leave below the theorem. E.g.: 3pt
	{\sl}% name of font to use in the body of the theorem
	{\parindent}% measure of space to indent
	{\bfseries}% name of head font
	{.}% punctuation between head and body
	{ }% space after theorem head; " " = normal interword space
	{\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}

\theoremstyle{VMMSHtheorem}

\newtheorem{theorem}{Теорема}
\newtheorem*{theorem*}{Теорема}
\newtheorem{lemma}[theorem]{Лемма}
\newtheorem*{lemma*}{Лемма}
\newtheorem{proposition}[theorem]{Утверждение}
\newtheorem*{proposition*}{Утверждение}
\newtheorem{corollary}[theorem]{Следствие}
\newtheorem*{corollary*}{Следствие}
\newtheorem{hypothesis}[theorem]{Гипотеза}
\newtheorem*{hypothesis*}{Гипотеза}

\newtheoremstyle{VMMSHremark}% name of the style to be used
	{\topsep}% measure of space to leave above the theorem. E.g.: 3pt
	{\topsep}% measure of space to leave below the theorem. E.g.: 3pt
	{}% name of font to use in the body of the theorem
	{\parindent}% measure of space to indent
	{\sc}% name of head font
	{.}% punctuation between head and body
	{ }% space after theorem head; " " = normal interword space
	{\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}

\theoremstyle{VMMSHremark}
\newtheorem{definition}[theorem]{Определение}
\newtheorem*{definition*}{Определение}
\newtheorem{designation}[theorem]{Обозначение}
\newtheorem*{designation*}{Обозначение}
\newtheorem{remark}[theorem]{Замечание}
\newtheorem*{remark*}{Замечание}
\newtheorem{example}[theorem]{Пример}
\newtheorem*{example*}{Пример}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{amscd}
\usepackage[english,russian]{babel}

\sloppy

\begin{document}

{
\center{МНОЖЕСТВА ТОЧЕК С ЦЕЛОЧИСЛЕННЫМИ РАССТОЯНИЯМИ НА ПЛОСКОСТИ И В ЕВКЛИДОВОМ ПРОСТРАНСТВЕ} % обязательное поле!
}

\vspace{2cm}

Множеством с целочисленными расстояниями будет называть такое подмножество плоскости $\mathbb{R}^2$,
не содержащееся ни в какой прямой,
что расстояние между любыми двумя точками $M$ есть целое число.

В \cite{anning1945integral,erdos1945integral} доказано, что любое МЦР конечно.
Следовательно, для МЦР естественным образом определяется диаметр:
$$
	diam(M) = \max_{A,B\in M} |A-B|,
$$
где $|A-B|$~--- обычное расстояние между $A$ и $B$.

Пусть $\mathfrak{M}_n$~--- множество всех МЦР, состоящих ровно из $n$ точек.
В~\cite{harborth1993upper} была рассмотрена функция
$$
	f(n) = \min_{M\in\mathfrak{M}_n} diam(M)
$$
и найдена верхняя оценка на эту функцию.
Кроме того, там же вычислены значения $f(n)$ для первых нескольких $n$.

В~\cite{solymosi2003note} была доказана нижняя оценка на $f(n)$:
\begin{equation}\label{Solymosi}
f(n) \geqslant \frac{1}{24} n
.
\end{equation}
В~\cite{our-mz-rus} коэффициент при $n$ был улучшен до 1/8.

Планируется:
\begin{itemize}
\item
ещё более улучшить оценку~\eqref{Solymosi};
\item
изучить связь задачи об отыскании $f(n)$ и задачи об упаковке точек в квадрат~\cite{locatelli2002packing};
\item
построить частные случаи оценки~\eqref{Solymosi}:
например, для МЦР, содержащихся в окружности, гиперболе и семействе софокусных гипербол;
\item
изучить такие МЦР $M\in\mathfrak{M}_n$, что $diam(M)=f(n)$.
\end{itemize}

\printbibliography

\end{document}

### 2019-05-08 pps-linear-bound/pps-linear-lower-bound.tex
Drafts for new paper about new linear bound using PPS

\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel} %используем русский и английский языки с переносами
\usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
\usepackage[pdftex,unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{indentfirst} % включить отступ у первого абзаца
\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
\graphicspath{{illustr/}}%путь к рисункам

\makeatletter
\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"

\usepackage{geometry} % Меняем поля страницы.
\geometry{left=2cm}% левое поле
\geometry{right=1cm}% правое поле
\geometry{top=2cm}% верхнее поле
\geometry{bottom=2cm}% нижнее поле

\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра

\sloppy

\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}

\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
\addbibresource{../common/notmy.bib}
\addbibresource{../common/my.bib}

\begin{document}
\renewcommand{\bibname}{Список цитированной литературы}
\renewcommand\refname{\bibname}
% !!!
% The text starts here

Avdeev N.N.

On existence and diameter bounds of integral point sets.

\paragraph{Abstract.}
A point set $M$ in $m$-dimensional Euclidean space is called an integral point set if all the distances between the
elements of $M$ are integers and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
We improve the linear lower bound for planar integral point sets.
This bound involves facts related to the Point Packing in a Square problem.
Then for arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
we give a construction of an integral point set $M$ of $n$ points in $m$-dimensional Euclidean space
where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.

\section{Introduction.}

\section{Lower bound for the diameter.}

\section{Constructing integral point sets}
For every integer $n\geq 3$ Solymosi~\cite{solymosi2003note} gives a construction of facher planar integral point set $M$
such as $|M| = n$ and for some $M_1, M_2 \in M$ holds $|M_1 - M_2| = 2$.
The construted set has both odd and even distances.

Now we will improve the Solymosi's result.

\paragraph{Construction 1.}
Let us choose a positive integer $k > 1$ and take
\begin{equation}
	a = 2^{2^k} - 1
	.
\end{equation}
Then
\begin{equation}
	\label{eq:a_equiv_3_mod_4}
	a \equiv 3 \mod 4
\end{equation}
and, moreover,
\begin{multline}
	a = \left(2^{2^{k-1}}\right)^2 - 1
	=
	\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-1}} - 1\right)
	=
	\\=
	\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-2}} + 1\right) \cdot ... \cdot \left(2^{2^1} + 1\right) \left(2^2 - 1\right)
	.
\end{multline}
We take $d_j = 2^{2^j} + 1$ for $1 \leq j \leq k-1$.
Then $d_j \equiv 1 \mod 4$.

Let $c_J = \prod_{j\in J} d_j$ for every subset of indexes $J\subset\{1,2,...,k-1\}$.
We get
\begin{equation}
	\label{eq:c_J_equiv_1_mod_4}
	c_J\equiv 1 \mod{4}
\end{equation}
and, moreover, $a$ is divisible by $c_J$.

From~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4} we obtain $a/c_J \equiv 3 \mod 4$.
Let $b_J = (c_J - a/c_J)/2$, then $b_J \equiv 1 \mod 2$.
Let us further take $g_J = (c_J + a/c_J)/2$, then $g_J \equiv 0 \mod 2$.

We will now define the coordinates of the points as following:
\begin{equation}
	M_J =\left(\frac{b_J}{2}, 0\right)
	,
	~~
	N   =\left(0, \frac{\sqrt{a}}{2}\right)
	,
	~~
	H_{1,2} = \left(\pm\frac{1}{2}, 0\right)
	.
\end{equation}
Then the distances are:
\begin{equation}
	|N - H_i|
	=
	\left(\frac{1}{4} + \frac{a}{4}\right)^{1/2}
	=
	\frac{1}{2}(a + 1)^{1/2}
	=
	\frac{1}{2}\cdot 2^{2^{k-1}}
	\in\mathbb{N}
	,
\end{equation}
\begin{multline}
	|N - M_J|
	=
	\left(\frac{b_J^2}{4} + \frac{a}{4}\right)^{1/2}
	=
	\frac{1}{2}\left(\left(\frac{c_J - a/c_J}{2}\right)^2 + a\right)^{1/2}
	=
	\\=
	\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 - \frac{a}{2} + \left(\frac{a}{c_j}\right)^2 + a\right)^{1/2}
	=
	\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 + \frac{a}{2} + \left(\frac{a}{c_j}\right)^2    \right)^{1/2}
	=
	\\=
	\frac{1}{2}\left(\left(\frac{c_J + a/c_J}{2}\right)^2\right)^{1/2}
	=
	\frac{1}{2}\left(\frac{c_J + a/c_J}{2}\right)
	=
	\frac{g_J}{2}
	\in\mathbb{N}
	,
\end{multline}
\begin{equation}
	|M_J - H_i|
	=
	\frac{b_J}{2} \pm \frac{1}{2}
	=
	\frac{b_J \pm 1}{2}
	\in\mathbb{N}
	.
\end{equation}

So, $M = \{ M_J, N, H_{1,2}\}$ is indeed a planar integral point set of $2^{k-1}+3$ points,
and distance 1 occurs in $M$.
Reminding that $k$ can be taken arbitrary large, we conclude that we can construct
a planar integral point set of arbitary large count of points so that the distance 1 occurs in that set.

TODO: $k=2$ and construction from~\cite{harborth1993upper}

TODO: remark on taking $k=1$.

Summing up the construction 1 above and~\cite[lemma 3 and proposition 6]{our-vmmsh-2018},
we obtain the following

TODO: should I write translated version of the results here?

\paragraph{Theorem 3.1.}
For every $n\geq 3$ there is a planar intergal set $M$ of $n$ points
such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds.
This set consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.

And vice versa, if $M$ is a planar integral point set of $n$ points
such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds,
then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.
Moreover, if $n > 3$, then the distance 1 occurs in $M$ only once.

Using the ``blowing up'' procedure described in~\cite[theorem 1.3]{kurz2008bounds},
we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,
with distance 1 occuring in it.
If we have an integral point set with distance 1 occuring,
then we can easily stretch it to turn 1 into the desired distance.
These facts, together with theorem 3.1, give the following

\paragraph{Theorem 3.2.}
For arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
there is an integral point set $M$ of $n$ points in $m$-dimensional Euclidean space,
so that for some $M_1, M_2\in M$ distance between $M_1$ and $M_2$ is exactly $d$.

\paragraph{Definition.}
We will call an integral point set $M$ \textit{irreducible}, if the greatest common divisor
of all the distances occuring in $M$ is 1.

If an integral point set is irreducible,
then it cannot be squashed to an integral point set of the same power and structure but smaller diameter.

\paragraph{Hypothesis 3.1.}
For every $m \geq 2$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{irreducible}
integral point set $M$ of $n$ points in $m$-dimensional Euclidean space
where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.

For $m \geq 3$, hypothesis 3.1 can be easily proved:
indeed, we can just take a facher integral point set acording to construction 1
with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[theorem 1.3]{kurz2008bounds},
replacing the out-of-line point with an $(m-2)$-dimensional simplex with proper side length.
That simplex will consist of $m-1$ points;
we can throw out up to all the points $M_J$ and still have points $H_{1,2}$ with distance 1.
So, the set consists of $m+1$ or more points and has distance 1 occuring in it.
As distance 1 occurs in the obtained set, it is irreducible.

So, hypothesis 3.1 needs proof only for $m=2$.

\section{Acknowledgements}
Author thanks Prof. E.M. Semenov for the fruitful discussion and ideas
and A.S. Chervinskaia for the idea of using the term ``facher''.

\printbibliography

\end{document}

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Linguistic issues

@@
 We improve the linear lower bound for planar integral point sets.
-This bound involves facts related to the Point Packing in a Square problem.
+This improvement takes into account some results related to the Point Packing in a Square problem.
 Then for arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
@@
 For $m \geq 3$, hypothesis 3.1 can be easily proved:
-indeed, we can just take a facher integral point set acording to construction 1
+indeed, we can just take a facher integral point set according to construction 1
 with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[theorem 1.3]{kurz2008bounds},
@@
 Author thanks Prof. E.M. Semenov for the fruitful discussion and ideas
-and A.S. Chervinskaia for the idea of using the term ``facher''.
+and A.S. Chervinskaia for the idea of using the term ``facher'' and other linguistic advices.
 

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Intro - first attempt

@@
 \section{Introduction.}
+Let $\mathbb{N}$ be the set of all positive integers and let $|M_1 - M_2|$ denote the Euclidean distance
+between points $M_1$ and $M_2$ in a finite-dimensional space $\mathbb{R}^m$.
+An \textit{integral point set} in $m$-dimensional Euclidean space is such a point set $M$ that all the distances between the
+points of $M$ are integers and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
+Erd\"os and Anning proved~\cite{anning1945integral,erdos1945integral} that every integral point set consists of a finite number of points.
+Taking the fact into account, we will denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space with
+$\mathfrak{M}(m,n)$ (following the notation of~\cite{our-vmmsh-2018}).
+The symbol $|M|$ will be used for cardinality of $M$, i.e. the quantity of points in $M$ in our case.
+
+For every finite point set, its diameter is naturally defined as
+\begin{equation}
+	\operatorname{diam} M = \max_{A,B\in M} |A-B|
+	.
+\end{equation}
+The next emerging question is: how the diameter of an integral point set depends on its cardinality?
+One can easily see that every $M\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = h$
+can be stretched to $M_p\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = ph$
+for every $p\in\mathbb{N}$.
+So, the question above should be rephrased:
+how \textit{the least possible} diameter of an integral point set depends on its cardinality?
+To answer it, the following function was introduced~\cite{kurz2008bounds,kurz2008minimum}:
+\begin{equation}
+	d(m,n) = \min_{M\in\mathfrak{M}(m,n)} \operatorname{diam} M = \min_{M\in\mathfrak{M}(m,n)} \max_{A,B\in M} |A-B|
+	.
+\end{equation}
+We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$
+and estimations; in the present paper, the case of $m=2$ will mostly be in the focus.
+
+The most significant breakthrough on case of $m=2$ was done by Solymosi~\cite{solymosi2003note},
+who proved that $cn \leq d(2,n)$ for a sufficiently small constant $c$.
+Followed the details of Solymosi's proof carefully,
+one can know out that the inequality holds at least for $c = 1/24$.
+The constant was improved in~\cite{our-mz-rus,our-vmmsh-2018}
+to $1/8$ for all $n$ and $3/8$ for sufficiently big $n$, respectively.
+
+The paper~\cite{solymosi2003note} contained one more interesting result.
+Let us define a function which is ``dual'' to $d(m,n)$ in some sense:
+\begin{equation}
+	l(m,n) = \min_{M\in\mathfrak{M}(m,n)} \min_{A,B\in M} |A-B|
+	.
+\end{equation}
+Solymosi proved that $l(2,n)\leq 2$.
+
+In the present paper we improve Solymosi's results:
+firstly, we obtain a greater constant $c$ in Theorem (TODO: the number of theorem and the exact constant),
+using the combined approach with the Point Packing in a Square problem;
+then, we prove that $l(m,n)=1$ for all $m$ and $n$ which it is correctly defined on.
 

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Using amsthm - WIP

@@
 \addbibresource{../common/my.bib}
+
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}[section]
 
@@
 
-\paragraph{Theorem 3.2.}
+\begin{theorem}
 	For arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
@@
 	so that for some $M_1, M_2\in M$ distance between $M_1$ and $M_2$ is exactly $d$.
+\end{theorem}
 

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Cosmetics

@@
 
-\section{Introduction.}
+\section{Introduction}
 Let $\mathbb{N}$ be the set of all positive integers and let $|M_1 - M_2|$ denote the Euclidean distance
@@
 
-\section{Lower bound for the diameter.}
+\section{Lower bound for the diameter}
 
@@
 indeed, we can just take a facher integral point set according to construction 1
-with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[theorem 1.3]{kurz2008bounds},
+with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
 replacing the out-of-line point with an $(m-2)$-dimensional simplex with proper side length.

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Linguistic cosmetics

@@
 
-The most significant breakthrough on case of $m=2$ was done by Solymosi~\cite{solymosi2003note},
+The most significant breakthrough on the planar case was done by Solymosi~\cite{solymosi2003note},
 who proved that $cn \leq d(2,n)$ for a sufficiently small constant $c$.
-Followed the details of Solymosi's proof carefully,
+Following the details of Solymosi's proof carefully,
 one can know out that the inequality holds at least for $c = 1/24$.

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Improve some citations

@@
 one can know out that the inequality holds at least for $c = 1/24$.
-The constant was improved in~\cite{our-mz-rus,our-vmmsh-2018}
-to $1/8$ for all $n$ and $3/8$ for sufficiently big $n$, respectively.
+(See~\cite[Exercise 2.6]{garibaldi2005erdos} for some remarks.)
+The constant was improved in~\cite{our-mz-rus}to $1/8$ for all $n$ and in~\cite{our-vmmsh-2018}
+to $3/8$ for sufficiently big $n$.
 

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
TODO note on using the term 'facher'
A theorem was cosmetically beautified

@@
 then, we prove that $l(m,n)=1$ for all $m$ and $n$ which it is correctly defined on.
+
+TODO: definition of facher IPS somewhere.
 
@@
 	For arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
-	there is an integral point set $M$ of $n$ points in $m$-dimensional Euclidean space,
-	so that for some $M_1, M_2\in M$ distance between $M_1$ and $M_2$ is exactly $d$.
+	there exists $M\in\mathfrak{M}(m,n)$,
+	such that for some $M_1, M_2\in M$ equality $|M_1 - M_2| = d$ holds.
 \end{theorem}

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Amsthm completed

@@
 \newtheorem{theorem}{Theorem}[section]
+\newtheorem{hypothesis}[theorem]{Hypothesis}
+\newtheorem{lemma}[theorem]{Lemma}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{construction}[theorem]{Construction}
 
@@
 
-\paragraph{Construction 1.}
+\begin{construction}
+	\label{con:planar_set_with_minimeter_1}
 	Let us choose a positive integer $k > 1$ and take
@@
 	a planar integral point set of arbitary large count of points so that the distance 1 occurs in that set.
+\end{construction}
 
@@
 
-\paragraph{Theorem 3.1.}
+\begin{theorem}
+	\label{thm:minimeter_1_planar}
 	For every $n\geq 3$ there is a planar intergal set $M$ of $n$ points
@@
 	Moreover, if $n > 3$, then the distance 1 occurs in $M$ only once.
+\end{theorem}
 
@@
 then we can easily stretch it to turn 1 into the desired distance.
-These facts, together with theorem 3.1, give the following
+These facts, together with Theorem~\ref{thm:minimeter_1_planar}, give the following
 
@@
 
-\paragraph{Definition.}
+\begin{definition}
 	We will call an integral point set $M$ \textit{irreducible}, if the greatest common divisor
 	of all the distances occuring in $M$ is 1.
+\end{definition}
 
@@
 
-\paragraph{Hypothesis 3.1.}
+\begin{hypothesis}
+	\label{hyp:irreducible_planar}
 	For every $m \geq 2$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{irreducible}
@@
 	where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
+\end{hypothesis}
 
-For $m \geq 3$, hypothesis 3.1 can be easily proved:
-indeed, we can just take a facher integral point set according to construction 1
+For $m \geq 3$, Hypothesis~\ref{hyp:irreducible_planar} can be easily proved:
+indeed, we can just take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
 with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
@@
 
-So, hypothesis 3.1 needs proof only for $m=2$.
+So, Hypothesis~\ref{hyp:irreducible_planar} needs proof only for $m=2$.
 

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Linguistic fix

@@
 using the combined approach with the Point Packing in a Square problem;
-then, we prove that $l(m,n)=1$ for all $m$ and $n$ which it is correctly defined on.
+then, we prove that $l(m,n)=1$ for all possible $m$ and $n$.
 

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Remarks on the construction

@@
 \newtheorem{construction}[theorem]{Construction}
+\newtheorem{remark}[theorem]{Remark}
 
@@
 
-TODO: $k=2$ and construction from~\cite{harborth1993upper}
+TODO: get rid of $H_{1,2}$; write signs $\pm$ to $M_J$; recalculate the cardinality; allow $k=1$.
 
-TODO: remark on taking $k=1$.
+\begin{remark}
+	Taking $k=1$ for Construction~\ref{con:planar_set_with_minimeter_1},
+	we obtain 3 points which form a right triangle with side length 1.
+	This triangle is obviously the optimal set in $\mathfrak{M}(2,3)$.
+\end{remark}
+
+\begin{remark}
+	Taking $k=2$ for Construction~\ref{con:planar_set_with_minimeter_1},
+	we obtain the optimal set in $\mathfrak{M}(2,5)$ presented in~\cite[Fig. 1]{harborth1993upper}.
+	If we truncate one point from it,
+	then we get on of the two optimal sets in $\mathfrak{M}(2,4)$.
+\end{remark}
+
+\begin{hypothesis}
+	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 - M_2|=1$ holds,
+	is not optimal.
+\end{hypothesis}
 

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Add definition of facher IPS

@@
 
-TODO: definition of facher IPS somewhere.
-
 \section{Lower bound for the diameter}
@@
 \section{Constructing integral point sets}
+
+\begin{definition}
+	A set $M\in\mathfrak{M}(2,n)$ is called \textit{facher}
+	if $M$ consists of $n-1$ points on a straight line
+	and one point out of the line.
+\end{definition}
+
+\begin{definition}
+	A set $M\in\mathfrak{M}(m,n)$ is called \textit{optimal}
+	if $\operatorname{diam}M=d(m,n)$.
+\end{definition}
+
+Facher sets are the simplest ones of planar integral point sets.
+It is shown that for $9\leq n \leq 122$ all the optimal sets are facher~\cite{kurz2008minimum}.
+For every cardinality $n$ (and every characteristic $q$~--- see~\cite{kurz2005characteristic} for details)
+there exists a facher set $M\in\mathfrak{M}(2,n)$ (with the desired characteristic).
+
 For every integer $n\geq 3$ Solymosi~\cite{solymosi2003note} gives a construction of facher planar integral point set $M$

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
TODO note
Minor edits

@@
 A point set $M$ in $m$-dimensional Euclidean space is called an integral point set if all the distances between the
-elements of $M$ are integers and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
-We improve the linear lower bound for planar integral point sets.
+elements of $M$ are integers, and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
+We improve the linear lower bound for diameter of planar integral point sets.
 This improvement takes into account some results related to the Point Packing in a Square problem.
 Then for arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
-we give a construction of an integral point set $M$ of $n$ points in $m$-dimensional Euclidean space
+we give a construction of an integral point set $M$ of $n$ points in $m$-dimensional Euclidean space,
 where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
@@
 (See~\cite[Exercise 2.6]{garibaldi2005erdos} for some remarks.)
-The constant was improved in~\cite{our-mz-rus}to $1/8$ for all $n$ and in~\cite{our-vmmsh-2018}
+The constant was improved in~\cite{our-mz-rus} to $1/8$ for all $n$ and in~\cite{our-vmmsh-2018}
 to $3/8$ for sufficiently big $n$.
@@
 \begin{remark}
-	Taking $k=2$ for Construction~\ref{con:planar_set_with_minimeter_1},
+	For $k=2$ in Construction~\ref{con:planar_set_with_minimeter_1},
 	we obtain the optimal set in $\mathfrak{M}(2,5)$ presented in~\cite[Fig. 1]{harborth1993upper}.
 	If we truncate one point from it,
-	then we get on of the two optimal sets in $\mathfrak{M}(2,4)$.
+	then we get one of the two optimal sets in $\mathfrak{M}(2,4)$.
 \end{remark}
@@
 
+TODO: construction with unique minimum edge?
+
 \begin{definition}

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Drafts of results to be used

@@
 \section{Lower bound for the diameter}
+
+We will need the following results and notions from~\cite{our-vmmsh-2018}.
+
+\begin{lemma}
+	\cite[Corollary 1]{our-vmmsh-2018}
+	\label{lem:max_points_on_line}
+	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
+	Then there is no straight line which contains more than $(d+3)/2$ points of $M$.
+\end{lemma}
+
+\begin{lemma}
+	\cite[Lemma 4]{our-vmmsh-2018}
+	\label{lem:square_container}
+	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
+	Then $M$ is situated in a square with side length $d$.
+\end{lemma}
+
+\begin{definition}
+	\textit{Cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is union of two straight lines:
+	the line which the points $M_1$ and $M_2$ are situated on,
+	and the bisector of line segment $M_1 M_2$.
+\end{definition}
+
+\begin{lemma}
+	\label{lem:intervals_cross}
+	If \textbf{open} line segments $M_1 M_2$ and $M_3 M_4$ does not intersect,
+	$M_1 \neq M_2$, $M_3 \neq M_4$,
+	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains not more than 4 points.
+\end{lemma}
+
+\begin{lemma}
+	\label{lemma:quadr_diag_edges}
+	Let $ABCD$ be a convex quadrilateral on the plane.
+	Then $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
+	i.e. at least one diagonal is greater than at least one side.
+\end{lemma}
+
+Определим теперь коэффициенты упаковки точек в квадрат $\varphi_k$.
+Пусть
+\begin{equation*}
+	\Phi_k = \{ P \subset [0;1]^2 : |P|=k\}
+	,
+\end{equation*}
+где $[0;1]^2$~--- замкнутый единичный квадрат на плоскости.
+Тогда
+\begin{equation*}
+	\varphi_k = \max_{P \in \Phi_k} \min_{A,B \in P} |A - B|
+	.
+\end{equation*}
+Иначе говоря, через $\varphi_k$ будем обозначать наибольшее число, такое,
+что в единичном квадрате нельзя разместить $k$ точек так,
+чтобы расстояние между любыми двумя точками было не менее $\varphi_k$.
+Проблема отыскания $\varphi_k$ носит название проблемы упаковки точек в квадрат~\cite{locatelli2002packing,costa2013valid}.
+
+\begin{theorem}%[\cite{costa2013valid}]
+	\label{thm:varphi_k_bounds}
+	\cite{costa2013valid}
+	Для $k\geq 2$ выполнено
+	\begin{equation*}
+		\sqrt{\frac{2}{k\sqrt{3}}}
+		\leq
+		\varphi_k
+		\leq
+		\frac{1}{k-1} +
+		\sqrt{
+			\frac{1}{(k-1)^2}
+			+
+			\frac{2}{(k-1)\sqrt{3}}
+		}
+	\end{equation*}
+\end{theorem}
 

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Add one more citation

@@
 For every cardinality $n$ (and every characteristic $q$~--- see~\cite{kurz2005characteristic} for details)
-there exists a facher set $M\in\mathfrak{M}(2,n)$ (with the desired characteristic).
+there exists a facher set $M\in\mathfrak{M}(2,n)$ (with the desired characteristic~\cite[Theorem 5]{our-vmmsh-2018}).
 

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Cosmetic edits

@@
 
-For every integer $n\geq 3$ Solymosi~\cite{solymosi2003note} gives a construction of facher planar integral point set $M$
-such as $|M| = n$ and for some $M_1, M_2 \in M$ holds $|M_1 - M_2| = 2$.
-The construted set has both odd and even distances.
+For every integer $n\geq 3$ Solymosi~\cite{solymosi2003note} gives a construction of a facher integral point set
+$M\in\mathfrak{M}(2,n)$
+such that equality $|M_1 - M_2| = 2$ holds for some $M_1, M_2 \in M$.
+The constructed set has both odd and even distances.
 

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Edits to the construction

@@
 
-	Let $c_J = \prod_{j\in J} d_j$ for every subset of indexes $J\subset\{1,2,...,k-1\}$.
+	Let $c_J = \prod_{j\in J} d_j$ for every subset of indexes $J\subset I = \{1,2,...,k-1\}$.
 	We get
@@
 	\begin{equation}
-		M_J =\left(\frac{b_J}{2}, 0\right)
+		M_{J\pm} =\left(\pm\frac{b_J}{2}, 0\right)
 		,
@@
 		N   =\left(0, \frac{\sqrt{a}}{2}\right)
-		,
-		~~
-		H_{1,2} = \left(\pm\frac{1}{2}, 0\right)
 		.
@@
 	Then the distances are:
-	\begin{equation}
-		|N - H_i|
-		=
-		\left(\frac{1}{4} + \frac{a}{4}\right)^{1/2}
-		=
-		\frac{1}{2}(a + 1)^{1/2}
-		=
-		\frac{1}{2}\cdot 2^{2^{k-1}}
-		\in\mathbb{N}
-		,
-	\end{equation}
 	\begin{multline}
-		|N - M_J|
+		|N - M_{J\pm}|
 		=
@@
 	\begin{equation}
-		|M_J - H_i|
+		|M_{J_1 \pm} - M_{J_2 \pm}|
 		=
-		\frac{b_J}{2} \pm \frac{1}{2}
+		\left|\frac{b_{J_1}}{2} \pm \frac{b_{J_2}}{2}\right|
 		=
-		\frac{b_J \pm 1}{2}
+		\left|\frac{b_{J_1} \pm b_{J_2}}{2}\right|
 		\in\mathbb{N}
@@
 	\end{equation}
+	In particular, for $k>1$ and $H = \{k-1\}$ we obtain $C_H = 2^{2^{k-1}}+1$
+	(for $k=1$, we consider $H=\varnothing$ and $c_H = 1$).
+	Moreover
+	\begin{equation}
+		b_H =
+		\left( 2^{2^{k-1}}+1 - \frac{a}{2^{2^{k-1}}+1} \right)/2
+		=
+		\left(2^{2^{k-1}}+1 - \left(  2^{2^{k-1}}-1 \right) \right)/2
+		=
+		1
+		.
+	\end{equation}
+	Then one of the distances is
+	\begin{equation}
+		|M_{H+} - M_{H-}|
+		=
+		\left|\frac{b_{H}}{2} - \frac{-b_{H}}{2}\right|
+		=
+		\left|\frac{1}{2} - \frac{-1}{2}\right|
+		= 1
+		.
+	\end{equation}
 
-	So, $M = \{ M_J, N, H_{1,2}\}$ is indeed a planar integral point set of $2^{k-1}+3$ points,
+	So, $M = \{ M_{J\pm}, N\}$ is indeed a planar integral point set of $2^k+1$ points,
 	and distance 1 occurs in $M$.
@@
 \begin{remark}
-	Taking $k=1$ for Construction~\ref{con:planar_set_with_minimeter_1},
+	For $k=1$ in Construction~\ref{con:planar_set_with_minimeter_1},
 	we obtain 3 points which form a right triangle with side length 1.
@@
 That simplex will consist of $m-1$ points;
-we can throw out up to all the points $M_J$ and still have points $H_{1,2}$ with distance 1.
+we can throw out up to all points $M_J$, except the two points with distance 1.
 So, the set consists of $m+1$ or more points and has distance 1 occuring in it.

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Removing k=1 from the construction

@@
 
-	Let $c_J = \prod_{j\in J} d_j$ for every subset of indexes $J\subset I = \{1,2,...,k-1\}$.
+	Let $c_J = \prod_{j\in J} d_j$ for every subset of indexes $J\subset I = \{1,2,...,k-1\}$
+	(and  $c_\varnothing = 1$).
 	We get
@@
 	\end{equation}
-	In particular, for $k>1$ and $H = \{k-1\}$ we obtain $C_H = 2^{2^{k-1}}+1$
-	(for $k=1$, we consider $H=\varnothing$ and $c_H = 1$).
-	Moreover
+	In particular, for $H = \{k-1\}$ we obtain $C_H = 2^{2^{k-1}}+1$ and
 	\begin{equation}
@@
 
-TODO: get rid of $H_{1,2}$; write signs $\pm$ to $M_J$; recalculate the cardinality; allow $k=1$.
+TODO: recalculate the cardinality.
 
 \begin{remark}
-	For $k=1$ in Construction~\ref{con:planar_set_with_minimeter_1},
-	we obtain 3 points which form a right triangle with side length 1.
+	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ literally,
+	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then
+	obtain a right triangle with side length 1.
 	This triangle is obviously the optimal set in $\mathfrak{M}(2,3)$.

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Cosmetics
stretch -> dilate (because dilation preserves shape, which is our case)

@@
 One can easily see that every $M\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = h$
-can be stretched to $M_p\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = ph$
+can be dilated to $M_p\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = ph$
 for every $p\in\mathbb{N}$.
@@
 	\end{equation}
-	Then one of the distances is
+	Then, one of the distances is
 	\begin{equation}
@@
 If we have an integral point set with distance 1 occuring,
-then we can easily stretch it to turn 1 into the desired distance.
+then we can easily dilate it to turn 1 into the desired distance.
 These facts, together with Theorem~\ref{thm:minimeter_1_planar}, give the following

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Proper reference to construction
Grant notice

@@
 On existence and diameter bounds of integral point sets.
+\footnote{
+	This work was carried out at Voronezh State University and supported by the Russian Science
+	Foundation under grant 19-11-00197.
+}
 
@@
 
-Summing up the construction 1 above and~\cite[lemma 3 and proposition 6]{our-vmmsh-2018},
+Summing up Construction~\ref{con:planar_set_with_minimeter_1} above and~\cite[lemma 3 and proposition 6]{our-vmmsh-2018},
 we obtain the following

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
One more link to extrenal paper
Add note about symmetry to the construction

@@
 there exists a facher set $M\in\mathfrak{M}(2,n)$ (with the desired characteristic~\cite[Theorem 5]{our-vmmsh-2018}).
+In~\cite{antonov2008maximal}, the facher sets of characteristic 1 are called \textit{semi-crabs} and investigated.
 
@@
 
-Summing up Construction~\ref{con:planar_set_with_minimeter_1} above and~\cite[lemma 3 and proposition 6]{our-vmmsh-2018},
+Summing up Construction~\ref{con:planar_set_with_minimeter_1} above with~\cite[lemma 3 and proposition 6]{our-vmmsh-2018}
+and~\cite[Section 6]{antonov2008maximal},
 we obtain the following
@@
 	such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds,
-	then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.
-	Moreover, if $n > 3$, then the distance 1 occurs in $M$ only once.
+	then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
+	and one point out of the line, on the bisector of line segment $M_1 M_2$.
+	There is only one maximal integral point set $M' \supseteq M$,
+	and the bisector is the axis of symmetry for $M'$.
+	Moreover, if $n > 3$, then the distance 1 occurs in $M$ (and $M'$) only once.
 \end{theorem}

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Note about charactristic

@@
 
+\begin{definition}
+	\cite{kurz2005characteristic}
+	A \textit{characteristic} of an integral point set $M$ is such a squarefree number $q$,
+	that for any points $M_1, M_2, M_3 \in M$ the area of triangle $M_1 M_2 M_3$
+	is $p\sqrt{q}$, where $p$ is rational.
+\end{definition}
+For given integral point set, the characteristic is determined uniquely.
+
 Facher sets are the simplest ones of planar integral point sets.
 It is shown that for $9\leq n \leq 122$ all the optimal sets are facher~\cite{kurz2008minimum}.
-For every cardinality $n$ (and every characteristic $q$~--- see~\cite{kurz2005characteristic} for details)
-there exists a facher set $M\in\mathfrak{M}(2,n)$ (with the desired characteristic~\cite[Theorem 5]{our-vmmsh-2018}).
+For every cardinality $n$ and every characteristic $q$
+there exists a facher set $M\in\mathfrak{M}(2,n)$ with characteristic $q$~\cite[Theorem 5]{our-vmmsh-2018}.
 In~\cite{antonov2008maximal}, the facher sets of characteristic 1 are called \textit{semi-crabs} and investigated.

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
The definition af maximal IPS
The PPS coefficients

@@
 \newtheorem{remark}[theorem]{Remark}
+\newtheorem{problem}[theorem]{Problem}
 
@@
 
-We will need the following results and notions from~\cite{our-vmmsh-2018}.
+In this section,
+we improve a better lower bound for minimum diameter of planar integral point sets by
+involving Point Packing in a Square problem.
+Below we introduce the problem, basic notions and results.
+
+\begin{problem}[Point Packing in a Square (PPS)\cite{locatelli2002packing,costa2013valid}]
+	\label{problem:PPS}
+	Given an integer $k > 1$, place $k$ points in the unit square $U = [ 0 , 1 ]^2$ such that their
+	minimum pairwise distance $m$ is maximal.
+\end{problem}
+
+\begin{definition}
+	For each $k > 1$, the corresponding maximal distance $m$ from Problem~\ref{problem:PPS}
+	is called the $k$-th PPS coefficient and denoted by $\varphi_k$.
+\end{definition}
+So, it's impossible to place $k$ points in a unit square such that each pairwise distance of the points is greater than $\varphi_k$.
+
+\begin{theorem}
+	\label{thm:varphi_k_bounds}
+	\cite{costa2013valid}
+	For every $k\geq 2$ the following inequality holds:
+	\begin{equation*}
+		\sqrt{\frac{2}{k\sqrt{3}}}
+		\leq
+		\varphi_k
+		\leq
+		\frac{1}{k-1} +
+		\sqrt{
+			\frac{1}{(k-1)^2}
+			+
+			\frac{2}{(k-1)\sqrt{3}}
+		}
+	\end{equation*}
+\end{theorem}
+
+For building the bound, we also require the following results and notions from~\cite{our-vmmsh-2018}.
 
@@
 \end{lemma}
-
-Определим теперь коэффициенты упаковки точек в квадрат $\varphi_k$.
-Пусть
-\begin{equation*}
-	\Phi_k = \{ P \subset [0;1]^2 : |P|=k\}
-	,
-\end{equation*}
-где $[0;1]^2$~--- замкнутый единичный квадрат на плоскости.
-Тогда
-\begin{equation*}
-	\varphi_k = \max_{P \in \Phi_k} \min_{A,B \in P} |A - B|
-	.
-\end{equation*}
-Иначе говоря, через $\varphi_k$ будем обозначать наибольшее число, такое,
-что в единичном квадрате нельзя разместить $k$ точек так,
-чтобы расстояние между любыми двумя точками было не менее $\varphi_k$.
-Проблема отыскания $\varphi_k$ носит название проблемы упаковки точек в квадрат~\cite{locatelli2002packing,costa2013valid}.
-
-\begin{theorem}%[\cite{costa2013valid}]
-	\label{thm:varphi_k_bounds}
-	\cite{costa2013valid}
-	Для $k\geq 2$ выполнено
-	\begin{equation*}
-		\sqrt{\frac{2}{k\sqrt{3}}}
-		\leq
-		\varphi_k
-		\leq
-		\frac{1}{k-1} +
-		\sqrt{
-			\frac{1}{(k-1)^2}
-			+
-			\frac{2}{(k-1)\sqrt{3}}
-		}
-	\end{equation*}
-\end{theorem}
 
@@
 
+%Now we need two more definitions.
+\begin{definition}
+	\cite{antonov2008maximal}
+	A set $M\in \mathfrak{M}(m,n)$ is \textit{maximal},
+	if there is no such a set $M'\in \mathfrak{M}(m,n+1)$
+	that $M \subsetneq M'$.
+\end{definition}
+
+%\begin{definition}
+%	A set $M\in \mathfrak{M}(m,n)$ is \textit{submaximal},
+%	if there is only one maximal integral point set $M'\in \mathfrak{M}(m,n+1)$,
+%	such that $M \subset M'$.
+%\end{definition}
+
 Summing up Construction~\ref{con:planar_set_with_minimeter_1} above with~\cite[lemma 3 and proposition 6]{our-vmmsh-2018}

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Remove a redundant lemma
Fix a lemma formulation

@@
 \begin{lemma}
-	\cite[Corollary 1]{our-vmmsh-2018}
-	\label{lem:max_points_on_line}
-	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
-	Then there is no straight line which contains more than $(d+3)/2$ points of $M$.
-\end{lemma}
-
-\begin{lemma}
 	\cite[Lemma 4]{our-vmmsh-2018}
@@
 	\label{lem:intervals_cross}
-	If \textbf{open} line segments $M_1 M_2$ and $M_3 M_4$ does not intersect,
-	$M_1 \neq M_2$, $M_3 \neq M_4$,
+	If \textbf{open} line segments $M_1 M_2$ and $M_3 M_4$ do not intersect,
+	%$M_1 \neq M_2$, $M_3 \neq M_4$,
 	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains not more than 4 points.

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
The main boundary theorem in terms of epsilon

@@
 \end{lemma}
+
+We can improve the constant in Theorem~(TODO: ref) if new exact values of $d(2,n)$ are found;
+however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$.
+To be precise, we have the following
+
+\begin{theorem}
+	For every $\varepsilon > 0$ there exists such a number $k_0$, that for every $k>k_0$ we have the inequality
+	\begin{equation}
+		d(2,k) \geq k\cdot(3^{1/4}\cdot2^{-3/2} - \varepsilon)
+		.
+	\end{equation}
+\end{theorem}
 

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Add observation about known diameters
Cosmetics

@@
 \newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{observation}[theorem]{Observation}
 
@@
 
+Basing on the exact values of $d(2,n)$ for $ 3 \geq n\leq 122$ and the estimate $d(2,123)>10000$
+\cite{kurz2008bounds}, we derive
+\begin{observation}
+	The inequality
+	\begin{equation}
+		d(2,n) \geq 3^{1/4}\cdot2^{-3/2} \cdot n
+	\end{equation}
+	holds for $4 \leq n \leq 21491$.
+\end{observation}
+
 We can improve the constant in Theorem~(TODO: ref) if new exact values of $d(2,n)$ are found;
@@
 \begin{theorem}
-	For every $\varepsilon > 0$ there exists such a number $k_0$, that for every $k>k_0$ we have the inequality
+	For every $\varepsilon > 0$ there exists such a number $n_0$, that the inequality
 	\begin{equation}
-		d(2,k) \geq k\cdot(3^{1/4}\cdot2^{-3/2} - \varepsilon)
-		.
+		d(2,n) \geq n\cdot(3^{1/4}\cdot2^{-3/2} - \varepsilon)
 	\end{equation}
+	holds for every $n>n_0$.
 \end{theorem}

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Linguistic cosmetics
The estimate for varphi_n

@@
 
+Performing some trivial manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
+one can make the following
+\begin{observation}
+	For $n \geq 21492$ we have
+	\begin{equation}
+		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{k-2}}
+		,
+	\end{equation}
+	where
+	\begin{equation}
+		\beta = \frac{1}{\sqrt{21490}} + \sqrt{ \frac{2}{\sqrt{3}} + \frac{1}{21490} } < 1.07464
+		.
+	\end{equation}
+\end{observation}
+
 We can improve the constant in Theorem~(TODO: ref) if new exact values of $d(2,n)$ are found;
@@
 there exists a facher set $M\in\mathfrak{M}(2,n)$ with characteristic $q$~\cite[Theorem 5]{our-vmmsh-2018}.
-In~\cite{antonov2008maximal}, the facher sets of characteristic 1 are called \textit{semi-crabs} and investigated.
+In~\cite{antonov2008maximal}, the facher sets of characteristic 1 were investigated; they were called \textit{semi-crabs}.
 

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Note about points in Construction are distinct

@@
 
+	Note that all the points $M_{J\pm}$ are distinct:
+	the equality $b_J = b_K$ implies $J=K$;
+	the equality $b_J = b_K$ implies $c_J = -c_K$ or $c_J = a / c_k$.
+	The first case is impossible because both $c_J$ and $c_K$ are positive;
+	the second case contradicts with~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4}.
+
 	So, $M = \{ M_{J\pm}, N\}$ is indeed a planar integral point set of $2^k+1$ points,

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Switch language to English
English: such that

@@
 \usepackage[utf8]{inputenc}
-\usepackage[english,russian]{babel} %используем русский и английский языки с переносами
+\usepackage[russian,english]{babel} %используем русский и английский языки с переносами
 \usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
@@
 \begin{document}
-\renewcommand{\bibname}{Список цитированной литературы}
-\renewcommand\refname{\bibname}
+
+%\renewcommand{\bibname}{Список цитированной литературы}
+%\renewcommand\refname{\bibname}
 % !!!
@@
 between points $M_1$ and $M_2$ in a finite-dimensional space $\mathbb{R}^m$.
-An \textit{integral point set} in $m$-dimensional Euclidean space is such a point set $M$ that all the distances between the
+An \textit{integral point set} in $m$-dimensional Euclidean space is a point set $M$ such that all the distances between the
 points of $M$ are integers and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
@@
 \end{definition}
-So, it's impossible to place $k$ points in a unit square such that each pairwise distance of the points is greater than $\varphi_k$.
+So, it's impossible to place $k$ points in a unit square in such a way that each pairwise distance of the points is greater than $\varphi_k$.
 
@@
 \begin{theorem}
-	For every $\varepsilon > 0$ there exists such a number $n_0$, that the inequality
+	For every $\varepsilon > 0$ there exists a number $n_0$ such that the inequality
 	\begin{equation}
@@
 	A set $M\in \mathfrak{M}(m,n)$ is \textit{maximal},
-	if there is no such a set $M'\in \mathfrak{M}(m,n+1)$
-	that $M \subsetneq M'$.
+	if there is no set $M'\in \mathfrak{M}(m,n+1)$
+	such that $M \subsetneq M'$.
 \end{definition}

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Typo fix
Just labels
The main construction of the bound

@@
 
-Basing on the exact values of $d(2,n)$ for $ 3 \geq n\leq 122$ and the estimate $d(2,123)>10000$
+Basing on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
 \cite{kurz2008bounds}, we derive
 \begin{observation}
+	\label{obs:4_leq_n_leq_21491}
 	The inequality
@@
 \begin{observation}
+	\label{varphi_n_where_n_geq_21492}
 	For $n \geq 21492$ we have
 	\begin{equation}
-		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{k-2}}
+		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{n-2}}
 		,
@@
 
-We can improve the constant in Theorem~(TODO: ref) if new exact values of $d(2,n)$ are found;
+In [TODO:cite], TODO: theorem about sublinear quantity of points on the line.
+However, we need the values of all the constants in such an estimate.
+
+\begin{definition}
+	A $\rho(k,M_1 M_2)$-curve, where $-|M_1 - M_2| < k < |M_1 - M_2|$ and $k$ is integer, is the set of points $N$,
+	such that for given points $M_1$ and $M_2$
+	the equality $|N - M_1| - |N-M_2|$ holds.
+\end{definition}
+So, $\rho(k,M_1 M_2)$-curve is a branch of a hyperbola for $k\neq 0$ and the bisector of line segment $M_1 M_2$ for $k=0$.
+
+\begin{observation}
+	\label{obs:rho_curves}
+	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
+	and equality $|M_1 - M_2| = |M_3 - M_4|$ holds but line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
+	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and $\rho(k,M_3 M_4)$-curve
+	do not intersect.
+\end{observation}
+
+\begin{lemma}
+	\label{lem:2k-1_segments}
+	Let $M \in \mathfrak{M}(2,n)$ for some $n$ and let $m$ be a straight line.
+	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$,
+	such that the length of $\Delta_i$ is exactly $k$.
+\end{lemma}
+\begin{proof}
+	Consider a point $N\in M \setminus m$.
+	Then for each $\Delta_i$ there is a $\rho(n_i,\Delta_i)$-curve containing $N$.
+	Due to Observation~\ref{obs:rho_curves}, all $n_i$ are distinct;
+	otherwise $\rho(n_i,\Delta_i)$-curve and $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
+	There can be only $2k-1$ different values for $n_i$,
+	so there are at most $2k-1$ distinct segments $\Delta_i$.
+\end{proof}
+
+TODO: different vs. distinct
+
+\begin{lemma}
+	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n$.
+	Let $|\Delta \cap M| = n^2 + 1$.
+	Then $l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$
+	.
+\end{lemma}
+
+TODO: $|\cdot|$ for length, $\#$ for cardinality?
+
+\begin{proof}
+	$n^2+1$ points, including the ends of $\Delta$, split the segment $\Delta$ into $n^2$
+	sequential segments $\Delta_i$.
+	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
+	at most three segments of length 2 etc.
+	The following two values of sums conclude the proof:
+	\begin{equation}
+		1 + \sum_{k=1}^n (2k-1) = n^2 + 1
+		,
+	\end{equation}
+	\begin{equation}
+		\sum_{k=1}^n k(2k-1) = \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
+		.
+	\end{equation}
+\end{proof}
+
+Now we will estimate the length of a line segment, which shares an arbitrary number of points
+with an integral point set.
+
+\begin{lemma}
+	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n>122$.
+	Let $|\Delta \cap M| = k > 10000$.
+	Then $l\geq \gamma_1 k^{3/2}$,
+	where
+
+	TODO!
+\end{lemma}
+
+This lemma leads to the following observation.
+
+\begin{observation}
+	\label{obs:estimate_points_on_straight_line}
+	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n>21491$.
+	Let $|\Delta \cap M| = k > 10000$.
+	Then $k \leq \gamma_2 b + 6$,
+	where
+
+	TODO!
+\end{observation}
+
+Now we are ready to proof the main theorem of the section.
+
+\begin{theorem}
+	\label{thm:main_estimate}
+	If $n\geq 4$, then $d(2,n) \geq \gamma (n - 2)$,
+	where
+	\begin{equation}
+		\gamma = ... < \frac{3^{1/4}}{2^{3/2}}
+	\end{equation}
+	TODO!
+\end{theorem}
+
+\begin{proof}
+	For $4 \leq n \leq 21491$, the claim of the theorem follows immediately from Observation~\ref{obs:4_leq_n_leq_21491}.
+	Let us consider $M\in \mathfrak{M}(2,n)$, $n \geq 21492$, $\operatorname{diam} M = b$.
+	Lemma~\ref{lem:square_container} yields that $M$ is situated is a square with side length $b$.
+	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1-M_2|$ and $|M_3-M_4|$
+	are minimal in $M$ ($M_2$ and $M_3$ may be the same point).
+	Then $|M_1-M_2| \leq b\varphi_{n-1}$ and $|M_3-M_4| \leq b\varphi_{n-1}$.
+	Due to Lemma~\ref{lemma:quadr_diag_edges}, open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect
+	(otherwise they are not minimal).
+
+	Let $C = cr(M_1 M_2) \cap cr(M_3 M_4)$.
+	Each point of $n\in M$ satisfies one of the following conditions:
+
+	a) $N$ belongs to $C$~--- totally at most $\gamma_2 b + 6$ points due to Lemma~\ref{lem:intervals_cross} and Observation~\ref{obs:estimate_points_on_straight_line};
+
+	b) $N$ belongs to the intersection of one of $|M_1 - M_2| - 1$ hyperbolas
+	with one of $|M_3 - M_4| - 1$ hyperbolas~--- totally at most $4 (|M_1 - M_2| - 1)(|M_3 - M_4| - 1)$ points;
+
+	c) $N$ belongs to the intersection of one of $|M_1 - M_2| - 1$ hyperbolas and $cr(M_3 M_4)$~---
+	totally at most $4 (|M_1 - M_2| - 1)$ points;
+
+	d) $N$ belongs to the intersection of one of $|M_3 - M_4| - 1$ hyperbolas and $cr(M_1 M_2)$~---
+	totally at most $4 (|M_3 - M_4| - 1)$ points;
+
+	(see~\cite[]{erdos1945integral} for details).
+
+	Summing up the cases above, we get the following estimate:
+	\begin{equation}
+		\label{eq:estimate_n_varphi_n_squared}
+		n \leq 4 b^2 \varphi_{n-1}^2 - 4 + \gamma_2 b + 6
+		.
+	\end{equation}
+	Observation~\ref{varphi_n_where_n_geq_21492} turns estimate~\eqref{eq:estimate_n_varphi_n_squared}
+	into the following:
+	\begin{equation}
+		\label{eq:estimate_n_sqrt_n-2}
+		n-2 \leq 4 b^2 \frac{\beta^2}{n-2} + \gamma_2 b
+		,
+	\end{equation}
+	which obviously leads to inequality
+	\begin{equation}
+		1 \leq 4\beta^2 \left(\frac{b}{n-2}\right)^2  + \gamma_2 \frac{b}{n-2}
+		.
+	\end{equation}
+	Let us denote $\lambda = b/(n-2)$ and solve the following square inequality with respect to $\lambda$:
+	\begin{equation}
+		\label{eq:square_inequality_lambda}
+		4\beta^2 \lambda^2  + \gamma_2 \lambda - 1 \geq 0
+		.
+	\end{equation}
+	The discriminant is $\gamma_2^2 + 16 \cdot \beta^2$,
+	so we obtain the following estimate:
+	\begin{equation}
+		\frac{b}{n-2} = \lambda \geq \frac{-\gamma_2 + \sqrt{\gamma_2^2 + 16 \cdot \beta^2}}{8\beta^2}
+		.
+	\end{equation}
+	Calculating the expression above for our $\beta$ and $\gamma_2$, we conclude the proof.
+\end{proof}
+
+We can improve the constant in Theorem~\ref{thm:main_estimate} if new exact values of $d(2,n)$ are found;
 however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$.

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Rewrite the definition of characteristic

@@
 	\cite{kurz2005characteristic}
-	A \textit{characteristic} of an integral point set $M$ is such a squarefree number $q$,
-	that for any points $M_1, M_2, M_3 \in M$ the area of triangle $M_1 M_2 M_3$
-	is $p\sqrt{q}$, where $p$ is rational.
+	Let $M$ be a planar integral point set.
+	A squarefree number $q$ is called the \textit{characteristic} of $M$,
+	if for any points $M_1, M_2, M_3 \in M$ the area of triangle $M_1 M_2 M_3$
+	is $p_{1,2,3}\sqrt{q}$ for some rational $p_{1,2,3}$.
 \end{definition}

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Updating TODO notes
Iteration: typos and cosmetics

@@
 
-\begin{problem}[Point Packing in a Square (PPS)\cite{locatelli2002packing,costa2013valid}]
+\begin{problem}[Point Packing in a Square (PPS)~\cite{locatelli2002packing,costa2013valid}]
 	\label{problem:PPS}
@@
 
-Performing some trivial manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
+Performing some simple manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
 one can make the following
@@
 	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
-	at most three segments of length 2 etc.
+	at most three segments of length 2, etc.
 	The following two values of sums conclude the proof:
@@
 	Let us consider $M\in \mathfrak{M}(2,n)$, $n \geq 21492$, $\operatorname{diam} M = b$.
-	Lemma~\ref{lem:square_container} yields that $M$ is situated is a square with side length $b$.
+	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square with side length $b$.
 	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1-M_2|$ and $|M_3-M_4|$
@@
 	Let $C = cr(M_1 M_2) \cap cr(M_3 M_4)$.
-	Each point of $n\in M$ satisfies one of the following conditions:
+	Each point of $N\in M$ satisfies one of the following conditions:
 
@@
 
-For every integer $n\geq 3$ Solymosi~\cite{solymosi2003note} gives a construction of a facher integral point set
+For every integer $n\geq 3$ Solymosi presented~\cite{solymosi2003note} a construction of a facher integral point set
 $M\in\mathfrak{M}(2,n)$
@@
 	the equality $b_J =  b_K$ implies $J=K$;
-	the equality $b_J = b_K$ implies $c_J = -c_K$ or $c_J = a / c_k$.
+	the equality $b_J = -b_K$ implies $c_J = -c_K$ or $c_J = a / c_K$.
 	The first case is impossible because both $c_J$ and $c_K$ are positive;
@@
 	Reminding that $k$ can be taken arbitrary large, we conclude that we can construct
-	a planar integral point set of arbitary large count of points so that the distance 1 occurs in that set.
+	a planar integral point set of arbitrary large count of points so that the distance 1 occurs in that set.
 \end{construction}
-
-TODO: recalculate the cardinality.
 
@@
 
-TODO: should I write translated version of the results here?
+TODO: write the translated version of the results here
 

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
irreducible -> prime

@@
 \begin{definition}
-	We will call an integral point set $M$ \textit{irreducible}, if the greatest common divisor
+	\cite{noll1989nclusters}
+	We will call an integral point set $M$ \textit{prime}, if the greatest common divisor
 	of all the distances occuring in $M$ is 1.
@@
 
-If an integral point set is irreducible,
+If an integral point set is prime,
 then it cannot be squashed to an integral point set of the same power and structure but smaller diameter.
@@
 \begin{hypothesis}
-	\label{hyp:irreducible_planar}
-	For every $m \geq 2$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{irreducible}
+	\label{hyp:prime_planar}
+	For every $m \geq 2$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{prime}
 	integral point set $M$ of $n$ points in $m$-dimensional Euclidean space
@@
 
-For $m \geq 3$, Hypothesis~\ref{hyp:irreducible_planar} can be easily proved:
+For $m \geq 3$, Hypothesis~\ref{hyp:prime_planar} can be easily proved:
 indeed, we can just take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
@@
 So, the set consists of $m+1$ or more points and has distance 1 occuring in it.
-As distance 1 occurs in the obtained set, it is irreducible.
+As distance 1 occurs in the obtained set, it is prime.
 
-So, Hypothesis~\ref{hyp:irreducible_planar} needs proof only for $m=2$.
+So, Hypothesis~\ref{hyp:prime_planar} needs proof only for $m=2$.
 

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Cardinality: |...| -> #(...)
TODO note
Cardinality notation - one more place
Switch from |A-B| distance notation to |AB| in order to apply it to line segments

@@
 \section{Introduction}
-Let $\mathbb{N}$ be the set of all positive integers and let $|M_1 - M_2|$ denote the Euclidean distance
-between points $M_1$ and $M_2$ in a finite-dimensional space $\mathbb{R}^m$.
+Let $\mathbb{N}$ be the set of all positive integers and let $|M_1 M_2|$ denote the Euclidean distance
+between points $M_1$ and $M_2$ in a finite-dimensional space $\mathbb{R}^m$
+(and, more generally, let $|\Delta|$ denote the length of line segment $\Delta$).
 An \textit{integral point set} in $m$-dimensional Euclidean space is a point set $M$ such that all the distances between the
@@
 $\mathfrak{M}(m,n)$ (following the notation of~\cite{our-vmmsh-2018}).
-The symbol $|M|$ will be used for cardinality of $M$, i.e. the quantity of points in $M$ in our case.
+The symbol $\# M$ will be used for cardinality of $M$, i.e. the quantity of points in $M$ in our case.
 
@@
 \begin{equation}
-	\operatorname{diam} M = \max_{A,B\in M} |A-B|
+	\operatorname{diam} M = \max_{A,B\in M} |AB|
 	.
@@
 \begin{equation}
-	d(m,n) = \min_{M\in\mathfrak{M}(m,n)} \operatorname{diam} M = \min_{M\in\mathfrak{M}(m,n)} \max_{A,B\in M} |A-B|
+	d(m,n) = \min_{M\in\mathfrak{M}(m,n)} \operatorname{diam} M = \min_{M\in\mathfrak{M}(m,n)} \max_{A,B\in M} |AB|
 	.
@@
 \begin{equation}
-	l(m,n) = \min_{M\in\mathfrak{M}(m,n)} \min_{A,B\in M} |A-B|
+	l(m,n) = \min_{M\in\mathfrak{M}(m,n)} \min_{A,B\in M} |AB|
 	.
@@
 \begin{definition}
-	A $\rho(k,M_1 M_2)$-curve, where $-|M_1 - M_2| < k < |M_1 - M_2|$ and $k$ is integer, is the set of points $N$,
+	A $\rho(k,M_1 M_2)$-curve, where $-|M_1 M_2| < k < |M_1 M_2|$ and $k$ is integer, is the set of points $N$,
 	such that for given points $M_1$ and $M_2$
-	the equality $|N - M_1| - |N-M_2|$ holds.
+	the equality $|N M_1| - |N M_2|$ holds.
 \end{definition}
@@
 	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
-	and equality $|M_1 - M_2| = |M_3 - M_4|$ holds but line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
+	and equality $|M_1 M_2| = |M_3 M_4|$ holds but line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
 	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and $\rho(k,M_3 M_4)$-curve
@@
 	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n$.
-	Let $|\Delta \cap M| = n^2 + 1$.
+	Let $\#(\Delta \cap M) = n^2 + 1$.
 	Then $l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$
@@
 	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n>122$.
-	Let $|\Delta \cap M| = k > 10000$.
+	Let $\#(\Delta \cap M) = k > 10000$.
 	Then $l\geq \gamma_1 k^{3/2}$,
+	TODO: why 10 000 ???
 	where
@@
 	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n>21491$.
-	Let $|\Delta \cap M| = k > 10000$.
+	Let $\#(\Delta \cap M) = k > 10000$.
 	Then $k \leq \gamma_2 b + 6$,
@@
 	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square with side length $b$.
-	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1-M_2|$ and $|M_3-M_4|$
+	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1 M_2|$ and $|M_3 M_4|$
 	are minimal in $M$ ($M_2$ and $M_3$ may be the same point).
-	Then $|M_1-M_2| \leq b\varphi_{n-1}$ and $|M_3-M_4| \leq b\varphi_{n-1}$.
+	Then $|M_1 M_2| \leq b\varphi_{n-1}$ and $|M_3 M_4| \leq b\varphi_{n-1}$.
 	Due to Lemma~\ref{lemma:quadr_diag_edges}, open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect
@@
 
-	b) $N$ belongs to the intersection of one of $|M_1 - M_2| - 1$ hyperbolas
-	with one of $|M_3 - M_4| - 1$ hyperbolas~--- totally at most $4 (|M_1 - M_2| - 1)(|M_3 - M_4| - 1)$ points;
+	b) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas
+	with one of $|M_3 M_4| - 1$ hyperbolas~--- totally at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
 
-	c) $N$ belongs to the intersection of one of $|M_1 - M_2| - 1$ hyperbolas and $cr(M_3 M_4)$~---
-	totally at most $4 (|M_1 - M_2| - 1)$ points;
+	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas and $cr(M_3 M_4)$~---
+	totally at most $4 (|M_1 M_2| - 1)$ points;
 
-	d) $N$ belongs to the intersection of one of $|M_3 - M_4| - 1$ hyperbolas and $cr(M_1 M_2)$~---
-	totally at most $4 (|M_3 - M_4| - 1)$ points;
+	d) $N$ belongs to the intersection of one of $|M_3 M_4| - 1$ hyperbolas and $cr(M_1 M_2)$~---
+	totally at most $4 (|M_3 M_4| - 1)$ points;
 
@@
 $M\in\mathfrak{M}(2,n)$
-such that equality $|M_1 - M_2| = 2$ holds for some $M_1, M_2 \in M$.
+such that equality $|M_1 M_2| = 2$ holds for some $M_1, M_2 \in M$.
 The constructed set has both odd and even distances.
@@
 	\begin{multline}
-		|N - M_{J\pm}|
+		|N M_{J\pm}|
 		=
@@
 	\begin{equation}
-		|M_{J_1 \pm} - M_{J_2 \pm}|
+		|M_{J_1 \pm}  M_{J_2 \pm}|
 		=
@@
 	\begin{equation}
-		|M_{H+} - M_{H-}|
+		|M_{H+}  M_{H-}|
 		=
@@
 \begin{hypothesis}
-	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 - M_2|=1$ holds,
+	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,
 	is not optimal.
@@
 	For every $n\geq 3$ there is a planar intergal set $M$ of $n$ points
-	such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds.
+	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds.
 	This set consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.
@@
 	And vice versa, if $M$ is a planar integral point set of $n$ points
-	such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds,
+	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
 	then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
@@
 	there exists $M\in\mathfrak{M}(m,n)$,
-	such that for some $M_1, M_2\in M$ equality $|M_1 - M_2| = d$ holds.
+	such that for some $M_1, M_2\in M$ equality $|M_1 M_2| = d$ holds.
 \end{theorem}

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Remove TODO note that has already been done
Implement a TODO note partially

@@
 In the present paper we improve Solymosi's results:
-firstly, we obtain a greater constant $c$ in Theorem (TODO: the number of theorem and the exact constant),
+firstly, we obtain a greater constant $c$ (TODO: the exact constant) in Theorem~\ref{thm:main_estimate},
 using the combined approach with the Point Packing in a Square problem;
@@
 \end{lemma}
-
-TODO: $|\cdot|$ for length, $\#$ for cardinality?
 

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
The remaining fixes of iteration

@@
 	such that for given points $M_1$ and $M_2$
-	the equality $|N M_1| - |N M_2|$ holds.
+	the equality $|N M_1| - |N M_2| = k$ holds.
 \end{definition}
-So, $\rho(k,M_1 M_2)$-curve is a branch of a hyperbola for $k\neq 0$ and the bisector of line segment $M_1 M_2$ for $k=0$.
+So, a $\rho(k,M_1 M_2)$-curve is a branch of a hyperbola for $k\neq 0$ and the bisector of line segment $M_1 M_2$ for $k=0$.
 

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Hypothesis -> Conjecture

@@
 \newtheorem{theorem}{Theorem}[section]
-\newtheorem{hypothesis}[theorem]{Hypothesis}
+\newtheorem{conjecture}[theorem]{Conjecture}
 \newtheorem{lemma}[theorem]{Lemma}
@@
 
-\begin{hypothesis}
+\begin{conjecture}
 	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,
 	is not optimal.
-\end{hypothesis}
+\end{conjecture}
 
@@
 
-\begin{hypothesis}
+\begin{conjecture}
 	\label{hyp:prime_planar}
@@
 	where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
-\end{hypothesis}
+\end{conjecture}
 
-For $m \geq 3$, Hypothesis~\ref{hyp:prime_planar} can be easily proved:
+For $m \geq 3$, Conjecture~\ref{hyp:prime_planar} can be easily proved:
 indeed, we can just take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
@@
 
-So, Hypothesis~\ref{hyp:prime_planar} needs proof only for $m=2$.
+So, Conjecture~\ref{hyp:prime_planar} needs proof only for $m=2$.
 

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
literally -> naively

@@
 \begin{remark}
-	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ literally,
+	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ naively,
 	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
with side length -> of side length

@@
 	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
-	Then $M$ is situated in a square with side length $d$.
+	Then $M$ is situated in a square of side length $d$.
 \end{lemma}
@@
 	Let us consider $M\in \mathfrak{M}(2,n)$, $n \geq 21492$, $\operatorname{diam} M = b$.
-	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square with side length $b$.
+	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square of side length $b$.
 	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1 M_2|$ and $|M_3 M_4|$
@@
 	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then
-	obtain a right triangle with side length 1.
+	obtain a right triangle of side length 1.
 	This triangle is obviously the optimal set in $\mathfrak{M}(2,3)$.
@@
 with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
-replacing the out-of-line point with an $(m-2)$-dimensional simplex with proper side length.
+replacing the out-of-line point with an $(m-2)$-dimensional simplex of proper side length.
 That simplex will consist of $m-1$ points;

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Use shorter notaton for lengths of line segments

@@
 	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$,
-	such that the length of $\Delta_i$ is exactly $k$.
+	such that $|\Delta_i| = k$.
 \end{lemma}
@@
 \begin{lemma}
-	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n$.
+	Let $\Delta$ be a straight line segment, $|\Delta|=l$ and $M \in \mathfrak{M}(2,n)$ for some $n$.
 	Let $\#(\Delta \cap M) = n^2 + 1$.
@@
 \begin{lemma}
-	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n>122$.
+	Let $\Delta$ be a straight line segment, $|\Delta| = l$ and $M \in \mathfrak{M}(2,n)$ for some $n>122$.
 	Let $\#(\Delta \cap M) = k > 10000$.
@@
 	\label{obs:estimate_points_on_straight_line}
-	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n>21491$.
+	Let $\Delta$ be a straight line segment, $|\Delta| = l$ and $M \in \mathfrak{M}(2,n)$ for some $n>21491$.
 	Let $\#(\Delta \cap M) = k > 10000$.

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Cosmetics
Code cosmetics
Add construction with unique minimum edge

@@
 	Let $\#(\Delta \cap M) = n^2 + 1$.
-	Then $l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$
+	Then
+	\begin{equation}
+		l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
 		.
+	\end{equation}
 \end{lemma}
@@
 
-	(see~\cite[]{erdos1945integral} for details).
-
+	(see~\cite{erdos1945integral} for details).
 	Summing up the cases above, we get the following estimate:
@@
 
-TODO: construction with unique minimum edge?
-
 \begin{definition}
@@
 So, Conjecture~\ref{hyp:prime_planar} needs proof only for $m=2$.
+For $m>2$, a bit more strong result can also be claimed.
+
+\begin{theorem}
+	For arbitrary integers $m > 2$, $n \geq m+1$, $d \geq 1$
+	there exists a prime set $M\in\mathfrak{M}(m,n)$,
+	such that for some $M_1, M_2\in M$ equality $|M_1 M_2| = d$ holds,
+	the distance $d$ is minimal and occurs in $M$ only once.
+\end{theorem}
+
+\begin{proof}
+	Exploit Construction~\ref{con:planar_set_with_minimeter_1} to get $M'\in \mathfrak{M}(2,n-m+2)$
+	with $M'_1, M'_2 \in M'$ and $|M'_1 M'_2| = 1$.
+	Dilate $M'$ to $M''$ in such a way that $M'_1$ and $M'_2$ turn into $M''_1$ and $M''_2$ resp.
+	and $|M''_1 M''_2| = d$.
+	Then ``blow up'' $M''$ to $M\in \mathfrak{M}(m,n)$ using a simplex of side length $d+1$.
+	The greatest common divisor of $d$ and $d+1$ is $1$, so $M$ is prime.
+\end{proof}
 

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Rephrase a theorem to use a shorter notation

@@
 	For every $m \geq 2$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{prime}
-	integral point set $M$ of $n$ points in $m$-dimensional Euclidean space
-	where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
+	integral point set $M\in\mathfrak{M}(m,n)$ which
+	contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
 \end{conjecture}

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Give the exact value if points in intersection of crosses
Rephrase the definition of \rho-curve

@@
 	%$M_1 \neq M_2$, $M_3 \neq M_4$,
-	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains not more than 4 points.
+	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains 2 or 4 points.
 \end{lemma}
@@
 \begin{definition}
-	A $\rho(k,M_1 M_2)$-curve, where $-|M_1 M_2| < k < |M_1 M_2|$ and $k$ is integer, is the set of points $N$,
-	such that for given points $M_1$ and $M_2$
-	the equality $|N M_1| - |N M_2| = k$ holds.
+	For given line segment $M_1 M_2$ and an integer $k$, $-|M_1 M_2| < k < |M_1 M_2|$,
+	we define a $\rho(k,M_1 M_2)$-curve as the set of points $N$
+	such that the equality $|N M_1| - |N M_2| = k$ holds.
 \end{definition}

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
A note

@@
 firstly, we obtain a greater constant $c$ (TODO: the exact constant) in Theorem~\ref{thm:main_estimate},
-using the combined approach with the Point Packing in a Square problem;
+using the combined approach with the Point Packing in a Square problem
+(we notice that such an approach is different from Solymosi's one);
 then, we prove that $l(m,n)=1$ for all possible $m$ and $n$.

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
A note
A label

@@
 However, we need the values of all the constants in such an estimate.
+To obtain such values, we will now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.
 
@@
 \begin{conjecture}
+	\label{con:no_optimal_with_edge_1}
 	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
Add motivation for a conjecture

@@
 
+The motivation of Conjecture~\ref{con:no_optimal_with_edge_1} is based on the results of~\cite[Section 5]{kurz2008minimum},
+especially the following theorem.
+
+\begin{theorem}
+	For every $12 \leq n \leq 122$, there exist a facher optimal sets $M_n\in\mathfrak{M}(2,n)$
+	that consist of $n-1$ points $\{ (b_1,0), ..., (b_{n-1},0)\}$
+	and the point $(0,\sqrt{A_n})$,
+	where $b_1,...,b_{n-1}, A_n$ are integer and $\sqrt{A_n}\notin\mathbb{N}$.
+\end{theorem}
+Such a set can not contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar} below,
+the points with distance 1 should be $(\pm 1/2, 0)$.
+
 %Now we need two more definitions.

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
Translated version of the involved results

@@
 \newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
 \newtheorem{observation}[theorem]{Observation}
@@
 
-Summing up Construction~\ref{con:planar_set_with_minimeter_1} above with~\cite[lemma 3 and proposition 6]{our-vmmsh-2018}
-and~\cite[Section 6]{antonov2008maximal},
+If the take $k=1$ in Lemma~\ref{lem:2k-1_segments},
+we obtan the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018}).
+\begin{corollary}
+	\label{cor:only_one_distance_1_on_straight_line}
+	Let $M \in \mathfrak{M}(2,n)$ for some $n$ and let $m$ be a straight line.
+	Then there is at most one pair of points $M_1,M_2\in M \cap m$
+	such that $|M_1 M_2| = 1$.
+\end{corollary}
+
+To describe all the sets $M \in \mathfrak{M}(2,n)$ with distance 1,
+we will need~\cite[Proposition 6]{our-vmmsh-2018}.
+For the sake of completeness, we will give here a slightly rephrased version of it with the proof as
+\begin{lemma}
+	\label{lem:no_4_points_in_semigeneral_position_with_distance_1}
+	Let $M=\{M_1,M_2,M_3,M_4\}\in \mathfrak{M}(2,4)$ such that $|M_1 M_2|=1$
+	and let $l$ be the straight line containing points $M_1$ and $M_2$.
+	Then one of the points $M_3$ or $M_4$ belongs to $l$.
+\end{lemma}
+
+\begin{proof}
+	Let $m$ denote the bisector of line segment $M_1 M_2$.
+	Due to the triangle inequality, there are two possibilities to place each of points $M_i$, $i=3,4$:
+
+	a) $M_i$ belongs to $l$ and $|M_i M_1|-|M_i M_2| = \pm 1$;
+
+	b) $M_i$ belongs to $m$ and $|M_i M_1| = |M_i M_2|$.
+
+	If (a) is true for both points, then $M\subset l$ and thus $M$ is not an integral point set.
+	If (a) is true for one point and (b) is true for another, then the claim of the lemma follows.
+	So, suppose the contrary, i.e. both points $M_3$ and $M_4$ belong to $m$.
+
+	The area of the triangle $M_1 M_3 M_4$ is rational because $|M_3 - M_4| \in \mathbb{Z}$.
+	Thus, the set $M$ is of characteristic 1,
+	and there is a Cartesian coordinate system such that $M_1=(-1/2,0)$, $M_2=(1/2,0)$, $M_3=(0, a/2)$.
+	(TODO: ref?)
+	It is clear that $a\neq 0$.
+	We set $b = |M_1 - M_3|$ and $O=(0,0)$.
+	Applying the Pythagorean theorem to triangle $OM_1M_3$, we obtain the following Diophantine equation:
+	\begin{equation}
+		\frac{1}{4} + \frac{a^2}{4} = b^2
+		,
+	\end{equation}
+	or, which is the same,
+	\begin{equation}
+		1 + a^2 = (2b)^2
+		,
+	\end{equation}
+	that has no integral solutions.
+	This contradiction concludes the proof.
+\end{proof}
+
+Using Construction~\ref{con:planar_set_with_minimeter_1},
+Lemma~\ref{lem:no_4_points_in_semigeneral_position_with_distance_1} and Corollary~\ref{cor:only_one_distance_1_on_straight_line}
+together with results of~\cite[Section 6]{antonov2008maximal},
 we obtain the following
-
-TODO: write the translated version of the results here
 

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
Just a label
The first estimate for a line segment

@@
 \begin{lemma}
+	\label{lem:line_segment_with n_squared_plus_one_points}
 	Let $\Delta$ be a straight line segment, $|\Delta|=l$ and $M \in \mathfrak{M}(2,n)$ for some $n$.
@@
 \begin{lemma}
-	Let $\Delta$ be a straight line segment, $|\Delta| = l$ and $M \in \mathfrak{M}(2,n)$ for some $n>122$.
-	Let $\#(\Delta \cap M) = k > 10000$.
-	Then $l\geq \gamma_1 k^{3/2}$,
-	TODO: why 10 000 ???
-	where
+	Let $\Delta$ be a straight line segment, $|\Delta| = l$ and $M \in \mathfrak{M}(2,n)$.
+	Let $\#(\Delta \cap M) = t$.
+	Then $l\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}$.
+\end{lemma}
 
-	TODO!
-\end{lemma}
+\begin{proof}
+	Let $f(k)$ denote the mininal length of a line segment
+	that shares $k$ points with and integral point set.
+	We observe that $f(k) > f (k-1)$.
+	Due to Lemma~\ref{lem:line_segment_with n_squared_plus_one_points},
+	$f(n^2+1) \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$.
+
+	Consider $t$ such that $n^2 + 1 \leq t \leq (n+1)^2$. %,
+	%then $f(t) > f(n^2 +1)$.% and $n \geq \sqrt{t} - 1$.
+	For $t\in\mathbb{N}$ the inequality $(\sqrt{t} - 1)^2 +1 \leq t$ holds,
+	thus
+	%and derive that
+	\begin{equation}
+		\label{eq:line_segment_length}
+		f(t) \geq f((\sqrt{t} - 1)^2 +1) \geq \frac{2}{3}(\sqrt{t} - 1)^3+\frac{1}{2}(\sqrt{t} - 1)^2-\frac{1}{6}(\sqrt{t} - 1)
+		=
+		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2}
+		.
+	\end{equation}
+\end{proof}
 

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
Language issues to be resolved before publication
Rephrase a bit

@@
 \begin{definition}
-	For given line segment $M_1 M_2$ and an integer $k$, $-|M_1 M_2| < k < |M_1 M_2|$,
+	For a line segment $M_1 M_2$ and an integer $k$, $-|M_1 M_2| < k < |M_1 M_2|$,
 	we define a $\rho(k,M_1 M_2)$-curve as the set of points $N$
-	such that the equality $|N M_1| - |N M_2| = k$ holds.
+	for which the equality $|N M_1| - |N M_2| = k$ holds.
 \end{definition}
-So, a $\rho(k,M_1 M_2)$-curve is a branch of a hyperbola for $k\neq 0$ and the bisector of line segment $M_1 M_2$ for $k=0$.
+So, in the planar case a $\rho(k,M_1 M_2)$-curve
+is a branch of a hyperbola for $k\neq 0$ and the bisector of line segment $M_1 M_2$ for $k=0$.
 
@@
 \end{proof}
-
-TODO: different vs. distinct
 
@@
 
+\section{Language issues to be resolved before publication}
+
+\begin{enumerate}
+	\item
+		there are at most $2k-1$ distinct segments $\Delta_i$
+
+		different vs. distinct ?
+
+	\item
+		Summing up
+
+		is it correct to say so ?
+
+	\item
+		... the following \\ Theorem X.
+
+		is it correct to split the sentence in such a way?
+
+	\item
+		For the sake of completeness.
+
+		Is the phrase correct?
+		There should be an example in one of papers about Banach limits...
+
+	\item
+		to belong vs. to be situated
+
+	\item
+		intersection of ... and ...
+		OR
+		intersection of ... with ...
+		?
+
+\end{enumerate}
+
 \section{Acknowledgements}

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
Final remarks and open problems - the first one
square inequality -> quadratic inequality

@@
 	\end{equation}
-	Let us denote $\lambda = b/(n-2)$ and solve the following square inequality with respect to $\lambda$:
+	Let us denote $\lambda = b/(n-2)$ and solve the following quadratic inequality with respect to $\lambda$:
 	\begin{equation}
@@
 
+\section{Final remarks and open problems}
+
+\begin{remark}
+	The estimate of Theorem~\ref{thm:main_estimate} is not tight.
+	The derivation of tight bounds for the minimum diameter $d(2, n)$
+	is still a challenging task for the future~\cite[Section 7]{kurz2008minimum}.
+\end{remark}
+TODO: should I give the known upper bound here?
+
 \section{Language issues to be resolved before publication}

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
which -> that (due to importance)
Edit a lemma cosmetically

@@
 
-Now we will estimate the length of a line segment, which shares an arbitrary number of points
+Now we will estimate the length of a line segment that shares an arbitrary number of points
 with an integral point set.
@@
 \begin{lemma}
-	Let $\Delta$ be a straight line segment, $|\Delta| = l$ and $M \in \mathfrak{M}(2,n)$.
+	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,n)$ for some $n$.
 	Let $\#(\Delta \cap M) = t$.
-	Then $l\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}$.
+	Then
+	\begin{equation}
+		\label{eq:estimate_for_segment_length}
+		b\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}
+		.
+	\end{equation}
 \end{lemma}

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
Shortify at one place
Thelinear estimate for the quantity of points on a line segment

@@
 
-	Consider $t$ such that $n^2 + 1 \leq t \leq (n+1)^2$. %,
+	%Consider $t$ such that $n^2 + 1 \leq t \leq (n+1)^2$. %,
 	%then $f(t) > f(n^2 +1)$.% and $n \geq \sqrt{t} - 1$.
@@
 	\label{obs:estimate_points_on_straight_line}
-	Let $\Delta$ be a straight line segment, $|\Delta| = l$ and $M \in \mathfrak{M}(2,n)$ for some $n>21491$.
-	Let $\#(\Delta \cap M) = k > 10000$.
+	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,n)$ for some $n$.
+	Let $\#(\Delta \cap M) = k$ and $b>10000$.
 	Then $k \leq \gamma_2 b + 6$,
 	where
-
-	TODO!
+	\begin{equation}
+		\gamma_2 = \frac{3846}{2593 \sqrt{647}-5823}
+	\end{equation}
 \end{observation}
 
-Now we are ready to proof the main theorem of the section.
+\begin{proof}
+	We are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
+	yields to $b > 10000$, because we know the maximum quantity of points for integral point sets of diameters at most $10000$.
+	So, let us consider $t\geq 647$ and find a coefficient $\gamma_2$,
+	such that the inequality
+	\begin{equation}
+		\label{eq:inequality_for_linear_estimate_of_segment_length}
+		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2} \geq \frac{t-6}{\gamma_2}
+	\end{equation}
+	holds for all $t\geq 647$.
+	For these $t$,
+	the left part of~\eqref{eq:inequality_for_linear_estimate_of_segment_length} obviously grows faster than the right one.
+	Turning~\eqref{eq:inequality_for_linear_estimate_of_segment_length} into the same equality
+	and solving it for $t=647$, we obtain the claimed estimate.
+\end{proof}
+
+Now we are ready to prove the main theorem of the section.
 

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
The exact value of coefficient in the main estimate
Uniform linear bound

@@
 In the present paper we improve Solymosi's results:
-firstly, we obtain a greater constant $c$ (TODO: the exact constant) in Theorem~\ref{thm:main_estimate},
+firstly, we obtain a greater constant $c = 5/11$ in Theorem~\ref{thm:main_estimate},
 using the combined approach with the Point Packing in a Square problem
@@
 	where
-	\begin{equation}
-		\gamma = ... < \frac{3^{1/4}}{2^{3/2}}
-	\end{equation}
-	TODO!
+	\begin{multline}
+		\frac{3^{1/4}}{2^{3/2}} >
+		\gamma = \frac{\sqrt{16 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}+\frac{14791716}{{{\left( 2593 \sqrt{647}-5823\right) }^{2}}}}-\frac{3846}{2593 \sqrt{647}-5823}}{8 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}}
+		>\\
+		> 0.45557
+		> \frac{5}{11}
+	\end{multline}
 \end{theorem}
@@
 	Calculating the expression above for our $\beta$ and $\gamma_2$, we conclude the proof.
+\end{proof}
+
+\begin{corollary}
+	If $n\geq 4$, then $d(2,n) > \frac{5}{11} n$.
+\end{corollary}
+
+\begin{proof}
+	For $4 \leq n \leq 21491$, the claim follows from Observation~\ref{obs:4_leq_n_leq_21491} immediately.
+	For $n > 21491$, the inequality $0.45557(n-2) > \frac{5}{11}n$ holds.
 \end{proof}

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
Rephrase a bit

@@
 using the combined approach with the Point Packing in a Square problem
-(we notice that such an approach is different from Solymosi's one);
+(this approach is different from Solymosi's one);
 then, we prove that $l(m,n)=1$ for all possible $m$ and $n$.

### 2019-06-01 pps-linear-bound/pps-linear-lower-bound.tex
Kurz's estimate of the diameter for sets with many collinear points

@@
 
-In [TODO:cite], TODO: theorem about sublinear quantity of points on the line.
-However, we need the values of all the constants in such an estimate.
-To obtain such values, we will now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.
+Now we need to estimate the quanity of points in an intersection of an integral point set and a line segment.
+There is the following estimate for minimum diameter of planar integral point sets with many collinear points:
+\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
+	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
+	at least $n^\delta$ collinear points there exists a $n_0 (\varepsilon)$
+	such that for all $n \geq n_0 (\varepsilon)$ we have
+	\begin{equation}
+		\operatorname{diam} P \geq n^{\frac{\delta}{4 log 2(1+\varepsilon)}\log \log n}
+		.
+	\end{equation}
+\end{theorem}
+However, the estimate is rather unsuitable for our needs,
+as it does not provide the values of all the constants.
+To obtain the desired estimate, we will now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.
 

### 2019-06-02 pps-linear-bound/pps-linear-lower-bound.tex
Labels

@@
 \begin{remark}
+	\label{rem:planar_set_with_minimeter_1_of_3_points}
 	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ naively,
@@
 \begin{remark}
+	\label{rem:planar_set_with_minimeter_1_of_4_and_5_points}
 	For $k=2$ in Construction~\ref{con:planar_set_with_minimeter_1},

### 2019-06-02 pps-linear-bound/pps-linear-lower-bound.tex
Language problems
Putting all the conjectures together

@@
 
-\begin{conjecture}
-	\label{con:no_optimal_with_edge_1}
-	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,
-	is not optimal.
-\end{conjecture}
-
-The motivation of Conjecture~\ref{con:no_optimal_with_edge_1} is based on the results of~\cite[Section 5]{kurz2008minimum},
-especially the following theorem.
-
-\begin{theorem}
-	For every $12 \leq n \leq 122$, there exist a facher optimal sets $M_n\in\mathfrak{M}(2,n)$
-	that consist of $n-1$ points $\{ (b_1,0), ..., (b_{n-1},0)\}$
-	and the point $(0,\sqrt{A_n})$,
-	where $b_1,...,b_{n-1}, A_n$ are integer and $\sqrt{A_n}\notin\mathbb{N}$.
-\end{theorem}
-Such a set can not contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar} below,
-the points with distance 1 should be $(\pm 1/2, 0)$.
-
 %Now we need two more definitions.
@@
 \end{remark}
-TODO: should I give the known upper bound here?
+The known upper bound is $d(2,n)\leq 2^{c \log n \log \log n}$~\cite{harborth1993upper}.
+However, we hope that the present article can give a framework for future better estimates.
+
+\begin{conjecture}
+	The approach of Theorem~\ref{thm:main_estimate} can be generalized to higher dimensions.
+\end{conjecture}
+For the overview of current lower bounds for higher dimensions, we refer the reader to~\cite{nozaki2013lower}.
+
+There are also the following ways to improve the current bound,
+based on Lemma~\ref{lem:square_container}.
+
+\begin{problem}
+	\label{prb:square_side_length_and_diameter_power}
+	Is there a number $\delta < 1$ such that any planar integral point set $M$ with $\operatorname{diam} M = d$
+	belongs to a square of side length $d^\delta$?
+\end{problem}
+If the answer for Problem~\ref{prb:square_side_length_and_diameter_power} is affirmative,
+then we can build an estimate better than linear.
+
+\begin{problem}
+	\label{prb:square_side_length_and_diameter_linear}
+	What are the maximal numbers $\delta \leq 1$ and $h \leq 0$
+	such that any planar integral point set $M$ with $\operatorname{diam} M = d$
+	belongs to a square of side length $\delta d - h$?
+\end{problem}
+
+If the answer for Problem~\ref{prb:square_side_length_and_diameter_linear} is not $\delta = 1$ and $h = 0$,
+then the linear estimate can be slightly improved.
+
+However, Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}
+do not seem to show the way for improving the estimate
+due to the following well-known theorem (TODO: refs!).
+\begin{theorem}
+	\label{thm:rational_set_on_circle}
+	There exists a dense subset $P$ of the unit circle such that for any $P_1, P_2 \in P$
+	the distance $|P_1 P_2|$ is rational.
+\end{theorem}
+We conjecture that thorough investigation of existing examples of circular integral point sets
+(TODO: refs!)
+will lead to the negative answers for the
+Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}.
+
+Another possibility to increase the constant in out lower bound is introduced by the following problem.
+
+\begin{problem}
+	\label{prb:minimal_shape}
+	What is the minimal shape $\mathcal{S}$ such that any planar integral point set $M$ with $\operatorname{diam} M = d$
+	belongs to $\mathcal{S}$?
+\end{problem}
+This problem in another generalization of Lemma~\ref{lem:square_container}.
+We know that $\mathcal{S}$ is not a circle of diameter $d$
+(the counterexample is a right triangle of side length $d$)
+and we conjecture that $\mathcal{S}$ is not a Reuleaux triangle of width $d$
+(the probable counterexample belongs to a circle of diameter $d$, see Theorem~\ref{thm:rational_set_on_circle}).
+
+The solution of Problem~\ref{prb:minimal_shape} (if ever found) will lead to sophisticated packing problem,
+something like ``How to put $n$ points in a convex hull of concentric circle and Reuleaux triangle?''
+Obviously, Problem~\ref{prb:minimal_shape} can be generalized to higher dimensions.
+
+For $n=3,4,5$, there are optimal sets in $\mathfrak{M}(2,n)$ that contain distance 1
+(see Remarks~\ref{rem:planar_set_with_minimeter_1_of_3_points} and~\ref{rem:planar_set_with_minimeter_1_of_4_and_5_points}).
+However, we can put forward the following conjecture.
+\begin{conjecture}
+	\label{con:no_optimal_with_edge_1}
+	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,
+	is not optimal.
+\end{conjecture}
+
+The motivation of Conjecture~\ref{con:no_optimal_with_edge_1} is based on the results of~\cite[Section 5]{kurz2008minimum},
+especially the following theorem.
+
+\begin{theorem}
+	For every $12 \leq n \leq 122$, there exist a facher optimal sets $M_n\in\mathfrak{M}(2,n)$
+	that consist of $n-1$ points $\{ (b_1,0), ..., (b_{n-1},0)\}$
+	and the point $(0,\sqrt{A_n})$,
+	where $b_1,...,b_{n-1}, A_n$ are integer and $\sqrt{A_n}\notin\mathbb{N}$.
+\end{theorem}
+Such a set can not contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar},
+the points with distance 1 should be $(\pm 1/2, 0)$.
+
+When we talk about integral point sets containing distance 1, another problem arises.
+
+\begin{problem}
+	For $M\in\mathfrak{M}(m,n)$, how many times can the distance 1 occur in $M$?
+\end{problem}
+The answer for $m=2$ is given by Theorem~\ref{thm:minimeter_1_planar}.
+For greater $m$, we will be brave enough to put forward
+\begin{conjecture}
+	Consider $M\in\mathfrak{M}(m,n)$, $M_1,M_2\in M$, $|M_1 M_2| =1$.
+	Then either $n = m+1$ or $M$ is a ``blowup'' of a facher set and distance 1 occurs in $m$
+	not more than $1+\frac{m^2 - m}{2}$ times.
+\end{conjecture}
+Such as estimate is based on the quantity of edges in $(m-1)$-dimensional simplex.
 
@@
 
+	\item
+		Observation vs. Proposition
+
+	\item
+		estimate vs. bound
+
 \end{enumerate}

### 2019-06-06 pps-linear-bound/pps-linear-lower-bound.tex
Move the last conjecture to the appropriate section

@@
 
-\begin{conjecture}
-	\label{hyp:prime_planar}
-	For every $m \geq 2$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{prime}
+\begin{theorem}
+	\label{thm:prime_with_desired_distance}
+	For every $m \geq 3$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{prime}
 	integral point set $M\in\mathfrak{M}(m,n)$ which
 	contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
-\end{conjecture}
+\end{theorem}
 
-For $m \geq 3$, Conjecture~\ref{hyp:prime_planar} can be easily proved:
-indeed, we can just take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
+\begin{proof}
+	Take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
 	with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
@@
 	As distance 1 occurs in the obtained set, it is prime.
+\end{proof}
 
-So, Conjecture~\ref{hyp:prime_planar} needs proof only for $m=2$.
-For $m>2$, a bit more strong result can also be claimed.
+A bit more strong result can also be claimed.
 
 \begin{theorem}
+	\label{thm:prime_with_unique_desired_distance}
 	For arbitrary integers $m > 2$, $n \geq m+1$, $d \geq 1$
@@
 \end{conjecture}
-Such as estimate is based on the quantity of edges in $(m-1)$-dimensional simplex.
+Such as estimate is based on the quantity of edges in $(m-2)$-dimensional simplex.
+
+We want to end the list of conjectures with the one which looks the easiest to deal with.
+
+\begin{conjecture}
+	\label{hyp:prime_planar}
+	For every $n \geq 3$, $d \geq 1$ there exists an \textbf{prime}
+	set $M\in\mathfrak{M}(2,n)$ which
+	contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
+\end{conjecture}
+This conjecture gives the same claim with Theorem~\ref{thm:prime_with_desired_distance}, but for $m=2$.
+Obviously, Theorem~\ref{thm:prime_with_unique_desired_distance} can be generalized (for $m=2$) into a similar conjecture.
 

### 2019-06-06 pps-linear-bound/pps-linear-lower-bound.tex
Language cosmetics

@@
 	\label{thm:prime_with_desired_distance}
-	For every $m \geq 3$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{prime}
+	For every $m \geq 3$, $n \geq m+1$, $d \geq 1$ there exists a \textbf{prime}
 	integral point set $M\in\mathfrak{M}(m,n)$ which

### 2019-06-09 pps-linear-bound/pps-linear-lower-bound.tex
different -> distinct
Fix typo: log -> \log
Avoid starting a proof with an equation

@@
 	\begin{equation}
-		\operatorname{diam} P \geq n^{\frac{\delta}{4 log 2(1+\varepsilon)}\log \log n}
+		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
 		.
@@
 	otherwise $\rho(n_i,\Delta_i)$-curve and $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
-	There can be only $2k-1$ different values for $n_i$,
+	There can be only $2k-1$ distinct values for $n_i$,
 	so there are at most $2k-1$ distinct segments $\Delta_i$.
@@
 \begin{proof}
-	$n^2+1$ points, including the ends of $\Delta$, split the segment $\Delta$ into $n^2$
+	Any $n^2+1$ points, including the ends of $\Delta$, split the segment $\Delta$ into $n^2$
 	sequential segments $\Delta_i$.
@@
 	\item
-		there are at most $2k-1$ distinct segments $\Delta_i$
-
-		different vs. distinct ?
-
-	\item
 		Summing up

### 2019-06-09 pps-linear-bound/pps-linear-lower-bound.tex
Fix language issue

@@
 
-We can improve the constant in Theorem~\ref{thm:main_estimate} if new exact values of $d(2,n)$ are found;
+We can improve the constant in Theorem~\ref{thm:main_estimate},
+if some new exact values of $d(2,n)$ are found;
 however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$.

### 2019-06-09 pps-linear-bound/pps-linear-lower-bound.tex
Fix a typo
Add new language issues
know out -> derive

@@
 Following the details of Solymosi's proof carefully,
-one can know out that the inequality holds at least for $c = 1/24$.
+one can derive that the inequality holds at least for $c = 1/24$.
 (See~\cite[Exercise 2.6]{garibaldi2005erdos} for some remarks.)
@@
 If the take $k=1$ in Lemma~\ref{lem:2k-1_segments},
-we obtan the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018}).
+we obtain the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018}).
 \begin{corollary}
@@
 	\item
+		contradict with or to ?
+
+	\item
 		Observation vs. Proposition
@@
 	\item
-		estimate vs. bound
+		estimate vs. bound vs. estimation
 

### 2019-06-10 pps-linear-bound/pps-linear-lower-bound.tex
contradict with -> contradict

@@
 	The first case is impossible because both $c_J$ and $c_K$ are positive;
-	the second case contradicts with~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4}.
+	the second case contradicts~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4}.
 
@@
 	\item
-		contradict with or to ?
-
-	\item
 		Observation vs. Proposition

### 2019-06-16 pps-linear-bound/pps-linear-lower-bound.tex
Advices to language issues list
Proper acknowledgements to Dr. Usachev

@@
 		is it correct to split the sentence in such a way?
-
-	\item
-		For the sake of completeness.
-
-		Is the phrase correct?
-		There should be an example in one of papers about Banach limits...
+		We should avoid it if possible.
 
@@
 	\item
-		Observation vs. Proposition
+		Observation -> Proposition
 
@@
 \section{Acknowledgements}
-Author thanks Prof. E.M. Semenov for the fruitful discussion and ideas
-and A.S. Chervinskaia for the idea of using the term ``facher'' and other linguistic advices.
+Author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
+A.S. Chervinskaia for the idea of using the term ``facher'' and other linguistic advices,
+and Dr. A.S. Usachev for proofreading.
 

### 2019-06-17 pps-linear-bound/pps-linear-lower-bound.tex
Fix a typo
Decimal values for some constants
Cosmetics
observation -> proposition

@@
 \newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{observation}[theorem]{Observation}
+\newtheorem{proposition}[theorem]{Proposition}
 
@@
 \cite{kurz2008bounds}, we derive
-\begin{observation}
+\begin{proposition}
 	\label{obs:4_leq_n_leq_21491}
@@
 	holds for $4 \leq n \leq 21491$.
-\end{observation}
+\end{proposition}
 
@@
 one can make the following
-\begin{observation}
+\begin{proposition}
 	\label{varphi_n_where_n_geq_21492}
@@
 	\end{equation}
-\end{observation}
+\end{proposition}
 
@@
 
-\begin{observation}
+\begin{proposition}
 	\label{obs:rho_curves}
@@
 	do not intersect.
-\end{observation}
+\end{proposition}
 
@@
 	Then for each $\Delta_i$ there is a $\rho(n_i,\Delta_i)$-curve containing $N$.
-	Due to Observation~\ref{obs:rho_curves}, all $n_i$ are distinct;
+	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
 	otherwise $\rho(n_i,\Delta_i)$-curve and $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
@@
 
-This lemma leads to the following observation.
+This lemma leads to the following proposition.
 
-\begin{observation}
+\begin{proposition}
 	\label{obs:estimate_points_on_straight_line}
@@
 	\begin{equation}
-		\gamma_2 = \frac{3846}{2593 \sqrt{647}-5823}
+		\gamma_2 = \frac{3846}{2593 \sqrt{647}-5823} = 0.063958...
 	\end{equation}
-\end{observation}
+\end{proposition}
 
@@
 	\begin{multline}
+		0.46530... =
 		\frac{3^{1/4}}{2^{3/2}} >
+		\\ >
 		\gamma = \frac{\sqrt{16 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}+\frac{14791716}{{{\left( 2593 \sqrt{647}-5823\right) }^{2}}}}-\frac{3846}{2593 \sqrt{647}-5823}}{8 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}}
@@
 		> \frac{5}{11}
+		.
 	\end{multline}
@@
 \begin{proof}
-	For $4 \leq n \leq 21491$, the claim of the theorem follows immediately from Observation~\ref{obs:4_leq_n_leq_21491}.
+	For $4 \leq n \leq 21491$, the claim of the theorem follows immediately from Proposition~\ref{obs:4_leq_n_leq_21491}.
 	Let us consider $M\in \mathfrak{M}(2,n)$, $n \geq 21492$, $\operatorname{diam} M = b$.
@@
 
-	a) $N$ belongs to $C$~--- totally at most $\gamma_2 b + 6$ points due to Lemma~\ref{lem:intervals_cross} and Observation~\ref{obs:estimate_points_on_straight_line};
+	a) $N$ belongs to $C$~--- totally at most $\gamma_2 b + 6$ points due to Lemma~\ref{lem:intervals_cross} and Proposition~\ref{obs:estimate_points_on_straight_line};
 
@@
 	\end{equation}
-	Observation~\ref{varphi_n_where_n_geq_21492} turns estimate~\eqref{eq:estimate_n_varphi_n_squared}
+	Proposition~\ref{varphi_n_where_n_geq_21492} turns estimate~\eqref{eq:estimate_n_varphi_n_squared}
 	into the following:
@@
 \begin{proof}
-	For $4 \leq n \leq 21491$, the claim follows from Observation~\ref{obs:4_leq_n_leq_21491} immediately.
+	For $4 \leq n \leq 21491$, the claim follows from Proposition~\ref{obs:4_leq_n_leq_21491} immediately.
 	For $n > 21491$, the inequality $0.45557(n-2) > \frac{5}{11}n$ holds.
@@
 
-Another possibility to increase the constant in out lower bound is introduced by the following problem.
+Another possibility to increase the constant in our lower bound is introduced by the following problem.
 
@@
 	\item
-		Observation -> Proposition
-
-	\item
 		estimate vs. bound vs. estimation

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
Add a ref about dense circular rational set

@@
 do not seem to show the way for improving the estimate
-due to the following well-known theorem (TODO: refs!).
+due to the following well-known theorem,
+which can be proved by an application of inversion~\cite{solymosi2010question}.
 \begin{theorem}

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
A small note
Replace the bold selection of 'open' with a remark on importance

@@
 	\label{lem:intervals_cross}
-	If \textbf{open} line segments $M_1 M_2$ and $M_3 M_4$ do not intersect,
+	If open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect,
 	%$M_1 \neq M_2$, $M_3 \neq M_4$,
@@
 \end{lemma}
+It is important that we consider the intersection of open line segments,
+so e.g. the cases $M_1 = M_3$ and $M_3 \in M_1 M_2$ satisfy the conditions of Lemma~\ref{lem:intervals_cross}.
 
@@
 \end{proposition}
+
+Therefore, we will focus on planar integral points sets of more than 21491 points.
 

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
ends (of line segment) -> endpoints
end the list -> conclude the list

@@
 \begin{proof}
-	Any $n^2+1$ points, including the ends of $\Delta$, split the segment $\Delta$ into $n^2$
+	Any $n^2+1$ points, including the endpoints of $\Delta$, split the segment $\Delta$ into $n^2$
 	sequential segments $\Delta_i$.
@@
 
-We want to end the list of conjectures with the one which looks the easiest to deal with.
+We want to conclude the list of conjectures with the one which looks the easiest to deal with.
 

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
Refs for circular sets

@@
 We conjecture that thorough investigation of existing examples of circular integral point sets
-(TODO: refs!)
+~\cite{bat2018number,harborth1993upper,piepmeyer1996maximum}
 will lead to the negative answers for the

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
Missing article 'a'

@@
 
-The solution of Problem~\ref{prb:minimal_shape} (if ever found) will lead to sophisticated packing problem,
+The solution of Problem~\ref{prb:minimal_shape} (if ever found) will lead to a sophisticated packing problem,
 something like ``How to put $n$ points in a convex hull of concentric circle and Reuleaux triangle?''

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
Fix the quantity of edges
A ref to the theorem about lattice

@@
 	Thus, the set $M$ is of characteristic 1,
-	and there is a Cartesian coordinate system such that $M_1=(-1/2,0)$, $M_2=(1/2,0)$, $M_3=(0, a/2)$.
-	(TODO: ref?)
+	and there is a Cartesian coordinate system such that $M_1=(-1/2,0)$, $M_2=(1/2,0)$, $M_3=(0, a/2)$
+	(see~\cite[Theorem 4]{our-vmmsh-2018}).
 	It is clear that $a\neq 0$.
@@
 	Consider $M\in\mathfrak{M}(m,n)$, $M_1,M_2\in M$, $|M_1 M_2| =1$.
-	Then either $n = m+1$ or $M$ is a ``blowup'' of a facher set and distance 1 occurs in $m$
-	not more than $1+\frac{m^2 - m}{2}$ times.
+	Then either $n = m+1$ or $M$ is a ``blowup'' of a facher set and distance 1 occurs in $M$
+	not more than $1+\frac{(m-1)^2 - (m-1)}{2}$ times.
 \end{conjecture}

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
Minor fix

@@
 	We are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
-	yields to $b > 10000$, because we know the maximum quantity of points for integral point sets of diameters at most $10000$.
+	yields to $b > 10000$, because we know the maximum quantity of points for planar integral point sets of diameters at most $10000$.
 	So, let us consider $t\geq 647$ and find a coefficient $\gamma_2$,

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
Switch to the new notation for all the integral point sets

@@
 Taking the fact into account, we will denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space with
-$\mathfrak{M}(m,n)$ (following the notation of~\cite{our-vmmsh-2018}).
+$\mathfrak{M}(m,n)$ (using the notation in~\cite{our-vmmsh-2018}),
+and let the symbol $\mathfrak{M}(m,\mathbb{N})$ stand for the set of all integral point sets in $m$-dimensional Euclidean space.
 The symbol $\# M$ will be used for cardinality of $M$, i.e. the quantity of points in $M$ in our case.
@@
 	\label{lem:2k-1_segments}
-	Let $M \in \mathfrak{M}(2,n)$ for some $n$ and let $m$ be a straight line.
+	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
 	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$,
@@
 	\label{lem:line_segment_with n_squared_plus_one_points}
-	Let $\Delta$ be a straight line segment, $|\Delta|=l$ and $M \in \mathfrak{M}(2,n)$ for some $n$.
+	Let $\Delta$ be a straight line segment, $|\Delta|=l$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
 	Let $\#(\Delta \cap M) = n^2 + 1$.
@@
 \begin{lemma}
-	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,n)$ for some $n$.
+	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
 	Let $\#(\Delta \cap M) = t$.
@@
 	\label{obs:estimate_points_on_straight_line}
-	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,n)$ for some $n$.
+	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
 	Let $\#(\Delta \cap M) = k$ and $b>10000$.
@@
 	\cite{kurz2005characteristic}
-	Let $M$ be a planar integral point set.
-	A squarefree number $q$ is called the \textit{characteristic} of $M$,
+	A squarefree number $q$ is called the \textit{characteristic} of $M\in\mathfrak{M}(2,\mathbb{N})$,
 	if for any points $M_1, M_2, M_3 \in M$ the area of triangle $M_1 M_2 M_3$
@@
 \end{definition}
-For given integral point set, the characteristic is determined uniquely.
+For given $M\in\mathfrak{M}(2,\mathbb{N})$, the characteristic is determined uniquely.
 
@@
 	\label{cor:only_one_distance_1_on_straight_line}
-	Let $M \in \mathfrak{M}(2,n)$ for some $n$ and let $m$ be a straight line.
+	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
 	Then there is at most one pair of points $M_1,M_2\in M \cap m$
@@
 	\label{prb:square_side_length_and_diameter_power}
-	Is there a number $\delta < 1$ such that any planar integral point set $M$ with $\operatorname{diam} M = d$
+	Is there a number $\delta < 1$ such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
 	belongs to a square of side length $d^\delta$?
@@
 	What are the maximal numbers $\delta \leq 1$ and $h \leq 0$
-	such that any planar integral point set $M$ with $\operatorname{diam} M = d$
+	such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
 	belongs to a square of side length $\delta d - h$?
@@
 	\label{prb:minimal_shape}
-	What is the minimal shape $\mathcal{S}$ such that any planar integral point set $M$ with $\operatorname{diam} M = d$
+	What is the minimal shape $\mathcal{S}$ such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
 	belongs to $\mathcal{S}$?

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
belongs -> is situated when needed

@@
 	Is there a number $\delta < 1$ such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	belongs to a square of side length $d^\delta$?
+	is situated in a square of side length $d^\delta$?
 \end{problem}
@@
 	such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	belongs to a square of side length $\delta d - h$?
+	is situated in a square of side length $\delta d - h$?
 \end{problem}
@@
 	What is the minimal shape $\mathcal{S}$ such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	belongs to $\mathcal{S}$?
+	is situated in  $\mathcal{S}$?
 \end{problem}
@@
 	\item
-		to belong vs. to be situated
-
-	\item
 		intersection of ... and ...

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
estimation -> estimate

@@
 We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$
-and estimations; in the present paper, the case of $m=2$ will mostly be in the focus.
+and estimates; in the present paper, the case of $m=2$ will mostly be in the focus.
 
@@
 	\item
-		estimate vs. bound vs. estimation
+		estimate vs. bound
 

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
intersection of ... and ... -> intersection of ... with ...

@@
 
-Now we need to estimate the quanity of points in an intersection of an integral point set and a line segment.
+Now we need to estimate the quanity of points in an intersection of an integral point set with a line segment.
 There is the following estimate for minimum diameter of planar integral point sets with many collinear points:
@@
 
-	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas and $cr(M_3 M_4)$~---
+	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas with $cr(M_3 M_4)$~---
 	totally at most $4 (|M_1 M_2| - 1)$ points;
 
-	d) $N$ belongs to the intersection of one of $|M_3 M_4| - 1$ hyperbolas and $cr(M_1 M_2)$~---
+	d) $N$ belongs to the intersection of one of $|M_3 M_4| - 1$ hyperbolas with $cr(M_1 M_2)$~---
 	totally at most $4 (|M_3 M_4| - 1)$ points;
@@
 	\item
-		intersection of ... and ...
-		OR
-		intersection of ... with ...
-		?
-
-	\item
 		estimate vs. bound

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
Let us treat 'summing up' to be a correct phrase
estimate -> bound where needed

@@
 We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$
-and estimates; in the present paper, the case of $m=2$ will mostly be in the focus.
+and bounds; in the present paper, the case of $m=2$ will mostly be in the focus.
 
@@
 \begin{remark}
-	The estimate of Theorem~\ref{thm:main_estimate} is not tight.
+	The bound of Theorem~\ref{thm:main_estimate} is not tight.
 	The derivation of tight bounds for the minimum diameter $d(2, n)$
@@
 If the answer for Problem~\ref{prb:square_side_length_and_diameter_power} is affirmative,
-then we can build an estimate better than linear.
+then we can build a bound better than linear.
 
@@
 	\item
-		Summing up
-
-		is it correct to say so ?
-
-	\item
 		... the following \\ Theorem X.
@@
 		We should avoid it if possible.
-
-	\item
-		estimate vs. bound
-
 \end{enumerate}

### 2019-06-18 pps-linear-bound/pps-linear-lower-bound.tex
The last language issue

@@
 Basing on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
-\cite{kurz2008bounds}, we derive
+\cite{kurz2008bounds}, we derive the following proposition.
 \begin{proposition}
@@
 Performing some simple manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
-one can make the following
+one can make the following proposition.
 \begin{proposition}
@@
 Now we need to estimate the quanity of points in an intersection of an integral point set with a line segment.
-There is the following estimate for minimum diameter of planar integral point sets with many collinear points:
+There is the following estimate for minimum diameter of planar integral point sets with many collinear points.
 \begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
@@
 however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$.
-To be precise, we have the following
+To be precise, we have the following theorem.
 
@@
 we will need~\cite[Proposition 6]{our-vmmsh-2018}.
-For the sake of completeness, we will give here a slightly rephrased version of it with the proof as
+For the sake of completeness, we will give here a slightly rephrased version of it with the proof as the following lemma.
 \begin{lemma}
@@
 together with results of~\cite[Section 6]{antonov2008maximal},
-we obtain the following
+we obtain the following theorem.
 
@@
 then we can easily dilate it to turn 1 into the desired distance.
-These facts, together with Theorem~\ref{thm:minimeter_1_planar}, give the following
+These facts, together with Theorem~\ref{thm:minimeter_1_planar}, give the following theorem.
 
@@
 The answer for $m=2$ is given by Theorem~\ref{thm:minimeter_1_planar}.
-For greater $m$, we will be brave enough to put forward
+For greater $m$, we will be brave enough to put forward the following conjecture.
 \begin{conjecture}
@@
 
-\section{Language issues to be resolved before publication}
-
-\begin{enumerate}
-	\item
-		... the following \\ Theorem X.
-
-		is it correct to split the sentence in such a way?
-		We should avoid it if possible.
-\end{enumerate}
-
 \section{Acknowledgements}

### 2019-06-19 pps-linear-bound/pps-linear-lower-bound.tex
Final fixes after self-proofreading

@@
 Taking the fact into account, we will denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space with
-$\mathfrak{M}(m,n)$ (using the notation in~\cite{our-vmmsh-2018}),
-and let the symbol $\mathfrak{M}(m,\mathbb{N})$ stand for the set of all integral point sets in $m$-dimensional Euclidean space.
+$\mathfrak{M}(m,n)$ (using the notation in~\cite{our-vmmsh-2018})
+and denote the set of all integral point sets in $m$-dimensional Euclidean space with $\mathfrak{M}(m,\mathbb{N})$.
 The symbol $\# M$ will be used for cardinality of $M$, i.e. the quantity of points in $M$ in our case.
@@
 In this section,
-we improve a better lower bound for minimum diameter of planar integral point sets by
+we improve the lower bound for minimum diameter of planar integral point sets by
 involving Point Packing in a Square problem.
@@
 \begin{definition}
-	\textit{Cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is union of two straight lines:
+	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
 	the line which the points $M_1$ and $M_2$ are situated on,
@@
 Performing some simple manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
-one can make the following proposition.
+one can derive the following proposition.
 \begin{proposition}
@@
 
-Now we need to estimate the quanity of points in an intersection of an integral point set with a line segment.
+Now we need to estimate the cardinality of an intersection of an integral point set with a line segment.
 There is the following estimate for minimum diameter of planar integral point sets with many collinear points.
@@
 	and equality $|M_1 M_2| = |M_3 M_4|$ holds but line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
-	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and $\rho(k,M_3 M_4)$-curve
+	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and the $\rho(k,M_3 M_4)$-curve
 	do not intersect.
@@
 	Let $f(k)$ denote the mininal length of a line segment
-	that shares $k$ points with and integral point set.
+	that shares $k$ points with an integral point set.
 	We observe that $f(k) > f (k-1)$.
@@
 		0.46530... =
-		\frac{3^{1/4}}{2^{3/2}} >
+		3^{1/4} \cdot 2^{-3/2} >
 		\\ >
@@
 It is shown that for $9\leq n \leq 122$ all the optimal sets are facher~\cite{kurz2008minimum}.
-For every cardinality $n$ and every characteristic $q$
+For every cardinality $n$ and every squarefree number $q$
 there exists a facher set $M\in\mathfrak{M}(2,n)$ with characteristic $q$~\cite[Theorem 5]{our-vmmsh-2018}.
@@
 	\end{equation}
-	Then, one of the distances is
+	Thus, one of the distances is
 	\begin{equation}
@@
 	Reminding that $k$ can be taken arbitrary large, we conclude that we can construct
-	a planar integral point set of arbitrary large count of points so that the distance 1 occurs in that set.
+	a planar integral point set of arbitrary large cardinality so that the distance 1 occurs in that set.
 \end{construction}
@@
 	The area of the triangle $M_1 M_3 M_4$ is rational because $|M_3 - M_4| \in \mathbb{Z}$.
-	Thus, the set $M$ is of characteristic 1,
+	Thus, the characteristic of $M$ is 1,
 	and there is a Cartesian coordinate system such that $M_1=(-1/2,0)$, $M_2=(1/2,0)$, $M_3=(0, a/2)$
@@
 If the answer for Problem~\ref{prb:square_side_length_and_diameter_linear} is not $\delta = 1$ and $h = 0$,
-then the linear estimate can be slightly improved.
+then the linear bound can be slightly improved.
 
 However, Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}
-do not seem to show the way for improving the estimate
+do not seem to show the way for improving the bound
 due to the following well-known theorem,
@@
 \end{problem}
-This problem in another generalization of Lemma~\ref{lem:square_container}.
+This problem is another generalization of Lemma~\ref{lem:square_container}.
 We know that $\mathcal{S}$ is not a circle of diameter $d$
@@
 
-When we talk about integral point sets containing distance 1, another problem arises.
+For planar integral point sets containing distance 1, we can put forward the following problem.
+\begin{problem}
+	For given $n\geq 3$ and squarefree number $q \neq 1$,
+	is there $M\in\mathfrak{M}(2,n)$, such that characteristic of $M$ is $q$
+	and distance 1 occurs in $M$?
+\end{problem}
+
+When we talk about integral point sets containing distance 1 of higher dimensions, another problem arises.
 
@@
 \end{conjecture}
-Such as estimate is based on the quantity of edges in $(m-2)$-dimensional simplex.
+Such as estimate is based on the quantity of edges in an $(m-2)$-dimensional simplex.
 

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Language cosmetics
Proofreading by Usachev - pp. 1,2

@@
 
-Avdeev N.N.
-
-On existence and diameter bounds of integral point sets.
+On existence of integral point sets and their diameter bounds.
 \footnote{
 	This work was carried out at Voronezh State University and supported by the Russian Science
-	Foundation under grant 19-11-00197.
+	Foundation grant 19-11-00197.
 }
+
+N.N. Avdeev
 
@@
 Erd\"os and Anning proved~\cite{anning1945integral,erdos1945integral} that every integral point set consists of a finite number of points.
-Taking the fact into account, we will denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space with
+Taking this into account, we denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space by
 $\mathfrak{M}(m,n)$ (using the notation in~\cite{our-vmmsh-2018})
-and denote the set of all integral point sets in $m$-dimensional Euclidean space with $\mathfrak{M}(m,\mathbb{N})$.
-The symbol $\# M$ will be used for cardinality of $M$, i.e. the quantity of points in $M$ in our case.
+and denote the set of all integral point sets in $m$-dimensional Euclidean space by $\mathfrak{M}(m,\mathbb{N})$.
+The symbol $\# M$ will be used for cardinality of $M$, that is the number of points in $M$ in our case.
 
@@
 \end{equation}
-The next emerging question is: how the diameter of an integral point set depends on its cardinality?
+Another emerging question is: how does the diameter of an integral point set depend on its cardinality?
 One can easily see that every $M\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = h$
@@
 for every $p\in\mathbb{N}$.
-So, the question above should be rephrased:
-how \textit{the least possible} diameter of an integral point set depends on its cardinality?
-To answer it, the following function was introduced~\cite{kurz2008bounds,kurz2008minimum}:
+So, the above question should be rephrased:
+how does \textit{the least possible} diameter of an integral point set depend on its cardinality?
+In order to answer this question, the following function was introduced~\cite{kurz2008bounds,kurz2008minimum}:
 \begin{equation}
@@
 We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$
-and bounds; in the present paper, the case of $m=2$ will mostly be in the focus.
+and its bounds; in the present paper, the case of $m=2$ will mostly be in the focus.
 
@@
 who proved that $cn \leq d(2,n)$ for a sufficiently small constant $c$.
-Following the details of Solymosi's proof carefully,
+Following Solymosi's proof carefully,
 one can derive that the inequality holds at least for $c = 1/24$.
@@
 The constant was improved in~\cite{our-mz-rus} to $1/8$ for all $n$ and in~\cite{our-vmmsh-2018}
-to $3/8$ for sufficiently big $n$.
+to $3/8$ for sufficiently large $n$.
 
-The paper~\cite{solymosi2003note} contained one more interesting result.
+The paper~\cite{solymosi2003note} contains one more interesting result.
 Let us define a function which is ``dual'' to $d(m,n)$ in some sense:
@@
 In the present paper we improve Solymosi's results:
-firstly, we obtain a greater constant $c = 5/11$ in Theorem~\ref{thm:main_estimate},
+first, we obtain a larger constant $c = 5/11$ in Theorem~\ref{thm:main_estimate},
 using the combined approach with the Point Packing in a Square problem
 (this approach is different from Solymosi's one);
-then, we prove that $l(m,n)=1$ for all possible $m$ and $n$.
+second, we prove that $l(m,n)=1$ for all possible $m$ and $n$.
 
@@
 In this section,
-we improve the lower bound for minimum diameter of planar integral point sets by
-involving Point Packing in a Square problem.
+we improve the lower bound for minimum diameter of planar integral point sets
+employing Point Packing in a Square problem.
 Below we introduce the problem, basic notions and results.
@@
 	We are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
-	yields to $b > 10000$, because we know the maximum quantity of points for planar integral point sets of diameters at most $10000$.
+	yields to $b > 10000$, because we know the maximum number of points for planar integral point sets of diameters at most $10000$.
 	So, let us consider $t\geq 647$ and find a coefficient $\gamma_2$,
@@
 
-When we talk about integral point sets containing distance 1 of higher dimensions, another problem arises.
+When we talk about integral point sets containing distance 1 in higher dimensions, another problem arises.
 
@@
 \end{conjecture}
-Such as estimate is based on the quantity of edges in an $(m-2)$-dimensional simplex.
+Such as estimate is based on the number of edges in an $(m-2)$-dimensional simplex.
 

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
bisector -> perpendicular bisector

@@
 	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
-	the line which the points $M_1$ and $M_2$ are situated on,
-	and the bisector of line segment $M_1 M_2$.
+	the line through $M_1$ and $M_2$,
+	and the perpendicular bisector of line segment $M_1 M_2$.
 \end{definition}
@@
 So, in the planar case a $\rho(k,M_1 M_2)$-curve
-is a branch of a hyperbola for $k\neq 0$ and the bisector of line segment $M_1 M_2$ for $k=0$.
+is a branch of a hyperbola for $k\neq 0$ and the perpendicular bisector of line segment $M_1 M_2$ for $k=0$.
 
@@
 \begin{proof}
-	Let $m$ denote the bisector of line segment $M_1 M_2$.
+	Let $m$ denote the perpendicular bisector of line segment $M_1 M_2$.
 	Due to the triangle inequality, there are two possibilities to place each of points $M_i$, $i=3,4$:
@@
 	then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
-	and one point out of the line, on the bisector of line segment $M_1 M_2$.
+	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
 	There is only one maximal integral point set $M' \supseteq M$,
-	and the bisector is the axis of symmetry for $M'$.
+	and the perpendicular bisector is the axis of symmetry for $M'$.
 	Moreover, if $n > 3$, then the distance 1 occurs in $M$ (and $M'$) only once.

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 3

@@
 
-For building the bound, we also require the following results and notions from~\cite{our-vmmsh-2018}.
+To prove the bound on $d(2,n)$, we also need the following results and notions from~\cite{our-vmmsh-2018}.
 
@@
 	Then $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
-	i.e. at least one diagonal is greater than at least one side.
+	that is at least one diagonal is greater than at least one side.
 \end{lemma}

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 4

@@
 Now we need to estimate the cardinality of an intersection of an integral point set with a line segment.
-There is the following estimate for minimum diameter of planar integral point sets with many collinear points.
+Assuming that the planar integral point sets contains many collinear points,
+the following result holds.
 \begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
@@
 as it does not provide the values of all the constants.
-To obtain the desired estimate, we will now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.
+To obtain the needed estimate, we now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.
 
 \begin{definition}
-	For a line segment $M_1 M_2$ and an integer $k$, $-|M_1 M_2| < k < |M_1 M_2|$,
+	For a line segment $M_1 M_2$ and an integer $k$, such that $-|M_1 M_2| < k < |M_1 M_2|$,
 	we define a $\rho(k,M_1 M_2)$-curve as the set of points $N$
@@
 	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
-	and equality $|M_1 M_2| = |M_3 M_4|$ holds but line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
+	the equality $|M_1 M_2| = |M_3 M_4|$ holds and line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
 	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and the $\rho(k,M_3 M_4)$-curve
@@
 	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
-	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$,
+	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$ with endpoints in $M$,
 	such that $|\Delta_i| = k$.

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 5

@@
 	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
-	otherwise $\rho(n_i,\Delta_i)$-curve and $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
+	otherwise the $\rho(n_i,\Delta_i)$-curve and the $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
 	There can be only $2k-1$ distinct values for $n_i$,
@@
 \begin{proof}
-	Any $n^2+1$ points, including the endpoints of $\Delta$, split the segment $\Delta$ into $n^2$
+	Any $n^2+1$ points, including the endpoints of $\Delta$, partition the segment $\Delta$ into $n^2$
 	sequential segments $\Delta_i$.
@@
 	at most three segments of length 2, etc.
-	The following two values of sums conclude the proof:
+	The following two expressions for sums conclude the proof:
 	\begin{equation}
@@
 
-Now we will estimate the length of a line segment that shares an arbitrary number of points
-with an integral point set.
+Now we will estimate the length of a line segment that intersects an integral point set
+by an arbitrary number of points.
 
@@
 	Let $f(k)$ denote the mininal length of a line segment
-	that shares $k$ points with an integral point set.
+	that intersects an integral point set by $k$ points.
 	We observe that $f(k) > f (k-1)$.

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 6

@@
 \begin{proof}
-	We are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
-	yields to $b > 10000$, because we know the maximum number of points for planar integral point sets of diameters at most $10000$.
+	We know the maximum number of points for planar integral point sets of diameters at most $10000$.
+	Thus, we are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
+	holds for $b > 10000$.
 	So, let us consider $t\geq 647$ and find a coefficient $\gamma_2$,
@@
 	holds for all $t\geq 647$.
-	For these $t$,
-	the left part of~\eqref{eq:inequality_for_linear_estimate_of_segment_length} obviously grows faster than the right one.
+	For such $t$,
+	the left-hand side of~\eqref{eq:inequality_for_linear_estimate_of_segment_length} obviously grows faster than the right-hand side.
 	Turning~\eqref{eq:inequality_for_linear_estimate_of_segment_length} into the same equality
-	and solving it for $t=647$, we obtain the claimed estimate.
+	and solving it for $t=647$, we obtain the required estimate.
 \end{proof}
@@
 \begin{proof}
-	For $4 \leq n \leq 21491$, the claim of the theorem follows immediately from Proposition~\ref{obs:4_leq_n_leq_21491}.
+	For $4 \leq n \leq 21491$, the assertion of the theorem follows immediately from Proposition~\ref{obs:4_leq_n_leq_21491}.
 	Let us consider $M\in \mathfrak{M}(2,n)$, $n \geq 21492$, $\operatorname{diam} M = b$.
@@
 	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1 M_2|$ and $|M_3 M_4|$
-	are minimal in $M$ ($M_2$ and $M_3$ may be the same point).
+	are minimal in $M$ ($M_2$ and $M_3$ may coincide).
 	Then $|M_1 M_2| \leq b\varphi_{n-1}$ and $|M_3 M_4| \leq b\varphi_{n-1}$.
@@
 	Let $C = cr(M_1 M_2) \cap cr(M_3 M_4)$.
-	Each point of $N\in M$ satisfies one of the following conditions:
+	Each point $N\in M$ satisfies one of the following conditions:
 

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 7

@@
 
-	a) $N$ belongs to $C$~--- totally at most $\gamma_2 b + 6$ points due to Lemma~\ref{lem:intervals_cross} and Proposition~\ref{obs:estimate_points_on_straight_line};
+	a) $N$ belongs to $C$~--- overall at most $\gamma_2 b + 6$ points by Lemma~\ref{lem:intervals_cross} and Proposition~\ref{obs:estimate_points_on_straight_line};
 
 	b) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas
-	with one of $|M_3 M_4| - 1$ hyperbolas~--- totally at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
+	with one of $|M_3 M_4| - 1$ hyperbolas~--- overall at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
 
 	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas with $cr(M_3 M_4)$~---
-	totally at most $4 (|M_1 M_2| - 1)$ points;
+	overall at most $4 (|M_1 M_2| - 1)$ points;
 
 	d) $N$ belongs to the intersection of one of $|M_3 M_4| - 1$ hyperbolas with $cr(M_1 M_2)$~---
-	totally at most $4 (|M_3 M_4| - 1)$ points;
+	overall at most $4 (|M_3 M_4| - 1)$ points;
 
 	(see~\cite{erdos1945integral} for details).
-	Summing up the cases above, we get the following estimate:
+	Summing up the above cases, we obtain the following estimate:
 	\begin{equation}
@@
 	\end{equation}
-	Let us denote $\lambda = b/(n-2)$ and solve the following quadratic inequality with respect to $\lambda$:
+	Let us denote $\lambda = b/(n-2)$ and solve the following quadratic inequality for $\lambda$:
 	\begin{equation}
@@
 if some new exact values of $d(2,n)$ are found;
-however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$.
-To be precise, we have the following theorem.
+however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$ above.
+More precisely, we have the following theorem.
 

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 8

@@
 	\begin{equation}
-		d(2,n) \geq n\cdot(3^{1/4}\cdot2^{-3/2} - \varepsilon)
+		d(2,n) \geq n\cdot(3^{1/4}\cdot 2^{-3/2} - \varepsilon)
 	\end{equation}
-	holds for every $n>n_0$.
+	holds for every $n>n_0$,
+	that is
+	\begin{equation}
+		\lim_{n\to\infty} \frac{d(2,n)}{n} \geq 3^{1/4}\cdot 2^{-3/2}
+		.
+	\end{equation}
 \end{theorem}
@@
 \end{definition}
-For given $M\in\mathfrak{M}(2,\mathbb{N})$, the characteristic is determined uniquely.
+For a given $M\in\mathfrak{M}(2,\mathbb{N})$, the characteristic is determined uniquely.
 
-Facher sets are the simplest ones of planar integral point sets.
-It is shown that for $9\leq n \leq 122$ all the optimal sets are facher~\cite{kurz2008minimum}.
+Facher sets are the simplest planar integral point sets.
+It is known that for $9\leq n \leq 122$ all the optimal sets are facher~\cite{kurz2008minimum}.
 For every cardinality $n$ and every squarefree number $q$
@@
 
-Now we will improve the Solymosi's result.
+Now we improve Solymosi's result.
 
@@
 	\label{con:planar_set_with_minimeter_1}
-	Let us choose a positive integer $k > 1$ and take
+	Let us choose a positive integer $k > 1$ and set
 	\begin{equation}

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 9

@@
 	\end{multline}
-	We take $d_j = 2^{2^j} + 1$ for $1 \leq j \leq k-1$.
+	We set $d_j = 2^{2^j} + 1$ for $1 \leq j \leq k-1$.
 	Then $d_j \equiv 1 \mod 4$.
 
-	Let $c_J = \prod_{j\in J} d_j$ for every subset of indexes $J\subset I = \{1,2,...,k-1\}$
+	Let $c_J = \prod_{j\in J} d_j$ for every subset of indices $J\subset I = \{1,2,...,k-1\}$
 	(and  $c_\varnothing = 1$).
-	We get
+	We obtain
 	\begin{equation}
@@
 
-	From~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4} we obtain $a/c_J \equiv 3 \mod 4$.
+	Statements~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4} yield $a/c_J \equiv 3 \mod 4$.
 	Let $b_J = (c_J - a/c_J)/2$, then $b_J \equiv 1 \mod 2$.
@@
 
-	We will now define the coordinates of the points as following:
+	Next we define the coordinates of the points as following:
 	\begin{equation}

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 10

@@
 	and distance 1 occurs in $M$.
-	Reminding that $k$ can be taken arbitrary large, we conclude that we can construct
+	Since $k$ can be taken arbitrary large, it follows that we can construct
 	a planar integral point set of arbitrary large cardinality so that the distance 1 occurs in that set.
@@
 	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then
-	obtain a right triangle of side length 1.
+	obtain an equilateral triangle of side length 1.
 	This triangle is obviously the optimal set in $\mathfrak{M}(2,3)$.
@@
 	we obtain the optimal set in $\mathfrak{M}(2,5)$ presented in~\cite[Fig. 1]{harborth1993upper}.
-	If we truncate one point from it,
+	If we remove one point from it,
 	then we get one of the two optimal sets in $\mathfrak{M}(2,4)$.
@@
 
-If the take $k=1$ in Lemma~\ref{lem:2k-1_segments},
+Setting $k=1$ in Lemma~\ref{lem:2k-1_segments},
 we obtain the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018}).
@@
 
-To describe all the sets $M \in \mathfrak{M}(2,n)$ with distance 1,
-we will need~\cite[Proposition 6]{our-vmmsh-2018}.
-For the sake of completeness, we will give here a slightly rephrased version of it with the proof as the following lemma.
+In order to describe all the sets $M \in \mathfrak{M}(2,n)$ with distance 1,
+we need~\cite[Proposition 6]{our-vmmsh-2018}.
+For the reader's convenience, we will state a slightly rephrased version of it and provide the proof.
 \begin{lemma}
 	\label{lem:no_4_points_in_semigeneral_position_with_distance_1}
-	Let $M=\{M_1,M_2,M_3,M_4\}\in \mathfrak{M}(2,4)$ such that $|M_1 M_2|=1$
-	and let $l$ be the straight line containing points $M_1$ and $M_2$.
-	Then one of the points $M_3$ or $M_4$ belongs to $l$.
+	Let $M=\{M_1,M_2,M_3,M_4\}\in \mathfrak{M}(2,4)$ such that $|M_1 M_2|=1$.
+	If $l$ is the straight line through points $M_1$ and $M_2$,
+	then one of the points $M_3$ or $M_4$ belongs to $l$.
 \end{lemma}
@@
 We know that $\mathcal{S}$ is not a circle of diameter $d$
-(the counterexample is a right triangle of side length $d$)
+(the counterexample is an equilateral triangle of side length $d$)
 and we conjecture that $\mathcal{S}$ is not a Reuleaux triangle of width $d$

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 11

@@
 	If (a) is true for one point and (b) is true for another, then the claim of the lemma follows.
-	So, suppose the contrary, i.e. both points $M_3$ and $M_4$ belong to $m$.
+	So, suppose to the contrary, that both points $M_3$ and $M_4$ belong to $m$.
 
@@
 	\end{equation}
-	or, which is the same,
+	or, equivalently,
 	\begin{equation}
 		1 + a^2 = (2b)^2
-		,
+		.
 	\end{equation}
-	that has no integral solutions.
+	This equation has no integral solutions.
 	This contradiction concludes the proof.
@@
 	\label{thm:minimeter_1_planar}
-	For every $n\geq 3$ there is a planar intergal set $M$ of $n$ points
+	For every $n\geq 3$ there is a planar integral set $M$ of $n$ points
 	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds.

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 12

@@
 
-A bit more strong result can also be claimed.
+A slightly stronger result can also be claimed.
 
@@
 \begin{proof}
-	Exploit Construction~\ref{con:planar_set_with_minimeter_1} to get $M'\in \mathfrak{M}(2,n-m+2)$
+	We apply Construction~\ref{con:planar_set_with_minimeter_1} to obtain $M'\in \mathfrak{M}(2,n-m+2)$
 	with $M'_1, M'_2 \in M'$ and $|M'_1 M'_2| = 1$.

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 13

@@
 	The derivation of tight bounds for the minimum diameter $d(2, n)$
-	is still a challenging task for the future~\cite[Section 7]{kurz2008minimum}.
+	is still a challenging task for the forthcoming investigation~\cite[Section 7]{kurz2008minimum}.
 \end{remark}
 The known upper bound is $d(2,n)\leq 2^{c \log n \log \log n}$~\cite{harborth1993upper}.
-However, we hope that the present article can give a framework for future better estimates.
+However, we hope that the present article can provide a framework for possible better estimates.
 
@@
 
-There are also the following ways to improve the current bound,
+There are also several ways to improve the current bound,
 based on Lemma~\ref{lem:square_container}.
@@
 If the answer for Problem~\ref{prb:square_side_length_and_diameter_power} is affirmative,
-then we can build a bound better than linear.
+then we can obtain a bound better than linear.
 

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 14

@@
 and we conjecture that $\mathcal{S}$ is not a Reuleaux triangle of width $d$
-(the probable counterexample belongs to a circle of diameter $d$, see Theorem~\ref{thm:rational_set_on_circle}).
+(the possible counterexample belongs to a circle of diameter $d$, see Theorem~\ref{thm:rational_set_on_circle}).
 
-The solution of Problem~\ref{prb:minimal_shape} (if ever found) will lead to a sophisticated packing problem,
+The solution to Problem~\ref{prb:minimal_shape} (if ever found) will lead to a sophisticated packing problem,
 something like ``How to put $n$ points in a convex hull of concentric circle and Reuleaux triangle?''
@@
 (see Remarks~\ref{rem:planar_set_with_minimeter_1_of_3_points} and~\ref{rem:planar_set_with_minimeter_1_of_4_and_5_points}).
-However, we can put forward the following conjecture.
+However, we can suggest the following conjecture.
 \begin{conjecture}
@@
 
-The motivation of Conjecture~\ref{con:no_optimal_with_edge_1} is based on the results of~\cite[Section 5]{kurz2008minimum},
+The motivation for Conjecture~\ref{con:no_optimal_with_edge_1} is based on the results of~\cite[Section 5]{kurz2008minimum},
 especially the following theorem.
@@
 
-For planar integral point sets containing distance 1, we can put forward the following problem.
+For planar integral point sets containing distance 1, we can emphasise on the following problem.
 \begin{problem}
@@
 
-When we talk about integral point sets containing distance 1 in higher dimensions, another problem arises.
+The consideration of integral point sets containing distance 1 in higher dimensions gives a rise to another problem.
 

### 2019-06-27 pps-linear-bound/pps-linear-lower-bound.tex
Proofreading by Usachev - p. 15

@@
 \begin{problem}
-	For $M\in\mathfrak{M}(m,n)$, how many times can the distance 1 occur in $M$?
+	How many times can the distance 1 occur in a set $M\in\mathfrak{M}(m,n)$?
 \end{problem}
 The answer for $m=2$ is given by Theorem~\ref{thm:minimeter_1_planar}.
-For greater $m$, we will be brave enough to put forward the following conjecture.
+For greater $m$, we will be bold enough to suggest the following conjecture.
 \begin{conjecture}
@@
 	Then either $n = m+1$ or $M$ is a ``blowup'' of a facher set and distance 1 occurs in $M$
-	not more than $1+\frac{(m-1)^2 - (m-1)}{2}$ times.
+	no more than $1+\frac{(m-1)^2 - (m-1)}{2}$ times.
 \end{conjecture}
-Such as estimate is based on the number of edges in an $(m-2)$-dimensional simplex.
+Such estimate is based on the number of edges in an $(m-2)$-dimensional simplex.
 
-We want to conclude the list of conjectures with the one which looks the easiest to deal with.
+We want to conclude the list of conjectures with the one which appears the easiest to deal with.
 
@@
 \end{conjecture}
-This conjecture gives the same claim with Theorem~\ref{thm:prime_with_desired_distance}, but for $m=2$.
+This conjecture gives the same claim as Theorem~\ref{thm:prime_with_desired_distance} does, but for $m=2$.
 Obviously, Theorem~\ref{thm:prime_with_unique_desired_distance} can be generalized (for $m=2$) into a similar conjecture.

### 2019-07-18 pps-linear-bound/pps-linear-lower-bound.tex
Title for PPS linear lower bound

@@
 \usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
-\usepackage[pdftex,unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
+\usepackage[unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
 %\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
@@
 
-On existence of integral point sets and their diameter bounds.
+\title{
+	On existence of integral point sets and their diameter bounds
 	\footnote{
@@
 	}
+}
 
+\author{
 	N.N. Avdeev
+	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
+}
+
+\maketitle
 

### 2019-07-18 mkmitu2016_and_arXiv/statya.tex
MKMITU - just for consistence

\documentclass{article}

\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage[tbtags]{amsmath}
\usepackage{amsfonts,amssymb}

\hoffset -10mm
\voffset -7mm

\usepackage{mathrsfs}
\usepackage{graphicx}

\usepackage{float} % Иначе ругается на таблицу

\usepackage{amssymb,amsmath,amsthm,amsfonts}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Theorem}

\begin{document}

N.N. Avdeev
\begin{center}
	\textbf{
		On the search of planar integral point sets
	}
\end{center}

\paragraph{Abstract.}
	The search of integral point sets usually employs some combinatorial techniques,
	i.e. combining simplices or known sets of smaller cardinality.
	In this note we (mostly due to methodical purposes) discuss another approach that is based on coordinate methods,
	graph theory and clique search.
	Some time estimations and given and multi-thread optimization is discussed.
	Finally, we check our software by employing it to find the minimum diameters
	of planar integral point sets and comparing the obtained values with known ones.

\vspace{3em}

УДК 511.95 + 514.112

Авдеев Н. Н.

\begin{center}
	\textbf{
	Об отыскании множеств точек на плоскости с целочисленными расстояниями}
\end{center}

\paragraph{Аннотация.}
	При отыскании множеств точек на плоскости с целочисленными расстояниями обычно используются
	методы комбинаторики, а именно сочетания симплексов или известных множеств меньшей мощности.
	В настоящей заметке (большей частью в методических целях)
	обсуждается другой подход, основанный на координатных методах,
	теории графов и поиске клик.
	Приводятся некоторые оценки по времени и обсуждается возможность применения параллельных вычислений.
	Разработанное ПО проверяется путём вычисления минимального возможного диаметра для множеств
	заданной мощности и сравнения полученных значений с уже известными.

\vspace{6em}

\footnotetext{Работа выполнена в Воронежском университете при поддержке РНФ, грант 16-11-101-25.}

Обозначим через $\mathbb{N}$ множество натуральных чисел, через $\mathbb{Z}$ --- множество целых чисел, через $\mathbb{R}$ --- множество действительных чисел, через $[x]$ --- наименьшее целое число, не превосходящее $x$, через $|A,B|$ --- расстояние между точками плоскости $A$ и $B$.
Для заданного $n\in \mathbb{N}, n\geq 3$ обозначим через $C_n$ множество таких последовательностей $M_1,M_2,...,M_n \in \mathbb{R}^2$, что $|M_i,M_j|\in\mathbb{N}$ для всех $1\leq i < j  \leq n$ и  $M_1,M_2,...,M_n$ не принадлежат никакой прямой.
Положим
$$
F(n)=\min\limits_{A\in C_n} d(A),
$$
где $d(A)$ --- диаметр $A$, т. е.
$$
d(A)=\max\limits_{x,y\in A}|x,y|
$$

В настоящей статье представлен алгоритм отыскания таких $A \in C_n$, что $d(A) = F(n)$.
Оценки для $F(n)$ были получены в \cite{nashaStatya}:
для $n \geq 17$
	$$
		\max\left( \frac{n-5}{8}, 1 \right) \leq F(n) \leq \left( \frac{b (1+\log_2 n )\ln (1 + \log_2 n)}{e}\right)^{1+\log_2 n},
	$$
где $b$ --- константа из неравенств Чебышева.

Рассмотрим сначала более общую задачу: для заданного $m \in \mathbb{N}$ отыщем (с точностью до движения) все такие множества $A \in C_n$, что $d(A) = m$.

Предположим, что такое множество найдено и его диаметр достигается на точках $P_1$ и $P_2$.
Введём на плоскости прямоугольную декартову систему координат так, что $P_1 = \left(-\frac{m}{2}; 0 \right)$, $P_2 = \left(\frac{m}{2}; 0\right)$.
Сведём задачу к конечному перебору, показав, что $A \subset P$, где $P$ --- конечное множество.
Действительно, пусть $P_{j} \in A$.
Тогда $|P_{j} P_1| \leq m$, $|P_{j} P_2| \leq m$, $|P_{j} P_1| \in \mathbb{Z}$ и $|P_{j} P_1| \in \mathbb{Z}$.
В качестве $P$ возьмём множество всех точек, удовлетворяющих вышеуказанному условию, т. е. множество точек пересечения пар всевозможных окружностей целого радиуса, не превосходящего $m$, с центрами в $P_1$ и $P_2$ соответственно.
Понятно, что $P$ содержит не более $2(m+1)^2$ точек; точное количество приведено ниже.

Координаты любой точки из $P$ могут быть выражены через радикалы:
$$
	P = \left\{
		\left(\left.  \frac{a^2-b^2}{2m} ; \pm\sqrt{a^2  - \left( \frac{a^2-b^2+m^2}{2m}\right)^2}  \right)	\right| 0 \leq a, b \leq m ; a+b \geq m
	\right\}
$$

Для оптимизации перебора сформируем три непересекающихся подмножества в множестве $P$:
$Q = \{P_j \in P | P_j = (x_j, 0)\}$;
$R = \{P_j \in P | P_j = (0, x_j), x_j \neq 0\}$;
$S = P \setminus \left( Q \cup R\right)$,
т. е. в множество $S$ попадают точки, у которых ни одна из координат не равна нулю.
Множества $Q$, $R$ и $S$ содержат $m+1$, $2\left\lbrack\frac{m+1}{2}\right\rbrack$ и $m^2$ точек соответственно.

Построим теперь граф $G$, вершины которого --- точки $P$;
вершины являются смежными, если расстояние между ними есть целое число, не превосходящее $m$.

На данном этапе задачу нельзя непосредственно свести к задаче о клике, решение которой приведено, например, в \cite{BronKerbosh}:
клика может оказаться подмножеством множества $Q$, а значит, и оси абсцисс, что не удовлетворяет условию задачи (такую клику в дальнейшем будем называть тривиальной);
однако решение задачи о клике, описанное в \cite{BronKerbosh} --- алгоритм Брона-Кербоша --- может быть модифицирован таким образом, что не найдёт тривиальные клики.
Для этого достаточно на начальном этапе каждой ветви перебора не использовать вершины, принадлежащие $Q$, как клики мощности 1.
При такой модификации алгоритма его сложность, очевидно, не увеличивается и, как доказано в работе \cite{Tomita}, не превосходит $O(3^{m^2/3})$.

Автором найден способ оптимизировать дальнейший перебор с учётом того, что рёбер в изучаемом графе достаточно мало.
Будем исключать вершины $P_j$ из графа по следующим правилам до тех пор, пока это возможно:

а) Если $P_j \in Q$ и $P_j$ не соединена ребром ни с какой точкой из $S \cup R$, то $P_j$ можно исключить, и вместе с ней можно исключить точку, симметричную $P_j$ относительно начала координат.

б) Если $P_j \in P$ и $P_j$ имеет менее $n-1$ смежных вершин, то $P_j$ можно исключить вместе с одной (в случае, когда $P_j \in Q \cup R$) или даже тремя (в случае, когда $P_j \in S$) точками, симметричными $P_j$ относительно осей координат.

Использование симметрии позволяет не проводить подсчёт смежных вершин для некоторых точек, что немаловажно для ускорения работы алгоритма на ЭВМ.
Заметим, что при реализации алгоритма симметричные точки целесообразно хранить парами или группами по 4.

Сложность каждой проверки правила (а) для множества $Q$ не превосходит $O(m^3)$, правила (б) для всего множества $P$ --- $O(m^4)$.

Для примера: однократное применение правила (б) к каждой точке из множества $Q$ для $m=413$ и $n=29$ заняло на используемой ЭВМ менее семи минут и привело к уменьшению количества вершин графа с 171396 до 2084.

После того, как дальнейшее сокращение графа $G_k$, полученного из графа $G$ применением этих правил $k$ раз, станет невозможным, начнём применять следующее рекуррентное правило, фактически обобщающее правила (а) и (б):

в)
Для каждой вершины $P_j$ через $\varphi(P_j,G_k)$ обозначим подграф графа $G_k$, состоящий из вершины $P_j$ и всех вершин, смежных с ней.
Если в графе $\varphi(P_j,G_k)$ отсутствует нетривиальная клика мощности $n$ (это определяется с помощью модифицированного алгоритма Брона-Кербоша), то вершину $P_j$ и вершины, симметричные ей относительно осей координат, можно исключить.

Если же нетривиальная клика найдена, то задача решена.

Целесообразно начать применение этого правила с вершины, имеющей минимальное число смежных вершин.

Очевидно, что рекурсия конечна.
Она заканчивается в одном из следующих случаев:

1) мощность графа $G_k$, полученного на $k$-й итерации, меньше $n$;

2) граф $G_k \subset Q$;

3) нетривиальная клика найдена.

С помощью приведённого алгоритма были найдены $F(n)$ для $n \leq 43 $.

Очевидно, что $F(3) = 1$.
Положим $m_1=1$, $n_1=3$.
На $k$-м шаге будем искать $A \in C_{n_k}$, такое, что $d(A) = m_k$.
Если такое множество найдено, это означает, что $F(n_k)=m_k$,
в таком случае увеличим количество точек, т. е. положим $n_{k+1} = n_k +1$, $m_{k+1} = m_k$;
в противном случае увеличим диаметр, т. е. $n_{k+1} = n_k$, $m_{k+1} = m_k +1$.

Автору удалось создать программу для ЭВМ и провести расчёты для множеств, диаметр которых не превышает 959.
Результаты вычислений представлены в таблице 1.

\begin{table}[H]
\caption{Значения $F(n)$}
\label{tabular:pc_counted}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
\textbf{n}    &  3 & 4 & 5 & 6 &  7 &  8 &  9 & 10 & 11 & 12 & 13 \\
\hline
\textbf{F(n)} &  1 & 4 & 7 & 8 & 17 & 21 & 29 & 40 & 51 & 63 & 74  \\
\hline
\hline
\textbf{n}    &  14 &  15 &  16 &  17 &  18 &  19 &  20 &  21 &  22 &  23 &  24 \\
\hline
\textbf{F(n)} &  91 & 104 & 121 & 134 & 153 & 164 & 196 & 212 & 228 & 244 & 272  \\
\hline
\hline
\textbf{n}    &  25 &  26 &  27 &  28 &  29 &  30 &  31 &  32 &  33 &  34 &  35 \\
\hline
\textbf{F(n)} & 288 & 319 & 332 & 364 & 396 & 437 & 464 & 494 & 524 & 553 & 578  \\
\hline
\end{tabular}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
\textbf{n}    &  36 &  37 &  38 &  39 &  40 &  41 &  42 &  43 \\
\hline
\textbf{F(n)} & 608 & 642 & 667 & 692 & 754 & 816 & 897 & 959 \\
\hline
\end{tabular}
\end{center}
\end{table}

Исходный код программы для ЭВМ, с помощью которой производились расчёты, размещён по адресам https://github.com/nickkolok/sct-algo и https://www.math.vsu.ru/source/sct-algo.tar.bz2 на условиях лицензии GNU GPLv3.

При проведении расчётов на ЭВМ, поддерживающих параллельные вычисления (например, лабораторных ЭВМ, имеющих многоядерный процессор) наиболее целесообразным представляется поиск систем с большим диаметром <<заранее>>.

Пусть имеется возможность исполнять $q$ параллельных процессов.
Всякий раз, когда количество выполняющихся процессов меньше максимально допустимого (т. е. после того, как предыдущий процесс завершил свою работу, либо перед началом вычислений), определим мощность и диаметр множества, которое будет искать запускаемый процесс, применяя следующие три правила до тех пор, пока это возможно:

(i)
Изначально положим $m_1=1$, $n_1=3$.

(ii)
Если множество с диаметром $m_k$ мощности $n_k$ было найдено ранее, положим $n_{k+1} = n_k +1$, $m_{k+1} = m_k$.

(iii)
Если ранее было установлено, что искомого множества с диаметром $m_k$ мощности, не превосходящей $n_k$, не существует,
или если в данный момент выполняется процесс, отыскивающий такое множество, положим $n_{k+1} = n_k$, $m_{k+1} = m_k+1$.

Как только достигнуто такое $k$, что данные правила применить нельзя, следует запустить процесс, отыскивающий множество с диаметром $m_k$ мощности $n_k$.

С ростом диаметра множества время поиска увеличивается очень значительно.
Например, отыскание оптимального мноества из 4 точек занимает менее 0.1 секунды на рабочей ЭВМ автора;
проверка наличия множества мощности 14 с диаметром 91 --- около 0.5 секунды;
вычисление же $F(43)$ при известной $F(42)$ потребовало нескольких суток работы ЭВМ.

Приведём наиболее простые, на взгляд автора, примеры оптимальных множеств.

Так, для $n=4$ оптимальных множеств два:
$$
\left\{\left( -2 ; 0\right),\left( 2 ; 0\right),\left( \frac{3}{2} ; \frac{\sqrt{15}}{2}\right),\left( 1 ; 0\right)\right\}
$$
и
$$
\left\{\left( -2 ; 0\right),\left( 2 ; 0\right),\left( \frac{3}{2} ; \frac{\sqrt{15}}{2}\right),\left( -\frac{3}{2} ; \frac{\sqrt{15}}{2}\right)\right\}
$$

\begin{figure}[h]
\center{\includegraphics[width=0.45\linewidth]{sct-4-4_1.png}}
\caption{$n=4$, первый случай}
\label{ris:image}
\end{figure}

\begin{figure}[h]
\center{\includegraphics[width=0.45\linewidth]{sct-4-4_2.png}}
\caption{$n=4$, второй случай}
\label{ris:image}
\end{figure}

Для $n=5$ оптимальное расположение единственно с точностью до движения:
$$
\left\{\left( -\frac{7}{2} ; 0\right),\left( \frac{7}{2} ; 0\right),\left( 0 ; \frac{\sqrt{15}}{2}\right),\left( \frac{1}{2} ; 0\right),\left( -\frac{1}{2} ; 0\right)\right\}
$$

\begin{figure}[h]
\center{\includegraphics[width=0.45\linewidth]{sct-5-7.png}}
\caption{$n=5$}
\label{ris:image}
\end{figure}

Автор выражает благодарность проф. Е. М. Семёнову за ценные советы и плодотворное обсуждение.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\small{
	First published as:
	Авдеев Н. Н.
	Об отыскании множеств точек на плоскости с целочисленными расстояниями
	// Математическое и компьютерное моделирование,
	информационные технологии управления: Сб. тр. Школы для студентов,
	аспирантов и молодых ученых <<МКМИТУ-2016>>. --- 2016. --- С. 15--19.
}

%\newpage

\addcontentsline{toc}{chapter}{\bibname}
\begin{thebibliography}{99}

\bibitem{nashaStatya} Авдеев Н. Н., Семёнов Е. М. О множествах точек на плоскости с целочисленными расстояниями. --- в печати.

\bibitem{BronKerbosh} Bron C., Kerbosh J. (1973), Algorithm 457 — Finding all cliques of an undirected graph, Comm. of ACM, 16, p. 575—577.

\bibitem{Tomita} Etsuji Tomita, Akira Tanaka, Haruhisa Takahashi (2006), The worst-case time complexity for generating all maximal cliques and computational experiments, Theoretical Computer Science, Vol 363, Issue 1, ISSN:0304-3975, p. 28-42.

\end{thebibliography}

%Замечание
%\begin{remark} Отметим, что доказательство теоремы \ref{theo1}
%может быть распространено на случай ...
%\end{remark}

%Определение
%\begin{definit} Будем называть ...
%\end{definit}

%Предложение
%\begin{propos} Пусть выполняются следующие неравенства...
%\end{propos}

%\begin{figure}[ht]
%\centering
%\includegraphics{1kriv80}
%\caption{Многообразие точек смены кратности.}
%\end{figure}

%===============Список литературы==================

\end{document}

### 2019-07-19 semi-general-suplinear-bound/semi-general-suplinear-bound.tex
First drafts of suplinear bound for sets in semi-general position

\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel} %используем русский и английский языки с переносами
\usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
\usepackage[unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{indentfirst} % включить отступ у первого абзаца
\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
\graphicspath{{illustr/}}%путь к рисункам

\makeatletter
\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"

\usepackage{geometry} % Меняем поля страницы.
\geometry{left=2cm}% левое поле
\geometry{right=1cm}% правое поле
\geometry{top=2cm}% верхнее поле
\geometry{bottom=2cm}% нижнее поле

\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра

\sloppy

\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}

\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
\addbibresource{../common/notmy.bib}
\addbibresource{../common/my.bib}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{construction}[theorem]{Construction}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{problem}[theorem]{Problem}

\begin{document}

%\renewcommand{\bibname}{Список цитированной литературы}
%\renewcommand\refname{\bibname}
% !!!
% The text starts here

\title{
	On diameter bounds for planar integral point sets in semi-general position
	\footnote{
		This work was carried out at Voronezh State University and supported by the Russian Science
		Foundation grant 19-11-00197.
	}
}

\author{
	N.N. Avdeev
	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
}

\maketitle

\paragraph{Abstract.}
A point set $M$ in the Euclidean plane is called a planar integral point set if all the distances between the
elements of $M$ are integers, and $M$ is not situated on a straight line.
A planar integral point set is called to be in semi-general position, if it does not contain collinear triples.
The existing bound for mininum diameter of planar integral point sets is linear.
We prove a new bound for mininum diameter of planar integral point sets in semi-general position
which is greater than linear.

\section{Introduction}

Kurz - facher
Bat-Ochir - circular
Kurz - precise values
Kurz - about many collinear points

Kurz - semi-general and general position, precise values

Piepmeyer(?) - upper bound \cite{harborth1993upper}

Particular cases - Huff, Monographies (2?), my-amm, my-ped

\section{Preliminary results}

In this section, we give some lemmas which will be used for the proof.

\begin{lemma}
	\cite[Observation 1]{solymosi2003note}
	If a triangle $T$ has integer side-lengths $a \leq b \leq c$,
	then the minimal height $m$ of it is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
\end{lemma}

\begin{definition}
	The part of a plane between two parallel straight lines with distance $\rho$ between the lines
	is called a strip of width $\rho$.
\end{definition}

\begin{lemma}
	TODO: cite Kvant

	If a triange $T$ with minimal height $\rho$ is situated in a strip,
	then the width of a strip is at least $\rho$.
\end{lemma}

\begin{lemma}
	\cite[Lemma 4]{our-vmmsh-2018};
	\cite[TODO]{my-pps-linear-bound-2019}
	\label{lem:square_container}
	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
	Then $M$ is situated in a square of side length $d$.
\end{lemma}

\begin{definition}
	\cite[TODO]{my-pps-linear-bound-2019}
	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
	the line through $M_1$ and $M_2$,
	and the perpendicular bisector of line segment $M_1 M_2$.
\end{definition}

\begin{lemma}
	\cite[Theorem TODO]{my-pps-linear-bound-2019}
	\label{lem:no_distance_one}
	Each set $M\in\mathfrak{M}(2,n)$
	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
	consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
\end{lemma}

\begin{lemma}
	\label{lem:count_of_points_on_hyperbolas}
	Let $\{M_1, M_2, M_3, M_4\} \subset M\in\overline\mathfrak{M}(2,n)$
	(points $M_2$ and $M_3$ may coincide, other points may not), $n\geq 4$.
	Then $\# M \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
\end{lemma}

\begin{remark}
	Lemma~\ref{lem:count_of_points_on_hyperbolas} is one of the variations of~\cite{erdos1945integral},
	\cite[TODO]{my-pps-linear-bound-2019}, etc.
\end{remark}

\begin{proof}
	Each point $N\in M$ satisfies one of the following conditions:

	a) $N$ belongs to $cr(M_1,M_2)$~--- overall at most 4 points;

	b) $N$ belongs to $cr(M_3,M_4)$~--- overall at most 4 points;

	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas
	with one of $|M_3 M_4| - 1$ hyperbolas~--- overall at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;

	Due to Lemma~\ref{lem:no_distance_one} we have $|M_1 M_2| \geq 2$ and $|M_3 M_4| \geq 2$.
	Since
	\begin{multline}
		4 (|M_1 M_2| - 1)(|M_3 M_4| - 1) + 4 + 4
		=
		4 ( (|M_1 M_2| - 1)(|M_3 M_4| - 1) + 2)
		=
		4 ( |M_1 M_2| \cdot |M_3 M_4| + 1 - |M_1 M_2| - |M_3 M_4| + 2)
		=
		4 ( |M_1 M_2| \cdot |M_3 M_4| + 1 - |M_1 M_2| - |M_3 M_4| + 2)
		=
		4 ( |M_1 M_2| \cdot |M_3 M_4| + (1 - |M_1 M_2|) + (2 - |M_3 M_4|))
		\leq
		4 |M_1 M_2| \cdot |M_3 M_4|
		,
	\end{multline}
	we are done.
\end{proof}

\section{The main result}

\begin{theorem}
	For every $n\in\mathbb{N}$ we have
	\begin{equation}
		\overline{d}(2,n) \geq c n^{5/4}
		,
	\end{equation}
	where $c = $ TODO!
\end{theorem}

\begin{proof}
	For $n = 3$ we have $\overline{d}(2,n) = 1$.
	Consider $M\in\overline\mathfrak{M}(2,n)$, $n \geq 4$, $\operatorname{diam} M = p$.

	Let us choose points $M_1, M_2, M_3, M_4 \in M$
	(points $M_2$ and $M_3$ may coincide, other points may not), such that
	\begin{equation}
		\min_{A, B \in M} |AB| = |M_1 M_2|
		,
	\end{equation}
	\begin{equation}
		\min_{A, B \in M \setminus \{M_1\}} |AB| = |M_3 M_4| = m
		.
	\end{equation}

	For $m \leq p^{2/5}$, Lemma~\ref{lem:count_of_points_on_hyperbolas} yields that
	\begin{equation}
		n \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| \leq  4 p^{4/5}
		,
	\end{equation}
	or, that is the same,
	\begin{equation}
		p \geq (n/4) ^ {5/4}
		.
	\end{equation}

	So, let us consider $m > p^{2/5}$.
	Due to Lemma~\ref{}

\end{proof}

\section{Conclusion}
The presented bound is the first special lower bound for sets in semi-general position.
Thus, we did not accepted the challenge to make the constant in Theorem~\ref{thm:main_result} as large as possible,
in order to keep the ideas of the proof clear and understandable.
A more thorough research can be done in the future to enlarge the constant.
However, the upper and lower bounds are still not tight.

\section{Acknowledgements}
Author thanks Dr. Prof. E.M. Semenov for proofreading and valuable advice.

\printbibliography

\end{document}

### 2019-07-20 semi-general-suplinear-bound/semi-general-suplinear-bound.tex
The paper is almost ready - we need only intro and references

@@
 We prove a new bound for mininum diameter of planar integral point sets in semi-general position
-which is greater than linear.
+which is better than linear.
 
@@
 
+The upper bound
 Piepmeyer(?) - upper bound \cite{harborth1993upper}
 
-Particular cases - Huff, Monographies (2?), my-amm, my-ped
+Particular cases - Huff, Monographies (2?), \cite{our-pmm-2018}, \cite{our-ped-2018}.
 
@@
 \begin{lemma}
-	TODO: cite Kvant
-
+	\cite{smurov1998stripcoverings}
 	If a triange $T$ with minimal height $\rho$ is situated in a strip,
@@
 \end{lemma}
+
+\begin{corollary}
+	\label{cor:solymosi_strip}
+	If a triangle $T$ with integer side-lengths $a \leq b \leq c$ is situated in a strip,
+	then the width of a strip is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
+\end{corollary}
 
@@
 	\label{lem:count_of_points_on_hyperbolas}
-	Let $\{M_1, M_2, M_3, M_4\} \subset M\in\overline\mathfrak{M}(2,n)$
+	Let $\{M_1, M_2, M_3, M_4\} \subset M\in\overline{\mathfrak{M}}(2,n)$
 	(points $M_2$ and $M_3$ may coincide, other points may not), $n\geq 4$.
@@
 		=
+		\\=
 		4 ( |M_1 M_2| \cdot |M_3 M_4| + 1 - |M_1 M_2| - |M_3 M_4| + 2)
 		=
-		4 ( |M_1 M_2| \cdot |M_3 M_4| + 1 - |M_1 M_2| - |M_3 M_4| + 2)
-		=
+		\\=
 		4 ( |M_1 M_2| \cdot |M_3 M_4| + (1 - |M_1 M_2|) + (2 - |M_3 M_4|))
@@
 \begin{theorem}
+	\label{thm:main_result}
 	For every $n\in\mathbb{N}$ we have
@@
 	\end{equation}
-	where $c = $ TODO!
+	where $c = 5^{-5/4}$.
 \end{theorem}
@@
 \begin{proof}
-	For $n = 3$ we have $\overline{d}(2,n) = 1$.
-	Consider $M\in\overline\mathfrak{M}(2,n)$, $n \geq 4$, $\operatorname{diam} M = p$.
+	For $n = 3$ we have $\overline{d}(2,3) = 1$.
+	Consider $M\in\overline{\mathfrak{M}}(2,n)$, $n \geq 4$, $\operatorname{diam} M = p$.
 
@@
 	\begin{equation}
-		p \geq (n/4) ^ {5/4}
+		\label{eq:hyperbolas_5_4}
+		p \geq (n/4) ^ {5/4} > (n/5) ^ {5/4}
 		.
@@
 	So, let us consider $m > p^{2/5}$.
-	Due to Lemma~\ref{}
+	Then for any $A,B \in M\setminus\{M_1\}$ the inequality $|AB| > p^{2/5}$ holds.
+	Due to Corollary~\ref{cor:solymosi_strip}, no three points of $M\setminus\{M_1\}$
+	are located in a strip of width $p^{1/5} / 2$.
 
+	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square with side length $p$.
+	Let us partition this square into $q$ strips, $2p^{4/5} \leq q < 2p^{4/5} + 1$, each of width at most $p^{1/5} / 2$.
+	Every strip contains at most two points of  $M\setminus\{M_1\}$,
+	thus
+	\begin{equation}
+		\label{eq:strips_4_5}
+		n \leq 2(2p^{4/5} + 1) + 1
+		= 4p^{4/5}+3
+		\leq 5 p^{4/5}
+		.
+	\end{equation}
+	The latter inequality holds because $\overline{d}(2,n) \geq 4$ for $n\geq 4$ [TODO: Kurz!]
+	and $4^{4/5}>3$.
+	From the inequality~\eqref{eq:strips_4_5} one can easily derive that
+	\begin{equation}
+		\label{eq:strips_5_4}
+		p \geq (n/5) ^ {5/4}
+		.
+	\end{equation}
 \end{proof}
+
+%TODO: a remark on how to easily derive 6/5 from Solymosi's corollary.
 

### 2019-09-07 AvdMomZvol_rails/Article.tex
Drafts of the paper about particular bounds

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian,english]{babel}
\usepackage{amsmath,amsfonts,amssymb,euscript,graphicx,wrapfig,multirow}
\usepackage{dsfont}
\usepackage{amsthm}
\inputencoding{utf8}
\bibliographystyle{unsrt}
\textheight=240mm \textwidth=170mm
\hoffset=-17mm
\voffset=-17mm

\usepackage{hyperref}

\makeatletter
%\renewcommand{\fnum@figure}{Figure \thefigure}
\renewcommand{\@biblabel}[1]{#1.}
\makeatother

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}
\theoremstyle{dfn}
\newtheorem{dfn}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}
%\renewcommand\refname{\centering References}

% Переключаем язык на английский.
% Очень полезно как в плане типографики (в том числе расстановки переносов),
% так и в плане того, что не надо переименовывать "Рисунок" в "Figure"
\selectlanguage{english}

\begin{center}
\emph{Voronezh State University}
\end{center}

\section{Introduction}

\begin{dfn}\label{dfn1}
	A planar integral point set (PIPS) is a set $\mathcal{P}$
	of non-collinear points in plane $\mathbb{R}^{2}$ such that
	for any pair of points $P_{1}, P_{2} \in \mathcal{P}$
	the Euclidean distance $|P_{1}P_{2}|$
	between points $P_{1}$ and $P_{2}$ is integral.
\end{dfn}

Definition \ref{dfn1} can be generalized as follows:

\begin{dfn}
	An integral point set $\mathcal{P}$ is a set of $n$ points in
	the $m$-dimensional Euclidean space $\mathbb{R}^{m}$ with pairwise
	integral distances.
\end{dfn}

How can we characterize an integral point set?
First of all, we can look at its dimension;
then, we can count the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
and is further called the cardinality;
finally, we can naturally% TODO: naturally/essentially? cf. arxiv/1
define the diameter of a finite point set.

\begin{dfn}
	A diameter of the integral point set $\mathcal{P}$ is defined as
	\begin{equation}
		\operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
		\mathcal{P}}{\max} |P_{1}P_{2}|
	\end{equation}
\end{dfn}

Every integral point set also has a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.

\begin{dfn}
	The function $d(m, n)$ is the minimum possible diameter of
	the integral point set $\mathcal{P}$ of $n$ points in
	$m$-dimensional Euclidean space $\mathbb{R}^{m}$.
\end{dfn}

%For a bit more sophisticated configurations, another functions are presented~\cite{kurz2008minimum,kreisel2008heptagon}.
%
%\begin{dfn}
%	The function $\overline{d}(m, n)$ is the minimum possible diameter of
%	the integral point set $\mathcal{P}$ of $n$ points in
%	$m$-dimensional Euclidean space $\mathbb{R}^{m}$
%	such that no $m+1$ points are located on an $(m-1)$-dimensional hyperplane.
%\end{dfn}

%\begin{dfn}
%	The function $\dot{d}(m, n)$ is the minimum possible diameter of
%	the integral point set $\mathcal{P}$ of $n$ points in
%	$m$-dimensional Euclidean space $\mathbb{R}^{m}$
%	such that no three points are located on a straight line
%	and no four points are located on a circle.
%\end{dfn}

For the list of known bounds for $d(m, n)$,
we refer the reader to~\cite[Theorem 1]{kurz2008bounds} or to~\cite{our-vmmsh-2018}.
We will discuss the following estimations presented at~\cite{kurz2008bounds}:
\begin{align}
	d(m, 2m + 1) \leq 8\\
	d(m, 2m + 2) \leq 13 \hypertarget{d2}\\
	d(m, 3m) \leq 109
\end{align}
and the following theorem \cite[Theorem 2.1]{kurz2008bounds}.

\begin{theorem}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - 2$ points on line $l$ and two points $P_{1}$ and $P_{2}$ on a
	parallel line $m$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
	exist positive integers $v$, $w$ with $f^{2} + v^{2}
	= w^{2}$ and $v < 2r$, where $|P_{1}P_{2}| = f$,
	then

	\begin{equation}\label{formula1}
		d(m, n - 2 + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
	\end{equation}

\end{theorem}

Firstly, we discuss the classification of planar integral points sets;
then, we present some bounds for $d(m,n)$ based on planar integral point sets of particular types
and provide some general constructions for such bounds.

\section{Classification of planar integral point sets}

\subsection{Integral point sets situated in two straight lines}

\begin{dfn}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.
\end{dfn}
Facher sets are very dominating examples of planar integral pont sets.
In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.

For non-facher integral point sets situated in two straight lines,
we can easily distinct the following three cases:

\begin{dfn}
	A planar integral point sets situated in two parallel straight lines
	is called a \textit{rails} set.
\end{dfn}

Among the rails sets, sets with 2 points on one line and all the other on another line dominate.

\begin{dfn}
	A planar integral point sets situated in two perpendicular straight lines
	is called a \textit{cross} set.
\end{dfn}
Every cross set has characteristic 1;
in~\cite{antonov2008maximal}, cross sets with only 2 points out of one of the lines are called \textit{crabs}.

\begin{dfn}
	A planar integral point sets situated in two straight lines
	that are not parallel nor perpendicular,
	is called a \textit{sciccors} set.
\end{dfn}

There is an important subclassof scissors sets.

\begin{dfn}
	A scissors set with an axis of symmetry,
	which is the angle bisector for the straight lines,
	is called a \textit{pyramid} set.
\end{dfn}

TODO: rephrase?

TODO: can a scissors set have another axis of symmetry?

\subsection{Other integral point sets}

\begin{dfn}
	A planar integral point sets that is situated on a circle is called a \textit{circular}
	point set.
\end{dfn}

Circular sets are very important examples
of integral point sets~\cite{harborth1993upper,piepmeyer1996maximum,bat2018number}.

\begin{dfn}
	A planar integral point sets that is situated on the conjunction of a circle with its center,
	is called a \textit{centered-circular} point set.
\end{dfn}

These six classes dominate among all the known planar integral point sets;
however, some sophisticated constructions are also known.

TODO: pictures!

\section{Bounds based on rails sets}

Let $i = \overline{1, k}$ denote the enumeration of all $i$
from $1$ to $k$.
Theorem \cite[Theorem 2.1]{kurz2008bounds} can be generalized as follows:

\begin{theorem}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - k$ points on line $l_1$ and $k$ points $P_1$, $P_2$, ..., $P_k$ on a
	parallel line $l_2$ with distance $r$ between $l_1$ and $l_2$. If there
	exist positive integers $v$, $w_{ij}$, with $f_{ij}^{2} + v^{2}
	= w_{ij}^{2}$ and $v < 2r$, where $i = \overline{1, k - 1}$, $j =
	\overline{i + 1, k}$, $|P_{i}P_{j}| = f_{ij}$,
	then

	\begin{equation}
		d(m, n - k + k(m - k)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
	\end{equation}

\end{theorem}

Below we give the planar integral point sets and the corresponding
estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 24$.
For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
which means that each abscissa is multiplied by $1/q$
and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
$$
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
$$

\begin{itemize}
\setlength{\itemsep}{-1mm}

\begin{figure}[htbp]
	\includegraphics[width=.48\linewidth]{picture_2.png}
	\hfill
	\includegraphics[width=.48\linewidth]{picture_3.png}
	\\
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 9 and diameter 158}
	\label{picture_2.png}}
	\hfill
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 10 and diameter 158}
	\label{picture_3.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
(-78 , 0),
(\pm 62, 0),
(\pm 55 , 0),
(\pm 10 , 0)\}
$
(Figure~\ref{picture_2.png})

$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,

which gives $d(m, 2m + 3) \leq 158$.

\item
$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
(\pm 78 , 0),
(\pm 62, 0),
(\pm 55 , 0),
(\pm 10 , 0)\}
$
(Figure~\ref{picture_3.png}).

$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,

which gives $d(m, 2m + 4) \leq 158$.

\begin{figure}[htbp]
	\includegraphics[width=.48\linewidth]{picture_4.png}
	\hfill
	\includegraphics[width=.48\linewidth]{picture_5.png}
	\\
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 11 and diameter 252}
	\label{picture_4.png}}
	\hfill
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 12 and diameter 297}
	\label{picture_5.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
(\pm 207 , 0),
(\pm 145 , 0),
(\pm 63 , 0),
(\pm 25 , 0),
(-297 , 0)\}
$
(Figure~\ref{picture_4.png}).

$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P})} = 252$,

which gives $d(m, 2m + 5) \leq 252$.

\begin{figure}[h!]
\center{\includegraphics[width=0.7\linewidth]{picture_6.png}}
\parbox{0.7\linewidth}{\caption{IPS of cardinality 13 and diameter 1073}
\label{picture_6.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
(\pm 297 , 0),
(\pm 207 , 0),
(\pm 145 , 0),
(\pm 63 , 0),
(\pm 25 , 0)\}
$
(Figure~\ref{picture_5.png}).

$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P})} = 297$,

which gives $d(m, 2m + 6) \leq 297$.

\item
$\mathcal{P}=\sqrt{19019}/{2} * \{ (\pm 200, 3),
(\pm 873 , 0),
(\pm 615 , 0),
(\pm 377 , 0),
(\pm 215 , 0),
(\pm 23 , 0),
$

$
(-1273 , 0)\}
$
(Figure~\ref{picture_6.png}).

$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P})} = 1073$,

which gives $d(m, 2m + 7) \leq 1073$.

\begin{figure}[h!]
\center{\includegraphics[width=0.7\linewidth]{picture_7.png}}
\parbox{0.7\linewidth}{\caption{IPS of cardinality 14 and diameter 1273}
\label{picture_7.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{19019}/{2} * \{ (\pm 200, 3),
(\pm 1273 , 0),
(\pm 873 , 0),
(\pm 615 , 0),
(\pm 377 , 0),
(\pm 215 , 0),
$

$
(\pm 23 , 0)\}
$
(Figure~\ref{picture_7.png}).

$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P})} = 1273$,

which gives $d(m, 2m + 8) \leq 1273$.

\begin{figure}[h!]
\center{\includegraphics[width=0.72\linewidth]{picture_8.png}}
\parbox{0.73\linewidth}{\caption{IPS of cardinality 15 and diameter 1888}
\label{picture_8.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{385}/{2} * \{ (\pm 1105, 48),
(\pm 1587 , 0),
(\pm 1269 , 0),
(\pm 763 , 0),
(\pm 623 , 0),
(\pm 529 , 0),
$

$
(\pm 339 , 0),
(-2189 , 0)\}
$
(Figure~\ref{picture_8.png}).

$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P})} = 1888$,

which gives $d(m, 2m + 9) \leq 1888$.

\begin{figure}[h!]
\center{\includegraphics[width=0.72\linewidth]{picture_9.png}}
\parbox{0.73\linewidth}{\caption{IPS of cardinality 16 and diameter 2189}
\label{picture_9.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{385}/{2} * \{ (\pm 1105, 48),
(\pm 2189 , 0),
(\pm 1587 , 0),
(\pm 1269 , 0),
(\pm 763 , 0),
(\pm 623 , 0),
$

$
(\pm 529 , 0),
(\pm 339 , 0)\}
$
(Figure~\ref{picture_9.png}).

$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P})} = 2189$,

which gives $d(m, 2m + 10) \leq 2189$.

\begin{figure}[h!]
\center{\includegraphics[width=0.9\linewidth]{picture_13.png}}
\parbox{0.9\linewidth}{\caption{IPS of cardinality 17 and diameter 3224}
\label{picture_13.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
(\pm 1376 , 0),
(\pm 1036 , 0),
(\pm 899 , 0),
(\pm 849 , 0),
(\pm 613 , 0),
$

$
(\pm 576 , 0),
(\pm 100 , 0),
(-1848 , 0)\}
$
(Figure~\ref{picture_13.png}).

$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 3224$,

which gives $d(m, 2m + 11) \leq 3224$.

\begin{figure}[h!]
\center{\includegraphics[width=0.9\linewidth]{picture_14.png}}
\parbox{0.9\linewidth}{\caption{IPS of cardinality 18 and diameter 3696}
\label{picture_14.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
(\pm 1848 , 0),
(\pm 1376 , 0),
(\pm 1036 , 0),
(\pm 899 , 0),
(\pm 849 , 0),
$

$
(\pm 613 , 0),
(\pm 576 , 0),
(\pm 100 , 0)\}
$
(Figure~\ref{picture_14.png}).

$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 3696$,

which gives $d(m, 2m + 12) \leq 3696$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_15.png}}
\parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 5141}
\label{picture_15.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
(\pm 1848 , 0),
(\pm 1376 , 0),
(\pm 1036 , 0),
(\pm 899 , 0),
(\pm 849 , 0),
$

$
(\pm 613 , 0),
(\pm 576 , 0),
(\pm 100 , 0),
(-3293 , 0)\}
$
(Figure~\ref{picture_15.png}).

$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 5141$,

which gives $d(m, 2m + 13) \leq 5141$.

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
(\pm 3293 , 0),
(\pm 1848 , 0),
(\pm 1376 , 0),
(\pm 1036 , 0),
(\pm 899 , 0),
$

$
(\pm 849 , 0),
(\pm 613 , 0),
(\pm 576 , 0),
(\pm 100 , 0)\}
$
(Figure~\ref{picture_16.png}).

$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 6586$,

which gives $d(m, 2m + 14) \leq 6586$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_16.png}}
\parbox{1\linewidth}{\caption{IPS of cardinality 20 and diameter 6586}
\label{picture_16.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
(\pm 5544 , 0),
(\pm 4128 , 0),
(\pm 3108 , 0),
(\pm 2697 , 0),
$

$
(\pm 2547 , 0),
(\pm 1839 , 0),
(\pm 1728 , 0),
(\pm 904 , 0),
(\pm 300 , 0),
(-6148 , 0)\}
$
(Figure~\ref{picture_17.png}).

$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 11692$,

which gives $d(m, 2m + 15) \leq 11692$.

\begin{figure}[h!]
\center{\includegraphics[width=0.9\linewidth]{picture_17.png}}
\parbox{0.9\linewidth}{\caption{IPS of cardinality 21 and diameter 11692}
\label{picture_17.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
(\pm 6148 , 0),
(\pm 5544 , 0),
(\pm 4128 , 0),
(\pm 3108 , 0),
$

$
(\pm 2697 , 0),
(\pm 2547 , 0),
(\pm 1839 , 0),
(\pm 1728 , 0),
(\pm 904 , 0),
(\pm 300 , 0)\}
$
(Figure~\ref{picture_18.png}).

$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 12296$,

which gives $d(m, 2m + 16) \leq 12296$.

\begin{figure}[h!]
\center{\includegraphics[width=0.9\linewidth]{picture_18.png}}
\parbox{0.9\linewidth}{\caption{IPS of cardinality 22 and diameter 12296}
\label{picture_18.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
(\pm 6148 , 0),
(\pm 5544 , 0),
(\pm 4128 , 0),
(\pm 3108 , 0),
$

$
(\pm 2697  , 0),
(\pm 2547 , 0),
(\pm 1839 , 0),
(\pm 1728 , 0),
(\pm 904 , 0),
(\pm 300 , 0),
(-9879 , 0)\}
$
(Figure~\ref{picture_19.png}).

$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 16027$,

which gives $d(m, 2m + 17) \leq 16027$.

\begin{figure}[h!]
\center{\includegraphics[width=0.8\linewidth]{picture_19.png}}
\parbox{0.8\linewidth}{\caption{IPS of cardinality 23 and diameter 16027}
\label{picture_19.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
(\pm 9879 , 0),
(\pm 6148 , 0),
(\pm 5544 , 0),
(\pm 4128 , 0),
$

$
(\pm 3108 , 0),
(\pm 2697 , 0),
(\pm 2547 , 0),
(\pm 1839 , 0),
(\pm 1728 , 0),
(\pm 904 , 0),
(\pm 300 , 0)\}
$
(Figure~\ref{picture_20.png}).

$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 19758$,

which gives $d(m, 2m + 18) \leq 19758$.

\begin{figure}[h!]
\center{\includegraphics[width=0.8\linewidth]{picture_20.png}}
\parbox{0.8\linewidth}{\caption{IPS of cardinality 24 and diameter 19758}
\label{picture_20.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 5244, 360),
(\pm 12296 , 0),
(\pm 11088 , 0),
(\pm 8579 , 0),
(\pm 8256 , 0),
$

$
(\pm 6216 , 0),
(\pm 5394 , 0),
(\pm 5094 , 0),
(\pm 3678 , 0),
(\pm 3456 , 0),
(\pm 1808 , 0),
(\pm 600 , 0),
$

$
(-19758 , 0)\}
$
(Figure~\ref{picture_21.png}).

$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P})} = 32054$,

which gives $d(m, 2m + 19) \leq 32054$.

\begin{figure}[h!]
\center{\includegraphics[width=0.9\linewidth]{picture_21.png}}
\parbox{0.9\linewidth}{\caption{IPS of cardinality 25 and diameter 32054}
\label{picture_21.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 5244, 360),
(\pm 19758 , 0),
(\pm 12296 , 0),
(\pm 11088 , 0),
(\pm 8579 , 0),
$

$
(\pm 8256 , 0),
(\pm 6216 , 0),
(\pm 5394 , 0),
(\pm 5094 , 0),
(\pm 3678 , 0),
(\pm 3456 , 0),
(\pm 1808 , 0),
$

$
(\pm 600 , 0)\}
$
(Figure~\ref{picture_22.png}).

$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P})} = 39516$,

which gives $d(m, 2m + 20) \leq 39516$.

\begin{figure}[h!]
\center{\includegraphics[width=0.95\linewidth]{picture_22.png}}
\parbox{0.95\linewidth}{\caption{IPS of cardinality 26 and diameter 39516}
\label{picture_22.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
(\pm 116058 , 0),
(\pm 86072 , 0),
(\pm 77616 , 0),
(\pm 60053 , 0),
$

$
(\pm 57792 , 0),
(\pm 43512 , 0),
(\pm 37758 , 0),
(\pm 35658 , 0),
(\pm 25746 , 0),
(\pm 24192 , 0),
$

$
(\pm 12656 , 0),
(\pm 4200 , 0),
(-138306 , 0)\}
$
(Figure~\ref{picture_23.png}).

$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 254364$,

which gives $d(m, 2m + 21) \leq 254364$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_23.png}}
\parbox{1\linewidth}{\caption{IPS of cardinality 27 and diameter 254364}
\label{picture_23.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
(\pm 138306 , 0),
(\pm 116058 , 0),
(\pm 86072 , 0),
(\pm 77616 , 0),
$

$
(\pm 60053 , 0),
(\pm 57792 , 0),
(\pm 43512 , 0),
(\pm 37758 , 0),
(\pm 35658 , 0),
(\pm 25746 , 0),
$

$
(\pm 24192 , 0),
(\pm 12656 , 0),
(\pm 4200 , 0)\}
$
(Figure~\ref{picture_24.png}).

$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 276612$,

which gives $d(m, 2m + 22) \leq 276612$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_24.png}}
\parbox{1\linewidth}{\caption{IPS of cardinality 28 and diameter 276612}
\label{picture_24.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
(\pm 138306 , 0),
(\pm 116058 , 0),
(\pm 86072 , 0),
(\pm 77616 , 0),
$

$
(\pm 60053 , 0),
(\pm 57792 , 0),
(\pm 43512 , 0),
(\pm 37758 , 0),
(\pm 35658 , 0),
(\pm 25746 , 0),
$

$
(\pm 24192 , 0),
(\pm 12656 , 0),
(\pm 4200 , 0),
(-199863 , 0)\}
$
(Figure~\ref{picture_25.png}).

$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 338169$,

which gives $d(m, 2m + 23) \leq 338169$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_25.png}}
\parbox{1\linewidth}{\caption{IPS of cardinality 29 and diameter 338169}
\label{picture_25.png}}
\end{figure}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
(\pm 199863 , 0),
(\pm 138306 , 0),
(\pm 116058 , 0),
(\pm 86072 , 0),
$

$
(\pm 77616 , 0),
(\pm 60053 , 0),
(\pm 57792 , 0),
(\pm 43512 , 0),
(\pm 37758 , 0),
(\pm 35658 , 0),
$

$
(\pm 25746 , 0),
(\pm 24192 , 0),
(\pm 12656 , 0),
(\pm 4200 , 0)\}
$
(Figure~\ref{picture_26.png}).

$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 399726$,

which gives $d(m, 2m + 24) \leq 399726$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_26.png}}
\parbox{1\linewidth}{\caption{IPS of cardinality 30 and diameter 399726}
\label{picture_26.png}}
\end{figure}

\end{itemize}

\begin{remark}
For all the examples above, maximum in the right part of (\ref{formula1})
evaluates to the diameter of the system.
The following example shows that the maximum can be evaluated to $w$:
\end{remark}

\begin{itemize}
\setlength{\itemsep}{-1mm}

\item
$\mathcal{P}=\sqrt{2}/{1} * \{ (\pm 76, 60),
(\pm 62 , 0),
(\pm 29 , 0),
(\pm 6 , 0)\}
$
(Figure~\ref{picture_10.png}).

$f = 152$, $v = 114$, $w = 190$, $\operatorname{diam(\mathcal{P})} = 162$.

We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
is worse that known one (\hyperlink{d2}{2})).

\begin{figure}[htbp]
	\includegraphics[width=.48\linewidth]{picture_10.png}
	\hfill
	\includegraphics[width=.48\linewidth]{picture.png}
	\\
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 8 and diameter 162}
	\label{picture_10.png}}
	\hfill
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 8 and diameter 409}
	\label{picture.png}}
\end{figure}

\end{itemize}

Guy gives \cite[D 20]{guy2013unsolved} a system of $8$ points located
on two parallel lines. The points of the system form a rectangle
(Figure~\ref{rectangle.pdf}):

\begin{figure}[htbp]
	\includegraphics[width=.6\linewidth]{rectangle.pdf}
	\hfill
	\includegraphics[width=.5\linewidth]{right_angle.pdf}
	\\
	\parbox{.65\linewidth}{\caption{Distances}
	\label{rectangle.pdf}}
	\hfill
	\parbox{.5\linewidth}{\caption{Counterexample}
	\label{right_angle.pdf}}
\end{figure}

\begin{itemize}
\setlength{\itemsep}{-1mm}

\item
$\mathcal{P}=\sqrt{1}/{1} * \{ (0, \pm 60),
(182 , \pm 60),
(209 , \pm 60),
(391 , \pm 60)\}
$
(Figure~\ref{picture.png}),

where $n = 8$, $\operatorname{diam(\mathcal{P})} = 409$. Using the ``blowing up''
construction, we obtain
\begin{equation}\label{result1}
d(m, 4m) \leq 409
\end{equation}
Figure~\ref{picture_11.pdf} shows an example for $m = 3$.

\begin{figure}[h!]
\center{\includegraphics[width=0.9\linewidth]{picture_11.pdf}}
\parbox{0.9\linewidth}{\caption{IPS of cardinality 12 and diameter 409}
\label{picture_11.pdf}}
\end{figure}

\end{itemize}

\begin{theorem}
Let $\mathcal{P}$ be a planar integral point set consisting of $k$
points on the line $l_{1}$ and $k$ points on the line $l_{2}$. Besides, these
points are symmetric with respect to one of the bisectors of the angles
formed by the intersection of lines $l_{1}$ and $l_{2}$, then

\begin{equation}\label{formula2}
d(m, (k - \alpha)m + \alpha) \leq \operatorname{diam(\mathcal{P})},
\end{equation}

where
\begin{equation*}
\alpha =
\begin{cases}
1, \text{when the intersection point} \in \mathcal{P}, \\
0, \text{when the intersection point} \notin \mathcal{P}.
\end{cases}
\end{equation*}

\end{theorem}

\begin{remark}
The angles can be acute or obtuse. Figure~\ref{right_angle.pdf} shows that the intersection angle of the lines $l_{1}$ and $l_{2}$ cannot be equal to ${\pi}/2$.
\end{remark}

Below we give the planar integral point sets and the corresponding
estimates of the function $d(m, n)$ for $n = 3m + 1$ and $n = 4m + 1$.

\begin{figure}[htbp]
	\includegraphics[width=.48\linewidth]{picture_56.png}
	\hfill
	\includegraphics[width=.48\linewidth]{picture_1260.png}
	\\
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 7 and diameter 56}
	\label{picture_56.png}}
	\hfill
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 9 and diameter 1260}
	\label{picture_1260.png}}
\end{figure}

\begin{figure}[h!]
\center{\includegraphics[width=0.9\linewidth]{picture_12.pdf}}
\parbox{0.9\linewidth}{\caption{IPS of cardinality 10 and diameter 56}
\label{picture_12.pdf}}
\end{figure}

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_1260_R3.pdf}}
\parbox{1\linewidth}{\caption{IPS of cardinality 13 and diameter 1260}
\label{picture_1260_R3.pdf}}
\end{figure}

\begin{itemize}
\setlength{\itemsep}{-1mm}

\item
$\mathcal{P}=\sqrt{3}/{2} * \{ (\pm 56, 0),
(14, 0),
(-34, 0),
(-10, 24),
(-21 , 35),
(35, -21)\}
$
(Figure~\ref{picture_56.png}),

where $n = 7$, $\operatorname{diam(\mathcal{P})} = 56$. Using the ``blowing up''
construction, we obtain
\begin{equation}\label{result2}
d(m, 3m + 1) \leq 56
\end{equation}
The resulting estimate improves the estimate for $d(m, 3m)$, which is presented
in \cite{kemnitz1988punktmengen}. Figure~\ref{picture_12.pdf} shows an example
for $m = 3$.

\item
$\mathcal{P}=\sqrt{39}/{8} * \{ (\pm 5040, 0),
(1911, 315),
(2352, 0),
(944, 0),
(336, 0),
(2940, -420),
(2044, 220),
$

$
(735, 1155)\}
$
(Figure~\ref{picture_1260.png}),

where $n = 9$, $\operatorname{diam(\mathcal{P})} = 1260$. Using the ``blowing up''
construction, we obtain
\begin{equation}\label{result3}
d(m, 4m + 1) \leq 1260
\end{equation}
Figure~\ref{picture_1260_R3.pdf} shows an example for $m = 3$.

\end{itemize}

\bibliography{literature}

\end{document}

### 2019-09-07 semi-general-suplinear-bound/semi-general-suplinear-bound.tex
Semi-general bound - arXiv version

@@
 A planar integral point set is called to be in semi-general position, if it does not contain collinear triples.
-The existing bound for mininum diameter of planar integral point sets is linear.
-We prove a new bound for mininum diameter of planar integral point sets in semi-general position
-which is better than linear.
+The existing lower bound for mininum diameter of planar integral point sets is linear.
+We prove a new lower bound for mininum diameter of planar integral point sets in semi-general position
+that is better than linear.
 
@@
 
-Kurz - facher
-Bat-Ochir - circular
-Kurz - precise values
-Kurz - about many collinear points
+An \textit{integral point set} in a plane is a point set $M$ such that all the usual (Euclidean) distances between the
+points of $M$ are integers and $M$ is not situated on a straight line.
+Every integral point set consists of a finite number of points~\cite{anning1945integral,erdos1945integral};
+thus, we denote the set of all planar integral point sets of $n$ points by
+$\mathfrak{M}(2,n)$ (using the notation in~\cite{our-vmmsh-2018})
+and define the diameter of $M\in\mathfrak{M}(2,n)$ in the following natural way:
+\begin{equation}
+	\operatorname{diam} M = \max_{A,B\in M} |AB|
+	,
+\end{equation}
+where $|AB|$ denotes the Euclidean distance.
+The symbol $\# M$ will be used for cardinality of $M$, that is the number of points in $M$ in our case.
 
-Kurz - semi-general and general position, precise values
+Since every integral point set can obviously be dilated to a set of larger diameter,
+minimal possible diameters of sets of given cardinality are in the focus.
+To be precise,
+the following function was introduced~\cite{kurz2008bounds,kurz2008minimum}:
+\begin{equation}
+	d(2,n) = \min_{M\in\mathfrak{M}(2,n)} \operatorname{diam} M
+	.
+\end{equation}
 
-The upper bound
-Piepmeyer(?) - upper bound \cite{harborth1993upper}
+It turned out to be very easy to construct a planar integral point set of $n$ points with $n-1$ collinear ones and one point out of the line
+(so-called \textit{facher} sets);
+the same holds for 2 points out of the line (we refer the reader to~\cite{antonov2008maximal}, where some of such sets are called \textit{crabs})
+and even for 4 points out of the line~\cite{huff1948diophantine}.
+For $9\leq n\leq 122$, the minimal possible diameter is achieved at a facher set~\cite{kurz2008bounds}.
 
-Particular cases - Huff, Monographies (2?), \cite{our-pmm-2018}, \cite{our-ped-2018}.
+\begin{definition}
+	A set $M\in\mathfrak{M}(2,n)$ is called to be in \textit{semi-general position},
+	if no three points of $M$ are collinear.
+	The set of all planar integral point sets in semi-general position
+	is denoted by $\overline{\mathfrak{M}}(2,n)$.
+\end{definition}
+
+Furthermore, the constructions of integral point sets in semi-general position of arbitrary cardinality
+appeared~\cite{harborth1993upper};
+such sets are situated on a circle.
+Also, there is a sophisticated construction of a circular integral point set of arbitrary cardinality
+that gives the possible numbers of odd integral distances
+between points in the plane~\cite{piepmeyer1996maximum}.
+
+\begin{definition}
+	A set $M\in\overline{\mathfrak{M}}(2,n)$ is called to be in \textit{general position},
+	if no four points of $M$ are concyclic.
+	The set of all planar integral point sets in general position
+	is denoted by $\dot{\mathfrak{M}}(2,n)$.
+\end{definition}
+
+It remains unknown if there are integral points sets in general position of arbitrary cardinality;
+however, some sets $M\in \dot{\mathfrak{M}}(2,7)$ are known~\cite{kreisel2008heptagon,kurz2013constructing}.
+
+The inequality
+\begin{equation*}
+	d(2,n) \leq \overline{d}(2,n) \leq \dot{d}(2,n)
+	,
+\end{equation*}
+where
+$
+	\overline{d}(2,n) = \min_{M\in\overline{\mathfrak{M}}(2,n)} \operatorname{diam} M
+$
+and
+$
+	\dot{d}(2,n) = \min_{M\in\dot{\mathfrak{M}}(2,n)} \operatorname{diam} M
+$,
+is obvious; however, a more interesting relation holds:
+\begin{equation*}
+	c_1 n \leq d(2,n) \leq \overline{d}(2,n) \leq n^{c_2 \log \log n}
+	.
+\end{equation*}
+The upper bound is presented in~\cite{harborth1993upper}.
+The lower bound was firstly introduced in~\cite{solymosi2003note};
+the largest known value for $c_1$ is $5/11$ for $n\geq 4$~\cite{my-pps-linear-bound-2019}.
+
+There are some bounds for minimal diameter of planar integral point sets in some special positions.
+Assuming that the planar integral point sets contains many collinear points,
+the following result holds.
+\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
+	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
+	at least $n^\delta$ collinear points, there exists a $n_0 (\varepsilon)$
+	such that for all $n \geq n_0 (\varepsilon)$ we have
+	\begin{equation}
+		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
+		.
+	\end{equation}
+\end{theorem}
+For diameter bounds for circular sets, we refer the reader to~\cite{bat2018number}.
+
+Particular cases of planar integral point sets are also discussed
+in~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved},~\cite{our-pmm-2018},~\cite{our-ped-2018}.
+For generalizaton in higher dimensions and the appropriate bounds, see~\cite{kurz2005characteristic,nozaki2013lower}.
+
+In the present paper we give a special bound for planar integral point sets in semi-general position.
+The condition of semi-general position is important in the given proof.
 
@@
 	\cite[Lemma 4]{our-vmmsh-2018};
-	\cite[TODO]{my-pps-linear-bound-2019}
+	\cite[Lemma 2.4]{my-pps-linear-bound-2019}
 	\label{lem:square_container}
@@
 \begin{definition}
-	\cite[TODO]{my-pps-linear-bound-2019}
+	\cite[Definition 2.5]{my-pps-linear-bound-2019}
 	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
@@
 \begin{lemma}
-	\cite[Theorem TODO]{my-pps-linear-bound-2019}
+	\cite[Theorem 3.10]{my-pps-linear-bound-2019}
 	\label{lem:no_distance_one}
@@
 \begin{remark}
-	Lemma~\ref{lem:count_of_points_on_hyperbolas} is one of the variations of~\cite{erdos1945integral},
-	\cite[TODO]{my-pps-linear-bound-2019}, etc.
+	Lemma~\ref{lem:count_of_points_on_hyperbolas} is one of the variations of~\cite{erdos1945integral}.
 \end{remark}
@@
 	\label{thm:main_result}
-	For every $n\in\mathbb{N}$ we have
+	For every integer $n \geq 3$ we have
 	\begin{equation}
-		\overline{d}(2,n) \geq c n^{5/4}
-		,
+		\overline{d}(2,n) \geq (n/5)^{5/4}
+		.
 	\end{equation}
-	where $c = 5^{-5/4}$.
 \end{theorem}
@@
 
+%TODO: Solymosi's proof accepts a=b; in our case, a=b leads to a better bound (?)
+% and the condition a < b may (??) remove -1/4 from the lemma about triangle
+
 	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square with side length $p$.
@@
 	\end{equation}
-	The latter inequality holds because $\overline{d}(2,n) \geq 4$ for $n\geq 4$ [TODO: Kurz!]
+	The latter inequality holds because $\overline{d}(2,n) \geq 4$ for $n\geq 4$~~\cite{kurz2008minimum}
+	% TODO: add two more links from the kurz2008minimum's intro in case of lack of references
 	and $4^{4/5}>3$.
@@
 
-%TODO: a remark on how to easily derive 6/5 from Solymosi's corollary.
+\begin{remark}
+	The following result in known:
+\end{remark}
+
+\begin{lemma}
+	\cite[Corollary 1]{solymosi2003note}
+	For $H \in \overline {\mathfrak{M}}(2,n)$, the minimum
+	distance in H is at least $n^{1/3}$.
+\end{lemma}
+Applying the same technique, one can easily derive that
+\begin{equation}
+	n \leq 3 \frac{\operatorname{diam} H }{n^{1/6}}
+	,
+\end{equation}
+which leads to the bound
+\begin{equation}
+	\overline{d}(2,n) \geq c_3 n^{7/6}
+	,
+\end{equation}
+which is less than the one from Theorem~\ref{thm:main_result}.
 

### 2019-09-07 semi-general-suplinear-bound/semi-general-suplinear-bound.tex
Suplinear bound for semi-general position: Russian abstract

@@
 that is better than linear.
+
+%\selectlanguage{russian}
+
+\paragraph{Аннотация.}
+Множество точек $M$ на плоскости называется плоским множеством с целочисленными расстояниями,
+если все расстояния между точками $M$ суть целые числа,
+и при этом $M$ не содержится ни в какой прямой.
+Говорят, что плоское множество с целочисленными расстояниями есть множество полуобщего положения,
+если никакие три его точки не лежат на одной прямой.
+Известная оценка снизу на диаметр плоского множества с целочисленными расстояниями
+линейна относительно его мощности.
+В статье доказывается надлинейная оценка снизу
+на диаметр плоского множества с целочисленными расстояниями полуобщего положения.
+
+%\selectlanguage{english}
 

### 2019-09-08 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
ChebSbornik: template as is

\documentclass[11pt,twoside,draft
]{article}
\usepackage{amsmath,amsfonts,amssymb,amsthm,indentfirst,enumerate,textcomp}
\usepackage[utf8]{inputenc}
\usepackage{chebsb}
\usepackage[russian]{babel}
\usepackage{indentfirst, array}
\usepackage{amscd,latexsym}
\usepackage{mathrsfs}
%\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage{tabularx}
\usepackage{multirow}

\usepackage{graphicx}
\usepackage{textcase}% Оформление страниц

\label{beg}
% заглавие стати и аннотация

\levkolonttl{%левый колонтитул - авторы
И.~И.~Иванов}
\prvkolonttl{%правый колонтитул - сокращенное название статьи
О квадратах и кубиках \ldots}

\UDK{% УДК статьи
517}

\DOI{
10.22405/2226-8383-2018-\tom-\iss-\pageref{beg}-\pageref{end}
}

\title
{%название статьи на русском языке с указанием источника финансирования при необходимости
О квадратах и кубиках в множествах конечного поля\footnote{Исследование выполнено за счет гранта Российского научного фонда (проект ХХ-ХХ-ХХХХХ).}}
{%название статьи на английском языке
On the squares and cubes in the set of finite fields}

\author
{%авторы статьи с указанием города на русском языке
И. И. Иванов (г. Москва)}
{%авторы статьи с указанием города на английском языке
I. I. Ivanov (Moscow)}

\Cite{
И. И. Иванов О квадратах и кубиках в множествах конечного поля // Чебышевcкий сборник, 2019, т.~\tom, вып.~\iss, с.~\pageref{beg}--\pageref{end}.
}
{I. I. Ivanov, 2019, "On the squares and cubes in the set of finite fields"\,, {\it Che\-by\-shev\-skii sbornik}, vol.~\tom, no.~\iss, pp.~\pageref{beg}--\pageref{end}.
}

\info
{%авторы статьи с указанием аффилиации на русском языке
\noindent {\bf И. И. Иванов}~---

\noindent
\emph{e-mail: ---}

}
{%авторы статьи с указанием аффилиации на английском языке
\noindent {\bf I. I. Ivanov}~---

\noindent
\emph{e-mail: ---}

}

\Abstract
{%Аннотация статьи на русском языке 150-250 слов с учетом ключевых слов
В статье...
}
{%Аннотация статьи на английском языке
In paper...
}

\keywords
{%ключевые слова на русском языке
конечные поля, квадраты, суммы.
}
{%ключевые слова на английском языке
finite fields, squares, sums.
}

%число наименований в библиографии
\Bibliography{18 названий.}{18 titles.}

\begin{document}

%генерация заглавия статьи
\maketitle

\enmaketitle

\section{Введение}

\section{Основной текст статьи}

\section{Заключение}

%библиография по ГОСТу
\begin{thebibliography}{99}

	\bibitem{1}
	Banks~W.\,D., Conflitti~A., Shparlinski~I.\,E. Character sums over integers with restricted $g$-ary digits // Illinois J. Math. 2002. Vol. 46, №3. P. 819-836.

\end{thebibliography}

%библиография по Гарвардскому стандарту
\begin{engbibliography}{99}

	\bibitem{en1}
	Banks, W.\,D., Conflitti, A. \& Shparlinski, I.\,E. 2002, ``Character sums over integers with restricted $g$-ary digits``, \textit{Illinois J. Math.}, vol. 46, no. 3, pp. 819-836.

\end{engbibliography}

\label{end}

\end{document}

### 2019-09-08 semi-general-suplinear-bound/semi-general-suplinear-bound.tex
Longer annotations
Chebyshevskiy Sbornik requires 150-250 words

@@
 A planar integral point set is called to be in semi-general position, if it does not contain collinear triples.
-The existing lower bound for mininum diameter of planar integral point sets is linear.
+The existing lower bound for mininum diameter of planar integral point sets is linear on its cardinality.
+There were no known special diameter bounds for planar integral point sets in semi-general position
+(the known upper bound for planar integral point sets is constructive
+and provides planar integral point sets in semi-general position).
 We prove a new lower bound for mininum diameter of planar integral point sets in semi-general position
-that is better than linear.
+that is better than linear (polynomial with power $5/4$).
+The proof is based on several lemmas observations, including the ones provided by Solymosi
+while proving the first linear lower bound for diameter of a planar integral point set.
+We have to admit that our bound is still not tight.
 
@@
 линейна относительно его мощности.
+Ранее не были известны отдельные оценки снизу на диаметр плоских множеств с целочисленными расстояниями полуобщего положения
+заданной мощности
+(известная конструктивная оценка сверху на диаметр плоских множеств с целочисленными расстояниями
+использует как раз множества полуобщего положения).
 В статье доказывается надлинейная оценка снизу
-на диаметр плоского множества с целочисленными расстояниями полуобщего положения.
+на диаметр плоского множества с целочисленными расстояниями полуобщего положения
+(полиномиальная с показателем $5/4$).
+Доказательство основано на относительно большом количестве лемм и наблюдений,
+включая результаты Солимоси
+из статьи, в которой была впервые доказана линейная оценка снизу
+на диаметр плоских множеств с целочисленными расстояниями.
+Полученная вданной статье оценка всё ещё не точна.
 

### 2019-09-08 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
The paper in ChebSb template

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% {{ biblatex
+%% Команда \Cite (с большой буквы) у biblatex своя.
+%% Сохраняем "нашу"
+\let\CiteChebSb=\Cite
+%% Убираем \Cite
+\let\Cite=\relax
+%% Вот теперь смело загружаем biblatex
+\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+\addbibresource{../common/notmy.bib}
+\addbibresource{../common/my.bib}
+%% Восстанавливаем "нашу" \Cite
+\let\Cite=\CiteChebSb
+%% }} biblatex
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Автоматическая нумерация теорем (коль скоро подключен amsthm)
+\theoremstyle{plain}
+%\newtheorem{theorem}{Theorem}[section]
+%\newtheorem{conjecture}[theorem]{Conjecture}
+%\newtheorem{lemma}[theorem]{Lemma}
+%\newtheorem{corollary}[theorem]{Corollary}
+%\newtheorem{proposition}[theorem]{Proposition}
+
+%\theoremstyle{definition}
+%\newtheorem{definition}[theorem]{Definition}
+%\newtheorem{construction}[theorem]{Construction}
+%\newtheorem{remark}[theorem]{Remark}
+%\newtheorem{problem}[theorem]{Problem}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \label{beg}
@@
 \levkolonttl{%левый колонтитул - авторы
-И.~И.~Иванов}
+N.~N.~Avdeev}
 \prvkolonttl{%правый колонтитул - сокращенное название статьи
-О квадратах и кубиках \ldots}
+On diameter bounds for planar integral point sets \ldots}
 
 \UDK{% УДК статьи
-517}
+519.146}
 
@@
 {%название статьи на русском языке с указанием источника финансирования при необходимости
-О квадратах и кубиках в множествах конечного поля\footnote{Исследование выполнено за счет гранта Российского научного фонда (проект ХХ-ХХ-ХХХХХ).}}
+Об оценках на диаметр плоских множеств с целочисленными расстояниями полуобщего положения%
+\footnote{Исследование выполнено за счет гранта Российского научного фонда (проект 19-11-00197).}}
 {%название статьи на английском языке
-On the squares and cubes in the set of finite fields}
+On diameter bounds for planar integral point sets in semi-general position}
 
@@
 {%авторы статьи с указанием города на русском языке
-И. И. Иванов (г. Москва)}
+Н. Н. Авдеев (г. Воронеж)}
 {%авторы статьи с указанием города на английском языке
-I. I. Ivanov (Moscow)}
+N. N. Avdeev (Voronezh)}
 
 \Cite{
-И. И. Иванов О квадратах и кубиках в множествах конечного поля // Чебышевcкий сборник, 2019, т.~\tom, вып.~\iss, с.~\pageref{beg}--\pageref{end}.
+Н. Н. Авдеев Об оценках на диаметр плоских множеств с целочисленными расстояниями полуобщего положения // Чебышевcкий сборник, 2019, т.~\tom, вып.~\iss, с.~\pageref{beg}--\pageref{end}.
 }
-{I. I. Ivanov, 2019, "On the squares and cubes in the set of finite fields"\,, {\it Che\-by\-shev\-skii sbornik}, vol.~\tom, no.~\iss, pp.~\pageref{beg}--\pageref{end}.
+{N. N. Avdeev, 2019, "On diameter bounds for planar integral point sets in semi-general position"\,, {\it Che\-by\-shev\-skii sbornik}, vol.~\tom, no.~\iss, pp.~\pageref{beg}--\pageref{end}.
 }
@@
 {%авторы статьи с указанием аффилиации на русском языке
-\noindent {\bf И. И. Иванов}~---
+\noindent {\bf Н. Н. Авдеев}~--- Воронежский государственный университет, кафедра теории функций и геометрии
 
 \noindent
-\emph{e-mail: ---}
+\emph{e-mail: nickkolok@mail.ru, avdeev@math.vsu.ru}
 
@@
 {%авторы статьи с указанием аффилиации на английском языке
-\noindent {\bf I. I. Ivanov}~---
+\noindent {\bf N. N. Avdeev}~--- Voronezh State University
 
 \noindent
-\emph{e-mail: ---}
+\emph{e-mail: nickkolok@mail.ru, avdeev@math.vsu.ru}
 
@@
 {%Аннотация статьи на русском языке 150-250 слов с учетом ключевых слов
-В статье...
+Множество точек $M$ на плоскости называется плоским множеством с целочисленными расстояниями,
+если все расстояния между точками $M$ суть целые числа,
+и при этом $M$ не содержится ни в какой прямой.
+Говорят, что плоское множество с целочисленными расстояниями есть множество полуобщего положения,
+если никакие три его точки не лежат на одной прямой.
+Известная оценка снизу на диаметр плоского множества с целочисленными расстояниями
+линейна относительно его мощности.
+Ранее не были известны отдельные оценки снизу на диаметр плоских множеств с целочисленными расстояниями полуобщего положения
+заданной мощности
+(известная конструктивная оценка сверху на диаметр плоских множеств с целочисленными расстояниями
+использует как раз множества полуобщего положения).
+В статье доказывается надлинейная оценка снизу
+на диаметр плоского множества с целочисленными расстояниями полуобщего положения
+(полиномиальная с показателем $5/4$).
+Доказательство основано на относительно большом количестве лемм и наблюдений,
+включая результаты Солимоси
+из статьи, в которой была впервые доказана линейная оценка снизу
+на диаметр плоских множеств с целочисленными расстояниями.
+Полученная вданной статье оценка всё ещё не точна.
 }
 {%Аннотация статьи на английском языке
-In paper...
+A point set $M$ in the Euclidean plane is called a planar integral point set if all the distances between the
+elements of $M$ are integers, and $M$ is not situated on a straight line.
+A planar integral point set is called to be in semi-general position, if it does not contain collinear triples.
+The existing lower bound for mininum diameter of planar integral point sets is linear on its cardinality.
+There were no known special diameter bounds for planar integral point sets in semi-general position
+(the known upper bound for planar integral point sets is constructive
+and provides planar integral point sets in semi-general position).
+We prove a new lower bound for mininum diameter of planar integral point sets in semi-general position
+that is better than linear (polynomial with power $5/4$).
+The proof is based on several lemmas observations, including the ones provided by Solymosi
+while proving the first linear lower bound for diameter of a planar integral point set.
+We have to admit that our bound is still not tight.
 }
@@
 {%ключевые слова на русском языке
-конечные поля, квадраты, суммы.
+комбинаторная геометрия, диаметр множества, множество с целочисленными расстояниями.
 }
 {%ключевые слова на английском языке
-finite fields, squares, sums.
+combinatorial geometry, diameter of a set, integral point set.
 }
@@
 %число наименований в библиографии
-\Bibliography{18 названий.}{18 titles.}
+\Bibliography{21 название.}{21 titles.}
 
@@
 
-\section{Введение}
+\section{Introduction}
 
-\section{Основной текст статьи}
+An \textit{integral point set} in a plane is a point set $M$ such that all the usual (Euclidean) distances between the
+points of $M$ are integers and $M$ is not situated on a straight line.
+Every integral point set consists of a finite number of points~\cite{anning1945integral,erdos1945integral};
+thus, we denote the set of all planar integral point sets of $n$ points by
+$\mathfrak{M}(2,n)$ (using the notation in~\cite{our-vmmsh-2018})
+and define the diameter of $M\in\mathfrak{M}(2,n)$ in the following natural way:
+\begin{equation}
+	\operatorname{diam} M = \max_{A,B\in M} |AB|
+	,
+\end{equation}
+where $|AB|$ denotes the Euclidean distance.
+The symbol $\# M$ will be used for cardinality of $M$, that is the number of points in $M$ in our case.
 
-\section{Заключение}
+Since every integral point set can obviously be dilated to a set of larger diameter,
+minimal possible diameters of sets of given cardinality are in the focus.
+To be precise,
+the following function was introduced~\cite{kurz2008bounds,kurz2008minimum}:
+\begin{equation}
+	d(2,n) = \min_{M\in\mathfrak{M}(2,n)} \operatorname{diam} M
+	.
+\end{equation}
+
+It turned out to be very easy to construct a planar integral point set of $n$ points with $n-1$ collinear ones and one point out of the line
+(so-called \textit{facher} sets);
+the same holds for 2 points out of the line (we refer the reader to~\cite{antonov2008maximal}, where some of such sets are called \textit{crabs})
+and even for 4 points out of the line~\cite{huff1948diophantine}.
+For $9\leq n\leq 122$, the minimal possible diameter is achieved at a facher set~\cite{kurz2008bounds}.
+
+\begin{definition}
+	A set $M\in\mathfrak{M}(2,n)$ is called to be in \textit{semi-general position},
+	if no three points of $M$ are collinear.
+	The set of all planar integral point sets in semi-general position
+	is denoted by $\overline{\mathfrak{M}}(2,n)$.
+\end{definition}
+
+Furthermore, the constructions of integral point sets in semi-general position of arbitrary cardinality
+appeared~\cite{harborth1993upper};
+such sets are situated on a circle.
+Also, there is a sophisticated construction of a circular integral point set of arbitrary cardinality
+that gives the possible numbers of odd integral distances
+between points in the plane~\cite{piepmeyer1996maximum}.
+
+\begin{definition}
+	A set $M\in\overline{\mathfrak{M}}(2,n)$ is called to be in \textit{general position},
+	if no four points of $M$ are concyclic.
+	The set of all planar integral point sets in general position
+	is denoted by $\dot{\mathfrak{M}}(2,n)$.
+\end{definition}
+
+It remains unknown if there are integral points sets in general position of arbitrary cardinality;
+however, some sets $M\in \dot{\mathfrak{M}}(2,7)$ are known~\cite{kreisel2008heptagon,kurz2013constructing}.
+
+The inequality
+\begin{equation*}
+	d(2,n) \leq \overline{d}(2,n) \leq \dot{d}(2,n)
+	,
+\end{equation*}
+where
+$
+	\overline{d}(2,n) = \min_{M\in\overline{\mathfrak{M}}(2,n)} \operatorname{diam} M
+$
+and
+$
+	\dot{d}(2,n) = \min_{M\in\dot{\mathfrak{M}}(2,n)} \operatorname{diam} M
+$,
+is obvious; however, a more interesting relation holds:
+\begin{equation*}
+	c_1 n \leq d(2,n) \leq \overline{d}(2,n) \leq n^{c_2 \log \log n}
+	.
+\end{equation*}
+The upper bound is presented in~\cite{harborth1993upper}.
+The lower bound was firstly introduced in~\cite{solymosi2003note};
+the largest known value for $c_1$ is $5/11$ for $n\geq 4$~\cite{my-pps-linear-bound-2019}.
+
+There are some bounds for minimal diameter of planar integral point sets in some special positions.
+Assuming that the planar integral point sets contains many collinear points,
+the following result holds.
+\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
+	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
+	at least $n^\delta$ collinear points, there exists a $n_0 (\varepsilon)$
+	such that for all $n \geq n_0 (\varepsilon)$ we have
+	\begin{equation}
+		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
+		.
+	\end{equation}
+\end{theorem}
+For diameter bounds for circular sets, we refer the reader to~\cite{bat2018number}.
+
+Particular cases of planar integral point sets are also discussed
+in~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved},~\cite{our-pmm-2018},~\cite{our-ped-2018}.
+For generalizaton in higher dimensions and the appropriate bounds, see~\cite{kurz2005characteristic,nozaki2013lower}.
+
+In the present paper we give a special bound for planar integral point sets in semi-general position.
+The condition of semi-general position is important in the given proof.
+
+\section{Preliminary results}
+
+In this section, we give some lemmas which will be used for the proof.
+
+\begin{lemma}
+	\cite[Observation 1]{solymosi2003note}
+	If a triangle $T$ has integer side-lengths $a \leq b \leq c$,
+	then the minimal height $m$ of it is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
+\end{lemma}
+
+\begin{definition}
+	The part of a plane between two parallel straight lines with distance $\rho$ between the lines
+	is called a strip of width $\rho$.
+\end{definition}
+
+\begin{lemma}
+	\cite{smurov1998stripcoverings}
+	If a triange $T$ with minimal height $\rho$ is situated in a strip,
+	then the width of a strip is at least $\rho$.
+\end{lemma}
+
+\begin{corollary}
+	\label{cor:solymosi_strip}
+	If a triangle $T$ with integer side-lengths $a \leq b \leq c$ is situated in a strip,
+	then the width of a strip is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
+\end{corollary}
+
+\begin{lemma}
+	\cite[Lemma 4]{our-vmmsh-2018};
+	\cite[Lemma 2.4]{my-pps-linear-bound-2019}
+	\label{lem:square_container}
+	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
+	Then $M$ is situated in a square of side length $d$.
+\end{lemma}
+
+\begin{definition}
+	\cite[Definition 2.5]{my-pps-linear-bound-2019}
+	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
+	the line through $M_1$ and $M_2$,
+	and the perpendicular bisector of line segment $M_1 M_2$.
+\end{definition}
+
+\begin{lemma}
+	\cite[Theorem 3.10]{my-pps-linear-bound-2019}
+	\label{lem:no_distance_one}
+	Each set $M\in\mathfrak{M}(2,n)$
+	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
+	consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
+	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
+\end{lemma}
+
+\begin{lemma}
+	\label{lem:count_of_points_on_hyperbolas}
+	Let $\{M_1, M_2, M_3, M_4\} \subset M\in\overline{\mathfrak{M}}(2,n)$
+	(points $M_2$ and $M_3$ may coincide, other points may not), $n\geq 4$.
+	Then $\# M \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
+\end{lemma}
+
+\begin{remark}
+	Lemma~\ref{lem:count_of_points_on_hyperbolas} is one of the variations of~\cite{erdos1945integral}.
+\end{remark}
+
+\begin{proof}
+	Each point $N\in M$ satisfies one of the following conditions:
+
+	a) $N$ belongs to $cr(M_1,M_2)$~--- overall at most 4 points;
+
+	b) $N$ belongs to $cr(M_3,M_4)$~--- overall at most 4 points;
+
+	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas
+	with one of $|M_3 M_4| - 1$ hyperbolas~--- overall at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
+
+	Due to Lemma~\ref{lem:no_distance_one} we have $|M_1 M_2| \geq 2$ and $|M_3 M_4| \geq 2$.
+	Since
+	\begin{multline}
+		4 (|M_1 M_2| - 1)(|M_3 M_4| - 1) + 4 + 4
+		=
+		4 ( (|M_1 M_2| - 1)(|M_3 M_4| - 1) + 2)
+		=
+		\\=
+		4 ( |M_1 M_2| \cdot |M_3 M_4| + 1 - |M_1 M_2| - |M_3 M_4| + 2)
+		=
+		\\=
+		4 ( |M_1 M_2| \cdot |M_3 M_4| + (1 - |M_1 M_2|) + (2 - |M_3 M_4|))
+		\leq
+		4 |M_1 M_2| \cdot |M_3 M_4|
+		,
+	\end{multline}
+	we are done.
+\end{proof}
+
+\section{The main result}
+
+\begin{theorem}
+	\label{thm:main_result}
+	For every integer $n \geq 3$ we have
+	\begin{equation}
+		\overline{d}(2,n) \geq (n/5)^{5/4}
+		.
+	\end{equation}
+\end{theorem}
+
+\begin{proof}
+	For $n = 3$ we have $\overline{d}(2,3) = 1$.
+	Consider $M\in\overline{\mathfrak{M}}(2,n)$, $n \geq 4$, $\operatorname{diam} M = p$.
+
+	Let us choose points $M_1, M_2, M_3, M_4 \in M$
+	(points $M_2$ and $M_3$ may coincide, other points may not), such that
+	\begin{equation}
+		\min_{A, B \in M} |AB| = |M_1 M_2|
+		,
+	\end{equation}
+	\begin{equation}
+		\min_{A, B \in M \setminus \{M_1\}} |AB| = |M_3 M_4| = m
+		.
+	\end{equation}
+
+	For $m \leq p^{2/5}$, Lemma~\ref{lem:count_of_points_on_hyperbolas} yields that
+	\begin{equation}
+		n \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| \leq  4 p^{4/5}
+		,
+	\end{equation}
+	or, that is the same,
+	\begin{equation}
+		\label{eq:hyperbolas_5_4}
+		p \geq (n/4) ^ {5/4} > (n/5) ^ {5/4}
+		.
+	\end{equation}
+
+	So, let us consider $m > p^{2/5}$.
+	Then for any $A,B \in M\setminus\{M_1\}$ the inequality $|AB| > p^{2/5}$ holds.
+	Due to Corollary~\ref{cor:solymosi_strip}, no three points of $M\setminus\{M_1\}$
+	are located in a strip of width $p^{1/5} / 2$.
+
+%TODO: Solymosi's proof accepts a=b; in our case, a=b leads to a better bound (?)
+% and the condition a < b may (??) remove -1/4 from the lemma about triangle
+
+	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square with side length $p$.
+	Let us partition this square into $q$ strips, $2p^{4/5} \leq q < 2p^{4/5} + 1$, each of width at most $p^{1/5} / 2$.
+	Every strip contains at most two points of  $M\setminus\{M_1\}$,
+	thus
+	\begin{equation}
+		\label{eq:strips_4_5}
+		n \leq 2(2p^{4/5} + 1) + 1
+		= 4p^{4/5}+3
+		\leq 5 p^{4/5}
+		.
+	\end{equation}
+	The latter inequality holds because $\overline{d}(2,n) \geq 4$ for $n\geq 4$~~\cite{kurz2008minimum}
+	% TODO: add two more links from the kurz2008minimum's intro in case of lack of references
+	and $4^{4/5}>3$.
+	From the inequality~\eqref{eq:strips_4_5} one can easily derive that
+	\begin{equation}
+		\label{eq:strips_5_4}
+		p \geq (n/5) ^ {5/4}
+		.
+	\end{equation}
+\end{proof}
+
+\begin{remark}
+	The following result in known:
+\end{remark}
+
+\begin{lemma}
+	\cite[Corollary 1]{solymosi2003note}
+	For $H \in \overline {\mathfrak{M}}(2,n)$, the minimum
+	distance in H is at least $n^{1/3}$.
+\end{lemma}
+Applying the same technique, one can easily derive that
+\begin{equation}
+	n \leq 3 \frac{\operatorname{diam} H }{n^{1/6}}
+	,
+\end{equation}
+which leads to the bound
+\begin{equation}
+	\overline{d}(2,n) \geq c_3 n^{7/6}
+	,
+\end{equation}
+which is less than the one from Theorem~\ref{thm:main_result}.
+
+\section{Conclusion}
+The presented bound is the first special lower bound for sets in semi-general position.
+Thus, we did not accepted the challenge to make the constant in Theorem~\ref{thm:main_result} as large as possible,
+in order to keep the ideas of the proof clear and understandable.
+A more thorough research can be done in the future to enlarge the constant.
+However, the upper and lower bounds are still not tight.
+
+\section{Acknowledgements}
+Author thanks Dr. Prof. E.M. Semenov for proofreading and valuable advice.
+
+\printbibliography
 

### 2019-09-11 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
English theorem-environments

@@
 %% Автоматическая нумерация теорем (коль скоро подключен amsthm)
-\theoremstyle{plain}
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{conjecture}[theorem]{Conjecture}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newtheorem{corollary}[theorem]{Corollary}
-%\newtheorem{proposition}[theorem]{Proposition}
 
-%\theoremstyle{definition}
-%\newtheorem{definition}[theorem]{Definition}
-%\newtheorem{construction}[theorem]{Construction}
-%\newtheorem{remark}[theorem]{Remark}
-%\newtheorem{problem}[theorem]{Problem}
+% К сожалению, английского ``Definition'' в стилевом файле нет.
+% Добавляю по аналогии.
+\newtheorem{Definition}{\indent {\sc Definition}}
+\renewcommand{\theDefinition}{\rm \arabic{Definition}}
+
+\newtheorem{Corollary}{\indent {\sc Corollary}}
+\renewcommand{\theCorollary}{\rm \arabic{Corollary}}
+
+\newtheorem{Remark}{\indent {\sc Remark}}
+\renewcommand{\theRemark}{\rm \arabic{Remark}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{definition}
+\begin{Definition}
 	A set $M\in\mathfrak{M}(2,n)$ is called to be in \textit{semi-general position},
@@
 	is denoted by $\overline{\mathfrak{M}}(2,n)$.
-\end{definition}
+\end{Definition}
 
@@
 
-\begin{definition}
+\begin{Definition}
 	A set $M\in\overline{\mathfrak{M}}(2,n)$ is called to be in \textit{general position},
@@
 	is denoted by $\dot{\mathfrak{M}}(2,n)$.
-\end{definition}
+\end{Definition}
 
@@
 the following result holds.
-\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
+\begin{Theorem}~\cite[Theorem 4]{kurz2008minimum}
 	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
@@
 	\end{equation}
-\end{theorem}
+\end{Theorem}
 For diameter bounds for circular sets, we refer the reader to~\cite{bat2018number}.
@@
 
-\begin{definition}
+\begin{Definition}
 	The part of a plane between two parallel straight lines with distance $\rho$ between the lines
 	is called a strip of width $\rho$.
-\end{definition}
+\end{Definition}
 
@@
 
-\begin{corollary}
+\begin{Corollary}
 	\label{cor:solymosi_strip}
@@
 	then the width of a strip is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
-\end{corollary}
+\end{Corollary}
 
@@
 
-\begin{definition}
+\begin{Definition}
 	\cite[Definition 2.5]{my-pps-linear-bound-2019}
@@
 	and the perpendicular bisector of line segment $M_1 M_2$.
-\end{definition}
+\end{Definition}
 
@@
 
-\begin{remark}
+\begin{Remark}
 	Lemma~\ref{lem:count_of_points_on_hyperbolas} is one of the variations of~\cite{erdos1945integral}.
-\end{remark}
+\end{Remark}
 
@@
 
-\begin{theorem}
+\begin{Theorem}
 	\label{thm:main_result}
@@
 	\end{equation}
-\end{theorem}
+\end{Theorem}
 
@@
 
-\begin{remark}
+\begin{Remark}
 	The following result in known:
-\end{remark}
+\end{Remark}
 

### 2019-09-11 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
ChebSb bibliography - WIP

@@
 \usepackage[utf8]{inputenc}
+\usepackage[T2A]{fontenc}
 \usepackage{chebsb}
-\usepackage[russian]{babel}
+\usepackage[english,russian]{babel}
 \usepackage{indentfirst, array}
@@
 %% Вот теперь смело загружаем biblatex
-\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+\usepackage[backend=biber,style=gost-numeric,sorting=none,dashed=false]{biblatex}
+%\usepackage[backend=biber,style=bath,sorting=ynt]{biblatex}
 \addbibresource{../common/notmy.bib}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\defbibenvironment{thebibliography}
+  {\textbackslash begin\{thebibliography\}\{99\}
+   % Page numbers and page headings are not wanted when copying
+   \thispagestyle{empty}
+   \pagestyle{empty}
+   \hyphenpenalty=10000
+   \exhyphenpenalty=10000
+   \renewcommand*{\&}{\textbackslash \symbol{38}}%
+   \renewcommand*{\mkbibemph}[1]{\textbackslash emph\{##1\}}%
+   \renewcommand*{\mkbibitalic}[1]{\textbackslash textit\{##1\}}%
+   \renewcommand*{\mkbibbold}[1]{\textbackslash textbf\{##1\}}%
+   \list{}{}}
+  {\endlist
+   \textbackslash end\{thebibliography\}
+   \thispagestyle{empty}
+   \pagestyle{empty}
+  }
+  {\item\textbackslash bibitem\{\printfield{entrykey}\}\\}
 
@@
 
-\printbibliography
+%\printbibliography
+
+%\selectlanguage{english}
+%\printbibliography[env=thebibliography]
 
 %библиография по ГОСТу
-\begin{thebibliography}{99}
 
-	\bibitem{1}
-	Banks~W.\,D., Conflitti~A., Shparlinski~I.\,E. Character sums over integers with restricted $g$-ary digits // Illinois J. Math. 2002. Vol. 46, №3. P. 819-836.
+% Мне библиографию в требуемом формате сгенерировал biblatex:
+% https://github.com/odomanov/biblatex-gost/issues/20
 
-\end{thebibliography}
+\input{blg-GOST}
+
+%\begin{thebibliography}{99}
+%
+%	\bibitem{1}
+%	Banks~W.\,D., Conflitti~A., Shparlinski~I.\,E. Character sums over integers with restricted $g$-ary digits // Illinois J. Math. 2002. Vol. 46, №3. P. 819-836.
+%
+%
+%\end{thebibliography}
 
 %библиография по Гарвардскому стандарту
-\begin{engbibliography}{99}
 
-	\bibitem{en1}
-	Banks, W.\,D., Conflitti, A. \& Shparlinski, I.\,E. 2002, ``Character sums over integers with restricted $g$-ary digits``, \textit{Illinois J. Math.}, vol. 46, no. 3, pp. 819-836.
+\input{blg-Harvard}
 
-\end{engbibliography}
+%\begin{engbibliography}{99}
+%
+%	\bibitem{en1}
+%	Banks, W.\,D., Conflitti, A. \& Shparlinski, I.\,E. 2002, ``Character sums over integers with restricted $g$-ary digits``, \textit{Illinois J. Math.}, vol. 46, no. 3, pp. 819-836.
+%
+%\end{engbibliography}
 

### 2019-09-11 semi-general-suplinear-bound/blg-Harvard.tex
ChebSb bibliography - WIP

\begin{engbibliography}{99}

\bibitem{ENGanning1945integral}
	Anning, N.H. \& Erdös, P. 1945.
	“Integral distances”.
	\emph{Bulletin of the American Mathematical Society}, vol. 51.8, pp. 598—600.

\bibitem{ENGerdos1945integral}
	Erdös, P. 1945.
	“Integral distances”.
	\emph{Bulletin of the American Mathematical Society}, vol. 51.12, p. 996.

\bibitem{ENGour-vmmsh-2018}
	Avdeev, N.N. \& Semenov, E.M. 2018.
	``Mnozhestva tochek c tselochislennymi rasstoyaniyami na ploskosti i v evklidovom prostranstve''
	(``Integral point sets on the plane and in Euclidean space'')
	\emph{Matematicheskiy forum (Itogi nauki. Yug Rossii)}, pp. 217—236.

\bibitem{ENGkurz2008bounds}
	Kurz, S. \& Laue, R. 2007.
	“Bounds for the minimum diameter of integral point sets”.
	\emph{Australasian Journal of Combinatorics}, vol. 39, pp. 233—240. arXiv: 0804.1296.

\bibitem{ENGkurz2008minimum}
	Kurz, S. \& Wassermann, A. 2011.
	“On the minimum diameter of plane integral point sets”.
	\emph{Ars Combinatoria}, vol. 101, pp. 265—287. arXiv: 0804.1307.

\bibitem{ENGantonov2008maximal}
	Antonov, A.R. \& Kurz, S. 2008.
	“Maximal integral point sets over $Z^2$”.
	\emph{International Journal of Computer Mathematics}, vol. 87.12, pp. 2653—2676. arXiv: 0804.1280.

\bibitem{ENGhuff1948diophantine}
	Huff, G.B. 1948.
	“Diophantine problems in geometry and elliptic ternary forms”.
	\emph{Duke Mathematical Journal}, vol. 15.2, pp. 443—453.

\bibitem{ENGharborth1993upper}
	Harborth, H., Kemnitz, A. \& Möller, M. 1993.
	“An upper bound for the minimum diameter of integral point sets”.
	\emph{Discrete \& Computational Geometry} vol. 9.4, pp. 427—432.

\bibitem{ENGpiepmeyer1996maximum}
	Piepmeyer, L. 1996.
	“The maximum number of odd integral distances between points in the plane”.
	\emph{Discrete \& Computational Geometry}, vol. 16.1, pp. 113—115.

\bibitem{ENGkreisel2008heptagon}
	Kreisel, T. \& Kurz, S. 2008.
	“There are integral heptagons, no three points on a line, no four on a circle”.
	\emph{Discrete \& Computational Geometry}, vol. 39.4, pp. 786—790.

\bibitem{ENGkurz2013constructing}
	Kurz, S., Noll, L.C., Rathbun, R, \& Simmons, C. 2014.
	“Constructing 7-clusters”.
	\emph{Serdica Journal of Computing}, vol. 8.1, pp. 47—70. arXiv: 1312.2318.

\bibitem{ENGsolymosi2003note}
	Solymosi, J. 2003.
	“Note on integral distances”.
	\emph{Discrete \& Computational Geometry}, vol. 30.2, pp. 337—342.

\bibitem{ENGmy-pps-linear-bound-2019}
	Avdeev, N. 2019.
	“On existence of integral point sets and their diameter bounds”.
	arXiv: 1906.11926

\bibitem{ENGbat2018number}
	Bat-Ochir, G. 2018.
	“On the number of points with pairwise integral distances on a circle”.
	\emph{Discrete Applied Mathematics}, vol. 254, pp. 17—32.

\bibitem{ENGbrass2006research}
	Brass, P., Moser, W.O.J. \& Pach, J. 2006.
	\emph{Research problems in discrete geometry}. Springer Science \& Business Media.

\bibitem{ENGguy2013unsolved}
	Guy, R. 2013.
	\emph{Unsolved problems in number theory}. Vol. 1.
	Springer Science \& Business Media.

\bibitem{ENGour-pmm-2018}
	Avdeev, N.N. 2018.
	“Ob otyskanii tseloudalennykh mnozhestv spetsial'nogo vida”
	(``On the search of integral point sets of a special type'').
	\emph{
		Aktual'nye problemy prikladnoj matematiki, informatiki i mekhaniki
		- sbornik trudov Mezhdunarodnoj nauchnoj konferencii.
	}
	(Actual problems of applied mathematics, informatics and mechanics
	- proc. of the int. conf.).
	Voronezh, pp. 492—498.

\bibitem{ENGour-ped-2018}
	Avdeev, N.N. 2018.
	“On integral point sets in special position”.
	\emph{Nekotorye voprosy analiza, algebry, geometrii i matematicheskogo obrazovaniya}
	(Some problems of analysis, algebra, geometry and mathematical education), vol. 8, pp. 5—6.

\bibitem{ENGkurz2005characteristic}
	Kurz, S. 2006.
	“On the characteristic of integral point sets in $E^m$”.
	\emph{Australasian Journal of Combinatorics}, vol. 36, pp. 241–248.
	arXiv: math/0511704.

\bibitem{ENGnozaki2013lower}
	Nozaki, H. 2013.
	“Lower bounds for the minimum diameter of integral point sets”.
	\emph{Australasian Journal of Combinatorics}, vol. 56, pp. 139—143.

\bibitem{ENGsmurov1998stripcoverings}
	Smurov, M. \& Spivak, A. 1998.
	“Pokrytiya poloskami” (``Covering by strips'').
	\emph{Kvant}, vol. 5, pp. 6--12

\end{engbibliography}

### 2019-09-20 AvdMomZvol_rails/Article.tex
From Zvolinsky
Final remarks

@@
 
+Avdeev N.N., Zvolinsky R.E., Momot E.A.
+
+On particular diameter bounds for integral point sets in higher dimensions
+
 \section{Introduction}
@@
 
-TODO: rephrase?
-
-TODO: can a scissors set have another axis of symmetry?
+%TODO: rephrase?
+
+%TODO: can a scissors set have another axis of symmetry?
 
@@
 
-TODO: pictures!
+%TODO: pictures!
 
@@
 
+\item
+$\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
+(\pm 207 , 0),
+(\pm 145 , 0),
+(\pm 63 , 0),
+(\pm 25 , 0),
+(-297 , 0)\}
+$
+(Figure~\ref{picture_4.png}).
+
+$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P})} = 252$,
+
+which gives $d(m, 2m + 5) \leq 252$.
+
 \begin{figure}[htbp]
@@
 $\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
-(\pm 207 , 0),
-(\pm 145 , 0),
-(\pm 63 , 0),
-(\pm 25 , 0),
-(-297 , 0)\}
-$
-(Figure~\ref{picture_4.png}).
-
-$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P})} = 252$,
-
-which gives $d(m, 2m + 5) \leq 252$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{picture_6.png}}
-\parbox{0.7\linewidth}{\caption{IPS of cardinality 13 and diameter 1073}
-\label{picture_6.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
 (\pm 297 , 0),
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{picture_7.png}}
-\parbox{0.7\linewidth}{\caption{IPS of cardinality 14 and diameter 1273}
-\label{picture_7.png}}
+\center{\includegraphics[width=0.7\linewidth]{picture_6.png}}
+\parbox{0.7\linewidth}{\caption{IPS of cardinality 13 and diameter 1073}
+\label{picture_6.png}}
 \end{figure}
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.72\linewidth]{picture_8.png}}
-\parbox{0.73\linewidth}{\caption{IPS of cardinality 15 and diameter 1888}
-\label{picture_8.png}}
-\end{figure}
-
 \item
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.72\linewidth]{picture_9.png}}
-\parbox{0.73\linewidth}{\caption{IPS of cardinality 16 and diameter 2189}
-\label{picture_9.png}}
+\center{\includegraphics[width=0.7\linewidth]{picture_7.png}}
+\parbox{0.7\linewidth}{\caption{IPS of cardinality 14 and diameter 1273}
+\label{picture_7.png}}
+\end{figure}
+
+\begin{figure}[h!]
+\center{\includegraphics[width=0.75\linewidth]{picture_8.png}}
+\parbox{0.75\linewidth}{\caption{IPS of cardinality 15 and diameter 1888}
+\label{picture_8.png}}
 \end{figure}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.9\linewidth]{picture_13.png}}
-\parbox{0.9\linewidth}{\caption{IPS of cardinality 17 and diameter 3224}
-\label{picture_13.png}}
+\center{\includegraphics[width=0.75\linewidth]{picture_9.png}}
+\parbox{0.75\linewidth}{\caption{IPS of cardinality 16 and diameter 2189}
+\label{picture_9.png}}
 \end{figure}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.9\linewidth]{picture_14.png}}
-\parbox{0.9\linewidth}{\caption{IPS of cardinality 18 and diameter 3696}
-\label{picture_14.png}}
+\center{\includegraphics[width=0.85\linewidth]{picture_13.png}}
+\parbox{0.85\linewidth}{\caption{IPS of cardinality 17 and diameter 3224}
+\label{picture_13.png}}
 \end{figure}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_15.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 5141}
-\label{picture_15.png}}
+\center{\includegraphics[width=0.85\linewidth]{picture_14.png}}
+\parbox{0.85\linewidth}{\caption{IPS of cardinality 18 and diameter 3696}
+\label{picture_14.png}}
 \end{figure}
@@
 
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_15.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 5141}
+\label{picture_15.png}}
+\end{figure}
+
 \item
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.9\linewidth]{picture_17.png}}
-\parbox{0.9\linewidth}{\caption{IPS of cardinality 21 and diameter 11692}
-\label{picture_17.png}}
-\end{figure}
-
 \item
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.9\linewidth]{picture_18.png}}
-\parbox{0.9\linewidth}{\caption{IPS of cardinality 22 and diameter 12296}
-\label{picture_18.png}}
+\center{\includegraphics[width=0.75\linewidth]{picture_17.png}}
+\parbox{0.75\linewidth}{\caption{IPS of cardinality 21 and diameter 11692}
+\label{picture_17.png}}
 \end{figure}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.8\linewidth]{picture_19.png}}
-\parbox{0.8\linewidth}{\caption{IPS of cardinality 23 and diameter 16027}
-\label{picture_19.png}}
+\center{\includegraphics[width=0.75\linewidth]{picture_18.png}}
+\parbox{0.75\linewidth}{\caption{IPS of cardinality 22 and diameter 12296}
+\label{picture_18.png}}
 \end{figure}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.8\linewidth]{picture_20.png}}
-\parbox{0.8\linewidth}{\caption{IPS of cardinality 24 and diameter 19758}
-\label{picture_20.png}}
+\center{\includegraphics[width=0.85\linewidth]{picture_19.png}}
+\parbox{0.85\linewidth}{\caption{IPS of cardinality 23 and diameter 16027}
+\label{picture_19.png}}
 \end{figure}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.9\linewidth]{picture_21.png}}
-\parbox{0.9\linewidth}{\caption{IPS of cardinality 25 and diameter 32054}
+\center{\includegraphics[width=0.85\linewidth]{picture_20.png}}
+\parbox{0.85\linewidth}{\caption{IPS of cardinality 24 and diameter 19758}
+\label{picture_20.png}}
+\end{figure}
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_21.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 25 and diameter 32054}
 \label{picture_21.png}}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.95\linewidth]{picture_22.png}}
-\parbox{0.95\linewidth}{\caption{IPS of cardinality 26 and diameter 39516}
+\center{\includegraphics[width=1\linewidth]{picture_22.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 26 and diameter 39516}
 \label{picture_22.png}}
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_24.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 28 and diameter 276612}
-\label{picture_24.png}}
-\end{figure}
-
 \item
@@
 \begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_24.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 28 and diameter 276612}
+\label{picture_24.png}}
+\end{figure}
+
+\begin{figure}[h!]
 \center{\includegraphics[width=1\linewidth]{picture_25.png}}
@@
 
-\begin{itemize}
-\setlength{\itemsep}{-1mm}
-
-\item
-$\mathcal{P}=\sqrt{2}/{1} * \{ (\pm 76, 60),
-(\pm 62 , 0),
-(\pm 29 , 0),
-(\pm 6 , 0)\}
-$
-(Figure~\ref{picture_10.png}).
-
-$f = 152$, $v = 114$, $w = 190$, $\operatorname{diam(\mathcal{P})} = 162$.
-
-We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
-is worse that known one (\hyperlink{d2}{2})).
-
-\begin{figure}[htbp]
+\begin{figure}[h!]
 	\includegraphics[width=.48\linewidth]{picture_10.png}
@@
 \end{figure}
+
+\begin{itemize}
+\setlength{\itemsep}{-1mm}
+
+\item
+$\mathcal{P}=\sqrt{2}/{1} * \{ (\pm 76, 60),
+(\pm 62 , 0),
+(\pm 29 , 0),
+(\pm 6 , 0)\}
+$
+(Figure~\ref{picture_10.png}).
+
+$f = 152$, $v = 114$, $w = 190$, $\operatorname{diam(\mathcal{P})} = 162$.
+
+We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
+is worse that known one (\hyperlink{d2}{2})).
 
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.9\linewidth]{picture_11.pdf}}
-\parbox{0.9\linewidth}{\caption{IPS of cardinality 12 and diameter 409}
+\center{\includegraphics[width=1\linewidth]{picture_11.pdf}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 12 and diameter 409}
 \label{picture_11.pdf}}
@@
 \end{itemize}
+
+\section{Bounds based on pyramid sets}
 
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.9\linewidth]{picture_12.pdf}}
-\parbox{0.9\linewidth}{\caption{IPS of cardinality 10 and diameter 56}
+\center{\includegraphics[width=1\linewidth]{picture_12.pdf}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 10 and diameter 56}
 \label{picture_12.pdf}}
@@
 
+\section{Final remarks}
+All the given PIPSs were obtained through a combination of computer search an intuition of the authors;
+so, the further search may lead to better bounds employing the same consructions.
+
+There is still no general construction for a rails or scissors PIPS of given cardinality.
+For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality with 2 points on one line
+and all the rest on the other;
+on the other hand, we still have not found any rails PIPSs with 4 and 5 points on the lines.
+
+As for today, we have found pyramid PIPS of cardinality at most 9.
+
+The source code can be obtained at https://gitlab.com/Nickkolok/ips-algo
+
 \bibliography{literature}

### 2019-10-13 ZvolMom_AMM2019/ZvolMom_AMM2019.tex
Drafts for AMM2019 - as received from Zvolinsky

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc} 
\usepackage[english, russian]{babel} 
\usepackage{amsmath,amsfonts,amssymb,euscript,graphicx,wrapfig,multirow}
\usepackage{dsfont}
\usepackage{amsthm}
\textheight=240mm \textwidth=170mm
\hoffset=-17mm % сдвиг влево
\voffset=-17mm % сдвиг вверх

\usepackage{hyperref}

\theoremstyle{theorem}
\newtheorem{theorem}{Теорема}
\theoremstyle{dfn}
\newtheorem{dfn}{Определение}
\theoremstyle{remark}
\newtheorem{remark}{Замечание}

\begin{document}

\selectlanguage{russian}

\noindent УДК

\begin{center}
\textbf{О некоторых оценках диаметра целоудалённых множеств точек более высоких размерностей} \\[3mm]
\textbf{Р.~Е.~Зволинский, Е.~А.~Момот} \\[2mm]
\emph{ВГУ}
\end{center}

\section{Введение}

\begin{dfn}\label{dfn1}
	Плоским целоудалённым множеством точек (ЦМ) называют такое 
	множество $\mathcal{P}$ не лежащих на одной прямой точек на 
	плоскости $\mathbb{R}^{2}$, таких, что для любой пары 
	точек $P_{1}, P_{2} \in \mathcal{P}$ 
	расстояние $|P_{1}P_{2}|$ между точками $P_{1}$ и $P_{2}$ есть целое число.
\end{dfn}

Определение \ref{dfn1} можно обобщить для общего случая:

\begin{dfn}
	Целоудалённым множеством $\mathcal{P}$ называют множество из $n$ точек в
	$m$-мерном Евклидовом пространстве $\mathbb{R}^{m}$ с
	попарными  целочисленными  расстояниями между точками данного множества.
\end{dfn}

Как можно охарактеризовать такое целоудалённое множество?
Во-первых, мы можем посмотреть на его размерность,
тогда у нас получится посчитать число точек данного множества, которое всегда конечно [1, 2], далее будем называть это число \textit{мощностью};
наконец, мы можем естественно определить диаметр такого конечного набора точек.

\begin{dfn}
	Диаметр ЦМ $\mathcal{P}$ опеделяется как
	\begin{equation}
		\operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
		\mathcal{P}}{\max} |P_{1}P_{2}|
	\end{equation}
\end{dfn}

У каждого ЦМ также есть характеристика
[3, 4].

\begin{dfn}
	Функция $d(m, n)$ есть минимальный возможный диаметр ЦМ $\mathcal{P}$ из $n$ 
	точек в $m$-мерном Евклиловом пространстве $\mathbb{R}^{m}$.
\end{dfn}

Спискок известных оценок функции $d(m, n)$
читатель может найти в [5, Theorem 1] или в [6].
Мы обсудим следующие оценки, представленные в [5]:
\begin{align}
	&d(m, 2m + 1) \leq 8\\
	&d(m, 2m + 2) \leq 13 \hypertarget{d2}\\
	&d(m, 3m) \leq 109
\end{align}
и следующую теорему [5, Theorem 2.1].

\begin{theorem}
	Пусть $\mathcal{P}$ это плоское целоудалённое множество точек, 
	где $n - 2$ точки  расположены на прямой $l_{1}$, а две точки 
	$P_{1}$ и  $P_{2}$ на параллельной ей прямой $l_{2}$ с расстоянием 
	$r$ между $l_{1}$ и $l_{2}$. Если существуют такие положительные целые числа 
	$v$, $w$, удовлетворяющие условиям $f^{2} + v^{2}
	= w^{2}$ и $v < 2r$, где $|P_{1}P_{2}| = f$,
	то

	\begin{equation}\label{formula1}
		d(m, n - 2 + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
	\end{equation}

\end{theorem}

В данной статье мы представим некоторые оценки функции $d(m,n)$, основанные на плоских ЦМ определённых типов.

\section{ЦМ, расположенные на двух прямых}

\begin{dfn}
    Плоские целоудалённые множества точек, расположенных на двух параллельных прямых, будем называть рельсовыми системами.
\end{dfn}

Среди рельсовых систем наиболее распространены ЦМ с двумя точками на одной прямой и остальными на другой.

\begin{dfn}
    Плоские целоудалённые множества точек, расположенных на двух прямых, 
    не являющихся ни параллельными, ни перпендикулярными, будем называть системами
    типа ножницы.
\end{dfn}

Существует важный подкласс наборов ножниц.

\begin{dfn}
    Системы типа ножницы с осью симметрии, 
    являющейся биссектрисой угла между прямыми, назовём пирамидальными.
\end{dfn}

\section{Оценки, основанные на рельсовых системах}

Пусть $i = \overline{1, k}$ обозначает перечисление всех $i$
от $1$ до $k$.
Теорему [5, Theorem 2.1] можно обобщить для общего случая
следующим образом:

\begin{theorem}
	Пусть $\mathcal{P}$ это плоское целоудалённое множество точек, где 
	$n - k$ точек расположены на прямой $l_1$, а $k$ точек $P_1$, $P_2$, $\ldots$,
	$P_k$ на параллельной ей прямой $l_2$ с расстоянием $r$ между $l_1$ и $l_2$. 
	Если существуют такие положительные целые числа $v$, $w_{ij}$, удовлетворяющие
	условиям $f_{ij}^{2} + v^{2} = w_{ij}^{2}$ и $v < 2r$, где 
	$i = \overline{1, k - 1}$, $j = \overline{i + 1, k}$, $|P_{i}P_{j}| = f_{ij}$,
	то

	\begin{equation}
		d(m, n - k + k(m - k)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
	\end{equation}

\end{theorem}

Ниже мы приводим оценки функции $d(m, n)$ для $n = 2m + k$, $3 \leq k \leq 24$, а также для $n = 4m$, основанные на рельсовых ЦМ (подробное описание данных ЦМ и конструкций для построения оценок можно найти в [7]).
\begin{align*}
&(d(m, 2m + k))_{k = 3, \ldots, 24} \leq 158,\, 158,\, 252,\, 297,\, 1073,\, 1273,\, 1888,\, 2189,\, 3224,\, 3696,\, 5141,\, 6586,\, \\
&11692,\, 12296,\, 16027,\, 19758,\, 32054,\, 39516,\, 254364,\, 276612,\, 338169,\, 399726
\end{align*}

Ричард Гай в своей работе [8, D 20] приводит ЦМ из $8$ точек, расположенных на двух параллельных прямых. На основании данной ЦМ там удалось получить оценку
\begin{equation}\label{result1}
d(m, 4m) \leq 409
\end{equation}

Также с помощью мощностей ЭВМ нам удалось получить рельсовую ЦМ мощности 40 (на данный момент это наибольшая по мощности рельсовая ЦМ, которую нам удалось найти).
Ниже приведен иллюстрированный пример данной ЦМ.

Для облегчения восприятия применяется запись вида 
[9, 10, 6]:
$\sqrt{p}/q * \{ (x_1,y_1), \ldots,$ $ (x_n, y_n)  \}$,
означающая, что каждую абсциссу нужно умножить на $1/q$,
а каждую ординату~--- умножить на $\sqrt{p}/q$, т.е.
$$
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
$$

\begin{figure}[h!]
	\begin{center}
	\includegraphics[width=1.1\linewidth]{picture_40.png}
	\parbox{1\linewidth}{\caption{ЦМ мощности 40, диаметр 107526848}
	\label{picture_40.png}}
	\end{center}
\end{figure}

\begin{itemize}
\setlength{\itemsep}{-1mm}

\item
$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 53763424, 60),
(\pm 51964380 , 0),
(\pm 35959560 , 0),
(\pm 30175080 , 0),
(\pm 22378720 , 0),
$

$
(\pm 20180160 , 0),
(\pm 15613780 , 0),
(\pm 15025920 , 0),
(\pm 14779380 , 0),
(\pm 11313120 , 0),
(\pm 9817080 , 0),
$

$
(\pm 9544080 , 655200),
(\pm 9271080 , 0),
(\pm 6693960 , 0),
(\pm 6289920 , 0),
(\pm 4006275 , 0),
(\pm 3290560 , 0),
$

$
(\pm 1938720 , 0),
(\pm 1092000 , 0),
(\pm 583648 , 0)\}
$
(Figure~\ref{picture_40.png}).

$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam(\mathcal{P})}
= 107526848$,

что даёт оценку $d(m, 2m + 34) \leq 107526848$.

\end{itemize}

\section{Оценки, основанные на пирамидальных системах}

\begin{theorem}
Пусть $\mathcal{P}$ это плоское целоудалённое множество точек, где
$k$ точек расположены на прямой $l_{1}$ и $k$ точек на прямой $l_{2}$.
Кроме того, эти точки симметричны относительно одной из биссектрис углов пересечения 
прямых $l_{1}$ и $l_{2}$, тогда

\begin{equation}\label{formula2}
d(m, (k - \alpha)m + \alpha) \leq \operatorname{diam(\mathcal{P})},
\end{equation}

где
\begin{equation*}
\alpha =
\begin{cases}
1, \text{если точка пересечения} \in \mathcal{P}, \\
0, \text{если точка пересечения} \notin \mathcal{P}.
\end{cases}
\end{equation*}

\end{theorem}

\begin{remark}
Углы могут быть острыми или тупыми. В [7] описано, почему угол пересечения прямых $l_{1}$ и $l_{2}$ не может быть равен ${\pi}/2$.
\end{remark}

Ниже мы приводим оценки функции $d(m, n)$ для $n = 3m + 1$ и $n = 4m + 1$, основанные на пирамидальных ЦМ (подробное описание данных ЦМ и конструкций для построения оценок можно найти в [7]).

\begin{align}
&d(m, 3m + 1) \leq 56 \label{result2} \\
&d(m, 4m + 1) \leq 1260 \label{result3}
\end{align}

Получившаяся оценка для функции $d(m, 3m + 1)$ улучшает оценку для $d(m, 3m)$, представленную в [3].

\section{Заключительное замечание}

Все данные ЦМ были получены путём сочетания копьютерного поиска и интуиции авторов;
таким образом, дальнейший поиск может привести к лучшим оценкам, использующим ту же
конструкцию.

До сих нет общей конструкции для рельсов или ножниц заданной мощности.

Для рельсовых ЦМ мы можем предположить, что существует множество произвольной мощности с $2$ точками на одной линии, а все остальные на другой; с другой стороны,
мы до сих пор не нашли ни одной рельсовой ЦМ с $4$ и $5$ точками на отделившихся прямых.

На сегодняшний день мы нашли пирамидальную ЦМ мощности не более $9$.

Исходный код можно получить по адресу https://gitlab.com/Nickkolok/ips-algo

\bigskip\centerline{\bf Литература}

1.	\emph{ Anning N.~H., Erd{\"o}s P.} Integral distances // Bulletin of the American Mathematical Society. -- 1945. -- Vol. 51, no. 8. -- Pp. 598-600. -- DOI: 10.1090/S0002-9904-1945-08407-9.

2.	\emph{ Erd{\"o}s P.} Integral distances // Bulletin of the American Mathematical Society. -- 1945. -- Vol. 51, no. 12. -- P. 996. -- DOI: 10.1090/S0002-9904-1945-08490-0.

3.  \emph{ Kemnitz A.} Punktmengen mit ganzzahligen Abst{\"a}nden. -- 1988.

4.  \emph{ Kurz S.} On the characteristic of integral point sets in $\mathbb{E}^m$ // Australasian Journal of Combinatorics. -- 2006. -- Vol.~36. -- Pp. 241-248. --
arXiv: math/0511704.

5.  \emph{ Kurz S., Laue R.} Bounds for the minimum diameter of integral point sets
// Australasian Journal of Combinatorics. -- 2007. -- Vol.~39. -- Pp. 233-240. --
arXiv: 0804.1296.

6.  \emph{ Авдеев Н.~Н., Семёнов Е.~М.} Множества точек с целочисленными расстояниями на плоскости и в евклидовом пространстве // Математический форум 
(Итоги науки. Юг России). -- 2018. -- Pp. 217-236.

7.  \emph{ Avdeev N.~N., Zvolinsky R.~E., Momot E.~A.} On particular diameter bounds
for integral point sets in higher dimensions. -- 2019. -- arXiv: 1909.10386v1

8.  \emph{ Guy R.} Unsolved problems in number theory. Vol. 1. -- Springer Science \& Business Media, 2013. -- DOI: 10.1007/978-1-4757-1738-9.

9.  \emph{ Авдеев Н.~Н.} On integral point sets in special position // Некоторые вопросы анализа, алгебры, геометрии и математического образования: материалы международной молодежной научной школы <<Актуальные направления математического анализа и смежные вопросы>>. -- 2018. -- Vol.~8. -- Pp. 5-6.

10.  \emph{ Авдеев Н.~Н.} Об отыскании целоудалённых множеств специального вида
// Актуальные проблемы прикладной математики, информатики и механики -- 
сборник трудов Международной научной конференции. -- Научно-исследовательские публикации, 2018. -- Pp. 492-498.

\end{document}

### 2019-10-14 ZvolMom_AMM2019/ZvolMom_AMM2019.tex
Subway iteration

@@
 
-\noindent УДК
+\noindent УДК~519.146
 
 \begin{center}
-\textbf{О некоторых оценках диаметра целоудалённых множеств точек более высоких размерностей} \\[3mm]
+\textbf{О некоторых оценках диаметра целоудалённых множеств точек в многомерных пространствах} \\[3mm]
 \textbf{Р.~Е.~Зволинский, Е.~А.~Момот} \\[2mm]
@@
 
-\begin{dfn}\label{dfn1}
-	Плоским целоудалённым множеством точек (ЦМ) называют такое 
-	множество $\mathcal{P}$ не лежащих на одной прямой точек на 
-	плоскости $\mathbb{R}^{2}$, таких, что для любой пары 
-	точек $P_{1}, P_{2} \in \mathcal{P}$ 
-	расстояние $|P_{1}P_{2}|$ между точками $P_{1}$ и $P_{2}$ есть целое число.
+Настоящая работа имеет целью кратко изложить (на русском языке) содержание работы [7], дополнив его некоторыми результатами
+вычислений и подробностями используемых алгоритмов.
+
+\begin{dfn}
+	Целоудалённым множеством $\mathcal{P}$ называют множество точек в
+	$m$-мерном евклидовом пространстве $\mathbb{R}^{m}$, такое, что расстояние
+	между любыми двумя точками этого множества есть целое число.
 \end{dfn}
 
-Определение \ref{dfn1} можно обобщить для общего случая:
+Если $m = 2$, то такое ЦМ будем называть \textit{плоским}.
+
+Как можно охарактеризовать целоудалённое множество?
+Во-первых, мы можем подсчитать количество точек в нём,
+которое всегда конечно [1, 2], далее будем называть это число \textit{мощностью};
+во-вторых, мы можем естественным образом определить диаметр конечного множества точек.
 
 \begin{dfn}
-	Целоудалённым множеством $\mathcal{P}$ называют множество из $n$ точек в
-	$m$-мерном Евклидовом пространстве $\mathbb{R}^{m}$ с
-	попарными  целочисленными  расстояниями между точками данного множества.
-\end{dfn}
-
-Как можно охарактеризовать такое целоудалённое множество?
-Во-первых, мы можем посмотреть на его размерность,
-тогда у нас получится посчитать число точек данного множества, которое всегда конечно [1, 2], далее будем называть это число \textit{мощностью};
-наконец, мы можем естественно определить диаметр такого конечного набора точек.
-
-\begin{dfn}
-	Диаметр ЦМ $\mathcal{P}$ опеделяется как
+	Диаметр конечного множества $\mathcal{P}$ определяется как
 	\begin{equation}
@@
 \begin{dfn}
-	Функция $d(m, n)$ есть минимальный возможный диаметр ЦМ $\mathcal{P}$ из $n$ 
-	точек в $m$-мерном Евклиловом пространстве $\mathbb{R}^{m}$.
+	Функция $d(m, n)$ есть минимальный возможный диаметр ЦМ $\mathcal{P}$ мощности $n$
+	в $m$-мерном евклидовом пространстве $\mathbb{R}^{m}$.
 \end{dfn}
 
-Спискок известных оценок функции $d(m, n)$
+Список известных оценок функции $d(m, n)$
 читатель может найти в [5, Theorem 1] или в [6].
+Кроме того, определённый прогресс был достигнут в работе [*]. %Nazaki 2018
 Мы обсудим следующие оценки, представленные в [5]:
@@
 \begin{theorem}
-	Пусть $\mathcal{P}$ это плоское целоудалённое множество точек, 
-	где $n - 2$ точки  расположены на прямой $l_{1}$, а две точки 
-	$P_{1}$ и  $P_{2}$ на параллельной ей прямой $l_{2}$ с расстоянием 
-	$r$ между $l_{1}$ и $l_{2}$. Если существуют такие положительные целые числа 
+	Пусть $\mathcal{P}$ -- плоское целоудалённое множество точек,
+	где $(n - 2)$ точки  расположены на прямой $l_{1}$, а две точки
+	$P_{1}$ и  $P_{2}$ на параллельной ей прямой $l_{2}$, при этом расстояние
+	между $l_{1}$ и $l_{2}$ равно $r$. Если существуют положительные целые числа
 	$v$, $w$, удовлетворяющие условиям $f^{2} + v^{2}
@@
 		d(m, n - 2 + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
+		.
 	\end{equation}
@@
 
-\section{ЦМ, расположенные на двух прямых}
+\section{Классификация ЦМ, расположенных на двух прямых}
 
-\begin{dfn}
-    Плоские целоудалённые множества точек, расположенных на двух параллельных прямых, будем называть рельсовыми системами.
-\end{dfn}
+Терминология, сложившаяся при класссификации ЦМ под влиянием С. Курца, весьма своеобразна.
+Он использовал такие наименования, как
+<<crab>> и даже <<semi-crab>> [**]. %TODO:Antonov, Kurz
+При переводе этой терминологии мы будем использовать прилагательные, следуя [***]. %TODO:обзорник
+
+\par
+\textbf{Определение 4.}
+Плоские целоудалённые множества мощности $n$, для которых можно найти такую прямую $l_1$, что
+$(n - 1)$ точек лежит на $l_1$, будем называть \textit{веерными}.
+
+\par
+\textbf{Определение 5.}
+Плоские целоудалённые множества мощности $n$, для которых можно единственным образом найти такие прямые $l_1$, $l_2$,
+что $k$ точек лежат на прямой $l_1$, а $(n - k)$ -- на прямой $l_2$, причём $k \geq 2$, будем называть \textit{рельсовыми}.
 
@@
 
-\begin{dfn}
-    Плоские целоудалённые множества точек, расположенных на двух прямых, 
-    не являющихся ни параллельными, ни перпендикулярными, будем называть системами
-    типа ножницы.
-\end{dfn}
+\par
+\textbf{Определение 6.}
+\textit{Ножничными} будем называть такие плоские целоудалённые множества мощности $n$, для которых можно единственным
+образом найти такие прямые $l_1$, $l_2$, что $k$ точек лежат на прямой $l_1$, а $(n - k)$ -- на прямой $l_2$, причём $k \geq 2$ и
+прямые $l_1$ и $l_2$ не являются ни параллельными, ни перпендикулярными.
 
-Существует важный подкласс наборов ножниц.
+Существует важный подкласс ножничных систем.
 
-\begin{dfn}
-    Системы типа ножницы с осью симметрии, 
-    являющейся биссектрисой угла между прямыми, назовём пирамидальными.
-\end{dfn}
+\par
+\textbf{Определение 7.}
+Ножничные системы с осью симметрии,
+являющейся биссектрисой угла между прямыми, назовём \textit{пирамидальными}.
+
+\par
+\textbf{Определение 8.}
+\textit{Крестовыми} будем называть такие плоские целоудалённые множества мощности $n$, для которых можно единственным
+образом найти такие прямые $l_1$, $l_2$, что $k$ точек лежат на прямой $l_1$, а $(n - k)$ -- на прямой $l_2$, причём $k \geq 2$ и
+$l_1$ перпендикулярна $l_2$.
 
@@
 от $1$ до $k$.
-Теорему [5, Theorem 2.1] можно обобщить для общего случая
+Теорему [5, Theorem 2.1] можно обобщить
 следующим образом:
@@
 \begin{theorem}
-	Пусть $\mathcal{P}$ это плоское целоудалённое множество точек, где 
-	$n - k$ точек расположены на прямой $l_1$, а $k$ точек $P_1$, $P_2$, $\ldots$,
-	$P_k$ на параллельной ей прямой $l_2$ с расстоянием $r$ между $l_1$ и $l_2$. 
-	Если существуют такие положительные целые числа $v$, $w_{ij}$, удовлетворяющие
+	Пусть $\mathcal{P}$ -- плоское целоудалённое множество точек, где
+	$k$ точек $P_1$, $P_2$, $\ldots$, $P_k$ последовательно расположены на прямой $l_1$,
+	а $(n - k)$ точек -- на параллельной ей прямой $l_2$, при этом расстояние между $l_1$ и $l_2$ равно $r$.
+	Если существуют положительные целые числа $v$, $w_{ij}$, удовлетворяющие
 	условиям $f_{ij}^{2} + v^{2} = w_{ij}^{2}$ и $v < 2r$, где
@@
 		d(m, n - k + k(m - k)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
+		.
 	\end{equation}
@@
 
-Ричард Гай в своей работе [8, D 20] приводит ЦМ из $8$ точек, расположенных на двух параллельных прямых. На основании данной ЦМ там удалось получить оценку
+Р. Гай в своей работе [8, D 20] приводит ЦМ из $8$ точек, расположенное на двух параллельных прямых. С помощью теоремы 2 можно построить следующую оценку:
 \begin{equation}\label{result1}
@@
 
-Также с помощью мощностей ЭВМ нам удалось получить рельсовую ЦМ мощности 40 (на данный момент это наибольшая по мощности рельсовая ЦМ, которую нам удалось найти).
+Также с помощью ЭВМ нам удалось получить рельсовую ЦМ мощности 40 (на данный момент это наибольшая по мощности рельсовая ЦМ, которую нам удалось найти).
 Ниже приведен иллюстрированный пример данной ЦМ.
@@
 Для облегчения восприятия применяется запись вида
-[9, 10, 6]:
+[6, 9, 10]:
 $\sqrt{p}/q * \{ (x_1,y_1), \ldots,$ $ (x_n, y_n)  \}$,
 означающая, что каждую абсциссу нужно умножить на $1/q$,
-а каждую ординату~--- умножить на $\sqrt{p}/q$, т.е.
+а каждую ординату -- на $\sqrt{p}/q$, т.е.
 $$
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 53763424, 60),
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 53763424, 0),
 (\pm 51964380 , 0),
@@
 $
-(Figure~\ref{picture_40.png}).
+(Рис.~\ref{picture_40.png}).
 
@@
 \begin{theorem}
-Пусть $\mathcal{P}$ это плоское целоудалённое множество точек, где
+Пусть $\mathcal{P}$ -- плоское целоудалённое множество точек, где
 $k$ точек расположены на прямой $l_{1}$ и $k$ точек на прямой $l_{2}$.
@@
 \begin{remark}
-Углы могут быть острыми или тупыми. В [7] описано, почему угол пересечения прямых $l_{1}$ и $l_{2}$ не может быть равен ${\pi}/2$.
+Углы могут быть острыми или тупыми, но, очевидно, не прямыми [7, Remark 2].
 \end{remark}
@@
 
-\section{Заключительное замечание}
+\section{Особенности алгоритмов}
+Используется алгоритм, описанный в [****], со следующими модификациями:
+во-первых, изначальный граф разбивается на несколько подграфов в зависимости от характеристик.
+Треугольники, образованные вершинами каждого подграфа, имеют одинаковую характеристику, т.е. каждому подграфу сооветствует своя
+характеристика.
+Во-вторых, мы стараемся как можно раньше исключать точки, входящие только в веерные системы, поскольку веерные системы
+нам не интересны.
+В-третьих, используется длинная целая арифметика и новый тип BigInt языка Javascript.
 
-Все данные ЦМ были получены путём сочетания копьютерного поиска и интуиции авторов;
+За представление ЦМ отвечает класс PointSet, следующий представлению (TODO:cсылка),
+что позволяет оперировать только с целыми числами.
+В дальнейшем операции проводятся лишь над точками, имеющими совместное представление (TODO:cсылка).
+
+Основной классификационный инструмент -- это функция, находящая все точки заданного множества, не лежащие на прямой,
+проходящей через две заданные точки.
+Реализованы и другие необходимые иструменты, такие как проверка прямых на параллельность в условиях представления каждой прямой парой точек
+с координатами вида (TODO: туда же),
+а также прочие проверки, позволяющие классифицировать ЦМ в терминах [TODO:arXiv(7)]
+(например, отыскивать центр окружности по заданным трём точкам в соответствующем представлении через целые числа).
+Большая часть функций покрыта юнит-тестами с использованием (TODO: списать из отчёта).
+Следуя идеологии NodeJS/NPM, часть переиспользуемого кода выделена в библиотеки (TODO: списать, какие, + ссылки как на электронные ресурсы).
+Визуализация найденных множеств производится с помощью gnuplot [TODO:ссылку!].
+
+\section{Заключительные замечания}
+
+Все данные ЦМ были получены путём сочетания компьютерного перебора и интуиции авторов;
 таким образом, дальнейший поиск может привести к лучшим оценкам, использующим ту же
@@
 
-До сих нет общей конструкции для рельсов или ножниц заданной мощности.
+До сих нет общей конструкции для рельсовых или ножничных систем заданной мощности.
 
 Для рельсовых ЦМ мы можем предположить, что существует множество произвольной мощности с $2$ точками на одной линии, а все остальные на другой; с другой стороны,
-мы до сих пор не нашли ни одной рельсовой ЦМ с $4$ и $5$ точками на отделившихся прямых.
+мы до сих пор не нашли ни одной рельсовой ЦМ с $4$ и $5$ точками на <<отделившихся>> прямых.
 
-На сегодняшний день мы нашли пирамидальную ЦМ мощности не более $9$.
+На сегодняшний день мы нашли пирамидальные ЦМ мощности не более $9$.
 
-Исходный код можно получить по адресу https://gitlab.com/Nickkolok/ips-algo
+Исходный код можно получить по адресам https://gitlab.com/Nickkolok/ips-algo,
+
+https://gitlab.com/Aisse-258/ips-algo
 
@@
 
+{\bf Зволинский Роман Евгеньевич},
+студент кафедры теории функций математического факультета ФГБОУ ВО <<Воронежский государственный университет>>, г. Воронеж.
+
+E-mail: roman-zvolinskiy@mail.ru
+
+Научный руководитель: Семенов Евгений Михайлович.
+
+{\bf Момот Екатерина Александровна},
+студент кафедры функционального анализа и операторных уравнений математического факультета
+ФГБОУ ВО <<Воронежский государственный университет>>, г. Воронеж.
+
 \end{document}

### 2019-10-14 ZvolMom_AMM2019/ZvolMom_AMM2019.tex
iteration from Zvolinsky

@@
 \usepackage{amsthm}
+\usepackage{textcomp}
 \textheight=240mm \textwidth=170mm
@@
 \usepackage{hyperref}
-
-\theoremstyle{theorem}
-\newtheorem{theorem}{Теорема}
-\theoremstyle{dfn}
-\newtheorem{dfn}{Определение}
-\theoremstyle{remark}
-\newtheorem{remark}{Замечание}
 
@@
 
-Настоящая работа имеет целью кратко изложить (на русском языке) содержание работы [7], дополнив его некоторыми результатами
+Настоящая работа имеет целью кратко изложить (на русском языке) содержание работы [1], дополнив его некоторыми результатами
 вычислений и подробностями используемых алгоритмов.
 
-\begin{dfn}
+\par
+\textbf{Определение 1.}
 Целоудалённым множеством $\mathcal{P}$ называют множество точек в
@@
 между любыми двумя точками этого множества есть целое число.
-\end{dfn}
 
@@
 Во-первых, мы можем подсчитать количество точек в нём,
-которое всегда конечно [1, 2], далее будем называть это число \textit{мощностью};
+которое всегда конечно [2, 3], далее будем называть это число \textit{мощностью};
 во-вторых, мы можем естественным образом определить диаметр конечного множества точек.
 
-\begin{dfn}
+\par
+\textbf{Определение 2.}
 Диаметр конечного множества $\mathcal{P}$ определяется как
@@
 \operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
-		\mathcal{P}}{\max} |P_{1}P_{2}|
+\mathcal{P}}{\max} |P_{1}P_{2}|,
 \end{equation}
-\end{dfn}
+где $|P_1 P_2|$ есть обычное расстояние между точками.
 
 У каждого ЦМ также есть характеристика
-[3, 4].
+[4, 5].
 
-\begin{dfn}
+\par
+\textbf{Определение 3.}
 Функция $d(m, n)$ есть минимальный возможный диаметр ЦМ $\mathcal{P}$ мощности $n$
 в $m$-мерном евклидовом пространстве $\mathbb{R}^{m}$.
-\end{dfn}
 
 Список известных оценок функции $d(m, n)$
-читатель может найти в [5, Theorem 1] или в [6].
-Кроме того, определённый прогресс был достигнут в работе [*]. %Nazaki 2018
-Мы обсудим следующие оценки, представленные в [5]:
+читатель может найти в [6, Theorem 1] или в [7].
+Кроме того, определённый прогресс был достигнут в работе [8]. %Nozaki 2018
+Мы обсудим следующие оценки, представленные в [6]:
 \begin{align}
-	&d(m, 2m + 1) \leq 8\\
-	&d(m, 2m + 2) \leq 13 \hypertarget{d2}\\
+	&d(m, 2m + 1) \leq 8 \\
+	&d(m, 2m + 2) \leq 13 \\
 	&d(m, 3m) \leq 109
 \end{align}
-и следующую теорему [5, Theorem 2.1].
+и следующую теорему [6, Theorem 2.1].
 
-\begin{theorem}
+\par
+\textbf{Теорема 1.}
 Пусть $\mathcal{P}$ -- плоское целоудалённое множество точек,
@@
 
-	\begin{equation}\label{formula1}
+\begin{equation}
 d(m, n - 2 + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
@@
 \end{equation}
-
-\end{theorem}
 
@@
 
-Терминология, сложившаяся при класссификации ЦМ под влиянием С. Курца, весьма своеобразна.
-Он использовал такие наименования, как
-<<crab>> и даже <<semi-crab>> [**]. %TODO:Antonov, Kurz
-При переводе этой терминологии мы будем использовать прилагательные, следуя [***]. %TODO:обзорник
+Терминология, сложившаяся при классификации ЦМ под влиянием С. Курца, весьма своеобразна. Он использовал такие наименования, как
+<<crab>> и даже <<semi-crab>> [9]. %TODO:Antonov, Kurz
+При переводе этой терминологии мы будем использовать прилагательные, следуя [7]. %TODO:обзорник
 
@@
 от $1$ до $k$.
-Теорему [5, Theorem 2.1] можно обобщить
+Теорему [6, Theorem 2.1] можно обобщить
 следующим образом:
 
-\begin{theorem}
+\par
+\textbf{Теорема 2.}
 Пусть $\mathcal{P}$ -- плоское целоудалённое множество точек, где
-	$k$ точек $P_1$, $P_2$, $\ldots$, $P_k$ последовательно расположены на прямой $l_1$,
-	а $(n - k)$ точек -- на параллельной ей прямой $l_2$, при этом расстояние между $l_1$ и $l_2$ равно $r$.
+$k$ точек $P_1$, $P_2$, $\ldots$, $P_k$ последовательно расположены на
+прямой $l_1$, а $(n - k)$ точек -- на параллельной ей прямой $l_2$,
+при этом расстояние между $l_1$ и $l_2$ равно $r$.
 Если существуют положительные целые числа $v$, $w_{ij}$, удовлетворяющие
@@
 \begin{equation}
-		d(m, n - k + k(m - k)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
+d(m, n + k(m - 2)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
 .
@@
 
-\end{theorem}
-
-Ниже мы приводим оценки функции $d(m, n)$ для $n = 2m + k$, $3 \leq k \leq 24$, а также для $n = 4m$, основанные на рельсовых ЦМ (подробное описание данных ЦМ и конструкций для построения оценок можно найти в [7]).
+Ниже мы приводим оценки функции $d(m, n)$ для $n = 2m + k$, $3 \leq k \leq 24$, а также для $n = 4m$, основанные на рельсовых ЦМ (подробное описание данных ЦМ и конструкций для построения оценок можно найти в [1]).
 \begin{align*}
@@
 
-Р. Гай в своей работе [8, D 20] приводит ЦМ из $8$ точек, расположенное на двух параллельных прямых. С помощью теоремы 2 можно построить следующую оценку:
-\begin{equation}\label{result1}
+Р. Гай в своей работе [10, D 20] приводит ЦМ из $8$ точек, расположенное на двух параллельных прямых. С помощью теоремы 2 можно построить следующую оценку:
+\begin{equation}
 d(m, 4m) \leq 409
@@
 Также с помощью ЭВМ нам удалось получить рельсовую ЦМ мощности 40 (на данный момент это наибольшая по мощности рельсовая ЦМ, которую нам удалось найти).
-Ниже приведен иллюстрированный пример данной ЦМ.
 
 Для облегчения восприятия применяется запись вида
-[6, 9, 10]:
+[7, 11, 12]:
 $\sqrt{p}/q * \{ (x_1,y_1), \ldots,$ $ (x_n, y_n)  \}$,
@@
 а каждую ординату -- на $\sqrt{p}/q$, т.е.
-$$
+\begin{align} \label{notation}
 	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
@@
 	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
-$$
+\end{align}
 
@@
 
-\begin{theorem}
+\par
+\textbf{Теорема 3.}
 Пусть $\mathcal{P}$ -- плоское целоудалённое множество точек, где
 $k$ точек расположены на прямой $l_{1}$ и $k$ точек на прямой $l_{2}$.
-Кроме того, эти точки симметричны относительно одной из биссектрис углов пересечения
+Пусть, кроме того, эти точки симметричны относительно одной из биссектрис углов пересечения
 прямых $l_{1}$ и $l_{2}$, тогда
 
-\begin{equation}\label{formula2}
+\begin{equation}
 d(m, (k - \alpha)m + \alpha) \leq \operatorname{diam(\mathcal{P})},
@@
 
-\end{theorem}
+\par
+\textbf{Замечание 1.}
+Углы могут быть острыми или тупыми, но, очевидно, не прямыми [1, Remark 2].
 
-\begin{remark}
-Углы могут быть острыми или тупыми, но, очевидно, не прямыми [7, Remark 2].
-\end{remark}
-
-Ниже мы приводим оценки функции $d(m, n)$ для $n = 3m + 1$ и $n = 4m + 1$, основанные на пирамидальных ЦМ (подробное описание данных ЦМ и конструкций для построения оценок можно найти в [7]).
+Ниже мы приводим оценки функции $d(m, n)$ для $n = 3m + 1$ и $n = 4m + 1$, основанные на пирамидальных ЦМ (подробное описание данных ЦМ и конструкций для построения оценок можно найти в [1]).
 
 \begin{align}
-&d(m, 3m + 1) \leq 56 \label{result2} \\
-&d(m, 4m + 1) \leq 1260 \label{result3}
+&d(m, 3m + 1) \leq 56 \\
+&d(m, 4m + 1) \leq 1260
 \end{align}
 
-Получившаяся оценка для функции $d(m, 3m + 1)$ улучшает оценку для $d(m, 3m)$, представленную в [3].
+Получившаяся оценка для функции $d(m, 3m + 1)$ улучшает оценку для $d(m, 3m)$, представленную в [4].
 
 \section{Особенности алгоритмов}
-Используется алгоритм, описанный в [****], со следующими модификациями:
-во-первых, изначальный граф разбивается на несколько подграфов в зависимости от характеристик.
-Треугольники, образованные вершинами каждого подграфа, имеют одинаковую характеристику, т.е. каждому подграфу сооветствует своя
+Используется алгоритм, описанный в [13], со следующими модификациями:
+
+Во-первых, изначальный граф разбивается на несколько подграфов в зависимости от характеристик.
+Треугольники, образованные вершинами каждого подграфа, имеют одинаковую характеристику, т.е.
+каждому подграфу сооветствует своя
 характеристика.
+
 Во-вторых, мы стараемся как можно раньше исключать точки, входящие только в веерные системы, поскольку веерные системы
 нам не интересны.
+
 В-третьих, используется длинная целая арифметика и новый тип BigInt языка Javascript.
 
-За представление ЦМ отвечает класс PointSet, следующий представлению (TODO:cсылка),
-что позволяет оперировать только с целыми числами.
-В дальнейшем операции проводятся лишь над точками, имеющими совместное представление (TODO:cсылка).
+За представление ЦМ отвечает класс PointSet, следующий представлению (\ref{notation}), что позволяет оперировать только с целыми числами.
+В дальнейшем операции проводятся лишь над точками, имеющими совместное представление (\ref{notation}).
 
-Основной классификационный инструмент -- это функция, находящая все точки заданного множества, не лежащие на прямой,
-проходящей через две заданные точки.
+Основной классификационный инструмент -- это функция, находящая все точки заданного множества, не лежащие на прямой, проходящей через две заданные точки.
+
 Реализованы и другие необходимые иструменты, такие как проверка прямых на параллельность в условиях представления каждой прямой парой точек
-с координатами вида (TODO: туда же),
-а также прочие проверки, позволяющие классифицировать ЦМ в терминах [TODO:arXiv(7)]
+с координатами вида (\ref{notation}), а также прочие проверки, позволяющие классифицировать ЦМ в терминах [9]
 (например, отыскивать центр окружности по заданным трём точкам в соответствующем представлении через целые числа).
-Большая часть функций покрыта юнит-тестами с использованием (TODO: списать из отчёта).
-Следуя идеологии NodeJS/NPM, часть переиспользуемого кода выделена в библиотеки (TODO: списать, какие, + ссылки как на электронные ресурсы).
-Визуализация найденных множеств производится с помощью gnuplot [TODO:ссылку!].
+Большая часть функций покрыта юнит-тестами с использованием библиотек ‘chai’ и
+<<мокко>>.
+%(TODO: списать из отчёта).
+
+Следуя идеологии NodeJS/NPM, часть переиспользуемого кода выделена в библиотеки
+‘json-bigint-native’ (см. https://www.npmjs.com/package/bigint-json-native), ‘bigint-isqrt’ (см. https://www.npmjs.com/package/bigint-isqrt)
+и ‘bigint-is-perfect-square’ (см.
+
+\noindent https://www.npmjs.com/package/bigint-is-perfect-square).
+ %(TODO: списать, какие, + ссылки как на электронные ресурсы).
+
+Визуализация найденных множеств производится с помощью gnuplot [14].
 
@@
 
-https://gitlab.com/Aisse-258/ips-algo
+https://gitlab.com/Aisse-258/ips-algo.
 
@@
 
-1.	\emph{ Anning N.~H., Erd{\"o}s P.} Integral distances // Bulletin of the American Mathematical Society. -- 1945. -- Vol. 51, no. 8. -- Pp. 598-600. -- DOI: 10.1090/S0002-9904-1945-08407-9.
+1.  \emph{ Avdeev N.~N., Zvolinsky R.~E., Momot E.~A.} On particular diameter bounds
+for integral point sets in higher dimensions. -- 2019. -- arXiv: 1909.10386v1
 
-2.	\emph{ Erd{\"o}s P.} Integral distances // Bulletin of the American Mathematical Society. -- 1945. -- Vol. 51, no. 12. -- P. 996. -- DOI: 10.1090/S0002-9904-1945-08490-0.
+2.	\emph{ Anning N.~H., Erd{\"o}s P.} Integral distances // Bulletin of the American Mathematical Society. -- 1945. -- Vol. 51, no. 8. -- Pp. 598-600. -- DOI: 10.1090/S0002-9904-1945-08407-9.
 
-3.  \emph{ Kemnitz A.} Punktmengen mit ganzzahligen Abst{\"a}nden. -- 1988.
+3.	\emph{ Erd{\"o}s P.} Integral distances // Bulletin of the American Mathematical Society. -- 1945. -- Vol. 51, no. 12. -- P. 996. -- DOI: 10.1090/S0002-9904-1945-08490-0.
 
-4.  \emph{ Kurz S.} On the characteristic of integral point sets in $\mathbb{E}^m$ // Australasian Journal of Combinatorics. -- 2006. -- Vol.~36. -- Pp. 241-248. --
+4.  \emph{ Kemnitz A.} Punktmengen mit ganzzahligen Abst{\"a}nden. -- 1988.
+
+5.  \emph{ Kurz S.} On the characteristic of integral point sets in $\mathbb{E}^m$ // Australasian Journal of Combinatorics. -- 2006. -- Vol.~36. -- Pp. 241-248. --
 arXiv: math/0511704.
 
-5.  \emph{ Kurz S., Laue R.} Bounds for the minimum diameter of integral point sets
+6.  \emph{ Kurz S., Laue R.} Bounds for the minimum diameter of integral point sets
 // Australasian Journal of Combinatorics. -- 2007. -- Vol.~39. -- Pp. 233-240. --
@@
 
-6.  \emph{ Авдеев Н.~Н., Семёнов Е.~М.} Множества точек с целочисленными расстояниями на плоскости и в евклидовом пространстве // Математический форум
+7.  \emph{ Авдеев Н.~Н., Семёнов Е.~М.} Множества точек с целочисленными расстояниями на плоскости и в евклидовом пространстве // Математический форум
 (Итоги науки. Юг России). -- 2018. -- Pp. 217-236.
 
-7.  \emph{ Avdeev N.~N., Zvolinsky R.~E., Momot E.~A.} On particular diameter bounds
-for integral point sets in higher dimensions. -- 2019. -- arXiv: 1909.10386v1
+8.  \emph{ Nozaki H.} Lower bounds for the minimum diameter of integral point sets
+// Australasian Journal of Combinatorics. -- 2013. -- Т.~56. -- С. 139-143.
 
-8.  \emph{ Guy R.} Unsolved problems in number theory. Vol. 1. -- Springer Science \& Business Media, 2013. -- DOI: 10.1007/978-1-4757-1738-9.
+9.  \emph{ Antonov A.~R., Kurz S.} Maximal integral point sets over
+$\mathbb{Z}^{2}$ // International Journal of Computer Mathematics. -- 2008. --
+Vol.~87, no.~12 -- Pp. 2653-2676. -- DOI:
 
-9.
+\noindent 10.1080/00207160902993636. -- arXiv: 0804.1280.
+
+10.  \emph{ Guy R.} Unsolved problems in number theory. Vol. 1. -- Springer Science \& Business Media, 2013. -- DOI: 10.1007/978-1-4757-1738-9.
+
+11.
 \emph{ Авдеев Н.~Н.} On integral point sets in special position // Некоторые вопросы анализа, алгебры, геометрии и математического образования: материалы международной молодежной научной школы <<Актуальные направления математического анализа и смежные вопросы>>.
@@
 
-10.  \emph{ Авдеев Н.~Н.} Об отыскании целоудалённых множеств специального вида
+12.  \emph{ Авдеев Н.~Н.} Об отыскании целоудалённых множеств специального вида
 // Актуальные проблемы прикладной математики, информатики и механики --
@@
 
+13.  \emph{ Авдеев Н.~Н.} Об отыскании множеств точек на плоскости с целочисленными
+расстояниями // Математическое и компьютерное моделирование, информационные технологии управления: сб. тр. Школы для студентов, аспирантов и молодых учёных
+<<МКМИТУ-2016>>. -- 2016. -- С. 15-19.
+
+14.  \emph{ Балдин Е.} Графики заказывали? // Системный администратор. -- 2007. --
+\textnumero.~4. -- С. 72-77.
+
 {\bf Зволинский Роман Евгеньевич},
-студент кафедры теории функций математического факультета ФГБОУ ВО <<Воронежский государственный университет>>, г. Воронеж.
+студент кафедры теории функций и геометрии математического факультета ФГБОУ ВО <<Воронежский государственный университет>>, г. Воронеж.
 

### 2019-10-15 ZvolMom_AMM2019/ZvolMom_AMM2019.tex
Minor fixes by Avdeev

@@
 \section{Особенности алгоритмов}
-Используется алгоритм, описанный в [13], со следующими модификациями:
+Используется алгоритм, описанный в [13], со следующими модификациями.
 
 Во-первых, изначальный граф разбивается на несколько подграфов в зависимости от характеристик.
-Треугольники, образованные вершинами каждого подграфа, имеют одинаковую характеристику, т.е.
-каждому подграфу сооветствует своя
-характеристика.
+Треугольники, образованные вершинами каждого подграфа, имеют одинаковую характеристику,
+т.е. каждому подграфу сооветствует своя характеристика.
 
 Во-вторых, мы стараемся как можно раньше исключать точки, входящие только в веерные системы, поскольку веерные системы
-нам не интересны.
+нам неинтересны.
 
@@
 
-До сих нет общей конструкции для рельсовых или ножничных систем заданной мощности.
+До сих пор нет общей конструкции для рельсовых или ножничных систем заданной мощности.
 
-Для рельсовых ЦМ мы можем предположить, что существует множество произвольной мощности с $2$ точками на одной линии, а все остальные на другой; с другой стороны,
-мы до сих пор не нашли ни одной рельсовой ЦМ с $4$ и $5$ точками на <<отделившихся>> прямых.
+Для рельсовых ЦМ мы можем предположить,
+что существует множество произвольной мощности с $2$ точками на одной линии и остальными на другой;
+с другой стороны,
+мы до сих пор не нашли ни одной рельсовой ЦМ с $4$ и $5$ точками на прямых соответственно.
 

### 2019-10-21 AvdMomZvol_rails/Article.tex
Last iteration

@@
 \inputencoding{utf8}
-\bibliographystyle{unsrt}
+%\bibliographystyle{unsrt}
 \textheight=240mm \textwidth=170mm
@@
 \voffset=-17mm
+
+\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+\addbibresource{../common/notmy.bib}
+\addbibresource{../common/my.bib}
 
@@
 
-\begin{center}
+\title{
+	On particular diameter bounds for integral point sets in higher dimensions
+	%\footnote{
+	%	This work was carried out at Voronezh State University and supported by the Russian Science
+	%	Foundation grant 19-11-00197.
+	%}
+}
+
+%% Прекрасно понимаю, что следующая команда - дичь и вордописчество, но время не ждёт, время жмёт
+\author{
+	Avdeev N.N.
+	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
+	, Zvolinsky R.E., Momot E.A.
+	\\
+	\\
 	\emph{Voronezh State University}
-\end{center}
-
-Avdeev N.N., Zvolinsky R.E., Momot E.A.
-
-On particular diameter bounds for integral point sets in higher dimensions
+}
+
+\maketitle
 
@@
 	Let $\mathcal{P}$ be a planar integral point set consisting of
-	$n - 2$ points on line $l$ and two points $P_{1}$ and $P_{2}$ on a
-	parallel line $m$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
+	$n - 2$ points on line $l_1$ and two points $P_{1}$ and $P_{2}$ on a
+	parallel line $l_2$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
 	exist positive integers $v$, $w$ with $f^{2} + v^{2}
@@
 
-\bibliography{literature}
+\printbibliography
+%\bibliography{literature}
 

### 2019-11-01 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
ChebSb bibliography - all but DOIs
Self-proofreading iteration

--- 2019-09-11 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
@@
 \usepackage{textcase}% Оформление страниц
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% {{ biblatex
-%% Команда \Cite (с большой буквы) у biblatex своя.
-%% Сохраняем "нашу"
-\let\CiteChebSb=\Cite
-%% Убираем \Cite
-\let\Cite=\relax
-%% Вот теперь смело загружаем biblatex
-\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
-\addbibresource{../common/notmy.bib}
-\addbibresource{../common/my.bib}
-%% Восстанавливаем "нашу" \Cite
-\let\Cite=\CiteChebSb
-%% }} biblatex
 
@@
 на диаметр плоских множеств с целочисленными расстояниями.
-Полученная вданной статье оценка всё ещё не точна.
+Полученная в данной статье оценка всё ещё не точна.
 }
@@
 A planar integral point set is called to be in semi-general position, if it does not contain collinear triples.
-The existing lower bound for mininum diameter of planar integral point sets is linear on its cardinality.
+The existing lower bound for mininum diameter of a planar integral point set is linear with respect to its cardinality.
 There were no known special diameter bounds for planar integral point sets in semi-general position
@@
 We prove a new lower bound for mininum diameter of planar integral point sets in semi-general position
-that is better than linear (polynomial with power $5/4$).
-The proof is based on several lemmas observations, including the ones provided by Solymosi
+that is better than linear (polynomial of power $5/4$).
+The proof is based on several lemmas and observations, including the ones provided by Solymosi
 while proving the first linear lower bound for diameter of a planar integral point set.
@@
 There are some bounds for minimal diameter of planar integral point sets in some special positions.
-Assuming that the planar integral point sets contains many collinear points,
+Assuming that a planar integral point set contains many collinear points,
 the following result holds.
@@
 in~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved},~\cite{our-pmm-2018},~\cite{our-ped-2018}.
-For generalizaton in higher dimensions and the appropriate bounds, see~\cite{kurz2005characteristic,nozaki2013lower}.
+For generalization to higher dimensions and the appropriate bounds, see~\cite{kurz2005characteristic,nozaki2013lower}.
 
@@
 	\cite[Observation 1]{solymosi2003note}
-	If a triangle $T$ has integer side-lengths $a \leq b \leq c$,
+	If a triangle $T$ has integer side lengths $a \leq b \leq c$,
 	then the minimal height $m$ of it is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
@@
 	If a triange $T$ with minimal height $\rho$ is situated in a strip,
-	then the width of a strip is at least $\rho$.
+	then the width of the strip is at least $\rho$.
 \end{lemma}
@@
 	\label{cor:solymosi_strip}
-	If a triangle $T$ with integer side-lengths $a \leq b \leq c$ is situated in a strip,
-	then the width of a strip is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
+	If a triangle $T$ with integer side lengths $a \leq b \leq c$ is situated in a strip,
+	then the width of the strip is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
 \end{Corollary}
@@
 		n \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| \leq  4 p^{4/5}
-		,
+		.
 	\end{equation}
-	or, that is the same,
+	This means that
 	\begin{equation}
@@
 The presented bound is the first special lower bound for sets in semi-general position.
-Thus, we did not accepted the challenge to make the constant in Theorem~\ref{thm:main_result} as large as possible,
+Thus, we have not accepted the challenge to make the constant in Theorem~\ref{thm:main_result} as large as possible,
 in order to keep the ideas of the proof clear and understandable.
@@
 
-\printbibliography
+%библиография по ГОСТу
 
-%библиография по ГОСТу
-\begin{thebibliography}{99}
+% Мне библиографию в требуемом формате сгенерировал biblatex:
+% https://github.com/odomanov/biblatex-gost/issues/20
 
-	\bibitem{1}
-	Banks~W.\,D., Conflitti~A., Shparlinski~I.\,E. Character sums over integers with restricted $g$-ary digits // Illinois J. Math. 2002. Vol. 46, №3. P. 819-836.
-
-\end{thebibliography}
+\input{blg-GOST}
 
 %библиография по Гарвардскому стандарту
-\begin{engbibliography}{99}
 
-	\bibitem{en1}
-	Banks, W.\,D., Conflitti, A. \& Shparlinski, I.\,E. 2002, ``Character sums over integers with restricted $g$-ary digits``, \textit{Illinois J. Math.}, vol. 46, no. 3, pp. 819-836.
-
-\end{engbibliography}
+\input{blg-Harvard}
 

### 2019-11-02 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
Author thanks  Dr. Usachev for remarks on English

@@
 {%Аннотация статьи на английском языке
-A point set $M$ in the Euclidean plane is called a planar integral point set if all the distances between the
+A point set $M$ in the Euclidean plane is said to be a planar integral point set if all the distances between the
 elements of $M$ are integers, and $M$ is not situated on a straight line.
-A planar integral point set is called to be in semi-general position, if it does not contain collinear triples.
-The existing lower bound for mininum diameter of a planar integral point set is linear with respect to its cardinality.
-There were no known special diameter bounds for planar integral point sets in semi-general position
+A planar integral point set is said to be a set in semi-general position, if it does not contain collinear triples.
+The existing lower bound for mininal diameter of planar integral point sets is linear with respect to its cardinality.
+There were no known special diameter bounds for planar integral point sets in semi-general position of given cardinality
 (the known upper bound for planar integral point sets is constructive
-and provides planar integral point sets in semi-general position).
-We prove a new lower bound for mininum diameter of planar integral point sets in semi-general position
+and employs planar integral point sets in semi-general position).
+We prove a new lower bound for minimal diameter of planar integral point sets in semi-general position
 that is better than linear (polynomial of power $5/4$).
-The proof is based on several lemmas and observations, including the ones provided by Solymosi
-while proving the first linear lower bound for diameter of a planar integral point set.
-We have to admit that our bound is still not tight.
+The proof is based on several lemmas and observations, including the ones established by Solymosi
+to prove the first linear lower bound for diameter of a planar integral point set.
+We have to admit that our bound is still not exact.
 }
@@
 Since every integral point set can obviously be dilated to a set of larger diameter,
-minimal possible diameters of sets of given cardinality are in the focus.
+minimal possible diameters of sets of given cardinality are of interest.
 To be precise,
-the following function was introduced~\cite{kurz2008bounds,kurz2008minimum}:
+the following function was introduced in~\cite{kurz2008bounds,kurz2008minimum}:
 \begin{equation}
@@
 and even for 4 points out of the line~\cite{huff1948diophantine}.
-For $9\leq n\leq 122$, the minimal possible diameter is achieved at a facher set~\cite{kurz2008bounds}.
+For $9\leq n\leq 122$, the minimal possible diameter is attained at a facher set~\cite{kurz2008bounds}.
 
@@
 Furthermore, the constructions of integral point sets in semi-general position of arbitrary cardinality
-appeared~\cite{harborth1993upper};
+appeared in~\cite{harborth1993upper};
 such sets are situated on a circle.
@@
 \begin{Definition}
-	A set $M\in\overline{\mathfrak{M}}(2,n)$ is called to be in \textit{general position},
+	A set $M\in\overline{\mathfrak{M}}(2,n)$ is said to be a set in \textit{general position},
 	if no four points of $M$ are concyclic.
@@
 
-It remains unknown if there are integral points sets in general position of arbitrary cardinality;
+It remains unknown whether there are integral points sets in general position of arbitrary cardinality;
 however, some sets $M\in \dot{\mathfrak{M}}(2,7)$ are known~\cite{kreisel2008heptagon,kurz2013constructing}.
@@
 \end{equation*}
-The upper bound is presented in~\cite{harborth1993upper}.
-The lower bound was firstly introduced in~\cite{solymosi2003note};
+The upper bound is established in~\cite{harborth1993upper}.
+The lower bound firstly appeared in~\cite{solymosi2003note};
 the largest known value for $c_1$ is $5/11$ for $n\geq 4$~\cite{my-pps-linear-bound-2019}.
@@
 In the present paper we give a special bound for planar integral point sets in semi-general position.
-The condition of semi-general position is important in the given proof.
+The condition of semi-general position is crucial in the given proof.
 
@@
 \begin{Definition}
-	The part of a plane between two parallel straight lines with distance $\rho$ between the lines
+	The part of a plane between two parallel straight lines with distance $\rho$ between them
 	is called a strip of width $\rho$.
@@
 	Let $\{M_1, M_2, M_3, M_4\} \subset M\in\overline{\mathfrak{M}}(2,n)$
-	(points $M_2$ and $M_3$ may coincide, other points may not), $n\geq 4$.
+	(points $M_2$ and $M_3$ may coincide, other points are distinct), $n\geq 4$.
 	Then $\# M \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
@@
 
-\begin{proof}
+\begin{proof}[Proof]
 	Each point $N\in M$ satisfies one of the following conditions:
@@
 	\end{multline}
-	we are done.
+	the assertion follows.
 \end{proof}
@@
 
-\begin{proof}
+\begin{proof}[Proof]
 	For $n = 3$ we have $\overline{d}(2,3) = 1$.
@@
 	Let us choose points $M_1, M_2, M_3, M_4 \in M$
-	(points $M_2$ and $M_3$ may coincide, other points may not), such that
+	(points $M_2$ and $M_3$ may coincide, other points are distinct), such that
 	\begin{equation}
@@
 	\end{equation}
-	This means that
+	Equivalently,
 	\begin{equation}
@@
 \begin{Remark}
-	The following result in known:
+	The following result appeared in~\cite{solymosi2003note} as Corollary 1:
 \end{Remark}
@@
 \begin{lemma}
-	\cite[Corollary 1]{solymosi2003note}
-	For $H \in \overline {\mathfrak{M}}(2,n)$, the minimum
+	For $H \in \overline {\mathfrak{M}}(2,n)$, the minimal
 	distance in H is at least $n^{1/3}$.
@@
 in order to keep the ideas of the proof clear and understandable.
-A more thorough research can be done in the future to enlarge the constant.
-However, the upper and lower bounds are still not tight.
+The more thorough research can be done in the future to improve the constant.
+However, the upper and lower bounds are still not exact.
 
 \section{Acknowledgements}
-Author thanks Dr. Prof. E.M. Semenov for proofreading and valuable advice.
+Author thanks Dr. Prof. E.M. Semenov for proofreading and valuable advices and Dr. A.S. Usachev for proofreading.
 

### 2019-11-02 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
We should not mention the author due to referring policy

@@
 \levkolonttl{%левый колонтитул - авторы
-N.~N.~Avdeev}
+~}%N.~N.~Avdeev}
 \prvkolonttl{%правый колонтитул - сокращенное название статьи
@@
 Об оценках на диаметр плоских множеств с целочисленными расстояниями полуобщего положения%
-\footnote{Исследование выполнено за счет гранта Российского научного фонда (проект 19-11-00197).}}
+~}%\footnote{Исследование выполнено за счет гранта Российского научного фонда (проект 19-11-00197).}}
 {%название статьи на английском языке
@@
 {%авторы статьи с указанием города на русском языке
-Н. Н. Авдеев (г. Воронеж)}
+~}%Н. Н. Авдеев (г. Воронеж)}
 {%авторы статьи с указанием города на английском языке
-N. N. Avdeev (Voronezh)}
+~}%N. N. Avdeev (Voronezh)}
 
 \Cite{
-Н. Н. Авдеев Об оценках на диаметр плоских множеств с целочисленными расстояниями полуобщего положения // Чебышевcкий сборник, 2019, т.~\tom, вып.~\iss, с.~\pageref{beg}--\pageref{end}.
+%Н. Н. Авдеев
+Об оценках на диаметр плоских множеств с целочисленными расстояниями полуобщего положения // Чебышевcкий сборник, 2019, т.~\tom, вып.~\iss, с.~\pageref{beg}--\pageref{end}.
 }
-{N. N. Avdeev, 2019, "On diameter bounds for planar integral point sets in semi-general position"\,, {\it Che\-by\-shev\-skii sbornik}, vol.~\tom, no.~\iss, pp.~\pageref{beg}--\pageref{end}.
+{
+%N. N. Avdeev,
+2019, "On diameter bounds for planar integral point sets in semi-general position"\,, {\it Che\-by\-shev\-skii sbornik}, vol.~\tom, no.~\iss, pp.~\pageref{beg}--\pageref{end}.
 }
@@
 {%авторы статьи с указанием аффилиации на русском языке
-\noindent {\bf Н. Н. Авдеев}~--- Воронежский государственный университет, кафедра теории функций и геометрии
-
+%\noindent {\bf Н. Н. Авдеев}~--- Воронежский государственный университет, кафедра теории функций и геометрии
+deleted
 \noindent
-\emph{e-mail: nickkolok@mail.ru, avdeev@math.vsu.ru}
+%\emph{e-mail: nickkolok@mail.ru, avdeev@math.vsu.ru}
 
@@
 {%авторы статьи с указанием аффилиации на английском языке
-\noindent {\bf N. N. Avdeev}~--- Voronezh State University
-
+%\noindent {\bf N. N. Avdeev}~--- Voronezh State University
+deleted
 \noindent
-\emph{e-mail: nickkolok@mail.ru, avdeev@math.vsu.ru}
+%\emph{e-mail: nickkolok@mail.ru, avdeev@math.vsu.ru}
 

### 2019-11-02 semi-general-suplinear-bound/blg-GOST.tex
We should not mention the author due to referring policy

\begin{thebibliography}{99}
\bibitem{anning1945integral}
Anning N. H., Erdös P. Integral distances // Bulletin of the American Mathematical
Society. — 1945. — т. 51, No 8. — с. 598—600.
\bibitem{erdos1945integral}
Erdös P. Integral distances // Bulletin of the American Mathematical Society. — 1945. —
т. 51, No 12. — с. 996.
\bibitem{our-vmmsh-2018}
Author et al., 2018
%Авдеев Н. Н., Семёнов Е. М. Множества точек с целочисленными расстояниями на
%плоскости и в евклидовом пространстве // Математический форум (Итоги науки. Юг
%России). — 2018. — с. 217—236.
\bibitem{kurz2008bounds}
Kurz S., Laue R. Bounds for the minimum diameter of integral point sets // Australasian
Journal of Combinatorics. — 2007. — т. 39. — с. 233—240. — arXiv: 0804.1296.
\bibitem{kurz2008minimum}
Kurz S., Wassermann A. On the minimum diameter of plane integral point sets // Ars
Combinatoria. — 2011. — т. 101. — с. 265—287. — arXiv: 0804.1307.
\bibitem{antonov2008maximal}
Antonov A. R., Kurz S. Maximal integral point sets over $\mathbb{Z}^2$ // International Journal of
Computer Mathematics. — 2008. — т. 87, No 12. — с. 2653—2676. — arXiv: 0804.1280.
\bibitem{huff1948diophantine}
Huff G. B. Diophantine problems in geometry and elliptic ternary forms // Duke
Mathematical Journal. — 1948. — т. 15, No 2. — с. 443—453.
\bibitem{harborth1993upper}
Harborth H., Kemnitz A., Möller M. An upper bound for the minimum diameter of integral
point sets // Discrete \& Computational Geometry. — 1993. — т. 9, No 4. — с. 427—432.
\bibitem{piepmeyer1996maximum}
Piepmeyer L. The maximum number of odd integral distances between points in the plane //
Discrete \& Computational Geometry. — 1996. — т. 16, No 1. — с. 113—115.
\bibitem{kreisel2008heptagon}
Kreisel T., Kurz S. There are integral heptagons, no three points on a line, no four on a
circle // Discrete \& Computational Geometry. — 2008. — т. 39, No 4. — с. 786—790.
\bibitem{kurz2013constructing}
Constructing 7-clusters / S. Kurz [и др.] // Serdica Journal of Computing. — 2014. — т. 8,
No 1. — с. 47—70. — arXiv: 1312.2318.
\bibitem{solymosi2003note}
Solymosi J. Note on integral distances // Discrete \& Computational Geometry. — 2003. —
т. 30, No 2. — с. 337—342.
\bibitem{my-pps-linear-bound-2019}
Author, 2019
%Avdeev N. On existence of integral point sets and their diameter bounds. — 2019. — arXiv:
%1906.11926.
\bibitem{bat2018number}
Bat-Ochir G. On the number of points with pairwise integral distances on a circle // Discrete
Applied Mathematics. — 2018. — т. 254. — с. 17—32.
\bibitem{brass2006research}
Brass P., Moser W. O., Pach J. Research problems in discrete geometry. — Springer Science
\& Business Media, 2006.
\bibitem{guy2013unsolved}
Guy R. Unsolved problems in number theory. т. 1. — Springer Science \& Business Media,
2013.
\bibitem{our-pmm-2018}
Author, 2018
%Авдеев Н. Н. Об отыскании целоудалённых множеств специального вида // Актуальные
%проблемы прикладной математики, информатики и механики - сборник трудов Международной
%научной конференции. — Научно-исследовательские публикации, 2018. — с. 492—498.
\bibitem{our-ped-2018}
Author, 2018
%Авдеев Н. Н. On integral point sets in special position // Некоторые вопросы анализа,
%алгебры, геометрии и математического образования: материалы международной молодежной
%научной школы «Актуальные направления математического анализа и смежные вопросы». —
%2018. — т. 8. — с. 5—6.
\bibitem{kurz2005characteristic}
Kurz S. On the characteristic of integral point sets in $E^m$ // Australasian Journal of
Combinatorics. — 2006. — т. 36. — с. 241. — arXiv: math/0511704.
\bibitem{nozaki2013lower}
Nozaki H. Lower bounds for the minimum diameter of integral point sets // Australasian
Journal of Combinatorics. — 2013. — т. 56. — с. 139—143.
\bibitem{smurov1998stripcoverings}
Смуров М., Спивак А. Покрытия полосками // Квант. — 1998. — No 5. — с. 6.
\end{thebibliography}

### 2019-11-02 semi-general-suplinear-bound/blg-Harvard.tex
We should not mention the author due to referring policy

@@
 	\emph{Bulletin of the American Mathematical Society}, vol. 51.8, pp. 598—600.
+	doi: 10.1090/S0002-9904-1945-08407-9
 
@@
 	\emph{Bulletin of the American Mathematical Society}, vol. 51.12, p. 996.
+	doi: 10.1090/S0002-9904-1945-08490-0
 
 \bibitem{ENGour-vmmsh-2018}
-	Avdeev, N.N. \& Semenov, E.M. 2018.
-	``Mnozhestva tochek c tselochislennymi rasstoyaniyami na ploskosti i v evklidovom prostranstve''
-	(``Integral point sets on the plane and in Euclidean space'')
-	\emph{Matematicheskiy forum (Itogi nauki. Yug Rossii)}, pp. 217—236.
+	Author et al., 2018
+	%Avdeev, N.N. \& Semenov, E.M. 2018.
+	%``Mnozhestva tochek c tselochislennymi rasstoyaniyami na ploskosti i v evklidovom prostranstve''
+	%(``Integral point sets on the plane and in Euclidean space'')
+	%\emph{Matematicheskiy forum (Itogi nauki. Yug Rossii)}, pp. 217—236.
+	% No DOI
 
@@
 	\emph{International Journal of Computer Mathematics}, vol. 87.12, pp. 2653—2676. arXiv: 0804.1280.
+	doi: 10.1080/00207160902993636
 
@@
 	\emph{Duke Mathematical Journal}, vol. 15.2, pp. 443—453.
+	doi:10.1215/S0012-7094-48-01543-9
 
@@
 	\emph{Discrete \& Computational Geometry} vol. 9.4, pp. 427—432.
+	doi: 10.1007/bf02189331
 
@@
 	\emph{Discrete \& Computational Geometry}, vol. 16.1, pp. 113—115.
+	doi: 10.1007/bf02711135
 
@@
 	\emph{Discrete \& Computational Geometry}, vol. 39.4, pp. 786—790.
+	doi: 10.1007/s00454-007-9038-6
 
@@
 	\emph{Discrete \& Computational Geometry}, vol. 30.2, pp. 337—342.
+	doi: 10.1007/s00454-003-0014-7
 
 \bibitem{ENGmy-pps-linear-bound-2019}
-	Avdeev, N. 2019.
-	“On existence of integral point sets and their diameter bounds”.
-	arXiv: 1906.11926
+	Author, 2019
+	%Avdeev, N. 2019.
+	%“On existence of integral point sets and their diameter bounds”.
+	%arXiv: 1906.11926
 
@@
 	\emph{Discrete Applied Mathematics}, vol. 254, pp. 17—32.
+	doi: 10.1016/j.dam.2018.07.004
 
@@
 	\emph{Research problems in discrete geometry}. Springer Science \& Business Media.
+	doi: 10.1007/0-387-29929-7
 
@@
 	Springer Science \& Business Media.
+	doi: 10.1007/978-1-4757-1738-9
 
 \bibitem{ENGour-pmm-2018}
-	Avdeev, N.N. 2018.
-	“Ob otyskanii tseloudalennykh mnozhestv spetsial'nogo vida”
-	(``On the search of integral point sets of a special type'').
-	\emph{
-		Aktual'nye problemy prikladnoj matematiki, informatiki i mekhaniki
-		- sbornik trudov Mezhdunarodnoj nauchnoj konferencii.
-	}
-	(Actual problems of applied mathematics, informatics and mechanics
-	- proc. of the int. conf.).
-	Voronezh, pp. 492—498.
+	Author, 2018
+	%Avdeev, N.N. 2018.
+	%“Ob otyskanii tseloudalennykh mnozhestv spetsial'nogo vida”
+	%(``On the search of integral point sets of a special type'').
+	%\emph{
+	%	Aktual'nye problemy prikladnoj matematiki, informatiki i mekhaniki
+	%	- sbornik trudov Mezhdunarodnoj nauchnoj konferencii.
+	%}
+	%(Actual problems of applied mathematics, informatics and mechanics
+	%- proc. of the int. conf.).
+	%Voronezh, pp. 492—498.
 
 \bibitem{ENGour-ped-2018}
-	Avdeev, N.N. 2018.
-	“On integral point sets in special position”.
-	\emph{Nekotorye voprosy analiza, algebry, geometrii i matematicheskogo obrazovaniya}
-	(Some problems of analysis, algebra, geometry and mathematical education), vol. 8, pp. 5—6.
+	Author, 2018
+	%Avdeev, N.N. 2018.
+	%“On integral point sets in special position”.
+	%\emph{Nekotorye voprosy analiza, algebry, geometrii i matematicheskogo obrazovaniya}
+	%(Some problems of analysis, algebra, geometry and mathematical education), vol. 8, pp. 5—6.
 

### 2019-11-02 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
Final fixes

@@
 A planar integral point set is said to be a set in semi-general position, if it does not contain collinear triples.
-The existing lower bound for mininal diameter of planar integral point sets is linear with respect to its cardinality.
+The existing lower bound for mininal diameter of a planar integral point set is linear with respect to its cardinality.
 There were no known special diameter bounds for planar integral point sets in semi-general position of given cardinality
@@
 	\cite{smurov1998stripcoverings}
-	If a triange $T$ with minimal height $\rho$ is situated in a strip,
+	If a triangle $T$ with minimal height $\rho$ is situated in a strip,
 	then the width of the strip is at least $\rho$.

### 2019-11-05 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Report 1, remark 2
I preferred the first suggested option, i.e. replacing c_j -> 2c_J
[J220] Report 2, starting the major revision
The referree has advices to try to better guide the reader through the long list of technical auxiliary results.
So, all the auxiliary results concerning the intersection of an integral point set with a line segment
have been separated into a new section.

@@
 
+\section{Collinear subsets of integral point sets}
+Integral point sets with many collinear ponts are very dominating examples.
+In this section we estimate the cardinality of an intersection of an integral point set with a line segment.
+
+Assuming that the planar integral point set contains many collinear points,
+the following result holds.
+\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
+	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
+	at least $n^\delta$ collinear points there exists a $n_0 (\varepsilon)$
+	such that for all $n \geq n_0 (\varepsilon)$ we have
+	\begin{equation}
+		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
+		.
+	\end{equation}
+\end{theorem}
+However, the estimate is rather unsuitable for our needs,
+as it does not provide the values of all the constants.
+To obtain the needed estimate, we now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.
+
+\begin{definition}
+	For a line segment $M_1 M_2$ and an integer $k$, such that $-|M_1 M_2| < k < |M_1 M_2|$,
+	we define a $\rho(k,M_1 M_2)$-curve as the set of points $N$
+	for which the equality $|N M_1| - |N M_2| = k$ holds.
+\end{definition}
+So, in the planar case a $\rho(k,M_1 M_2)$-curve
+is a branch of a hyperbola for $k\neq 0$ and the perpendicular bisector of line segment $M_1 M_2$ for $k=0$.
+
+\begin{proposition}
+	\label{obs:rho_curves}
+	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
+	the equality $|M_1 M_2| = |M_3 M_4|$ holds and line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
+	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and the $\rho(k,M_3 M_4)$-curve
+	do not intersect.
+\end{proposition}
+
+\begin{lemma}
+	\label{lem:2k-1_segments}
+	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
+	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$ with endpoints in $M$,
+	such that $|\Delta_i| = k$.
+\end{lemma}
+\begin{proof}
+	Consider a point $N\in M \setminus m$.
+	Then for each $\Delta_i$ there is a $\rho(n_i,\Delta_i)$-curve containing $N$.
+	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
+	otherwise the $\rho(n_i,\Delta_i)$-curve and the $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
+	There can be only $2k-1$ distinct values for $n_i$,
+	so there are at most $2k-1$ distinct segments $\Delta_i$.
+\end{proof}
+
+Setting $k=1$ in Lemma~\ref{lem:2k-1_segments},
+we obtain the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018}).
+\begin{corollary}
+	\label{cor:only_one_distance_1_on_straight_line}
+	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
+	Then there is at most one pair of points $M_1,M_2\in M \cap m$
+	such that $|M_1 M_2| = 1$.
+\end{corollary}
+
+\begin{lemma}
+	\label{lem:line_segment_with n_squared_plus_one_points}
+	Let $\Delta$ be a straight line segment, $|\Delta|=l$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
+	Let $\#(\Delta \cap M) = n^2 + 1$.
+	Then
+	\begin{equation}
+		l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
+		.
+	\end{equation}
+\end{lemma}
+
+\begin{proof}
+	Any $n^2+1$ points, including the endpoints of $\Delta$, partition the segment $\Delta$ into $n^2$
+	sequential segments $\Delta_i$.
+	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
+	at most three segments of length 2, etc.
+	The following two expressions for sums conclude the proof:
+	\begin{equation}
+		1 + \sum_{k=1}^n (2k-1) = n^2 + 1
+		,
+	\end{equation}
+	\begin{equation}
+		\sum_{k=1}^n k(2k-1) = \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
+		.
+	\end{equation}
+\end{proof}
+
+Now we will estimate the length of a line segment that intersects an integral point set
+by an arbitrary number of points.
+
+\begin{lemma}
+	\label{lem:line_segment_cardinality}
+	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
+	Let $\#(\Delta \cap M) = t$.
+	Then
+	\begin{equation}
+		\label{eq:estimate_for_segment_length}
+		b\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}
+		.
+	\end{equation}
+\end{lemma}
+
+\begin{proof}
+	Let $f(k)$ denote the mininal length of a line segment
+	that intersects an integral point set by $k$ points.
+	We observe that $f(k) > f (k-1)$.
+	Due to Lemma~\ref{lem:line_segment_with n_squared_plus_one_points},
+	$f(n^2+1) \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$.
+
+	%Consider $t$ such that $n^2 + 1 \leq t \leq (n+1)^2$. %,
+	%then $f(t) > f(n^2 +1)$.% and $n \geq \sqrt{t} - 1$.
+	For $t\in\mathbb{N}$ the inequality $(\sqrt{t} - 1)^2 +1 \leq t$ holds,
+	thus
+	%and derive that
+	\begin{equation}
+		\label{eq:line_segment_length}
+		f(t) \geq f((\sqrt{t} - 1)^2 +1) \geq \frac{2}{3}(\sqrt{t} - 1)^3+\frac{1}{2}(\sqrt{t} - 1)^2-\frac{1}{6}(\sqrt{t} - 1)
+		=
+		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2}
+		.
+	\end{equation}
+\end{proof}
+
+In~\cite{kurz2008bounds}, thorough investigation of planar integral point sets with diameter at most 10 000
+is carried out;
+thus, we can restrict our attention to intersections of planar integral point sets with lne segments of length more than 10 000.
+
+Lemma~\ref{lem:line_segment_cardinality} leads to the following proposition.
+
+\begin{proposition}
+	\label{obs:estimate_points_on_straight_line}
+	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
+	Let $\#(\Delta \cap M) = k$ and $b>10000$.
+	Then $k \leq \gamma_2 b + 6$,
+	where
+	\begin{equation}
+		\gamma_2 = \frac{3846}{2593 \sqrt{647}-5823} = 0.063958...
+	\end{equation}
+\end{proposition}
+
+\begin{proof}
+	We know the maximum number of points for planar integral point sets of diameters at most $10000$.
+	Thus, we are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
+	holds for $b > 10000$.
+	So, let us consider $t\geq 647$ and find a coefficient $\gamma_2$,
+	such that the inequality
+	\begin{equation}
+		\label{eq:inequality_for_linear_estimate_of_segment_length}
+		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2} \geq \frac{t-6}{\gamma_2}
+	\end{equation}
+	holds for all $t\geq 647$.
+	For such $t$,
+	the left-hand side of~\eqref{eq:inequality_for_linear_estimate_of_segment_length} obviously grows faster than the right-hand side.
+	Turning~\eqref{eq:inequality_for_linear_estimate_of_segment_length} into the same equality
+	and solving it for $t=647$, we obtain the required estimate.
+\end{proof}
+
+Concluding the section, we remark that
+Corollary~\ref{cor:only_one_distance_1_on_straight_line}
+and
+Proposition~\ref{obs:estimate_points_on_straight_line}
+are the only results of this section
+that are used in the proofs below;
+all the other are auxiliary.
+
 \section{Lower bound for the diameter}
@@
 Now we need to estimate the cardinality of an intersection of an integral point set with a line segment.
-Assuming that the planar integral point sets contains many collinear points,
-the following result holds.
-\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
-	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
-	at least $n^\delta$ collinear points there exists a $n_0 (\varepsilon)$
-	such that for all $n \geq n_0 (\varepsilon)$ we have
-	\begin{equation}
-		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
-		.
-	\end{equation}
-\end{theorem}
-However, the estimate is rather unsuitable for our needs,
-as it does not provide the values of all the constants.
-To obtain the needed estimate, we now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.
-
-\begin{definition}
-	For a line segment $M_1 M_2$ and an integer $k$, such that $-|M_1 M_2| < k < |M_1 M_2|$,
-	we define a $\rho(k,M_1 M_2)$-curve as the set of points $N$
-	for which the equality $|N M_1| - |N M_2| = k$ holds.
-\end{definition}
-So, in the planar case a $\rho(k,M_1 M_2)$-curve
-is a branch of a hyperbola for $k\neq 0$ and the perpendicular bisector of line segment $M_1 M_2$ for $k=0$.
-
-\begin{proposition}
-	\label{obs:rho_curves}
-	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
-	the equality $|M_1 M_2| = |M_3 M_4|$ holds and line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
-	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and the $\rho(k,M_3 M_4)$-curve
-	do not intersect.
-\end{proposition}
-
-\begin{lemma}
-	\label{lem:2k-1_segments}
-	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
-	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$ with endpoints in $M$,
-	such that $|\Delta_i| = k$.
-\end{lemma}
-\begin{proof}
-	Consider a point $N\in M \setminus m$.
-	Then for each $\Delta_i$ there is a $\rho(n_i,\Delta_i)$-curve containing $N$.
-	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
-	otherwise the $\rho(n_i,\Delta_i)$-curve and the $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
-	There can be only $2k-1$ distinct values for $n_i$,
-	so there are at most $2k-1$ distinct segments $\Delta_i$.
-\end{proof}
-
-\begin{lemma}
-	\label{lem:line_segment_with n_squared_plus_one_points}
-	Let $\Delta$ be a straight line segment, $|\Delta|=l$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = n^2 + 1$.
-	Then
-	\begin{equation}
-		l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
-		.
-	\end{equation}
-\end{lemma}
-
-\begin{proof}
-	Any $n^2+1$ points, including the endpoints of $\Delta$, partition the segment $\Delta$ into $n^2$
-	sequential segments $\Delta_i$.
-	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
-	at most three segments of length 2, etc.
-	The following two expressions for sums conclude the proof:
-	\begin{equation}
-		1 + \sum_{k=1}^n (2k-1) = n^2 + 1
-		,
-	\end{equation}
-	\begin{equation}
-		\sum_{k=1}^n k(2k-1) = \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
-		.
-	\end{equation}
-\end{proof}
-
-Now we will estimate the length of a line segment that intersects an integral point set
-by an arbitrary number of points.
-
-\begin{lemma}
-	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = t$.
-	Then
-	\begin{equation}
-		\label{eq:estimate_for_segment_length}
-		b\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}
-		.
-	\end{equation}
-\end{lemma}
-
-\begin{proof}
-	Let $f(k)$ denote the mininal length of a line segment
-	that intersects an integral point set by $k$ points.
-	We observe that $f(k) > f (k-1)$.
-	Due to Lemma~\ref{lem:line_segment_with n_squared_plus_one_points},
-	$f(n^2+1) \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$.
-
-	%Consider $t$ such that $n^2 + 1 \leq t \leq (n+1)^2$. %,
-	%then $f(t) > f(n^2 +1)$.% and $n \geq \sqrt{t} - 1$.
-	For $t\in\mathbb{N}$ the inequality $(\sqrt{t} - 1)^2 +1 \leq t$ holds,
-	thus
-	%and derive that
-	\begin{equation}
-		\label{eq:line_segment_length}
-		f(t) \geq f((\sqrt{t} - 1)^2 +1) \geq \frac{2}{3}(\sqrt{t} - 1)^3+\frac{1}{2}(\sqrt{t} - 1)^2-\frac{1}{6}(\sqrt{t} - 1)
-		=
-		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2}
-		.
-	\end{equation}
-\end{proof}
-
-This lemma leads to the following proposition.
-
-\begin{proposition}
-	\label{obs:estimate_points_on_straight_line}
-	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = k$ and $b>10000$.
-	Then $k \leq \gamma_2 b + 6$,
-	where
-	\begin{equation}
-		\gamma_2 = \frac{3846}{2593 \sqrt{647}-5823} = 0.063958...
-	\end{equation}
-\end{proposition}
-
-\begin{proof}
-	We know the maximum number of points for planar integral point sets of diameters at most $10000$.
-	Thus, we are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
-	holds for $b > 10000$.
-	So, let us consider $t\geq 647$ and find a coefficient $\gamma_2$,
-	such that the inequality
-	\begin{equation}
-		\label{eq:inequality_for_linear_estimate_of_segment_length}
-		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2} \geq \frac{t-6}{\gamma_2}
-	\end{equation}
-	holds for all $t\geq 647$.
-	For such $t$,
-	the left-hand side of~\eqref{eq:inequality_for_linear_estimate_of_segment_length} obviously grows faster than the right-hand side.
-	Turning~\eqref{eq:inequality_for_linear_estimate_of_segment_length} into the same equality
-	and solving it for $t=647$, we obtain the required estimate.
-\end{proof}
 
@@
 		\\=
-		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 - \frac{a}{2} + \left(\frac{a}{c_j}\right)^2 + a\right)^{1/2}
+		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 - \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2 + a\right)^{1/2}
 		=
-		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 + \frac{a}{2} + \left(\frac{a}{c_j}\right)^2    \right)^{1/2}
+		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 + \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2    \right)^{1/2}
 		=
@@
 %\end{definition}
-
-Setting $k=1$ in Lemma~\ref{lem:2k-1_segments},
-we obtain the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018}).
-\begin{corollary}
-	\label{cor:only_one_distance_1_on_straight_line}
-	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
-	Then there is at most one pair of points $M_1,M_2\in M \cap m$
-	such that $|M_1 M_2| = 1$.
-\end{corollary}
 

### 2019-11-26 AvdMomZvol_rails/Article.tex
Iteration from Zvolinsky
Grant note

@@
 	On particular diameter bounds for integral point sets in higher dimensions
-	%\footnote{
-	%	This work was carried out at Voronezh State University and supported by the Russian Science
-	%	Foundation grant 19-11-00197.
-	%}
+	\footnote{
+		This work was carried out at Voronezh State University and supported by the Russian Science
+		Foundation grant 19-11-00197.
+	}
 }
@@
 	\begin{equation}\label{formula1}
-		d(m, n - 2 + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
+		d(m, n + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
 	\end{equation}
@@
 	\begin{equation}
-		d(m, n - k + k(m - k)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
+		d(m, n + k(m - 2)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
 	\end{equation}
@@
 Below we give the planar integral point sets and the corresponding
-estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 24$.
+estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 36$.
 For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
@@
 \begin{figure}[htbp]
+	\includegraphics[width=.48\linewidth]{picture_d17.png}
+	\hfill
+	\includegraphics[width=.48\linewidth]{picture_d21.png}
+	\\
+	\parbox{.48\linewidth}{\caption{PIPS of cardinality 7 and diameter 17}
+	\label{picture_d17.png}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{PIPS of cardinality 8 and diameter 21}
+	\label{picture_d21.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{315}/{2} * \{ (13, 1),
+(29, 1),
+(0, 0),
+(10, 0),
+(16, 0),
+(26, 0),
+(32, 0)\}
+$
+(Figure~\ref{picture_d17.png})
+
+$f = 8$, $v = 6$, $w = 10$, $\operatorname{diam(\mathcal{P})} = 17$,
+
+which gives $d(m, 2m + 3) \leq 17$.
+
+\item
+$\mathcal{P}=\sqrt{315}/{2} * \{ (13, 1),
+(29, 1),
+(0, 0),
+(10, 0),
+(16, 0),
+(26, 0),
+(32, 0),
+(42, 0)\}
+$
+(Figure~\ref{picture_d21.png})
+
+$f = 8$, $v = 6$, $w = 10$, $\operatorname{diam(\mathcal{P})} = 21$,
+
+which gives $d(m, 2m + 4) \leq 21$.
+
+\item
+$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
+(-78 , 0),
+(\pm 62, 0),
+(\pm 55 , 0),
+(\pm 10 , 0)\}
+$
+(Figure~\ref{picture_2.png})
+
+$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,
+
+which gives $d(m, 2m + 5) \leq 158$.
+
+\item
+$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
+(\pm 78 , 0),
+(\pm 62, 0),
+(\pm 55 , 0),
+(\pm 10 , 0)\}
+$
+(Figure~\ref{picture_3.png}).
+
+$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,
+
+which gives $d(m, 2m + 6) \leq 158$.
+
+\begin{figure}[h!]
 	\includegraphics[width=.48\linewidth]{picture_2.png}
@@
 \item
-$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
-(-78 , 0),
-(\pm 62, 0),
-(\pm 55 , 0),
-(\pm 10 , 0)\}
-$
-(Figure~\ref{picture_2.png})
-
-$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,
-
-which gives $d(m, 2m + 3) \leq 158$.
-
-\item
-$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
-(\pm 78 , 0),
-(\pm 62, 0),
-(\pm 55 , 0),
-(\pm 10 , 0)\}
-$
-(Figure~\ref{picture_3.png}).
-
-$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,
-
-which gives $d(m, 2m + 4) \leq 158$.
-
-\item
 $\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
@@
 
-which gives $d(m, 2m + 5) \leq 252$.
+which gives $d(m, 2m + 7) \leq 252$.
 
@@
 
-which gives $d(m, 2m + 6) \leq 297$.
+which gives $d(m, 2m + 8) \leq 297$.
 
@@
 
-which gives $d(m, 2m + 7) \leq 1073$.
+which gives $d(m, 2m + 9) \leq 1073$.
 
@@
 \center{\includegraphics[width=0.7\linewidth]{picture_6.png}}
-\parbox{0.7\linewidth}{\caption{IPS of cardinality 13 and diameter 1073}
+\parbox{0.7\linewidth}{\caption{PIPS of cardinality 13 and diameter 1073}
 \label{picture_6.png}}
@@
 
-which gives $d(m, 2m + 8) \leq 1273$.
+which gives $d(m, 2m + 10) \leq 1273$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=0.7\linewidth]{picture_7.png}}
+\parbox{0.7\linewidth}{\caption{PIPS of cardinality 14 and diameter 1273}
+\label{picture_7.png}}
+\end{figure}
 
@@
 
-which gives $d(m, 2m + 9) \leq 1888$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{picture_7.png}}
-\parbox{0.7\linewidth}{\caption{IPS of cardinality 14 and diameter 1273}
-\label{picture_7.png}}
-\end{figure}
+which gives $d(m, 2m + 11) \leq 1888$.
 
@@
 \center{\includegraphics[width=0.75\linewidth]{picture_8.png}}
-\parbox{0.75\linewidth}{\caption{IPS of cardinality 15 and diameter 1888}
+\parbox{0.75\linewidth}{\caption{PIPS of cardinality 15 and diameter 1888}
 \label{picture_8.png}}
@@
 
-which gives $d(m, 2m + 10) \leq 2189$.
+which gives $d(m, 2m + 12) \leq 2189$.
 
@@
 \center{\includegraphics[width=0.75\linewidth]{picture_9.png}}
-\parbox{0.75\linewidth}{\caption{IPS of cardinality 16 and diameter 2189}
+\parbox{0.75\linewidth}{\caption{PIPS of cardinality 16 and diameter 2189}
 \label{picture_9.png}}
@@
 
-which gives $d(m, 2m + 11) \leq 3224$.
+which gives $d(m, 2m + 13) \leq 3224$.
 
@@
 \center{\includegraphics[width=0.85\linewidth]{picture_13.png}}
-\parbox{0.85\linewidth}{\caption{IPS of cardinality 17 and diameter 3224}
+\parbox{0.85\linewidth}{\caption{PIPS of cardinality 17 and diameter 3224}
 \label{picture_13.png}}
@@
 
-which gives $d(m, 2m + 12) \leq 3696$.
+which gives $d(m, 2m + 14) \leq 3696$.
 
@@
 \center{\includegraphics[width=0.85\linewidth]{picture_14.png}}
-\parbox{0.85\linewidth}{\caption{IPS of cardinality 18 and diameter 3696}
+\parbox{0.85\linewidth}{\caption{PIPS of cardinality 18 and diameter 3696}
 \label{picture_14.png}}
@@
 
-which gives $d(m, 2m + 13) \leq 5141$.
+which gives $d(m, 2m + 15) \leq 5141$.
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_15.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 5141}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 19 and diameter 5141}
 \label{picture_15.png}}
@@
 
-which gives $d(m, 2m + 14) \leq 6586$.
+which gives $d(m, 2m + 16) \leq 6586$.
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_16.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 20 and diameter 6586}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 20 and diameter 6586}
 \label{picture_16.png}}
@@
 
-which gives $d(m, 2m + 15) \leq 11692$.
+which gives $d(m, 2m + 17) \leq 11692$.
 
@@
 
-which gives $d(m, 2m + 16) \leq 12296$.
+which gives $d(m, 2m + 18) \leq 12296$.
 
@@
 \center{\includegraphics[width=0.75\linewidth]{picture_17.png}}
-\parbox{0.75\linewidth}{\caption{IPS of cardinality 21 and diameter 11692}
+\parbox{0.75\linewidth}{\caption{PIPS of cardinality 21 and diameter 11692}
 \label{picture_17.png}}
@@
 
-which gives $d(m, 2m + 17) \leq 16027$.
+which gives $d(m, 2m + 19) \leq 16027$.
 
@@
 \center{\includegraphics[width=0.75\linewidth]{picture_18.png}}
-\parbox{0.75\linewidth}{\caption{IPS of cardinality 22 and diameter 12296}
+\parbox{0.75\linewidth}{\caption{PIPS of cardinality 22 and diameter 12296}
 \label{picture_18.png}}
@@
 
-which gives $d(m, 2m + 18) \leq 19758$.
+which gives $d(m, 2m + 20) \leq 19758$.
 
@@
 \center{\includegraphics[width=0.85\linewidth]{picture_19.png}}
-\parbox{0.85\linewidth}{\caption{IPS of cardinality 23 and diameter 16027}
+\parbox{0.85\linewidth}{\caption{PIPS of cardinality 23 and diameter 16027}
 \label{picture_19.png}}
@@
 
-which gives $d(m, 2m + 19) \leq 32054$.
+which gives $d(m, 2m + 21) \leq 32054$.
 
@@
 \center{\includegraphics[width=0.85\linewidth]{picture_20.png}}
-\parbox{0.85\linewidth}{\caption{IPS of cardinality 24 and diameter 19758}
+\parbox{0.85\linewidth}{\caption{PIPS of cardinality 24 and diameter 19758}
 \label{picture_20.png}}
-\end{figure}
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_21.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 25 and diameter 32054}
-\label{picture_21.png}}
 \end{figure}
@@
 
-which gives $d(m, 2m + 20) \leq 39516$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_22.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 26 and diameter 39516}
-\label{picture_22.png}}
+which gives $d(m, 2m + 22) \leq 39516$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_21.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 25 and diameter 32054}
+\label{picture_21.png}}
 \end{figure}
@@
 
-which gives $d(m, 2m + 21) \leq 254364$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_23.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 27 and diameter 254364}
-\label{picture_23.png}}
+which gives $d(m, 2m + 23) \leq 254364$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_22.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 26 and diameter 39516}
+\label{picture_22.png}}
 \end{figure}
@@
 
-which gives $d(m, 2m + 22) \leq 276612$.
+which gives $d(m, 2m + 24) \leq 276612$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_23.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 27 and diameter 254364}
+\label{picture_23.png}}
+\end{figure}
 
@@
 
-which gives $d(m, 2m + 23) \leq 338169$.
+which gives $d(m, 2m + 25) \leq 338169$.
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_24.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 28 and diameter 276612}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 28 and diameter 276612}
 \label{picture_24.png}}
@@
 \center{\includegraphics[width=1\linewidth]{picture_25.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 29 and diameter 338169}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 29 and diameter 338169}
 \label{picture_25.png}}
@@
 
-which gives $d(m, 2m + 24) \leq 399726$.
+which gives $d(m, 2m + 26) \leq 399726$.
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_26.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 30 and diameter 399726}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 30 and diameter 399726}
 \label{picture_26.png}}
 \end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 580290, 0),
+(\pm 430360, 0),
+(\pm 388080, 0),
+(\pm 300265, 0),
+$
+
+$
+(\pm 288960, 0),
+(\pm 217560, 0),
+(\pm 188790, 0),
+(\pm 183540, 12600),
+(\pm 178290, 0),
+$
+
+$
+(\pm 128730, 0),
+(\pm 120960, 0),
+(\pm 63280, ),
+(\pm 21000, 0),
+(\pm 11224, 0),
+$
+
+$
+(-1033912, 0),
+(-999315, 0),
+(-691530  , 0)\}
+$
+(Figure~\ref{picture_31.png}).
+
+$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P})} = 1614202$,
+
+which gives $d(m, 2m + 27) \leq 1614202$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_31.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 31 and diameter 1614202}
+\label{picture_31.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 691530, 0),
+(\pm 580290, 0),
+(\pm 430360, 0),
+(\pm 388080, 0),
+$
+
+$
+(\pm 300265, 0),
+(\pm 288960, 0),
+(\pm 217560, 0),
+(\pm 188790, 0),
+(\pm 183540, 12600),
+$
+
+$
+(\pm 178290, 0),
+(\pm 128730, 0),
+(\pm 120960, 0),
+(\pm 63280, 0),
+(\pm 21000, 0),
+$
+
+$
+(\pm 11224, 0),
+(-1033912, 0),
+(-999315, 0)\}
+$
+(Figure~\ref{picture_32.png}).
+
+$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P})} = 1725442$,
+
+which gives $d(m, 2m + 28) \leq 1725442$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_32.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 32 and diameter 1725442}
+\label{picture_32.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 999315, 0),
+(\pm 691530, 0),
+(\pm 580290, 0),
+(\pm 430360, 0),
+$
+
+$
+(\pm 388080, 0),
+(\pm 300265, 0),
+(\pm 288960, 0),
+(\pm 217560, 0),
+(\pm 188790, 0),
+$
+
+$
+(\pm 183540, 12600),
+(\pm 178290, 0),
+(\pm 128730, 0),
+(\pm 120960, 0),
+(\pm 63280, 0),
+$
+
+$
+(\pm 21000, 0),
+(\pm 11224, 0),
+(-1033912, 0)\}
+$
+(Figure~\ref{picture_33.png}).
+
+$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P})} = 2033227$,
+
+which gives $d(m, 2m + 29) \leq 2033227$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_33.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 33 and diameter 2033227}
+\label{picture_33.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 1033912, 0),
+(\pm 999315, 0),
+(\pm 691530, 0),
+(\pm 580290, 0),
+$
+
+$
+(\pm 430360, 0),
+(\pm 388080, 0),
+(\pm 300265, 0),
+(\pm 288960, 0),
+(\pm 217560, 0),
+$
+
+$
+(\pm 188790, 0),
+(\pm 183540, 12600),
+(\pm 178290, 0),
+(\pm 128730, 0),
+(\pm 120960, 0),
+$
+
+$
+(\pm 63280, 0),
+(\pm 21000, 0),
+(\pm 11224, 0)\}
+$
+(Figure~\ref{picture_34.png}).
+
+$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P})} = 2067824$,
+
+which gives $d(m, 2m + 30) \leq 2067824$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_34.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 34 and diameter 2067824}
+\label{picture_34.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 3997260, 0),
+(\pm 2766120, 0),
+(\pm 2321160, 0),
+(\pm 1721440, 0),
+$
+
+$
+(\pm 1552320, 0),
+(\pm 1201060, 0),
+(\pm 1155840, 0),
+(\pm 870240, 0),
+(\pm 755160, 0),
+$
+
+$
+(\pm 734160, 50400),
+(\pm 713160, 0),
+(\pm 514920, 0),
+(\pm 483840, 0),
+(\pm 308175, 0),
+$
+
+$
+(\pm 253120, 0),
+(\pm 84000, 0),
+(\pm 44896, 0),
+(-4135648 , 0)\}
+$
+(Figure~\ref{picture_35.png}).
+
+$f = 1468320$, $v = 12824$, $w = 1468376$, $\operatorname{diam(\mathcal{P})} = 8132908$,
+
+which gives $d(m, 2m + 31) \leq 8132908$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_35.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 35 and diameter 8132908}
+\label{picture_35.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 4135648, 0),
+(\pm 3997260, 0),
+(\pm 2766120, 0),
+(\pm 2321160, 0),
+$
+
+$
+(\pm 1721440, 0),
+(\pm 1552320, 0),
+(\pm 1201060, 0),
+(\pm 1155840, 0),
+(\pm 870240, 0),
+$
+
+$
+(\pm 755160, 0),
+(\pm 734160, 50400),
+(\pm 713160, 0),
+(\pm 514920, 0),
+(\pm 483840, 0),
+$
+
+$
+(\pm 308175, 0),
+(\pm 253120, 0),
+(\pm 84000, 0),
+(\pm 44896, 0)\}
+$
+(Figure~\ref{picture_36.png}).
+
+$f = 1468320$, $v = 12824$, $w = 1468376$, $\operatorname{diam(\mathcal{P})} = 8271296$,
+
+which gives $d(m, 2m + 32) \leq 8271296$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_36.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 36 and diameter 8271296}
+\label{picture_36.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 12991095, 0),
+(\pm 8989890, 0),
+(\pm 7543770, 0),
+(\pm 5594680, 0),
+$
+
+$
+(\pm 5045040, 0),
+(\pm 3903445, 0),
+(\pm 3756480, 0),
+(\pm 3694845, 0),
+(\pm 2828280, 0),
+$
+
+$
+(\pm 2454270, 0),
+(\pm 2386020, 163800),
+(\pm 2317770, 0),
+(\pm 1673490, 0),
+(\pm 1572480, 0),
+$
+
+$
+(\pm 822640, 0),
+(\pm 484680, 0),
+(\pm 273000, 0),
+(\pm 145912, 0),
+$
+
+$
+(-13440856, 0)\}
+$
+(Figure~\ref{picture_37.png}).
+
+$f = 4772040$, $v = 41678$, $w = 4772222$, $\operatorname{diam(\mathcal{P})} = 26431951$,
+
+which gives $d(m, 2m + 33) \leq 26431951$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_37.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 37 and diameter 26431951}
+\label{picture_37.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 13440856, 0),
+(\pm 12991095, 0),
+(\pm 8989890, 0),
+(\pm 7543770, 0),
+$
+
+$
+(\pm 5594680, 0),
+(\pm 5045040, 0),
+(\pm 3903445, 0),
+(\pm 3756480, 0),
+(\pm 3694845, 0),
+$
+
+$
+(\pm 2828280, 0),
+(\pm 2454270, 0),
+(\pm 2386020, 163800),
+(\pm 2317770, 0),
+(\pm 1673490, 0),
+$
+
+$
+(\pm 1572480, 0),
+(\pm 822640, 0),
+(\pm 484680, 0),
+(\pm 273000, 0),
+$
+
+$
+(\pm 145912, 0)\}
+$
+(Figure~\ref{picture_38.png}).
+
+$f = 4772040$, $v = 41678$, $w = 4772222$, $\operatorname{diam(\mathcal{P})} = 26881712$,
+
+which gives $d(m, 2m + 34) \leq 26881712$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_38.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 38 and diameter 26881712}
+\label{picture_38.png}}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 51964380, 0),
+(\pm 35959560, 0),
+(\pm 30175080, 0),
+(\pm 22378720, 0),
+$
+
+$
+(\pm 20180160, 0),
+(\pm 15613780, 0),
+(\pm 15025920, 0),
+(\pm 14779380, 0),
+(\pm 11313120, 0),
+$
+
+$
+(\pm 9817080, 0),
+(\pm 9544080, 655200),
+(\pm 9271080, 0),
+(\pm 6693960, 0),
+(\pm 6289920, 0),
+$
+
+$
+(\pm 4006275, 0),
+(\pm 3290560, 0),
+(\pm 1938720, 0),
+(\pm 1092000, 0),
+$
+
+$
+(\pm 583648, 0),
+(-53763424 , 0)\}
+$
+(Figure~\ref{picture_39.png}).
+
+$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam(\mathcal{P})} = 105727804$,
+
+which gives $d(m, 2m + 35) \leq 105727804$.
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{picture_39.png}}
+\parbox{1\linewidth}{\caption{PIPS of cardinality 39 and diameter 105727804}
+\label{picture_39.png}}
+\end{figure}
+
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=1\linewidth]{picture_40.png}
+	\parbox{1\linewidth}{\caption{PIPS of cardinality 40 and diameter 107526848}
+	\label{picture_40.png}}
+	\end{center}
+\end{figure}
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 53763424, 0),
+(\pm 51964380 , 0),
+(\pm 35959560 , 0),
+(\pm 30175080 , 0),
+$
+
+$
+(\pm 22378720 , 0),
+(\pm 20180160 , 0),
+(\pm 15613780 , 0),
+(\pm 15025920 , 0),
+(\pm 14779380 , 0),
+$
+
+$
+(\pm 11313120 , 0),
+(\pm 9817080 , 0),
+(\pm 9544080 , 655200),
+(\pm 9271080 , 0),
+(\pm 6693960 , 0),
+$
+
+$
+(\pm 6289920 , 0),
+(\pm 4006275 , 0),
+(\pm 3290560 , 0),
+(\pm 1938720 , 0),
+(\pm 1092000 , 0),
+$
+
+$
+(\pm 583648 , 0)\}
+$
+(Figure~\ref{picture_40.png}).
+
+$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam(\mathcal{P})}
+= 107526848$,
+
+which gives $d(m, 2m + 36) \leq 107526848$.
 

### 2019-11-30 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Fix misprints

@@
 \section{Collinear subsets of integral point sets}
-Integral point sets with many collinear ponts are very dominating examples.
+Integral point sets with many collinear points are very dominating examples.
 In this section we estimate the cardinality of an intersection of an integral point set with a line segment.
@@
 \begin{proof}
-	Let $f(k)$ denote the mininal length of a line segment
+	Let $f(k)$ denote the minimal length of a line segment
 	that intersects an integral point set by $k$ points.
@@
 
-In~\cite{kurz2008bounds}, thorough investigation of planar integral point sets with diameter at most 10 000
+In~\cite{kurz2008bounds}, thorough investigation of planar integral point sets with diameter at most 10~000
 is carried out;
-thus, we can restrict our attention to intersections of planar integral point sets with lne segments of length more than 10 000.
+thus, we can restrict our attention to intersections of planar integral point sets with line segments of length more than 10~000.
 

### 2019-11-30 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Word usage, mostly SciLexic

@@
 \begin{proof}
-	Any $n^2+1$ points, including the endpoints of $\Delta$, partition the segment $\Delta$ into $n^2$
+	Any $n^2+1$ points, including the endpoints of $\Delta$, split the segment $\Delta$ into $n^2$
 	sequential segments $\Delta_i$.
@@
 
-Concluding the section, we remark that
+To conclude, we remark that
 Corollary~\ref{cor:only_one_distance_1_on_straight_line}
@@
 
-Basing on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
+Based on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
 \cite{kurz2008bounds}, we derive the following proposition.
@@
 
-Performing some simple manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
+By means of some simple manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
 one can derive the following proposition.
@@
 	\label{rem:planar_set_with_minimeter_1_of_3_points}
-	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ naively,
+	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ ad litteram,
 	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then
@@
 	with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
-	replacing the out-of-line point with an $(m-2)$-dimensional simplex of proper side length.
+	replacing the point that is not situated on the line, with an $(m-2)$-dimensional simplex of proper side length.
 	That simplex will consist of $m-1$ points;
-	we can throw out up to all points $M_J$, except the two points with distance 1.
+	we can exclude practically all $M_J$ points, except the two points with distance 1.
 	So, the set consists of $m+1$ or more points and has distance 1 occuring in it.
@@
 
-For planar integral point sets containing distance 1, we can emphasise on the following problem.
+For planar integral point sets containing distance 1, we can emphasize the following problem.
 \begin{problem}
@@
 \section{Acknowledgements}
-Author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
-A.S. Chervinskaia for the idea of using the term ``facher'' and other linguistic advices,
+The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
+A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
 and Dr. A.S. Usachev for proofreading.

### 2020-01-02 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Improving the section structure

@@
 
+TODO: note about Section structure.
+
 \section{Collinear subsets of integral point sets}
@@
 In this section we estimate the cardinality of an intersection of an integral point set with a line segment.
+
+\subsection{General results}
 
@@
 
+\subsection{Special results}
+
 In~\cite{kurz2008bounds}, thorough investigation of planar integral point sets with diameter at most 10~000
@@
 
-\section{Lower bound for the diameter}
+\section{Planar integral point set in a square}
 
-In this section,
-we improve the lower bound for minimum diameter of planar integral point sets
-employing Point Packing in a Square problem.
-Below we introduce the problem, basic notions and results.
+In this section, we develop some methods that employ the fact that every planar integral point set is situated in a square.
+Then we establish a connection between distances in integral point sets and the Point Packing in the Square problem.
+
+\subsection{General results}
+
+Let us introduce the Problem, basic notions and results.
 
@@
 
-To prove the bound on $d(2,n)$, we also need the following results and notions from~\cite{our-vmmsh-2018}.
+Now we will estimate the side length of a square which contains a given planar integral points set.
 
@@
 
-\begin{definition}
-	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
-	the line through $M_1$ and $M_2$,
-	and the perpendicular bisector of line segment $M_1 M_2$.
-\end{definition}
-
-\begin{lemma}
-	\label{lem:intervals_cross}
-	If open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect,
-	%$M_1 \neq M_2$, $M_3 \neq M_4$,
-	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains 2 or 4 points.
-\end{lemma}
-It is important that we consider the intersection of open line segments,
-so e.g. the cases $M_1 = M_3$ and $M_3 \in M_1 M_2$ satisfy the conditions of Lemma~\ref{lem:intervals_cross}.
-
-\begin{lemma}
-	\label{lemma:quadr_diag_edges}
-	Let $ABCD$ be a convex quadrilateral on the plane.
-	Then $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
-	that is at least one diagonal is greater than at least one side.
-\end{lemma}
+\subsection{Special results}
 
@@
 
-Now we need to estimate the cardinality of an intersection of an integral point set with a line segment.
+\section{Lower bound for the diameter}
+
+In this section,
+we improve the lower bound for minimum diameter of planar integral point sets
+employing the results of two previous sections.
+
+\subsection{General results}
+
+\begin{definition}
+	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
+	the line through $M_1$ and $M_2$,
+	and the perpendicular bisector of line segment $M_1 M_2$.
+\end{definition}
+
+\begin{lemma}
+	\label{lem:intervals_cross}
+	If open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect,
+	%$M_1 \neq M_2$, $M_3 \neq M_4$,
+	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains 2 or 4 points.
+\end{lemma}
+It is important that we consider the intersection of open line segments,
+so e.g. the cases $M_1 = M_3$ and $M_3 \in M_1 M_2$ satisfy the conditions of Lemma~\ref{lem:intervals_cross}.
+
+\begin{lemma}
+	\label{lemma:quadr_diag_edges}
+	Let $ABCD$ be a convex quadrilateral on the plane.
+	Then $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
+	that is at least one diagonal is greater than at least one side.
+\end{lemma}
+
+\subsection{Special results}
 

### 2020-01-02 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Improving structure of the Section devoted to sets with side 1
[J220] Note about Section structure

@@
 
-TODO: note about Section structure.
+Sections 2--4 are devoted to the linear lower bound.
+Each Section consists of 2 Subsections: General results an Special results.
+In General results, avoid attempts to provide tight estimates;
+such attempts are done in Special results.
+
+Section 5 is devoted to integral point sets with distance 1 and has its own structure.
 
@@
 
-\section{Constructing integral point sets}
+\section{Integral point sets with distance 1}
+
+\subsection{Constructing planar integral point sets with distance 1}
 
@@
 
+\subsection{Description of planar integral point sets with distance 1}
+
 %Now we need two more definitions.
@@
 \end{theorem}
+
+\subsection{Higher dimensions}
 

### 2020-01-02 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Proof of Lemma 3.4 (ex Lemma 2.4)
As requested in Review 1, p. 1 and Review 2, medium remark 4.

@@
 \end{lemma}
+
+\begin{proof}
+	Consider $M_1, M_2 \in M$ such that $|M_1 - M_2| = d$
+	and a Cartesian coordinate system such that
+	$M_1 = (-d/2; 0)$, $M_2 = (d/2; 0)$.
+
+	Then for $M_i =(x_i, y_i)\in M$, $i > 2$,
+	the inequality $|x_i| < d/2$ holds
+	(otherwise we get $|M_i - M_1| > d$ or $|M_i - M_2| > d$).
+	Moreover, $|y_i| < d$ (otherwise we get $|M_i - M_1| > d$ or $|M_i - M_2| > d$).
+
+	Set $y_{\max} = \max_{i} y_i$, $y_{\min} = \min_{i} y_i$, $M_{\max}=(x_{\max}, y_{\max})$, $M_{\min}=(x_{\min}, y_{\min})$
+	(in case of multiple points providing the maximum or the minimum, the choice is arbitrary).
+	Then
+	\begin{multline}
+		d \geq |M_{\max} - M_{\min}| = \sqrt{(x_{\max} - x_{\min})^2 + (y_{\max} - y_{\min})^2}
+		\geq
+		\\\geq
+		\sqrt{(y_{\max} - y_{\min})^2} =
+		y_{\max} - y_{\min}
+	\end{multline}
+	Thus, the sides of the square that contains $M$ are $x=\pm d/2$,
+	$y=y_{\max}$ and $y=y_{\max} - d$.
+\end{proof}
 

### 2020-01-02 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Proof on Lemma 4.2 (ex Lemma 2.6)
See Review 2, medium remark 5

@@
 \end{lemma}
-It is important that we consider the intersection of open line segments,
+
+\begin{proof}
+	Consider straight lines $l^1_1$, $l^1_2$, $l^2_1$, $l^2_2$
+	such that $cr(M_1,M_2) = l^1_1 \cup l^1_2$ and $cr(M_3,M_4) = l^2_1 \cup l^2_2$.
+
+	The cases $l^1_1 = l^2_1$, $l^1_2 = l^2_2$ and $l^1_1 = l^2_2$, $l^1_2 = l^2_1$ are impossible,
+	because the centers of open line segments $M_1 M_2$ and $M_3 M_4$ do not coincide as far as
+	the open line segments do not intersect.
+	So, all the possible cases (up to enumeration of the straight lines) are:
+	\begin{itemize}
+		\item
+			$l^1_1 = l^2_1$, $l^1_2 \neq l^2_2$.
+			The straight lines $l^1_2$ and $l^2_2$ are both orthogonal to $l^1_1$ and thus do not intersect;
+			thus, $cr(M_1,M_2) \cap cr(M_3,M_4) = l^1_1 = l^2_1$.
+		\item
+			$l^1_1$ is parallel to $l^2_1$ and $l^1_2$ is parallel to $l^2_2$.
+			Then
+			\begin{equation*}
+				cr(M_1,M_2) \cap cr(M_3,M_4) = (l^1_1 \cup l^1_2)\cap(l^2_1 \cup l^2_2) =
+				\\=
+				(l^1_1 \cap l^2_2) \cup (l^1_2 \cap l^2_1)
+				,
+			\end{equation*}
+			that is two points.
+		\item
+			$l^1_1$ is neither parallel nor orthogonal to $l^2_1$.
+			Then the intersection contains 4 points.
+	\end{itemize}
+\end{proof}
+
+\begin{remark}
+	It is important that we consider the intersection of open line segments in the claim of Lemma~\ref{lem:intervals_cross},
 	so e.g. the cases $M_1 = M_3$ and $M_3 \in M_1 M_2$ satisfy the conditions of Lemma~\ref{lem:intervals_cross}.
+\end{remark}
 

### 2020-01-02 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Add restriction m>2
Review 2, medium remark 1

@@
 (and, more generally, let $|\Delta|$ denote the length of line segment $\Delta$).
-An \textit{integral point set} in $m$-dimensional Euclidean space is a point set $M$ such that all the distances between the
+An \textit{integral point set} in $m$-dimensional Euclidean space, $m>2$, is a point set $M$ such that all the distances between the
 points of $M$ are integers and $M$ is not situated on an $(m-1)$-dimensional hyperplane.

### 2020-01-02 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Fix minor remarks 1 to 7 from Review 2

@@
 \end{equation}
-We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$
-and its bounds; in the present paper, the case of $m=2$ will mostly be in the focus.
+We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$ and its bounds.
+In the present paper, the case of $m=2$ will mostly be in the focus.
 
@@
 as it does not provide the values of all the constants.
-To obtain the needed estimate, we now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.
+To obtain the needed estimate, we now prove a generalization of~\cite[Lemma 3]{our-vmmsh-2018}.
 
@@
 	\label{lem:2k-1_segments}
-	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
-	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$ with endpoints in $M$,
+	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $l$ be a straight line.
+	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset l$ with endpoints in $M$,
 	such that $|\Delta_i| = k$.
@@
 \begin{proof}
-	Consider a point $N\in M \setminus m$.
+	Consider a point $N\in M \setminus l$.
 	Then for each $\Delta_i$ there is a $\rho(n_i,\Delta_i)$-curve containing $N$.
@@
 	\label{cor:only_one_distance_1_on_straight_line}
-	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $m$ be a straight line.
-	Then there is at most one pair of points $M_1,M_2\in M \cap m$
+	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $l$ be a straight line.
+	Then there is at most one pair of points $M_1,M_2\in M \cap l$
 	such that $|M_1 M_2| = 1$.
@@
 
-By means of some simple manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
-one can derive the following proposition.
+Now we need to estimate $\varphi_n$ with $\beta/\sqrt{n-n_0}$.
+Using the upper bound of Theorem~\ref{thm:varphi_k_bounds},
+one can easily find the values of $\beta$ and $n_0$.
+
 \begin{proposition}
@@
 In this section,
-we improve the lower bound for minimum diameter of planar integral point sets
+we improve the lower bound for the minimum diameter of planar integral point sets
 employing the results of two previous sections.
@@
 if some new exact values of $d(2,n)$ are found;
-however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$ above.
+however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$ from above.
 More precisely, we have the following theorem.
@@
 
-Using the ``blowing up'' procedure described in~\cite[theorem 1.3]{kurz2008bounds},
+Using the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
 we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,

### 2020-01-03 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Add the proof for Lemma 4.4 (ex Lemma 2.7)
Review 2, medium remark 5
[J220] Put the references in alphabetical order or authors
As was requested by the Editor

@@
 
-\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+\usepackage[backend=biber,style=gost-numeric,sorting=nyt]{biblatex}
 \addbibresource{../common/notmy.bib}
@@
 
+\begin{proof}
+	Suppose the contrary.
+	Then $AC$ in the minimal side of the triangle $ABC$ and, thus, the angle $ABC$ is acute.
+	So are also angles $BCD$, $CDA$ and $DAB$.
+	We conclude that all angles of a convex quadrilateral $ABCD$ are acute, which is impossible.
+	The contradiction concludes the proof.
+\end{proof}
+
 \subsection{Special results}

### 2020-01-03 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Translate the titles of Russian references
Review 2, medium remark 3

@@
 Taking this into account, we denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space by
-$\mathfrak{M}(m,n)$ (using the notation in~\cite{our-vmmsh-2018})
+$\mathfrak{M}(m,n)$ (using the notation in~\cite{our-vmmsh-2018-translit})
 and denote the set of all integral point sets in $m$-dimensional Euclidean space by $\mathfrak{M}(m,\mathbb{N})$.
@@
 (See~\cite[Exercise 2.6]{garibaldi2005erdos} for some remarks.)
-The constant was improved in~\cite{our-mz-rus} to $1/8$ for all $n$ and in~\cite{our-vmmsh-2018}
+The constant was improved in~\cite{our-mz-rus-translit} to $1/8$ for all $n$ and in~\cite{our-vmmsh-2018-translit}
 to $3/8$ for sufficiently large $n$.
@@
 as it does not provide the values of all the constants.
-To obtain the needed estimate, we now prove a generalization of~\cite[Lemma 3]{our-vmmsh-2018}.
+To obtain the needed estimate, we now prove a generalization of~\cite[Lemma 3]{our-vmmsh-2018-translit}.
 
@@
 Setting $k=1$ in Lemma~\ref{lem:2k-1_segments},
-we obtain the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018}).
+we obtain the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018-translit}).
 \begin{corollary}
@@
 \begin{lemma}
-	\cite[Lemma 4]{our-vmmsh-2018}
+	\cite[Lemma 4]{our-vmmsh-2018-translit}
 	\label{lem:square_container}
@@
 For every cardinality $n$ and every squarefree number $q$
-there exists a facher set $M\in\mathfrak{M}(2,n)$ with characteristic $q$~\cite[Theorem 5]{our-vmmsh-2018}.
+there exists a facher set $M\in\mathfrak{M}(2,n)$ with characteristic $q$~\cite[Theorem 5]{our-vmmsh-2018-translit}.
 In~\cite{antonov2008maximal}, the facher sets of characteristic 1 were investigated; they were called \textit{semi-crabs}.
@@
 In order to describe all the sets $M \in \mathfrak{M}(2,n)$ with distance 1,
-we need~\cite[Proposition 6]{our-vmmsh-2018}.
+we need~\cite[Proposition 6]{our-vmmsh-2018-translit}.
 For the reader's convenience, we will state a slightly rephrased version of it and provide the proof.
@@
 	and there is a Cartesian coordinate system such that $M_1=(-1/2,0)$, $M_2=(1/2,0)$, $M_3=(0, a/2)$
-	(see~\cite[Theorem 4]{our-vmmsh-2018}).
+	(see~\cite[Theorem 4]{our-vmmsh-2018-translit}).
 	It is clear that $a\neq 0$.

### 2020-01-03 pps-linear-bound/J220-AJC-review-reply-1.tex
[J220] Reply to the reviews

\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel} %используем русский и английский языки с переносами
\usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
\usepackage[unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{indentfirst} % включить отступ у первого абзаца
\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
\graphicspath{{illustr/}}%путь к рисункам

\makeatletter
\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"

\usepackage{geometry} % Меняем поля страницы.
\geometry{left=2cm}% левое поле
\geometry{right=1cm}% правое поле
\geometry{top=2cm}% верхнее поле
\geometry{bottom=2cm}% нижнее поле

%\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра

\sloppy

\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}

%\usepackage[backend=biber,style=gost-numeric,sorting=nyt]{biblatex}
%\addbibresource{../common/notmy.bib}
%\addbibresource{../common/my.bib}

\begin{document}

Dear Editor and Referees,

Thank you for your attention paid to the paper.
I've done my best to cover all the remarks and advices,
and the present document is a detaled description of the improvements done.

First of all, I'd like to say special thanks to Referee 2 for the advice
to better guide the reader through the long list of technical auxiliary results.
As it was recommended, I've performed a major revision:
the structure of the paper has been rearranged.
Section 2 has been splitted into 3 smaller sections;
moreover, each section of main part has been divided into subsections
(see last paragraphs of Introduction for detailed description).

Due to such a rearrangement, numbers of most Lemmas, Theorems etc. have been changed,
and many results have been permutted to fit the new structure.
Thus, it is rather impossible to mention the page and line numbers of every change that has been made,
as it had been requested by the Editor.
As a possible solution, I've published all the workflow at my GitHub:
\\
https://github.com/nickkolok/sct-statya/commits/master
\\
As far as I put all my work on the topic into the single repository,
I've marked the commits concerning this paper with ``[J220]''.

As for Review 1, I agree with all the remarks completely.
\begin{enumerate}
	\item
		> Lemma 2.4: the reference is Russian, you may write a proof of Lemma 2.4 here in English

		Done. Lemma 2.4 is now Lemma 3.4 and has a self-contained proof.

	\item

		> page 10, line 3: $c_j \rightarrow 2c_J$ or you may eliminate this line.

		I've preferred the first option, i.e. replacement $c_j \rightarrow 2c_J$
\end{enumerate}

As for Review 2, I agree with all the remarks completely, except one.

Medium remarks:

\begin{enumerate}
	\item
		> The statement that integral point sets have to be finite is only true if we assume a dimension of at least 2.

		Fixed. Page 1, line -7.

	\item

		> As the author’s recent preprint
		``On diameter bounds for planar integral point sets in semi-general position''
		is not very long and overlaps the content of the present paper,
		it might be a good idea to summarize both to a single paper.

		This this the only remark for which I have to write a rebuttal for,
		and the reason is pretty simple:
		the paper ``On diameter bounds for planar integral point sets in semi-general position''
		has already been sent to another journal.
		Thus, it seems to be impossible to include its content into the present paper.
		Moreover, if the referees insist, I can
		\textbf{
		try to split the present paper into two ones:
		}
		one about the lower bound and one about integral point sets with distance 1.
		The present paper looks rather long, and the requested proofs (see below)
		have made it even more longer.
		However, if the paper will be published as a single one, then I'll also be very glad.

	\item

		> Please translate the titles of the non-english reference, like [8] and [3],
		into english or provide a link where those papers can be accessed.

		The translation has been done.
		However, the numbers of references have been changed, because
		the Editor requested to put the references in alphabetical order of authors.

	\item

		> Please add a self-contained proof for Lemma 2.4 since not all readers, including me, can read the one in [3].

		Done. Lemma 2.4 is now Lemma 3.4 and has a self-contained proof.

	\item

		> Is Lemma 2.6 also from [3] or have an obvious proof? Same for Lemma 2.7.

		Lemma 2.6 is now Lemma 4.2 and has a self-contained proof.
		\\
		Lemma 2.7 is now Lemma 4.4 and has a self-contained proof.

	\item

		> Was the inequality $d(2, 123) \leq d(2, 21491)$ used in the proof of Proposition 2.8?

		Yes, it was.
		The comment has been added before the proposition (which is now Proposition 3.5).

	\item

		> Please add a proof for Proposition 2.12.

		Thank you ver much for this remark.
		I have to admit that the result is not formulated correctly.
		The edits affect Definition 2.2 (ex Defintion 2.11), Proposition 2.3 (ex Proposition 2.12) and Lemma 2.4 (ex Lemma 2.13).
		Proposition 2.3 now has a self-contained proof.

	\item

		> The “blow up” procedure needs an explanation – not everybody is an expert on this.

		Construction 5.10 now provides the explanation.

\end{enumerate}

Minor remarks have been implemented precisely as they had been written.
As for minor remark 3, a brief proof for the corresponding Proposition 3.6 has been added.

%\printbibliography

\end{document}

### 2020-01-04 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Review 2, medium remark 6
[J220] Thank referees in Acknowledgements
[J220] Proper umlaut for Erdos in Introduction

@@
 points of $M$ are integers and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
-Erd\"os and Anning proved~\cite{anning1945integral,erdos1945integral} that every integral point set consists of a finite number of points.
+Erd{\H{o}}s and Anning proved~\cite{anning1945integral,erdos1945integral} that every integral point set consists of a finite number of points.
 Taking this into account, we denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space by
@@
 
+It is clear that for $n>123$ we have $d(2,n) \geq d(2,123)$.
 Based on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
@@
 and Dr. A.S. Usachev for proofreading.
+Also the author would like to owe a special thanks to the anonymous referees
+for their helpful comments and suggestions on the paper.
 

### 2020-01-04 pps-linear-bound/pps-linear-lower-bound.tex
[J220] SciLexic to Acknowledgements
[J220] Review 2, medium remark 7

@@
 	For a line segment $M_1 M_2$ and an integer $k$, such that $-|M_1 M_2| < k < |M_1 M_2|$,
-	we define a $\rho(k,M_1 M_2)$-curve as the set of points $N$
+	we define a $\rho(k,M_1, M_2)$-curve as the set of points $N$
 	for which the equality $|N M_1| - |N M_2| = k$ holds.
 \end{definition}
-So, in the planar case a $\rho(k,M_1 M_2)$-curve
+So, in the planar case a $\rho(k,M_1, M_2)$-curve
 is a branch of a hyperbola for $k\neq 0$ and the perpendicular bisector of line segment $M_1 M_2$ for $k=0$.
@@
 	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
-	the equality $|M_1 M_2| = |M_3 M_4|$ holds and line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
-	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and the $\rho(k,M_3 M_4)$-curve
+	the equality $|M_1 M_2| = |M_3 M_4| =h $ holds, line segments $M_1 M_2$ and $M_3 M_4$ do not coincide
+	and vectors $\overrightarrow{M_1 M_2}$ and $\overrightarrow{M_3 M_4}$ are of equal direction,
+	then for a fixed integer $k$, $|k| < h$, the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve
 	do not intersect.
 \end{proposition}
+
+\begin{proof}
+	For $k=0$, the claim is obvious.
+	Indeed, in this case the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve
+	are distinct hyperplanes orthogonal to $l$ and thus do not intersect.
+
+	For $k\neq 0$, suppose the contrary.
+	Let $N$ be a point in the intersection.
+
+	Consider a plane $\Pi \supset (l \cup \{N\})$
+	with a Cartesian coordinate system such that $M_i = (x_i,0)$ for $1 \leq i \leq 4$.
+	Let $R_1$ and $R_2$ be the projections
+	of the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve onto $\Pi$.
+
+	Then $R_1$ is a branch of a hyperbola defined by the equation
+	\begin{equation}
+		x- \frac{x_1+x_2}{2}=a(h,k)\frac{\sqrt{y^2+b^2(h,k)}}{b(h,k)}
+	\end{equation}
+	and $R_2$ is a branch of a hyperbola defined by the equation
+	\begin{equation}
+		x- \frac{x_3+x_4}{2}=a(h,k)\frac{\sqrt{y^2+b^2(h,k)}}{b(h,k)}
+		,
+	\end{equation}
+	where $a(h,k)$ and $b(h,k)$ are coefficients which depend on $h$ and $k$.
+	For $N=(x_N, y_N) \in R_1 \cap R_2$ we have
+	\begin{equation}
+		\begin{cases}
+			x_N- \dfrac{x_1+x_2}{2}=a(h,k)\dfrac{\sqrt{y_N^2+b^2(h,k)}}{b(h,k)}
+			\\[4pt]\\
+			x_N- \dfrac{x_3+x_4}{2}=a(h,k)\dfrac{\sqrt{y_N^2+b^2(h,k)}}{b(h,k)}
+		\end{cases}
+	\end{equation}
+	and, thus, $x_1+x_2 = x_3+x_4$.
+	On the other hand, $x_2 - x_1 = x_4 - x_3 = h$,
+	and we conclude that $x_1 = x_3$ as well as $x_2 = x_4$.
+	So, line segments $M_1 M_2$ and $M_3 M_4$ coincide.
+	The contradiction concludes the proof.
+\end{proof}
 
@@
 	Consider a point $N\in M \setminus l$.
-	Then for each $\Delta_i$ there is a $\rho(n_i,\Delta_i)$-curve containing $N$.
+	Let $\Delta_i = M_i^1 M_i^2$, and
+	let vectors $\overrightarrow{M_i^1 M_i^2}$ and $\overrightarrow{M_j^1 M_j^2}$ be of equal direction for every $i$ and $j$.
+	Then for each $i$ there is a $\rho(n_i, M_i^1, M_i^2)$-curve containing $N$.
 	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
-	otherwise the $\rho(n_i,\Delta_i)$-curve and the $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
+	otherwise the $\rho(n_i,M_i^1, M_i^2)$-curve and the $\rho(n_i,M_j^1, M_j^2)$-curve, $j\neq i$, do not intersect.
 	There can be only $2k-1$ distinct values for $n_i$,
@@
 A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
-and Dr. A.S. Usachev for proofreading.
+Dr.~A.S.~Usachev for proofreading,
+and E.A. Momot for the ``SciLexic'' project that helped the author with some word usage issues.
 Also the author would like to owe a special thanks to the anonymous referees

### 2020-01-04 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Review 2, minor remark 3
Reformulaton and a brief proof
[J220] Review 2, medium remark 8

@@
 
-Now we need to estimate $\varphi_n$ with $\beta/\sqrt{n-n_0}$.
-Using the upper bound of Theorem~\ref{thm:varphi_k_bounds},
-one can easily find the values of $\beta$ and $n_0$.
+Now we will use the upper bound of Theorem~\ref{thm:varphi_k_bounds}
+to estimate $\varphi_n$ with $\beta/\sqrt{n-2}$.
 
@@
 \end{proposition}
+
+\begin{proof}
+	Indeed, for $n \geq 21492$ we have
+	\begin{multline*}
+		\varphi_n
+		\leq
+		\varphi_{n-1}
+		\leq
+		\frac{1}{n-2} +
+		\sqrt{
+			\frac{1}{(n-2)^2}
+			+
+			\frac{2}{(n-2)\sqrt{3}}
+		}
+		=
+		\\=
+		\frac{1}{\sqrt{n-2}}
+		\cdot
+		\frac{1}{\sqrt{n-2}}
+		 +
+		\sqrt{
+			\frac{1}{(n-2)^2}
+			+
+			\frac{2}{(n-2)\sqrt{3}}
+		}
+		\leq
+		\\ \leq
+		\frac{1}{\sqrt{n-2}}
+		\cdot
+		\frac{1}{\sqrt{21490}}
+		 +
+		\sqrt{
+			\frac{1}{(n-2)^2}
+			+
+			\frac{2}{(n-2)\sqrt{3}}
+		}
+		=
+		\\=
+		\frac{1}{\sqrt{n-2}}
+		\left(
+		\frac{1}{\sqrt{21490}}
+		 +
+		\sqrt{
+			\frac{1}{n-2}
+			+
+			\frac{2}{\sqrt{3}}
+		}
+		\right)
+		\leq
+		\\ \leq
+		\frac{1}{\sqrt{n-2}}
+		\left(
+		\frac{1}{\sqrt{21490}}
+		 +
+		\sqrt{
+			\frac{1}{21490}
+			+
+			\frac{2}{\sqrt{3}}
+		}
+		\;
+		\right)
+		.
+	\end{multline*}
+\end{proof}
 
@@
 
-Using the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
-we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,
+To build integral points sets with distance 1 in higher dimensions,
+we need the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds}.
+
+\begin{construction}
+	\label{con:blowing_up}
+	Let $M\in\mathfrak{M}(2,n)$ have many points on a straight line $l_1$
+	and $M\setminus l_1 \subset l_2$ for a straight line $l_2$.
+	Moreover, let $l_1$ be parallel to $l_2$.
+
+	Construct $M' \subset \mathbb{R}^m$ by replacing all the points $M_1,...,M_n\in M \cap l_2$
+	by equal $(m-2)$--dimensional simplices $S_1,...,S_n$
+	in such a way that
+	\begin{equation}
+		\forall(j=1,...,n) \forall(N_i \in M \cap l_1 )\forall\left(M^j_k, M^j_h \in S_j\right)\left[|N_i M^j_k| = |N_i M^j_h|\right].
+	\end{equation}
+	Then under certain conditions (in particular, if $\#(M \cap l_2)=1$ that we actually need)
+	we get $M'\in\mathfrak{M}(m,\mathbb{N})$.
+
+	For further examples we also refer the reader to~\cite{avdeev2019particular}.
+\end{construction}
+
+Now we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,
 with distance 1 occuring in it.
@@
 	Take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
-	with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
+	with sufficiently large height and then apply the ``blowing up'' procedure described in Construction~\ref{con:blowing_up},
 	replacing the point that is not situated on the line, with an $(m-2)$-dimensional simplex of proper side length.

### 2020-01-04 pps-linear-bound/pps-linear-lower-bound.tex
[J220] A self-found bug

@@
 	\label{prb:square_side_length_and_diameter_linear}
-	What are the maximal numbers $\delta \leq 1$ and $h \leq 0$
+	What are the minimal numbers $\delta \leq 1$ and $h \leq 0$
 	such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	is situated in a square of side length $\delta d - h$?
+	is situated in a square of side length $\delta d + h$?
 \end{problem}

### 2020-01-04 pps-linear-bound/J220-AJC-review-reply-1.tex
[J220] Reply - SciLexic

@@
 I've done my best to cover all the remarks and advices,
-and the present document is a detaled description of the improvements done.
+and the present document is a detailed description of the improvements done.
 
@@
 the structure of the paper has been rearranged.
-Section 2 has been splitted into 3 smaller sections;
+Section 2 has been split into 3 smaller sections;
 moreover, each section of main part has been divided into subsections
@@
 Due to such a rearrangement, numbers of most Lemmas, Theorems etc. have been changed,
-and many results have been permutted to fit the new structure.
+and many results have been permuted to fit the new structure.
 Thus, it is rather impossible to mention the page and line numbers of every change that has been made,
@@
 		I have to admit that the result is not formulated correctly.
-		The edits affect Definition 2.2 (ex Defintion 2.11), Proposition 2.3 (ex Proposition 2.12) and Lemma 2.4 (ex Lemma 2.13).
+		The edits affect Definition 2.2 (ex Definition 2.11), Proposition 2.3 (ex Proposition 2.12) and Lemma 2.4 (ex Lemma 2.13).
 		Proposition 2.3 now has a self-contained proof.

### 2020-01-11 pps-linear-bound/J220-AJC-review-reply-1.tex
[J220] Minor improvements to the reply

@@
 Thank you for your attention paid to the paper.
-I've done my best to cover all the remarks and advices,
+I've done my best to incorporate all the remarks and advices,
 and the present document is a detailed description of the improvements done.
@@
 
-		This this the only remark for which I have to write a rebuttal for,
-		and the reason is pretty simple:
+		This is the only remark for which I have to write a rebuttal,
+		and the reason is that
 		the paper ``On diameter bounds for planar integral point sets in semi-general position''
-		has already been sent to another journal.
+		has already been submitted to another journal.
 		Thus, it seems to be impossible to include its content into the present paper.
@@
 		Thank you ver much for this remark.
-		I have to admit that the result is not formulated correctly.
+		I have to admit that the result was not formulated correctly.
 		The edits affect Definition 2.2 (ex Definition 2.11), Proposition 2.3 (ex Proposition 2.12) and Lemma 2.4 (ex Lemma 2.13).
@@
 
-Minor remarks have been implemented precisely as they had been written.
+Minor remarks have been implemented precisely as they had been suggested.
 As for minor remark 3, a brief proof for the corresponding Proposition 3.6 has been added.

### 2020-02-21 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Make the reference list even more cute
[J220] Second round of remarks

@@
 
-\usepackage[backend=biber,style=gost-numeric,sorting=nyt]{biblatex}
+\usepackage[backend=biber,bibstyle=ieee,citestyle=numeric,sorting=nyt]{biblatex}
+\DeclareFieldFormat{sentencecase}{#1}
 \addbibresource{../common/notmy.bib}
@@
 Sections 2--4 are devoted to the linear lower bound.
-Each Section consists of 2 Subsections: General results an Special results.
-In General results, avoid attempts to provide tight estimates;
-such attempts are done in Special results.
+Each Section consists of 2 Subsections: General results and Special results.
+In General results, we do not aim to provide tight estimates;
+they are discussed in Special results.
 
@@
 due to the following well-known theorem,
-which can be proved by an application of inversion~\cite{solymosi2010question}.
+which can be proved by an application of circle inversion~\cite{solymosi2010question}.
 \begin{theorem}

### 2020-03-04 pps-linear-bound/pps-linear-lower-bound.tex
Employ biblatex2bibitem
[J220] Remarks from AJC, part 1

@@
 
+%\usepackage{cite}
+
+\usepackage{biblatex2bibitem}
+
 \theoremstyle{plain}
@@
 	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
+	\\
+	\textit{Voronezh State University, Voronezh, Russia}
 }
@@
 we give a construction of an integral point set $M$ of $n$ points in $m$-dimensional Euclidean space,
-where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
+where $M$ contains points $M_1$ and $M_2$ such that the distance between $M_1$ and $M_2$ is exactly $d$.
 
@@
 and denote the set of all integral point sets in $m$-dimensional Euclidean space by $\mathfrak{M}(m,\mathbb{N})$.
-The symbol $\# M$ will be used for cardinality of $M$, that is the number of points in $M$ in our case.
+The symbol $\# M$ will be used for the cardinality of $M$, that is the number of points in $M$ in our case.
 
@@
 We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$ and its bounds.
-In the present paper, the case of $m=2$ will mostly be in the focus.
+In the present paper, we mostly focus on the case $m=2$.
 
@@
 as it does not provide the values of all the constants.
-To obtain the needed estimate, we now prove a generalization of~\cite[Lemma 3]{our-vmmsh-2018-translit}.
+To obtain the required estimate, we now prove a generalization of~\cite[Lemma 3]{our-vmmsh-2018-translit}.
 
@@
 Now we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,
-with distance 1 occuring in it.
-If we have an integral point set with distance 1 occuring,
+with distance 1 occurring in it.
+If we have an integral point set with distance 1 occurring,
 then we can easily dilate it to turn 1 into the desired distance.
@@
 	We will call an integral point set $M$ \textit{prime}, if the greatest common divisor
-	of all the distances occuring in $M$ is 1.
+	of all the distances occurring in $M$ is 1.
 \end{definition}
@@
 	we can exclude practically all $M_J$ points, except the two points with distance 1.
-	So, the set consists of $m+1$ or more points and has distance 1 occuring in it.
+	So, the set consists of $m+1$ or more points and has distance 1 occurring in it.
 	As distance 1 occurs in the obtained set, it is prime.
@@
 \end{theorem}
-Such a set can not contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar},
+Such a set cannot contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar},
 the points with distance 1 should be $(\pm 1/2, 0)$.
@@
 \end{conjecture}
-Such estimate is based on the number of edges in an $(m-2)$-dimensional simplex.
+Such an estimate is based on the number of edges in an $(m-2)$-dimensional simplex.
 
@@
 	\label{hyp:prime_planar}
-	For every $n \geq 3$, $d \geq 1$ there exists an \textbf{prime}
+	For every $n \geq 3$, $d \geq 1$ there exists a \textbf{prime}
 	set $M\in\mathfrak{M}(2,n)$ which
@@
 
+\printbibitembibliography
+
 \end{document}

### 2020-03-25 pps-linear-bound/pps-linear-lower-bound.tex
[J220][auto] Only referenced equations are numbered
Switch to homemade AJC style
[J220] Switch to bibitems

@@
 
-\usepackage[backend=biber,bibstyle=ieee,citestyle=numeric,sorting=nyt]{biblatex}
-\DeclareFieldFormat{sentencecase}{#1}
-\addbibresource{../common/notmy.bib}
-\addbibresource{../common/my.bib}
-
-%\usepackage{cite}
-
-\usepackage{biblatex2bibitem}
+%\usepackage[backend=biber,style=ajc2020unofficial]{biblatex}
+%\addbibresource{../common/notmy.bib}
+%\addbibresource{../common/my.bib}
+
+\usepackage{cite}
+
+%\usepackage{biblatex2bibitem}
 
@@
 \newtheorem{problem}[theorem]{Problem}
+
+%Only referenced equations are numbered
+\let\etoolboxforlistloop\forlistloop % save the good meaning of \forlistloop
+\usepackage{mathtools}
+\mathtoolsset{showonlyrefs}
+\let\forlistloop\etoolboxforlistloop % restore the good meaning of \forlistloop
+
+% https://tex.stackexchange.com/questions/220268/biblatex-and-autonum-dont-work-together
+
+%\mathtoolsset{showonlyrefs=false}
+% (write an equation/multline to be force-numbered here)
+%\mathtoolsset{showonlyrefs=true}
 
@@
 
-\printbibliography
-
-\printbibitembibliography
+%\printbibliography
+
+%\printbibitembibliography
+
+\begin{thebibliography}{99}
+%
+\bibitem{anning1945integral}
+N. H. Anning and P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{8} (1945), 598–600, \textsc{doi}: \href
+{https://doi.org/10.1090/S0002-9904-1945-08407-9} {\nolinkurl {10.1090/S0002-9904-1945-08407-9}}.
+%
+\bibitem{antonov2008maximal}
+A. R. Antonov and S. Kurz, Maximal integral point sets over {$\mathbb {Z}^2$}, \emph{Int. J. Comput. Math.} 87.\textbf{12}
+(2008), 2653–2676, \textsc{doi}: \href {https://doi.org/10.1080/00207160902993636} {\nolinkurl {10.1080/00207160902993636}},
+arXiv: \href {http://arxiv.org/abs/0804.1280} {\nolinkurl {0804.1280}}.
+%
+\bibitem{our-mz-rus-translit}
+N. N. Avdeev and E. M. Semenov, On point sets in the plane with integral distances (О множествах точек на плоскости с
+целочисленными расстояниями), \emph{Matematicheskie zametki} 100.\textbf{5} (2016), 757–761.
+%
+\bibitem{our-vmmsh-2018-translit}
+N. N. Avdeev and E. M. Semenov, Integral point sets in the plane and Euclidean space (Множества точек с целочисленными
+расстояниями на плоскости и в евклидовом пространстве), \emph{Matematicheskij forum (Itogi nauki. Yug Rossii)} (2018),
+217–236.
+%
+\bibitem{avdeev2019particular}
+N. N. Avdeev, R. E. Zvolinsky, and E. A. Momot, On particular diameter bounds for integral point sets in higher dimensions
+(2019), arXiv: \href {http://arxiv.org/abs/1909.10386} {\nolinkurl {1909.10386}}, \textsc{url}: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190910386A/abstract}.
+%
+\bibitem{bat2018number}
+G. Bat-Ochir, On the number of points with pairwise integral distances on a circle, \emph{Discrete Appl. Math.} 254 (2018),
+17–32, \textsc{doi}: \href {https://doi.org/10.1016/j.dam.2018.07.004} {\nolinkurl {10.1016/j.dam.2018.07.004}}.
+%
+\bibitem{costa2013valid}
+A. Costa, Valid constraints for the point packing in a square problem, \emph{Discrete Appl. Math.} 161.\textbf{18} (2013),
+2901–2909.
+%
+\bibitem{erdos1945integral}
+P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{12} (1945), 996, \textsc{doi}: \href
+{https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl {10.1090/S0002-9904-1945-08490-0}}.
+%
+\bibitem{garibaldi2005erdos}
+J. Garibaldi and A. Iosevich, The Erdős distance problem: lecture notes (2005).%
+\bibitem{harborth1993upper}
+H. Harborth, A. Kemnitz, and M. Möller, An upper bound for the minimum diameter of integral point sets, \emph{Discrete
+Comput. Geom.} 9.\textbf{4} (1993), 427–432, \textsc{doi}: \href {https://doi.org/10.1007/bf02189331} {\nolinkurl
+{10.1007/bf02189331}}.
+%
+\bibitem{kurz2005characteristic}
+S. Kurz, On the characteristic of integral point sets in {$\mathbb {E}^m$}, \emph{Australas. J. Combin.} 36 (2006), 241–248,
+arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl {math/0511704}}.
+%
+\bibitem{kurz2008bounds}
+S. Kurz and R. Laue, Bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 39 (2007), 233–240,
+arXiv: \href {http://arxiv.org/abs/0804.1296} {\nolinkurl {0804.1296}}.
+%
+\bibitem{kurz2008minimum}
+S. Kurz and A. Wassermann, On the minimum diameter of plane integral point sets, \emph{Ars Combin.} 101 (2011), 265–287,
+arXiv: \href {http://arxiv.org/abs/0804.1307} {\nolinkurl {0804.1307}}.
+%
+\bibitem{locatelli2002packing}
+M. Locatelli and U. Raber, Packing equal circles in a square: a deterministic global optimization approach, \emph{Discrete Appl.
+Math.} 122.\textbf{1-3} (2002), 139–166.
+%
+\bibitem{noll1989nclusters}
+L. C. Noll and D. I. Bell, {$n$}-clusters for {$1<n<7$}, \emph{Math. Comp.} 53.\textbf{187} (1989), 439–444.
+%
+\bibitem{nozaki2013lower}
+H. Nozaki, Lower bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 56 (2013), 139–143.
+%
+\bibitem{piepmeyer1996maximum}
+L. Piepmeyer, The maximum number of odd integral distances between points in the plane, \emph{Discrete Comput. Geom.}
+16.\textbf{1} (1996), 113–115, \textsc{doi}: \href {https://doi.org/10.1007/bf02711135} {\nolinkurl {10.1007/bf02711135}}.
+%
+\bibitem{solymosi2003note}
+J. Solymosi, Note on integral distances, \emph{Discrete Comput. Geom.} 30.\textbf{2} (2003), 337–342, \textsc{doi}: \href
+{https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.
+%
+\bibitem{solymosi2010question}
+J. Solymosi and F. De Zeeuw, On a question of Erdős and Ulam, \emph{Discrete Comput. Geom.} 43.\textbf{2} (2010), 393–401,
+arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl {0806.3095}}.
+\end{thebibliography}
 

### 2020-03-25 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Bibliography: use []s instead of dots
[J220] Capitalization issues

@@
 \graphicspath{{illustr/}}%путь к рисункам
-
-\makeatletter
-\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
-\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"
 
@@
 Sections 2--4 are devoted to the linear lower bound.
-Each Section consists of 2 Subsections: General results and Special results.
+Each section consists of 2 subsections: General results and Special results.
 In General results, we do not aim to provide tight estimates;
@@
 
-Let us introduce the Problem, basic notions and results.
+Let us introduce the problem, basic notions and results.
 

### 2020-03-25 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Avoid dashes next to formulas

@@
 
-	a) $N$ belongs to $C$~--- overall at most $\gamma_2 b + 6$ points by Lemma~\ref{lem:intervals_cross} and Proposition~\ref{obs:estimate_points_on_straight_line};
+	a) $N$ belongs to $C$; overall at most $\gamma_2 b + 6$ points by Lemma~\ref{lem:intervals_cross} and Proposition~\ref{obs:estimate_points_on_straight_line};
 
 	b) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas
-	with one of $|M_3 M_4| - 1$ hyperbolas~--- overall at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
-
-	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas with $cr(M_3 M_4)$~---
+	with one of $|M_3 M_4| - 1$ hyperbolas; overall at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
+
+	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas with $cr(M_3 M_4)$;
 	overall at most $4 (|M_1 M_2| - 1)$ points;
 
-	d) $N$ belongs to the intersection of one of $|M_3 M_4| - 1$ hyperbolas with $cr(M_1 M_2)$~---
+	d) $N$ belongs to the intersection of one of $|M_3 M_4| - 1$ hyperbolas with $cr(M_1 M_2)$;
 	overall at most $4 (|M_3 M_4| - 1)$ points;
-
+	\\
 	(see~\cite{erdos1945integral} for details).

### 2021-02-23 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Very drafts

--- 2019-10-21 AvdMomZvol_rails/Article.tex
@@
 \title{
-	On particular diameter bounds for integral point sets in higher dimensions
-	%\footnote{
-	%	This work was carried out at Voronezh State University and supported by the Russian Science
-	%	Foundation grant 19-11-00197.
-	%}
+	On particular examples of planar integral point sets and their classification
+	\footnote{
+		This work was carried out at Voronezh State University and supported by the Russian Science
+		Foundation grant 19-11-00197.
+	}
 }
@@
 	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
-	, Zvolinsky R.E., Momot E.A.
+	, Momot E.A., Zvolinsky A.E.
 	\\
@@
 
-Definition \ref{dfn1} can be generalized as follows:
-
-\begin{dfn}
-	An integral point set $\mathcal{P}$ is a set of $n$ points in
-	the $m$-dimensional Euclidean space $\mathbb{R}^{m}$ with pairwise
-	integral distances.
-\end{dfn}
-
-How can we characterize an integral point set?
-First of all, we can look at its dimension;
-then, we can count the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
+How do we characterize an integral point set?
+For example, we can count the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
 and is further called the cardinality;
-finally, we can naturally% TODO: naturally/essentially? cf. arxiv/1
+alternatively, we can naturally% TODO: naturally/essentially? cf. arxiv/1
 define the diameter of a finite point set.
@@
 
-\section{Bounds based on rails sets}
-
-Let $i = \overline{1, k}$ denote the enumeration of all $i$
-from $1$ to $k$.
-Theorem \cite[Theorem 2.1]{kurz2008bounds} can be generalized as follows:
-
-\begin{theorem}
-	Let $\mathcal{P}$ be a planar integral point set consisting of
-	$n - k$ points on line $l_1$ and $k$ points $P_1$, $P_2$, ..., $P_k$ on a
-	parallel line $l_2$ with distance $r$ between $l_1$ and $l_2$. If there
-	exist positive integers $v$, $w_{ij}$, with $f_{ij}^{2} + v^{2}
-	= w_{ij}^{2}$ and $v < 2r$, where $i = \overline{1, k - 1}$, $j =
-	\overline{i + 1, k}$, $|P_{i}P_{j}| = f_{ij}$,
-	then
-
-	\begin{equation}
-		d(m, n - k + k(m - k)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
-	\end{equation}
-
-\end{theorem}
-
-Below we give the planar integral point sets and the corresponding
-estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 24$.
-For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
-$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
-which means that each abscissa is multiplied by $1/q$
-and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
-$$
-	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
-	=
-	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
-$$
-
-\begin{itemize}
-\setlength{\itemsep}{-1mm}
-
-\begin{figure}[htbp]
-	\includegraphics[width=.48\linewidth]{picture_2.png}
-	\hfill
-	\includegraphics[width=.48\linewidth]{picture_3.png}
-	\\
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 9 and diameter 158}
-	\label{picture_2.png}}
-	\hfill
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 10 and diameter 158}
-	\label{picture_3.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
-(-78 , 0),
-(\pm 62, 0),
-(\pm 55 , 0),
-(\pm 10 , 0)\}
-$
-(Figure~\ref{picture_2.png})
-
-$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,
-
-which gives $d(m, 2m + 3) \leq 158$.
-
-\item
-$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
-(\pm 78 , 0),
-(\pm 62, 0),
-(\pm 55 , 0),
-(\pm 10 , 0)\}
-$
-(Figure~\ref{picture_3.png}).
-
-$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,
-
-which gives $d(m, 2m + 4) \leq 158$.
-
-\item
-$\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
-(\pm 207 , 0),
-(\pm 145 , 0),
-(\pm 63 , 0),
-(\pm 25 , 0),
-(-297 , 0)\}
-$
-(Figure~\ref{picture_4.png}).
-
-$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P})} = 252$,
-
-which gives $d(m, 2m + 5) \leq 252$.
-
-\begin{figure}[htbp]
-	\includegraphics[width=.48\linewidth]{picture_4.png}
-	\hfill
-	\includegraphics[width=.48\linewidth]{picture_5.png}
-	\\
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 11 and diameter 252}
-	\label{picture_4.png}}
-	\hfill
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 12 and diameter 297}
-	\label{picture_5.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
-(\pm 297 , 0),
-(\pm 207 , 0),
-(\pm 145 , 0),
-(\pm 63 , 0),
-(\pm 25 , 0)\}
-$
-(Figure~\ref{picture_5.png}).
-
-$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P})} = 297$,
-
-which gives $d(m, 2m + 6) \leq 297$.
-
-\item
-$\mathcal{P}=\sqrt{19019}/{2} * \{ (\pm 200, 3),
-(\pm 873 , 0),
-(\pm 615 , 0),
-(\pm 377 , 0),
-(\pm 215 , 0),
-(\pm 23 , 0),
-$
-
-$
-(-1273 , 0)\}
-$
-(Figure~\ref{picture_6.png}).
-
-$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P})} = 1073$,
-
-which gives $d(m, 2m + 7) \leq 1073$.
+\section{Rails sets}
 
 \begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{picture_6.png}}
-\parbox{0.7\linewidth}{\caption{IPS of cardinality 13 and diameter 1073}
-\label{picture_6.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{19019}/{2} * \{ (\pm 200, 3),
-(\pm 1273 , 0),
-(\pm 873 , 0),
-(\pm 615 , 0),
-(\pm 377 , 0),
-(\pm 215 , 0),
-$
-
-$
-(\pm 23 , 0)\}
-$
-(Figure~\ref{picture_7.png}).
-
-$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P})} = 1273$,
-
-which gives $d(m, 2m + 8) \leq 1273$.
-
-\item
-$\mathcal{P}=\sqrt{385}/{2} * \{ (\pm 1105, 48),
-(\pm 1587 , 0),
-(\pm 1269 , 0),
-(\pm 763 , 0),
-(\pm 623 , 0),
-(\pm 529 , 0),
-$
-
-$
-(\pm 339 , 0),
-(-2189 , 0)\}
-$
-(Figure~\ref{picture_8.png}).
-
-$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P})} = 1888$,
-
-which gives $d(m, 2m + 9) \leq 1888$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{picture_7.png}}
-\parbox{0.7\linewidth}{\caption{IPS of cardinality 14 and diameter 1273}
-\label{picture_7.png}}
-\end{figure}
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.75\linewidth]{picture_8.png}}
-\parbox{0.75\linewidth}{\caption{IPS of cardinality 15 and diameter 1888}
-\label{picture_8.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{385}/{2} * \{ (\pm 1105, 48),
-(\pm 2189 , 0),
-(\pm 1587 , 0),
-(\pm 1269 , 0),
-(\pm 763 , 0),
-(\pm 623 , 0),
-$
-
-$
-(\pm 529 , 0),
-(\pm 339 , 0)\}
-$
-(Figure~\ref{picture_9.png}).
-
-$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P})} = 2189$,
-
-which gives $d(m, 2m + 10) \leq 2189$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.75\linewidth]{picture_9.png}}
-\parbox{0.75\linewidth}{\caption{IPS of cardinality 16 and diameter 2189}
-\label{picture_9.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
-(\pm 1376 , 0),
-(\pm 1036 , 0),
-(\pm 899 , 0),
-(\pm 849 , 0),
-(\pm 613 , 0),
-$
-
-$
-(\pm 576 , 0),
-(\pm 100 , 0),
-(-1848 , 0)\}
-$
-(Figure~\ref{picture_13.png}).
-
-$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 3224$,
-
-which gives $d(m, 2m + 11) \leq 3224$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.85\linewidth]{picture_13.png}}
-\parbox{0.85\linewidth}{\caption{IPS of cardinality 17 and diameter 3224}
-\label{picture_13.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
-(\pm 1848 , 0),
-(\pm 1376 , 0),
-(\pm 1036 , 0),
-(\pm 899 , 0),
-(\pm 849 , 0),
-$
-
-$
-(\pm 613 , 0),
-(\pm 576 , 0),
-(\pm 100 , 0)\}
-$
-(Figure~\ref{picture_14.png}).
-
-$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 3696$,
-
-which gives $d(m, 2m + 12) \leq 3696$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.85\linewidth]{picture_14.png}}
-\parbox{0.85\linewidth}{\caption{IPS of cardinality 18 and diameter 3696}
-\label{picture_14.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
-(\pm 1848 , 0),
-(\pm 1376 , 0),
-(\pm 1036 , 0),
-(\pm 899 , 0),
-(\pm 849 , 0),
-$
-
-$
-(\pm 613 , 0),
-(\pm 576 , 0),
-(\pm 100 , 0),
-(-3293 , 0)\}
-$
-(Figure~\ref{picture_15.png}).
-
-$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 5141$,
-
-which gives $d(m, 2m + 13) \leq 5141$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_15.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 5141}
-\label{picture_15.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
-(\pm 3293 , 0),
-(\pm 1848 , 0),
-(\pm 1376 , 0),
-(\pm 1036 , 0),
-(\pm 899 , 0),
-$
-
-$
-(\pm 849 , 0),
-(\pm 613 , 0),
-(\pm 576 , 0),
-(\pm 100 , 0)\}
-$
-(Figure~\ref{picture_16.png}).
-
-$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 6586$,
-
-which gives $d(m, 2m + 14) \leq 6586$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_16.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 20 and diameter 6586}
-\label{picture_16.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
-(\pm 5544 , 0),
-(\pm 4128 , 0),
-(\pm 3108 , 0),
-(\pm 2697 , 0),
-$
-
-$
-(\pm 2547 , 0),
-(\pm 1839 , 0),
-(\pm 1728 , 0),
-(\pm 904 , 0),
-(\pm 300 , 0),
-(-6148 , 0)\}
-$
-(Figure~\ref{picture_17.png}).
-
-$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 11692$,
-
-which gives $d(m, 2m + 15) \leq 11692$.
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
-(\pm 6148 , 0),
-(\pm 5544 , 0),
-(\pm 4128 , 0),
-(\pm 3108 , 0),
-$
-
-$
-(\pm 2697 , 0),
-(\pm 2547 , 0),
-(\pm 1839 , 0),
-(\pm 1728 , 0),
-(\pm 904 , 0),
-(\pm 300 , 0)\}
-$
-(Figure~\ref{picture_18.png}).
-
-$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 12296$,
-
-which gives $d(m, 2m + 16) \leq 12296$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.75\linewidth]{picture_17.png}}
-\parbox{0.75\linewidth}{\caption{IPS of cardinality 21 and diameter 11692}
-\label{picture_17.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
-(\pm 6148 , 0),
-(\pm 5544 , 0),
-(\pm 4128 , 0),
-(\pm 3108 , 0),
-$
-
-$
-(\pm 2697  , 0),
-(\pm 2547 , 0),
-(\pm 1839 , 0),
-(\pm 1728 , 0),
-(\pm 904 , 0),
-(\pm 300 , 0),
-(-9879 , 0)\}
-$
-(Figure~\ref{picture_19.png}).
-
-$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 16027$,
-
-which gives $d(m, 2m + 17) \leq 16027$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.75\linewidth]{picture_18.png}}
-\parbox{0.75\linewidth}{\caption{IPS of cardinality 22 and diameter 12296}
-\label{picture_18.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
-(\pm 9879 , 0),
-(\pm 6148 , 0),
-(\pm 5544 , 0),
-(\pm 4128 , 0),
-$
-
-$
-(\pm 3108 , 0),
-(\pm 2697 , 0),
-(\pm 2547 , 0),
-(\pm 1839 , 0),
-(\pm 1728 , 0),
-(\pm 904 , 0),
-(\pm 300 , 0)\}
-$
-(Figure~\ref{picture_20.png}).
-
-$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 19758$,
-
-which gives $d(m, 2m + 18) \leq 19758$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.85\linewidth]{picture_19.png}}
-\parbox{0.85\linewidth}{\caption{IPS of cardinality 23 and diameter 16027}
-\label{picture_19.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 5244, 360),
-(\pm 12296 , 0),
-(\pm 11088 , 0),
-(\pm 8579 , 0),
-(\pm 8256 , 0),
-$
-
-$
-(\pm 6216 , 0),
-(\pm 5394 , 0),
-(\pm 5094 , 0),
-(\pm 3678 , 0),
-(\pm 3456 , 0),
-(\pm 1808 , 0),
-(\pm 600 , 0),
-$
-
-$
-(-19758 , 0)\}
-$
-(Figure~\ref{picture_21.png}).
-
-$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P})} = 32054$,
-
-which gives $d(m, 2m + 19) \leq 32054$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.85\linewidth]{picture_20.png}}
-\parbox{0.85\linewidth}{\caption{IPS of cardinality 24 and diameter 19758}
-\label{picture_20.png}}
-\end{figure}
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_21.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 25 and diameter 32054}
-\label{picture_21.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 5244, 360),
-(\pm 19758 , 0),
-(\pm 12296 , 0),
-(\pm 11088 , 0),
-(\pm 8579 , 0),
-$
-
-$
-(\pm 8256 , 0),
-(\pm 6216 , 0),
-(\pm 5394 , 0),
-(\pm 5094 , 0),
-(\pm 3678 , 0),
-(\pm 3456 , 0),
-(\pm 1808 , 0),
-$
-
-$
-(\pm 600 , 0)\}
-$
-(Figure~\ref{picture_22.png}).
-
-$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P})} = 39516$,
-
-which gives $d(m, 2m + 20) \leq 39516$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_22.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 26 and diameter 39516}
-\label{picture_22.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
-(\pm 116058 , 0),
-(\pm 86072 , 0),
-(\pm 77616 , 0),
-(\pm 60053 , 0),
-$
-
-$
-(\pm 57792 , 0),
-(\pm 43512 , 0),
-(\pm 37758 , 0),
-(\pm 35658 , 0),
-(\pm 25746 , 0),
-(\pm 24192 , 0),
-$
-
-$
-(\pm 12656 , 0),
-(\pm 4200 , 0),
-(-138306 , 0)\}
-$
-(Figure~\ref{picture_23.png}).
-
-$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 254364$,
-
-which gives $d(m, 2m + 21) \leq 254364$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_23.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 27 and diameter 254364}
-\label{picture_23.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
-(\pm 138306 , 0),
-(\pm 116058 , 0),
-(\pm 86072 , 0),
-(\pm 77616 , 0),
-$
-
-$
-(\pm 60053 , 0),
-(\pm 57792 , 0),
-(\pm 43512 , 0),
-(\pm 37758 , 0),
-(\pm 35658 , 0),
-(\pm 25746 , 0),
-$
-
-$
-(\pm 24192 , 0),
-(\pm 12656 , 0),
-(\pm 4200 , 0)\}
-$
-(Figure~\ref{picture_24.png}).
-
-$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 276612$,
-
-which gives $d(m, 2m + 22) \leq 276612$.
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
-(\pm 138306 , 0),
-(\pm 116058 , 0),
-(\pm 86072 , 0),
-(\pm 77616 , 0),
-$
-
-$
-(\pm 60053 , 0),
-(\pm 57792 , 0),
-(\pm 43512 , 0),
-(\pm 37758 , 0),
-(\pm 35658 , 0),
-(\pm 25746 , 0),
-$
-
-$
-(\pm 24192 , 0),
-(\pm 12656 , 0),
-(\pm 4200 , 0),
-(-199863 , 0)\}
-$
-(Figure~\ref{picture_25.png}).
-
-$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 338169$,
-
-which gives $d(m, 2m + 23) \leq 338169$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_24.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 28 and diameter 276612}
-\label{picture_24.png}}
-\end{figure}
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_25.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 29 and diameter 338169}
-\label{picture_25.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
-(\pm 199863 , 0),
-(\pm 138306 , 0),
-(\pm 116058 , 0),
-(\pm 86072 , 0),
-$
-
-$
-(\pm 77616 , 0),
-(\pm 60053 , 0),
-(\pm 57792 , 0),
-(\pm 43512 , 0),
-(\pm 37758 , 0),
-(\pm 35658 , 0),
-$
-
-$
-(\pm 25746 , 0),
-(\pm 24192 , 0),
-(\pm 12656 , 0),
-(\pm 4200 , 0)\}
-$
-(Figure~\ref{picture_26.png}).
-
-$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 399726$,
-
-which gives $d(m, 2m + 24) \leq 399726$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_26.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 30 and diameter 399726}
-\label{picture_26.png}}
-\end{figure}
-
-\end{itemize}
-
-\begin{remark}
-For all the examples above, maximum in the right part of (\ref{formula1})
-evaluates to the diameter of the system.
-The following example shows that the maximum can be evaluated to $w$:
-\end{remark}
-
-\begin{figure}[h!]
-	\includegraphics[width=.48\linewidth]{picture_10.png}
-	\hfill
-	\includegraphics[width=.48\linewidth]{picture.png}
-	\\
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 8 and diameter 162}
-	\label{picture_10.png}}
-	\hfill
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 8 and diameter 409}
-	\label{picture.png}}
-\end{figure}
-
-\begin{itemize}
-\setlength{\itemsep}{-1mm}
-
-\item
-$\mathcal{P}=\sqrt{2}/{1} * \{ (\pm 76, 60),
-(\pm 62 , 0),
-(\pm 29 , 0),
-(\pm 6 , 0)\}
-$
-(Figure~\ref{picture_10.png}).
-
-$f = 152$, $v = 114$, $w = 190$, $\operatorname{diam(\mathcal{P})} = 162$.
-
-We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
-is worse that known one (\hyperlink{d2}{2})).
-
-\end{itemize}
-
-Guy gives \cite[D 20]{guy2013unsolved} a system of $8$ points located
-on two parallel lines. The points of the system form a rectangle
-(Figure~\ref{rectangle.pdf}):
-
-\begin{figure}[htbp]
-	\includegraphics[width=.6\linewidth]{rectangle.pdf}
-	\hfill
-	\includegraphics[width=.5\linewidth]{right_angle.pdf}
-	\\
-	\parbox{.65\linewidth}{\caption{Distances}
-	\label{rectangle.pdf}}
-	\hfill
-	\parbox{.5\linewidth}{\caption{Counterexample}
-	\label{right_angle.pdf}}
-\end{figure}
-
-\begin{itemize}
-\setlength{\itemsep}{-1mm}
-
-\item
-$\mathcal{P}=\sqrt{1}/{1} * \{ (0, \pm 60),
-(182 , \pm 60),
-(209 , \pm 60),
-(391 , \pm 60)\}
-$
-(Figure~\ref{picture.png}),
-
-where $n = 8$, $\operatorname{diam(\mathcal{P})} = 409$. Using the ``blowing up''
-construction, we obtain
-\begin{equation}\label{result1}
-d(m, 4m) \leq 409
-\end{equation}
-Figure~\ref{picture_11.pdf} shows an example for $m = 3$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_11.pdf}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 12 and diameter 409}
-\label{picture_11.pdf}}
-\end{figure}
-
-\end{itemize}
-
-\section{Bounds based on pyramid sets}
-
-\begin{theorem}
-Let $\mathcal{P}$ be a planar integral point set consisting of $k$
-points on the line $l_{1}$ and $k$ points on the line $l_{2}$. Besides, these
-points are symmetric with respect to one of the bisectors of the angles
-formed by the intersection of lines $l_{1}$ and $l_{2}$, then
-
-\begin{equation}\label{formula2}
-d(m, (k - \alpha)m + \alpha) \leq \operatorname{diam(\mathcal{P})},
-\end{equation}
-
-where
-\begin{equation*}
-\alpha =
-\begin{cases}
-1, \text{when the intersection point} \in \mathcal{P}, \\
-0, \text{when the intersection point} \notin \mathcal{P}.
-\end{cases}
-\end{equation*}
-
-\end{theorem}
-
-\begin{remark}
-The angles can be acute or obtuse. Figure~\ref{right_angle.pdf} shows that the intersection angle of the lines $l_{1}$ and $l_{2}$ cannot be equal to ${\pi}/2$.
-\end{remark}
-
-Below we give the planar integral point sets and the corresponding
-estimates of the function $d(m, n)$ for $n = 3m + 1$ and $n = 4m + 1$.
-
-\begin{figure}[htbp]
-	\includegraphics[width=.48\linewidth]{picture_56.png}
-	\hfill
-	\includegraphics[width=.48\linewidth]{picture_1260.png}
-	\\
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 7 and diameter 56}
-	\label{picture_56.png}}
-	\hfill
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 9 and diameter 1260}
-	\label{picture_1260.png}}
-\end{figure}
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_12.pdf}}
+%\center{\includegraphics[width=1\linewidth]{picture_12.pdf}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 10 and diameter 56}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_1260_R3.pdf}}
+%\center{\includegraphics[width=1\linewidth]{picture_1260_R3.pdf}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 13 and diameter 1260}
@@
 All the given PIPSs were obtained through a combination of computer search an intuition of the authors;
-so, the further search may lead to better bounds employing the same consructions.
 
@@
 on the other hand, we still have not found any rails PIPSs with 4 and 5 points on the lines.
-
-As for today, we have found pyramid PIPS of cardinality at most 9.
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Drafts of the intro

@@
 
+The main purpose of this work is to provide examples of planar integral point sets
+that may give the clue for development of further classification.
+
+It's notable that any of the examples that are discussed below
+is located on a union of at most three straight lines.
+In the most pictures these lines are shown as solid lines.
+For classification of planar IPS that are located on a union of two straight lines,
+we refer the reader to~\cite{avdeev2019particular}.
+
+There are some examples of planar IPS that are not contained in the union of any three straight lines:
+for examples, these are heptagons presented in~\cite{kreisel2008heptagon} and 7-clusters from~\cite{kurz2013constructing}.
+However, we have to keep in mind that the circular inversion under certain conditions
+translates an integral point set into an integral point set
+(although sometimes additional dilation is necessary).
+On the other hand, the circular inversion may translate a straight line into a circle and vice versa.
+
+Thus, we can consider \emph{generalized circles}, that are circles or straight lines;
+obviously, in that point of view all the examples from~\cite{kreisel2008heptagon} and~\cite{kurz2013constructing}
+are located on a union of three generalized circles,
+because each seven points are located on a union of a circle and two straight lines.
+
 \section{Rails sets}
@@
 \begin{figure}[h!]
-%\center{\includegraphics[width=1\linewidth]{picture_12.pdf}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 10 and diameter 56}
-\label{picture_12.pdf}}
+center{\includegraphics[width=1\linewidth]{./img/42_symm.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 42 and diameter 2473117504}
+\label{42_symm.png}}
 \end{figure}
@@
 \begin{figure}[h!]
-%\center{\includegraphics[width=1\linewidth]{picture_1260_R3.pdf}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 13 and diameter 1260}
-\label{picture_1260_R3.pdf}}
+center{\includegraphics[width=1\linewidth]{./img/46_symm.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 46 and diameter 3118278592}
+\label{46_symm.png}}
+\end{figure}
+
+\begin{figure}[h!]
+center{\includegraphics[width=1\linewidth]{./img/48_symm.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 48 and diameter 71720407616}
+\label{48_symm.png}}
 \end{figure}
@@
 \item
-$\mathcal{P}=\sqrt{3}/{2} * \{ (\pm 56, 0),
-(14, 0),
-(-34, 0),
-(-10, 24),
-(-21 , 35),
-(35, -21)\}
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 345596160; 0);
+(\pm219513840 -15069600);
+(\pm260201760; 0);
+(\pm225792840; 0);
+(\pm213234840; 0);
+(\pm153961080; 0);
+(\pm144668160; 0
+(\pm25116000; 0);
+(\pm694026840; 0);
+(\pm514710560; 0);
+(\pm359116940; 0);
+(\pm13423904; 0);
+(\pm75682880; 0);
+(\pm464143680; 0);
+(\pm827069880; 0);
+(\pm92144325; 0);
+(\pm1195180740; 0);
+(\pm1236558752; 0);
+(\pm44590560; 0);
+(\pm339925740; 0);
+(\pm117312468; 0)\}
 $
-(Figure~\ref{picture_56.png}),
+(Figure~\ref{42_symm.png}),
 
-where $n = 7$, $\operatorname{diam(\mathcal{P})} = 56$. Using the ``blowing up''
+where $n = 42$, $\operatorname{diam(\mathcal{P})} = 2473117504$. Using the ``blowing up''
 construction, we obtain
@@
 \item
-$\mathcal{P}=\sqrt{39}/{8} * \{ (\pm 5040, 0),
-(1911, 315),
-(2352, 0),
-(944, 0),
-(336, 0),
-(2940, -420),
-(2044, 220),
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm435751680; 0);
+(\pm276778320; -19000800);
+(\pm328080480; 0);
+(\pm284695320; 0);
+(\pm268861320; 0);
+(\pm194124840; 0);
+(\pm182407680; 0);
+(\pm31668000; 0);
+(\pm1559139296; 0);
+(\pm1506967020; 0);
+(\pm1042827240; 0);
+(\pm875077320; 0);
+(\pm648982880; 0);
+(\pm585224640; 0);
+(\pm452799620; 0);
+(\pm95426240; 0);
+(\pm16925792; 0);
+(\pm116181975; 0);
+(\pm428602020; 0);
+(\pm56222880; 0);
+(\pm769560480; 0);
+(\pm626458560; 0);
+(\pm130761918; 0)\}
 $
+(Figure~\ref{46_symm.png}),
 
-$
-(735, 1155)\}
-$
-(Figure~\ref{picture_1260.png}),
-
-where $n = 9$, $\operatorname{diam(\mathcal{P})} = 1260$. Using the ``blowing up''
+where $n = 46$, $\operatorname{diam(\mathcal{P})} = 3118278592$. Using the ``blowing up''
 construction, we obtain
@@
 Figure~\ref{picture_1260_R3.pdf} shows an example for $m = 3$.
+
+\item
+$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm10022288640; 0);
+( \pm6365901360 ; -437018400); 
+( \pm7545851040 ; 0); 
+( \pm6547992360 ; 0); 
+( \pm6183810360 ; 0); 
+( \pm4464871320 ; 0); 
+( \pm4195376640 ; 0); 
+( \pm728364000 ; 0); 
+( \pm35860203808 ; 0); 
+( \pm34660241460 ; 0); 
+( \pm23985026520 ; 0); 
+( \pm20126778360 ; 0); 
+( \pm14926606240 ; 0); 
+( \pm13460166720 ; 0); 
+( \pm10414391260 ; 0); 
+( \pm2194803520 ; 0); 
+( \pm389293216 ; 0); 
+( \pm2672185425 ; 0); 
+( \pm9857846460 ; 0); 
+( \pm1293126240 ; 0); 
+( \pm17699891040 ; 0); 
+( \pm14408546880 ; 0); 
+( \pm3007524114 ; 0); 
+( \pm3402061572 ; 0)\}
+$
+(Figure~\ref{48_symm.png}),
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Plan for rails
[AMZAE] Remove TODO notes
[AMZAE] Just EOLs
[AMZAE] Drafts for two more sections

@@
 
+Two IPSs below have been obtained by dilating~\cite[Fig. 34]{avdeev2019particular} by $23$ and $29$  resp.;
+third one has been constructed by dilation and merge.
+
 \begin{figure}[h!]
@@
 
+\section{Example of sets with many common points that canot be merged}
+
+TODO: three ''romantic'' pictures in a row.
+
+TODO: specify the irrational distance:
+
+Note that distance between the non-adoptable points is
+\begin{equation}
+	\sqrt{\left(\frac{}{} - \frac{}{}\right)^2 + \left(\frac{}{} - \frac{}{}\right)^2}
+\end{equation}
+
+\section{Integral point sets with two axes of symmetry}
+
+TODO:
+
+TODO: 11_14450_1_8693fad981c82dc92d7c9365bcb9945e
+
+The set on Fig. TODO has been obtained from the set on Fig. TODO by dilation by 5.
+
+The set on Fig. TODO has been obtained from the set on Fig. TODO by dilation by 42 (or less? TODO).
+
+\section{Other examples}
+
+TODO: 8_2535_1_d680ff2d172ebe40250a54ca9419af34
+
 \section{Final remarks}
-All the given PIPSs were obtained through a combination of computer search an intuition of the authors;
-
-There is still no general construction for a rails or scissors PIPS of given cardinality.
-For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality with 2 points on one line
-and all the rest on the other;
-on the other hand, we still have not found any rails PIPSs with 4 and 5 points on the lines.
+All the given planar integral point sets were obtained through a combination of computer search an intuition of the authors;
 
@@
 \printbibliography
-%\bibliography{literature}
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Remove old intro

@@
 
-\begin{dfn}
-	The function $d(m, n)$ is the minimum possible diameter of
-	the integral point set $\mathcal{P}$ of $n$ points in
-	$m$-dimensional Euclidean space $\mathbb{R}^{m}$.
-\end{dfn}
-
-%For a bit more sophisticated configurations, another functions are presented~\cite{kurz2008minimum,kreisel2008heptagon}.
-%
-%\begin{dfn}
-%	The function $\overline{d}(m, n)$ is the minimum possible diameter of
-%	the integral point set $\mathcal{P}$ of $n$ points in
-%	$m$-dimensional Euclidean space $\mathbb{R}^{m}$
-%	such that no $m+1$ points are located on an $(m-1)$-dimensional hyperplane.
-%\end{dfn}
-
-%\begin{dfn}
-%	The function $\dot{d}(m, n)$ is the minimum possible diameter of
-%	the integral point set $\mathcal{P}$ of $n$ points in
-%	$m$-dimensional Euclidean space $\mathbb{R}^{m}$
-%	such that no three points are located on a straight line
-%	and no four points are located on a circle.
-%\end{dfn}
-
-For the list of known bounds for $d(m, n)$,
-we refer the reader to~\cite[Theorem 1]{kurz2008bounds} or to~\cite{our-vmmsh-2018}.
-We will discuss the following estimations presented at~\cite{kurz2008bounds}:
-\begin{align}
-	d(m, 2m + 1) \leq 8\\
-	d(m, 2m + 2) \leq 13 \hypertarget{d2}\\
-	d(m, 3m) \leq 109
-\end{align}
-and the following theorem \cite[Theorem 2.1]{kurz2008bounds}.
-
-\begin{theorem}
-	Let $\mathcal{P}$ be a planar integral point set consisting of
-	$n - 2$ points on line $l_1$ and two points $P_{1}$ and $P_{2}$ on a
-	parallel line $l_2$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
-	exist positive integers $v$, $w$ with $f^{2} + v^{2}
-	= w^{2}$ and $v < 2r$, where $|P_{1}P_{2}| = f$,
-	then
-
-	\begin{equation}\label{formula1}
-		d(m, n + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
-	\end{equation}
-
-\end{theorem}
-
-Firstly, we discuss the classification of planar integral points sets;
-then, we present some bounds for $d(m,n)$ based on planar integral point sets of particular types
-and provide some general constructions for such bounds.
-
-\section{Classification of planar integral point sets}
-
-\subsection{Integral point sets situated in two straight lines}
-
-\begin{dfn}
-	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
-	a \textit{facher} set.
-\end{dfn}
-Facher sets are very dominating examples of planar integral pont sets.
-In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
-For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.
-
-For non-facher integral point sets situated in two straight lines,
-we can easily distinct the following three cases:
-
-\begin{dfn}
-	A planar integral point sets situated in two parallel straight lines
-	is called a \textit{rails} set.
-\end{dfn}
-
-Among the rails sets, sets with 2 points on one line and all the other on another line dominate.
-
-\begin{dfn}
-	A planar integral point sets situated in two perpendicular straight lines
-	is called a \textit{cross} set.
-\end{dfn}
-Every cross set has characteristic 1;
-in~\cite{antonov2008maximal}, cross sets with only 2 points out of one of the lines are called \textit{crabs}.
-
-\begin{dfn}
-	A planar integral point sets situated in two straight lines
-	that are not parallel nor perpendicular,
-	is called a \textit{sciccors} set.
-\end{dfn}
-
-There is an important subclassof scissors sets.
-
-\begin{dfn}
-	A scissors set with an axis of symmetry,
-	which is the angle bisector for the straight lines,
-	is called a \textit{pyramid} set.
-\end{dfn}
-
-%TODO: rephrase?
-
-%TODO: can a scissors set have another axis of symmetry?
-
-\subsection{Other integral point sets}
-
-\begin{dfn}
-	A planar integral point sets that is situated on a circle is called a \textit{circular}
-	point set.
-\end{dfn}
-
-Circular sets are very important examples
-of integral point sets~\cite{harborth1993upper,piepmeyer1996maximum,bat2018number}.
-
-\begin{dfn}
-	A planar integral point sets that is situated on the conjunction of a circle with its center,
-	is called a \textit{centered-circular} point set.
-\end{dfn}
-
-These six classes dominate among all the known planar integral point sets;
-however, some sophisticated constructions are also known.
-
-%TODO: pictures!
+While the minimal possible diameter for planar integral point sets of given cardinality was being counted,
+it was noticed~\cite{kurz2008minimum} that such diameter is attained at sets with many points on a straight line;
+for some estimations on this tendency, we also refer the reader to~\cite{solymosi2003note}.
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Remark on rails set

@@
 \section{Rails sets}
+
+\begin{dfn}
+	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
+	a \textit{facher} set.
+\end{dfn}
+Facher sets are very dominating examples of planar integral pont sets.
+In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
+For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.
+
+\begin{dfn}[\cite{avdeev2019particular}]
+	A non-facher planar integral point sets situated in two parallel straight lines
+	is called a \textit{rails} set.
+\end{dfn}
+
+Among the rails sets, sets with 2 points on one line and all the other on another line dominate.
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Remove the remnants of estimates

@@
 
-where $n = 42$, $\operatorname{diam(\mathcal{P})} = 2473117504$. Using the ``blowing up''
-construction, we obtain
-\begin{equation}\label{result2}
-d(m, 3m + 1) \leq 56
-\end{equation}
-The resulting estimate improves the estimate for $d(m, 3m)$, which is presented
-in \cite{kemnitz1988punktmengen}. Figure~\ref{picture_12.pdf} shows an example
-for $m = 3$.
+where $n = 42$, $\operatorname{diam(\mathcal{P})} = 2473117504$.
 
@@
 
-where $n = 46$, $\operatorname{diam(\mathcal{P})} = 3118278592$. Using the ``blowing up''
-construction, we obtain
-\begin{equation}\label{result3}
-d(m, 4m + 1) \leq 1260
-\end{equation}
-Figure~\ref{picture_1260_R3.pdf} shows an example for $m = 3$.
+where $n = 46$, $\operatorname{diam(\mathcal{P})} = 3118278592$.
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] - WIP

@@
 
-\section{Example of sets with many common points that canot be merged}
+Taking
+
+\section{Example of sets with many common points that cannot be merged}
 
@@
 
-TODO: 11_14450_1_8693fad981c82dc92d7c9365bcb9945e
+TODO: 11\_14450\_1\_8693fad981c82dc92d7c9365bcb9945e
 
@@
 
+\section{Arrow-like integral point sets with one axis of symmetry}
+
+TODO:
+
+на каждое такое множество: исходный вариант, все этапы растращаивания с коэффициентом растаращивания и пояснением, кто из кого.
+
 \section{Other examples}
 
-TODO: 8_2535_1_d680ff2d172ebe40250a54ca9419af34
+TODO: 8\_2535\_1\_d680ff2d172ebe40250a54ca9419af34
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Intro is almost done

@@
 
-While the minimal possible diameter for planar integral point sets of given cardinality was being counted,
+While the minimal possible diameter for planar integral point sets of given cardinality was being computed,
 it was noticed~\cite{kurz2008minimum} that such diameter is attained at sets with many points on a straight line;
 for some estimations on this tendency, we also refer the reader to~\cite{solymosi2003note}.
+Thus, the following classification was introduced:
+\begin{dfn}
+	A planar integral point set $M$ is said to be in \emph{semi-general position}
+	if no three points of $M$ are located in a straight line.
+\end{dfn}
+
+The most dominating examples of PIPS in semi-general position are circular sets.
+\begin{dfn}
+	A planar integral point sets that is situated on a circle is said to be a \textit{circular}
+	point set.
+\end{dfn}
+
+So, the following constraint appeared.
+\begin{dfn}
+	A planar integral point set $M$ is said to be in \emph{general position}
+	if no three points of $M$ are located in a straight line
+	and no four points of $M$ are located in a circle.
+\end{dfn}
+
+Planar integral point sets in general position are very difficult to find;
+the first known examples were presented in~\cite{kreisel2008heptagon}.
+As for now, there is no known example of PIPS of cardinality 8 in general position.
 
@@
 is located on a union of at most three straight lines.
-In the most pictures these lines are shown as solid lines.
+%In the most pictures these lines are shown as solid lines.
 For classification of planar IPS that are located on a union of two straight lines,
@@
 On the other hand, the circular inversion may translate a straight line into a circle and vice versa.
-
 Thus, we can consider \emph{generalized circles}, that are circles or straight lines;

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Cosmetics edits to the intro

@@
 
-How do we characterize an integral point set?
-For example, we can count the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
-and is further called the cardinality;
-alternatively, we can naturally% TODO: naturally/essentially? cf. arxiv/1
+How do we describe an integral point set?
+For example, with the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
+and is said to be the \emph{cardinality} of the IPS.
+Furthermore, we can naturally % TODO: naturally/essentially? cf. arxiv/1
 define the diameter of a finite point set.
@@
 \begin{dfn}
-	A diameter of the integral point set $\mathcal{P}$ is defined as
+	The diameter of the integral point set $\mathcal{P}$ is defined as
 	\begin{equation}
@@
 Every integral point set also has a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.
+Characteristic of an IPS does not change if the set is moved, dilated ot flipped;
+moreover, even addition or deletion of a point does not change the characteristic of IPS.
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Abstract

@@
 \maketitle
+
+\paragraph{Abstract.}
+	A planar integral point set is a set of non-collinear points in plane
+	such that for any pair of the  points the Euclidean distance
+	between the points is integral.
+	We discuss the classification of planar integral point sets
+	and provide examples of sets that are not covered by the existent classification.
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Add indices to point sets
[AMZAE] Formulate the conjecture about infinite rails

@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 345596160; 0);
+$\mathcal{P_42}=\sqrt{154}/{1} * \{ (\pm 345596160; 0);
 (\pm219513840 -15069600);
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm435751680; 0);
+$\mathcal{P_46}=\sqrt{154}/{1} * \{ (\pm435751680; 0);
 (\pm276778320; -19000800);
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm10022288640; 0);
+$\mathcal{P_48}=\sqrt{154}/{1} * \{ (\pm10022288640; 0);
 ( \pm6365901360 ; -437018400);
@@
 
-Taking
+Taking the examples into consideration,
+we can conjecture that there is an infinite set with rational distances
+that contains $\mathcal{P_48}$.
+(However, it is known that\cite{solymosi2010question}
+if a rational set S has infinitely many points on a line or on a circle,
+then all but 4 resp. 3 points of S are on the line or on the circle.)
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
8ки которые не могут быть вместе
Пробелы и переносы
Системы с 2мя осями симметрии

@@
 \item
-$\mathcal{P_42}=\sqrt{154}/{1} * \{ (\pm 345596160; 0);
+$\mathcal{P_42}=\sqrt{154}/{1} * \{
+(\pm 345596160; 0);
 (\pm219513840 -15069600);
-(\pm260201760; 0);
+(\pm260201760; 0);\\
 (\pm225792840; 0);
@@
 (\pm153961080; 0);
-(\pm144668160; 0
-(\pm25116000; 0);
+(\pm144668160; 0);
+(\pm25116000; 0);\\
 (\pm694026840; 0);
@@
 (\pm13423904; 0);
-(\pm75682880; 0);
+(\pm75682880; 0);\\
 (\pm464143680; 0);
@@
 (\pm1195180740; 0);
-(\pm1236558752; 0);
+(\pm1236558752; 0);\\
 (\pm44590560; 0);
@@
 \item
-$\mathcal{P_46}=\sqrt{154}/{1} * \{ (\pm435751680; 0);
+$\mathcal{P_46}=\sqrt{154}/{1} * \{
+(\pm435751680; 0);
 (\pm276778320; -19000800);
 (\pm328080480; 0);
-(\pm284695320; 0);
+(\pm284695320; 0);\\
 (\pm268861320; 0);
@@
 (\pm31668000; 0);
-(\pm1559139296; 0);
+(\pm1559139296; 0);\\
 (\pm1506967020; 0);
@@
 (\pm648982880; 0);
-(\pm585224640; 0);
+(\pm585224640; 0);\\
 (\pm452799620; 0);
@@
 (\pm116181975; 0);
-(\pm428602020; 0);
+(\pm428602020; 0);\\
 (\pm56222880; 0);
@@
 \item
-$\mathcal{P_48}=\sqrt{154}/{1} * \{ (\pm10022288640; 0);
+$\mathcal{P_48}=\sqrt{154}/{1} * \{
+( \pm10022288640; 0);
 ( \pm6365901360 ; -437018400);
-( \pm7545851040 ; 0);
+( \pm7545851040 ; 0);\\
 ( \pm6547992360 ; 0);
@@
 ( \pm4195376640 ; 0);
-( \pm728364000 ; 0);
+( \pm728364000 ; 0);\\
 ( \pm35860203808 ; 0);
@@
 ( \pm23985026520 ; 0);
-( \pm20126778360 ; 0);
+( \pm20126778360 ; 0);\\
 ( \pm14926606240 ; 0);
@@
 ( \pm2194803520 ; 0);
-( \pm389293216 ; 0);
+( \pm389293216 ; 0);\\
 ( \pm2672185425 ; 0);
@@
 ( \pm17699891040 ; 0);
-( \pm14408546880 ; 0);
+( \pm14408546880 ; 0);\\
 ( \pm3007524114 ; 0);
@@
 
+\begin{figure}[h!]
+	\begin{minipage}[h]{0.32\linewidth}
+		\begin{center}
+			\includegraphics[width=1\linewidth]{./img/8_2520_143_symm1.png}\\ a)
+		\end{center}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[h]{0.32\linewidth}
+		\begin{center}
+			\includegraphics[width=1\linewidth]{./img/8_2520_143_symm2.png}\\ b)
+		\end{center}
+	\end{minipage}
+	\begin{minipage}[h]{0.32\linewidth}
+		\begin{center}
+			\includegraphics[width=1\linewidth]{./img/8_2520_143_4a27_other.png}\\ c)
+		\end{center}
+	\end{minipage}
+	\hfill
+	\caption{IPSs with cardinality 8 and diameter 2520 with many common points}
+	\label{8_with_many_common}
+\end{figure}
+
+\begin{itemize}
+\item
+$\mathcal{P}=\sqrt{143}/2*\{
+( \pm1620 ; 0);
+( \pm1920 ; 300);
+( \pm735 ; 75);
+( \pm340 ; 0);
+\}$
+
+\item
+$\mathcal{P}=\sqrt{143}/2*\{
+( \pm1620 ; 0);
+( \pm1920 ; 300);
+( \pm735 ; 75);
+( \pm1767 ; 147);
+\}$
+
+\item
+$\mathcal{P}=
+\sqrt{143}/2*\{
+( \pm1620 ; 0);
+( \pm1920 ; 300);
+( \pm735 ; 75);
+( -340 ; 0);
+( 1767 ; 147);
+\}$
+
+\end{itemize}
 TODO: specify the irrational distance:
@@
 \begin{equation}
-	\sqrt{\left(\frac{}{} - \frac{}{}\right)^2 + \left(\frac{}{} - \frac{}{}\right)^2}
+	\sqrt{\left(\frac{1767}{2} - \frac{340}{2}\right)^2 + \left(\frac{147}{2}\right)^2\cdot143} = 2\sqrt{320401}
 \end{equation}
+
+Where 320401 is prime.
 
@@
 
-TODO:
+\begin{figure}[h!]
+center{\includegraphics[width=1\linewidth]{./img/15_61960444_1_other.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 61960444}
+\label{15_61960444_1.png}}
+\end{figure}
 
-TODO: 11\_14450\_1\_8693fad981c82dc92d7c9365bcb9945e
+\begin{figure}[h!]
+center{\includegraphics[width=1\linewidth]{./img/15_45970652_1_other.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 45970652}
+\label{15_45970652_1.png}}
+\end{figure}
+
+\begin{itemize}
+\item
+$\mathcal{P}=\sqrt{1}/1*\{
+( \pm9406950 ; 0);
+( \pm7063350 ; 0);
+( 2962050 ; \pm3437280);
+( 0 ; 0);
+( -2962050 ; \pm3437280);\\
+( \pm2188662 ; 0);
+( \pm30980222 ; 0);
+( \pm2914854 ; 0);
+\}$
+
+\item
+$\mathcal{P}=\sqrt{1}/1*\{
+( \pm6979350 ; 0);
+( \pm5240550 ; 0);
+( 2197650 ; \pm2550240);
+( 0 ; 0);
+( -2197650 ; \pm2550240);\\
+( \pm1623846 ; 0);
+( \pm22985326 ; 0);
+( \pm2476152 ; 0);
+\}$
+\end{itemize}
+
+%TODO:
+
+%TODO: 11_14450_1_8693fad981c82dc92d7c9365bcb9945e
 
@@
 
-TODO: 8\_2535\_1\_d680ff2d172ebe40250a54ca9419af34
+%TODO: 8_2535_1_d680ff2d172ebe40250a54ca9419af34
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
Cosmetic fixes

@@
 that contains $\mathcal{P_48}$.
-(However, it is known that\cite{solymosi2010question}
-if a rational set S has infinitely many points on a line or on a circle,
-then all but 4 resp. 3 points of S are on the line or on the circle.)
+(However, it is known~\cite{solymosi2010question} that
+if a rational set $S$ has infinitely many points on a line or on a circle,
+then all but 4 resp. 3 points of $S$ are on the line or on the circle.)
 
@@
 	\sqrt{\left(\frac{1767}{2} - \frac{340}{2}\right)^2 + \left(\frac{147}{2}\right)^2\cdot143} = 2\sqrt{320401}
+	.
 \end{equation}
-
-Where 320401 is prime.
+It's notable that 320401 is a prime number.
 
@@
 \section{Final remarks}
-All the given planar integral point sets were obtained through a combination of computer search an intuition of the authors;
-
+All the given planar integral point sets were obtained through a combination of computer search an intuition of the authors.
 The source code can be obtained at https://gitlab.com/Nickkolok/ips-algo

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
Cosmetic fixes
[AMZAE] Remove TODO notes

@@
 
-TODO: three ''romantic'' pictures in a row.
-
 \begin{figure}[h!]
@@
 \end{itemize}
-TODO: specify the irrational distance:
 
@@
 \section{Final remarks}
-All the given planar integral point sets were obtained through a combination of computer search an intuition of the authors.
+All the given planar integral point sets were obtained through a combination of computer search and intuition of the authors.
 The source code can be obtained at https://gitlab.com/Nickkolok/ips-algo

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] One more sophisticated example

@@
 
-%TODO: 8_2535_1_d680ff2d172ebe40250a54ca9419af34
+Figure~\ref{8_2535_1_d680} displays an IPS with
+no axis of symmetry;
+alothough the set is of characteristic 1,
+it cannot be extended by reflecting relatively to the $x$ axis.
+Moreover, we failed to extend it by dilation and looking for extra points on the $x$ axis.
+
+%TODO: write about points which prevent the set from reflection
+
+$\mathcal{P}_8=
+\sqrt{1}/13*
+\{
+( 29575 ; 0);
+( -2366 ; -8112);
+( 12844 ; 0);
+( 10647 ; -14196);
+( 8450 ; 0);
+( 15022 ; -3696);
+( 21294 ; 0);
+( 0 ; 0);
+\}
+$
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{./img/8_2535_1_d680ff2d172ebe40250a54ca9419af34_other.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 8 and diameter 2535}
+\label{8_2535_1_d680}}
+\end{figure}
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Introduce the star notation

@@
 that may give the clue for development of further classification.
+
+For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
+$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
+which means that each abscissa is multiplied by $1/q$
+and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
+$$
+	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
+	=
+	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}
+	.
+$$
+Here $q$ is the characteristic of the PIPS;
+every PIPS can be represented in such way~\cite{our-vmmsh-2018-translit}[Theorem 4].
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Cosmetics

@@
 		\mathcal{P}}{\max} |P_{1}P_{2}|
+		.
 	\end{equation}
@@
 In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
-For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.
+%For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.
 
@@
 \begin{figure}[h!]
-center{\includegraphics[width=1\linewidth]{./img/42_symm.png}}
+\center{\includegraphics[width=1\linewidth]{./img/42_symm.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 42 and diameter 2473117504}
@@
 \begin{figure}[h!]
-center{\includegraphics[width=1\linewidth]{./img/46_symm.png}}
+\center{\includegraphics[width=1\linewidth]{./img/46_symm.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 46 and diameter 3118278592}
@@
 \begin{figure}[h!]
-center{\includegraphics[width=1\linewidth]{./img/48_symm.png}}
+\center{\includegraphics[width=1\linewidth]{./img/48_symm.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 48 and diameter 71720407616}
@@
 \item
-$\mathcal{P_42}=\sqrt{154}/{1} * \{
+Figure~\ref{42_symm.png}:
+\\
+$\mathcal{P}_{42}=\sqrt{154}/{1} * \{
 (\pm 345596160; 0);
-(\pm219513840 -15069600);
+(\pm219513840; -15069600);
 (\pm260201760; 0);\\
@@
 $
-(Figure~\ref{42_symm.png}),
-
-where $n = 42$, $\operatorname{diam(\mathcal{P})} = 2473117504$.
 
 \item
-$\mathcal{P_46}=\sqrt{154}/{1} * \{
+Figure~\ref{46_symm.png}:
+\\
+$\mathcal{P}_{46}=\sqrt{154}/{1} * \{
 (\pm435751680; 0);
@@
 $
-(Figure~\ref{46_symm.png}),
-
-where $n = 46$, $\operatorname{diam(\mathcal{P})} = 3118278592$.
 
 \item
-$\mathcal{P_48}=\sqrt{154}/{1} * \{
+Figure~\ref{48_symm.png}:
+\\
+$\mathcal{P}_{48}=\sqrt{154}/{1} * \{
 ( \pm10022288640; 0);
@@
 $
-(Figure~\ref{48_symm.png}),
 
@@
 Taking the examples into consideration,
-we can conjecture that there is an infinite set with rational distances
-that contains $\mathcal{P_48}$.
+we can conjecture that there is an infinite point set with rational distances
+that contains $\mathcal{P}_{48}$.
 (However, it is known~\cite{solymosi2010question} that
-if a rational set $S$ has infinitely many points on a line or on a circle,
+if a point set $S$ with rational distances has infinitely many points on a line or on a circle,
 then all but 4 resp. 3 points of $S$ are on the line or on the circle.)
@@
 \section{Example of sets with many common points that cannot be merged}
+
+Figure~\ref{8_with_many_common} shows an example of three PIPS of cardinality 8,
+each pair of that shares 6 or 7 points but cannot be united into another PIPS.
 
@@
 
-Note that distance between the non-adoptable points is
+The distance between the non-adoptable points is
 \begin{equation}
@@
 \begin{figure}[h!]
-center{\includegraphics[width=1\linewidth]{./img/15_61960444_1_other.png}}
+\center{\includegraphics[width=1\linewidth]{./img/15_61960444_1_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 61960444}
@@
 \begin{figure}[h!]
-center{\includegraphics[width=1\linewidth]{./img/15_45970652_1_other.png}}
+\center{\includegraphics[width=1\linewidth]{./img/15_45970652_1_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 45970652}

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Add one more peculiar example

@@
 and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
-$$
+\begin{equation}
+	\label{eq:char_lattice}
 	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
@@
 	.
-$$
+\end{equation}
 Here $q$ is the characteristic of the PIPS;
@@
 
+The set shown on Figure~\ref{8_2400_42_56f3} has an axis of symmetry, but it is $y$ axis, not $x$ axis
+and due to the fact that its characteristic is not 1,
+the set cannot be rotated by $90^\circ$ but still stay on lattice~\eqref{eq:char_lattice}:
+\begin{equation}
+	\mathcal{P}_{8y}=
+	\sqrt{42}/1*\{( \pm1200 ; 0);
+	( \pm529 ; 182);
+	( \pm814 ; 152);
+	( \pm440 ; 80)
+	\}
+\end{equation}
+
+\begin{figure}[h!]
+	\center{\includegraphics[width=1\linewidth]{./img/8_2400_42_56f33d8dc5df552595e807c438bc0f55_other.png}}
+	\parbox{1\linewidth}{\caption{IPS of cardinality 8 and diameter 2400}
+	\label{8_2400_42_56f3}}
+\end{figure}
+
 \section{Final remarks}

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Cosmetics

@@
 no axis of symmetry;
-alothough the set is of characteristic 1,
+although the set is of characteristic 1,
 it cannot be extended by reflecting relatively to the $x$ axis.
 Moreover, we failed to extend it by dilation and looking for extra points on the $x$ axis.
-
+%
 %TODO: write about points which prevent the set from reflection
-
-$\mathcal{P}_8=
+%
+\begin{multline}
+	\mathcal{P}_8=
 	\sqrt{1}/13*
 	\{
+	( 0 ; 0);
+	( 8450 ; 0);
+	( 12844 ; 0);
+	( 21294 ; 0);
 	( 29575 ; 0);
+	\\
 	( -2366 ; -8112);
-( 12844 ; 0);
 	( 10647 ; -14196);
-( 8450 ; 0);
-( 15022 ; -3696);
-( 21294 ; 0);
-( 0 ; 0);
+	( 15022 ; -3696)
 	\}
-$
-
+\end{multline}
+%
 \begin{figure}[h!]
@@
 All the given planar integral point sets were obtained through a combination of computer search and intuition of the authors.
+
 The source code can be obtained at https://gitlab.com/Nickkolok/ips-algo

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Transitional commit: prepare the first version to be submitted

@@
 
-\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
-\addbibresource{../common/notmy.bib}
-\addbibresource{../common/my.bib}
+%\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+%\addbibresource{../common/notmy.bib}
+%\addbibresource{../common/my.bib}
+%\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
 
 \usepackage{hyperref}
+\usepackage{cite}
 
@@
 
-The set on Fig. TODO has been obtained from the set on Fig. TODO by dilation by 5.
+%The set on Fig. TODO has been obtained from the set on Fig. TODO by dilation by 5.
 
-The set on Fig. TODO has been obtained from the set on Fig. TODO by dilation by 42 (or less? TODO).
+%The set on Fig. TODO has been obtained from the set on Fig. TODO by dilation by 42 (or less? TODO).
 
-\section{Arrow-like integral point sets with one axis of symmetry}
+%\section{Arrow-like integral point sets with one axis of symmetry}
 
-TODO:
+%TODO:
 
-на каждое такое множество: исходный вариант, все этапы растращаивания с коэффициентом растаращивания и пояснением, кто из кого.
+%на каждое такое множество: исходный вариант, все этапы растращаивания с коэффициентом растаращивания и пояснением, кто из кого.
 
@@
 
-\printbibliography
+%\printbibliography
+%\printbibitembibliography
+
+\begin{thebibliography}{99}
+%
+\bibitem{anning1945integral}
+\emph{Anning} \emph{N. H.}, \emph{Erdős} \emph{P.}. Integral distances //
+Bulletin of the American Mathematical Society. — 1945. — Vol. 51, no. 8. —
+Pp. 598–600. — DOI: \href {https://doi.org/10.1090/S0002-9904-1945-08407-9}
+{\nolinkurl {10.1090/S0002-9904-1945-08407-9}}.
+%
+\bibitem{erdos1945integral}
+\emph{Erdős} \emph{P.}. Integral distances // Bulletin of the American Mathematical
+Society. — 1945. — Vol. 51, no. 12. — P. 996. — DOI: \href
+{https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl
+{10.1090/S0002-9904-1945-08490-0}}.%
+\bibitem{kemnitz1988punktmengen}
+\emph{Kemnitz} \emph{A.}. Punktmengen mit ganzzahligen Abständen. — 1988.
+%
+\bibitem{kurz2005characteristic}
+\emph{Kurz} \emph{S.}. On the characteristic of integral point sets in {$\mathbb
+{E}^m$} // Australasian Journal of Combinatorics. — 2006. — Vol. 36. —
+Pp. 241–248. — arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl
+{math/0511704}}.
+%
+\bibitem{kurz2008minimum}
+\emph{Kurz} \emph{S.}, \emph{Wassermann} \emph{A.}. On the minimum diameter
+of plane integral point sets // Ars Combinatoria. — 2011. — Vol. 101. — Pp. 265–287. —
+arXiv: \href {http://arxiv.org/abs/0804.1307} {\nolinkurl {0804.1307}}.
+%
+\bibitem{solymosi2003note}
+\emph{Solymosi} \emph{J.}. Note on integral distances // Discrete \& Computational
+Geometry. — 2003. — Vol. 30, no. 2. — Pp. 337–342. — DOI: \href
+{https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.
+%
+\bibitem{kreisel2008heptagon}
+\emph{Kreisel} \emph{T.}, \emph{Kurz} \emph{S.}. There are integral heptagons, no
+three points on a line, no four on a circle // Discrete \& Computational Geometry. —
+2008. — Vol. 39, no. 4. — Pp. 786–790. — DOI: \href
+{https://doi.org/10.1007/s00454-007-9038-6} {\nolinkurl {10.1007/s00454-007-9038-6}}.
+%
+\bibitem{our-ped-2018}
+\emph{Авдеев} \emph{Н. Н.}. On integral point sets in special position // Некоторые
+вопросы анализа, алгебры, геометрии и математического образования: материалы
+международной молодежной научной школы «Актуальные направления
+математического анализа и смежные вопросы». — 2018. — Vol. 8. — Pp. 5–6.
+%
+\bibitem{our-pmm-2018}
+\emph{Авдеев} \emph{Н. Н.}. Об отыскании целоудалённых множеств
+специального вида // Актуальные проблемы прикладной математики,
+информатики и механики - сборник трудов Международной научной
+конференции. — Научно-исследовательские публикации, 2018. — Pp. 492–498.
+%
+\bibitem{our-vmmsh-2018}
+\emph{Авдеев} \emph{Н. Н.}, \emph{Семёнов} \emph{Е. М.}. Множества точек с
+целочисленными расстояниями на плоскости и в евклидовом пространстве //
+Математический форум (Итоги науки. Юг России). — 2018. — Pp. 217–236.%
+\bibitem{our-vmmsh-2018-translit}
+\emph{Avdeev} \emph{N. N.}, \emph{Semenov} \emph{E. M.}. Integral point sets in
+the plane and Euclidean space (Множества точек с целочисленными расстояниями
+на плоскости и в евклидовом пространстве) // Matematicheskij forum (Itogi nauki.
+Yug Rossii). — 2018. — Pp. 217–236.
+%
+\bibitem{avdeev2019particular}
+\emph{Avdeev} \emph{N. N.}, \emph{Zvolinsky} \emph{R. E.},
+\emph{Momot} \emph{E. A.}. On particular diameter bounds for integral point sets in
+higher dimensions. — 2019. — arXiv: \href {http://arxiv.org/abs/1909.10386}
+{\nolinkurl {1909.10386}}. — URL: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190910386A/abstract}.
+%
+\bibitem{kurz2013constructing}
+Constructing $7$-clusters / S. Kurz [et al.] // Serdica Journal of Computing. — 2014. —
+Vol. 8, no. 1. — Pp. 47–70. — arXiv: \href {http://arxiv.org/abs/1312.2318} {\nolinkurl
+{1312.2318}}.
+%
+\bibitem{antonov2008maximal}
+\emph{Antonov} \emph{A. R.}, \emph{Kurz} \emph{S.}. Maximal integral point sets
+over {$\mathbb {Z}^2$} // International Journal of Computer Mathematics. —
+2008. — Vol. 87, no. 12. — Pp. 2653–2676. — DOI: \href
+{https://doi.org/10.1080/00207160902993636} {\nolinkurl
+{10.1080/00207160902993636}}. — arXiv: \href {http://arxiv.org/abs/0804.1280}
+{\nolinkurl {0804.1280}}.
+%
+\bibitem{solymosi2010question}
+\emph{Solymosi} \emph{J.}, \emph{De Zeeuw} \emph{F.}. On a question of Erdős
+and Ulam // Discrete \& Computational Geometry. — 2010. — Vol. 43, no. 2. —
+Pp. 393–401. — arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl
+{0806.3095}}.
+\end{thebibliography}
 

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Beautify coordinates
Обновление картинок
[AMZAE] Add example of arrow-like set

@@
 Figure~\ref{42_symm.png}:
+\begin{multline}
+	\mathcal{P}_{42}=\sqrt{154}/{1} * \{
+		(\pm219513840; -15069600);
+		(\pm 345596160; 0);
+		(\pm260201760; 0);
 		\\
-$\mathcal{P}_{42}=\sqrt{154}/{1} * \{
-(\pm 345596160; 0);
-(\pm219513840; -15069600);
-(\pm260201760; 0);\\
 		(\pm225792840; 0);
@@
 		(\pm144668160; 0);
-(\pm25116000; 0);\\
+		(\pm25116000; 0);
+		\\
 		(\pm694026840; 0);
@@
 		(\pm13423904; 0);
-(\pm75682880; 0);\\
+		(\pm75682880; 0);
+		\\
 		(\pm464143680; 0);
@@
 		(\pm1195180740; 0);
-(\pm1236558752; 0);\\
+		(\pm1236558752; 0);
+		\\
 		(\pm44590560; 0);
 		(\pm339925740; 0);
-(\pm117312468; 0)\}
-$
+		(\pm117312468; 0)
+	\}
+\end{multline}
 
@@
 Figure~\ref{46_symm.png}:
+\begin{multline}
+	\mathcal{P}_{46}=
+	\sqrt{154}/{1} * \{
+		(\pm276778320; -19000800);
+		(\pm435751680; 0);
+		(\pm328080480; 0);
 		\\
-$\mathcal{P}_{46}=\sqrt{154}/{1} * \{
-(\pm435751680; 0);
-(\pm276778320; -19000800);
-(\pm328080480; 0);
-(\pm284695320; 0);\\
 		(\pm268861320; 0);
@@
 		(\pm182407680; 0);
-(\pm31668000; 0);
-(\pm1559139296; 0);\\
+		(\pm1559139296; 0);
+		(\pm284695320; 0);
+		\\
 		(\pm1506967020; 0);
@@
 		(\pm648982880; 0);
-(\pm585224640; 0);\\
+		(\pm585224640; 0);
+		\\
 		(\pm452799620; 0);
@@
 		(\pm116181975; 0);
-(\pm428602020; 0);\\
+		(\pm428602020; 0);
+		\\
 		(\pm56222880; 0);
@@
 		(\pm626458560; 0);
-(\pm130761918; 0)\}
-$
+		(\pm31668000; 0);
+		(\pm130761918; 0)
+	\}
+\end{multline}
 
@@
 Figure~\ref{48_symm.png}:
+\begin{multline}
+	\mathcal{P}_{48}=
+	\sqrt{154}/{1} * \{
+		( \pm6365901360 ; -437018400);
+		( \pm10022288640; 0);
+		( \pm23985026520 ; 0);
 		\\
-$\mathcal{P}_{48}=\sqrt{154}/{1} * \{
-( \pm10022288640; 0);
-( \pm6365901360 ; -437018400);
-( \pm7545851040 ; 0);\\
-( \pm6547992360 ; 0);
+		( \pm389293216 ; 0);
 		( \pm6183810360 ; 0);
@@
 		( \pm4195376640 ; 0);
-( \pm728364000 ; 0);\\
+		( \pm728364000 ; 0);
+		\\
 		( \pm35860203808 ; 0);
 		( \pm34660241460 ; 0);
-( \pm23985026520 ; 0);
-( \pm20126778360 ; 0);\\
+		( \pm7545851040 ; 0);
+		( \pm20126778360 ; 0);
+		\\
 		( \pm14926606240 ; 0);
@@
 		( \pm2194803520 ; 0);
-( \pm389293216 ; 0);\\
+		\\
 		( \pm2672185425 ; 0);
@@
 		( \pm17699891040 ; 0);
-( \pm14408546880 ; 0);\\
+		\\
+		( \pm14408546880 ; 0);
 		( \pm3007524114 ; 0);
-( \pm3402061572 ; 0)\}
-$
+		( \pm6547992360 ; 0);
+		( \pm3402061572 ; 0)
+	\}
+\end{multline}
 
@@
 
-%\section{Arrow-like integral point sets with one axis of symmetry}
+\section{Arrow-like integral point sets with one axis of symmetry}
+
+Figure~\ref{17_1024296_1_639b}:
+\begin{multline}
+	\mathcal{P}_8=
+	\sqrt{1}/1*\{
+		( -2847 ; \pm72072);
+		( 47073 ; \pm124488);
+		( 47073 ; 0);
+		( \pm98943 ; 0);
+		\\
+		( -71487 ; 0);
+		( -694668 ; 0);
+		( 15457 ; 0);
+		( 23582 ; 0);
+		( 125307 ; 0);
+		\\
+		( 329628 ; 0);
+		( -50367 ; 0);
+		( -14943 ; 0);
+		( 63073 ; 0);
+		( 172207 ; 0)
+	\}
+\end{multline}
+
+\begin{figure}[h!]
+\center{\includegraphics[width=1\linewidth]{./img/17_1024296_1_639b7a0452227fde1aceabf02e8778e4_other.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 17 and diameter 1024296}
+\label{17_1024296_1_639b}}
+\end{figure}
 
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{./img/8_2535_1_d680ff2d172ebe40250a54ca9419af34_other.png}}
+\center{\includegraphics[width=1\linewidth]{./img/8_2535_1_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 8 and diameter 2535}
@@
 \begin{figure}[h!]
-	\center{\includegraphics[width=1\linewidth]{./img/8_2400_42_56f33d8dc5df552595e807c438bc0f55_other.png}}
+	\center{\includegraphics[width=1\linewidth]{./img/8_2400_42_symm.png}}
 	\parbox{1\linewidth}{\caption{IPS of cardinality 8 and diameter 2400}

### 2021-03-15 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Fixes before arXiv

@@
 
+The set $\mathcal{P}_{19}$ shown on Figure~\ref{19_41327616148} was obtained from the set shown on Figure~\ref{9_2890_1_02e6af}
+by dilation and looking for points on $x$ axis.
+
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{./img/15_61960444_1_other.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 61960444}
-\label{15_61960444_1.png}}
+\center{\includegraphics[width=1\linewidth]{./img/9_2890_1_02e6afb1c19846c6c4debb44e52cc468_other.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 9 and diameter 2890}
+\label{9_2890_1_02e6af}}
 \end{figure}
 
+\begin{equation}
+	\mathcal{P}_9=\sqrt{1}/1*\{
+	( 0 ; 0);
+	( \pm1445 ; 0);
+	( \pm1085 ; 0);
+	( -455 ; \pm528);
+	( 455 ; \pm528)
+	\}
+\end{equation}
+
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{./img/15_45970652_1_other.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 45970652}
-\label{15_45970652_1.png}}
+\center{\includegraphics[width=1\linewidth]{./img/19_41327616148_1_88f7a023e19224a633e4213ec953bd8d_other.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 41327616148}
+\label{19_41327616148}}
 \end{figure}
 
-\begin{itemize}
-\item
-$\mathcal{P}=\sqrt{1}/1*\{
-( \pm9406950 ; 0);
-( \pm7063350 ; 0);
-( 2962050 ; \pm3437280);
-( 0 ; 0);
-( -2962050 ; \pm3437280);\\
-( \pm2188662 ; 0);
-( \pm30980222 ; 0);
-( \pm2914854 ; 0);
-\}$
+\begin{multline}
+	\mathcal{P}_{19}=
+	\sqrt{1}/1*\{(
+	\pm5492845050 ; 0);
+	( 1194096750 ; \pm2292665760);
+	( -2757277950 ; \pm2292665760);
+	\\
+	( 3929663850 ; 0);
+	( -781590600 ; 0);
+	( -7056026250 ; 0);
+	( -2241428154 ; 0);
+	\\
+	( 678246954 ; 0);
+	( -21445398674 ; 0);
+	( 19882217474 ; 0);
+	( -2725798218 ; 0);
+	\\
+	( 1162617018 ; 0);
+	( -3007651248 ; 0);
+	( 1444470048 ; 0);
+	( -3581930898 ; 0);
+	( 2018749698 ; 0)
+\}
+\end{multline}
 
-\item
-$\mathcal{P}=\sqrt{1}/1*\{
-( \pm6979350 ; 0);
-( \pm5240550 ; 0);
-( 2197650 ; \pm2550240);
-( 0 ; 0);
-( -2197650 ; \pm2550240);\\
-( \pm1623846 ; 0);
-( \pm22985326 ; 0);
-( \pm2476152 ; 0);
-\}$
-\end{itemize}
-
-%TODO:
-
-%TODO: 11_14450_1_8693fad981c82dc92d7c9365bcb9945e
-
-%The set on Fig. TODO has been obtained from the set on Fig. TODO by dilation by 5.
-
-%The set on Fig. TODO has been obtained from the set on Fig. TODO by dilation by 42 (or less? TODO).
-
-\section{Arrow-like integral point sets with one axis of symmetry}
+\section{Arrow-like integral point set with one axis of symmetry}
 
@@
 \begin{multline}
-	\mathcal{P}_8=
+	\mathcal{P}_{17}=
 	\sqrt{1}/1*\{

### 2021-03-27 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] A bunch of edits

@@
 
-Every integral point set also has a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.
+Every integral point set has also a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.
 Characteristic of an IPS does not change if the set is moved, dilated ot flipped;
@@
 \end{dfn}
-Facher sets are very dominating examples of planar integral pont sets.
+Facher sets are very dominating examples of planar integral point sets.
 In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
@@
 
-The set $\mathcal{P}_{19}$ shown on Figure~\ref{19_41327616148} was obtained from the set shown on Figure~\ref{9_2890_1_02e6af}
+The set $\mathcal{P}_{19}$ shown on Figure~\ref{19_20663808074} was obtained from the set $\mathcal{P}_{9}$ shown on Figure~\ref{9_2890_1_02e6af}
 by dilation and looking for points on $x$ axis.
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{./img/19_41327616148_1_88f7a023e19224a633e4213ec953bd8d_other.png}}
-\parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 41327616148}
-\label{19_41327616148}}
+\center{\includegraphics[width=1\linewidth]{./img/19_20663808074_1_0070b97f26bb194a7de2ddc37ee38b08_symm.png}}
+\parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 20663808074}
+\label{19_20663808074}}
 \end{figure}
 
-\begin{multline}
+\begin{multline*}
 	\mathcal{P}_{19}=
-	\sqrt{1}/1*\{(
-	\pm5492845050 ; 0);
-	( 1194096750 ; \pm2292665760);
-	( -2757277950 ; \pm2292665760);
+	\sqrt{1}/1*\{
+		( 0 ; 0);
+		( -987843675 ; \pm1146332880);
+		( 987843675 ; \pm1146332880);
 		\\
-	( 3929663850 ; 0);
-	( -781590600 ; 0);
-	( -7056026250 ; 0);
-	( -2241428154 ; 0);
+		( \pm729918777 ; 0);
+		( \pm972103809 ; 0);
+		( \pm1113030324 ; 0);
+		( \pm1400170149 ; 0);
 		\\
-	( 678246954 ; 0);
-	( -21445398674 ; 0);
-	( 19882217474 ; 0);
-	( -2725798218 ; 0);
-	\\
-	( 1162617018 ; 0);
-	( -3007651248 ; 0);
-	( 1444470048 ; 0);
-	( -3581930898 ; 0);
-	( 2018749698 ; 0)
+		( \pm3137217825 ; 0);
+		( \pm2355627225 ; 0);
+		( \pm10331904037 ; 0);
 	\}
-\end{multline}
+\end{multline*}
 
-\section{Arrow-like integral point set with one axis of symmetry}
+\section{Arrow-like integral point sets with one axis of symmetry}
 
-Figure~\ref{17_1024296_1_639b}:
+On Figure~\ref{17_1024296_1_639b}, the following IPS is shown:
 \begin{multline}
@@
 		\\
+		( -694668 ; 0);
 		( -71487 ; 0);
-		( -694668 ; 0);
-		( 15457 ; 0);
-		( 23582 ; 0);
-		( 125307 ; 0);
-		\\
-		( 329628 ; 0);
 		( -50367 ; 0);
 		( -14943 ; 0);
+		( 15457 ; 0);
+		\\
+		( 23582 ; 0);
 		( 63073 ; 0);
+		( 125307 ; 0);
 		( 172207 ; 0)
+		( 329628 ; 0);
 	\}
 \end{multline}
+Note that the axis of symmetry for $\mathcal{P}_{17}$ is the $x$ axis;
+all such sets are of characteristic 1.
+Moreover, note that the set contains three points with the same first coordinates.
 
@@
 
-%TODO:
+On Figures~\ref{10_2431_1_0606} and~\ref{15_19203744_1_80db}, another examples of arrow-like IPSs are shown.
+The one on Figure~\ref{15_19203744_1_80db} is obtained from the one on Figure~\ref{10_2431_1_0606}
+by dilation and looking for points on $x$ axis.
 
-%на каждое такое множество: исходный вариант, все этапы растращаивания с коэффициентом растаращивания и пояснением, кто из кого.
+\begin{figure}[h!]
+	\center{\includegraphics[width=1\linewidth]{./img/10_2431_1_0606659e825917be0c3c431c1375bf21_other.png}}
+	\parbox{1\linewidth}{\caption{IPS of cardinality 10 and diameter 2431}
+	\label{10_2431_1_0606}}
+\end{figure}
+
+\begin{figure}[h!]
+	\center{\includegraphics[width=1\linewidth]{./img/15_19203744_1_80dbf8e05b25f69b8366e50d1d7463f6_other.png}}
+	\parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 19203744}
+	\label{15_19203744_1_80db}}
+\end{figure}
+
+Figure~\ref{10_2431_1_0606}:
+\begin{multline*}
+	\mathcal{P}_{10} =
+	\sqrt{1}/1*\{
+		( 0 ; 0);
+		( 0 ; \pm252);
+		( 960 ; \pm468);
+		( -1120 ; 0);
+		( -405 ; 0);
+		( 336 ; 0);
+		( 561 ; 0);
+		( 1311 ; 0)
+	\}
+\end{multline*}
+
+Figure~\ref{15_19203744_1_80db}:
+\begin{multline*}
+	\mathcal{P}_{15} =
+	\sqrt{1}/1*\{
+		( 0 ; 0);
+		( 0 ; \pm1413720);
+		( 5385600 ; \pm2625480);
+		\\
+		( -6283200 ; 0);
+		( -2272050 ; 0);
+		( -1971915 ; 0);
+		( -635040 ; 0);
+		( 1884960 ; 0);
+		\\
+		( 3147210 ; 0);
+		( 4558176 ; 0);
+		( 5976333 ; 0);
+		( 7354710 ; 0);
+		( 12920544 ; 0);
+	\}
+\end{multline*}
 
@@
 
-The set shown on Figure~\ref{8_2400_42_56f3} has an axis of symmetry, but it is $y$ axis, not $x$ axis
-and due to the fact that its characteristic is not 1,
-the set cannot be rotated by $90^\circ$ but still stay on lattice~\eqref{eq:char_lattice}:
+The set shown on Figure~\ref{8_2400_42_56f3} has an axis of symmetry, but it is $y$ axis, not $x$ axis.
+Due to the fact that its characteristic is not 1,
+the set cannot be rotated by $90^\circ$ but still be on lattice~\eqref{eq:char_lattice}:
 \begin{equation}

### 2021-03-27 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Add the definition of characteristic

@@
 
-Every integral point set has also a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.
+Every planar IPS aso has a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.
+
+\begin{dfn}
+	The characteristic of a planar integral point set $M$ is the least positive integer $q$
+	such that the area of any triangle with vertices $M_1, M_2, M_3 \in M$
+	is commensurable with $\sqrt{q}$.
+\end{dfn}
 Characteristic of an IPS does not change if the set is moved, dilated ot flipped;

### 2021-03-27 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Edits from Usachev

@@
 	such that for any pair of the  points the Euclidean distance
-	between the points is integral.
+	between them is integral.
 	We discuss the classification of planar integral point sets
@@
 How do we describe an integral point set?
-For example, with the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
+For example, by the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
 and is said to be the \emph{cardinality} of the IPS.
@@
 
-It's notable that any of the examples that are discussed below
-is located on a union of at most three straight lines.
+Note that all examples that are discussed below are
+located on a union of at most three straight lines.
 %In the most pictures these lines are shown as solid lines.
@@
 \end{dfn}
-Facher sets are very dominating examples of planar integral point sets.
+Facher sets are predominant examples of planar integral point sets.
 In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
@@
 Figure~\ref{8_with_many_common} shows an example of three PIPS of cardinality 8,
-each pair of that shares 6 or 7 points but cannot be united into another PIPS.
+each pair of that shares 6 or 7 points but cannot be combined into another PIPS.
 
@@
 
-On Figures~\ref{10_2431_1_0606} and~\ref{15_19203744_1_80db}, another examples of arrow-like IPSs are shown.
+On Figures~\ref{10_2431_1_0606} and~\ref{15_19203744_1_80db}, other examples of arrow-like IPSs are shown.
 The one on Figure~\ref{15_19203744_1_80db} is obtained from the one on Figure~\ref{10_2431_1_0606}
@@
 although the set is of characteristic 1,
-it cannot be extended by reflecting relatively to the $x$ axis.
+it cannot be extended by the reflection in the $x$ axis.
 Moreover, we failed to extend it by dilation and looking for extra points on the $x$ axis.

### 2021-03-27 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Acknowledgements
[AMZAE] Name correction
[AMZAE] Keywords

@@
 	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
-	, Momot E.A., Zvolinsky A.E.
+	, Momot E.A., Zvolinskiy A.E.
 	\\
@@
 	and provide examples of sets that are not covered by the existent classification.
+
+\paragraph{Keywords.}
+	integral point sets,
+	planar integral point sets,
+	classification of planar integral point sets,
+	diameter of an integral point set
 
@@
 
+\section{Acknowledgements}
+The authors thank Dr. Prof. E.M. Semenov and Dr.~A.S.~Usachev for proofreading.
+
 %\printbibliography

### 2021-05-18 vestnikpmm20210518/Lushina_Avdeev.tex
[pmm2021] The very drafts

\documentclass[a4paper,openbib]{article}
\pagestyle{empty} % нумерация выкл.
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{amsfonts,amssymb}
\usepackage{latexsym}
\usepackage{euscript}
\usepackage{enumerate}
\usepackage{graphics}
\usepackage[dvips]{graphicx}
\usepackage{geometry}
\graphicspath{{examples/}}
\usepackage{wrapfig}
\usepackage[colorlinks=true,urlcolor=blue]{hyperref}
\usepackage{indentfirst}

\righthyphenmin=2

\usepackage[14pt]{extsizes}

\geometry{left=2.4cm}% левое поле
\geometry{right=2.4cm}% правое поле
\geometry{top=2.2cm}% верхнее поле
\geometry{bottom=3.2cm}% нижнее поле

\renewcommand{\baselinestretch}{1.3}

\renewcommand{\leq}{\leqslant}
\renewcommand{\geq}{\geqslant}

\newcommand{\longcomment}[1]{}

\begin{document}
\noindent УДК 514.112.3
%\clubpenalty=10000
%\widowpenalty=10000

\begin{center}
	\bfО ХАРАКТЕРИСТИКЕ ПЛОСКИХ МНОЖЕСТВ С ЦЕЛОЧИСЛЕННЫМИ РАССТОЯНИЯМИ С РЁБРАМИ 1 И 2
	\footnote{
		Работа выполнена в Воронежском университете при поддержке РНФ, грант 19-11-00197.
	}

	\bfЛушина Е. А., Авдеев Н. Н.

	\itВоронежский Государственный Университет
\end{center}

Аннотация: Разработано ПО для выборочного сжатия специфичных изображений (фотографий печатного и рукописного текста) с потерями, распределяемыми по изображению неравномерно.
Приведены критерии определения областей изображения, не содержащих полезной информации. Рассмотрены популярные методы сжатия изображений и возможность их комбинирования с
представленным в данной статье методом. Выдвинуты идеи и планы по совершенствованию разработанного ПО.

Ключевые слова: разработка ПО, Javascript, сжатие изображений, сжатие JPEG, однородные области изображений, цветовое пространство RGB, статистические методы обработки изображений

We develop software for selective lossy compression for images of special type (photos of print or handwriten text) with uneven loss distribution. We provide two criteria of
image regions without valuable information. Well-known methods of image compression are reviewed, and their compatibility with our approach is discussed. Ways to improve our
software are proposed.

Keywords: software development, Javascript, image compression, JPEG compression, homogenous regions of images, RGB color space, statistical methods of image processing

Обозначим через $\mathfrak{M}$ множество таких подмножеств $M\left\{M_{1}, M_{2}, \ldots\right\} \subset$ $\mathbb{R}^{2}$, что $\left|M_{i}-M_{j}\right| \in \mathbb{N}$ для всех $\mathrm{i}, \mathrm{j}$, где $\mathbb{N}-$ множество натуральных чисел и $\left|M_{i}-M_{j}\right|-$ расстояние между точками $M_{i}$ и $M_{j} .$

Доказано [1,2], что всякое бесконечное множество $M \in \mathfrak{M}$ содержится в некоторой прямой $L \subset \mathbb{R}^{2}$, и поэтому оно является предметом изучения не столько комбинаторной геометрии, сколько теории чисел, что выходит за рамки рассматриваемой проблемы.

Определение. Мощностью множества М, состоящего из конечного
числа элементов, называется количество элементов этого множества. Обозначается |M|.

Для заданного $n \in \mathbb{N}$ обозначим через $\mathfrak{M}_{n}$ множество таких $M \in \mathfrak{M}$, что |M|=n и $M \not \subset L$ для любой прямой $L \subset \mathbb{R}^{2}$ Множество $M \in \mathfrak{M}$ характеризуется такими параметрами, как
1. Мощность
2. Диаметр
3. Характеристика
4. Минимальное ребро Определение. Диаметром множества М будем называть наибольшее расстояние между любыми двумя точками множества М.
$\operatorname{diam} M=\max |A-B|$, где $A, B \in M$
Определение. Характеристикой множества $M \in \mathfrak{M}_{n} \quad$ называется свободное от квадратов число р такое, что плошадь любого треугольника $\mathrm{ABC}$, где $A, B, C \in M$ соизмерима $\mathrm{c} \sqrt{p} .$ Обозначается char $\mathrm{M}=\mathrm{p} .$

Определение. Ребром множества М будем называть любой отрезок $M_{1} M_{2}$, где $M_{1}, M_{2} \in M$
Множество М характеристики р может быть размещено на решетке
$$
\left\{\left(\frac{a_{i}}{2 m} ; \frac{b_{i} \sqrt{p}}{2 m}\right)\right\}
$$
где $a_{i}, b_{i} \in \mathbb{Z}(\mathbb{Z}-$ множество целых чисел), в качестве $\mathrm{m}$ можно взять длину любого ребра множества $\mathrm{M}$ [3, Теорема 4].
Приведем несколько известных фактов о множествах $M \in \mathfrak{M}_{n}:$
\begin{itemize}
- Для любого $n \in \mathbb{N}, n \geq 3$ выполнено $\mathfrak{M}_{n} \neq \emptyset$ [3, Теорема 2].
$>\quad \mathrm{C}$ возрастанием мощности диаметр возрастает не менее чем линейно [4,6].
- Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого свободного от квадратов числа р существует множество $M \in \mathfrak{M}_{n}$ такое, что char $\mathrm{M}=\mathrm{p}[3,$, Teopema 5 ].

\item Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого натурального числа $\mathrm{k}$ существует множество $M \in \mathfrak{M}_{n}$, содержащее ребро $\mathrm{k}[5] .$
$>\quad$ Не существует множества $M \in \mathfrak{M}_{n}$ характеристики 1 с ребром 1
$[3] .$
\end{itemize}
В статье «О существовании специальных множеств с целочисленными расстояниями с ребром 1» получены следующие результаты исследования множества с характеристикой 2,3 и $5:$
1. $\quad$ Не существует множества $M \in \mathfrak{M}_{3}$ характеристики 2 или $5 \mathrm{c}$ ребром $1 .$
2. $\quad$ Существует бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{3}$, char M=3
с ребром 1 и char $\mathrm{M}=5$ с ребром $2 .$ В данной работе рассматривается множество $M \in \mathfrak{M}_{3}$, char M=6 и char

М=7 с ребром 1. Доказывается, что не существует множества $M \in \mathfrak{M}_{3} \mathrm{c}$ характеристикой вида $4 k+1,4 k+2$, при $k=0,1,2, \ldots$ Вместе с тем, мы строим бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{3}$, char $\mathrm{M}=4 k+3$, при $k=$ $0,1,2, \ldots$ с ребром $1 .$
Утверждение 1. Не существует множества $M \in \mathfrak{M}_{3}$ характеристики $6 \mathrm{c}$ ребром $1 .$ Доказательство. По теореме о сетке введем систему координат так, что $\mathrm{O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{b \sqrt{6}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
Пифагора:
$$
\begin{array}{c}
\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{6}}{2}\right)^{2}=r^{2} \\
1+6 b^{2}=4 r^{2}
\end{array}
$$
Очевидно, что в уравнении (1) левая часть - нечетное число при любом $b \in \mathbb{Z}$, а правая $-$ четное число при любом $r \in \mathbb{Z} .$ Следовательно, уравнение
(1) неразрешимо в целых числах. $\mathbf{y}_{\text {тверждение }}$ 2. Существует бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{n}$ характеристики 7 с ребром $1 .$

Доказательство. По теореме о сетке введем систему координат так, что $\mathrm{O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{\mathrm{b} \sqrt{7}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
Пифагора:
$$
\begin{array}{c}
\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{7}}{2}\right)^{2}=r^{2} \\
1+7 b^{2}=4 r^{2}
\end{array}
$$
Уравнение (2) имеет решение в целых числах, если $\left(\frac{1+7 b^{2}}{4}\right)-$ полный квадрат. Например, $b^{2}=9, \mathrm{r}^{2}=16 .$ Таким образом, $(4 ; 3)-$ решение уравнения (2).
Уравнение (2) равносильно следующему уравнению:
$4 r^{2}-7 b^{2}=1$
Введем замену: $n=2 r .$ Получим
$$
n^{2}-7 b^{2}=1
$$
Уравнение (4) называется уравнением Пелля. Определение. Уравнение Пелля - диофантово уравнение вида
$$
x^{2}-m y^{2}=1
$$
где $\mathrm{m}-$ натуральное число, не являющееся квадратом. Ясно, что уравнение Пелля имеет тривиальное решение $\mathrm{x}=\pm 1, \mathrm{y}=0 .$ Теорема 1. Любое уравнение Пелля имеет нетривиальное решение [7]. Теорема 2. Все нетривиальные положительные решения получаются многократным умножением основного решения на себя [7].

Применим формулу разности квадратов к уравнению (4) и подставим вместо неизвестных найденное решение:
$$
\begin{array}{l}
(n-b \sqrt{7})(n+b \sqrt{7})=1 \\
(8-3 \sqrt{7})(8+3 \sqrt{7})=1
\end{array}
$$
Согласное Теореме 2 :
$$
(8-3 \sqrt{7})^{\mathrm{k}}(8+3 \sqrt{7})^{\mathrm{k}}=1^{\mathrm{k}} \quad \mathrm{k} \in \mathbb{N}
$$
Заметим, что при возведении в четные степени \textrm{K} , ~ п о л у ч а ю т с я ~ н е ч е т н ы е ~

n, а следовательно, $r-$ нецелые числа. Поэтому будем возводить выражение
(5) в степени вида $\mathrm{k}=2 \mathrm{~m}+1, \mathrm{~m}=0,1,2 \ldots$
Приведем несколько решений уравнения (4):
При $\mathrm{k}=3:(2024-765 \sqrt{7})(2024+765 \sqrt{7})=1 \quad$ пара $\quad$ чисел
(2024;765) является решением уравнения (4), следовательно, (1012;765) решение уравнения (3).
$$
\text { При } \mathrm{k}=5:(514088-194307 \sqrt{7})(514088+194307 \sqrt{7})=1
$$
(514088; 194307) - решением уравнения (4), (257044; 194307) - решение уравнения (3).

Продолжая возводить в нечетные степени обе части уравнения, получим бесконечную последовательность решений для исходного уравнения (2):
$$
\left(\frac{x_{n}}{2} ; y_{n}\right), \text { где } x_{n}+y_{n} \sqrt{7}=(8-3 \sqrt{7})^{2 n+1}, n=0,1,2 \ldots
$$
Таким образом, существует бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{n}$ характеристики 7 с ребром $1 .$

Утверждение 3. Всякое множество $M \in \mathfrak{M}_{3}$ с минимальным ребром имеет характеристику вида $\mathrm{p}=4 \mathrm{k}+3, \mathrm{k}=0,1,2 \ldots$
Доказательство. Пусть минимальное ребро множества $\mathfrak{M}_{3}$ равно $1 .$ Воспользуемся теоремой о сетке и введем систему координат так, что О (0;0),
$A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{b \sqrt{p}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме Пифагора:
$$
\begin{array}{c}
\left(\frac{1}{2}\right)^{2}+\left(\frac{\mathrm{b} \sqrt{\mathrm{p}}}{2}\right)^{2}=\mathrm{r}^{2} \\
1+p b^{2}=4 r^{2}
\end{array}
$$
Для того, чтобы $\mathrm{r}^{2}$ было целым числом необходимо, чтобы левая часть уравнения была кратна 4 и переменные $\mathrm{p}, \mathrm{b}-$ нечетные числа. Следовательно, произведение $p b^{2}$ должно удовлетворять следующему условию: $p b^{2}=4 k+$ $+3, k=0,1,2, \ldots$ Тогда $b^{2}=(2 \mathrm{~m}+1)^{2}=4 \mathrm{~m}^{2}+4 \mathrm{~m}+1=4\left(\mathrm{~m}^{2}+\mathrm{m}\right)+1=$
$4 n+1, \mathrm{n}=0,1,2, \ldots$ Так как это равенство выполнено при любом $\mathrm{n}$, оно выполнено и при $\mathrm{n=0.}$ Отсюда, $p=4 k+3, k=0,1,2, \ldots$ $\mathbf{y}_{\text {тверждение } 4 . \text { Не существует множества } M \in \mathfrak{M}_{3} \mathrm{c} \text { характеристикой }}$
$4 k+1, k=0,1,2, \ldots$ и ребром $1 .$
Доказательство. По теореме о сетке введем систему координат так, что $\mathrm{O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{b \sqrt{4 k+1}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
Пифагора:
$$
\begin{array}{c}
\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{4 k+1}}{2}\right)^{2}=r^{2} \\
1+(4 k+1) b^{2}=4 r^{2}
\end{array}
$$
Заметим, что обе части уравнения (5) должны быть одной четности, следовательно, $b=2 \mathrm{~m}+1 .$ Тогда $b^{2}=4 n+1, \mathrm{n}=0,1,2, \ldots$
$$
\begin{aligned}
1+(4 k+1)(4 n+1) &=4 r^{2} \\
& 16 n k+4 k+4 n+2=4 r^{2}
\end{aligned}
$$
Очевидно, что левая часть уравнения (6) ни при каком $n, k \in \mathbb{Z}$ не делится на 4. Таким образом, уравнение (5) неразрешимо в целых числах. $\mathbf{y}_{\mathbf{T в е р} \boldsymbol{ж}}$ ение 4. Не существует множества $M \in \mathfrak{M}_{3} \mathrm{c}$ характеристикой $4 k+2, k=0,1,2, \ldots$ и ребром $1 .$
Доказательство. Воспользуемся способом задания системы координат, указанным в утверждении $3 .$
$$
\begin{array}{c}
\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{4 k+2}}{2}\right)^{2}=r^{2} \\
1+(4 k+2) b^{2}=4 r^{2}
\end{array}
$$
Заметим, что правая часть уравнения (7) четна при любом $r \in \mathbb{Z}, \mathrm{a}$ левая часть - нечетна при любом $b \in \mathbb{Z} .$ Таким образом, уравнение (7) не имеет решений в целых числах. Утверждение 5. Для всякой характеристики р найдется система с ребром 2
Доказательство. По теореме о сетке введем следующую систему
координат: O $(0 ; 0), A(-1 ; 0), B(1 ; 0), C\left(0 ; \frac{b \sqrt{\mathrm{p}}}{4}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
Пифагора:
$$
1+\left(\frac{\mathrm{b} \sqrt{\mathrm{p}}}{4}\right)^{2}=\mathrm{r}^{2}
$$
Отсюда,
$$
1+\frac{\mathrm{pb}^{2}}{16}=\mathrm{r}^{2}
$$
Правая часть уравнения (8) является целым числом в том случае, если b=4n, $\mathrm{n}=0,1,2, \ldots$ Тогда
$$
\begin{array}{l}
1+\mathrm{pn}^{2}=r^{2} \\
r^{2}-\mathrm{pn}^{2}=1
\end{array}
$$
Очевидно, равенство (9) является уравнением Пелля. По теореме 1 [7], уравнение (9) имеет нетривиальное решение, при любом свободном оТ квадратов числе $\mathrm{p} .$

Список использованной литературы:
1. Erdös P. Integral distances // Bull. Amer. Math. Soc. --- 1945. --- Vol. 51, N 12 . ---
P. 996.

2. Anning N. H., Erdös P. Integral distances // Bull. Amer. Math. Soc.-1945.Vol. 51, № 8.-P. 598-600.

3. Авдеев Н.Н., Семенов Е.М. Множества точек с целочисленными расстояниями на плоскости и в Евклидовом пространстве.

4. Solymosi J. Note on integral distances //Discrete \& Computational Geometry. --- 2003. --- T. 30. --- №. 2. --- C. 337-342.

5. Зволинский P. E. Facher integral point sets with particular distances of arbitrary cardinality.
В сборнике: Актуальные проблемы прикладной математики, информатики и механики сборник трудов Международной научной конференции. 2021. С. 668-674

6. Avdeev N. N. On existence of integral point sets and their diameter bounds //Australasian Journal of Combinatorics. – 2020. – Т. 77. – С. 100-116.

7. Бугаенко В. О. Уравнение Пелля. Серия: «Библиотека «Математическое просвещение»».М.: МЦНМО, 2001. --- 32с.: ил.

On characteristic of planar integral point sets with edges 1 and 2

{\bf Лушина Екатерина Александровна}, студент математического факультета ФГБОУ ВО «Воронежский государственный университет», г.~Воронеж.

{\bf Авдеев Николай Николаевич}, аспирант кафедры теории функций и геометрии математического факультета ФГБОУ ВО «Воронежский государственный университет», г.~Воронеж.

e-mail: nickkolok@mail.ru

Научный руководитель:
{\bf Семенов Евгений Михайлович},
заведующий кафедрой теории функций и геометрии математического факультета ФГБОУ ВО «Воронежский государственный
университет», г. Воронеж.

\end{document}

### 2021-05-18 vestnikpmm20210518/Lushina_Avdeev.tex
[pmm2021] Keywords

@@
 \begin{center}
-	\bfО ХАРАКТЕРИСТИКЕ ПЛОСКИХ МНОЖЕСТВ С ЦЕЛОЧИСЛЕННЫМИ РАССТОЯНИЯМИ С РЁБРАМИ 1 И 2
+	\bfО ХАРАКТЕРИСТИКЕ ПЛОСКИХ МНОЖЕСТВ С ЦЕЛОЧИСЛЕННЫМИ РАССТОЯНИЯМИ\\С РЁБРАМИ 1 И 2
 	\footnote{
@@
 
-Аннотация: Разработано ПО для выборочного сжатия специфичных изображений (фотографий печатного и рукописного текста) с потерями, распределяемыми по изображению неравномерно.
-Приведены критерии определения областей изображения, не содержащих полезной информации. Рассмотрены популярные методы сжатия изображений и возможность их комбинирования с
-представленным в данной статье методом. Выдвинуты идеи и планы по совершенствованию разработанного ПО.
+Аннотация:
 
-Ключевые слова: разработка ПО, Javascript, сжатие изображений, сжатие JPEG, однородные области изображений, цветовое пространство RGB, статистические методы обработки изображений
+Ключевые слова:
+	плоские множества с целочисленными расстояниями,
+	характеристика треугольника,
+	уравнение Пелля
 
-We develop software for selective lossy compression for images of special type (photos of print or handwriten text) with uneven loss distribution. We provide two criteria of
-image regions without valuable information. Well-known methods of image compression are reviewed, and their compatibility with our approach is discussed. Ways to improve our
-software are proposed.
+Abstract:
 
-Keywords: software development, Javascript, image compression, JPEG compression, homogenous regions of images, RGB color space, statistical methods of image processing
+Keywords:
+	planar integral point sets,
+	characteristic of a triangle,
+	Pell equation
 

### 2021-05-18 vestnikpmm20210518/Lushina_Avdeev.tex
[pmm2021] WIP

@@
 
-Обозначим через $\mathfrak{M}$ множество таких подмножеств $M\left\{M_{1}, M_{2}, \ldots\right\} \subset$ $\mathbb{R}^{2}$, что $\left|M_{i}-M_{j}\right| \in \mathbb{N}$ для всех $\mathrm{i}, \mathrm{j}$, где $\mathbb{N}-$ множество натуральных чисел и $\left|M_{i}-M_{j}\right|-$ расстояние между точками $M_{i}$ и $M_{j} .$
+Обозначим через $\mathfrak{M}$ множество таких подмножеств $M\left\{M_{1}, M_{2}, \ldots\right\} \subset$ $\mathbb{R}^{2}$,
+что $\left|M_{i}-M_{j}\right| \in \mathbb{N}$ для всех $\mathrm{i}, \mathrm{j}$,
+где $\mathbb{N}$ "--- множество натуральных чисел,
+$\mathbb{R}^2$ "--- евклидова плоскость
+и $\left|M_{i}-M_{j}\right|$ "--- расстояние между точками $M_{i}$ и $M_{j}$.
 
-Доказано [1,2], что всякое бесконечное множество $M \in \mathfrak{M}$ содержится в некоторой прямой $L \subset \mathbb{R}^{2}$, и поэтому оно является предметом изучения не столько комбинаторной геометрии, сколько теории чисел, что выходит за рамки рассматриваемой проблемы.
+Доказано [1,2], что всякое бесконечное множество $M \in \mathfrak{M}$ содержится в некоторой прямой $L \subset \mathbb{R}^{2}$,
+и потому может быть отождествлено с некоторым подмножеством множества $\mathbb{Z}$ целых чисел,
+
+является предметом изучения не столько комбинаторной геометрии,
+сколько теории чисел, что выходит за рамки рассматриваемой проблемы.
 
@@
 Определение.
-Характеристикой множества $M \in \mathfrak{M}_{n} \quad$ называется свободное от квадратов число р такое, что плошадь любого треугольника $\mathrm{ABC}$, где $A, B, C \in M$ соизмерима $\mathrm{c} \sqrt{p} .$ Обозначается char $\mathrm{M}=\mathrm{p} .$
+Характеристикой множества $M \in \mathfrak{M}_{n} \quad$ называется свободное от квадратов число р такое, что плошадь любого треугольника $\mathrm{ABC}$, где $A, B, C \in M$ соизмерима $\mathrm{c} \sqrt{p} .$ Обозначается char $\mathrm{M}={p} .$
 
@@
 $$
-где $a_{i}, b_{i} \in \mathbb{Z}(\mathbb{Z}-$ множество целых чисел), в качестве $\mathrm{m}$ можно взять длину любого ребра множества $\mathrm{M}$ [3, Теорема 4].
+где $a_{i}, b_{i} \in \mathbb{Z}(\mathbb{Z}$ "--- множество целых чисел), в качестве $m$ можно взять длину любого ребра множества $\mathrm{M}$ [3, Теорема 4].
 Приведем несколько известных фактов о множествах $M \in \mathfrak{M}_{n}:$
 \begin{itemize}
-- Для любого $n \in \mathbb{N}, n \geq 3$ выполнено $\mathfrak{M}_{n} \neq \emptyset$ [3, Теорема 2].
-$>\quad \mathrm{C}$ возрастанием мощности диаметр возрастает не менее чем линейно [4,6].
-- Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого свободного от квадратов числа р существует множество $M \in \mathfrak{M}_{n}$ такое, что char $\mathrm{M}=\mathrm{p}[3,$, Teopema 5 ].
+	\item
+		Для любого $n \in \mathbb{N}, n \geq 3$ выполнено $\mathfrak{M}_{n} \neq \emptyset$ [3, Теорема 2].
+	\item
+		С возрастанием мощности диаметр возрастает не менее чем линейно [4,6].
+	\item
+		Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого свободного от квадратов числа $р$ существует множество $M \in \mathfrak{M}_{n}$ такое, что char $\mathrm{M}={p}[3,$, Teopema 5 ].
 
-\item Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого натурального числа $\mathrm{k}$ существует множество $M \in \mathfrak{M}_{n}$, содержащее ребро $\mathrm{k}[5] .$
-$>\quad$ Не существует множества $M \in \mathfrak{M}_{n}$ характеристики 1 с ребром 1
-$[3] .$
+	\item
+		Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого натурального числа $k$ существует множество $M \in \mathfrak{M}_{n}$, содержащее ребро ${k}$[5].
+	\item
+		Не существует множества $M \in \mathfrak{M}_{n}$ характеристики 1 с ребром 1 [3].
 \end{itemize}
@@
 Доказательство.
-По теореме о сетке введем систему координат так, что $\mathrm{O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{\mathrm{b} \sqrt{7}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
+По теореме о сетке введем систему координат так, что $\mathrm{O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{{b} \sqrt{7}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
 Пифагора:
@@
 $$
-Уравнение (2) имеет решение в целых числах, если $\left(\frac{1+7 b^{2}}{4}\right)-$ полный квадрат. Например, $b^{2}=9, \mathrm{r}^{2}=16 .$ Таким образом, $(4 ; 3)-$ решение уравнения (2).
+Уравнение (2) имеет решение в целых числах, если $\left(\frac{1+7 b^{2}}{4}\right)-$ полный квадрат. Например, $b^{2}=9, {r}^{2}=16 .$ Таким образом, $(4 ; 3)-$ решение уравнения (2).
 Уравнение (2) равносильно следующему уравнению:
@@
 $$
-Согласное Теореме 2 :
+Согласно теореме 2,
 $$
-(8-3 \sqrt{7})^{\mathrm{k}}(8+3 \sqrt{7})^{\mathrm{k}}=1^{\mathrm{k}} \quad \mathrm{k} \in \mathbb{N}
+(8-3 \sqrt{7})^{{k}}(8+3 \sqrt{7})^{{k}}=1^{{k}} \quad {k} \in \mathbb{N}
 $$
@@
 n, а следовательно, $r-$ нецелые числа. Поэтому будем возводить выражение
-(5) в степени вида $\mathrm{k}=2 \mathrm{~m}+1, \mathrm{~m}=0,1,2 \ldots$
+(5) в степени вида ${k}=2 \mathrm{~m}+1, \mathrm{~m}=0,1,2 \ldots$
 Приведем несколько решений уравнения (4):
-При $\mathrm{k}=3:(2024-765 \sqrt{7})(2024+765 \sqrt{7})=1 \quad$ пара $\quad$ чисел
+При ${k}=3:(2024-765 \sqrt{7})(2024+765 \sqrt{7})=1 \quad$ пара $\quad$ чисел
 (2024;765) является решением уравнения (4), следовательно, (1012;765) решение уравнения (3).
 $$
-\text { При } \mathrm{k}=5:(514088-194307 \sqrt{7})(514088+194307 \sqrt{7})=1
+\text { При } {k}=5:(514088-194307 \sqrt{7})(514088+194307 \sqrt{7})=1
 $$
@@
 $$
-Таким образом, существует бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{n}$ характеристики 7 с ребром $1 .$
+Таким образом, существует бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{n}$ характеристики 7 с ребром 1.
 
-Утверждение 3. Всякое множество $M \in \mathfrak{M}_{3}$ с минимальным ребром имеет характеристику вида $\mathrm{p}=4 \mathrm{k}+3, \mathrm{k}=0,1,2 \ldots$
-Доказательство. Пусть минимальное ребро множества $\mathfrak{M}_{3}$ равно $1 .$ Воспользуемся теоремой о сетке и введем систему координат так, что О (0;0),
+\paragraph{Утверждение 3.}
+Всякое множество $M \in \mathfrak{M}_{3}$ с минимальным ребром 1 имеет характеристику вида
+$p=4k+3, k=0,1,2 \ldots$
+\paragraph{Доказательство.}
+Пусть минимальное ребро множества $\mathfrak{M}_{3}$ равно $1 .$ Воспользуемся теоремой о сетке и введем систему координат так, что О (0;0),
 $A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{b \sqrt{p}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме Пифагора:
@@
 \begin{array}{c}
-\left(\frac{1}{2}\right)^{2}+\left(\frac{\mathrm{b} \sqrt{\mathrm{p}}}{2}\right)^{2}=\mathrm{r}^{2} \\
+\left(\frac{1}{2}\right)^{2}+\left(\frac{{b} \sqrt{{p}}}{2}\right)^{2}={r}^{2} \\
 1+p b^{2}=4 r^{2}
@@
 $$
-Для того, чтобы $\mathrm{r}^{2}$ было целым числом необходимо, чтобы левая часть уравнения была кратна 4 и переменные $\mathrm{p}, \mathrm{b}-$ нечетные числа.
-Следовательно, произведение $p b^{2}$ должно удовлетворять следующему условию: $p b^{2}=4 k+$ $+3, k=0,1,2, \ldots$ Тогда $b^{2}=(2 \mathrm{~m}+1)^{2}=4 \mathrm{~m}^{2}+4 \mathrm{~m}+1=4\left(\mathrm{~m}^{2}+\mathrm{m}\right)+1=$
-$4 n+1, \mathrm{n}=0,1,2, \ldots$ Так как это равенство выполнено при любом $\mathrm{n}$, оно выполнено и при $\mathrm{n=0.}$ Отсюда, $p=4 k+3, k=0,1,2, \ldots$ $\mathbf{y}_{\text {тверждение } 4 .
-\text { Не существует множества } M \in \mathfrak{M}_{3} \mathrm{c} \text { характеристикой }}$
-$4 k+1, k=0,1,2, \ldots$ и ребром $1 .$
-Доказательство.
-По теореме о сетке введем систему координат так, что $\mathrm{O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{b \sqrt{4 k+1}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
-Пифагора:
+Для того, чтобы ${r}^{2}$ было целым числом необходимо,
+чтобы левая часть уравнения была кратна 4 и ${p}, {b}$ были  нечетными.
+Следовательно,  $p b^{2}\equiv3(\operatorname{mod}4)$.
+Но $b^{2}\equiv1(\operatorname{mod}4)$,
+откуда $p=4 k+3, k\in\mathbb{N}$.
+
+\paragraph{Утверждение 4.}
+Для всякого свободного от квадратов числа $р$ существует бесконечное семейство множеств $\{M_j\}\in\mathfrak{M}_3$ с ребром 2
+таких, что $\operatorname{char}M_j = p$.
+\paragraph{Доказательство.}
+По теореме о сетке введем следующую систему
+координат:  $O(0 ; 0), A(-1 ; 0), B(1 ; 0), C\left(0 ; \frac{b \sqrt{p}}{4}\right)$ и $A, B, C \in M$.
+По теореме Пифагора:
 $$
-\begin{array}{c}
-\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{4 k+1}}{2}\right)^{2}=r^{2} \\
-1+(4 k+1) b^{2}=4 r^{2}
-\end{array}
+1+\left(\frac{{b} \sqrt{{p}}}{4}\right)^{2}={r}^{2}
 $$
-Заметим, что обе части уравнения (5) должны быть одной четности, следовательно, $b=2 \mathrm{~m}+1 .$ Тогда $b^{2}=4 n+1, \mathrm{n}=0,1,2, \ldots$
+Отсюда
 $$
-\begin{aligned}
-1+(4 k+1)(4 n+1) &=4 r^{2} \\
-& 16 n k+4 k+4 n+2=4 r^{2}
-\end{aligned}
+1+\frac{\mathrm{pb}^{2}}{16}={r}^{2}\eqno{(8)}
 $$
-Очевидно, что левая часть уравнения (6) ни при каком $n, k \in \mathbb{Z}$ не делится на 4.
-Таким образом, уравнение (5) неразрешимо в целых числах.
-$\mathbf{y}_{\mathbf{T в е р} \boldsymbol{ж}}$ ение 4.
-Не существует множества $M \in \mathfrak{M}_{3} \mathrm{c}$ характеристикой $4 k+2, k=0,1,2, \ldots$ и ребром $1 .$
-Доказательство. Воспользуемся способом задания системы координат, указанным в утверждении $3 .$
+Правая часть уравнения (8) является целым числом в том случае, если $b=4n$, $n\in\mathbb{N}$.
+Тогда
+$
+1+pn^{2}=r^{2}
+$ и
 $$
-\begin{array}{c}
-\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{4 k+2}}{2}\right)^{2}=r^{2} \\
-1+(4 k+2) b^{2}=4 r^{2}
-\end{array}
+r^{2}-pn^{2}=1\eqno{(9)}
 $$
-Заметим, что правая часть уравнения (7) четна при любом $r \in \mathbb{Z}, \mathrm{a}$ левая часть - нечетна при любом $b \in \mathbb{Z} .$ Таким образом, уравнение (7) не имеет решений в целых числах.
-Утверждение 5.
-Для всякой характеристики р найдется система с ребром 2
-Доказательство. По теореме о сетке введем следующую систему
-координат: O $(0 ; 0), A(-1 ; 0), B(1 ; 0), C\left(0 ; \frac{b \sqrt{\mathrm{p}}}{4}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
-Пифагора:
-$$
-1+\left(\frac{\mathrm{b} \sqrt{\mathrm{p}}}{4}\right)^{2}=\mathrm{r}^{2}
-$$
-Отсюда,
-$$
-1+\frac{\mathrm{pb}^{2}}{16}=\mathrm{r}^{2}
-$$
-Правая часть уравнения (8) является целым числом в том случае, если b=4n, $\mathrm{n}=0,1,2, \ldots$ Тогда
-$$
-\begin{array}{l}
-1+\mathrm{pn}^{2}=r^{2} \\
-r^{2}-\mathrm{pn}^{2}=1
-\end{array}
-$$
-Очевидно, равенство (9) является уравнением Пелля. По теореме 1 [7], уравнение (9) имеет нетривиальное решение, при любом свободном оТ квадратов числе $\mathrm{p} .$
+Очевидно, равенство (9) является уравнением Пелля и имеет нетривиальное решение при любом свободном от квадратов числе ${p}$.
 
-Список использованной литературы:
+Список использованной литературы
+
 1. Erdös P. Integral distances // Bull. Amer. Math. Soc. --- 1945. --- Vol. 51, N 12 . ---

### 2021-05-18 vestnikpmm20210518/Lushina_Avdeev.tex
[pmm2021] WIP

@@
 
-Обозначим через $\mathfrak{M}$ множество таких подмножеств $M\left\{M_{1}, M_{2}, \ldots\right\} \subset$ $\mathbb{R}^{2}$,
+Обозначим через $\mathfrak{M}$ множество таких подмножеств $M=\left\{M_{1}, M_{2}, \ldots\right\} \subset$ $\mathbb{R}^{2}$,
 что $\left|M_{i}-M_{j}\right| \in \mathbb{N}$ для всех $\mathrm{i}, \mathrm{j}$,
@@
 
-Доказано [1,2], что всякое бесконечное множество $M \in \mathfrak{M}$ содержится в некоторой прямой $L \subset \mathbb{R}^{2}$,
+Доказано [1,2], что всякое бесконечное множество $M \in \mathfrak{M}$ содержится в некоторой прямой $L \subset \mathbb{R}^{2}$
 и потому может быть отождествлено с некоторым подмножеством множества $\mathbb{Z}$ целых чисел,
+которое, в свою очередь, является объектом изучения теории чисел, но не геометрии.
+Таким образом, нас интересуют только конечные множества $M \in \mathfrak{M}$.
 
-является предметом изучения не столько комбинаторной геометрии,
-сколько теории чисел, что выходит за рамки рассматриваемой проблемы.
+\paragraph{Определение.}
+Мощностью множества $M$ (обозначается $|M|$), состоящего из конечного
+числа элементов, называется количество элементов этого множества.
 
-Определение. Мощностью множества М, состоящего из конечного
-числа элементов, называется количество элементов этого множества. Обозначается |M|.
+Для заданного $n \in \mathbb{N}$ обозначим через $\mathfrak{M}_{n}$ множество таких $M \in \mathfrak{M}$,
+что $|M|=n$ и $M \not \subset L$ для любой прямой $L \subset \mathbb{R}^{2}$.
 
-Для заданного $n \in \mathbb{N}$ обозначим через $\mathfrak{M}_{n}$ множество таких $M \in \mathfrak{M}$, что |M|=n и $M \not \subset L$ для любой прямой $L \subset \mathbb{R}^{2}$ Множество $M \in \mathfrak{M}$ характеризуется такими параметрами, как
-1. Мощность
-2. Диаметр
-3. Характеристика
-4. Минимальное ребро Определение. Диаметром множества М будем называть наибольшее расстояние между любыми двумя точками множества М.
-$\operatorname{diam} M=\max |A-B|$, где $A, B \in M$
-Определение.
-Характеристикой множества $M \in \mathfrak{M}_{n} \quad$ называется свободное от квадратов число р такое, что плошадь любого треугольника $\mathrm{ABC}$, где $A, B, C \in M$ соизмерима $\mathrm{c} \sqrt{p} .$ Обозначается char $\mathrm{M}={p} .$
+Для заданного множество $M \in \mathfrak{M}$ можно вычислить следующие величины,
+которые дают важную информацию о нём:
+\begin{enumerate}
+	\item
+		мощность;
+	\item
+		диаметр;
+	\item
+		характеристика;
+	\item
+		минимальное ребро.
+\end{enumerate}
 
-Определение. Ребром множества М будем называть любой отрезок $M_{1} M_{2}$, где $M_{1}, M_{2} \in M$
-Множество М характеристики р может быть размещено на решетке
+\paragraph{Определение.}
+Диаметром множества М будем называть наибольшее расстояние между любыми двумя точками множества $M$:
+$$
+	\operatorname{diam} M=\max_{A, B \in M} |A-B|
+	.
+$$
+
+\paragraph{Определение.}
+Характеристикой множества $M \in \mathfrak{M}_{n}$ называется свободное от квадратов число $р$ такое,
+что для любых $A, B, C \in M$ площадь треугольника $ABC$ соизмерима с $\sqrt{p}$.
+Пишут: $\operatorname{char}{M}={p}$.
+
+\paragraph{Определение.}
+Ребром множества $М$ будем называть любой отрезок $M_{1} M_{2}$, где $M_{1}, M_{2} \in M$.
+
+При исследовании множеств $M\in\mathfrak{M}_n$ часто бывает удобно ввести систему координат.
+
+\paragraph{Теорема о сетке} [3, теорема 4].
+Множество $М\in\mathfrak{M}_n$, $\operatorname{char} M = p$ может быть размещено на решётке
 $$
 	\left\{\left(\frac{a_{i}}{2 m} ; \frac{b_{i} \sqrt{p}}{2 m}\right)\right\}
+	,
 $$
-где $a_{i}, b_{i} \in \mathbb{Z}(\mathbb{Z}$ "--- множество целых чисел), в качестве $m$ можно взять длину любого ребра множества $\mathrm{M}$ [3, Теорема 4].
-Приведем несколько известных фактов о множествах $M \in \mathfrak{M}_{n}:$
+где $a_{i}, b_{i} \in \mathbb{Z}$, в качестве $m$ можно взять длину любого ребра множества $M$.
+
+Возникает закономерный вопрос о существовании множества $M\in\mathfrak{M}_n$
+с заданными характеристикой, диаметром и минимальным ребром.
+Известны следующие факты.
 \begin{itemize}
 	\item
-		Для любого $n \in \mathbb{N}, n \geq 3$ выполнено $\mathfrak{M}_{n} \neq \emptyset$ [3, Теорема 2].
+		Для любого $n \in \mathbb{N}, n \geq 3$ выполнено $\mathfrak{M}_{n} \neq \varnothing$ [3, теорема 2].
 	\item
@@
 	\item
-		Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого свободного от квадратов числа $р$ существует множество $M \in \mathfrak{M}_{n}$ такое, что char $\mathrm{M}={p}[3,$, Teopema 5 ].
-
+		Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого свободного от квадратов числа $р$
+		существует множество $M \in \mathfrak{M}_{n}$ такое, что $\operatorname{char}{M}={p}$ [3, теорема 5].
 	\item
-		Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого натурального числа $k$ существует множество $M \in \mathfrak{M}_{n}$, содержащее ребро ${k}$[5].
+		Для любой мощности $n \in \mathbb{N}, n \geq 3$ и любого натурального числа $k$ существует множество
+		$M \in \mathfrak{M}_{n}$, содержащее ребро ${k}$~[5].
 	\item
@@
 \end{itemize}
-В статье «О существовании специальных множеств с целочисленными расстояниями с ребром 1» получены следующие результаты исследования множества с характеристикой 2,3 и $5:$
-1. $\quad$ Не существует множества $M \in \mathfrak{M}_{3}$ характеристики 2 или $5 \mathrm{c}$ ребром $1 .$
-2. $\quad$ Существует бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{3}$, char M=3
-с ребром 1 и char $\mathrm{M}=5$ с ребром $2 .$ В данной работе рассматривается множество $M \in \mathfrak{M}_{3}$, char M=6 и char
+В работе [8] доказано, что
+не существует множества $M \in \mathfrak{M}_{3}$ характеристики 2 или 5 с ребром 1,
+но существует бесконечные семейства множеств $M_{i} \in \mathfrak{M}_{3}$ с $\operatorname{char} M=3$
+и ребром 1 и с $\operatorname{char} M=5$ с ребром $2$.
+В данной заметке продолжается исследование множеств $M \in \mathfrak{M}_{3}$ с ребром 1 или 2
+и обобщаются результаты [8].
+Доказывается, что любое множество $M \in \mathfrak{M}_{3}$ с ребром 1
+имеет харакеристику вида $4k+3$, $k\in\mathbb{N}$.
+Вместе с тем, для любого свободного от квадратов $p>1$ мы строим бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{3}$,
+$\operatorname{char} M=p$ с ребром 2.
 
-М=7 с ребром 1.
-Доказывается, что не существует множества $M \in \mathfrak{M}_{3} \mathrm{c}$ характеристикой вида $4 k+1,4 k+2$, при $k=0,1,2, \ldots$ Вместе с тем, мы строим бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{3}$, char $\mathrm{M}=4 k+3$, при $k=$ $0,1,2, \ldots$ с ребром $1 .$
-Утверждение 1.
-Не существует множества $M \in \mathfrak{M}_{3}$ характеристики $6 \mathrm{c}$ ребром $1 .$ Доказательство.
-По теореме о сетке введем систему координат так, что $\mathrm{O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{b \sqrt{6}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
-Пифагора:
+Для наглядности рассмотрим следующие частные случаи.
+
+\paragraph{Утверждение 1.}
+Не существует множества $M \in \mathfrak{M}_{3}$ характеристики 6  с ребром 1.
+\paragraph{Доказательство.}
+По теореме о сетке введем систему координат так, что
+${O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{b \sqrt{6}}{2}\right)$
+и $\{A, B, C\}= M$.
+По теореме Пифагора
 $$
-\begin{array}{c}
-\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{6}}{2}\right)^{2}=r^{2} \\
-1+6 b^{2}=4 r^{2}
-\end{array}
+	\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{6}}{2}\right)^{2}=r^{2}
+	,
 $$
-Очевидно, что в уравнении (1) левая часть - нечетное число при любом $b \in \mathbb{Z}$, а правая $-$ четное число при любом $r \in \mathbb{Z} .$ Следовательно, уравнение
-(1) неразрешимо в целых числах. $\mathbf{y}_{\text {тверждение }}$ 2. Существует бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{n}$ характеристики 7 с ребром $1 .$
+$$
+	1+6 b^{2}=4 r^{2}\eqno{(1)}
+	.
+$$
+Очевидно, что в уравнении (1) левая часть "--- нечётное число при любом $b \in \mathbb{Z}$,
+а правая "--- чётное число при любом $r \in \mathbb{Z}$.
+Следовательно, уравнение (1) неразрешимо в целых числах.
 
-Доказательство.
-По теореме о сетке введем систему координат так, что $\mathrm{O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{{b} \sqrt{7}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме
-Пифагора:
+\paragraph{Утверждение 2.}
+Существует бесконечное семейство множеств $M_{i} \in \mathfrak{M}_{n}$ характеристики 7 с ребром 1.
+
+\paragraph{Доказательство.}
+По теореме о сетке введем систему координат так, что
+${O}(0 ; 0), A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{{b} \sqrt{7}}{2}\right)$ и
+и $\{A, B, C\}= M$.
+По теореме Пифагора:
 $$
-\begin{array}{c}
-\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{7}}{2}\right)^{2}=r^{2} \\
+	\left(\frac{1}{2}\right)^{2}+\left(\frac{b \sqrt{7}}{2}\right)^{2}=r^{2}
+$$
+$$
 	1+7 b^{2}=4 r^{2}
-\end{array}
+	.
+	\eqno{(2)}
 $$
-Уравнение (2) имеет решение в целых числах, если $\left(\frac{1+7 b^{2}}{4}\right)-$ полный квадрат. Например, $b^{2}=9, {r}^{2}=16 .$ Таким образом, $(4 ; 3)-$ решение уравнения (2).
+Уравнение (2) имеет решение в целых числах, если $\frac{1+7 b^{2}}{4}$ есть полный квадрат.
+Например, $b^{2}=9, {r}^{2}=16$.
+Таким образом, $(4 ; 3)$ "--- решение уравнения (2).
 Уравнение (2) равносильно следующему уравнению:
-$4 r^{2}-7 b^{2}=1$
-Введем замену: $n=2 r .$ Получим
+$$
+	4 r^{2}-7 b^{2}=1
+	.
+	\eqno{(3)}
+$$
+Введём замену: $n=2 r$.
+Получим
 $$
 	n^{2}-7 b^{2}=1
+	\eqno{(4)}
 $$
-Уравнение (4) называется уравнением Пелля. Определение. Уравнение Пелля - диофантово уравнение вида
+Уравнение (4) называется уравнением Пелля.
+
+\paragraph{Определение.}
+Уравнение Пелля "--- диофантово уравнение вида
 $$
 x^{2}-m y^{2}=1
+,
 $$
-где $\mathrm{m}-$ натуральное число, не являющееся квадратом.
-Ясно, что уравнение Пелля имеет тривиальное решение $\mathrm{x}=\pm 1, \mathrm{y}=0 .$ Теорема 1.
-Любое уравнение Пелля имеет нетривиальное решение [7].
-Теорема 2.
-Все нетривиальные положительные решения получаются многократным умножением основного решения на себя [7].
+где $\mathrm{m}$ "--- натуральное число, не являющееся квадратом.
+Ясно, что уравнение Пелля имеет тривиальное решение $\mathrm{x}=\pm 1, \mathrm{y}=0$.
+\paragraph{Теорема 1} [7].
+Любое уравнение Пелля имеет нетривиальное решение.
+\paragraph{Теорема 2} [7].
+Все нетривиальные положительные решения получаются многократным умножением основного решения на себя.
 
@@
 $$
-(8-3 \sqrt{7})^{{k}}(8+3 \sqrt{7})^{{k}}=1^{{k}} \quad {k} \in \mathbb{N}
+	(8-3 \sqrt{7})^{{k}}(8+3 \sqrt{7})^{{k}}=1^{{k}}, \quad {k} \in \mathbb{N}
+	\eqno{(5)}
 $$
-Заметим, что при возведении в четные степени \textrm{K} , ~ п о л у ч а ю т с я ~ н е ч е т н ы е ~
+Заметим, что при возведении в четные степени $k$ получаются нечетные
+$n$, и, следовательно, $r$ "--- нецелые числа.
+Поэтому будем возводить выражение
+(5) в нечётные степени: ${k}=2 m+1$, $m\in\mathbb{N}$.
+Приведем несколько решений уравнения (4).
 
-n, а следовательно, $r-$ нецелые числа. Поэтому будем возводить выражение
-(5) в степени вида ${k}=2 \mathrm{~m}+1, \mathrm{~m}=0,1,2 \ldots$
-Приведем несколько решений уравнения (4):
-При ${k}=3:(2024-765 \sqrt{7})(2024+765 \sqrt{7})=1 \quad$ пара $\quad$ чисел
-(2024;765) является решением уравнения (4), следовательно, (1012;765) решение уравнения (3).
+При ${k}=3$:
 $$
-\text { При } {k}=5:(514088-194307 \sqrt{7})(514088+194307 \sqrt{7})=1
+	(2024-765 \sqrt{7})(2024+765 \sqrt{7})=1
+	,
 $$
-(514088; 194307) - решением уравнения (4), (257044; 194307) - решение уравнения (3).
+пара чисел
+$(2024;765)$ является решением уравнения (4), следовательно, $(1012;765)$ есть решение уравнения (3).
 
-Продолжая возводить в нечетные степени обе части уравнения, получим бесконечную последовательность решений для исходного уравнения (2):
+При $k=5$:
+$$
+	(514088-194307 \sqrt{7})(514088+194307 \sqrt{7})=1
+	.
+$$
+$(514088; 194307)$ "--- решение уравнения (4), $(257044; 194307)$ "--- решение уравнения (3).
+
+Продолжая возводить в нечетные степени обе части уравнения,
+получим бесконечную последовательность решений для исходного уравнения (2):
 $$
@@
 Всякое множество $M \in \mathfrak{M}_{3}$ с минимальным ребром 1 имеет характеристику вида
-$p=4k+3, k=0,1,2 \ldots$
+$p=4k+3$, $k\in\mathbb{N}$.
 \paragraph{Доказательство.}
-Пусть минимальное ребро множества $\mathfrak{M}_{3}$ равно $1 .$ Воспользуемся теоремой о сетке и введем систему координат так, что О (0;0),
-$A\left(-\frac{1}{2} ; 0\right), B\left(\frac{1}{2} ; 0\right), C\left(0 ; \frac{b \sqrt{p}}{2}\right)$ и $A, B, \mathrm{C} \in M .$ По теореме Пифагора:
+Пусть минимальное ребро множества $\mathfrak{M}_{3}$ равно 1.
+Воспользуемся теоремой о сетке и введем систему координат так, что $О (0;0)$,
+$A\left(-\frac{1}{2} ; 0\right)$, $B\left(\frac{1}{2} ; 0\right)$, $C\left(0 ; \frac{b \sqrt{p}}{2}\right)$
+и $\{A, B, C\}= M$. По теореме Пифагора:
 $$
-\begin{array}{c}
-\left(\frac{1}{2}\right)^{2}+\left(\frac{{b} \sqrt{{p}}}{2}\right)^{2}={r}^{2} \\
+	\left(\frac{1}{2}\right)^{2}+\left(\frac{{b} \sqrt{{p}}}{2}\right)^{2}={r}^{2}
+$$
+$$
 	1+p b^{2}=4 r^{2}
-\end{array}
 $$
 Для того, чтобы ${r}^{2}$ было целым числом необходимо,
-чтобы левая часть уравнения была кратна 4 и ${p}, {b}$ были  нечетными.
+чтобы левая часть уравнения была кратна 4 и ${p}, {b}$ были  нечётными.
 Следовательно,  $p b^{2}\equiv3(\operatorname{mod}4)$.
@@
 \paragraph{Утверждение 4.}
-Для всякого свободного от квадратов числа $р$ существует бесконечное семейство множеств $\{M_j\}\in\mathfrak{M}_3$ с ребром 2
+Для всякого свободного от квадратов числа $p>1$ существует бесконечное семейство множеств $\{M_j\}\subset\mathfrak{M}_3$ с ребром 2
 таких, что $\operatorname{char}M_j = p$.
@@
 $$
-1+\frac{\mathrm{pb}^{2}}{16}={r}^{2}\eqno{(8)}
+1+\frac{{pb}^{2}}{16}={r}^{2}\eqno{(8)}
 $$
@@
 $$
-Очевидно, равенство (9) является уравнением Пелля и имеет нетривиальное решение при любом свободном от квадратов числе ${p}$.
+Очевидно, равенство (9) является уравнением Пелля и имеет нетривиальное решение при любом свободном от квадратов числе $p>1$.
+
+\paragraph{Замечание.}
+Тот факт, что при $p=1$ уравнение (9) имеет только тривиальное решение,
+сам по себе не позволяет опровергнуть существование множеств
+$M\in\mathfrak{M}_3$, $\operatorname{char}M=1$ c минимальным ребром 2,
+поскольку (в отличие от ситуации с минимальным ребром 1)
+искомая третья точка может лежать не только на серединном перпендикуляре к ребру,
+но и на гиперболе специального вида.
 
@@
 
+8. Лушина Е. А. О существовании специальных множеств с целочисленными расстояниями с ребром 1. --- в печати.
+
 On characteristic of planar integral point sets with edges 1 and 2

### 2021-05-18 vestnikpmm20210518/Lushina_Avdeev.tex
[pmm2021] Final

@@
 Аннотация:
+	Изучаются множества точек на плоскости с целочисленными расстояниями.
+	Доказывается, что если некоторые две точки такого множества расположены на расстоянии 1,
+	то множество имеет характеристику $4k+3$ для некоторого натурального $k$.
+	Для любой характеристики $p>1$ строится множество точек на плоскости с целочисленными расстояниями,
+	среди которых расстояние 2.
 
@@
 Abstract:
+	The note is devoted to planar integral point sets.
+	We prove that any planar integral point sets which contains distance 1
+	has charactristic $4k+3$ for some integer $k$.
+	For any given characteristic $p>1$ we construct a planar integral point set
+	with distance 2.
 
@@
 
-Обозначим через $\mathfrak{M}$ множество таких подмножеств $M=\left\{M_{1}, M_{2}, \ldots\right\} \subset$ $\mathbb{R}^{2}$,
-что $\left|M_{i}-M_{j}\right| \in \mathbb{N}$ для всех $\mathrm{i}, \mathrm{j}$,
+Обозначим через $\mathfrak{M}$ множество таких подмножеств $$M=\left\{M_{1}, M_{2}, \ldots\right\} \subset \mathbb{R}^{2},$$
+что $\left|M_{i}-M_{j}\right| \in \mathbb{N}$ для всех ${i},{j}$,
 где $\mathbb{N}$ "--- множество натуральных чисел,
@@
 
-Доказано [1,2], что всякое бесконечное множество $M \in \mathfrak{M}$ содержится в некоторой прямой $L \subset \mathbb{R}^{2}$
-и потому может быть отождествлено с некоторым подмножеством множества $\mathbb{Z}$ целых чисел,
+Всякое бесконечное множество $M \in \mathfrak{M}$ содержится [1,2] в некоторой прямой $L \subset \mathbb{R}^{2}$
+и потому может быть отождествлено (с точностью до сдвига) с некоторым подмножеством множества $\mathbb{Z}$ целых чисел,
 которое, в свою очередь, является объектом изучения теории чисел, но не геометрии.
@@
 но существует бесконечные семейства множеств $M_{i} \in \mathfrak{M}_{3}$ с $\operatorname{char} M=3$
-и ребром 1 и с $\operatorname{char} M=5$ с ребром $2$.
+c ребром 1 и с $\operatorname{char} M=5$ с ребром $2$.
 В данной заметке продолжается исследование множеств $M \in \mathfrak{M}_{3}$ с ребром 1 или 2
@@
 По теореме о сетке введем следующую систему
-координат:  $O(0 ; 0), A(-1 ; 0), B(1 ; 0), C\left(0 ; \frac{b \sqrt{p}}{4}\right)$ и $A, B, C \in M$.
+координат:  $O(0 ; 0), A(-1 ; 0), B(1 ; 0), C\left(0 ; \frac{b \sqrt{p}}{4}\right)$
+и $\{A, B, C\} = M$.
 По теореме Пифагора:
@@
 $$
-1+\frac{{pb}^{2}}{16}={r}^{2}\eqno{(8)}
+1+\frac{{pb}^{2}}{16}={r}^{2}\eqno{(6)}
 $$
-Правая часть уравнения (8) является целым числом в том случае, если $b=4n$, $n\in\mathbb{N}$.
+Правая часть уравнения (6) является целым числом в том случае, если $b=4n$, $n\in\mathbb{N}$.
 Тогда
@@
 $$
-r^{2}-pn^{2}=1\eqno{(9)}
+r^{2}-pn^{2}=1\eqno{(7)}
 $$
-Очевидно, равенство (9) является уравнением Пелля и имеет нетривиальное решение при любом свободном от квадратов числе $p>1$.
+Очевидно, равенство (7) является уравнением Пелля и имеет нетривиальное решение при любом свободном от квадратов числе $p>1$.
 
 \paragraph{Замечание.}
-Тот факт, что при $p=1$ уравнение (9) имеет только тривиальное решение,
+Тот факт, что при $p=1$ уравнение (7) имеет только тривиальное решение,
 сам по себе не позволяет опровергнуть существование множеств

### 2021-08-28 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Declare the absense of conflicts
Minor edits

@@
-\documentclass[12pt]{article}
+\documentclass[11pt,A4paper]{article}
 \usepackage[utf8]{inputenc}
@@
 %\bibliographystyle{unsrt}
-\textheight=240mm \textwidth=170mm
-\hoffset=-17mm
-\voffset=-17mm
+
+\usepackage{geometry} % Меняем поля страницы.
+\geometry{left=1cm}% левое поле
+\geometry{right=1cm}% правое поле
+\geometry{top=1cm}% верхнее поле
+\geometry{bottom=1.5cm}% нижнее поле
 
@@
 Planar integral point sets in general position are very difficult to find;
-the first known examples were presented in~\cite{kreisel2008heptagon}.
+the first known examples of such sets with cardinaity 7 were presented in~\cite{kreisel2008heptagon}.
 As for now, there is no known example of PIPS of cardinality 8 in general position.
@@
 Here $q$ is the characteristic of the PIPS;
-every PIPS can be represented in such way~\cite{our-vmmsh-2018-translit}[Theorem 4].
+every PIPS can be represented in such way~\cite[Theorem 4]{our-vmmsh-2018-translit}.
 
@@
 
+The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+
 \end{document}

### 2021-10-31 semi-general-suplinear-bound/blg-GOST.tex
[CHEB] Now the paper is accepted, and we can mention the author
Revert "We should not mention the author due to referring policy"

This reverts commit e2fa42531fa8f4e2330024a62bf335bfd5d8a0ef.

@@
 \bibitem{our-vmmsh-2018}
-Author et al., 2018
-%Авдеев Н. Н., Семёнов Е. М. Множества точек с целочисленными расстояниями на
-%плоскости и в евклидовом пространстве // Математический форум (Итоги науки. Юг
-%России). — 2018. — с. 217—236.
+Авдеев Н. Н., Семёнов Е. М. Множества точек с целочисленными расстояниями на
+плоскости и в евклидовом пространстве // Математический форум (Итоги науки. Юг
+России). — 2018. — с. 217—236.
 \bibitem{kurz2008bounds}
@@
 \bibitem{my-pps-linear-bound-2019}
-Author, 2019
-%Avdeev N. On existence of integral point sets and their diameter bounds. — 2019. — arXiv:
-%1906.11926.
+Avdeev N. On existence of integral point sets and their diameter bounds. — 2019. — arXiv:
+1906.11926.
 \bibitem{bat2018number}
@@
 \bibitem{our-pmm-2018}
-Author, 2018
-%Авдеев Н. Н. Об отыскании целоудалённых множеств специального вида // Актуальные
-%проблемы прикладной математики, информатики и механики - сборник трудов Международной
-%научной конференции. — Научно-исследовательские публикации, 2018. — с. 492—498.
+Авдеев Н. Н. Об отыскании целоудалённых множеств специального вида // Актуальные
+проблемы прикладной математики, информатики и механики - сборник трудов Международной
+научной конференции. — Научно-исследовательские публикации, 2018. — с. 492—498.
 \bibitem{our-ped-2018}
-Author, 2018
-%Авдеев Н. Н. On integral point sets in special position // Некоторые вопросы анализа,
-%алгебры, геометрии и математического образования: материалы международной молодежной
-%научной школы «Актуальные направления математического анализа и смежные вопросы». —
-%2018. — т. 8. — с. 5—6.
+Авдеев Н. Н. On integral point sets in special position // Некоторые вопросы анализа,
+алгебры, геометрии и математического образования: материалы международной молодежной
+научной школы «Актуальные направления математического анализа и смежные вопросы». —
+2018. — т. 8. — с. 5—6.
 \bibitem{kurz2005characteristic}

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] Start the translation: Conclusion & Acknowledgements

--- 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
@@
 
-\section{Introduction}
+\section{Введение}
 
@@
 
-\section{Conclusion}
-The presented bound is the first special lower bound for sets in semi-general position.
-Thus, we have not accepted the challenge to make the constant in Theorem~\ref{thm:main_result} as large as possible,
-in order to keep the ideas of the proof clear and understandable.
-The more thorough research can be done in the future to improve the constant.
-However, the upper and lower bounds are still not exact.
+\section{Заключение}
+Полученная в данной статье оценка ---
+первая специальная оценка диаметра снизу для множества с целочисленными расстояниями получобщего положения.
+Поэтому автор не преследовал цели получить как можно б\'ольшую константу в теореме~\ref{thm:main_result},
+предпочитая сохранить ясность и прозрачность доказательства.
+Очевидно, что в результате более тщательного исследования эта константа может быть увеличена.
+Кроме того, и верхние, и нижние оценки по-прежнему, увы, остаются далеки от точных.
 
-\section{Acknowledgements}
-Author thanks Dr. Prof. E.M. Semenov for proofreading and valuable advices and Dr. A.S. Usachev for proofreading.
+\section{Благодарности}
+Автор благодарит проф. Е.М. Семёнова за вычитку и статьи и ценные советы;
+к.ф.-м.н. А.С. Усачёва за вычитку английского варианта статьи.
 

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
Translate the intro

@@
 
-An \textit{integral point set} in a plane is a point set $M$ such that all the usual (Euclidean) distances between the
-points of $M$ are integers and $M$ is not situated on a straight line.
-Every integral point set consists of a finite number of points~\cite{anning1945integral,erdos1945integral};
-thus, we denote the set of all planar integral point sets of $n$ points by
-$\mathfrak{M}(2,n)$ (using the notation in~\cite{our-vmmsh-2018})
-and define the diameter of $M\in\mathfrak{M}(2,n)$ in the following natural way:
+\textit{Множеством с целочисленными расстояниями (МЦР)} на плоскости
+называется такое множество точек $M$, что расстояние (евклидово) между любыми двумя точками из $M$
+есть целое число, и при этом $M$ не является подмножеством никакой прямой.
+Каждое МЦР состоит лишьиз конечного числа точек~\cite{anning1945integral,erdos1945integral};
+поэтому мы можем обозначить множество всех множеств из $n$ точек с целочисленными расстояниями через
+$\mathfrak{M}(2,n)$ (придерживаясь обозначений~\cite{our-vmmsh-2018} и в целом для краткости изложения)
+и естественным образом определить диаметр множества $M\in\mathfrak{M}(2,n)$:
 \begin{equation}
@@
 \end{equation}
-where $|AB|$ denotes the Euclidean distance.
-The symbol $\# M$ will be used for cardinality of $M$, that is the number of points in $M$ in our case.
+где $|AB|$ обозначает обычное (евклидово) расстояние между точками $A$ и $B$.
+Через $\# M$ будем обозначать мощность множества $M$, которая в нашем случае всегда совпадает с количеством точек в $M$.
 
-Since every integral point set can obviously be dilated to a set of larger diameter,
-minimal possible diameters of sets of given cardinality are of interest.
-To be precise,
-the following function was introduced in~\cite{kurz2008bounds,kurz2008minimum}:
+Поскольку любое МЦР можно очевидным образом растянуть
+до множества большего диаметра,
+интерес представляет в первую очередь минимальный возможный диаметр при заданной мощности множества.
+Это рассуждение формализуется следующей функцией, введённой в~\cite{kurz2008bounds,kurz2008minimum}:
 \begin{equation}
@@
 
-It turned out to be very easy to construct a planar integral point set of $n$ points with $n-1$ collinear ones and one point out of the line
-(so-called \textit{facher} sets);
-the same holds for 2 points out of the line (we refer the reader to~\cite{antonov2008maximal}, where some of such sets are called \textit{crabs})
-and even for 4 points out of the line~\cite{huff1948diophantine}.
-For $9\leq n\leq 122$, the minimal possible diameter is attained at a facher set~\cite{kurz2008bounds}.
+Выяснилось, что достаточно просто сконструировать МЦР на плоскости,
+все точки которого, кроме одной, лежат на некоторой прямой
+(такие множества называются \textit{веерными});
+аналогично --- для 2 точек вне прямой (см., напр., ~\cite{antonov2008maximal}, где такие множества названы \textit{крабовыми})
+и даже для 4 точек вне прямой~\cite{huff1948diophantine}.
+Для $9\leq n\leq 122$ минимальный возможный диаметр МЦР мощности $n$ достигаетя именно на веерном множестве~\cite{kurz2008bounds}.
 
 \begin{Definition}
-	A set $M\in\mathfrak{M}(2,n)$ is called to be in \textit{semi-general position},
-	if no three points of $M$ are collinear.
-	The set of all planar integral point sets in semi-general position
-	is denoted by $\overline{\mathfrak{M}}(2,n)$.
+	Множество $M\in\mathfrak{M}(2,n)$ есть МЦР \textit{полуобщего положения},
+	если никакие три точки из $M$ не лежат на одной прямой.
+	Кратко будем писать, что $M\in \overline{\mathfrak{M}}(2,n)$.
 \end{Definition}
 
-Furthermore, the constructions of integral point sets in semi-general position of arbitrary cardinality
-appeared in~\cite{harborth1993upper};
-such sets are situated on a circle.
-Also, there is a sophisticated construction of a circular integral point set of arbitrary cardinality
-that gives the possible numbers of odd integral distances
-between points in the plane~\cite{piepmeyer1996maximum}.
+В~\cite{harborth1993upper} приводится конструкция МЦР полуобщего положения для любой наперёд заданной мощности;
+получаемое МЦР содержится в некоторой окружности.
+Кроме того, в~\cite{piepmeyer1996maximum} приведена изящная конструкция МЦР $M$ любой наперёд заданной мощности,
+содержащегося в окружности,
+для любого возможного количества нечётных расстояний между точками $M$.
 
 \begin{Definition}
-	A set $M\in\overline{\mathfrak{M}}(2,n)$ is said to be a set in \textit{general position},
-	if no four points of $M$ are concyclic.
-	The set of all planar integral point sets in general position
-	is denoted by $\dot{\mathfrak{M}}(2,n)$.
+	Множество $M\in\overline{\mathfrak{M}}(2,n)$ есть МЦР \textit{общего положения},
+	если никакие четыре точки из $M$ не лежат на одной окружности.
+	Кратко будем писать, что $M\in \dot{\mathfrak{M}}(2,n)$.
 \end{Definition}
 
-It remains unknown whether there are integral points sets in general position of arbitrary cardinality;
-however, some sets $M\in \dot{\mathfrak{M}}(2,7)$ are known~\cite{kreisel2008heptagon,kurz2013constructing}.
+Остаётся неясным, любую ли мощность может имет МЦР общего положения;
+однако, известны~\cite{kreisel2008heptagon,kurz2013constructing} примеры МЦР $M\in \dot{\mathfrak{M}}(2,7)$.
 
-The inequality
+Очевидно неравенство
 \begin{equation*}
@@
 \end{equation*}
-where
+где
 $
@@
 $
-and
+и
 $
 	\dot{d}(2,n) = \min_{M\in\dot{\mathfrak{M}}(2,n)} \operatorname{diam} M
-$,
-is obvious; however, a more interesting relation holds:
+$.
+Выполнено, однако, и более интересное соотношение:
 \begin{equation*}
@@
 \end{equation*}
-The upper bound is established in~\cite{harborth1993upper}.
-The lower bound firstly appeared in~\cite{solymosi2003note};
-the largest known value for $c_1$ is $5/11$ for $n\geq 4$~\cite{my-pps-linear-bound-2019}.
+Верхняя оценка установлена в~\cite{harborth1993upper}.
+Нижняя оценка впервые была получена в~\cite{solymosi2003note};
+наибольшее известное значение для $c_1$ равно $5/11$ (для $n\geq 4$)~\cite{my-pps-linear-bound-2019}.
 
-There are some bounds for minimal diameter of planar integral point sets in some special positions.
-Assuming that a planar integral point set contains many collinear points,
-the following result holds.
+Известны некоторые оценки на диаметр МЦР специального вида.
+В частности, для множеств, в которых достаточно много точек лежит на одной прямой,
+доказана следующая теорема.
 \begin{Theorem}~\cite[Theorem 4]{kurz2008minimum}
-	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
-	at least $n^\delta$ collinear points, there exists a $n_0 (\varepsilon)$
-	such that for all $n \geq n_0 (\varepsilon)$ we have
+	Для любых $\delta > 0$, $\varepsilon > 0$ и множества $P\in\mathfrak{M}(2,n)$,
+	в котором не менее $n^\delta$ лежат на некоторой прямой, существует число $n_0 (\varepsilon)$
+	такое, что для всех $n \geq n_0 (\varepsilon)$ верно неравенство
 	\begin{equation}
@@
 \end{Theorem}
-For diameter bounds for circular sets, we refer the reader to~\cite{bat2018number}.
+Об оценках на диаметр МЦР,
+содержащихся в окружности, можно прочесть в~\cite{bat2018number}.
 
-Particular cases of planar integral point sets are also discussed
-in~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved},~\cite{our-pmm-2018},~\cite{our-ped-2018}.
-For generalization to higher dimensions and the appropriate bounds, see~\cite{kurz2005characteristic,nozaki2013lower}.
+Отдельные случаи МЦР на плоскости также обсуждались
+в~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved},~\cite{our-pmm-2018},~\cite{our-ped-2018}.
+За обобщением на пространства более высоких размерностей и соответствующими оценками
+отсылаем читателя к~\cite{kurz2005characteristic,nozaki2013lower}.
 
-In the present paper we give a special bound for planar integral point sets in semi-general position.
-The condition of semi-general position is crucial in the given proof.
+В настоящей статье мы доказываем отдельную оценку на диаметр множества с целочисленными расстояниями полуобщего положения.
+Требование полуобщего положения является существенным для доказательства.
 

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
Translate the names of the rest sections
[CHEB] Translate the main results

@@
 
-\section{Preliminary results}
+\section{Вспомогательные утверждения}
 
@@
 
-\section{The main result}
+\section{Основной результат}
 
@@
 	\label{thm:main_result}
-	For every integer $n \geq 3$ we have
+	Для всякого целого $n \geq 3$ выполнено неравенство
 	\begin{equation}
@@
 
-\begin{proof}[Proof]
-	For $n = 3$ we have $\overline{d}(2,3) = 1$.
-	Consider $M\in\overline{\mathfrak{M}}(2,n)$, $n \geq 4$, $\operatorname{diam} M = p$.
+\begin{proof}[Доказательство]
+	Для $n = 3$ имеем $\overline{d}(2,3) = 1$ (достигается на равностороннем треугольнике со стороной 1),
+	и утверждение теоремы очевидно.
+	Рассмотрим $M\in\overline{\mathfrak{M}}(2,n)$, $n \geq 4$, $\operatorname{diam} M = p$.
 
-	Let us choose points $M_1, M_2, M_3, M_4 \in M$
-	(points $M_2$ and $M_3$ may coincide, other points are distinct), such that
+	Выберем точки $M_1, M_2, M_3, M_4 \in M$
+	(точки $M_2$ и $M_3$ могут совпадать, остальные должны быть попарно различны) таким образом, что
 	\begin{equation}
@@
 
-	For $m \leq p^{2/5}$, Lemma~\ref{lem:count_of_points_on_hyperbolas} yields that
+	Для $m \leq p^{2/5}$ по лемме~\ref{lem:count_of_points_on_hyperbolas}
 	\begin{equation}
@@
 	\end{equation}
-	Equivalently,
+	Или, что то же самое,
 	\begin{equation}
@@
 
-	So, let us consider $m > p^{2/5}$.
-	Then for any $A,B \in M\setminus\{M_1\}$ the inequality $|AB| > p^{2/5}$ holds.
-	Due to Corollary~\ref{cor:solymosi_strip}, no three points of $M\setminus\{M_1\}$
-	are located in a strip of width $p^{1/5} / 2$.
+	Рассмотрим теперь $m > p^{2/5}$.
+	Тогда для любых точек $A,B \in M\setminus\{M_1\}$ выполнено неравенство $|AB| > p^{2/5}$.
+	Согласно следствию~\ref{cor:solymosi_strip}, никакие три точки из  $M\setminus\{M_1\}$
+	не лежат в полосе шириной $p^{1/5} / 2$.
 
@@
 
-	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square with side length $p$.
-	Let us partition this square into $q$ strips, $2p^{4/5} \leq q < 2p^{4/5} + 1$, each of width at most $p^{1/5} / 2$.
-	Every strip contains at most two points of  $M\setminus\{M_1\}$,
-	thus
+	В силу леммы~\ref{lem:square_container} множетво $M$ расположено в квадрате со стороной $p$.
+	покроем этот квадрат объединением $q$ полос, $2p^{4/5} \leq q < 2p^{4/5} + 1$ так,
+	что ширина каждой полосы не превышает $p^{1/5} / 2$.
+	Каждая из полученных полос содержит не более двух точек из  $M\setminus\{M_1\}$,
+	поэтому
 	\begin{equation}
@@
 	\end{equation}
-	The latter inequality holds because $\overline{d}(2,n) \geq 4$ for $n\geq 4$~~\cite{kurz2008minimum}
+	Последнее неравенство выполнено в силу того, что $\overline{d}(2,n) \geq 4$ для $n\geq 4$~~\cite{kurz2008minimum}
 	% TODO: add two more links from the kurz2008minimum's intro in case of lack of references
-	and $4^{4/5}>3$.
-	From the inequality~\eqref{eq:strips_4_5} one can easily derive that
+	и $4^{4/5}>3$.
+	Из неравенства~\eqref{eq:strips_4_5} легко получаем, что
 	\begin{equation}
@@
 \begin{Remark}
-	The following result appeared in~\cite{solymosi2003note} as Corollary 1:
+	Известен следующий факт~\cite[Следствие 1]{solymosi2003note}:
 \end{Remark}
@@
 \begin{lemma}
-	For $H \in \overline {\mathfrak{M}}(2,n)$, the minimal
-	distance in H is at least $n^{1/3}$.
+	Пусть $H \in \overline {\mathfrak{M}}(2,n)$.
+	Тогда минимальное расстояние между точками $H$ не менее $n^{1/3}$.
 \end{lemma}
-Applying the same technique, one can easily derive that
+Применяя аналогичные приёмы, можно получить, что
 \begin{equation}
@@
 \end{equation}
-which leads to the bound
+откуда выводится оценка
 \begin{equation}
@@
 \end{equation}
-which is less than the one from Theorem~\ref{thm:main_result}.
+которая, очевидно, слабее оценки в теореме~\ref{thm:main_result}.
 

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] Translate the preliminary results

@@
 
-In this section, we give some lemmas which will be used for the proof.
+В этом параграфе приводятся определения и леммы, которые в дальнейшем потребуются
+для доказательства основного результата.
 
@@
 	\cite[Observation 1]{solymosi2003note}
-	If a triangle $T$ has integer side lengths $a \leq b \leq c$,
-	then the minimal height $m$ of it is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
+	Пусть стороны треугольник $T$ есть целые числа $a \leq b \leq c$,
+	тогда наимаеньшая из высот $m$ этого треугольника не менее $\left(a - \frac{1}{4}\right)^{1/2}$.
 \end{lemma}
@@
 \begin{Definition}
-	The part of a plane between two parallel straight lines with distance $\rho$ between them
-	is called a strip of width $\rho$.
+	Полосой ширины $\rho$ будем называть часть плоскости,
+	заключённую между двумя параллельными прямыми,
+	отстоящими друг от друга на расстояние $\rho$.
 \end{Definition}
@@
 	\cite{smurov1998stripcoverings}
-	If a triangle $T$ with minimal height $\rho$ is situated in a strip,
-	then the width of the strip is at least $\rho$.
+	Пусть треугольник $T$, наименьшая из высот которого равна $\rho$, расположен в полосе.
+	Тогда ширина этой полосы не менее $\rho$.
 \end{lemma}
@@
 	\label{cor:solymosi_strip}
-	If a triangle $T$ with integer side lengths $a \leq b \leq c$ is situated in a strip,
-	then the width of the strip is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
+	Пусть треугольник $T$ с целыми сторонами $a \leq b \leq c$ лежит в некоторой полосе,
+	тогда ширина этой полосы не менее $\left(a - \frac{1}{4}\right)^{1/2}$.
 \end{Corollary}
@@
 	\label{lem:square_container}
-	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
-	Then $M$ is situated in a square of side length $d$.
+	Пусть $M\in\mathfrak{M}(2,n)$ и $\operatorname{diam} M = d$.
+	Тогда $M$ расположено в некотором квадрате со стороной $d$.
 \end{lemma}
@@
 	\cite[Definition 2.5]{my-pps-linear-bound-2019}
-	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
-	the line through $M_1$ and $M_2$,
-	and the perpendicular bisector of line segment $M_1 M_2$.
+	Будем называть \textit{крестом} точек $M_1$ и $M_2$ и обозначать через $cr(M_1,M_2)$ объединение двух прямых:
+	прямой, проходящей через точки $M_1$ и $M_2$,
+	и серединного перпендикуляра к отрезку $M_1 M_2$.
 \end{Definition}
@@
 	\label{lem:no_distance_one}
-	Each set $M\in\mathfrak{M}(2,n)$
-	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
-	consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
-	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
+	Каждое множество $M\in\mathfrak{M}(2,n)$,
+	для некоторых точек которого $M_1,M_2 \in M$ выполнено равенство $|M_1 M_2|=1$,
+	состоит из $n-1$ точек, включая $M_1$ и $M_2$, лежащих на некоторой прямой $m$,
+	и ещё одной точки, лежащей вне $m$ на серединном перпендикуляре к отрезку $M_1 M_2$.
 \end{lemma}
@@
 	\label{lem:count_of_points_on_hyperbolas}
-	Let $\{M_1, M_2, M_3, M_4\} \subset M\in\overline{\mathfrak{M}}(2,n)$
-	(points $M_2$ and $M_3$ may coincide, other points are distinct), $n\geq 4$.
-	Then $\# M \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
+	Пусть $\{M_1, M_2, M_3, M_4\} \subset M\in\overline{\mathfrak{M}}(2,n)$
+	(точки $M_2$ и $M_3$ могут совпадать, остальные попарно различны), $n\geq 4$.
+	Тогда $\# M \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
 \end{lemma}
@@
 \begin{Remark}
-	Lemma~\ref{lem:count_of_points_on_hyperbolas} is one of the variations of~\cite{erdos1945integral}.
+	Лемма~\ref{lem:count_of_points_on_hyperbolas} явлется вариацией~\cite{erdos1945integral}.
 \end{Remark}
 
-\begin{proof}[Proof]
-	Each point $N\in M$ satisfies one of the following conditions:
+\begin{proof}[Доказательство]
+	Для каждой точки $N\in M$ выполнено одно из следующих условий:
 
-	a) $N$ belongs to $cr(M_1,M_2)$~--- overall at most 4 points;
+	a) $N$ принадлежит $cr(M_1,M_2)$~--- итого не более 4 точек (не более 2 на каждой из прямых);
 
-	b) $N$ belongs to $cr(M_3,M_4)$~--- overall at most 4 points;
+	b) $N$ принадлежит $cr(M_3,M_4)$~--- итого не более 4 точек (не более 2 на каждой из прямых);
 
-	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas
-	with one of $|M_3 M_4| - 1$ hyperbolas~--- overall at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
+	c) $N$ принадлежит пересечению одной из $|M_1 M_2| - 1$ гипербол
+	с одной из $|M_3 M_4| - 1$ гипербол~--- итого не более  $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ точек;
 
-	Due to Lemma~\ref{lem:no_distance_one} we have $|M_1 M_2| \geq 2$ and $|M_3 M_4| \geq 2$.
-	Since
+	По лемме~\ref{lem:no_distance_one} $|M_1 M_2| \geq 2$ и $|M_3 M_4| \geq 2$.
+	Из того, что
 	\begin{multline}
@@
 	\end{multline}
-	the assertion follows.
+	и следует утверждение леммы.
 \end{proof}

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] Fix typos

@@
 Полученная в данной статье оценка ---
-первая специальная оценка диаметра снизу для множества с целочисленными расстояниями получобщего положения.
-Поэтому автор не преследовал цели получить как можно б\'ольшую константу в теореме~\ref{thm:main_result},
+первая специальная оценка диаметра снизу для множества с целочисленными расстояниями полуобщего положения.
+Поэтому автор не преследовал цели получить как можно б\'{о}льшую константу в теореме~\ref{thm:main_result},
 предпочитая сохранить ясность и прозрачность доказательства.

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] Minor fixes

@@
 \begin{lemma}
-	\cite[Observation 1]{solymosi2003note}
+	\cite[Наблюдение 1]{solymosi2003note}
 	Пусть стороны треугольник $T$ есть целые числа $a \leq b \leq c$,
-	тогда наимаеньшая из высот $m$ этого треугольника не менее $\left(a - \frac{1}{4}\right)^{1/2}$.
+	тогда наименьшая из высот $m$ этого треугольника не менее $\left(a - \frac{1}{4}\right)^{1/2}$.
 \end{lemma}
@@
 \begin{Definition}
-	\cite[Definition 2.5]{my-pps-linear-bound-2019}
+	\cite[Определение 2.5]{my-pps-linear-bound-2019}
 	Будем называть \textit{крестом} точек $M_1$ и $M_2$ и обозначать через $cr(M_1,M_2)$ объединение двух прямых:

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] Minor fixes

@@
 \Cite{
-Н. Н. Авдеев Об оценках на диаметр плоских множеств с целочисленными расстояниями полуобщего положения // Чебышевcкий сборник, 2019, т.~\tom, вып.~\iss, с.~\pageref{beg}--\pageref{end}.
+Н. Н. Авдеев. Об оценках на диаметр плоских множеств с целочисленными расстояниями полуобщего положения // Чебышевcкий сборник, 2019, т.~\tom, вып.~\iss, с.~\pageref{beg}--\pageref{end}.
 }
@@
 	В силу леммы~\ref{lem:square_container} множетво $M$ расположено в квадрате со стороной $p$.
-	покроем этот квадрат объединением $q$ полос, $2p^{4/5} \leq q < 2p^{4/5} + 1$ так,
+	Покроем этот квадрат объединением $q$ полос, $2p^{4/5} \leq q < 2p^{4/5} + 1$ так,
 	что ширина каждой полосы не превышает $p^{1/5} / 2$.

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] Translate the captions of Remarks, Definitions and Corollaries

@@
 \usepackage{textcase}% Оформление страниц
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Автоматическая нумерация теорем (коль скоро подключен amsthm)
-
-% К сожалению, английского ``Definition'' в стилевом файле нет.
-% Добавляю по аналогии.
-\newtheorem{Definition}{\indent {\sc Definition}}
-\renewcommand{\theDefinition}{\rm \arabic{Definition}}
-
-\newtheorem{Corollary}{\indent {\sc Corollary}}
-\renewcommand{\theCorollary}{\rm \arabic{Corollary}}
-
-\newtheorem{Remark}{\indent {\sc Remark}}
-\renewcommand{\theRemark}{\rm \arabic{Remark}}
 
@@
 
-\begin{Definition}
+\begin{definition}
 	Множество $M\in\mathfrak{M}(2,n)$ есть МЦР \textit{полуобщего положения},
@@
 	Кратко будем писать, что $M\in \overline{\mathfrak{M}}(2,n)$.
-\end{Definition}
+\end{definition}
 
@@
 
-\begin{Definition}
+\begin{definition}
 	Множество $M\in\overline{\mathfrak{M}}(2,n)$ есть МЦР \textit{общего положения},
@@
 	Кратко будем писать, что $M\in \dot{\mathfrak{M}}(2,n)$.
-\end{Definition}
+\end{definition}
 
@@
 
-\begin{Definition}
+\begin{definition}
 	Полосой ширины $\rho$ будем называть часть плоскости,
@@
 	отстоящими друг от друга на расстояние $\rho$.
-\end{Definition}
+\end{definition}
 
@@
 
-\begin{Corollary}
+\begin{corollary}
 	\label{cor:solymosi_strip}
@@
 	тогда ширина этой полосы не менее $\left(a - \frac{1}{4}\right)^{1/2}$.
-\end{Corollary}
+\end{corollary}
 
@@
 
-\begin{Definition}
+\begin{definition}
 	\cite[Определение 2.5]{my-pps-linear-bound-2019}
@@
 	и серединного перпендикуляра к отрезку $M_1 M_2$.
-\end{Definition}
+\end{definition}
 
@@
 
-\begin{Remark}
+\begin{remark}
 	Лемма~\ref{lem:count_of_points_on_hyperbolas} явлется вариацией~\cite{erdos1945integral}.
-\end{Remark}
+\end{remark}
 
@@
 
-\begin{Remark}
+\begin{remark}
 	Известен следующий факт~\cite[Следствие 1]{solymosi2003note}:
-\end{Remark}
+\end{remark}
 

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] Translate captions of Theorems and Lemmas

@@
 доказана следующая теорема.
-\begin{Theorem}~\cite[Theorem 4]{kurz2008minimum}
+\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
 	Для любых $\delta > 0$, $\varepsilon > 0$ и множества $P\in\mathfrak{M}(2,n)$,
@@
 	\end{equation}
-\end{Theorem}
+\end{theorem}
 Об оценках на диаметр МЦР,
@@
 
-\begin{lemma}
+\begin{lemm}
 	\cite[Наблюдение 1]{solymosi2003note}
@@
 	тогда наименьшая из высот $m$ этого треугольника не менее $\left(a - \frac{1}{4}\right)^{1/2}$.
-\end{lemma}
+\end{lemm}
 
@@
 
-\begin{lemma}
+\begin{lemm}
 	\cite{smurov1998stripcoverings}
@@
 	Тогда ширина этой полосы не менее $\rho$.
-\end{lemma}
+\end{lemm}
 
@@
 
-\begin{lemma}
+\begin{lemm}
 	\cite[Lemma 4]{our-vmmsh-2018};
@@
 	Тогда $M$ расположено в некотором квадрате со стороной $d$.
-\end{lemma}
+\end{lemm}
 
@@
 
-\begin{lemma}
+\begin{lemm}
 	\cite[Theorem 3.10]{my-pps-linear-bound-2019}
@@
 	и ещё одной точки, лежащей вне $m$ на серединном перпендикуляре к отрезку $M_1 M_2$.
-\end{lemma}
+\end{lemm}
 
-\begin{lemma}
+\begin{lemm}
 	\label{lem:count_of_points_on_hyperbolas}
@@
 	Тогда $\# M \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4|$.
-\end{lemma}
+\end{lemm}
 
@@
 
-\begin{Theorem}
+\begin{theorem}
 	\label{thm:main_result}
@@
 	\end{equation}
-\end{Theorem}
+\end{theorem}
 
@@
 
-\begin{lemma}
+\begin{lemm}
 	Пусть $H \in \overline {\mathfrak{M}}(2,n)$.
 	Тогда минимальное расстояние между точками $H$ не менее $n^{1/3}$.
-\end{lemma}
+\end{lemm}
 Применяя аналогичные приёмы, можно получить, что

### 2021-10-31 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] Fix enhanced cite-s

@@
 доказана следующая теорема.
-\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
+\begin{theorem}~\cite[теорема 4]{kurz2008minimum}
 	Для любых $\delta > 0$, $\varepsilon > 0$ и множества $P\in\mathfrak{M}(2,n)$,
@@
 \begin{lemm}
-	\cite[Наблюдение 1]{solymosi2003note}
+	\cite[наблюдение 1]{solymosi2003note}
 	Пусть стороны треугольник $T$ есть целые числа $a \leq b \leq c$,
@@
 \begin{lemm}
-	\cite[Lemma 4]{our-vmmsh-2018};
-	\cite[Lemma 2.4]{my-pps-linear-bound-2019}
+	\cite[лемма 4]{our-vmmsh-2018};
+	\cite[лемма 2.4]{my-pps-linear-bound-2019}
 	\label{lem:square_container}
@@
 \begin{definition}
-	\cite[Определение 2.5]{my-pps-linear-bound-2019}
+	\cite[определение 2.5]{my-pps-linear-bound-2019}
 	Будем называть \textit{крестом} точек $M_1$ и $M_2$ и обозначать через $cr(M_1,M_2)$ объединение двух прямых:
@@
 \begin{lemm}
-	\cite[Theorem 3.10]{my-pps-linear-bound-2019}
+	\cite[теорема 3.10]{my-pps-linear-bound-2019}
 	\label{lem:no_distance_one}
@@
 \begin{remark}
-	Известен следующий факт~\cite[Следствие 1]{solymosi2003note}:
+	Известен следующий факт~\cite[следствие 1]{solymosi2003note}:
 \end{remark}

### 2021-11-01 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] After self-proofreading

@@
 есть целое число, и при этом $M$ не является подмножеством никакой прямой.
-Каждое МЦР состоит лишьиз конечного числа точек~\cite{anning1945integral,erdos1945integral};
+Каждое МЦР состоит лишь из конечного числа точек~\cite{anning1945integral,erdos1945integral};
 поэтому мы можем обозначить множество всех множеств из $n$ точек с целочисленными расстояниями через
@@
 и даже для 4 точек вне прямой~\cite{huff1948diophantine}.
-Для $9\leq n\leq 122$ минимальный возможный диаметр МЦР мощности $n$ достигаетя именно на веерном множестве~\cite{kurz2008bounds}.
+Для $9\leq n\leq 122$ минимальный возможный диаметр МЦР мощности $n$ достигается именно на веерном множестве~\cite{kurz2008bounds}.
 
@@
 
-Остаётся неясным, любую ли мощность может имет МЦР общего положения;
+Остаётся неясным, любую ли мощность может иметь МЦР общего положения;
 однако, известны~\cite{kreisel2008heptagon,kurz2013constructing} примеры МЦР $M\in \dot{\mathfrak{M}}(2,7)$.
@@
 	Для любых $\delta > 0$, $\varepsilon > 0$ и множества $P\in\mathfrak{M}(2,n)$,
-	в котором не менее $n^\delta$ лежат на некоторой прямой, существует число $n_0 (\varepsilon)$
+	в котором не менее $n^\delta$ точек лежат на некоторой прямой, существует число $n_0 (\varepsilon)$
 	такое, что для всех $n \geq n_0 (\varepsilon)$ верно неравенство
@@
 
-В настоящей статье мы доказываем отдельную оценку на диаметр множества с целочисленными расстояниями полуобщего положения.
+В настоящей статье мы доказываем отдельную оценку снизу на диаметр множества с целочисленными расстояниями полуобщего положения.
 Требование полуобщего положения является существенным для доказательства.
@@
 	\cite[наблюдение 1]{solymosi2003note}
-	Пусть стороны треугольник $T$ есть целые числа $a \leq b \leq c$,
-	тогда наименьшая из высот $m$ этого треугольника не менее $\left(a - \frac{1}{4}\right)^{1/2}$.
+	Пусть стороны треугольника $T$ есть целые числа $a \leq b \leq c$,
+	тогда наименьшая из высот треугольника $T$ не менее $\left(a - \frac{1}{4}\right)^{1/2}$.
 \end{lemm}
@@
 
-	По лемме~\ref{lem:no_distance_one} $|M_1 M_2| \geq 2$ и $|M_3 M_4| \geq 2$.
+	По лемме~\ref{lem:no_distance_one} выполнено $|M_1 M_2| \geq 2$ и $|M_3 M_4| \geq 2$.
 	Из того, что
@@
 
-	Для $m \leq p^{2/5}$ по лемме~\ref{lem:count_of_points_on_hyperbolas}
+	Если $m \leq p^{2/5}$, то по лемме~\ref{lem:count_of_points_on_hyperbolas}
 	\begin{equation}
@@
 
-	Рассмотрим теперь $m > p^{2/5}$.
+	Пусть теперь $m > p^{2/5}$.
 	Тогда для любых точек $A,B \in M\setminus\{M_1\}$ выполнено неравенство $|AB| > p^{2/5}$.
@@
 
-	В силу леммы~\ref{lem:square_container} множетво $M$ расположено в квадрате со стороной $p$.
+	В силу леммы~\ref{lem:square_container} множество $M$ расположено в квадрате со стороной $p$.
 	Покроем этот квадрат объединением $q$ полос, $2p^{4/5} \leq q < 2p^{4/5} + 1$ так,

### 2021-11-01 semi-general-suplinear-bound/CHEBSB_Avdeev_rus.tex
[CHEB] After Semenov's advice

@@
 на диаметр плоских множеств с целочисленными расстояниями.
-Полученная в данной статье оценка всё ещё не точна.
+%Полученная в данной статье оценка всё ещё не точна.
 }
@@
 to prove the first linear lower bound for diameter of a planar integral point set.
-We have to admit that our bound is still not exact.
+%We have to admit that our bound is still not exact.
 }
@@
 \begin{remark}
-	Лемма~\ref{lem:count_of_points_on_hyperbolas} явлется вариацией~\cite{erdos1945integral}.
+	Лемма~\ref{lem:count_of_points_on_hyperbolas} основана на работе~\cite{erdos1945integral}.
 \end{remark}
@@
 \section{Благодарности}
-Автор благодарит проф. Е.М. Семёнова за вычитку и статьи и ценные советы;
-к.ф.-м.н. А.С. Усачёва за вычитку английского варианта статьи.
+Автор благодарит проф. Е.М. Семёнова и к.ф.-м.н. А.С. Усачёва за обсуждение статьи и ценные советы.
 

### 2021-11-03 AvdMomZvol_rails/Article.tex
[AMZr] Edits from Usachev - p. 4

@@
 	\footnote{
-		This work was carried out at Voronezh State University and supported by the Russian Science
+		This work was carried out at Voronezh State University and supported by Russian Science
 		Foundation grant 19-11-00197.
@@
 	A planar integral point set (PIPS) is a set $\mathcal{P}$
-	of non-collinear points in plane $\mathbb{R}^{2}$ such that
+	of non-collinear points in the plane $\mathbb{R}^{2}$ such that
 	for any pair of points $P_{1}, P_{2} \in \mathcal{P}$
@@
 First of all, we can look at its dimension;
-then, we can count the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
-and is further called the cardinality;
-finally, we can naturally% TODO: naturally/essentially? cf. arxiv/1
+then, we can find its cardinality, which is always finite~\cite{anning1945integral,erdos1945integral};
+finally, we can naturally % TODO: naturally/essentially? cf. arxiv/1
 define the diameter of a finite point set.
@@
 \begin{dfn}
-	A diameter of the integral point set $\mathcal{P}$ is defined as
+	The diameter of an integral point set $\mathcal{P}$ is defined as
 	\begin{equation}
@@
 
-Every integral point set also has a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.
+Every integral point set has also a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.
 
 \begin{dfn}
-	The function $d(m, n)$ is the minimum possible diameter of
+	The function $d(m, n)$ is the minimal possible diameter of
 	the integral point set $\mathcal{P}$ of $n$ points in
@@
 we refer the reader to~\cite[Theorem 1]{kurz2008bounds} or to~\cite{our-vmmsh-2018}.
-We will discuss the following estimations presented at~\cite{kurz2008bounds}:
+We will discuss the following ones presented at~\cite{kurz2008bounds}:
 \begin{align}
 	d(m, 2m + 1) \leq 8\\
-	d(m, 2m + 2) \leq 13 \hypertarget{d2}\\
+	d(m, 2m + 2) \leq 13 \label{eq:d_m_2m+2}\\
 	d(m, 3m) \leq 109
@@
 	Let $\mathcal{P}$ be a planar integral point set consisting of
-	$n - 2$ points on line $l_1$ and two points $P_{1}$ and $P_{2}$ on a
+	$n - 2$ points on the line $l_1$ and two points $P_{1}$ and $P_{2}$ on a
 	parallel line $l_2$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
@@
 
-Firstly, we discuss the classification of planar integral points sets;
-then, we present some bounds for $d(m,n)$ based on planar integral point sets of particular types
-and provide some general constructions for such bounds.
+First, we discuss the classification of planar integral points sets;
+then, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
+and provide some general constructions of such bounds.
 
@@
 \end{dfn}
-Facher sets are very dominating examples of planar integral pont sets.
-In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
+Facher sets are very dominating examples of planar integral point sets.
+In~\cite{antonov2008maximal}, facher sets with characteristic 1 are called \textit{semi-crabs}.
 For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.
@@
 \begin{dfn}
-	A planar integral point sets situated in two parallel straight lines
+	A planar integral point set situated in two parallel straight lines
 	is called a \textit{rails} set.
@@
 \begin{dfn}
-	A planar integral point sets situated in two perpendicular straight lines
+	A planar integral point set situated in two perpendicular straight lines
 	is called a \textit{cross} set.
@@
 \begin{dfn}
-	A planar integral point sets situated in two straight lines
+	A planar integral point set situated in two straight lines
 	that are not parallel nor perpendicular,
-	is called a \textit{sciccors} set.
+	is called a \textit{scissors} set.
 \end{dfn}
 
-There is an important subclassof scissors sets.
+There is an important subclass of scissors sets.
 
@@
 \begin{dfn}
-	A planar integral point sets that is situated on a circle is called a \textit{circular}
+	A planar integral point set that is situated on a circle is called a \textit{circular}
 	point set.
@@
 \begin{dfn}
-	A planar integral point sets that is situated on the conjunction of a circle with its center,
+	A planar integral point set that is situated on the conjunction of a circle with its center,
 	is called a \textit{centered-circular} point set.
@@
 	Let $\mathcal{P}$ be a planar integral point set consisting of
-	$n - k$ points on line $l_1$ and $k$ points $P_1$, $P_2$, ..., $P_k$ on a
+	$n - k$ points on the line $l_1$ and $k$ points $P_1$, $P_2$, ..., $P_k$ on a
 	parallel line $l_2$ with distance $r$ between $l_1$ and $l_2$. If there
@@
 estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 36$.
-For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
+For the sake of brevity, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
 $\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
 which means that each abscissa is multiplied by $1/q$
-and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
+and each ordinate is multiplied by $\sqrt{p}/q$,  that is
 $$
@@
 
-$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam(\mathcal{P})}
+$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam}(\mathcal{P})
 = 107526848$,
@@
 \begin{remark}
-For all the examples above, maximum in the right part of (\ref{formula1})
-evaluates to the diameter of the system.
-The following example shows that the maximum can be evaluated to $w$:
+In all the examples above, the maximum in the right-hand side of (\ref{formula1})
+is attained on the diameter of the system.
+The following example shows that the maximum can be attained on $w$:
 \end{remark}
@@
 We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
-is worse that known one (\hyperlink{d2}{2})).
+is weaker than the one given in~\eqref{eq:d_m_2m+2}.
 
@@
 
-Guy gives \cite[D 20]{guy2013unsolved} a system of $8$ points located
-on two parallel lines. The points of the system form a rectangle
+Guy gives \cite[D 20]{guy2013unsolved} a PIPS of $8$ points located
+on two parallel lines. The points of the set form a rectangle
 (Figure~\ref{rectangle.pdf}):
@@
 	\hfill
-	\parbox{.5\linewidth}{\caption{Counterexample}
+	\parbox{.5\linewidth}{\caption{Contradiction}
 	\label{right_angle.pdf}}
@@
 \section{Final remarks}
-All the given PIPSs were obtained through a combination of computer search an intuition of the authors;
-so, the further search may lead to better bounds employing the same consructions.
-
-There is still no general construction for a rails or scissors PIPS of given cardinality.
-For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality with 2 points on one line
+All the discussed PIPSs were obtained through a combination of computer search and intuition of the authors;
+so, the further search may lead to better bounds employing the same constructions.
+
+There is still no general construction for a rails or scissors PIPS of arbitrary cardinality.
+For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality, with 2 points on one line
 and all the rest on the other;

### 2021-11-03 AvdMomZvol_rails/Article.tex
[AMZr] Edits from Usachev - p. 3

@@
 \begin{theorem}
+	\label{thm:rails_blowup}
 	Let $\mathcal{P}$ be a planar integral point set consisting of
@@
 
-Below we give the planar integral point sets and the corresponding
+Below we give the examples of planar integral point sets and the corresponding
 estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 36$.
-For the sake of brevity, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
-$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
+For the sake of brevity, we use the notation
+$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$
+from~\cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018},
 which means that each abscissa is multiplied by $1/q$
@@
 $$
-
-\begin{itemize}
-\setlength{\itemsep}{-1mm}
 
@@
 
+Below $f$ and $w$ are those from Theorem~\ref{thm:rails_blowup}.
+
+\begin{itemize}
+\setlength{\itemsep}{-1mm}
+
 \item
@@
 
-Below we give the planar integral point sets and the corresponding
+Below we give the examples of planar integral point sets and the corresponding
 estimates of the function $d(m, n)$ for $n = 3m + 1$ and $n = 4m + 1$.

### 2021-11-04 AvdMomZvol_rails/Article.tex
[AMZr] Edits from Dr. Usachev - page 2

@@
 \begin{dfn}
+	The characteristic of a PIPS $\mathcal{P}$ is a squarefree number $q$
+	such that the area of any triangle $M_1M_2M_3$, $\{M_1,M_2,M_3\}\subset \mathcal{P}$,
+	is commensurable with $\sqrt{q}$.
+\end{dfn}
+
+\begin{dfn}
 	The function $d(m, n)$ is the minimal possible diameter of
@@
 \begin{theorem}
+	\label{thm:Kurz_blowup}
 	Let $\mathcal{P}$ be a planar integral point set consisting of
@@
 
-First, we discuss the classification of planar integral points sets;
-then, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
+In this paper we further study the various features of IPS.
+In Section 2, we discuss the classification of planar integral points sets;
+in Section 3, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
 and provide some general constructions of such bounds.
@@
 \end{dfn}
-Facher sets are very dominating examples of planar integral point sets.
+Most of the examples of planar integral point sets turns to be facher sets.
 In~\cite{antonov2008maximal}, facher sets with characteristic 1 are called \textit{semi-crabs}.
-For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.
-
-For non-facher integral point sets situated in two straight lines,
-we can easily distinct the following three cases:
+For $9 \leq n \leq 122$, the diameter $d(2,n)$ is attained on a facher point set~\cite{kurz2008minimum}.
+
+Non-facher integral point sets situated in two straight lines
+can be further classificated in the following three cases:
 
@@
 
-Among the rails sets, sets with 2 points on one line and all the other on another line dominate.
+Among the rails sets, sets with 2 points on one line and all the other on another line prevail.
 
@@
 Circular sets are very important examples
-of integral point sets~\cite{harborth1993upper,piepmeyer1996maximum,bat2018number}.
+of integral point sets~\cite{harborth1993upper,piepmeyer1996maximum,bat2018number};
+for example, the best known upper bound for $d(2,n)$
+is attained on a circular set.
 
@@
 
-These six classes dominate among all the known planar integral point sets;
-however, some sophisticated constructions are also known.
+Most of the known planar integral point sets fall into these six classes;
+however, there are some sophisticated examples which do not~\cite{kreisel2008heptagon,kurz2013constructing,avdeev2021particular}.
+A lower bound for the diameter of a PIPS with no collinear triples (so-called PIPS \textit{in semi-general position})
+is presented in~\cite{my-semi-general-5-4-bound-2019}.
 
@@
 
-Below $f$ and $w$ are those from Theorem~\ref{thm:rails_blowup}.
+Below $f$ and $w$ are those from Theorem~\ref{thm:Kurz_blowup}.
 

### 2021-11-06 AvdMomZvol_rails/Article.tex
[AMZr] Abstract

@@
 
+\paragraph{Absctract.}
+	An integral point set $\mathcal{P}$ is a set of $n$ points in
+	the $m$-dimensional Euclidean space $\mathbb{R}^{m}$ with pairwise
+	integral distances,
+	such that $\mathcal{P}$ is not contained in a $(m-1)$-dimensional hyperplane.
+	In the present paper we present some classes of planar integral point sets ($m=2$);
+	we mostly focus on the subsets of a union of two straight lines:
+	facher, rails, scissors and pyramid sets.
+	We use rails and pyramid sets to construct lower bounds for a diameter
+	of integral point sets in higher dimensions.
+
 \section{Introduction}

### 2021-11-06 AvdMomZvol_rails/Article.tex
[AMZr] Edits from Dr. Usachev - page 1

@@
 	the $m$-dimensional Euclidean space $\mathbb{R}^{m}$ with pairwise
-	integral distances.
+	integral distances,
+	such that $\mathcal{P}$ is not contained in a $(m-1)$-dimensional hyperplane.
 \end{dfn}
@@
 then, we can find its cardinality, which is always finite~\cite{anning1945integral,erdos1945integral};
-finally, we can naturally % TODO: naturally/essentially? cf. arxiv/1
-define the diameter of a finite point set.
+finally, we can compute the diameter of a finite point set,
+which is naturally defined as follows:
 
@@
 
+The following notion was introduced in~\cite{kurz2008bounds}.
+
 \begin{dfn}
@@
 \end{dfn}
+
+The computation of exact values of $d(m,n)$ is a difficult problem.
+In 2003, Solymosi proved~\cite{solymosi2003note} that $d(2,n) \geq cn$ for some constant $c$;
+now it is known~\cite{my-pps-linear-bound-2019} that $c>5/11$.
+So, even finding the estimates is not easy.
 
@@
 
-For the list of known bounds for $d(m, n)$,
-we refer the reader to~\cite[Theorem 1]{kurz2008bounds} or to~\cite{our-vmmsh-2018}.
-We will discuss the following ones presented at~\cite{kurz2008bounds}:
+A long list of known bounds for $d(m, n)$ and some exact values
+can be found in~\cite[Theorem 1]{kurz2008bounds} or in~\cite{our-vmmsh-2018}.
+Below we will discuss the following ones presented at~\cite{kurz2008bounds}:
 \begin{align}
@@
 	\begin{equation}\label{formula1}
-		d(m, n + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
+		d(m, n + 2(m - 2)) \leq \max(w, \operatorname{diam}(\mathcal{P}))
 	\end{equation}
@@
 In Section 2, we discuss the classification of planar integral points sets;
-in Section 3, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
+in Sections 3 and 4, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
 and provide some general constructions of such bounds.

### 2021-11-06 AvdMomZvol_rails/Article.tex
[AMZr] Acknowledgements
[AMZr] Section refs - just in case

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 In this paper we further study the various features of IPS.
-In Section 2, we discuss the classification of planar integral points sets;
-in Sections 3 and 4, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
+In Section~\ref{sec:classif}, we discuss the classification of planar integral points sets;
+in Sections~\ref{sec:bounds_rails}~and~\ref{sec:bounds_pyramid}, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
 and provide some general constructions of such bounds.
@@
 \section{Classification of planar integral point sets}
+\label{sec:classif}
 
@@
 \section{Bounds based on rails sets}
+\label{sec:bounds_rails}
 
@@
 \section{Bounds based on pyramid sets}
+\label{sec:bounds_pyramid}
 
@@
 
+\section{Acknowledgements}
+The author thanks Dr. A.S. Usachev for his help in polishing the text of the paper.
+
 \printbibliography

### 2021-11-07 AvdMomZvol_rails/Article.tex
[AMZr] Mention higher dimensions
[AMZr] Code cosmetics
[AMZr] A brief remark on the characteristic of a cen-circ PIPS

@@
 now it is known~\cite{my-pps-linear-bound-2019} that $c>5/11$.
+The most recent advances for higher dimensions can be found in~\cite{nozaki2013lower}.
 So, even finding the estimates is not easy.
@@
 \end{dfn}
+
+\begin{remark}
+	Every centered-circular point set has characteristic 1.
+	Indeed, let $\mathcal{P} = \{M_0, M_1,$ $ M_2, M_3, ..., M_k\}$ be a centered-circular point set
+	with $M_0$ being the center.
+	According to the sine formula, in the triangle $M_1 M_2 M_3$
+	\begin{equation}
+		\sin \angle M_1 M_2 M_3 = \frac{|M_1M_3|}{2R}
+		,
+	\end{equation}
+	where $R = |M_0 M_1| $ is the radius of the triangle's circumcircle.
+	Then the area of the triangle $M_1 M_2 M_3$ is
+	\begin{equation}
+		S = \frac12 |M_2 M_1| \cdot |M_2 M_3| \cdot \sin \angle M_1 M_2 M_3 =
+		\frac12 |M_2 M_1| \cdot |M_2 M_3| \cdot \frac{|M_1M_3|}{2|M_0 M_1|} \in\mathbb Q
+		.
+	\end{equation}
+\end{remark}
 
@@
 \begin{equation}\label{formula2}
-d(m, (k - \alpha)m + \alpha) \leq \operatorname{diam(\mathcal{P})},
+d(m, (k - \alpha)m + \alpha) \leq \operatorname{diam}(\mathcal{P}),
 \end{equation}

### 2021-11-07 AvdMomZvol_rails/Article.tex
[AMZr] A ref to existent rails sets

@@
 Among the rails sets, sets with 2 points on one line and all the other on another line prevail.
+In Section~\ref{sec:bounds_rails} below, some examples of such sets are discussed;
+they include an example of rails sets with 2 points on one line and 38 point on the other line
+(Fig.~\ref{picture_40.png})
+and
+an example of rails set with 4 points on one line and 4 points on the other
+(Fig.~\ref{picture.png}).
+We have no examples of rails set with 4 points on one line and 5 points on the other;
+and we have no examples of rails sets with 3 points on one line and 9 points on the other.
 

### 2021-11-07 AvdMomZvol_rails/Article.tex
[AMZr] A picture with cen-circ set

@@
 
+\begin{figure}[h!]
+\center{\includegraphics[width=0.7\linewidth]{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png}}
+\parbox{0.7\linewidth}{\caption{A centered-circular PIPS of cardinality 17 and diameter 2210}
+\label{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png}}
+\end{figure}
+
+On Fig.~\ref{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png} the following centered-circular PIPS is shown:
+\begin{multline*}
+	\sqrt{1}/1105*\{
+	( 0 ; 0);
+	( \pm1221025 ; 0);
+	( 113953 ; 1215696);
+	( -113953 ; -1215696);
+	\\
+	( 341887 ; 1172184);
+	( -341887 ; -1172184);
+	( 680225 ; 1014000);
+	( -680225 ; -1014000);
+	\\
+	( 782977 ; -936936);
+	( -782977 ; 936936);
+	( 859775 ; -867000);
+	( -859775 ; 867000);
+	\\
+	( 1073057 ; 582624);
+	( -1073057 ; -582624);
+	( 1198975 ; 231000);
+	( -1198975 ; -231000);
+\}
+\end{multline*}
+
 Most of the known planar integral point sets fall into these six classes;

### 2021-11-07 AvdMomZvol_rails/Article.tex
[AMZr] Remove a TODO note
[AMZr] Move the notation to where itisused first

@@
 
+For the sake of brevity, we will hereafter use the notation
+$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$
+from~\cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018},
+which means that each abscissa is multiplied by $1/q$
+and each ordinate is multiplied by $\sqrt{p}/q$,  that is
+$$
+	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
+	=
+	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
+$$
+
 \begin{dfn}
@@
 \end{dfn}
-
-%TODO: rephrase?
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png}}
-\parbox{0.7\linewidth}{\caption{A centered-circular PIPS of cardinality 17 and diameter 2210}
+\begin{figure}[htbp]
+	\includegraphics[width=.42\linewidth]{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png}
+	\hfill
+	\includegraphics[width=.545\linewidth]{15_3364_91_30e991f82a54fb9caeca5b6a6776b981_X.png}
+	\\
+	\parbox{.37\linewidth}{\caption{A centered-circular PIPS of cardinality 17 and diameter 2210}
 	\label{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png}}
+	\hfill
+	\parbox{.5\linewidth}{\caption{A scissors PIPS of cardinality 15 and diameter 3364}
+	\label{15_3364_91_30e991f82a54fb9caeca5b6a6776b981_X.png}}
 \end{figure}
@@
 estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 36$.
-For the sake of brevity, we use the notation
-$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$
-from~\cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018},
-which means that each abscissa is multiplied by $1/q$
-and each ordinate is multiplied by $\sqrt{p}/q$,  that is
-$$
-	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
-	=
-	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
-$$
 

### 2021-11-07 AvdMomZvol_rails/Article.tex
[AMZr] Add reference for pyramid sets
[AMZr] Complete the example of a scissors set

@@
 
-%TODO: can a scissors set have another axis of symmetry?
+Examples of pyramid sets are given in Section~\ref{sec:bounds_pyramid} on
+Fig.~\ref{picture_56.png} and Fig.~\ref{picture_1260.png}.
+We don't have any examples of pyramid sets with cardinality more than 9.
+
+It still remain unknown whether a scissors set can have another axis of symmetry;
+on Fig.~\ref{15_3364_91_30e991f82a54fb9caeca5b6a6776b981_X.png} the following scissors set with central symmetry is shown:
+\begin{multline*}
+	\sqrt{91}/1*\{
+	( 0 ; 0);
+	( 792 ; 120);
+	( -792 ; -120);
+\\
+	( \pm256 ; 0);
+	( \pm646 ; 0);
+	( \pm693 ; 0);
+	( \pm1152 ; 0);
+	( \pm1242 ; 0);
+	( \pm1682 ; 0)
+\}
+\end{multline*}
 

### 2021-11-07 AvdMomZvol_rails/Article.tex
[AMZr] An example of 3-8-rails set

@@
 $$
+
+On Fig.~\ref{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png} we show the following rails set with 3 points on one line and 8 points on the other:
+\begin{multline*}
+\mathcal{P}_{3,8}=\sqrt{255255}/2*\{
+	( 1767 ; -3);
+	( 2791 ; -3);
+	( 4071 ; -3);
+\\
+	( -306 ; 0);
+	( 0 ; 0);
+	( 1798 ; 0);
+	( 2304 ; 0);
+	( 2760 ; 0);
+	( 3534 ; 0);
+	( 4040 ; 0);
+	( 4558 ; 0)
+\}
+\end{multline*}
+It's notable that $\operatorname{char} \mathcal{P}_{3,8}=255255 = 3 \cdot 5 \cdot 7 \cdot 11 \cdot 13 \cdot 17$
+is a production of the first 6 odd primes,
+and the distances between the three points on the bottom line are $512=2^{9}$, $640=2^7 \cdot 5$ and
+$512+640 = 1152 = 2^7 \cdot 3^2$.
+
+\begin{figure}[h!]
+	\center{\includegraphics[width=0.9\linewidth]{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}}
+	\parbox{0.9\linewidth}{\caption{PIPS of cardinality 11 and diameter 2423}
+	\label{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}}
+\end{figure}
 

### 2021-11-08 AvdMomZvol_rails/Article.tex
[AMZr] Probably the last example of 4-4-rails
[AMZr] The bunch of final edits

@@
 	such that $\mathcal{P}$ is not contained in a $(m-1)$-dimensional hyperplane.
-	In the present paper we present some classes of planar integral point sets ($m=2$);
+	In the present paper we discuss some classes of planar integral point sets ($m=2$);
 	we mostly focus on the subsets of a union of two straight lines:
@@
 \label{sec:intro}
+
+The study of integral point sets belongs to the classical analysis.
+It is motivated by both pure theoretical aspects~\cite{ascher2020erdos,solymosi2010question}
+and applications~\cite{kurz2007enumeration}.
 
@@
 	The function $d(m, n)$ is the minimal possible diameter of
-	the integral point set $\mathcal{P}$ of $n$ points in
+	an integral point set $\mathcal{P}$ of $n$ points in
 	$m$-dimensional Euclidean space $\mathbb{R}^{m}$.
@@
 
-%For a bit more sophisticated configurations, another functions are presented~\cite{kurz2008minimum,kreisel2008heptagon}.
-%
-%\begin{dfn}
-%	The function $\overline{d}(m, n)$ is the minimum possible diameter of
-%	the integral point set $\mathcal{P}$ of $n$ points in
-%	$m$-dimensional Euclidean space $\mathbb{R}^{m}$
-%	such that no $m+1$ points are located on an $(m-1)$-dimensional hyperplane.
-%\end{dfn}
-
-%\begin{dfn}
-%	The function $\dot{d}(m, n)$ is the minimum possible diameter of
-%	the integral point set $\mathcal{P}$ of $n$ points in
-%	$m$-dimensional Euclidean space $\mathbb{R}^{m}$
-%	such that no three points are located on a straight line
-%	and no four points are located on a circle.
-%\end{dfn}
-
 A long list of known bounds for $d(m, n)$ and some exact values
@@
 \begin{align}
-	d(m, 2m + 1) \leq 8\\
+	d(m, 2m + 1) \leq 8  \label{eq:d_m_2m+1}\\
 	d(m, 2m + 2) \leq 13 \label{eq:d_m_2m+2}\\
-	d(m, 3m) \leq 109
+	d(m, 3m) \leq 109    \label{eq:d_m_3m}
 \end{align}
@@
 In Section~\ref{sec:classif}, we discuss the classification of planar integral points sets;
-in Sections~\ref{sec:bounds_rails}~and~\ref{sec:bounds_pyramid}, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
+in Sections~\ref{sec:bounds_rails}~and~\ref{sec:bounds_pyramid}, we present some upper bounds for $d(m,n)$
+based on planar integral point sets of special types
 and provide some general constructions of such bounds.
+All the bounds, like~\eqref{eq:d_m_2m+1}--\eqref{eq:d_m_3m}, are of the form $d(m, km+p)\leq c_{k,p}$.
 
@@
 
-\subsection{Integral point sets situated in two straight lines}
+\subsection{Integral point sets situated on two straight lines}
 
@@
 
-Non-facher integral point sets situated in two straight lines
+Non-facher integral point sets situated on two straight lines
 can be further classificated in the following three cases:
@@
 \begin{dfn}
-	A planar integral point set situated in two parallel straight lines
+	A planar integral point set situated on two parallel straight lines
 	is called a \textit{rails} set.
@@
 (Fig.~\ref{picture.png}).
-We have no examples of rails set with 4 points on one line and 5 points on the other;
-and we have no examples of rails sets with 3 points on one line and 9 points on the other.
+We still do not know whether there are rails set with 4 points on one line and 5 points on the other;
+and the same for rails sets with 3 points on one line and 9 points on the other.
 
@@
 It's notable that $\operatorname{char} \mathcal{P}_{3,8}=255255 = 3 \cdot 5 \cdot 7 \cdot 11 \cdot 13 \cdot 17$
-is a production of the first 6 odd primes,
-and the distances between the three points on the bottom line are $512=2^{9}$, $640=2^7 \cdot 5$ and
+is a product of the first 6 odd primes,
+and the distances between the three points on the lower line are $512=2^{9}$, $640=2^7 \cdot 5$ and
 $512+640 = 1152 = 2^7 \cdot 3^2$.
@@
 
+Rails sets with 4 and 4 points on the lines are somewhat rather common;
+an example of such a set is shown on Fig.~\ref{8_1111_770_5666c83a03a80e1d3f8f8e8186a145f6__rails_4.png}:
+\begin{equation*}
+	\sqrt{770}/1*\{
+		( 0 ; 0);
+		( 380 ; 0);
+		( 731 ; 0);
+		( 1111 ; 0);
+		( 354 ; -12);
+		( 377 ; -12);
+		( 734 ; -12);
+		( 757 ; -12)%;
+	\}
+\end{equation*}
+
+\begin{figure}[h!]
+	\center{\includegraphics[width=0.9\linewidth]{8_1111_770_5666c83a03a80e1d3f8f8e8186a145f6__rails_4.png}}
+	\parbox{0.9\linewidth}{\caption{PIPS of cardinality 8 and diameter 1111}
+	\label{8_1111_770_5666c83a03a80e1d3f8f8e8186a145f6__rails_4.png}}
+\end{figure}
+
 \begin{dfn}
-	A planar integral point set situated in two perpendicular straight lines
+	A planar integral point set situated on two perpendicular straight lines
 	is called a \textit{cross} set.
@@
 Every cross set has characteristic 1;
-in~\cite{antonov2008maximal}, cross sets with only 2 points out of one of the lines are called \textit{crabs}.
+in~\cite{antonov2008maximal}, cross sets with only 2 points out of one of the lines that have two axes of symmetry are called \textit{crabs}.
 
 \begin{dfn}
-	A planar integral point set situated in two straight lines
+	A planar integral point set situated on two straight lines
 	that are not parallel nor perpendicular,
@@
 
-It still remain unknown whether a scissors set can have another axis of symmetry;
+It still remains unknown whether a scissors set can have another axis of symmetry;
 on Fig.~\ref{15_3364_91_30e991f82a54fb9caeca5b6a6776b981_X.png} the following scissors set with central symmetry is shown:
@@
 and all the rest on the other;
-on the other hand, we still have not found any rails PIPSs with 4 and 5 points on the lines.
+on the other hand, it is still unknown whether there are any rails PIPSs with 4 and 5 points on the lines.
 
@@
 \section{Acknowledgements}
-The author thanks Dr. A.S. Usachev for his help in polishing the text of the paper.
+The authors thank A.S. Usachev for his help in polishing the text of the paper.
 

### 2021-11-09 AvdMomZvol_rails/Article.tex
[AMZr] Rename the paper
[AMZr] Hopefully the last fix

@@
 \title{
-	On particular diameter bounds for integral point sets in higher dimensions
+	On classification of integral point sets and particular upper bounds for their minimal diameter in higher dimensions
 	\footnote{
@@
 All the bounds, like~\eqref{eq:d_m_2m+1}--\eqref{eq:d_m_3m}, are of the form $d(m, km+p)\leq c_{k,p}$.
+
+This paper is an extended version of preprint~\cite{avdeev2019particular}.
 
@@
 	with $M_0$ being the center.
-	According to the sine formula, in the triangle $M_1 M_2 M_3$
+	by the Law of sines, in the triangle $M_1 M_2 M_3$
 	\begin{equation}

### 2021-11-14 2021_Lob/Article.tex
[2021Lob] As is from Zvolinsky

\documentclass[12pt]{article}

\oddsidemargin=-1mm
\evensidemargin=-1mm
\textheight=220mm
\topmargin=-10mm
\textwidth=165mm
\headsep=10mm

\usepackage{amssymb,amsmath,amsfonts}
\usepackage[T2A]{fontenc}
\usepackage[cp1251]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{graphicx}
\usepackage{epstopdf}

\let\chapter\section
\sloppy

\begin{document}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\noindent
	УДК: 514.822

	{\large
		\begin{center}
			\textbf{Примеры множеств точек с целочисленными расстояниями, найденные на ЭВМ}\\
			Р.Е.~Зволинский, Н.Н.~Авдеев, Е.А.~Момот, А.Е.~Зволинский
		\end{center}
	}

	{\footnotesize
		\noindent
		1) roman.zvolinskiy@gmail.com; Воронежский государственный университет\\[1em]
		\noindent
		2) nickkolok@mail.ru; avdeev@vsu.ru; Воронежский государственный университет\\[1em]
		\noindent
		3) y-kate@ukr.net; Воронежский государственный университет\\[1em]
		\noindent
		4) aezv.global@gmail.com; Воронежский государственный университет
	}

	%\hrule

	\begin{abstract}
		Множеством точек на плоскости с целочисленными расстояниями называют такое множество не лежащих на одной прямой точек,
		что расстояние между любыми двумя точками этого множества есть целое число. В настоящей статье представлены новые 
		результаты по поиску подобных множеств.
	\end{abstract}

	\noindent
	\textit{Ключевые слова:} множество точек на плоскости с целочисленными расстояниями, диаметр множества, 
	мощность множества, характеристика множества, отыскание множеств точек с целочисленными расстояниями.

	\vspace{1em}
	\hrule
	\vspace{2em}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\medskip

	\textbf{Определение 1.} {\it Множеством $\mathcal P$ точек на плоскости с целочисленными расстояниями (сокращённо МЦР) называют
	такое множество не лежащих на одной прямой прямой точек, что расстояние между любыми двумя точками
	из $\mathcal P$ есть целое число.}

	\medskip

	П. Эрдёш и Н. Эннинг [1, 2] доказали, что подобные множества состоят из конечного числа точек, т.е. 
	имеют конечную мощность (cardinality). Кроме того, данные множества обладают и конечным диаметром
	(diameter), определяемым как
	$$
	    \operatorname{diam} (\mathcal P) = \max_{P_1, P_2 \in \mathcal P}
	    |P_1 P_2|,
	$$
	где $|P_1 P_2|$ есть обычное расстояние между точками. Позже А. Кемнитц
	[3] обнаружил, что у каждого такого множества есть ещё и характеристика.

	Множества точек на плоскости с целочисленными расстояниями весьма разнообразны. Взять к примеру веерные множества (facher sets), которые в
	свою очередь также бывают очень разные [4, 5].

	\medskip

	\textbf{Определение 2.} {\it Множества точек на плоскости с целочисленными расстояниями мощности $n$, для которых можно найти
		такую прямую $\ell_1$, что $(n - 1)$ точек лежит на $\ell_1$, будем называть веерными (facher sets).}

	\medskip

	Существуют и более интересные множества: крестовые (cross sets), пирамидальные (pyramid set), ножничные (scissors sets). В статье [6] введена классификация подобных множеств, 
	а в [7] даются примеры, в классификацию не ложащиеся.

	Для построения каких-либо общих рассуждений, касающихся данных множеств, необходимы наглядные примеры. Поиск таких 
	примеров осуществляется с помощью аппаратных средств ЭВМ.

	\begin{center}
		{\bf Ножничные множества.}
	\end{center}

    \medskip

    \textbf{Определение 3.} {\it Ножничными (scissors sets) будем называть такие множества точек на плоскости с целочисленными 
    расстояниями мощности $n$, для которых можно единственным образом найти такие прямые $\ell_1$, $\ell_2$, 
    что $k$ точек лежат на прямой $\ell_1$, а $(n - k)$~--- на прямой $\ell_2$, причём $k \geqslant 2$ и прямые 
    $\ell_1$ и $\ell_2$ не являются ни параллельными, ни перпендикулярными.}

    \medskip

    Для облегчения восприятия применяется используемая в [5, 8, 9] запись вида:
    $\sqrt p/q * \{(x_1, y_1), \ldots, (x_n, y_n)\}$ ($p$~--- это характеристика), означающая, 
    что каждую абсциссу нужно умножить на $1/q$, а каждую ординату~--- на $\sqrt p/q$, т. е.
    $$
        \sqrt p/q * \{(x_1, y_1), \ldots, (x_n, y_n)\} = 
        \left\{\left(\frac{x_1}q, \frac{y_1\sqrt p}q\right),
        \ldots, \left(\frac{x_n}q, \frac{y_n\sqrt p}q\right)\right\}.
    $$

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=1\linewidth]{25_1544076_91.pdf}
    	\caption{МЦР мощности 25, диаметр 1544076}
    \end{figure}

    \begin{align*}
        &\sqrt{91}/1 * \{(\pm 772038, 0),
        (\pm 470016, 0),
        (\pm 193732, 0),
        (\pm 171564, 0),
        (\pm 126684, 0),
        (\pm 117504, 0), \\
        &(80784, 12240),
        (-80784, -12240),
        (\pm 72960, 0),
        (\pm 70686, 0),
        (\pm 65892, 0),
        (\pm 42891, 0), \\
        &(\pm 26112, 0),
        (0, 0)\} \text{ (Рис. 1)}. 
    \end{align*}

    \begin{center}
        {\bf Рельсовые множества.}	
    \end{center}

    \medskip

    \textbf{Определение 4.} {\it Рельсовыми (rails sets) будем называть такие множества точек на плоскости с целочисленными расстояниями мощности $n$, 
    для которых можно единственным образом найти такие прямые $\ell_1$, $\ell_2$, что $k$ точек лежат на прямой $\ell_1$,
    а $(n - k)$~--- на прямой $\ell_2$, причём $k \geqslant 2$.}

	\medskip

	\begin{figure}[h!]
		\centering
		\includegraphics[width=1\linewidth]{28_158064_154.pdf}
		\caption{МЦР мощности 28, диаметр 158064}
	\end{figure}

    \begin{align*}
    	&\sqrt{154}/1 * \{(\pm 79032, 0),
    	(\pm 49184, 0),
    	(\pm 44352, 0),
    	(\pm 34316, 0),
    	(\pm 33024, 0),
    	(\pm 24864, 0), \\
    	&(\pm 21576, 0),
    	(\pm 20976, -1440),
    	(\pm 20376, 0),
    	(\pm 14712, 0),
    	(\pm 13824, 0),
    	(\pm 8805, 0), \\
    	&(\pm 7232, 0),
    	(\pm 2400, 0)
    	\} \text{ (Рис. 2)}.
    \end{align*}

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=1\linewidth]{32_1598904_154.pdf}
    	\caption{МЦР мощности 32, диаметр 1598904}
    \end{figure}

    \begin{align*}
    	&\sqrt{154}/1 * \{(\pm 799452, 0),
    	(\pm 553224, 0),
    	(\pm 464232, 0),
    	(\pm 344288, 0),
    	(\pm 310464, 0), \\
    	&(\pm 240212, 0),
    	(\pm 231168, 0),
    	(\pm 174048, 0),
    	(\pm 151032, 0),
    	(\pm 146832, -10080), \\
    	&(\pm 142632, 0),
    	(\pm 102984, 0),
    	(\pm 96768, 0),
    	(\pm 61635, 0),
    	(\pm 50624, 0),
    	(\pm 16800, 0),
    	\} \text{ (Рис. 3)}.
    \end{align*}

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=1\linewidth]{44_456989104_154.pdf}
    	\caption{МЦР мощности 44, диаметр 456989104}
    \end{figure}

    \begin{align*}
    	&\sqrt{154}/1 * \{(\pm 228494552, 0),
    	(\pm 220848615, 0),
    	(\pm 194867400, 0),
    	(\pm 152828130, 0), \\
    	&(\pm 128244090, 0),
    	(\pm 95109560, 0),
    	(\pm 85765680, 0),
    	(\pm 66358565, 0), 
    	(\pm 63860160, 0), \\
    	&(\pm 62812365, 0),
    	(\pm 48080760, 0),
    	(\pm 41722590, 0),
    	(\pm 40562340, -2784600), 
    	(\pm 39402090, 0), \\
    	&(\pm 28449330, 0),
    	(\pm 26732160, 0),
    	(\pm 16126320, 0),
    	(\pm 13984880, 0),
    	(\pm 8239560, 0), \\
    	&(\pm 4641000, 0),
    	(\pm 3136770, 0),
    	(\pm 2480504, 0)
    	\} \text{ (Рис. 4)}.
    \end{align*}

    \begin{center}
    	{\bf Множества точек на плоскости с целочисленными расстояниями, не входящие в общую классификацию.}
    \end{center}

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=1\linewidth]{23_243663771_1.pdf}
    	\caption{МЦР мощности 23, диаметр 243663771}
    \end{figure}

    \begin{align*}
    	&\sqrt 1/1 * \{(125570211, 0),
    	(56036760, 0),
    	(44950698, 0),
    	(29275190, 0),
    	(21302190, 0), \\
    	&(\pm 16820310, 0), 
    	(12380862, 0),
    	(10722410, 0),
    	(9463662, 0),
    	(8002410, \pm 21162960), \\
    	&(8002410, 0), 
    	(4008940, 0), 
    	(2627690, 0),
    	(1748940, 0),
    	(-118093560, 0),
    	(-48727185, 0), \\
    	&(-12152790, 0), 
    	(-8562390, 0), 
    	(-2540310, 0),
    	(-483990, \pm 12252240)
    	\} \text{ (Рис. 5)}.
    \end{align*}

	{\footnotesize Работа выполнена в Воронежском университете при поддержке РНФ, грант 19-11-00197.}

	\begin{thebibliography}{99}

		\bibitem{bib:01}
		Anning\,N.\,H., Erd{\"o}s\,P. {\it Integral distances} //~Bulletin of the American Mathematical Society.~-- 
		1945.~-- Vol.~51.~-- no.~8.~-- Pp.~598--600.~-- DOI: 10.1090/S0002-9904-1945-08407-9.

		\bibitem{bib:02}
		Erd{\"o}s\,P. {\it Integral distances} //~Bulletin of the American Mathematical Society.~-- 
		1945.~-- Vol.~51.~-- no.~12.~-- P. 996.~-- DOI: 10.1090/S0002-9904-1945-08490-0.

		\bibitem{bib:03}
		Kemnitz\,A. {\it Punktmengen mit ganzzahligen Abst{\"a}nden.}~-- 1988.

		\bibitem{bib:04}
		Zvolinsky\,R.\,E. {\it Facher integral point sets with particular distances of arbitrary cardinality} //~Актуальные 
		проблемы прикладной математики, информатики и механики~-- сборник трудов Международной научной конференции.~--
		Научно-исследовательские публикации, 2020.~ Pp.~668--674.

		\bibitem{bib:05}
		Авдеев\,Н.\,Н., Семёнов\,Е.\,М. {\it Множества точек с целочисленными расстояниями на плоскости и в евклидовом пространстве} //~Математический форум (Итоги науки. Юг России).~-- 2018.~-- Pp.~217--236.

		\bibitem{bib:06}
		Avdeev\,N.\,N., Zvolinsky\,R.\,E., Momot\,E.\,A. {\it On particular diameter bounds for integral point sets in higher dimensions.}~-- 
		2019.~-- arXiv: 1909.10386.~-- URL: https://ui.adsabs.harvard.edu/abs/2019arXiv190910386A/abstract.

		\bibitem{bib:07}
		Avdeev\,N.\,N., Momot\,E.\,A., Zvolinsky\,A.\,E. {\it On particular examples of planar integral point sets and their classification.}~--
		2021.~-- arXiv: 2102.12462[math.CO].

		\bibitem{bib:08}
		Авдеев\,Н.\,Н. {\it On integral point sets in special position} //~Некоторые вопросы анализа, алгебры, геометрии и 
		математического образования: материалы международной молодежной научной школы <<Актуальные направления математического 
		анализа и смежные вопросы>>.~-- 2018.~-- Vol.~8.~-- Pp.~5--6.

		\bibitem{bib:09}
		Авдеев\,Н.\,Н. {\it Об отыскании целоудалённых множеств специального вида}
		//~Актуальные проблемы прикладной математики, информатики и механики~-- сборник трудов 
		Международной научной конференции.~-- Научно-исследовательские публикации, 2018.~--
		Pp.~492--498.

	\end{thebibliography}

    \begin{otherlanguage}{english}

    	{\large
    		\begin{center}
    			\textbf{Examples of integral point sets found by computer search}\\
    			R.E. Zvolinsky, N.N. Avdeev, E.A. Momot, A.E. Zvolinsky
    		\end{center}
    	}

    	\begin{abstract}
    		An integral point set is a set of points on the plane such that any distance between the points of the set is integral, and there is no straight line which contains the whole set. We present new examples of integral point sets found.
    	\end{abstract}

    	\textit{Keywords:} planar integral point sets, diameter of an integral point set, cardinality of an integral point set, characteristic of an integral point set, search of integral point sets.

    \end{otherlanguage}

\end{document}

### 2021-11-22 AvdMomZvol_rails/Article.tex
[AMZr] Add note that our bounds are weaker
Revert "[AMZr] Rename the paper"
We are not going to rename it any more

This reverts commit d2ded6ad373e2aef111b7c4c63a5abc3052b314d.

--- 2021-11-08 AvdMomZvol_rails/Article.tex
@@
 All the bounds, like~\eqref{eq:d_m_2m+1}--\eqref{eq:d_m_3m}, are of the form $d(m, km+p)\leq c_{k,p}$.
+%In particular, we improve the bound~\eqref{eq:d_m_3m} to
+%\begin{equation}
+%	d(m, 3m) \leq d(m, 3m + 1) \leq 56
+%\end{equation}
+%and present a new bound
+%\begin{equation}
+%	d(m, 4m + 1) \leq 1260
+%	.
+%\end{equation}
 
@@
 
+However, we have to admit that both bounds which employ pyramid sets
+are indeed weaker than the one
+\begin{equation}
+	d(m, m^2+m)\leq 17
+\end{equation}
+proved in~\cite{kurz2008bounds},
+so the constructions of bounds based on pyramid sets tend to be only for the conceptual purpose.
+
 \end{itemize}

### 2021-11-22 AvdMomZvol_rails/Article.tex
Inject the bibliography into TeX code

@@
 
-\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
-\addbibresource{../common/notmy.bib}
-\addbibresource{../common/my.bib}
+%\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+%\addbibresource{../common/notmy.bib}
+%\addbibresource{../common/my.bib}
+
+%\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
+
+\usepackage{cite}
 
@@
 
-\printbibliography
+%\printbibliography
+%\printbibitembibliography
 %\bibliography{literature}
 
+\begin{thebibliography}{99}
+{}
+\bibitem{ascher2020erdos}
+\emph{Ascher} \emph{K.}, \emph{Braune} \emph{L.}, \emph{Turchet} \emph{A.}
+The Erdős–Ulam problem, Lang’s conjecture and uniformity // Bulletin of the London
+Mathematical Society. — 2020. — Vol. 52, no. 6. — Pp. 1053–1063. — arXiv: \href
+{http://arxiv.org/abs/1809.02037} {\nolinkurl {1809.02037}}.
+{}\bibitem{solymosi2010question}
+\emph{Solymosi} \emph{J.}, \emph{De Zeeuw} \emph{F.} On a question of Erdős
+and Ulam // Discrete \& Computational Geometry. — 2010. — Vol. 43, no. 2. —
+Pp. 393–401. — arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl
+{0806.3095}}.
+{}
+\bibitem{kurz2007enumeration}
+\emph{Kurz} \emph{S.} Enumeration of Integral Tetrahedra // Journal of Integer
+Sequences. — 2007. — Vol. 10, no. 2. — P. 3.
+{}
+\bibitem{anning1945integral}
+\emph{Anning} \emph{N. H.}, \emph{Erdős} \emph{P.} Integral distances //
+Bulletin of the American Mathematical Society. — 1945. — Vol. 51, no. 8. —
+Pp. 598–600. — DOI: \href {https://doi.org/10.1090/S0002-9904-1945-08407-9}
+{\nolinkurl {10.1090/S0002-9904-1945-08407-9}}.
+{}
+\bibitem{erdos1945integral}
+\emph{Erdős} \emph{P.} Integral distances // Bulletin of the American Mathematical
+Society. — 1945. — Vol. 51, no. 12. — P. 996. — DOI: \href
+{https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl
+{10.1090/S0002-9904-1945-08490-0}}.
+{}
+\bibitem{kemnitz1988punktmengen}
+\emph{Kemnitz} \emph{A.} Punktmengen mit ganzzahligen Abständen. — 1988.
+{}
+\bibitem{kurz2005characteristic}
+\emph{Kurz} \emph{S.} On the characteristic of integral point sets in {$\mathbb
+{E}^m$} // Australasian Journal of Combinatorics. — 2006. — Vol. 36. —
+Pp. 241–248. — arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl
+{math/0511704}}.
+{}
+\bibitem{kurz2008bounds}
+\emph{Kurz} \emph{S.}, \emph{Laue} \emph{R.} Bounds for the minimum diameter
+of integral point sets // Australasian Journal of Combinatorics. — 2007. — Vol. 39. —
+Pp. 233–240. — arXiv: \href {http://arxiv.org/abs/0804.1296} {\nolinkurl
+{0804.1296}}.
+{}
+\bibitem{solymosi2003note}
+\emph{Solymosi} \emph{J.} Note on integral distances // Discrete \& Computational
+Geometry. — 2003. — Vol. 30, no. 2. — Pp. 337–342. — DOI: \href
+{https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.{}
+\bibitem{my-pps-linear-bound-2019}
+\emph{Avdeev} \emph{N.} On existence of integral point sets and their diameter
+bounds // Australasian Journal of Combinatorics. — 2020. — Vol. 77, no. 1. —
+Pp. 100–116. — arXiv: \href {http://arxiv.org/abs/1906.11926} {\nolinkurl
+{1906.11926}}. — URL: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190611926A/abstract}.
+{}
+\bibitem{nozaki2013lower}
+\emph{Nozaki} \emph{H.} Lower bounds for the minimum diameter of integral point
+sets // Australasian Journal of Combinatorics. — 2013. — Vol. 56. — Pp. 139–143.
+{}
+\bibitem{our-vmmsh-2018}
+\emph{Авдеев} \emph{Н. Н.}, \emph{Семёнов} \emph{Е. М.} Множества точек с
+целочисленными расстояниями на плоскости и в евклидовом пространстве //
+Математический форум (Итоги науки. Юг России). — 2018. — Pp. 217–236.
+{}
+\bibitem{antonov2008maximal}
+\emph{Antonov} \emph{A. R.}, \emph{Kurz} \emph{S.} Maximal integral point sets
+over {$\mathbb {Z}^2$} // International Journal of Computer Mathematics. —
+2008. — Vol. 87, no. 12. — Pp. 2653–2676. — DOI: \href
+{https://doi.org/10.1080/00207160902993636} {\nolinkurl
+{10.1080/00207160902993636}}. — arXiv: \href {http://arxiv.org/abs/0804.1280}
+{\nolinkurl {0804.1280}}.
+{}
+\bibitem{kurz2008minimum}
+\emph{Kurz} \emph{S.}, \emph{Wassermann} \emph{A.} On the minimum diameter
+of plane integral point sets // Ars Combinatoria. — 2011. — Vol. 101. — Pp. 265–287. —
+arXiv: \href {http://arxiv.org/abs/0804.1307} {\nolinkurl {0804.1307}}.
+{}
+\bibitem{our-ped-2018}
+\emph{Авдеев} \emph{Н. Н.} On integral point sets in special position // Некоторые
+вопросы анализа, алгебры, геометрии и математического образования: материалы
+международной молодежной научной школы «Актуальные направления
+математического анализа и смежные вопросы». — 2018. — Vol. 8. — Pp. 5–6.
+{}
+\bibitem{our-pmm-2018}
+\emph{Авдеев} \emph{Н. Н.} Об отыскании целоудалённых множеств
+специального вида // Актуальные проблемы прикладной математики,
+информатики и механики - сборник трудов Международной научной
+конференции. — Научно-исследовательские публикации, 2018. — Pp. 492–498.
+{}\bibitem{harborth1993upper}
+\emph{Harborth} \emph{H.}, \emph{Kemnitz} \emph{A.},
+\emph{Möller} \emph{M.} An upper bound for the minimum diameter of integral
+point sets // Discrete \& Computational Geometry. — 1993. — Vol. 9, no. 4. —
+Pp. 427–432. — DOI: \href {https://doi.org/10.1007/bf02189331} {\nolinkurl
+{10.1007/bf02189331}}.
+{}
+\bibitem{piepmeyer1996maximum}
+\emph{Piepmeyer} \emph{L.} The maximum number of odd integral distances
+between points in the plane // Discrete \& Computational Geometry. — 1996. —
+Vol. 16, no. 1. — Pp. 113–115. — DOI: \href {https://doi.org/10.1007/bf02711135}
+{\nolinkurl {10.1007/bf02711135}}.
+{}
+\bibitem{bat2018number}
+\emph{Bat-Ochir} \emph{G.} On the number of points with pairwise integral
+distances on a circle // Discrete Applied Mathematics. — 2018. — Vol. 254. —
+Pp. 17–32. — DOI: \href {https://doi.org/10.1016/j.dam.2018.07.004} {\nolinkurl
+{10.1016/j.dam.2018.07.004}}.
+{}
+\bibitem{kreisel2008heptagon}
+\emph{Kreisel} \emph{T.}, \emph{Kurz} \emph{S.} There are integral heptagons, no
+three points on a line, no four on a circle // Discrete \& Computational Geometry. —
+2008. — Vol. 39, no. 4. — Pp. 786–790. — DOI: \href
+{https://doi.org/10.1007/s00454-007-9038-6} {\nolinkurl {10.1007/s00454-007-9038-6}}.
+{}
+\bibitem{kurz2013constructing}
+Constructing $7$-clusters / S. Kurz [et al.] // Serdica Journal of Computing. — 2014. —
+Vol. 8, no. 1. — Pp. 47–70. — arXiv: \href {http://arxiv.org/abs/1312.2318} {\nolinkurl
+{1312.2318}}.
+{}
+\bibitem{avdeev2021particular}
+\emph{Avdeev} \emph{N.}, \emph{Momot} \emph{E.},
+\emph{Zvolinsky} \emph{A.} On particular examples of planar integral point sets and
+their classification. —. — arXiv: \href {http://arxiv.org/abs/2102.12462 [math.CO]}
+{\nolinkurl {2102.12462 [math.CO]}}.
+{}
+\bibitem{my-semi-general-5-4-bound-2019}
+\emph{Avdeev} \emph{N.} On diameter bounds for planar integral point sets in
+semi-general position. — 2019. — arXiv: \href {http://arxiv.org/abs/1907.09331}
+{\nolinkurl {1907.09331}}. — URL: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190709331A/abstract}.
+{}\bibitem{guy2013unsolved}
+\emph{Guy} \emph{R.} Unsolved problems in number theory. Vol. 1. — Springer
+Science \& Business Media, 2013. — DOI: \href
+{https://doi.org/10.1007/978-1-4757-1738-9} {\nolinkurl {10.1007/978-1-4757-1738-9}}.
+\end{thebibliography}
+
 \end{document}

### 2021-12-03 2021_Lob/report_L.tex
[lob2021] The report

\documentclass[10pt,pdf,hyperref={unicode},aspectratio=169]{beamer}

\usepackage{amssymb,amsmath,amsfonts}
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{graphicx}
\usepackage{epstopdf}

\let\chapter\section
\sloppy

\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{llemma}{Лемма}
\newtheorem{ttheorem}[llemma]{Теорема}
\newtheorem{eexample}[llemma]{Пример}
\newtheorem{property}[llemma]{Свойство}
\newtheorem{remark}[llemma]{Замечание}
\newtheorem{ddefinition}[llemma]{Определение}
\newtheorem{ccorollary}{Следствие}[llemma]

\begin{document}

\title{
	Примеры множеств точек\\с целочисленными расстояниями,\\найденные на ЭВМ
}
\author{Р.Е.~Зволинский, Н.Н.~Авдеев, Е.А.~Момот, А.Е.~Зволинский}
\institute{ Воронежский Государственный Университет \\
		{Работа выполнена  при поддержке РНФ, грант 19-11-00197.
	}}
\date{Воронеж 2021}

\maketitle

\setbeamertemplate{footline}[frame number]
\setbeamertemplate{navigation symbols}{\large{}}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

	\begin{ddefinition}
		Множеством $\mathcal P$ точек на плоскости с целочисленными расстояниями (сокращённо МЦР) называют
		такое множество не лежащих на одной прямой точек, что расстояние между любыми двумя точками
		из $\mathcal P$ есть целое число.
	\end{ddefinition}

	\vfill

	Всякое МЦР конечно~\cite{bib:01, bib:02} и, следовательно, имеет конечный диаметр
	\begin{equation*}
	    \operatorname{diam} (\mathcal P) = \max_{P_1, P_2 \in \mathcal P}
	    |P_1 P_2|,
	\end{equation*}
	где $|P_1 P_2|$ есть обычное расстояние между точками.

	\vfill

	$\operatorname{char} M = p \quad \Leftrightarrow \quad S_{\triangle ABC} \in \mathbb{Q}\sqrt{p}, ~~ \{A,B,C\}\subset M$

	\vfill
	Пример: характеристика треугольника со сторонами $(1,1,1)$ равна $3$.

\end{frame}

\begin{frame}
	\begin{ddefinition}
		\emph{Веерными (facher)}%
		\footnote{
			По типично европейской традиции, заложенной С. Курцем, классам МЦР даются оригинальные названия;
			например, в~\cite{crabs} изучаются МЦР классов {\it crab} и даже {\it semi-crab}
		}
		называют МЦР мощности $n$, для которых можно найти
		такую прямую $\ell_1$, что $(n - 1)$ точек лежит на $\ell_1$.
	\end{ddefinition}

	\vfill

	Веерные множества встречаются чаще остальных,
	и вопрос о существовании веерного множества с заданными параметрами хорошо изучен в~\cite{bib:04} и \cite[Теорема 5]{bib:05}.

	\vfill

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=0.7\linewidth]{facher-5.png}
    \end{figure}

\end{frame}

\begin{frame}
	Частичная классификация МЦР введена в~\cite{bib:06}; примеры множеств, в неё не входящих, даны в~\cite{bib:07}.

	\vfill

	Для построения каких-либо общих рассуждений, касающихся существования МЦР некоторого специального вида,
	необходимы наглядные примеры.
	Поиск таких примеров ведётся на ЭВМ и может занимать продолжительное время.
	Авторами разработано ПО на языке NodeJS и опубликовано по адресу gitlab.com/nickkolok/ips-algo
	под свободной лицензией GNU GPLv3.

	\vfill

	Цель данной работы~--- привести примеры МЦР специального вида достаточно большой мощности,
	найденные на ЭВМ.

	\vfill

    \begin{equation*}
        \sqrt p/q * \{(x_1, y_1), \ldots, (x_n, y_n)\} =
        \left\{\left(\frac{x_1}q, \frac{y_1\sqrt p}q\right),
        \ldots, \left(\frac{x_n}q, \frac{y_n\sqrt p}q\right)\right\}.
    \end{equation*}

\end{frame}

\begin{frame}

	\begin{ddefinition}
		\emph{ Ножничными (scissors)} назовём МЦР мощности $n$,
		для которых можно единственным образом найти такие прямые $\ell_1$, $\ell_2$,
		что $k$ точек лежат на $\ell_1$, а $(n - k)$~--- на $\ell_2$, причём $k \geqslant 2$ и
		$\ell_1$ и $\ell_2$ не параллельны и не перпендикулярны.
	\end{ddefinition}

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=1\linewidth]{25_1544076_91.pdf}
    	%\caption
    	\\~\\{Ножничная МЦР мощности 25, диаметр 1 544 076}
    \end{figure}

    \begin{align*}\footnotesize
        \sqrt{91}/1 * \{ &
		\pm(80784, 12240),
        (0, 0),
        (\pm 772038, 0),
        (\pm 470016, 0),
        \\&
        (\pm 193732, 0),
        (\pm 171564, 0),
        (\pm 126684, 0),
        (\pm 117504, 0),
        \\&
        (\pm 72960, 0),
        (\pm 70686, 0),
        (\pm 65892, 0),
        (\pm 42891, 0),
        (\pm 26112, 0)
        \}
    \end{align*}
\end{frame}

\begin{frame}

    \begin{ddefinition}
		\emph{ Рельсовыми (rails)} называют МЦР
		мощности $n$, для которых можно единственным образом найти такие прямые $\ell_1 \| \ell_2$,
		что $k$ точек лежат на $\ell_1$,
		а $(n - k)$~--- на  $\ell_2$, причём $k \geqslant 2$.
	\end{ddefinition}

	\begin{figure}
		\centering
		\includegraphics[width=1\linewidth]{28_158064_154.pdf}
		\\~\\{Рельсовая МЦР мощности 28, диаметр 158 064}
	\end{figure}

    \begin{align*}
    	\sqrt{154}/1 * \{
		&
    	(\pm 79032, 0),
    	(\pm 49184, 0),
    	(\pm 44352, 0),
    	(\pm 34316, 0),
    	(\pm 33024, 0),
    	(\pm 24864, 0), \\
    	&(\pm 21576, 0),
    	(\pm 20976, -1440),
    	(\pm 20376, 0),
    	(\pm 14712, 0),
    	(\pm 13824, 0),
    	(\pm 8805, 0), \\
    	&(\pm 7232, 0),
    	(\pm 2400, 0)
    	\}
    \end{align*}

\end{frame}

\begin{frame}

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=1\linewidth]{32_1598904_154.pdf}
    	\\~\\{Рельсовая МЦР мощности 32, диаметр 1 598 904}
    \end{figure}

    \begin{align*}
    	\sqrt{154}/1 * \{
		&
    	(\pm 799452, 0),
    	(\pm 553224, 0),
    	(\pm 464232, 0),
    	(\pm 344288, 0),
    	(\pm 310464, 0), \\
    	&(\pm 240212, 0),
    	(\pm 231168, 0),
    	(\pm 174048, 0),
    	(\pm 151032, 0),
    	(\pm 146832, -10080), \\
    	&(\pm 142632, 0),
    	(\pm 102984, 0),
    	(\pm 96768, 0),
    	(\pm 61635, 0),
    	(\pm 50624, 0),
    	(\pm 16800, 0)
    	\}
    \end{align*}

\end{frame}

\begin{frame}

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=1\linewidth]{44_456989104_154.pdf}
    	\\~\\{Рельсовая МЦР мощности 44, диаметр 456 989 104}
    \end{figure}

    \begin{align*}
    	\sqrt{154}/1 * \{
		&
    	(\pm 228494552, 0),
    	(\pm 220848615, 0),
    	(\pm 194867400, 0),
    	(\pm 152828130, 0), \\
    	&(\pm 128244090, 0),
    	(\pm 95109560, 0),
    	(\pm 85765680, 0),
    	(\pm 66358565, 0),
    	(\pm 63860160, 0), \\
    	&(\pm 62812365, 0),
    	(\pm 48080760, 0),
    	(\pm 41722590, 0),
    	(\pm 40562340, -2784600),
		\\&
    	(\pm 39402090, 0),
    	(\pm 28449330, 0),
    	(\pm 26732160, 0),
    	(\pm 16126320, 0),
    	(\pm 13984880, 0),
		\\&
    	(\pm 8239560, 0),
    	(\pm 4641000, 0),
    	(\pm 3136770, 0),
    	(\pm 2480504, 0)
    	\}
    \end{align*}

\end{frame}

\begin{frame}
    \frametitle{Пример  МЦР, не входящего в существующую классификацию.}

    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=1\linewidth]{23_243663771_1.pdf}
    	\caption{МЦР мощности 23, диаметр 243663771, не входящее в классификацию}
    \end{figure}

    \begin{align*}
    	&\sqrt 1/1 * \{(125570211, 0),
    	(56036760, 0),
    	(44950698, 0),
    	(29275190, 0),
    	(21302190, 0), \\
    	&(\pm 16820310, 0),
    	(12380862, 0),
    	(10722410, 0),
    	(9463662, 0),
    	(8002410, \pm 21162960), \\
    	&(8002410, 0),
    	(4008940, 0),
    	(2627690, 0),
    	(1748940, 0),
    	(-118093560, 0),
    	(-48727185, 0), \\
    	&(-12152790, 0),
    	(-8562390, 0),
    	(-2540310, 0),
    	(-483990, \pm 12252240)
    	\}
    \end{align*}

\end{frame}

\begin{frame}
	Все вычисления "--- с целыми числами \texttt{BigInt}

	\vfill

	Основные проблемы целочисленных вычислений:

	\begin{itemize}
		\item
			Проверка числа на полный квадрат
		\item
			Факторизация
		\item
			Освобождение от квадратов \textit(для вычисления характеристики)
		\item
			Перечисление делителей \textit(для вставки точек)
		\item
			Хранение списков простых чисел
		\item
			Сериализация в JSON
	\end{itemize}

	Опубликованы библиотеки на NPM:

	\begin{itemize}
		\item
			\texttt{bigint-isqrt}
		\item
			\texttt{bigint-is-perfect-square}
		\item
			\texttt{bigint-json-native} \textit(форк)
	\end{itemize}

\end{frame}

\setbeamertemplate{bibliography item}{\insertbiblabel\footnotesize}

\begin{frame}
	\begin{thebibliography}{99}
		\setlength\itemsep{0pt}\scriptsize

		\bibitem{bib:01}
		Anning\,N.\,H., Erd{\"o}s\,P. {\it Integral distances} //~Bulletin of the American Mathematical Society.~--
		1945.~-- Vol.~51.~-- no.~8.~-- С.~598--600.~-- DOI: 10.1090/S0002-9904-1945-08407-9.

		\bibitem{bib:02}
		Erd{\"o}s\,P. {\it Integral distances} //~Bulletin of the American Mathematical Society.~--
		1945.~-- Vol.~51.~-- no.~12.~-- С. 996.~-- DOI: 10.1090/S0002-9904-1945-08490-0.

		\bibitem{crabs}
		Antonov\,A.\,R., Kurz\,S. {\it Maximal integral point sets over $\mathbb{Z}^2$}
		//International Journal of Computer Mathematics.~-- 2010.~-- Т. 87.~-- №. 12.~-- С. 2653-2676.~-- arXiv: 0804.1280[math.CO].

		\bibitem{bib:04}
		Zvolinsky\,R.\,E. {\it Facher integral point sets with particular distances of arbitrary cardinality} //~Актуальные
		проблемы прикладной математики, информатики и механики~-- сб. тр. Междунар. науч. конф.~-- 2020.~-- С.~668--674.

		\bibitem{bib:05}
		Авдеев\,Н.\,Н., Семёнов\,Е.\,М. {\it Множества точек с целочисленными расстояниями на плоскости и в евклидовом пространстве}
		//~Математический форум (Итоги науки. Юг России).~-- 2018.~-- С.~217--236.

		\bibitem{bib:06}
		Avdeev\,N.\,N., Zvolinsky\,R.\,E., Momot\,E.\,A. {\it On particular diameter bounds for integral point sets in higher dimensions.}~--
		2019.~-- arXiv: 1909.10386.

		\bibitem{bib:07}
		Avdeev\,N.\,N., Momot\,E.\,A., Zvolinsky\,A.\,E. {\it On particular examples of planar integral point sets and their classification.}~--
		2021.~-- arXiv: 2102.12462[math.CO].

		\bibitem{bib:08}
		Авдеев\,Н.\,Н. {\it On integral point sets in special position} //~Некоторые вопросы анализа, алгебры, геометрии и
		математического образования: мат-лы междунар. молодежной научной школы <<Актуальные направления математического
		анализа и смежные вопросы>>.~-- 2018.~-- Vol.~8.~-- Pp.~5--6.

		\bibitem{bib:09}
		Авдеев\,Н.\,Н. {\it Об отыскании целоудалённых множеств специального вида}
		//~Актуальные проблемы прикладной математики, информатики и механики~-- сб. тр.
		Международной науч. конф.~-- 2018.~-- С.~492--498.

		\bibitem{bib:03}
		Kemnitz\,A. {\it Punktmengen mit ganzzahligen Abst{\"a}nden.}~-- 1988.
	\end{thebibliography}

\end{frame}

\setbeamertemplate{navigation symbols}{}

\begin{frame}
	{
		\huge\centering
		~\\~\\~\\
		Спасибо за внимание
	}
	~\\
	\vspace{6.28em}
	nickkolok@mail.ru, avdeev@math.vsu.ru
	\\
	github.com/nickkolok
	\\
	gitlab.com/nickkolok
	\\
	arxiv.org/a/avdeev\_n\_1.html
\end{frame}

\end{document}

### 2021-12-03 2021_Lob/Zvolinsky_et_al.tex
[2021Lob] Rename the main file

--- 2021-11-14 2021_Lob/Article.tex
@@
 	\noindent
-	УДК: 514.822
+	УДК: УДК 511.95 + 514.112
 
@@
 	\textit{Ключевые слова:} множество точек на плоскости с целочисленными расстояниями, диаметр множества,
-	мощность множества, характеристика множества, отыскание множеств точек с целочисленными расстояниями.
+	мощность множества.
 
@@
 	\textbf{Определение 1.} {\it Множеством $\mathcal P$ точек на плоскости с целочисленными расстояниями (сокращённо МЦР) называют
-	такое множество не лежащих на одной прямой прямой точек, что расстояние между любыми двумя точками
+	такое множество не лежащих на одной прямой точек, что расстояние между любыми двумя точками
 	из $\mathcal P$ есть целое число.}
@@
 
-	П. Эрдёш и Н. Эннинг [1, 2] доказали, что подобные множества состоят из конечного числа точек, т.е. 
-	имеют конечную мощность (cardinality). Кроме того, данные множества обладают и конечным диаметром
-	(diameter), определяемым как
-	$$
+	Всякое МЦР конечно~\cite{bib:01, bib:02} и, следовательно, имеет конечный диаметр
+	\begin{equation*}
 	    \operatorname{diam} (\mathcal P) = \max_{P_1, P_2 \in \mathcal P}
 	    |P_1 P_2|,
-	$$
-	где $|P_1 P_2|$ есть обычное расстояние между точками. Позже А. Кемнитц
-	[3] обнаружил, что у каждого такого множества есть ещё и характеристика.
-
-	Множества точек на плоскости с целочисленными расстояниями весьма разнообразны. Взять к примеру веерные множества (facher sets), которые в
-	свою очередь также бывают очень разные [4, 5].
+	\end{equation*}
+	где $|P_1 P_2|$ есть обычное расстояние между точками.
 
@@
 
-	\textbf{Определение 2.} {\it Множества точек на плоскости с целочисленными расстояниями мощности $n$, для которых можно найти
-		такую прямую $\ell_1$, что $(n - 1)$ точек лежит на $\ell_1$, будем называть веерными (facher sets).}
+	\textbf{Определение 2.} {\it {\bf Веерными (facher)}%
+		\footnote{
+			По типично европейской традиции, заложенной С. Курцем, классам МЦР даются оригинальные названия;
+			например, в~\cite{crabs} изучаются МЦР классов {\it crab} и даже {\it semi-crab}
+		}
+		называют МЦР мощности $n$, для которых можно найти
+		такую прямую $\ell_1$, что $(n - 1)$ точек лежит на $\ell_1$.}
 
@@
 
-	Существуют и более интересные множества: крестовые (cross sets), пирамидальные (pyramid set), ножничные (scissors sets). В статье [6] введена классификация подобных множеств, 
-	а в [7] даются примеры, в классификацию не ложащиеся.
+	Веерные множества встречаются чаще остальных,
+	и вопрос о существовании веерного множества с заданными параметрами хорошо изучен в~\cite{bib:04} и \cite[Теорема 5]{bib:05}.
+	Частичная классификация МЦР введена в~\cite{bib:06}; примеры множеств, в неё не входящих, даны в~\cite{bib:07}.
+	Для построения каких-либо общих рассуждений, касающихся существования МЦР некоторого специального вида,
+	необходимы наглядные примеры.
+	Поиск таких примеров ведётся на ЭВМ и может занимать продолжительное время.
+	Авторами разработано ПО на языке NodeJS и опубликовано по адресу gitlab.com/nickkolok/ips-algo
+	под свободной лицензией GNU GPLv3.
 
-	Для построения каких-либо общих рассуждений, касающихся данных множеств, необходимы наглядные примеры. Поиск таких 
-	примеров осуществляется с помощью аппаратных средств ЭВМ.
-
-	\begin{center}
-		{\bf Ножничные множества.}
-	\end{center}
+	Цель данной работы~--- привести примеры МЦР специального вида достаточно большой мощности,
+	найденные на ЭВМ.
 
@@
 
-    \textbf{Определение 3.} {\it Ножничными (scissors sets) будем называть такие множества точек на плоскости с целочисленными 
-    расстояниями мощности $n$, для которых можно единственным образом найти такие прямые $\ell_1$, $\ell_2$, 
-    что $k$ точек лежат на прямой $\ell_1$, а $(n - k)$~--- на прямой $\ell_2$, причём $k \geqslant 2$ и прямые 
-    $\ell_1$ и $\ell_2$ не являются ни параллельными, ни перпендикулярными.}
+    \textbf{Определение 3.} {\it {\bf Ножничными (scissors)} назовём МЦР мощности $n$,
+    для которых можно единственным образом найти такие прямые $\ell_1$, $\ell_2$,
+    что $k$ точек лежат на $\ell_1$, а $(n - k)$~--- на $\ell_2$, причём $k \geqslant 2$ и
+    $\ell_1$ и $\ell_2$ не параллельны и не перпендикулярны.}
 
@@
 
-    Для облегчения восприятия применяется используемая в [5, 8, 9] запись вида:
-    $\sqrt p/q * \{(x_1, y_1), \ldots, (x_n, y_n)\}$ ($p$~--- это характеристика), означающая, 
+    Для облегчения восприятия применяется используемая в \cite{bib:05, bib:08, bib:09} запись вида:
+    $\sqrt p/q * \{(x_1, y_1), \ldots, (x_n, y_n)\}$ ($p$~--- это характеристика МЦР, введённая в~\cite{bib:03}), означающая,
     что каждую абсциссу нужно умножить на $1/q$, а каждую ординату~--- на $\sqrt p/q$, т. е.
-    $$
+    \begin{equation*}
         \sqrt p/q * \{(x_1, y_1), \ldots, (x_n, y_n)\} =
@@
         \ldots, \left(\frac{x_n}q, \frac{y_n\sqrt p}q\right)\right\}.
-    $$
+    \end{equation*}
 
@@
     	\includegraphics[width=1\linewidth]{25_1544076_91.pdf}
-    	\caption{МЦР мощности 25, диаметр 1544076}
+    	\caption{Ножничная МЦР мощности 25, диаметр 1544076}
     \end{figure}
@@
     \begin{align*}
-        &\sqrt{91}/1 * \{(\pm 772038, 0),
+        &\sqrt{91}/1 * \{
+		\pm(80784, 12240),
+        (0, 0),
+        (\pm 772038, 0),
         (\pm 470016, 0),
         (\pm 193732, 0),
-        (\pm 171564, 0),
+        (\pm 171564, 0),\\&
         (\pm 126684, 0),
-        (\pm 117504, 0), \\
-        &(80784, 12240),
-        (-80784, -12240),
+        (\pm 117504, 0),
         (\pm 72960, 0),
@@
         (\pm 65892, 0),
-        (\pm 42891, 0), \\
-        &(\pm 26112, 0),
-        (0, 0)\} \text{ (Рис. 1)}. 
+        (\pm 42891, 0),
+        (\pm 26112, 0)
+        \} \\&
+        \text{ (Рис. 1)}.
     \end{align*}
 
-    \begin{center}
-        {\bf Рельсовые множества.}	
-    \end{center}
+    \textbf{Определение 4.} {\it {\bf Рельсовыми (rails)} называют МЦР
+    мощности $n$, для которых можно единственным образом найти такие прямые $\ell_1 \| \ell_2$,
+    что $k$ точек лежат на $\ell_1$,
+    а $(n - k)$~--- на  $\ell_2$, причём $k \geqslant 2$.}
 
-    \medskip
-
-    \textbf{Определение 4.} {\it Рельсовыми (rails sets) будем называть такие множества точек на плоскости с целочисленными расстояниями мощности $n$, 
-    для которых можно единственным образом найти такие прямые $\ell_1$, $\ell_2$, что $k$ точек лежат на прямой $\ell_1$,
-    а $(n - k)$~--- на прямой $\ell_2$, причём $k \geqslant 2$.}
-
-	\medskip
-
-	\begin{figure}[h!]
+	\begin{figure}
 		\centering
 		\includegraphics[width=1\linewidth]{28_158064_154.pdf}
-		\caption{МЦР мощности 28, диаметр 158064}
+		\caption{Рельсовая МЦР мощности 28, диаметр 158064}
 	\end{figure}
@@
     	\includegraphics[width=1\linewidth]{32_1598904_154.pdf}
-    	\caption{МЦР мощности 32, диаметр 1598904}
+    	\caption{Рельсовая МЦР мощности 32, диаметр 1598904}
     \end{figure}
@@
     	\includegraphics[width=1\linewidth]{44_456989104_154.pdf}
-    	\caption{МЦР мощности 44, диаметр 456989104}
+    	\caption{Рельсовая МЦР мощности 44, диаметр 456989104}
     \end{figure}
@@
 
-    \begin{center}
-    	{\bf Множества точек на плоскости с целочисленными расстояниями, не входящие в общую классификацию.}
-    \end{center}
+    {\bf Пример  МЦР, не входящего в существующую классификацию.}
 
@@
     	\includegraphics[width=1\linewidth]{23_243663771_1.pdf}
-    	\caption{МЦР мощности 23, диаметр 243663771}
+    	\caption{МЦР мощности 23, диаметр 243663771, не входящее в классификацию}
     \end{figure}
@@
 
-	{\footnotesize Работа выполнена в Воронежском университете при поддержке РНФ, грант 19-11-00197.}
+	{Работа выполнена в ВГУ при поддержке РНФ, грант 19-11-00197.}
 
@@
 		Anning\,N.\,H., Erd{\"o}s\,P. {\it Integral distances} //~Bulletin of the American Mathematical Society.~--
-		1945.~-- Vol.~51.~-- no.~8.~-- Pp.~598--600.~-- DOI: 10.1090/S0002-9904-1945-08407-9.
+		1945.~-- Vol.~51.~-- no.~8.~-- С.~598--600.~-- DOI: 10.1090/S0002-9904-1945-08407-9.
 
@@
 		Erd{\"o}s\,P. {\it Integral distances} //~Bulletin of the American Mathematical Society.~--
-		1945.~-- Vol.~51.~-- no.~12.~-- P. 996.~-- DOI: 10.1090/S0002-9904-1945-08490-0.
+		1945.~-- Vol.~51.~-- no.~12.~-- С. 996.~-- DOI: 10.1090/S0002-9904-1945-08490-0.
 
-		\bibitem{bib:03}
-		Kemnitz\,A. {\it Punktmengen mit ganzzahligen Abst{\"a}nden.}~-- 1988.
+		\bibitem{crabs}
+		Antonov\,A.\,R., Kurz\,S. {\it Maximal integral point sets over $\mathbb{Z}^2$}
+		//International Journal of Computer Mathematics.~-- 2010.~-- Т. 87.~-- №. 12.~-- С. 2653-2676.~-- arXiv: 0804.1280[math.CO].
 
@@
 		Zvolinsky\,R.\,E. {\it Facher integral point sets with particular distances of arbitrary cardinality} //~Актуальные
-		проблемы прикладной математики, информатики и механики~-- сборник трудов Международной научной конференции.~--
-		Научно-исследовательские публикации, 2020.~ Pp.~668--674.
+		проблемы прикладной математики, информатики и механики~-- сб. тр. Междунар. науч. конф.~-- 2020.~-- С.~668--674.
 
 		\bibitem{bib:05}
-		Авдеев\,Н.\,Н., Семёнов\,Е.\,М.
-{\it Множества точек с целочисленными расстояниями на плоскости и в евклидовом пространстве} //~Математический форум (Итоги науки.
-Юг России).~-- 2018.~-- Pp.~217--236.
+		Авдеев\,Н.\,Н., Семёнов\,Е.\,М. {\it Множества точек с целочисленными расстояниями на плоскости и в евклидовом пространстве}
+		//~Математический форум (Итоги науки. Юг России).~-- 2018.~-- С.~217--236.
 
@@
 		Avdeev\,N.\,N., Zvolinsky\,R.\,E., Momot\,E.\,A. {\it On particular diameter bounds for integral point sets in higher dimensions.}~--
-		2019.~-- arXiv: 1909.10386.~-- URL: https://ui.adsabs.harvard.edu/abs/2019arXiv190910386A/abstract.
+		2019.~-- arXiv: 1909.10386.
 
@@
 		Авдеев\,Н.\,Н. {\it On integral point sets in special position} //~Некоторые вопросы анализа, алгебры, геометрии и
-		математического образования: материалы международной молодежной научной школы <<Актуальные направления математического 
+		математического образования: мат-лы междунар. молодежной научной школы <<Актуальные направления математического
 		анализа и смежные вопросы>>.~-- 2018.~-- Vol.~8.~-- Pp.~5--6.
@@
 		Авдеев\,Н.\,Н. {\it Об отыскании целоудалённых множеств специального вида}
-		//~Актуальные проблемы прикладной математики, информатики и механики~-- сборник трудов 
-		Международной научной конференции.~-- Научно-исследовательские публикации, 2018.~--
-		Pp.~492--498.
+		//~Актуальные проблемы прикладной математики, информатики и механики~-- сб. тр.
+		Международной науч. конф.~-- 2018.~-- С.~492--498.
 
+		\bibitem{bib:03}
+		Kemnitz\,A. {\it Punktmengen mit ganzzahligen Abst{\"a}nden.}~-- 1988.
 	\end{thebibliography}
@@
 
-    	\textit{Keywords:} planar integral point sets, diameter of an integral point set, cardinality of an integral point set, characteristic of an integral point set, search of integral point sets.
+    	\textit{Keywords:} planar integral point sets, diameter of an integral point set, cardinality of an integral point set.
 

### 2024-06-19 char-4k-bound/char-4k-bound.tex
[char4k] A stub: AJC

--- 2020-03-25 pps-linear-bound/pps-linear-lower-bound.tex
@@
 
-%\usepackage[backend=biber,style=ajc2020unofficial]{biblatex}
-%\addbibresource{../common/notmy.bib}
-%\addbibresource{../common/my.bib}
-
-\usepackage{cite}
+\usepackage[backend=biber,style=ajc2020unofficial]{biblatex}
+\addbibresource{../common/notmy.bib}
+\addbibresource{../common/my.bib}
+
+%\usepackage{cite}
 
@@
 		d \geq |M_{\max} - M_{\min}| = \sqrt{(x_{\max} - x_{\min})^2 + (y_{\max} - y_{\min})^2}
-		\geq
 		\\\geq
@@
 	Indeed, for $n \geq 21492$ we have
-	\begin{multline*}
+	\begin{align*}
 		\varphi_n
@@
 		\frac{1}{n-2} +
+		&
 		\sqrt{
@@
 		}
-		=
-		\\=
+		\\&=
 		\frac{1}{\sqrt{n-2}}
@@
 		}
-		\leq
-		\\ \leq
+		\\&\leq
 		\frac{1}{\sqrt{n-2}}
@@
 		\right)
-		\leq
-		\\ \leq
+		\\&\leq
 		\frac{1}{\sqrt{n-2}}
@@
 		.
-	\end{multline*}
+	\end{align*}
 \end{proof}
@@
 				cr(M_1,M_2) \cap cr(M_3,M_4) = (l^1_1 \cup l^1_2)\cap(l^2_1 \cup l^2_2) =
-				\\=
 				(l^1_1 \cap l^2_2) \cup (l^1_2 \cap l^2_1)
@@
 		0.46530... =
-		3^{1/4} \cdot 2^{-3/2} >
+		3^{1/4} \cdot 2^{-3/2}
 		\\ >
 		\gamma = \frac{\sqrt{16 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}+\frac{14791716}{{{\left( 2593 \sqrt{647}-5823\right) }^{2}}}}-\frac{3846}{2593 \sqrt{647}-5823}}{8 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}}
-		>\\
+		\\
 		> 0.45557
@@
 		\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-1}} - 1\right)
-		=
 		\\=
@@
 		\frac{1}{2}\left(\left(\frac{c_J - a/c_J}{2}\right)^2 + a\right)^{1/2}
-		=
 		\\=
@@
 		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 + \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2    \right)^{1/2}
-		=
 		\\=
@@
 
-%\printbibliography
-
-%\printbibitembibliography
-
-\begin{thebibliography}{99}
-%
-\bibitem{anning1945integral}
-N. H. Anning and P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{8} (1945), 598–600, \textsc{doi}: \href
-{https://doi.org/10.1090/S0002-9904-1945-08407-9} {\nolinkurl {10.1090/S0002-9904-1945-08407-9}}.
-%
-\bibitem{antonov2008maximal}
-A. R. Antonov and S. Kurz, Maximal integral point sets over {$\mathbb {Z}^2$}, \emph{Int. J. Comput. Math.} 87.\textbf{12}
-(2008), 2653–2676, \textsc{doi}: \href {https://doi.org/10.1080/00207160902993636} {\nolinkurl {10.1080/00207160902993636}},
-arXiv: \href {http://arxiv.org/abs/0804.1280} {\nolinkurl {0804.1280}}.
-%
-\bibitem{our-mz-rus-translit}
-N. N. Avdeev and E. M. Semenov, On point sets in the plane with integral distances (О множествах точек на плоскости с
-целочисленными расстояниями), \emph{Matematicheskie zametki} 100.\textbf{5} (2016), 757–761.
-%
-\bibitem{our-vmmsh-2018-translit}
-N. N. Avdeev and E. M. Semenov, Integral point sets in the plane and Euclidean space (Множества точек с целочисленными
-расстояниями на плоскости и в евклидовом пространстве), \emph{Matematicheskij forum (Itogi nauki. Yug Rossii)} (2018),
-217–236.
-%
-\bibitem{avdeev2019particular}
-N. N. Avdeev, R. E. Zvolinsky, and E. A. Momot, On particular diameter bounds for integral point sets in higher dimensions
-(2019), arXiv: \href {http://arxiv.org/abs/1909.10386} {\nolinkurl {1909.10386}}, \textsc{url}: \url
-{https://ui.adsabs.harvard.edu/abs/2019arXiv190910386A/abstract}.
-%
-\bibitem{bat2018number}
-G. Bat-Ochir, On the number of points with pairwise integral distances on a circle, \emph{Discrete Appl. Math.} 254 (2018),
-17–32, \textsc{doi}: \href {https://doi.org/10.1016/j.dam.2018.07.004} {\nolinkurl {10.1016/j.dam.2018.07.004}}.
-%
-\bibitem{costa2013valid}
-A. Costa, Valid constraints for the point packing in a square problem, \emph{Discrete Appl. Math.} 161.\textbf{18} (2013),
-2901–2909.
-%
-\bibitem{erdos1945integral}
-P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{12} (1945), 996, \textsc{doi}: \href
-{https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl {10.1090/S0002-9904-1945-08490-0}}.
-%
-\bibitem{garibaldi2005erdos}
-J. Garibaldi and A. Iosevich, The Erdős distance problem: lecture notes (2005).%
-\bibitem{harborth1993upper}
-H. Harborth, A. Kemnitz, and M. Möller, An upper bound for the minimum diameter of integral point sets, \emph{Discrete
-Comput. Geom.} 9.\textbf{4} (1993), 427–432, \textsc{doi}: \href {https://doi.org/10.1007/bf02189331} {\nolinkurl
-{10.1007/bf02189331}}.
-%
-\bibitem{kurz2005characteristic}
-S. Kurz, On the characteristic of integral point sets in {$\mathbb {E}^m$}, \emph{Australas. J. Combin.} 36 (2006), 241–248,
-arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl {math/0511704}}.
-%
-\bibitem{kurz2008bounds}
-S. Kurz and R. Laue, Bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 39 (2007), 233–240,
-arXiv: \href {http://arxiv.org/abs/0804.1296} {\nolinkurl {0804.1296}}.
-%
-\bibitem{kurz2008minimum}
-S. Kurz and A. Wassermann, On the minimum diameter of plane integral point sets, \emph{Ars Combin.} 101 (2011), 265–287,
-arXiv: \href {http://arxiv.org/abs/0804.1307} {\nolinkurl {0804.1307}}.
-%
-\bibitem{locatelli2002packing}
-M. Locatelli and U. Raber, Packing equal circles in a square: a deterministic global optimization approach, \emph{Discrete Appl.
-Math.} 122.\textbf{1-3} (2002), 139–166.
-%
-\bibitem{noll1989nclusters}
-L. C. Noll and D. I. Bell, {$n$}-clusters for {$1<n<7$}, \emph{Math. Comp.} 53.\textbf{187} (1989), 439–444.
-%
-\bibitem{nozaki2013lower}
-H. Nozaki, Lower bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 56 (2013), 139–143.
-%
-\bibitem{piepmeyer1996maximum}
-L. Piepmeyer, The maximum number of odd integral distances between points in the plane, \emph{Discrete Comput. Geom.}
-16.\textbf{1} (1996), 113–115, \textsc{doi}: \href {https://doi.org/10.1007/bf02711135} {\nolinkurl {10.1007/bf02711135}}.
-%
-\bibitem{solymosi2003note}
-J. Solymosi, Note on integral distances, \emph{Discrete Comput. Geom.} 30.\textbf{2} (2003), 337–342, \textsc{doi}: \href
-{https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.
-%
-\bibitem{solymosi2010question}
-J. Solymosi and F. De Zeeuw, On a question of Erdős and Ulam, \emph{Discrete Comput. Geom.} 43.\textbf{2} (2010), 393–401,
-arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl {0806.3095}}.
-\end{thebibliography}
+\printbibliography
 

### 2024-06-27 char-4k-bound/char-4k-bound.tex
[4k3] WIP - sets of integers

@@
 
+TODO
+
 \section{Introduction}
@@
 
-\section{Collinear subsets of integral point sets}
-Integral point sets with many collinear points are very dominating examples.
-In this section we estimate the cardinality of an intersection of an integral point set with a line segment.
+\section{Basic Notions and Results}
 
-\subsection{General results}
+In this Section, we provide proper definitions and list some known results.
 
-Assuming that the planar integral point set contains many collinear points,
-the following result holds.
-\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
-	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
-	at least $n^\delta$ collinear points there exists a $n_0 (\varepsilon)$
-	such that for all $n \geq n_0 (\varepsilon)$ we have
+For the sake of brevity, the following designations will be used for numeric sets
+of positive integers, non-negative integers and all the integers resp.:
 \begin{equation}
-		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
-		.
+	\mathbb{N} = \{1,2,3,4,...\},\quad \mathbb{N}_0 = \mathbb{N} \cup \{0\},
+	\quad
+	\mathbb{N}_\pm = \{0,\pm 1,\pm 2,\pm 3,\pm 4,...\}
 \end{equation}
-\end{theorem}
-However, the estimate is rather unsuitable for our needs,
-as it does not provide the values of all the constants.
-To obtain the required estimate, we now prove a generalization of~\cite[Lemma 3]{our-vmmsh-2018-translit}.
-
-\begin{definition}
-	For a line segment $M_1 M_2$ and an integer $k$, such that $-|M_1 M_2| < k < |M_1 M_2|$,
-	we define a $\rho(k,M_1, M_2)$-curve as the set of points $N$
-	for which the equality $|N M_1| - |N M_2| = k$ holds.
-\end{definition}
-So, in the planar case a $\rho(k,M_1, M_2)$-curve
-is a branch of a hyperbola for $k\neq 0$ and the perpendicular bisector of line segment $M_1 M_2$ for $k=0$.
-
-\begin{proposition}
-	\label{obs:rho_curves}
-	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
-	the equality $|M_1 M_2| = |M_3 M_4| =h $ holds, line segments $M_1 M_2$ and $M_3 M_4$ do not coincide
-	and vectors $\overrightarrow{M_1 M_2}$ and $\overrightarrow{M_3 M_4}$ are of equal direction,
-	then for a fixed integer $k$, $|k| < h$, the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve
-	do not intersect.
-\end{proposition}
-
-\begin{proof}
-	For $k=0$, the claim is obvious.
-	Indeed, in this case the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve
-	are distinct hyperplanes orthogonal to $l$ and thus do not intersect.
-
-	For $k\neq 0$, suppose the contrary.
-	Let $N$ be a point in the intersection.
-
-	Consider a plane $\Pi \supset (l \cup \{N\})$
-	with a Cartesian coordinate system such that $M_i = (x_i,0)$ for $1 \leq i \leq 4$.
-	Let $R_1$ and $R_2$ be the projections
-	of the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve onto $\Pi$.
-
-	Then $R_1$ is a branch of a hyperbola defined by the equation
-	\begin{equation}
-		x- \frac{x_1+x_2}{2}=a(h,k)\frac{\sqrt{y^2+b^2(h,k)}}{b(h,k)}
-	\end{equation}
-	and $R_2$ is a branch of a hyperbola defined by the equation
-	\begin{equation}
-		x- \frac{x_3+x_4}{2}=a(h,k)\frac{\sqrt{y^2+b^2(h,k)}}{b(h,k)}
-		,
-	\end{equation}
-	where $a(h,k)$ and $b(h,k)$ are coefficients which depend on $h$ and $k$.
-	For $N=(x_N, y_N) \in R_1 \cap R_2$ we have
-	\begin{equation}
-		\begin{cases}
-			x_N- \dfrac{x_1+x_2}{2}=a(h,k)\dfrac{\sqrt{y_N^2+b^2(h,k)}}{b(h,k)}
-			\\[4pt]\\
-			x_N- \dfrac{x_3+x_4}{2}=a(h,k)\dfrac{\sqrt{y_N^2+b^2(h,k)}}{b(h,k)}
-		\end{cases}
-	\end{equation}
-	and, thus, $x_1+x_2 = x_3+x_4$.
-	On the other hand, $x_2 - x_1 = x_4 - x_3 = h$,
-	and we conclude that $x_1 = x_3$ as well as $x_2 = x_4$.
-	So, line segments $M_1 M_2$ and $M_3 M_4$ coincide.
-	The contradiction concludes the proof.
-\end{proof}
 
 \begin{lemma}
-	\label{lem:2k-1_segments}
-	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $l$ be a straight line.
-	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset l$ with endpoints in $M$,
-	such that $|\Delta_i| = k$.
-\end{lemma}
-\begin{proof}
-	Consider a point $N\in M \setminus l$.
-	Let $\Delta_i = M_i^1 M_i^2$, and
-	let vectors $\overrightarrow{M_i^1 M_i^2}$ and $\overrightarrow{M_j^1 M_j^2}$ be of equal direction for every $i$ and $j$.
-	Then for each $i$ there is a $\rho(n_i, M_i^1, M_i^2)$-curve containing $N$.
-	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
-	otherwise the $\rho(n_i,M_i^1, M_i^2)$-curve and the $\rho(n_i,M_j^1, M_j^2)$-curve, $j\neq i$, do not intersect.
-	There can be only $2k-1$ distinct values for $n_i$,
-	so there are at most $2k-1$ distinct segments $\Delta_i$.
-\end{proof}
-
-Setting $k=1$ in Lemma~\ref{lem:2k-1_segments},
-we obtain the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018-translit}).
-\begin{corollary}
-	\label{cor:only_one_distance_1_on_straight_line}
-	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $l$ be a straight line.
-	Then there is at most one pair of points $M_1,M_2\in M \cap l$
-	such that $|M_1 M_2| = 1$.
-\end{corollary}
-
-\begin{lemma}
-	\label{lem:line_segment_with n_squared_plus_one_points}
-	Let $\Delta$ be a straight line segment, $|\Delta|=l$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = n^2 + 1$.
-	Then
-	\begin{equation}
-		l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
-		.
-	\end{equation}
+	\label{lem:triangle_4k_plus_3}
+	Any triangle with sides $a \leq b \leq c$, where $c=a+b-1$ and $a, b, c \in \mathbb{N}$,
+	has a characteristic of the form $p=4k+3$, $k\in \mathbb{N}_0$.
 \end{lemma}
@@
 \begin{proof}
-	Any $n^2+1$ points, including the endpoints of $\Delta$, split the segment $\Delta$ into $n^2$
-	sequential segments $\Delta_i$.
-	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
-	at most three segments of length 2, etc.
-	The following two expressions for sums conclude the proof:
-	\begin{equation}
-		1 + \sum_{k=1}^n (2k-1) = n^2 + 1
-		,
-	\end{equation}
-	\begin{equation}
-		\sum_{k=1}^n k(2k-1) = \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
-		.
-	\end{equation}
+	Indeed, let the triangle $ABC$ satisfy: $|BC|=a$, $|AC|=b$.
+Then, from the triangle inequality $|AB| < |BC|+|AC|$, we can represent the length of side $AB$ as: $|AB|=a+b-s$, $s \in \mathbb{N}$ and $s \leq a$.
+For $s=1$, the conditions of statement 6 are satisfied: $s$ is an odd number and $s \leq a$.
+	Therefore, the triangle with sides $a$, $b$, and $c=a+b-1$ has a characteristic of the form $p=4k+3$, $k\in \mathbb{N}_0$.
 \end{proof}
-
-Now we will estimate the length of a line segment that intersects an integral point set
-by an arbitrary number of points.
-
-\begin{lemma}
-	\label{lem:line_segment_cardinality}
-	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = t$.
-	Then
-	\begin{equation}
-		\label{eq:estimate_for_segment_length}
-		b\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}
-		.
-	\end{equation}
-\end{lemma}
-
-\begin{proof}
-	Let $f(k)$ denote the minimal length of a line segment
-	that intersects an integral point set by $k$ points.
-	We observe that $f(k) > f (k-1)$.
-	Due to Lemma~\ref{lem:line_segment_with n_squared_plus_one_points},
-	$f(n^2+1) \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$.
-
-	%Consider $t$ such that $n^2 + 1 \leq t \leq (n+1)^2$. %,
-	%then $f(t) > f(n^2 +1)$.% and $n \geq \sqrt{t} - 1$.
-	For $t\in\mathbb{N}$ the inequality $(\sqrt{t} - 1)^2 +1 \leq t$ holds,
-	thus
-	%and derive that
-	\begin{equation}
-		\label{eq:line_segment_length}
-		f(t) \geq f((\sqrt{t} - 1)^2 +1) \geq \frac{2}{3}(\sqrt{t} - 1)^3+\frac{1}{2}(\sqrt{t} - 1)^2-\frac{1}{6}(\sqrt{t} - 1)
-		=
-		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2}
-		.
-	\end{equation}
-\end{proof}
-
-\subsection{Special results}
-
-In~\cite{kurz2008bounds}, thorough investigation of planar integral point sets with diameter at most 10~000
-is carried out;
-thus, we can restrict our attention to intersections of planar integral point sets with line segments of length more than 10~000.
-
-Lemma~\ref{lem:line_segment_cardinality} leads to the following proposition.
-
-\begin{proposition}
-	\label{obs:estimate_points_on_straight_line}
-	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = k$ and $b>10000$.
-	Then $k \leq \gamma_2 b + 6$,
-	where
-	\begin{equation}
-		\gamma_2 = \frac{3846}{2593 \sqrt{647}-5823} = 0.063958...
-	\end{equation}
-\end{proposition}
-
-\begin{proof}
-	We know the maximum number of points for planar integral point sets of diameters at most $10000$.
-	Thus, we are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
-	holds for $b > 10000$.
-	So, let us consider $t\geq 647$ and find a coefficient $\gamma_2$,
-	such that the inequality
-	\begin{equation}
-		\label{eq:inequality_for_linear_estimate_of_segment_length}
-		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2} \geq \frac{t-6}{\gamma_2}
-	\end{equation}
-	holds for all $t\geq 647$.
-	For such $t$,
-	the left-hand side of~\eqref{eq:inequality_for_linear_estimate_of_segment_length} obviously grows faster than the right-hand side.
-	Turning~\eqref{eq:inequality_for_linear_estimate_of_segment_length} into the same equality
-	and solving it for $t=647$, we obtain the required estimate.
-\end{proof}
-
-To conclude, we remark that
-Corollary~\ref{cor:only_one_distance_1_on_straight_line}
-and
-Proposition~\ref{obs:estimate_points_on_straight_line}
-are the only results of this section
-that are used in the proofs below;
-all the other are auxiliary.
-
-\section{Planar integral point set in a square}
-
-In this section, we develop some methods that employ the fact that every planar integral point set is situated in a square.
-Then we establish a connection between distances in integral point sets and the Point Packing in the Square problem.
-
-\subsection{General results}
-
-Let us introduce the problem, basic notions and results.
-
-\begin{problem}[Point Packing in a Square (PPS)~\cite{locatelli2002packing,costa2013valid}]
-	\label{problem:PPS}
-	Given an integer $k > 1$, place $k$ points in the unit square $U = [ 0 , 1 ]^2$ such that their
-	minimum pairwise distance $m$ is maximal.
-\end{problem}
-
-\begin{definition}
-	For each $k > 1$, the corresponding maximal distance $m$ from Problem~\ref{problem:PPS}
-	is called the $k$-th PPS coefficient and denoted by $\varphi_k$.
-\end{definition}
-So, it's impossible to place $k$ points in a unit square in such a way that each pairwise distance of the points is greater than $\varphi_k$.
-
-\begin{theorem}
-	\label{thm:varphi_k_bounds}
-	\cite{costa2013valid}
-	For every $k\geq 2$ the following inequality holds:
-	\begin{equation*}
-		\sqrt{\frac{2}{k\sqrt{3}}}
-		\leq
-		\varphi_k
-		\leq
-		\frac{1}{k-1} +
-		\sqrt{
-			\frac{1}{(k-1)^2}
-			+
-			\frac{2}{(k-1)\sqrt{3}}
-		}
-	\end{equation*}
-\end{theorem}
-
-Now we will estimate the side length of a square which contains a given planar integral points set.
-
-\begin{lemma}
-	\cite[Lemma 4]{our-vmmsh-2018-translit}
-	\label{lem:square_container}
-	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
-	Then $M$ is situated in a square of side length $d$.
-\end{lemma}
-
-\begin{proof}
-	Consider $M_1, M_2 \in M$ such that $|M_1 - M_2| = d$
-	and a Cartesian coordinate system such that
-	$M_1 = (-d/2; 0)$, $M_2 = (d/2; 0)$.
-
-	Then for $M_i =(x_i, y_i)\in M$, $i > 2$,
-	the inequality $|x_i| < d/2$ holds
-	(otherwise we get $|M_i - M_1| > d$ or $|M_i - M_2| > d$).
-	Moreover, $|y_i| < d$ (otherwise we get $|M_i - M_1| > d$ or $|M_i - M_2| > d$).
-
-	Set $y_{\max} = \max_{i} y_i$, $y_{\min} = \min_{i} y_i$, $M_{\max}=(x_{\max}, y_{\max})$, $M_{\min}=(x_{\min}, y_{\min})$
-	(in case of multiple points providing the maximum or the minimum, the choice is arbitrary).
-	Then
-	\begin{multline}
-		d \geq |M_{\max} - M_{\min}| = \sqrt{(x_{\max} - x_{\min})^2 + (y_{\max} - y_{\min})^2}
-		\\\geq
-		\sqrt{(y_{\max} - y_{\min})^2} =
-		y_{\max} - y_{\min}
-	\end{multline}
-	Thus, the sides of the square that contains $M$ are $x=\pm d/2$,
-	$y=y_{\max}$ and $y=y_{\max} - d$.
-\end{proof}
-
-\subsection{Special results}
-
-It is clear that for $n>123$ we have $d(2,n) \geq d(2,123)$.
-Based on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
-\cite{kurz2008bounds}, we derive the following proposition.
-\begin{proposition}
-	\label{obs:4_leq_n_leq_21491}
-	The inequality
-	\begin{equation}
-		d(2,n) \geq 3^{1/4}\cdot2^{-3/2} \cdot n
-	\end{equation}
-	holds for $4 \leq n \leq 21491$.
-\end{proposition}
-
-Therefore, we will focus on planar integral points sets of more than 21491 points.
-
-Now we will use the upper bound of Theorem~\ref{thm:varphi_k_bounds}
-to estimate $\varphi_n$ with $\beta/\sqrt{n-2}$.
-
-\begin{proposition}
-	\label{varphi_n_where_n_geq_21492}
-	For $n \geq 21492$ we have
-	\begin{equation}
-		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{n-2}}
-		,
-	\end{equation}
-	where
-	\begin{equation}
-		\beta = \frac{1}{\sqrt{21490}} + \sqrt{ \frac{2}{\sqrt{3}} + \frac{1}{21490} } < 1.07464
-		.
-	\end{equation}
-\end{proposition}
-
-\begin{proof}
-	Indeed, for $n \geq 21492$ we have
-	\begin{align*}
-		\varphi_n
-		\leq
-		\varphi_{n-1}
-		\leq
-		\frac{1}{n-2} +
-		&
-		\sqrt{
-			\frac{1}{(n-2)^2}
-			+
-			\frac{2}{(n-2)\sqrt{3}}
-		}
-		\\&=
-		\frac{1}{\sqrt{n-2}}
-		\cdot
-		\frac{1}{\sqrt{n-2}}
-		 +
-		\sqrt{
-			\frac{1}{(n-2)^2}
-			+
-			\frac{2}{(n-2)\sqrt{3}}
-		}
-		\\&\leq
-		\frac{1}{\sqrt{n-2}}
-		\cdot
-		\frac{1}{\sqrt{21490}}
-		 +
-		\sqrt{
-			\frac{1}{(n-2)^2}
-			+
-			\frac{2}{(n-2)\sqrt{3}}
-		}
-		\\&=
-		\frac{1}{\sqrt{n-2}}
-		\left(
-		\frac{1}{\sqrt{21490}}
-		 +
-		\sqrt{
-			\frac{1}{n-2}
-			+
-			\frac{2}{\sqrt{3}}
-		}
-		\right)
-		\\&\leq
-		\frac{1}{\sqrt{n-2}}
-		\left(
-		\frac{1}{\sqrt{21490}}
-		 +
-		\sqrt{
-			\frac{1}{21490}
-			+
-			\frac{2}{\sqrt{3}}
-		}
-		\;
-		\right)
-		.
-	\end{align*}
-\end{proof}
-
-\section{Lower bound for the diameter}
-
-In this section,
-we improve the lower bound for the minimum diameter of planar integral point sets
-employing the results of two previous sections.
-
-\subsection{General results}
-
-\begin{definition}
-	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
-	the line through $M_1$ and $M_2$,
-	and the perpendicular bisector of line segment $M_1 M_2$.
-\end{definition}
-
-\begin{lemma}
-	\label{lem:intervals_cross}
-	If open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect,
-	%$M_1 \neq M_2$, $M_3 \neq M_4$,
-	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains 2 or 4 points.
-\end{lemma}
-
-\begin{proof}
-	Consider straight lines $l^1_1$, $l^1_2$, $l^2_1$, $l^2_2$
-	such that $cr(M_1,M_2) = l^1_1 \cup l^1_2$ and $cr(M_3,M_4) = l^2_1 \cup l^2_2$.
-
-	The cases $l^1_1 = l^2_1$, $l^1_2 = l^2_2$ and $l^1_1 = l^2_2$, $l^1_2 = l^2_1$ are impossible,
-	because the centers of open line segments $M_1 M_2$ and $M_3 M_4$ do not coincide as far as
-	the open line segments do not intersect.
-	So, all the possible cases (up to enumeration of the straight lines) are:
-	\begin{itemize}
-		\item
-			$l^1_1 = l^2_1$, $l^1_2 \neq l^2_2$.
-			The straight lines $l^1_2$ and $l^2_2$ are both orthogonal to $l^1_1$ and thus do not intersect;
-			thus, $cr(M_1,M_2) \cap cr(M_3,M_4) = l^1_1 = l^2_1$.
-		\item
-			$l^1_1$ is parallel to $l^2_1$ and $l^1_2$ is parallel to $l^2_2$.
-			Then
-			\begin{equation*}
-				cr(M_1,M_2) \cap cr(M_3,M_4) = (l^1_1 \cup l^1_2)\cap(l^2_1 \cup l^2_2) =
-				(l^1_1 \cap l^2_2) \cup (l^1_2 \cap l^2_1)
-				,
-			\end{equation*}
-			that is two points.
-		\item
-			$l^1_1$ is neither parallel nor orthogonal to $l^2_1$.
-			Then the intersection contains 4 points.
-	\end{itemize}
-\end{proof}
-
-\begin{remark}
-	It is important that we consider the intersection of open line segments in the claim of Lemma~\ref{lem:intervals_cross},
-	so e.g. the cases $M_1 = M_3$ and $M_3 \in M_1 M_2$ satisfy the conditions of Lemma~\ref{lem:intervals_cross}.
-\end{remark}
-
-\begin{lemma}
-	\label{lemma:quadr_diag_edges}
-	Let $ABCD$ be a convex quadrilateral on the plane.
-	Then $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
-	that is at least one diagonal is greater than at least one side.
-\end{lemma}
-
-\begin{proof}
-	Suppose the contrary.
-	Then $AC$ in the minimal side of the triangle $ABC$ and, thus, the angle $ABC$ is acute.
-	So are also angles $BCD$, $CDA$ and $DAB$.
-	We conclude that all angles of a convex quadrilateral $ABCD$ are acute, which is impossible.
-	The contradiction concludes the proof.
-\end{proof}
-
-\subsection{Special results}
-
-Now we are ready to prove the main theorem of the section.
-
-\begin{theorem}
-	\label{thm:main_estimate}
-	If $n\geq 4$, then $d(2,n) \geq \gamma (n - 2)$,
-	where
-	\begin{multline}
-		0.46530... =
-		3^{1/4} \cdot 2^{-3/2}
-		\\ >
-		\gamma = \frac{\sqrt{16 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}+\frac{14791716}{{{\left( 2593 \sqrt{647}-5823\right) }^{2}}}}-\frac{3846}{2593 \sqrt{647}-5823}}{8 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}}
-		\\
-		> 0.45557
-		> \frac{5}{11}
-		.
-	\end{multline}
-\end{theorem}
-
-\begin{proof}
-	For $4 \leq n \leq 21491$, the assertion of the theorem follows immediately from Proposition~\ref{obs:4_leq_n_leq_21491}.
-	Let us consider $M\in \mathfrak{M}(2,n)$, $n \geq 21492$, $\operatorname{diam} M = b$.
-	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square of side length $b$.
-	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1 M_2|$ and $|M_3 M_4|$
-	are minimal in $M$ ($M_2$ and $M_3$ may coincide).
-	Then $|M_1 M_2| \leq b\varphi_{n-1}$ and $|M_3 M_4| \leq b\varphi_{n-1}$.
-	Due to Lemma~\ref{lemma:quadr_diag_edges}, open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect
-	(otherwise they are not minimal).
-
-	Let $C = cr(M_1 M_2) \cap cr(M_3 M_4)$.
-	Each point $N\in M$ satisfies one of the following conditions:
-
-	a) $N$ belongs to $C$; overall at most $\gamma_2 b + 6$ points by Lemma~\ref{lem:intervals_cross} and Proposition~\ref{obs:estimate_points_on_straight_line};
-
-	b) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas
-	with one of $|M_3 M_4| - 1$ hyperbolas; overall at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
-
-	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas with $cr(M_3 M_4)$;
-	overall at most $4 (|M_1 M_2| - 1)$ points;
-
-	d) $N$ belongs to the intersection of one of $|M_3 M_4| - 1$ hyperbolas with $cr(M_1 M_2)$;
-	overall at most $4 (|M_3 M_4| - 1)$ points;
-	\\
-	(see~\cite{erdos1945integral} for details).
-	Summing up the above cases, we obtain the following estimate:
-	\begin{equation}
-		\label{eq:estimate_n_varphi_n_squared}
-		n \leq 4 b^2 \varphi_{n-1}^2 - 4 + \gamma_2 b + 6
-		.
-	\end{equation}
-	Proposition~\ref{varphi_n_where_n_geq_21492} turns estimate~\eqref{eq:estimate_n_varphi_n_squared}
-	into the following:
-	\begin{equation}
-		\label{eq:estimate_n_sqrt_n-2}
-		n-2 \leq 4 b^2 \frac{\beta^2}{n-2} + \gamma_2 b
-		,
-	\end{equation}
-	which obviously leads to inequality
-	\begin{equation}
-		1 \leq 4\beta^2 \left(\frac{b}{n-2}\right)^2  + \gamma_2 \frac{b}{n-2}
-		.
-	\end{equation}
-	Let us denote $\lambda = b/(n-2)$ and solve the following quadratic inequality for $\lambda$:
-	\begin{equation}
-		\label{eq:square_inequality_lambda}
-		4\beta^2 \lambda^2  + \gamma_2 \lambda - 1 \geq 0
-		.
-	\end{equation}
-	The discriminant is $\gamma_2^2 + 16 \cdot \beta^2$,
-	so we obtain the following estimate:
-	\begin{equation}
-		\frac{b}{n-2} = \lambda \geq \frac{-\gamma_2 + \sqrt{\gamma_2^2 + 16 \cdot \beta^2}}{8\beta^2}
-		.
-	\end{equation}
-	Calculating the expression above for our $\beta$ and $\gamma_2$, we conclude the proof.
-\end{proof}
-
-\begin{corollary}
-	If $n\geq 4$, then $d(2,n) > \frac{5}{11} n$.
-\end{corollary}
-
-\begin{proof}
-	For $4 \leq n \leq 21491$, the claim follows from Proposition~\ref{obs:4_leq_n_leq_21491} immediately.
-	For $n > 21491$, the inequality $0.45557(n-2) > \frac{5}{11}n$ holds.
-\end{proof}
-
-We can improve the constant in Theorem~\ref{thm:main_estimate},
-if some new exact values of $d(2,n)$ are found;
-however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$ from above.
-More precisely, we have the following theorem.
-
-\begin{theorem}
-	For every $\varepsilon > 0$ there exists a number $n_0$ such that the inequality
-	\begin{equation}
-		d(2,n) \geq n\cdot(3^{1/4}\cdot 2^{-3/2} - \varepsilon)
-	\end{equation}
-	holds for every $n>n_0$,
-	that is
-	\begin{equation}
-		\lim_{n\to\infty} \frac{d(2,n)}{n} \geq 3^{1/4}\cdot 2^{-3/2}
-		.
-	\end{equation}
-\end{theorem}
-
-\section{Integral point sets with distance 1}
-
-\subsection{Constructing planar integral point sets with distance 1}
-
-\begin{definition}
-	A set $M\in\mathfrak{M}(2,n)$ is called \textit{facher}
-	if $M$ consists of $n-1$ points on a straight line
-	and one point out of the line.
-\end{definition}
-
-\begin{definition}
-	A set $M\in\mathfrak{M}(m,n)$ is called \textit{optimal}
-	if $\operatorname{diam}M=d(m,n)$.
-\end{definition}
-
-\begin{definition}
-	\cite{kurz2005characteristic}
-	A squarefree number $q$ is called the \textit{characteristic} of $M\in\mathfrak{M}(2,\mathbb{N})$,
-	if for any points $M_1, M_2, M_3 \in M$ the area of triangle $M_1 M_2 M_3$
-	is $p_{1,2,3}\sqrt{q}$ for some rational $p_{1,2,3}$.
-\end{definition}
-For a given $M\in\mathfrak{M}(2,\mathbb{N})$, the characteristic is determined uniquely.
-
-Facher sets are the simplest planar integral point sets.
-It is known that for $9\leq n \leq 122$ all the optimal sets are facher~\cite{kurz2008minimum}.
-For every cardinality $n$ and every squarefree number $q$
-there exists a facher set $M\in\mathfrak{M}(2,n)$ with characteristic $q$~\cite[Theorem 5]{our-vmmsh-2018-translit}.
-In~\cite{antonov2008maximal}, the facher sets of characteristic 1 were investigated; they were called \textit{semi-crabs}.
-
-For every integer $n\geq 3$ Solymosi presented~\cite{solymosi2003note} a construction of a facher integral point set
-$M\in\mathfrak{M}(2,n)$
-such that equality $|M_1 M_2| = 2$ holds for some $M_1, M_2 \in M$.
-The constructed set has both odd and even distances.
-
-Now we improve Solymosi's result.
-
-\begin{construction}
-	\label{con:planar_set_with_minimeter_1}
-	Let us choose a positive integer $k > 1$ and set
-	\begin{equation}
-		a = 2^{2^k} - 1
-		.
-	\end{equation}
-	Then
-	\begin{equation}
-		\label{eq:a_equiv_3_mod_4}
-		a \equiv 3 \mod 4
-	\end{equation}
-	and, moreover,
-	\begin{multline}
-		a = \left(2^{2^{k-1}}\right)^2 - 1
-		=
-		\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-1}} - 1\right)
-		\\=
-		\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-2}} + 1\right) \cdot ... \cdot \left(2^{2^1} + 1\right) \left(2^2 - 1\right)
-		.
-	\end{multline}
-	We set $d_j = 2^{2^j} + 1$ for $1 \leq j \leq k-1$.
-	Then $d_j \equiv 1 \mod 4$.
-
-	Let $c_J = \prod_{j\in J} d_j$ for every subset of indices $J\subset I = \{1,2,...,k-1\}$
-	(and  $c_\varnothing = 1$).
-	We obtain
-	\begin{equation}
-		\label{eq:c_J_equiv_1_mod_4}
-		c_J\equiv 1 \mod{4}
-	\end{equation}
-	and, moreover, $a$ is divisible by $c_J$.
-
-	Statements~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4} yield $a/c_J \equiv 3 \mod 4$.
-	Let $b_J = (c_J - a/c_J)/2$, then $b_J \equiv 1 \mod 2$.
-	Let us further take $g_J = (c_J + a/c_J)/2$, then $g_J \equiv 0 \mod 2$.
-
-	Next we define the coordinates of the points as following:
-	\begin{equation}
-		M_{J\pm} =\left(\pm\frac{b_J}{2}, 0\right)
-		,
-		~~
-		N   =\left(0, \frac{\sqrt{a}}{2}\right)
-		.
-	\end{equation}
-	Then the distances are:
-	\begin{multline}
-		|N M_{J\pm}|
-		=
-		\left(\frac{b_J^2}{4} + \frac{a}{4}\right)^{1/2}
-		=
-		\frac{1}{2}\left(\left(\frac{c_J - a/c_J}{2}\right)^2 + a\right)^{1/2}
-		\\=
-		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 - \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2 + a\right)^{1/2}
-		=
-		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 + \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2    \right)^{1/2}
-		\\=
-		\frac{1}{2}\left(\left(\frac{c_J + a/c_J}{2}\right)^2\right)^{1/2}
-		=
-		\frac{1}{2}\left(\frac{c_J + a/c_J}{2}\right)
-		=
-		\frac{g_J}{2}
-		\in\mathbb{N}
-		,
-	\end{multline}
-	\begin{equation}
-		|M_{J_1 \pm}  M_{J_2 \pm}|
-		=
-		\left|\frac{b_{J_1}}{2} \pm \frac{b_{J_2}}{2}\right|
-		=
-		\left|\frac{b_{J_1} \pm b_{J_2}}{2}\right|
-		\in\mathbb{N}
-		.
-	\end{equation}
-	In particular, for $H = \{k-1\}$ we obtain $C_H = 2^{2^{k-1}}+1$ and
-	\begin{equation}
-		b_H =
-		\left( 2^{2^{k-1}}+1 - \frac{a}{2^{2^{k-1}}+1} \right)/2
-		=
-		\left(2^{2^{k-1}}+1 - \left(  2^{2^{k-1}}-1 \right) \right)/2
-		=
-		1
-		.
-	\end{equation}
-	Thus, one of the distances is
-	\begin{equation}
-		|M_{H+}  M_{H-}|
-		=
-		\left|\frac{b_{H}}{2} - \frac{-b_{H}}{2}\right|
-		=
-		\left|\frac{1}{2} - \frac{-1}{2}\right|
-		= 1
-		.
-	\end{equation}
-
-	Note that all the points $M_{J\pm}$ are distinct:
-	the equality $b_J =  b_K$ implies $J=K$;
-	the equality $b_J = -b_K$ implies $c_J = -c_K$ or $c_J = a / c_K$.
-	The first case is impossible because both $c_J$ and $c_K$ are positive;
-	the second case contradicts~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4}.
-
-	So, $M = \{ M_{J\pm}, N\}$ is indeed a planar integral point set of $2^k+1$ points,
-	and distance 1 occurs in $M$.
-	Since $k$ can be taken arbitrary large, it follows that we can construct
-	a planar integral point set of arbitrary large cardinality so that the distance 1 occurs in that set.
-\end{construction}
-
-\begin{remark}
-	\label{rem:planar_set_with_minimeter_1_of_3_points}
-	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ ad litteram,
-	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then
-	obtain an equilateral triangle of side length 1.
-	This triangle is obviously the optimal set in $\mathfrak{M}(2,3)$.
-\end{remark}
-
-\begin{remark}
-	\label{rem:planar_set_with_minimeter_1_of_4_and_5_points}
-	For $k=2$ in Construction~\ref{con:planar_set_with_minimeter_1},
-	we obtain the optimal set in $\mathfrak{M}(2,5)$ presented in~\cite[Fig. 1]{harborth1993upper}.
-	If we remove one point from it,
-	then we get one of the two optimal sets in $\mathfrak{M}(2,4)$.
-\end{remark}
-
-\subsection{Description of planar integral point sets with distance 1}
-
-%Now we need two more definitions.
-\begin{definition}
-	\cite{antonov2008maximal}
-	A set $M\in \mathfrak{M}(m,n)$ is \textit{maximal},
-	if there is no set $M'\in \mathfrak{M}(m,n+1)$
-	such that $M \subsetneq M'$.
-\end{definition}
-
-%\begin{definition}
-%	A set $M\in \mathfrak{M}(m,n)$ is \textit{submaximal},
-%	if there is only one maximal integral point set $M'\in \mathfrak{M}(m,n+1)$,
-%	such that $M \subset M'$.
-%\end{definition}
-
-In order to describe all the sets $M \in \mathfrak{M}(2,n)$ with distance 1,
-we need~\cite[Proposition 6]{our-vmmsh-2018-translit}.
-For the reader's convenience, we will state a slightly rephrased version of it and provide the proof.
-\begin{lemma}
-	\label{lem:no_4_points_in_semigeneral_position_with_distance_1}
-	Let $M=\{M_1,M_2,M_3,M_4\}\in \mathfrak{M}(2,4)$ such that $|M_1 M_2|=1$.
-	If $l$ is the straight line through points $M_1$ and $M_2$,
-	then one of the points $M_3$ or $M_4$ belongs to $l$.
-\end{lemma}
-
-\begin{proof}
-	Let $m$ denote the perpendicular bisector of line segment $M_1 M_2$.
-	Due to the triangle inequality, there are two possibilities to place each of points $M_i$, $i=3,4$:
-
-	a) $M_i$ belongs to $l$ and $|M_i M_1|-|M_i M_2| = \pm 1$;
-
-	b) $M_i$ belongs to $m$ and $|M_i M_1| = |M_i M_2|$.
-
-	If (a) is true for both points, then $M\subset l$ and thus $M$ is not an integral point set.
-	If (a) is true for one point and (b) is true for another, then the claim of the lemma follows.
-	So, suppose to the contrary, that both points $M_3$ and $M_4$ belong to $m$.
-
-	The area of the triangle $M_1 M_3 M_4$ is rational because $|M_3 - M_4| \in \mathbb{Z}$.
-	Thus, the characteristic of $M$ is 1,
-	and there is a Cartesian coordinate system such that $M_1=(-1/2,0)$, $M_2=(1/2,0)$, $M_3=(0, a/2)$
-	(see~\cite[Theorem 4]{our-vmmsh-2018-translit}).
-	It is clear that $a\neq 0$.
-	We set $b = |M_1 - M_3|$ and $O=(0,0)$.
-	Applying the Pythagorean theorem to triangle $OM_1M_3$, we obtain the following Diophantine equation:
-	\begin{equation}
-		\frac{1}{4} + \frac{a^2}{4} = b^2
-		,
-	\end{equation}
-	or, equivalently,
-	\begin{equation}
-		1 + a^2 = (2b)^2
-		.
-	\end{equation}
-	This equation has no integral solutions.
-	This contradiction concludes the proof.
-\end{proof}
-
-Using Construction~\ref{con:planar_set_with_minimeter_1},
-Lemma~\ref{lem:no_4_points_in_semigeneral_position_with_distance_1} and Corollary~\ref{cor:only_one_distance_1_on_straight_line}
-together with results of~\cite[Section 6]{antonov2008maximal},
-we obtain the following theorem.
-
-\begin{theorem}
-	\label{thm:minimeter_1_planar}
-	For every $n\geq 3$ there is a planar integral set $M$ of $n$ points
-	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds.
-	This set consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.
-
-	And vice versa, if $M$ is a planar integral point set of $n$ points
-	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
-	then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
-	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
-	There is only one maximal integral point set $M' \supseteq M$,
-	and the perpendicular bisector is the axis of symmetry for $M'$.
-	Moreover, if $n > 3$, then the distance 1 occurs in $M$ (and $M'$) only once.
-\end{theorem}
-
-\subsection{Higher dimensions}
-
-To build integral points sets with distance 1 in higher dimensions,
-we need the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds}.
-
-\begin{construction}
-	\label{con:blowing_up}
-	Let $M\in\mathfrak{M}(2,n)$ have many points on a straight line $l_1$
-	and $M\setminus l_1 \subset l_2$ for a straight line $l_2$.
-	Moreover, let $l_1$ be parallel to $l_2$.
-
-	Construct $M' \subset \mathbb{R}^m$ by replacing all the points $M_1,...,M_n\in M \cap l_2$
-	by equal $(m-2)$--dimensional simplices $S_1,...,S_n$
-	in such a way that
-	\begin{equation}
-		\forall(j=1,...,n) \forall(N_i \in M \cap l_1 )\forall\left(M^j_k, M^j_h \in S_j\right)\left[|N_i M^j_k| = |N_i M^j_h|\right].
-	\end{equation}
-	Then under certain conditions (in particular, if $\#(M \cap l_2)=1$ that we actually need)
-	we get $M'\in\mathfrak{M}(m,\mathbb{N})$.
-
-	For further examples we also refer the reader to~\cite{avdeev2019particular}.
-\end{construction}
-
-Now we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,
-with distance 1 occurring in it.
-If we have an integral point set with distance 1 occurring,
-then we can easily dilate it to turn 1 into the desired distance.
-These facts, together with Theorem~\ref{thm:minimeter_1_planar}, give the following theorem.
-
-\begin{theorem}
-	For arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
-	there exists $M\in\mathfrak{M}(m,n)$,
-	such that for some $M_1, M_2\in M$ equality $|M_1 M_2| = d$ holds.
-\end{theorem}
-
-\begin{definition}
-	\cite{noll1989nclusters}
-	We will call an integral point set $M$ \textit{prime}, if the greatest common divisor
-	of all the distances occurring in $M$ is 1.
-\end{definition}
-
-If an integral point set is prime,
-then it cannot be squashed to an integral point set of the same power and structure but smaller diameter.
-
-\begin{theorem}
-	\label{thm:prime_with_desired_distance}
-	For every $m \geq 3$, $n \geq m+1$, $d \geq 1$ there exists a \textbf{prime}
-	integral point set $M\in\mathfrak{M}(m,n)$ which
-	contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
-\end{theorem}
-
-\begin{proof}
-	Take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
-	with sufficiently large height and then apply the ``blowing up'' procedure described in Construction~\ref{con:blowing_up},
-	replacing the point that is not situated on the line, with an $(m-2)$-dimensional simplex of proper side length.
-	That simplex will consist of $m-1$ points;
-	we can exclude practically all $M_J$ points, except the two points with distance 1.
-	So, the set consists of $m+1$ or more points and has distance 1 occurring in it.
-	As distance 1 occurs in the obtained set, it is prime.
-\end{proof}
-
-A slightly stronger result can also be claimed.
-
-\begin{theorem}
-	\label{thm:prime_with_unique_desired_distance}
-	For arbitrary integers $m > 2$, $n \geq m+1$, $d \geq 1$
-	there exists a prime set $M\in\mathfrak{M}(m,n)$,
-	such that for some $M_1, M_2\in M$ equality $|M_1 M_2| = d$ holds,
-	the distance $d$ is minimal and occurs in $M$ only once.
-\end{theorem}
-
-\begin{proof}
-	We apply Construction~\ref{con:planar_set_with_minimeter_1} to obtain $M'\in \mathfrak{M}(2,n-m+2)$
-	with $M'_1, M'_2 \in M'$ and $|M'_1 M'_2| = 1$.
-	Dilate $M'$ to $M''$ in such a way that $M'_1$ and $M'_2$ turn into $M''_1$ and $M''_2$ resp.
-	and $|M''_1 M''_2| = d$.
-	Then ``blow up'' $M''$ to $M\in \mathfrak{M}(m,n)$ using a simplex of side length $d+1$.
-	The greatest common divisor of $d$ and $d+1$ is $1$, so $M$ is prime.
-\end{proof}
-
-\section{Final remarks and open problems}
-
-\begin{remark}
-	The bound of Theorem~\ref{thm:main_estimate} is not tight.
-	The derivation of tight bounds for the minimum diameter $d(2, n)$
-	is still a challenging task for the forthcoming investigation~\cite[Section 7]{kurz2008minimum}.
-\end{remark}
-The known upper bound is $d(2,n)\leq 2^{c \log n \log \log n}$~\cite{harborth1993upper}.
-However, we hope that the present article can provide a framework for possible better estimates.
-
-\begin{conjecture}
-	The approach of Theorem~\ref{thm:main_estimate} can be generalized to higher dimensions.
-\end{conjecture}
-For the overview of current lower bounds for higher dimensions, we refer the reader to~\cite{nozaki2013lower}.
-
-There are also several ways to improve the current bound,
-based on Lemma~\ref{lem:square_container}.
-
-\begin{problem}
-	\label{prb:square_side_length_and_diameter_power}
-	Is there a number $\delta < 1$ such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	is situated in a square of side length $d^\delta$?
-\end{problem}
-If the answer for Problem~\ref{prb:square_side_length_and_diameter_power} is affirmative,
-then we can obtain a bound better than linear.
-
-\begin{problem}
-	\label{prb:square_side_length_and_diameter_linear}
-	What are the minimal numbers $\delta \leq 1$ and $h \leq 0$
-	such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	is situated in a square of side length $\delta d + h$?
-\end{problem}
-
-If the answer for Problem~\ref{prb:square_side_length_and_diameter_linear} is not $\delta = 1$ and $h = 0$,
-then the linear bound can be slightly improved.
-
-However, Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}
-do not seem to show the way for improving the bound
-due to the following well-known theorem,
-which can be proved by an application of circle inversion~\cite{solymosi2010question}.
-\begin{theorem}
-	\label{thm:rational_set_on_circle}
-	There exists a dense subset $P$ of the unit circle such that for any $P_1, P_2 \in P$
-	the distance $|P_1 P_2|$ is rational.
-\end{theorem}
-We conjecture that thorough investigation of existing examples of circular integral point sets
-~\cite{bat2018number,harborth1993upper,piepmeyer1996maximum}
-will lead to the negative answers for the
-Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}.
-
-Another possibility to increase the constant in our lower bound is introduced by the following problem.
-
-\begin{problem}
-	\label{prb:minimal_shape}
-	What is the minimal shape $\mathcal{S}$ such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	is situated in  $\mathcal{S}$?
-\end{problem}
-This problem is another generalization of Lemma~\ref{lem:square_container}.
-We know that $\mathcal{S}$ is not a circle of diameter $d$
-(the counterexample is an equilateral triangle of side length $d$)
-and we conjecture that $\mathcal{S}$ is not a Reuleaux triangle of width $d$
-(the possible counterexample belongs to a circle of diameter $d$, see Theorem~\ref{thm:rational_set_on_circle}).
-
-The solution to Problem~\ref{prb:minimal_shape} (if ever found) will lead to a sophisticated packing problem,
-something like ``How to put $n$ points in a convex hull of concentric circle and Reuleaux triangle?''
-Obviously, Problem~\ref{prb:minimal_shape} can be generalized to higher dimensions.
-
-For $n=3,4,5$, there are optimal sets in $\mathfrak{M}(2,n)$ that contain distance 1
-(see Remarks~\ref{rem:planar_set_with_minimeter_1_of_3_points} and~\ref{rem:planar_set_with_minimeter_1_of_4_and_5_points}).
-However, we can suggest the following conjecture.
-\begin{conjecture}
-	\label{con:no_optimal_with_edge_1}
-	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,
-	is not optimal.
-\end{conjecture}
-
-The motivation for Conjecture~\ref{con:no_optimal_with_edge_1} is based on the results of~\cite[Section 5]{kurz2008minimum},
-especially the following theorem.
-
-\begin{theorem}
-	For every $12 \leq n \leq 122$, there exist a facher optimal sets $M_n\in\mathfrak{M}(2,n)$
-	that consist of $n-1$ points $\{ (b_1,0), ..., (b_{n-1},0)\}$
-	and the point $(0,\sqrt{A_n})$,
-	where $b_1,...,b_{n-1}, A_n$ are integer and $\sqrt{A_n}\notin\mathbb{N}$.
-\end{theorem}
-Such a set cannot contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar},
-the points with distance 1 should be $(\pm 1/2, 0)$.
-
-For planar integral point sets containing distance 1, we can emphasize the following problem.
-\begin{problem}
-	For given $n\geq 3$ and squarefree number $q \neq 1$,
-	is there $M\in\mathfrak{M}(2,n)$, such that characteristic of $M$ is $q$
-	and distance 1 occurs in $M$?
-\end{problem}
-
-The consideration of integral point sets containing distance 1 in higher dimensions gives a rise to another problem.
-
-\begin{problem}
-	How many times can the distance 1 occur in a set $M\in\mathfrak{M}(m,n)$?
-\end{problem}
-The answer for $m=2$ is given by Theorem~\ref{thm:minimeter_1_planar}.
-For greater $m$, we will be bold enough to suggest the following conjecture.
-\begin{conjecture}
-	Consider $M\in\mathfrak{M}(m,n)$, $M_1,M_2\in M$, $|M_1 M_2| =1$.
-	Then either $n = m+1$ or $M$ is a ``blowup'' of a facher set and distance 1 occurs in $M$
-	no more than $1+\frac{(m-1)^2 - (m-1)}{2}$ times.
-\end{conjecture}
-Such an estimate is based on the number of edges in an $(m-2)$-dimensional simplex.
-
-We want to conclude the list of conjectures with the one which appears the easiest to deal with.
-
-\begin{conjecture}
-	\label{hyp:prime_planar}
-	For every $n \geq 3$, $d \geq 1$ there exists a \textbf{prime}
-	set $M\in\mathfrak{M}(2,n)$ which
-	contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
-\end{conjecture}
-This conjecture gives the same claim as Theorem~\ref{thm:prime_with_desired_distance} does, but for $m=2$.
-Obviously, Theorem~\ref{thm:prime_with_unique_desired_distance} can be generalized (for $m=2$) into a similar conjecture.
 

### 2024-07-09 char-4k-bound/char-4k-bound.tex
[4k3] WIP, rather ready

@@
 \usepackage[dvips]{graphicx} %хотим вставлять рисунки?
-\graphicspath{{illustr/}}%путь к рисункам
+\graphicspath{{img/}}%путь к рисункам
 
@@
 
+\DeclareMathOperator{\diam}{diam}
+\DeclareMathOperator{\chr}{char}
+
 \begin{document}
@@
 \title{
-	On existence of integral point sets and their diameter bounds
+	On the characteristic and diameter of planar integral point sets
 	\footnote{
-		This work was carried out at Voronezh State University and supported by the Russian Science
-		Foundation grant 19-11-00197.
+		The work of the first author was supported by the Theoretical Physics
+		and Mathematics Advancement Foundation “BASIS”.
 	}
@@
 	N.N. Avdeev
-	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
+	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru},
+	E.A. Lushina
+	\footnote{TODO}
 	\\
@@
 \paragraph{Abstract.}
-A point set $M$ in $m$-dimensional Euclidean space is called an integral point set if all the distances between the
-elements of $M$ are integers, and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
-We improve the linear lower bound for diameter of planar integral point sets.
-This improvement takes into account some results related to the Point Packing in a Square problem.
-Then for arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
-we give a construction of an integral point set $M$ of $n$ points in $m$-dimensional Euclidean space,
-where $M$ contains points $M_1$ and $M_2$ such that the distance between $M_1$ and $M_2$ is exactly $d$.
-
-TODO
+A point set $M$ in Euclidean plane is called an integral point set in semi-general position if all the distances between the
+elements of $M$ are integers, and $M$ does not contain collinear triples.
+We improve the linear lower bound for diameter of such sets in the particular case
+when the characteristic of the set is of the form $4k+1$ or $4k+2$.
+To achieve that, we combine hyperbolae-based and grid-based toolsets.
 
 \section{Introduction}
-Let $\mathbb{N}$ be the set of all positive integers and let $|M_1 M_2|$ denote the Euclidean distance
-between points $M_1$ and $M_2$ in a finite-dimensional space $\mathbb{R}^m$
-(and, more generally, let $|\Delta|$ denote the length of line segment $\Delta$).
-An \textit{integral point set} in $m$-dimensional Euclidean space, $m>2$, is a point set $M$ such that all the distances between the
-points of $M$ are integers and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
-Erd{\H{o}}s and Anning proved~\cite{anning1945integral,erdos1945integral} that every integral point set consists of a finite number of points.
-Taking this into account, we denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space by
-$\mathfrak{M}(m,n)$ (using the notation in~\cite{our-vmmsh-2018-translit})
-and denote the set of all integral point sets in $m$-dimensional Euclidean space by $\mathfrak{M}(m,\mathbb{N})$.
-The symbol $\# M$ will be used for the cardinality of $M$, that is the number of points in $M$ in our case.
+	A planar integral point set (IPS) is a set of points in the plane, such that the distance between any pair of its points is an integer, and at least one triple of its points is non-collinear.
+The latter condition is essential to avoid subsets of a straight line; those are de-facto equivalent to subsets of integer numbers and form a completely different combinatorial object.
 
-For every finite point set, its diameter is naturally defined as
+	In 1945, Erdös gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
+	In any IPS he chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolae, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
+Applying the same argument to the line segment $M_1M_3$, Erdös concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.
+
+	Thus, we can easily infer the lower bound for diameter of an IPS $M$:
+	$$
+		\operatorname{diam} M \ge \frac{\sqrt{\#M}}{2}.
+	$$
+	That was the first lower bound for the diameter ever;
+	however, Erdös did not only estimate the diameter,
+	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolae.
+
+	Using this toolset, in 2003 Solymosi proved~\cite{solymosi2003note} that
+	$$
+		\operatorname{diam} M \ge c \cdot \#M.
+	$$
+	Although Solymosi did not give the constant $c$ explicitly,
+	it can be inferred from his proof that $c=\frac{1}{24}$ [TODO].
+
+	In~\cite{our-vmmsh-2018-translit} the constant (for $n\geq 4$) was improved to $0.3457$
+	employing Point Packing in a Square Problem~\cite{markot2005newverified,costa2013valid}.
+	In~\cite{my-pps-linear-bound-2019} the approach has been further developed, and the constant has been tightened to $\frac{5}{11}$.
+	Finally, in~\cite{my-semi-general-5-4-bound-2019} for IPS $M$ in semi-general position (that is an IPS with no collinear triples) it was proved that
 	\begin{equation}
-	\operatorname{diam} M = \max_{A,B\in M} |AB|
+		\label{eq:chebsb_lower_bound}
+		\diam M \geq \left(\frac n 5 \right)^{5/4}
+	\end{equation}
+	(in assumption that the set has at east 4 points).
+	All these lower bounds are based on the Erdos's framework --- cofocal hyperbolae.
+
+	Meanwhile in 1988 Kemnitz introduced~\cite{kemnitz1988punktmengen}
+	a characteristic of an IPS, that is a squarefree integer $q$ such as the area
+	of any triangle formed by any triple of points from the IPS is comparable with $\sqrt{q}$.
+	(Indeed, Kemnitz proved this fact for any point set with rational distances ---
+	in contrast with IPS, those can be infinite even they contain non-collinear triples,
+	see~\cite{huff1948diophantine} for an example of construction
+	and~\cite{solymosi2010question} for some known limitations.)
+	In 2000s, Kurz introduced~\cite{kurz2005characteristic} the function $d(2,n)$ that evaluates
+	to the minimal possible diameter of planar IPS of cardinality $n$.
+	Then Kurz employed Kemnitz's results and found~\cite[Subsection 4.2]{kurz2008minimum} exact values of $d(2,n)$
+	up to $n=122$ by exhaustive computer search.
+	(Taking the characteristic into consideration allows to boost such search significantly ---
+	basically because all triples of an IPS form triangles with equal characteristics.)
+
+	For generalization in higher dimensions, we refer the reader to~\cite{nozaki2013lower}.
+
+	In the present paper, we put the power of Erdös and Kemnitz approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
+	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
+	the bound~\eqref{eq:chebsb_lower_bound} can be improved to
+	\begin{equation}
+		\operatorname{diam} M \ge \left( \frac{25}{36} n \right)^{5/4}
 		.
 	\end{equation}
-Another emerging question is: how does the diameter of an integral point set depend on its cardinality?
-One can easily see that every $M\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = h$
-can be dilated to $M_p\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = ph$
-for every $p\in\mathbb{N}$.
-So, the above question should be rephrased:
-how does \textit{the least possible} diameter of an integral point set depend on its cardinality?
-In order to answer this question, the following function was introduced~\cite{kurz2008bounds,kurz2008minimum}:
-\begin{equation}
-	d(m,n) = \min_{M\in\mathfrak{M}(m,n)} \operatorname{diam} M = \min_{M\in\mathfrak{M}(m,n)} \max_{A,B\in M} |AB|
-	.
-\end{equation}
-We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$ and its bounds.
-In the present paper, we mostly focus on the case $m=2$.
 
-The most significant breakthrough on the planar case was done by Solymosi~\cite{solymosi2003note},
-who proved that $cn \leq d(2,n)$ for a sufficiently small constant $c$.
-Following Solymosi's proof carefully,
-one can derive that the inequality holds at least for $c = 1/24$.
-(See~\cite[Exercise 2.6]{garibaldi2005erdos} for some remarks.)
-The constant was improved in~\cite{our-mz-rus-translit} to $1/8$ for all $n$ and in~\cite{our-vmmsh-2018-translit}
-to $3/8$ for sufficiently large $n$.
-
-The paper~\cite{solymosi2003note} contains one more interesting result.
-Let us define a function which is ``dual'' to $d(m,n)$ in some sense:
-\begin{equation}
-	l(m,n) = \min_{M\in\mathfrak{M}(m,n)} \min_{A,B\in M} |AB|
-	.
-\end{equation}
-Solymosi proved that $l(2,n)\leq 2$.
-
-In the present paper we improve Solymosi's results:
-first, we obtain a larger constant $c = 5/11$ in Theorem~\ref{thm:main_estimate},
-using the combined approach with the Point Packing in a Square problem
-(this approach is different from Solymosi's one);
-second, we prove that $l(m,n)=1$ for all possible $m$ and $n$.
-
-Sections 2--4 are devoted to the linear lower bound.
-Each section consists of 2 subsections: General results and Special results.
-In General results, we do not aim to provide tight estimates;
-they are discussed in Special results.
-
-Section 5 is devoted to integral point sets with distance 1 and has its own structure.
+	In Section 2, we give all the required notions and known results.
+In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological.
+Section 4 is devoted to the connection between Erdös curves and characteristic.
+In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.
 
@@
 
-In this Section, we provide proper definitions and list some known results.
+In this Section, we provide rigorous definitions and list some known results.
 
-For the sake of brevity, the following designations will be used for numeric sets
+For the sake of brevity, the following notations will be used for sets
 of positive integers, non-negative integers and all the integers resp.:
@@
 
+For a finite set $M$, we will denote its cardinality by $\#M$.
+
+\begin{definition}
+	\label{def:IPS}
+	A planar integral point set (IPS) is a set $M$
+	of non-collinear points in the plane $\mathbb{R}^{2}$ such that
+	for any pair of points $M_{1}, M_{2} \in \mathcal{P}$
+	the Euclidean distance $|M_{1}M_{2}|$
+	between points $M_{1}$ and $M_{2}$ is integral.
+	Notation: $M\in\mathfrak{M}$, and also $M\in\mathfrak{M}_n$ for $n=\#M$.
+\end{definition}
+
+When we say that a set in non-collinear,
+we mean that it has at least one non-collinear triple.
+If we tighten the condition and require all the triples to be non-collinear,
+we get the next definition.
+
+\begin{definition}
+	\label{def:IPS_semi_general}
+	A planar IPS $M$ is said to be in \emph{semi-general position}
+	if no three points of $M$ are collinear.
+	Notation: $M\in\overline{\mathfrak{M}}$, and also $M\in\overline{\mathfrak{M}}_n$ for $n=\#M$.
+\end{definition}
+
+In the present paper, we mostly focus on IPS in semi-general position.
+However, the next restriction step can be done as the following.
+
+\begin{definition}
+	\label{def:IPS_general}
+	A planar IPS $M$ is said to be in \emph{general position}
+	if no three points of $M$ are collinear
+	and no four points of $M$ are concircular.
+	Notation: $M\in\overline{\mathfrak{M}}$, and also $M\in\overline{\mathfrak{M}}_n$ for $n=\#M$.
+\end{definition}
+
+\begin{definition}
+	The diameter of an integral point set $M$ is defined by setting
+	\begin{equation}
+		\operatorname{diam M} = \underset{M_{1}, M_{2} \in
+		M}{\max} |M_{1}M_{2}|
+		.
+	\end{equation}
+\end{definition}
+
+\begin{definition}
+	Two numbers $a$ and $b$ are \emph{commensurable} if their ratio $a/b$
+	is a rational number.
+\end{definition}
+
+For example, the pair $(7, 2/3)$ is commensurable,
+the pair $(\sqrt3/3, \sqrt{12})$ is also commensurable,
+but the pair $(\sqrt2, \sqrt3)$ is not.
+
+\begin{definition}
+	A number is called \emph{squarefree} if its only perfect square divisor is 1.
+	The first squarefree numbers are: 1, 2, 3, 5, 6, 7, 10, 11, 13,...
+\end{definition}
+
+\begin{definition}
+	The characteristic of a planar IPS $\mathcal{M}$ is a squarefree number $q$
+	such that the area of any triangle $M_1M_2M_3$, $\{M_1,M_2,M_3\}\subset M$,
+	is commensurable with $\sqrt{q}$.
+	Notation: $\chr M = q$.
+\end{definition}
+
+\begin{definition}
+	If points $M_1, M_2 \in M \in \mathcal{M}$,
+	then the line segment $M_1 M_2$ is said to be an \emph{edge} of $M$.
+\end{definition}
+
+The following result is due to Kemnitz:
+
+\begin{theorem}[the Grid Theorem]
+	A set $M \in \mathfrak{M}_n$ with characteristic $p$ can be placed on the grid
+	$$
+	\left\{\left(\frac{a_i}{2m} ; \frac{b_i \sqrt{p}}{2m}\right)\right\},
+	$$
+	where $a_i, b_i \in \mathbb{N}_\pm$, and $m$ can be taken as the length of any edge of the set $M$.
+	\label{tm:1}
+\end{theorem}
+
+For the sake of completeness, we provide the proof for the Grid Theorem below.
+
+\begin{proof}
+	Let $M_1, M_2 \in M \in \mathfrak M$ and $|M_1 M_2| = m$.
+	Set $M_1=(-m/2, 0)$, $M_2=(m/2, 0)$.
+	Then for any $M_i \in M$, $M_i = (x,y)$ one has
+	$|M_i M_1| = k \in \mathbb N_0$, $|M_i M_2| = n \in \mathbb N_0$.
+	The point $M_i$ belongs to the intersection of two circles,
+	whose equations are
+	\begin{gather}
+		\left(-\frac{m}{2} - x\right)^2 + y ^2 = k^2,
+	\\
+		\left( \frac{m}{2} - x\right)^2 + y ^2 = n^2,
+	\end{gather}
+	where $a+b\geq m$.
+
+	The solution is
+	\begin{gather}
+		x = \frac{k^2 - n^2}{2 m} = \frac{a_i}{2m},
+	\\
+	%	y = \frac{\pm \sqrt{-(a^2 - b^2)^2 + m^2 (2 a^2 + 2 b^2  - m^2)} }{2m}.
+		y = \pm\sqrt{k^2 - \left(\frac{m}{2}+x\right)^2} =
+		\pm\sqrt{k^2 - \left(\frac{m}{2}+\frac{a_i}{2m}\right)^2} =
+		\frac{\pm b_i \sqrt{q}}{2m},
+	\end{gather}
+	and the claim follows.
+\end{proof}
+
+\begin{definition}
+	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
+	For $N\in\mathbb N_\pm$, $|n| < |M_1 M_2|$
+	we will say that the set of points
+	\begin{equation}
+		\{M_0 : |M_0 M_1| - |M_0 M_2| = n\}
+	\end{equation}
+	is called the \emph{$n$-th Erdos curve}.
+\end{definition}
+
+Obviously, the 0-th Erdos curve is the perpendicular bisector of $M_1 M_2$,
+and all the other Erdos curves are branches of cofocal hyperbolae.
+
+On Figure~\ref{fig:Erdos_curves_numbered}, the Erdos curves are shown for an edge of length 3.
+
+\begin{figure}
+\center{\includegraphics[width=0.6\linewidth]{Erdos_curves_numbered.png}}
+\caption{Erdos curves}
+\label{fig:Erdos_curves_numbered}
+\end{figure}
+
+Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erdos curves.
+For the sake of brevity, Erdos curves with odd numbers are named odd Erdos curves,
+and the ones with even numbers are named even Erdos curves.
+
+Two following definitions are used to classify integral point sets
+with many collinear triples~\cite{avdeev2019particular}:
+
+\begin{definition}
+	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
+	a \emph{facher} set.
+\end{definition}
+For $9\leq n\leq 122$, the minimal possible diameter is attained at a facher set~\cite{kurz2008bounds}.
+
+\begin{definition}
+	A planar integral point set situated on two parallel straight lines
+	is called a \emph{rails} set.
+\end{definition}
+
+\begin{definition}
+	The part of a plane between two parallel straight lines with distance $\rho$ between them
+	is called a strip of width $\rho$.
+\end{definition}
+
+\begin{lemma}
+	\cite{smurov1998stripcoverings}
+	\label{lem:smurov_min_height}
+	If a triangle $T$ with minimal height $\rho$ is situated in a strip,
+	then the width of the strip is at least $\rho$.
+\end{lemma}
+
+\begin{lemma}
+	\cite[Lemma 4]{our-vmmsh-2018};
+	\cite[Lemma 2.4]{my-pps-linear-bound-2019}
+	\label{lem:square_container}
+	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
+	Then $M$ is situated in a square of side length $d$.
+\end{lemma}
+
+\begin{definition}
+	\cite[Definition 2.5]{my-pps-linear-bound-2019}
+	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
+	the line through $M_1$ and $M_2$,
+	and the perpendicular bisector of line segment $M_1 M_2$.
+\end{definition}
+
+\begin{lemma}
+	\cite[Theorem 3.10]{my-pps-linear-bound-2019}
+	\label{lem:no_distance_one}
+	Each set $M\in\mathfrak{M}_n$
+	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
+	consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
+	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
+\end{lemma}
+
+As far as our attention will be mostly restricted to planar integral point sets
+with characteristic of the form $4k+1$ and $4k+2$ in semi-general position.
+If an IPS $M$ satisfies these conditions,
+we will write that
+$M\in\overline{\mathfrak{M}}'$, and also $M\in\overline{\mathfrak{M}}'_n$ for $n=\#M$
+
+%3
+\section{Integral Point Sets with Various Characteristics}
+
+Let us demonstrate that the classes $4k+1$, $4k+2$ nor $4k+3$
+are neither too exotic nor pathological.
+In order to do this, we will provide important examples of IPS for each class.
+
+For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
+$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
+which means that each abscissa is multiplied by $1/q$
+and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
+\begin{equation}
+	\label{eq:char_lattice}
+	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
+	=
+	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}
+	,
+\end{equation}
+where $q$ is the characteristic of the IPS.
+Such notation is made possible by the Grid Theorem.
+
+When we think about IPS with characteristic $4k+1$,
+the very first example that comes to our mind is Egyptian triangle,
+that is the triangle with sides $(3;4;5)$.
+Egyptian triangle is obviously an IPS with characteristic $1 = 4 \cdot 0 +1$.
+
+Facher sets with characteristic 1, that are called \emph{semi-crabs},
+are investigated in~\cite{antonov2008maximal}.
+There are also much more complex IPS with characteristic 1;
+for those from $\dot{\mathfrak{M}}_7$, we refer the reader to~\cite{kurz2013constructing}.
+
+Figure~\ref{fig:char385} shows a rails IPS of characteristic $385 = 4 \cdot 96 + 1$ presented in~\cite{avdeev2019particular}.
+Its coordinates are
+\begin{multline}
+	\sqrt{385}/{2} * \{ (\pm 1105, 48),
+	(\pm 2189 ; 0),
+	(\pm 1587 ; 0),
+	(\pm 1269 ; 0),
+	\\
+	(\pm 763 ; 0),
+	(\pm 623 ; 0),
+	(\pm 529 ; 0),
+	(\pm 339 ; 0)\}
+\end{multline}
+
+\begin{figure}
+\center{\includegraphics[width=0.6\linewidth]{char385.png}}
+\caption{IPS of cardinality 16 and diameter 2189}
+\label{fig:char385}
+\end{figure}
+
+In~\cite{kreisel2008heptagon}, the first ever known
+planar integral point set $M_7 \in \dot{\mathfrak{M}}_7$ is given,
+see Fig.~\ref{fig:heptagon2}.
+Its coordinates are
+\begin{multline}
+\sqrt{2002}/2227 * \{ (0;0), (2227^2\cdot10;0), (26127018;932064), (32 142 553; 411 864),
+\\
+	 (17615968;238464), (7 344 908;411 864), (19079044; 54168)\}
+	 ,
+\end{multline}
+and
+$
+	\chr M_7 = 2002 = 4\cdot 500 + 2
+$.
+
+It is noticeable that the second example of an IPS from $\dot{\mathfrak{M}}_7$
+given in the same article has the same characteristic.
+
+Also the largest known rails IPS presented in~\cite{}[TODO:Omsk!]
+that has 104 points on one straight line and the rest 2 on another
+(that is the cardinality of 106) has the characteristic of $154 = 4 \cdot 38 + 2$.
+(We do not list the coordinates of that set here due to its diameter which is 2745754098774581800288844387372160.)
+
+\begin{figure}
+\center{\includegraphics[width=0.6\linewidth]{heptagon2.png}}
+\caption{The heptagon in general position}
+\label{fig:heptagon2}
+\end{figure}
+
+As for characteristic $4k+3$, we should mention that the \emph{upper} bound for the minimal diameter
+of planar integral point set given in~\cite{harborth1993upper} employs IPS with characteristic $3$.
+Moreover, all the IPS of minimal possible diameter provided in~\cite[\S 5, Fig. 1]{harborth1993upper}
+(for cardinalities from 3 to 9) have characteristic of form $4k+3$.
+
+For the sake of completeness, on Fig.~\ref{fig:11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}
+we give an example of rails set
+with 3 points on one line and 8 points on the other (first presented in~\cite{})
+whose characteristic is
+\begin{equation}
+	255255 = 3 \cdot 5 \cdot 7 \cdot 11 \cdot 13 \cdot 17
+	 = 4 \cdot 63813  + 3
+\end{equation}
+and whose coordinates are
+\begin{multline}
+\mathcal{P}_{3,8}=\sqrt{255255}/2*\{
+	( 1767 ; -3);
+	( 2791 ; -3);
+	( 4071 ; -3);
+\\
+	( -306 ; 0);
+	( 0 ; 0);
+	( 1798 ; 0);
+	( 2304 ; 0);
+	( 2760 ; 0);
+	( 3534 ; 0);
+	( 4040 ; 0);
+	( 4558 ; 0)
+\}
+\end{multline}
+
+\begin{figure}
+\center{\includegraphics[width=0.6\linewidth]{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}}
+\caption{The heptagon in general position}
+\label{fig:11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}
+\end{figure}
+
+%4
+\section{Erdös Curves and Characteristic}
+
+\begin{lemma}
+	\label{lem:pre_weeding_I}
+	Any set $M = {A, B, C} \in \mathfrak{M}_3$ with edge $|AB| = m$, where $|AC| - |BC| = m - s$, $m, s \in \mathbb{N}$, and $s$ is an odd number and $s \leq m$, has a characteristic of the form $p = 4k + 3$, $k \in \mathbb{N}_0$.
+\end{lemma}
+
+\begin{proof}
+	Let $|BC| = n$.
+Then, from the triangle inequality $|AC| < |AB| + |BC|$, we have that $|AC| = m + n - s$, $s \in \mathbb{N}$ and $s \leq m$.
+By the Grid Theorem, we introduce such coordinates that $A = (-\frac{m}{2}; 0)$, $B = (\frac{m}{2}; 0)$, and $C = (\frac{a}{2m}; \frac{b\sqrt{p}}{2m})$.
+We find the distance between points $A$ and $C$, and points $B$ and $C$ in coordinates, and form a system of equations
+	\begin{equation}
+		\left\{
+		\begin{aligned}
+		\frac{\sqrt{(a+m^2)^2+pb^2}}{2m}&=m+n-s.\\
+		\frac{\sqrt{(a-m^2)^2+pb^2}}{2m}&=n
+		\end{aligned}
+		\right.
+		\label{eq:29}
+	\end{equation}
+	Multiply each equation by $2m$ and square both sides of the equations in system (\ref{eq:29})
+	\begin{equation}
+	\left\{
+	\begin{aligned}
+	(a+m^2)^2+pb^2&=4m^2(m+n-s)^2,\\
+	(a-m^2)^2+pb^2&=4m^2n^2
+	\end{aligned}
+	\right.
+	\label{eq: 30}
+	\end{equation}
+	and then subtract the second equation from the first one:
+	\begin{equation}
+	(a+m^2)^2-(a-m^2)^2=4m^2(m+n-s)^2-4m^2n^2,
+	\label{eq: 31}
+	\end{equation}
+	$$
+	4m^2a=4m^2((m+n-s)^2-n^2),
+	$$
+	so
+	\begin{equation}
+	a=(m-s)(m+2n-s).
+	\label{eq: 32}
+	\end{equation}
+	Substitute the obtained expression (\ref{eq:32}) into the second equation of the system (\ref{eq:30}).
+	\begin{equation}
+		((m-s)(m+2n-s)-m^2)^2+pb^2=4m^2n^2,
+	\end{equation}
+	that immediately gives
+	\begin{equation}
+		\label{eq:33}
+		pb^2=s(2m+2n-s)(2n-s)(2m-s).
+	\end{equation}
+	According to the assumption, let $s=2t+1$, where $t\in\mathbb{N}_0$.
+	Then equation (\ref{eq:33}) can be rewritten as:
+	\begin{equation}
+		pb^2=(2t+1)(2m+2n-2t-1)(2n-2t-1)(2m-2t-1).
+		\label{eq:34.2}
+	\end{equation}
+	or
+	\begin{eqnarray}
+		pb^2+1 &=&4(-4t^4+8t^3m+8t^3n-8t^3-4t^2m^2-12t^2mn+12t^2m- {}\nonumber\\
+			 &     &-4t^2n^2+12t^2n-6t^2+4tm^2n-4tmn^2-12tmn+6tm- {}\nonumber\\
+			 &     &-4tn^2+6tn-2t+2m^2n-m^2+2mn^2-3mn+m-n^2+n).{}\nonumber\\
+	\label{eq: 35}
+	\end{eqnarray}
+	For equation (\ref{eq:35}) to have integer solutions, it is necessary for $p$ and $b$ to be odd.
+	The right-hand side is divisible by 4. Then $pb^2\equiv3\operatorname{mod}{4}$.
+	But $b^{2}\equiv1(\operatorname{mod}4)$.
+	Thus, $p=4k+3$, $k\in \mathbb{N}_{0}$.
+\end{proof}
+
+Then the following theorem immediately appears.
+
+\begin{theorem}[The Weeding Theorem I]
+	\label{thm:weeding_I}
+	Let $M = {M_1, M_2, M_3} \in \mathfrak{M}_3$ be an IPS of semi-general position with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
+Then, the length of the edge $M_1 M_2$ and its $n$-th Erdős curve, which contains the point $M_3$, have the same parity.
+\end{theorem}
+
+\begin{proof}
+	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and let the $n$-th Erdős curve satisfy $|M_1 M_3| - |M_2 M_3| = n$.
+	If $m$ is an odd number, then by Remark \ref{rm: 2}, the difference $n = |M_1 M_3| - |M_2 M_3| = m - s$, where $s$ is an even number, is also an odd number.
+	Let $m$ be an even number; then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is an even number, as the difference of two even numbers.
+\end{proof}
+
+An even edge is a bit special case.
+
+\begin{lemma}
+	\ref{lem:pre_weeding_II}
+	Every set $M = {A, B, C} \in \mathfrak{M}_3$ with an even edge $|AB| = 2q$, where $|AC| - |BC| = 2q - s$, $q, s \in \mathbb{N}$, $s$ is an odd number and $s \leq 2q$, has a characteristic of the form $p = 8k + 7$, $k \in \mathbb{N}_0$.
+\end{lemma}
+
+\begin{proof}
+	Let us set $m = 2q$ in the Lemma~\ref{lem:pre_weeding_I}.
+	Then equation \eqref{eq:33} turns into
+	\begin{equation}
+		\label{eq:33.1}
+		pb^2 = s(4q + 2n - s)(2n - s)(4q - s).
+	\end{equation}
+	Taking into account that $s = 2t + 1$, $t \in \mathbb{N}_0$,
+	we bring equation (\ref{eq:33.1}) to the form:
+	\begin{equation}
+	pb^2 = (2t + 1)(4q + 2n - 2t - 1)(2n - 2t - 1)(4q - 2t - 1).
+	\label{eq: 34.2}
+	\end{equation}
+	or, after expansion,
+	\begin{eqnarray}
+		pb^2 + 1&=& 8(-2t^4+8t^3q+4t^3n-4t^3-8t^2q^2-12t^2qn+12t^2q- {}\nonumber\\
+			 &     &-2t^2n^2+6t^2n-3t^2+8tq^2n-4tqn^2-12tqn+6tq- {}\nonumber\\
+			 &     &-2tn^2+3tn-t+4q^2n-2q^2+2qn^2-3qn+q)-  {}\nonumber\\
+			   &     &-4n(n-1).
+		\label{eq: 35.1}
+	\end{eqnarray}
+	Among two consecutive natural numbers $n$ and $(n-1)$,
+	exactly one of them is divisible by 2.
+	Therefore, the expression $4n(n-1)\equiv0\pmod{8}$.
+	Thus, the right-hand side of equation (\ref{eq:35.1}) is an even number and a multiple of 8.
+	Then $pb^2\equiv7\pmod{8}$.
+	A necessary condition for the existence of an integer solution to equation (\ref{eq:35.1})
+	is that numbers $p$ and $b$ must be odd.
+	But $b^2\equiv1\pmod{8}$, so $p=8k+7$, $k\in\mathbb{N}_0$.
+\end{proof}
+
+\begin{theorem}[The Weeding Theorem II]
+	\label{thm:weeding_II}
+	Let an IPS in semi-general position $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erdős curve of edge $M_1 M_2$.
+	Then the set $M$ has characteristic $p=8k+7$, $k\in \mathbb{N}_{0}$.
+\end{theorem}
+
+\begin{proof}
+	Indeed, let $|M_1 M_2|=2q$, $q \in \mathbb{N}$.
+	For the $n$-th Erdős curve we have $|M_1 M_3|-|M_2 M_3|=n$.
+	According to Lemma~\ref{lem:pre_weeding_II}, the difference $n=|M_1 M_3|-|M_2 M_3|=2q-s$,
+	where $s$ is an odd number,
+	is also an odd number and $\operatorname{char} M = 8k + 7$, $k\in \mathbb{N}_{0}$.
+\end{proof}
+
+%5
+\section{Auxiliary Results}
+
+\begin{lemma}
+	\label{lem:no_distance_2}
+	Every set $M \in \overline{\mathfrak{M}_n}'$ with an edge $|M_1M_2|=2$ has a cardinality $n=3$.
+\end{lemma}
+
+\begin{proof}
+	Let $M \in \overline{\mathfrak{M}_n}'$ and $M_1, M_2 \in M$.
+Then all points in the set $M$ lie on the circle $cr(M_1, M_2)$.
+Otherwise, according to Statement 4, $\operatorname{char}{M}={8k+7}$, $k\in \mathbb{N}_0$.
+
+	Assume the opposite.
+Let $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}_4}'$.
+Introduce Cartesian coordinates such that $M_1(-1;0), M_2(1;0), O(0;0)$.
+Then $M_3$ and $M_4$ lie on the median perpendicular to $M_1M_2$.
+Since the distance $|M_3M_4|$ is an integer, the area of triangle $M_1M_3M_4$ is a rational number.
+By Definition 6, this means that $\operatorname{char}{M}={1}$.
+Then by the theorem about the lattice we have $M_3(0;t/4)$.
+Without loss of generality, let us assume $t\in\mathbb{N}$.
+Let $|M_1M_3|=s$.
+Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
+	$$
+	1+\frac{t^2}{16}=s^2,
+	$$
+	or equivalently,
+	\begin{equation}
+	\label{eq: 42}
+	16s^2-t^2=16.
+	\end{equation}
+	Find the solutions to equation (\ref{eq: 42}) in natural numbers. To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$. Then equation (\ref{eq: 42}) takes the form:
+	\begin{equation}
+	\label{eq: 43}
+	s^2-k^2=1.
+	\end{equation}
+	Break down the left-hand side of equation (\ref{eq: 43}) into factors:
+	\begin{equation}
+	\label{eq: 44}
+	(s-k)(s+k)=1.
+	\end{equation}
+	Since $s,k\in\mathbb{N}$, the number 1 can be represented as the product of two integers in two ways: $1\cdot1$ and $-1\cdot(-1)$.
+In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$, which contradicts that $k\in\mathbb{N}$.
+	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1,k=0$, which contradicts that $s,k\in\mathbb{N}$.
+
+	Thus, equation (\ref{eq: 42}) has no natural solutions in natural numbers.
+	The obtained contradiction completes the proof of the lemma.
+\end{proof}
+
+The following result is a key detail of bounds presented
+in~\cite{solymosi2003note} and~\cite{my-semi-general-5-4-bound-2019}:
+
+\begin{lemma}
+	\cite[Observation 1]{solymosi2003note}
+	If a triangle $T$ has integer side lengths $a \leq b \leq c$,
+	then the minimal height $m$ of it is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
+\end{lemma}
+
+If we exclude characteristics of the form $4k+3$, this Solymosi's result can be sharpened a bit.
+
 \begin{lemma}
@@
 	Any triangle with sides $a \leq b \leq c$, where $c=a+b-1$ and $a, b, c \in \mathbb{N}$,
-	has a characteristic of the form $p=4k+3$, $k\in \mathbb{N}_0$.
+	has the characteristic of the form $p=4k+3$, $k\in \mathbb{N}_0$.
 \end{lemma}
@@
 \begin{proof}
-	Indeed, let the triangle $ABC$ satisfy: $|BC|=a$, $|AC|=b$.
-Then, from the triangle inequality $|AB| < |BC|+|AC|$, we can represent the length of side $AB$ as: $|AB|=a+b-s$, $s \in \mathbb{N}$ and $s \leq a$.
-For $s=1$, the conditions of statement 6 are satisfied: $s$ is an odd number and $s \leq a$.
-	Therefore, the triangle with sides $a$, $b$, and $c=a+b-1$ has a characteristic of the form $p=4k+3$, $k\in \mathbb{N}_0$.
+	Indeed, consider the triangle $ABC$ and let $|BC|=a$, $|AC|=b$.
+	From the triangle inequality $|AB| < |BC|+|AC|$,
+	we can represent the length of side $AB$ as $|AB|=a+b-s$, $s \in \mathbb{N}$ and $s \leq a$.
+	For $s=1$, the conditions of TODO are satisfied: $s$ is an odd number and $s \leq a$.
+	Therefore, the triangle with sides $a$, $b$, and $c=a+b-1$
+	has a characteristic of the form $p=4k+3$, $k\in \mathbb{N}_0$.
 \end{proof}
+
+\begin{lemma}
+	\label{lem:triangle_height_our}
+	Let a triangle $ABC$ have integer characteristic different from $4k+3$, $k\in \mathbb{N}_0$, with $a\leq b\leq c$. Then the smallest height of the triangle $ABC$ is at least $(2a-1)^{1/2}$.
+\end{lemma}
+
+\begin{proof}
+	By Corollary \ref{cr:3}, a triangle with side $c=a+b-1$ cannot have a characteristic different from $4k+3$, $k\in \mathbb{N}_0$. Therefore, in a triangle with integer sides, we have $a+b\geq c+2.$
+
+	The height $h$ of the triangle $ABC$, dropped onto side $c$, can be found from the formula for its area: $S=hc/2$, which gives $h=2S/c$.
+To find the area of the triangle, we use the Heron's formula in the following form:
+	$$
+	S=\frac{1}{4} \sqrt{4a^2c^2-(c^2+a^2-b^2)^2}.
+	$$
+	Then
+	\begin{multline}
+		\label{eq:39}
+		h^2=\left( \frac{2}{4c} \cdot \sqrt{4a^2c^2-(c^2+a^2-b^2)^2}\right)^2
+		=
+		\frac{1}{4c^2} \cdot \left( 4a^2c^2-(c^2+a^2-b^2)^2 \right)
+		\\
+		=
+		a^2- \left(\frac{c^2+a^2-b^2}{2c} \right)^2
+	\end{multline}
+
+	From three heights of the triangle, the smallest one is that dropped onto its largest side.
+	For fixed sides $a$ and $b$, let us set $c=a+b-2$. Then
+	\begin{multline}
+		\label{eq: 40}
+		\frac{c^2+a^2-b^2}{c}
+		=
+		c+\frac{a+b}{c}(a-b) = c+\frac{c+2}{c}(a-b)
+		=
+		\\
+		c+\left(1+\frac{2}{c}\right)(a-b) \leq c+a-b=2a-1
+	\end{multline}
+	and this equality is possible when $a=b$. We rewrite expression (\ref{eq: 39}) as follows:
+	$$
+	h^2 \geq a^2- \left(\frac{2a-1}{2} \right)^2=a^2-(a-1)^2=a^2-a^2+2a-1=2a-1.
+	$$
+
+	Thus, the smallest height of the triangle with characteristic different from $4k+3$ is at least $(2a-1)^{1/2}$.
+\end{proof}
+
+\begin{corollary}
+	\label{cor:sqrt_5_3}
+	For the height of a triangle $ABC$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}_4}'$, with sides $3 \leq a \leq b \leq c$, the following estimate holds:
+	$$
+		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2},
+	$$
+	where the difference between the left and right-hand sides increases with increasing $a$.
+\end{corollary}
+
+%6
+\section{The Main Bound}
+
+The following Lemma is a classic Erdos-style intersection-enumeration one,
+empowered by our Weeding Theorem I.
+
+\begin{lemma}
+	\label{lem:intersection_enumeration}
+	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}_n}$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$.
+Then $\# M \le |M_1 M_2| \cdot |M_3 M_4| - 2$.
+\end{lemma}
+
+\begin{proof}
+	We will distinguish three cases: two even edges, two odd edges and two edges with different parity.
+
+	Let's consider the first case. Suppose both edges $M_1M_2$ and $M_3M_4$ have even lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:
+
+	a) $N$ belongs to $cr(M_1,M_2)$, which implies that there are no more than 4 points (no more than 2 on each of the lines);
+
+	b) $N$ belongs to $cr(M_3,M_4)$, which implies that there are no more than 4 points (no more than 2 on each of the lines);
+
+	c) $N$ belongs to the intersection of one of the $(|M_1M_2|/2 - 1)$ hyperbolae with one of the $(|M_3M_4|/2 - 1)$ hyperbolae, which implies that there are no more than $4(|M_1M_2|/2 - 1)(|M_3M_4|/2 - 1)$ points.
+
+	Assuming that the edges $M_1M_2$ and $M_3M_4$ have even lengths,
+	we infer that $|M_1M_2|\geq4$ and $|M_3M_4|\geq4$.
+	Then,
+	\begin{multline}
+		4 \left(\frac{|M_1 M_2|}{2} - 1\right) \left(\frac{|M_3 M_4|}{2} - 1 \right) + 4 + 4
+		=
+		\\=
+		 4\left(\frac{|M_1 M_2| \cdot |M_3 M_4|}{4} - \frac{|M_1 M_2|}{2} - \frac{|M_3 M_4|}{2}+1\right) + 4 + 4
+		=
+		\\=
+		|M_1 M_2| \cdot |M_3 M_4| - 2 |M_1 M_2| - 2 |M_3 M_4| + 4 + 4 + 4
+		\leq
+		|M_1 M_2| \cdot |M_3 M_4| - 4 \\
+		<
+		|M_1 M_2| \cdot |M_3 M_4| - 2
+		.
+	\end{multline}
+
+	Let's consider the second case. Suppose both edges $M_1M_2$ and $M_3M_4$ have odd lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:
+
+	a) $N$ belongs to $cr(M_1,M_2)$, which implies that there are no more than 2 points (otherwise, $\operatorname{char} M = 4k+3$);
+
+	b) $N$ belongs to $cr(M_3,M_4)$, which implies that there are no more than 2 points;
+
+	c) $N$ belongs to the intersection of one of $(|M_1 M_2|-1)/2$ hyperbolae with one of $(|M_3M_4|-1)/2$ hyperbolae, which implies that there are no more than $(|M_1 M_2|-1)(|M_3 M_4|-1)$ points.
+
+	Assuming that the edges $M_1M_2$ and $M_3M_4$ have odd lengths,
+	we infer that that $|M_1M_2|\geq3$ and $|M_3M_4|\geq3$.
+	Then,
+	\begin{multline}
+		(|M_1 M_2| - 1)(|M_3 M_4| - 1) + 2 + 2
+		=
+		|M_1 M_2| \cdot |M_3 M_4| - |M_1 M_2| - |M_3 M_4| + 4
+		\leq \\
+		\leq |M_1 M_2| \cdot |M_3 M_4| - 2
+		.
+	\end{multline}
+
+	Let's consider the third case.
+Suppose the lengths of edges $M_1M_2$ and $M_3M_4$ are different and without loss of generality suppose edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
+	Then, for each point $N\in M$, one of the following conditions is satisfied:
+
+	a) $N$ belongs to $cr(M_1, M_2)$, which implies that there are no more than 4 points (no more than 2 on each of the lines);
+
+	b) $N$ belongs to $cr(M_3,M_4)$, which implies that there are no more than 2 points (otherwise $\operatorname{char} M = 4k+3$);
+
+	c) $N$ belongs to the intersection of one of $(|{M_1 M_2}|/2-1)$ hyperbolae with one of $(|M_3 M_4|-1)/2$ hyperbolae, which implies that there are no more than $4(|M_1 M_2|/2-1)(|M_3 M_4|-1)/2$ points.
+
+	From the parity assumption for edges $M_1 M_2$ and $M_3 M_4$
+	we can infer that $|M_1 M_2|\geq4$ and $|M_3M_4|\geq3$.
+	Then,
+	\begin{multline}
+		4\left(\frac{|M_1 M_2|}{2} - 1\right) \left(\frac{|M_3 M_4| - 1}{2}\right) + 2 + 4
+		=
+		\\=
+		(|M_1 M_2| - 2)(|M_3 M_4| - 1) + 6
+		=
+		\\=
+		|M_1 M_2| \cdot |M_3 M_4| - |M_1 M_2| - 2 |M_3 M_4| + 2 + 6
+		\leq
+		|M_1 M_2| \cdot |M_3 M_4| - 2
+		.
+	\end{multline}
+
+	Thus, the lemma is proved.
+\end{proof}
+
+The following function was introduced in~\cite{kurz2008minimum}
+\begin{equation}
+	\overline{d}(2,n) = \min_{M\in\overline{\mathfrak{M}_{n}}} \operatorname{diam} M
+	,
+\end{equation}
+and some values were given:
+
+\begin{eqnarray}
+\label{d}
+\overline{d}(2,3)&=& 1,{}\nonumber\\
+\overline{d}(2,4)&=& 4,{}\nonumber\\
+\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{}\nonumber\\
+\overline{d}(2,7)&=& 33,{}\nonumber\\
+\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{}\nonumber\\
+\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{}\nonumber\\
+\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\nonumber\\
+\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{}\nonumber\\
+\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{}\nonumber\\
+\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{}\nonumber\\
+\overline{d}(2,28)&=& 10672,{}\nonumber\\
+\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{}\nonumber\\
+\overline{d}(2,37) &>& 20000.
+\end{eqnarray}
+
+Now we are finally ready to proof our main result and improve the bound~\eqref{eq:chebsb_lower_bound}
+in our special case.
+
+\begin{theorem}
+	\label{thm:main_result}
+	Let $M \in \overline{\mathfrak{M}}_n'$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
+	Then, for every integer $n \geq 3$, the inequality
+	\begin{equation}
+		\operatorname{diam} M \geq \left( \frac{25n}{36} \right)^{5/4}
+	\end{equation}
+	holds.
+\end{theorem}
+
+\begin{proof}
+	For $n = 3$, we have $\operatorname{diam} M \geq 3$ (achieved by an isosceles triangle with sides 2, 3, 3), and the theorem statement is obvious.
+
+	Let us consider $M \in \overline{\mathfrak{M}_n}'$, $n \geq 4$, $\operatorname{diam} M = p$.
+
+	Choose points $M_1, M_2, M_3, M_4 \in M$ (points $M_2$ and $M_3$ may coincide, while the others must be pairwise distinct) such that
+
+	\begin{equation*}
+		\min_{A,B \in M} |AB| = |M_1 M_2|
+	\end{equation*}
+
+	\begin{equation*}
+		\min_{A,B \in M \setminus {M_1}} |AB| = |M_3 M_4| = m
+	\end{equation*}
+
+	If $m \leq \frac{6}{5}p^{2/5}$, then by Lemma~\ref{lem:intersection_enumeration},
+	\begin{equation*}
+		n \leq |M_1 M_2| \cdot |M_3 M_4| - 2 \leq \frac{36}{25}p^{4/5} - 2
+	\end{equation*}
+	or equivalently
+	\begin{equation}
+		p \geq \left(\frac{25(n+2)}{36}\right)^{5/4} \geq \left(\frac{25n}{36}\right)^{5/4}
+	\end{equation}
+	which is exactly the claim of the theorem.
+
+	Now we have to consider $m > \frac{6}{5}p^{2/5}$.
+	Then for any points $A, B \in M \setminus {M_1}$, we have $|AB| > \frac{6}{5}p^{2/5}$.
+	By Corollary~\ref{cor:sqrt_5_3} and Lemma~\ref{lem:smurov_min_height},
+	no three points from $M \setminus {M_1}$ lie in a strip of width
+	\begin{equation}
+		\frac{\sqrt{5}}{\sqrt{3}} \cdot \sqrt{\frac{6p^{2/5}}{5}} = \sqrt{2} \cdot p^{1/5}
+		.
+	\end{equation}
+
+	By Lemma~\ref{lem:square_container}, the set $M$ lies in a square with side length $p$.
+We cover this square by $q$ strips, $\frac{p^{4/5}}{\sqrt{2}} \leq q < \frac{p^{4/5}}{\sqrt{2}} + 1$, such that the width of each strip does not exceed $\sqrt{2} \cdot p^{1/5}$.
+	Each of the obtained strips contains no more than two points from $M \setminus {M_1}$, so
+	\begin{equation}
+		\label{eq: 46}
+		n \leq 2 \left( \frac{p^{4/5}}{\sqrt{2}} + 1\right) + 1=\frac{2p^{4/5}}{\sqrt{2}} +3 = \sqrt{2}p^{4/5}+3
+		.
+	\end{equation}
+
+	From inequality (\ref{eq:46}), we obtain
+	\begin{equation}
+		\label{eq:strips_5_4}
+		p \geq \left( \frac{n -3}{\sqrt{2}} \right)^{5/4}
+		.
+	\end{equation}
+
+	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
+	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20,000$ and our estimate on the diameter is also true.
+	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10,655$.
+	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.
+
+	For estimation (\ref{eq:strips_5_4}) and $n \geq 7^4$, we have
+	\begin{equation}
+		\label{eq:strips_5_4_7}
+		p \geq \left( \frac{n -3}{\sqrt{2}} \right)^{5/4}  \geq \left( \frac{25n}{36} \right)^{5/4}
+		.
+	\end{equation}
+
+	Thus, for any $n \geq 3$ the inequality $
+		\operatorname{diam} M \geq \left( \frac{25n}{36} \right)^{5/4}
+	$ holds.
+\end{proof}
+
+\section{Conclusion}
+	The bound proved above (as well as the one (TODO)) may appear to be far from precise values of $\overline{d}(2,n)$.
+	However, it's easy to see that values of $d(2;n)$ tend to repeat often;
+	thus, it is rather not unrealistic that the bounds may converge to the precise values.
+
+	Also, we should notice that our approach and especially Weeding Theorems do not require semi-general position and can be applied to tighten the bound from~\cite{my-pps-linear-bound-2019}
+	in the special case of characteristic $4k+1$ or $4k+2$.
+	However, we are not ready to accept this tedious challenge yet.
+
+	Another research area that is able to utilize our results fruitfully is the maximization of IPS.
+	The Weeding Theorems I and II can sometimes make the exhaustive search up to 4 times faster.
 
@@
 The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
-A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
+Dr. A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
 Dr.~A.S.~Usachev for proofreading,
 and E.A. Momot for the ``SciLexic'' project that helped the author with some word usage issues.
-Also the author would like to owe a special thanks to the anonymous referees
-for their helpful comments and suggestions on the paper.
+%Also the author would like to owe a special thanks to the anonymous referees
+%for their helpful comments and suggestions on the paper.
 

### 2024-07-11 char-4k-bound/char-4k-bound.tex
[4k3] Various minor fixes

@@
 	E.A. Lushina
-	\footnote{TODO}
+	%\footnote{TODO}
 	\\
@@
 elements of $M$ are integers, and $M$ does not contain collinear triples.
-We improve the linear lower bound for diameter of such sets in the particular case
+We improve the lower bound for diameter of such sets in the particular case
 when the characteristic of the set is of the form $4k+1$ or $4k+2$.
@@
 	Although Solymosi did not give the constant $c$ explicitly,
-	it can be inferred from his proof that $c=\frac{1}{24}$ [TODO].
+	it can be inferred~\cite{our-vmmsh-2018-translit} from his proof that $c=\frac{1}{24}$.
 
@@
 	a characteristic of an IPS, that is a squarefree integer $q$ such as the area
-	of any triangle formed by any triple of points from the IPS is comparable with $\sqrt{q}$.
+	of a triangle formed by any triple of points from the IPS is comparable with $\sqrt{q}$.
 	(Indeed, Kemnitz proved this fact for any point set with rational distances ---
-	in contrast with IPS, those can be infinite even they contain non-collinear triples,
-	see~\cite{huff1948diophantine} for an example of construction
+	in contrast with IPS, those can be infinite even if they contain non-collinear triples,
+	see~\cite{huff1948diophantine} for an example construction
 	and~\cite{solymosi2010question} for some known limitations.)
@@
 	(Taking the characteristic into consideration allows to boost such search significantly ---
-	basically because all triples of an IPS form triangles with equal characteristics.)
+	basically because all triples of an IPS form triangles with equal characteristic.)
 
@@
 
-	In the present paper, we put the power of Erdös and Kemnitz approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
+	In the present paper, we put the power of Erdös's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
 	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
@@
 	of non-collinear points in the plane $\mathbb{R}^{2}$ such that
-	for any pair of points $M_{1}, M_{2} \in \mathcal{P}$
+	for any pair of points $M_{1}, M_{2} \in M$
 	the Euclidean distance $|M_{1}M_{2}|$
@@
 	if no three points of $M$ are collinear.
-	Notation: $M\in\overline{\mathfrak{M}}$, and also $M\in\overline{\mathfrak{M}}_n$ for $n=\#M$.
+	Notation: $M\in\dot{\mathfrak{M}}$, and also $M\in\dot{\mathfrak{M}}_n$ for $n=\#M$.
 \end{definition}
@@
 	\begin{equation}
-		\operatorname{diam M} = \underset{M_{1}, M_{2} \in
+		\operatorname{diam} M = \underset{M_{1}, M_{2} \in
 		M}{\max} |M_{1}M_{2}|
@@
 	A number is called \emph{squarefree} if its only perfect square divisor is 1.
-	The first squarefree numbers are: 1, 2, 3, 5, 6, 7, 10, 11, 13,...
+	The first squarefree numbers are: 1, 2, 3, 5, 6, 7, 10, 11, 13, ...
 \end{definition}
@@
 \begin{definition}
-	If points $M_1, M_2 \in M \in \mathcal{M}$,
+	If points $M_1, M_2 \in M \in \mathfrak{M}$,
 	then the line segment $M_1 M_2$ is said to be an \emph{edge} of $M$.
@@
 
-The following result is due to Kemnitz:
+The following result is due to Kemnitz~\cite{kemnitz1988punktmengen}:
 
@@
 
-For the sake of completeness, we provide the proof for the Grid Theorem below.
+For the sake of completeness, we sketch the proof for the Grid Theorem below.
 
@@
 	\end{gather}
-	where $a+b\geq m$.
+	where $k+n\geq m$.
 
@@
 and all the other Erdos curves are branches of cofocal hyperbolae.
-
 On Figure~\ref{fig:Erdos_curves_numbered}, the Erdos curves are shown for an edge of length 3.
@@
 \begin{figure}
-\center{\includegraphics[width=0.6\linewidth]{Erdos_curves_numbered.png}}
+\center{\includegraphics[width=0.95\linewidth]{Erdos_curves_numbered.png}}
 \caption{Erdos curves}
@@
 \begin{lemma}
-	\cite[Lemma 4]{our-vmmsh-2018};
+	\cite[Lemma 4]{our-vmmsh-2018-translit};
 	\cite[Lemma 2.4]{my-pps-linear-bound-2019}
@@
 
-As far as our attention will be mostly restricted to planar integral point sets
+Our attention will be mostly restricted to planar integral point sets
 with characteristic of the form $4k+1$ and $4k+2$ in semi-general position.
-If an IPS $M$ satisfies these conditions,
+Thus, if an IPS $M$ satisfies these conditions,
 we will write that
-$M\in\overline{\mathfrak{M}}'$, and also $M\in\overline{\mathfrak{M}}'_n$ for $n=\#M$
+$M\in\overline{\mathfrak{M}}'$, and also $M\in\overline{\mathfrak{M}}'_n$ for $n=\#M$.
 
@@
 \begin{figure}
-\center{\includegraphics[width=0.6\linewidth]{char385.png}}
+\center{\includegraphics[width=0.95\linewidth]{char385.png}}
 \caption{IPS of cardinality 16 and diameter 2189}
@@
 planar integral point set $M_7 \in \dot{\mathfrak{M}}_7$ is given,
-see Fig.~\ref{fig:heptagon2}.
+see Figure~\ref{fig:heptagon2}.
 Its coordinates are
@@
 
-Also the largest known rails IPS presented in~\cite{}[TODO:Omsk!]
-that has 104 points on one straight line and the rest 2 on another
-(that is the cardinality of 106) has the characteristic of $154 = 4 \cdot 38 + 2$.
+Also the largest known rails IPS presented in~\cite{momot2022example}
+with 104 points on one straight line and the rest 2 on another
+(that gives the cardinality of 106) has the characteristic of $154 = 4 \cdot 38 + 2$.
 (We do not list the coordinates of that set here due to its diameter which is 2745754098774581800288844387372160.)
@@
 \begin{figure}
-\center{\includegraphics[width=0.6\linewidth]{heptagon2.png}}
+\center{\includegraphics[width=0.95\linewidth]{heptagon2.png}}
 \caption{The heptagon in general position}
@@
 of planar integral point set given in~\cite{harborth1993upper} employs IPS with characteristic $3$.
-Moreover, all the IPS of minimal possible diameter provided in~\cite[\S 5, Fig. 1]{harborth1993upper}
+Moreover, all the IPS of minimal possible diameter provided in~\cite[\S 5, Figure 1]{harborth1993upper}
 (for cardinalities from 3 to 9) have characteristic of form $4k+3$.
 
-For the sake of completeness, on Fig.~\ref{fig:11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}
+For the sake of completeness, on Figure~\ref{fig:11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}
 we give an example of rails set
-with 3 points on one line and 8 points on the other (first presented in~\cite{})
+with 3 points on one line and 8 points on the other (first presented in~\cite[Figure 1]{avdeev2019particular})
 whose characteristic is
@@
 \begin{figure}
-\center{\includegraphics[width=0.6\linewidth]{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}}
-\caption{The heptagon in general position}
+\center{\includegraphics[width=0.95\linewidth]{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}}
+\caption{The rails set with characteristic 255255}
 \label{fig:11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}
@@
 	\label{lem:pre_weeding_I}
-	Any set $M = {A, B, C} \in \mathfrak{M}_3$ with edge $|AB| = m$, where $|AC| - |BC| = m - s$, $m, s \in \mathbb{N}$, and $s$ is an odd number and $s \leq m$, has a characteristic of the form $p = 4k + 3$, $k \in \mathbb{N}_0$.
+	Any set $M = \{A, B, C\} \in \mathfrak{M}_3$ with edge $|AB| = m$, where $|AC| - |BC| = m - s$, $m, s \in \mathbb{N}$, and $s$ is an odd number and $s \leq m$, has a characteristic of the form $p = 4k + 3$, $k \in \mathbb{N}_0$.
 \end{lemma}
@@
 	\begin{equation}
-		\left\{
-		\begin{aligned}
+		\label{eq:29}
+		\begin{cases}
 		\frac{\sqrt{(a+m^2)^2+pb^2}}{2m}&=m+n-s.\\
 		\frac{\sqrt{(a-m^2)^2+pb^2}}{2m}&=n
-		\end{aligned}
-		\right.
-		\label{eq:29}
+		\end{cases}
 	\end{equation}
@@
 	\begin{equation}
-	\left\{
-	\begin{aligned}
+		\label{eq:30_pre_weeding_I}
+		\begin{cases}
 		(a+m^2)^2+pb^2&=4m^2(m+n-s)^2,\\
 		(a-m^2)^2+pb^2&=4m^2n^2
-	\end{aligned}
-	\right.
-	\label{eq: 30}
+		\end{cases}
 	\end{equation}
@@
 	\begin{equation}
+		\label{eq:32_pre_weeding_I}
 		a=(m-s)(m+2n-s).
-	\label{eq: 32}
 	\end{equation}
-	Substitute the obtained expression (\ref{eq:32}) into the second equation of the system (\ref{eq:30}).
+	Substitute the obtained expression~\eqref{eq:32_pre_weeding_I}
+	into the second equation of the system~\eqref{eq:30_pre_weeding_I}:
 	\begin{equation}
@@
 	\begin{equation}
-		pb^2=(2t+1)(2m+2n-2t-1)(2n-2t-1)(2m-2t-1).
 		\label{eq:34.2}
+		pb^2=(2t+1)(2m+2n-2t-1)(2n-2t-1)(2m-2t-1)
 	\end{equation}
@@
 			 &     &-4t^2n^2+12t^2n-6t^2+4tm^2n-4tmn^2-12tmn+6tm- {}\nonumber\\
-			 &     &-4tn^2+6tn-2t+2m^2n-m^2+2mn^2-3mn+m-n^2+n).{}\nonumber\\
-	\label{eq: 35}
+			 &     &-4tn^2+6tn-2t+2m^2n-m^2+2mn^2-3mn+m-n^2+n).%{}\nonumber\\
+	\label{eq:35}
 	\end{eqnarray}
-	For equation (\ref{eq:35}) to have integer solutions, it is necessary for $p$ and $b$ to be odd.
+	If equation (\ref{eq:35}) has integer solutions, then both $p$ and $b$ are odd.
 	The right-hand side is divisible by 4. Then $pb^2\equiv3\operatorname{mod}{4}$.
@@
 	\label{thm:weeding_I}
-	Let $M = {M_1, M_2, M_3} \in \mathfrak{M}_3$ be an IPS of semi-general position with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
+	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$ be an IPS of semi-general position with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
 Then, the length of the edge $M_1 M_2$ and its $n$-th Erdős curve, which contains the point $M_3$, have the same parity.
@@
 	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and let the $n$-th Erdős curve satisfy $|M_1 M_3| - |M_2 M_3| = n$.
-	If $m$ is an odd number, then by Remark \ref{rm: 2}, the difference $n = |M_1 M_3| - |M_2 M_3| = m - s$, where $s$ is an even number, is also an odd number.
+	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3| = m - s$, where $s$ is an even number, is also an odd number.
 	Let $m$ be an even number; then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is an even number, as the difference of two even numbers.
@@
 \begin{lemma}
-	\ref{lem:pre_weeding_II}
-	Every set $M = {A, B, C} \in \mathfrak{M}_3$ with an even edge $|AB| = 2q$, where $|AC| - |BC| = 2q - s$, $q, s \in \mathbb{N}$, $s$ is an odd number and $s \leq 2q$, has a characteristic of the form $p = 8k + 7$, $k \in \mathbb{N}_0$.
+	\label{lem:pre_weeding_II}
+	Every set $M = \{A, B, C\} \in \mathfrak{M}_3$ with an even edge $|AB| = 2q$, where $|AC| - |BC| = 2q - s$, $q, s \in \mathbb{N}$, $s$ is an odd number and $s \leq 2q$, has a characteristic of the form $p = 8k + 7$, $k \in \mathbb{N}_0$.
 \end{lemma}
@@
 			   &     &-4n(n-1).
-		\label{eq: 35.1}
+		\label{eq:35.1}
 	\end{eqnarray}
@@
 	we can represent the length of side $AB$ as $|AB|=a+b-s$, $s \in \mathbb{N}$ and $s \leq a$.
-	For $s=1$, the conditions of TODO are satisfied: $s$ is an odd number and $s \leq a$.
+	For $s=1$, the conditions of Lemma~\ref{lem:pre_weeding_I} are satisfied: $s$ is an odd number and $s \leq a$.
 	Therefore, the triangle with sides $a$, $b$, and $c=a+b-1$
@@
 \section{Conclusion}
-	The bound proved above (as well as the one (TODO)) may appear to be far from precise values of $\overline{d}(2,n)$.
+	The bound proved above (as well as~\eqref{eq:chebsb_lower_bound}) may appear to be far from precise values of $\overline{d}(2,n)$.
 	However, it's easy to see that values of $d(2;n)$ tend to repeat often;

### 2024-07-11 char-4k-bound/char-4k-bound.tex
[4k3] Embrace some fixes from Dr. A.S. Usachev

@@
 	\footnote{
-		The work of the first author was supported by the Theoretical Physics
+		The work of the first author was partially supported by the Theoretical Physics
 		and Mathematics Advancement Foundation “BASIS”.
@@
 	\label{lem:pre_weeding_I}
-	Any set $M = \{A, B, C\} \in \mathfrak{M}_3$ with edge $|AB| = m$, where $|AC| - |BC| = m - s$, $m, s \in \mathbb{N}$, and $s$ is an odd number and $s \leq m$, has a characteristic of the form $p = 4k + 3$, $k \in \mathbb{N}_0$.
+	Any set $M = \{A, B, C\} \in \mathfrak{M}_3$ with edge $|AB| = m$, where $|AC| - |BC| = m - s$, $m, s \in \mathbb{N}$, $s$ is an odd number and $s \leq m$, has a characteristic of the form $p = 4k + 3$, $k \in \mathbb{N}_0$.
 \end{lemma}

### 2024-07-11 char-4k-bound/char-4k-bound.tex
[4k3] Embrace some more fixes from Dr. A.S. Usachev

@@
 \begin{definition}
+	\label{def:char}
 	The characteristic of a planar IPS $\mathcal{M}$ is a squarefree number $q$
@@
 \begin{proof}
-	Let $|BC| = n$.
-Then, from the triangle inequality $|AC| < |AB| + |BC|$, we have that $|AC| = m + n - s$, $s \in \mathbb{N}$ and $s \leq m$.
-By the Grid Theorem, we introduce such coordinates that $A = (-\frac{m}{2}; 0)$, $B = (\frac{m}{2}; 0)$, and $C = (\frac{a}{2m}; \frac{b\sqrt{p}}{2m})$.
+	Let $|BC| = n$. Then, from the triangle inequality $|AC| < |AB| + |BC|$, we have that $|AC| = m + n - s$, $s \in \mathbb{N}$ and $s \leq m$.
+	By the Grid Theorem, we can assume that $A = (-\frac{m}{2}; 0)$, $B = (\frac{m}{2}; 0)$, and $C = (\frac{a}{2m}; \frac{b\sqrt{p}}{2m})$.
 We find the distance between points $A$ and $C$, and points $B$ and $C$ in coordinates, and form a system of equations
@@
 	\end{equation}
-	According to the assumption, let $s=2t+1$, where $t\in\mathbb{N}_0$.
+	Since $s$ is assumed to be odd, let $s=2t+1$, where $t\in\mathbb{N}_0$.
 	Then equation (\ref{eq:33}) can be rewritten as:
@@
 	If equation (\ref{eq:35}) has integer solutions, then both $p$ and $b$ are odd.
-	The right-hand side is divisible by 4. Then $pb^2\equiv3\operatorname{mod}{4}$.
-	But $b^{2}\equiv1(\operatorname{mod}4)$.
-	Thus, $p=4k+3$, $k\in \mathbb{N}_{0}$.
+	The right-hand side is divisible by 4, which implies that  $pb^2\equiv3\operatorname{mod}{4}$.
+	Since $b^{2}\equiv1(\operatorname{mod}4)$, it follows that $p=4k+3$, $k\in \mathbb{N}_{0}$.
 \end{proof}
 
-Then the following theorem immediately appears.
+The following theorem follows immediately.
 
@@
 	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and let the $n$-th Erdős curve satisfy $|M_1 M_3| - |M_2 M_3| = n$.
-	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3| = m - s$, where $s$ is an even number, is also an odd number.
-	Let $m$ be an even number; then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is an even number, as the difference of two even numbers.
+	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is an even number, is also an odd number.
+	Let $m$ be an even number; then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is an even number, as a difference of two even numbers.
 \end{proof}
 
-An even edge is a bit special case.
+%An even edge is a bit special case.
+The case when the length of an edge is even is slightly more specific and allows a more precise statement.
 
@@
 	\end{eqnarray}
-	Among two consecutive natural numbers $n$ and $(n-1)$,
-	exactly one of them is divisible by 2.
+	Among two consecutive natural numbers,
+	exactly one is even.
 	Therefore, the expression $4n(n-1)\equiv0\pmod{8}$.
-	Thus, the right-hand side of equation (\ref{eq:35.1}) is an even number and a multiple of 8.
+	Thus, the right-hand side of equation (\ref{eq:35.1}) is a multiple of 8.
 	Then $pb^2\equiv7\pmod{8}$.
@@
 	is that numbers $p$ and $b$ must be odd.
-	But $b^2\equiv1\pmod{8}$, so $p=8k+7$, $k\in\mathbb{N}_0$.
+	Since $b^2\equiv1\pmod{8}$, it follows that $p=8k+7$, $k\in\mathbb{N}_0$.
 \end{proof}
@@
 
-	Assume the opposite.
-Let $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}_4}'$.
-Introduce Cartesian coordinates such that $M_1(-1;0), M_2(1;0), O(0;0)$.
-Then $M_3$ and $M_4$ lie on the median perpendicular to $M_1M_2$.
+	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}_4}'$.
+	Employ th Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
+	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
 	Since the distance $|M_3M_4|$ is an integer, the area of triangle $M_1M_3M_4$ is a rational number.
-By Definition 6, this means that $\operatorname{char}{M}={1}$.
-Then by the theorem about the lattice we have $M_3(0;t/4)$.
-Without loss of generality, let us assume $t\in\mathbb{N}$.
-Let $|M_1M_3|=s$.
-Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
+	By Definition~\ref{def:char}, this means that $\operatorname{char}{M}={1}$.
+	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
 	$$
@@
 	\end{equation}
-	Find the solutions to equation (\ref{eq: 42}) in natural numbers. To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$. Then equation (\ref{eq: 42}) takes the form:
+	Now, we find the solutions to equation (\ref{eq: 42}) in natural numbers.
+	To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
+	Then equation (\ref{eq: 42}) takes the form:
 	\begin{equation}
@@
 In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$, which contradicts that $k\in\mathbb{N}$.
-	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1,k=0$, which contradicts that $s,k\in\mathbb{N}$.
+	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$, which also contradicts that $s,k\in\mathbb{N}$.
 
-	Thus, equation (\ref{eq: 42}) has no natural solutions in natural numbers.
-	The obtained contradiction completes the proof of the lemma.
+	Thus, equation (\ref{eq: 42}) has no natural solutions.
+	The obtained contradiction completes the proof.
 \end{proof}
 
-The following result is a key detail of bounds presented
+The following result is a key element for bounds presented
 in~\cite{solymosi2003note} and~\cite{my-semi-general-5-4-bound-2019}:
@@
 	If a triangle $T$ has integer side lengths $a \leq b \leq c$,
-	then the minimal height $m$ of it is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
+	then its minimal height $m$ is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
 \end{lemma}
 
-If we exclude characteristics of the form $4k+3$, this Solymosi's result can be sharpened a bit.
+If we exclude characteristics of the form $4k+3$, this Solymosi's result can be sharpened slightly.
 
@@
 	Indeed, consider the triangle $ABC$ and let $|BC|=a$, $|AC|=b$.
-	From the triangle inequality $|AB| < |BC|+|AC|$,
+	Using the triangle inequality $|AB| < |BC|+|AC|$,
 	we can represent the length of side $AB$ as $|AB|=a+b-s$, $s \in \mathbb{N}$ and $s \leq a$.

### 2024-07-11 char-4k-bound/char-4k-bound.tex
[4k3] Embrace all the fixes from Dr. A.S. Usachev

@@
 	\label{lem:triangle_height_our}
-	Let a triangle $ABC$ have integer characteristic different from $4k+3$, $k\in \mathbb{N}_0$, with $a\leq b\leq c$. Then the smallest height of the triangle $ABC$ is at least $(2a-1)^{1/2}$.
+	Let a triangle $ABC$ have the characteristic different from $4k+3$, $k\in \mathbb{N}_0$, with $a\leq b\leq c$. Then the smallest height of the triangle $ABC$ is at least $(2a-1)^{1/2}$.
 \end{lemma}
@@
 \begin{proof}
-	By Corollary \ref{cr:3}, a triangle with side $c=a+b-1$ cannot have a characteristic different from $4k+3$, $k\in \mathbb{N}_0$. Therefore, in a triangle with integer sides, we have $a+b\geq c+2.$
+	By Lemma~\ref{lem:triangle_4k_plus_3}, a triangle with a side $c=a+b-1$ cannot have a characteristic different from $4k+3$, $k\in \mathbb{N}_0$.
+Therefore, in a triangle with integer sides, we have $a+b\geq c+2.$
 
@@
 
-The following Lemma is a classic Erdos-style intersection-enumeration one,
+The following lemma is a classical Erdos-style intersection-enumeration one,
 empowered by our Weeding Theorem I.
@@
 	Assuming that the edges $M_1M_2$ and $M_3M_4$ have odd lengths,
-	we infer that that $|M_1M_2|\geq3$ and $|M_3M_4|\geq3$.
+	we infer that $|M_1M_2|\geq3$ and $|M_3M_4|\geq3$.
 	Then,
@@
 	Let's consider the third case.
-Suppose the lengths of edges $M_1M_2$ and $M_3M_4$ are different and without loss of generality suppose edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
+Suppose the lengths of edges $M_1M_2$ and $M_3M_4$ are different.
+Without loss of generality suppose the edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
 	Then, for each point $N\in M$, one of the following conditions is satisfied:
@@
 
-	Thus, the lemma is proved.
+	This proves the assertion.
 \end{proof}
@@
 \overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{}\nonumber\\
-\overline{d}(2,37) &>& 20000.
+\overline{d}(2,37) &>& 20\ 000.
 \end{eqnarray}
 
-Now we are finally ready to proof our main result and improve the bound~\eqref{eq:chebsb_lower_bound}
+Now we are finally ready to proof our main result.
+It improves the bound~\eqref{eq:chebsb_lower_bound}
 in our special case.
@@
 \begin{proof}
-	For $n = 3$, we have $\operatorname{diam} M \geq 3$ (achieved by an isosceles triangle with sides 2, 3, 3), and the theorem statement is obvious.
+	For $n = 3$, we have $\operatorname{diam} M \geq 3$ (achieved by an isosceles triangle with sides 2, 3, 3), and the assertion is obvious.
 
@@
 	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
-	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20,000$ and our estimate on the diameter is also true.
+	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
 	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10,655$.
@@
 
-	Also, we should notice that our approach and especially Weeding Theorems do not require semi-general position and can be applied to tighten the bound from~\cite{my-pps-linear-bound-2019}
+	Also, we should notice that our approach and  specifically Weeding Theorems do not require semi-general position and can be applied to tighten the bound from~\cite{my-pps-linear-bound-2019}
 	in the special case of characteristic $4k+1$ or $4k+2$.

### 2024-07-25 char-4k-bound/char-4k-bound.tex
[4k3] Shorten the affiliation
[4k3] ref -> eqref

@@
 	\\
-	\textit{Voronezh State University, Voronezh, Russia}
+	\textit{Voronezh State University}
 }
@@
 	\end{equation}
-	Multiply each equation by $2m$ and square both sides of the equations in system (\ref{eq:29})
+	Multiply each equation by $2m$ and square both sides of the equations in system~\eqref{eq:29}
 	\begin{equation}
@@
 	Since $s$ is assumed to be odd, let $s=2t+1$, where $t\in\mathbb{N}_0$.
-	Then equation (\ref{eq:33}) can be rewritten as:
+	Then equation \eqref{eq:33} can be rewritten as:
 	\begin{equation}
@@
 	\end{eqnarray}
-	If equation (\ref{eq:35}) has integer solutions, then both $p$ and $b$ are odd.
+	If equation~\eqref{eq:35} has integer solutions, then both $p$ and $b$ are odd.
 	The right-hand side is divisible by 4, which implies that  $pb^2\equiv3\operatorname{mod}{4}$.
@@
 	Taking into account that $s = 2t + 1$, $t \in \mathbb{N}_0$,
-	we bring equation (\ref{eq:33.1}) to the form:
+	we bring equation \eqref{eq:33.1} to the form:
 	\begin{equation}
@@
 	Then $pb^2\equiv7\pmod{8}$.
-	A necessary condition for the existence of an integer solution to equation (\ref{eq:35.1})
+	A necessary condition for the existence of an integer solution to equation~\eqref{eq:35.1}
 	is that numbers $p$ and $b$ must be odd.
@@
 	\end{equation}
-	Now, we find the solutions to equation (\ref{eq: 42}) in natural numbers.
+	Now, we find the solutions to equation~\eqref{eq: 42} in positive integers.
 	To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
-	Then equation (\ref{eq: 42}) takes the form:
+	Then equation~\eqref{eq: 42} takes the form:
 	\begin{equation}
@@
 	\end{equation}
-	Break down the left-hand side of equation (\ref{eq: 43}) into factors:
+	Break down the left-hand side of equation~\eqref{eq: 43} into factors:
 	\begin{equation}
@@
 
-	Thus, equation (\ref{eq: 42}) has no natural solutions.
+	Thus, equation~\eqref{eq: 42} has no positive integer solutions.
 	The obtained contradiction completes the proof.
@@
 	\end{multline}
-	and this equality is possible when $a=b$. We rewrite expression (\ref{eq: 39}) as follows:
+	and this equality is possible when $a=b$. We rewrite expression~\eqref{eq:39} as follows:
 	$$
@@
 
-	From inequality (\ref{eq:46}), we obtain
+	From inequality~\eqref{eq:46}, we obtain
 	\begin{equation}
@@
 
-	For estimation (\ref{eq:strips_5_4}) and $n \geq 7^4$, we have
+	For estimation~\eqref{eq:strips_5_4} and $n \geq 7^4$, we have
 	\begin{equation}

### 2024-07-25 char-4k-bound/char-4k-bound.tex
[4k3] 11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3 -> rails_3_8

@@
 
-For the sake of completeness, on Figure~\ref{fig:11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}
+For the sake of completeness, on Figure~\ref{fig:rails_3_8.png}
 we give an example of rails set
@@
 \begin{figure}
-\center{\includegraphics[width=0.95\linewidth]{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}}
+\center{\includegraphics[width=0.95\linewidth]{rails_3_8.png}}
 \caption{The rails set with characteristic 255255}
-\label{fig:11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}
+\label{fig:rails_3_8.png}
 \end{figure}

### 2024-07-25 char-4k-bound/char-4k-bound.tex
[4k3] Just fix a label
[4k3] A bunch of minor fixes

@@
 	\begin{equation}
-	pb^2 = (2t + 1)(4q + 2n - 2t - 1)(2n - 2t - 1)(4q - 2t - 1).
+	pb^2 = (2t + 1)(4q + 2n - 2t - 1)(2n - 2t - 1)(4q - 2t - 1)
 	\label{eq: 34.2}
@@
 	Among two consecutive natural numbers,
-	exactly one is even.
-	Therefore, the expression $4n(n-1)\equiv0\pmod{8}$.
-	Thus, the right-hand side of equation (\ref{eq:35.1}) is a multiple of 8.
+	exactly one is even,
+	so $4n(n-1)\equiv0\pmod{8}$.
+	Thus, the right-hand side of equation~\eqref{eq:35.1} is a multiple of 8.
 	Then $pb^2\equiv7\pmod{8}$.
@@
 	Let $M \in \overline{\mathfrak{M}_n}'$ and $M_1, M_2 \in M$.
-Then all points in the set $M$ lie on the circle $cr(M_1, M_2)$.
-Otherwise, according to Statement 4, $\operatorname{char}{M}={8k+7}$, $k\in \mathbb{N}_0$.
+	Then all points in the set $M$ lie on $cr(M_1, M_2)$.
+	Otherwise, according to Lemma~\ref{lem:pre_weeding_II}, $\operatorname{char}{M}={8k+7}$, $k\in \mathbb{N}_0$.
 
 	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}_4}'$.
-	Employ th Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
+	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
 	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
@@
 	\begin{equation}
-		\label{eq: 46}
+		\label{eq:46}
 		n \leq 2 \left( \frac{p^{4/5}}{\sqrt{2}} + 1\right) + 1=\frac{2p^{4/5}}{\sqrt{2}} +3 = \sqrt{2}p^{4/5}+3

### 2024-07-25 char-4k-bound/char-4k-bound.tex
[4k3] char385.png -> rails_2_14.png
[4k3] Fix duplicate ref

@@
 
-For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
+For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018-translit}:
 $\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
@@
 
-Figure~\ref{fig:char385} shows a rails IPS of characteristic $385 = 4 \cdot 96 + 1$ presented in~\cite{avdeev2019particular}.
+Figure~\ref{fig:rails_2_14} shows a rails IPS of characteristic $385 = 4 \cdot 96 + 1$ presented in~\cite{avdeev2019particular}.
 Its coordinates are
@@
 \begin{figure}
-\center{\includegraphics[width=0.95\linewidth]{char385.png}}
+\center{\includegraphics[width=0.95\linewidth]{rails_2_14.png}}
 \caption{IPS of cardinality 16 and diameter 2189}
-\label{fig:char385}
+\label{fig:rails_2_14}
 \end{figure}

### 2024-07-25 char-4k-bound/char-4k-bound.tex
[4k3] hyperbolae -> hyperbolas (because others write so)

@@
 when the characteristic of the set is of the form $4k+1$ or $4k+2$.
-To achieve that, we combine hyperbolae-based and grid-based toolsets.
+To achieve that, we combine hyperbolas-based and grid-based toolsets.
 
@@
 	In 1945, Erdös gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
-	In any IPS he chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolae, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
+	In any IPS he chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
 Applying the same argument to the line segment $M_1M_3$, Erdös concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.
@@
 	however, Erdös did not only estimate the diameter,
-	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolae.
+	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolas.
 
@@
 	(in assumption that the set has at east 4 points).
-	All these lower bounds are based on the Erdos's framework --- cofocal hyperbolae.
+	All these lower bounds are based on the Erdos's framework --- cofocal hyperbolas.
 
@@
 Obviously, the 0-th Erdos curve is the perpendicular bisector of $M_1 M_2$,
-and all the other Erdos curves are branches of cofocal hyperbolae.
+and all the other Erdos curves are branches of cofocal hyperbolas.
 On Figure~\ref{fig:Erdos_curves_numbered}, the Erdos curves are shown for an edge of length 3.
@@
 
-	c) $N$ belongs to the intersection of one of the $(|M_1M_2|/2 - 1)$ hyperbolae with one of the $(|M_3M_4|/2 - 1)$ hyperbolae, which implies that there are no more than $4(|M_1M_2|/2 - 1)(|M_3M_4|/2 - 1)$ points.
+	c) $N$ belongs to the intersection of one of the $(|M_1M_2|/2 - 1)$ hyperbolas with one of the $(|M_3M_4|/2 - 1)$ hyperbolas, which implies that there are no more than $4(|M_1M_2|/2 - 1)(|M_3M_4|/2 - 1)$ points.
 
@@
 
-	c) $N$ belongs to the intersection of one of $(|M_1 M_2|-1)/2$ hyperbolae with one of $(|M_3M_4|-1)/2$ hyperbolae, which implies that there are no more than $(|M_1 M_2|-1)(|M_3 M_4|-1)$ points.
+	c) $N$ belongs to the intersection of one of $(|M_1 M_2|-1)/2$ hyperbolas with one of $(|M_3M_4|-1)/2$ hyperbolas, which implies that there are no more than $(|M_1 M_2|-1)(|M_3 M_4|-1)$ points.
 
@@
 
-	c) $N$ belongs to the intersection of one of $(|{M_1 M_2}|/2-1)$ hyperbolae with one of $(|M_3 M_4|-1)/2$ hyperbolae, which implies that there are no more than $4(|M_1 M_2|/2-1)(|M_3 M_4|-1)/2$ points.
+	c) $N$ belongs to the intersection of one of $(|{M_1 M_2}|/2-1)$ hyperbolas with one of $(|M_3 M_4|-1)/2$ hyperbolas, which implies that there are no more than $4(|M_1 M_2|/2-1)(|M_3 M_4|-1)/2$ points.
 

### 2024-09-27 char-4k-bound/char-4k-bound.tex
[4k3] Last edits from Usachev

@@
 	In 1945, Erdös gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
-	In any IPS he chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
+	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
 Applying the same argument to the line segment $M_1M_3$, Erdös concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.
@@
 
-	In~\cite{our-vmmsh-2018-translit} the constant (for $n\geq 4$) was improved to $0.3457$
+	In~\cite{our-vmmsh-2018-translit} the constant (for $\# M \geq 4$) was improved to $0.3457$
 	employing Point Packing in a Square Problem~\cite{markot2005newverified,costa2013valid}.
@@
 		\label{eq:chebsb_lower_bound}
-		\diam M \geq \left(\frac n 5 \right)^{5/4}
+		\diam M \geq \left(\frac {\# M} 5 \right)^{5/4}
 	\end{equation}
@@
 	\label{def:char}
-	The characteristic of a planar IPS $\mathcal{M}$ is a squarefree number $q$
+	The characteristic of a planar IPS $M$ is a squarefree number $q$
 	such that the area of any triangle $M_1M_2M_3$, $\{M_1,M_2,M_3\}\subset M$,

### 2024-09-27 char-4k-bound/references.tex
Switch to pregenerated references

--- 2020-03-25 pps-linear-bound/pps-linear-lower-bound.tex
@@
-\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
-\usepackage[T2A]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[russian,english]{babel} %используем русский и английский языки с переносами
-\usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
-\usepackage[unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
-%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
-\usepackage{indentfirst} % включить отступ у первого абзаца
-\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
-\graphicspath{{illustr/}}%путь к рисункам
-
-\usepackage{geometry} % Меняем поля страницы.
-\geometry{left=2cm}% левое поле
-\geometry{right=1cm}% правое поле
-\geometry{top=2cm}% верхнее поле
-\geometry{bottom=2cm}% нижнее поле
-
-\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
-\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра
-
-\sloppy
-
-\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}
-
-%\usepackage[backend=biber,style=ajc2020unofficial]{biblatex}
-%\addbibresource{../common/notmy.bib}
-%\addbibresource{../common/my.bib}
-
-\usepackage{cite}
-
-%\usepackage{biblatex2bibitem}
-
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{construction}[theorem]{Construction}
-\newtheorem{remark}[theorem]{Remark}
-\newtheorem{problem}[theorem]{Problem}
-
-%Only referenced equations are numbered
-\let\etoolboxforlistloop\forlistloop % save the good meaning of \forlistloop
-\usepackage{mathtools}
-\mathtoolsset{showonlyrefs}
-\let\forlistloop\etoolboxforlistloop % restore the good meaning of \forlistloop
-
-% https://tex.stackexchange.com/questions/220268/biblatex-and-autonum-dont-work-together
-
-%\mathtoolsset{showonlyrefs=false}
-% (write an equation/multline to be force-numbered here)
-%\mathtoolsset{showonlyrefs=true}
-
-\begin{document}
-
-%\renewcommand{\bibname}{Список цитированной литературы}
-%\renewcommand\refname{\bibname}
-% !!!
-% The text starts here
-
-\title{
-	On existence of integral point sets and their diameter bounds
-	\footnote{
-		This work was carried out at Voronezh State University and supported by the Russian Science
-		Foundation grant 19-11-00197.
-	}
-}
-
-\author{
-	N.N. Avdeev
-	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
-	\\
-	\textit{Voronezh State University, Voronezh, Russia}
-}
-
-\maketitle
-
-\paragraph{Abstract.}
-A point set $M$ in $m$-dimensional Euclidean space is called an integral point set if all the distances between the
-elements of $M$ are integers, and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
-We improve the linear lower bound for diameter of planar integral point sets.
-This improvement takes into account some results related to the Point Packing in a Square problem.
-Then for arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
-we give a construction of an integral point set $M$ of $n$ points in $m$-dimensional Euclidean space,
-where $M$ contains points $M_1$ and $M_2$ such that the distance between $M_1$ and $M_2$ is exactly $d$.
-
-\section{Introduction}
-Let $\mathbb{N}$ be the set of all positive integers and let $|M_1 M_2|$ denote the Euclidean distance
-between points $M_1$ and $M_2$ in a finite-dimensional space $\mathbb{R}^m$
-(and, more generally, let $|\Delta|$ denote the length of line segment $\Delta$).
-An \textit{integral point set} in $m$-dimensional Euclidean space, $m>2$, is a point set $M$ such that all the distances between the
-points of $M$ are integers and $M$ is not situated on an $(m-1)$-dimensional hyperplane.
-Erd{\H{o}}s and Anning proved~\cite{anning1945integral,erdos1945integral} that every integral point set consists of a finite number of points.
-Taking this into account, we denote the set of all integral point sets of $n$ points in $m$-dimensional Euclidean space by
-$\mathfrak{M}(m,n)$ (using the notation in~\cite{our-vmmsh-2018-translit})
-and denote the set of all integral point sets in $m$-dimensional Euclidean space by $\mathfrak{M}(m,\mathbb{N})$.
-The symbol $\# M$ will be used for the cardinality of $M$, that is the number of points in $M$ in our case.
-
-For every finite point set, its diameter is naturally defined as
-\begin{equation}
-	\operatorname{diam} M = \max_{A,B\in M} |AB|
-	.
-\end{equation}
-Another emerging question is: how does the diameter of an integral point set depend on its cardinality?
-One can easily see that every $M\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = h$
-can be dilated to $M_p\in\mathfrak{M}(m,n)$ with $\operatorname{diam} M = ph$
-for every $p\in\mathbb{N}$.
-So, the above question should be rephrased:
-how does \textit{the least possible} diameter of an integral point set depend on its cardinality?
-In order to answer this question, the following function was introduced~\cite{kurz2008bounds,kurz2008minimum}:
-\begin{equation}
-	d(m,n) = \min_{M\in\mathfrak{M}(m,n)} \operatorname{diam} M = \min_{M\in\mathfrak{M}(m,n)} \max_{A,B\in M} |AB|
-	.
-\end{equation}
-We also refer to~\cite{kurz2008bounds} for a list of known exact values of $d(m,n)$ and its bounds.
-In the present paper, we mostly focus on the case $m=2$.
-
-The most significant breakthrough on the planar case was done by Solymosi~\cite{solymosi2003note},
-who proved that $cn \leq d(2,n)$ for a sufficiently small constant $c$.
-Following Solymosi's proof carefully,
-one can derive that the inequality holds at least for $c = 1/24$.
-(See~\cite[Exercise 2.6]{garibaldi2005erdos} for some remarks.)
-The constant was improved in~\cite{our-mz-rus-translit} to $1/8$ for all $n$ and in~\cite{our-vmmsh-2018-translit}
-to $3/8$ for sufficiently large $n$.
-
-The paper~\cite{solymosi2003note} contains one more interesting result.
-Let us define a function which is ``dual'' to $d(m,n)$ in some sense:
-\begin{equation}
-	l(m,n) = \min_{M\in\mathfrak{M}(m,n)} \min_{A,B\in M} |AB|
-	.
-\end{equation}
-Solymosi proved that $l(2,n)\leq 2$.
-
-In the present paper we improve Solymosi's results:
-first, we obtain a larger constant $c = 5/11$ in Theorem~\ref{thm:main_estimate},
-using the combined approach with the Point Packing in a Square problem
-(this approach is different from Solymosi's one);
-second, we prove that $l(m,n)=1$ for all possible $m$ and $n$.
-
-Sections 2--4 are devoted to the linear lower bound.
-Each section consists of 2 subsections: General results and Special results.
-In General results, we do not aim to provide tight estimates;
-they are discussed in Special results.
-
-Section 5 is devoted to integral point sets with distance 1 and has its own structure.
-
-\section{Collinear subsets of integral point sets}
-Integral point sets with many collinear points are very dominating examples.
-In this section we estimate the cardinality of an intersection of an integral point set with a line segment.
-
-\subsection{General results}
-
-Assuming that the planar integral point set contains many collinear points,
-the following result holds.
-\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
-	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
-	at least $n^\delta$ collinear points there exists a $n_0 (\varepsilon)$
-	such that for all $n \geq n_0 (\varepsilon)$ we have
-	\begin{equation}
-		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
-		.
-	\end{equation}
-\end{theorem}
-However, the estimate is rather unsuitable for our needs,
-as it does not provide the values of all the constants.
-To obtain the required estimate, we now prove a generalization of~\cite[Lemma 3]{our-vmmsh-2018-translit}.
-
-\begin{definition}
-	For a line segment $M_1 M_2$ and an integer $k$, such that $-|M_1 M_2| < k < |M_1 M_2|$,
-	we define a $\rho(k,M_1, M_2)$-curve as the set of points $N$
-	for which the equality $|N M_1| - |N M_2| = k$ holds.
-\end{definition}
-So, in the planar case a $\rho(k,M_1, M_2)$-curve
-is a branch of a hyperbola for $k\neq 0$ and the perpendicular bisector of line segment $M_1 M_2$ for $k=0$.
-
-\begin{proposition}
-	\label{obs:rho_curves}
-	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
-	the equality $|M_1 M_2| = |M_3 M_4| =h $ holds, line segments $M_1 M_2$ and $M_3 M_4$ do not coincide
-	and vectors $\overrightarrow{M_1 M_2}$ and $\overrightarrow{M_3 M_4}$ are of equal direction,
-	then for a fixed integer $k$, $|k| < h$, the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve
-	do not intersect.
-\end{proposition}
-
-\begin{proof}
-	For $k=0$, the claim is obvious.
-	Indeed, in this case the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve
-	are distinct hyperplanes orthogonal to $l$ and thus do not intersect.
-
-	For $k\neq 0$, suppose the contrary.
-	Let $N$ be a point in the intersection.
-
-	Consider a plane $\Pi \supset (l \cup \{N\})$
-	with a Cartesian coordinate system such that $M_i = (x_i,0)$ for $1 \leq i \leq 4$.
-	Let $R_1$ and $R_2$ be the projections
-	of the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve onto $\Pi$.
-
-	Then $R_1$ is a branch of a hyperbola defined by the equation
-	\begin{equation}
-		x- \frac{x_1+x_2}{2}=a(h,k)\frac{\sqrt{y^2+b^2(h,k)}}{b(h,k)}
-	\end{equation}
-	and $R_2$ is a branch of a hyperbola defined by the equation
-	\begin{equation}
-		x- \frac{x_3+x_4}{2}=a(h,k)\frac{\sqrt{y^2+b^2(h,k)}}{b(h,k)}
-		,
-	\end{equation}
-	where $a(h,k)$ and $b(h,k)$ are coefficients which depend on $h$ and $k$.
-	For $N=(x_N, y_N) \in R_1 \cap R_2$ we have
-	\begin{equation}
-		\begin{cases}
-			x_N- \dfrac{x_1+x_2}{2}=a(h,k)\dfrac{\sqrt{y_N^2+b^2(h,k)}}{b(h,k)}
-			\\[4pt]\\
-			x_N- \dfrac{x_3+x_4}{2}=a(h,k)\dfrac{\sqrt{y_N^2+b^2(h,k)}}{b(h,k)}
-		\end{cases}
-	\end{equation}
-	and, thus, $x_1+x_2 = x_3+x_4$.
-	On the other hand, $x_2 - x_1 = x_4 - x_3 = h$,
-	and we conclude that $x_1 = x_3$ as well as $x_2 = x_4$.
-	So, line segments $M_1 M_2$ and $M_3 M_4$ coincide.
-	The contradiction concludes the proof.
-\end{proof}
-
-\begin{lemma}
-	\label{lem:2k-1_segments}
-	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $l$ be a straight line.
-	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset l$ with endpoints in $M$,
-	such that $|\Delta_i| = k$.
-\end{lemma}
-\begin{proof}
-	Consider a point $N\in M \setminus l$.
-	Let $\Delta_i = M_i^1 M_i^2$, and
-	let vectors $\overrightarrow{M_i^1 M_i^2}$ and $\overrightarrow{M_j^1 M_j^2}$ be of equal direction for every $i$ and $j$.
-	Then for each $i$ there is a $\rho(n_i, M_i^1, M_i^2)$-curve containing $N$.
-	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
-	otherwise the $\rho(n_i,M_i^1, M_i^2)$-curve and the $\rho(n_i,M_j^1, M_j^2)$-curve, $j\neq i$, do not intersect.
-	There can be only $2k-1$ distinct values for $n_i$,
-	so there are at most $2k-1$ distinct segments $\Delta_i$.
-\end{proof}
-
-Setting $k=1$ in Lemma~\ref{lem:2k-1_segments},
-we obtain the following result (which is exactly~\cite[Lemma 3]{our-vmmsh-2018-translit}).
-\begin{corollary}
-	\label{cor:only_one_distance_1_on_straight_line}
-	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $l$ be a straight line.
-	Then there is at most one pair of points $M_1,M_2\in M \cap l$
-	such that $|M_1 M_2| = 1$.
-\end{corollary}
-
-\begin{lemma}
-	\label{lem:line_segment_with n_squared_plus_one_points}
-	Let $\Delta$ be a straight line segment, $|\Delta|=l$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = n^2 + 1$.
-	Then
-	\begin{equation}
-		l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
-		.
-	\end{equation}
-\end{lemma}
-
-\begin{proof}
-	Any $n^2+1$ points, including the endpoints of $\Delta$, split the segment $\Delta$ into $n^2$
-	sequential segments $\Delta_i$.
-	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
-	at most three segments of length 2, etc.
-	The following two expressions for sums conclude the proof:
-	\begin{equation}
-		1 + \sum_{k=1}^n (2k-1) = n^2 + 1
-		,
-	\end{equation}
-	\begin{equation}
-		\sum_{k=1}^n k(2k-1) = \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
-		.
-	\end{equation}
-\end{proof}
-
-Now we will estimate the length of a line segment that intersects an integral point set
-by an arbitrary number of points.
-
-\begin{lemma}
-	\label{lem:line_segment_cardinality}
-	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = t$.
-	Then
-	\begin{equation}
-		\label{eq:estimate_for_segment_length}
-		b\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}
-		.
-	\end{equation}
-\end{lemma}
-
-\begin{proof}
-	Let $f(k)$ denote the minimal length of a line segment
-	that intersects an integral point set by $k$ points.
-	We observe that $f(k) > f (k-1)$.
-	Due to Lemma~\ref{lem:line_segment_with n_squared_plus_one_points},
-	$f(n^2+1) \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$.
-
-	%Consider $t$ such that $n^2 + 1 \leq t \leq (n+1)^2$. %,
-	%then $f(t) > f(n^2 +1)$.% and $n \geq \sqrt{t} - 1$.
-	For $t\in\mathbb{N}$ the inequality $(\sqrt{t} - 1)^2 +1 \leq t$ holds,
-	thus
-	%and derive that
-	\begin{equation}
-		\label{eq:line_segment_length}
-		f(t) \geq f((\sqrt{t} - 1)^2 +1) \geq \frac{2}{3}(\sqrt{t} - 1)^3+\frac{1}{2}(\sqrt{t} - 1)^2-\frac{1}{6}(\sqrt{t} - 1)
-		=
-		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2}
-		.
-	\end{equation}
-\end{proof}
-
-\subsection{Special results}
-
-In~\cite{kurz2008bounds}, thorough investigation of planar integral point sets with diameter at most 10~000
-is carried out;
-thus, we can restrict our attention to intersections of planar integral point sets with line segments of length more than 10~000.
-
-Lemma~\ref{lem:line_segment_cardinality} leads to the following proposition.
-
-\begin{proposition}
-	\label{obs:estimate_points_on_straight_line}
-	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
-	Let $\#(\Delta \cap M) = k$ and $b>10000$.
-	Then $k \leq \gamma_2 b + 6$,
-	where
-	\begin{equation}
-		\gamma_2 = \frac{3846}{2593 \sqrt{647}-5823} = 0.063958...
-	\end{equation}
-\end{proposition}
-
-\begin{proof}
-	We know the maximum number of points for planar integral point sets of diameters at most $10000$.
-	Thus, we are interested in integral $t$ such that estimate~\eqref{eq:estimate_for_segment_length}
-	holds for $b > 10000$.
-	So, let us consider $t\geq 647$ and find a coefficient $\gamma_2$,
-	such that the inequality
-	\begin{equation}
-		\label{eq:inequality_for_linear_estimate_of_segment_length}
-		\frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5 }{6}t^{1/2} \geq \frac{t-6}{\gamma_2}
-	\end{equation}
-	holds for all $t\geq 647$.
-	For such $t$,
-	the left-hand side of~\eqref{eq:inequality_for_linear_estimate_of_segment_length} obviously grows faster than the right-hand side.
-	Turning~\eqref{eq:inequality_for_linear_estimate_of_segment_length} into the same equality
-	and solving it for $t=647$, we obtain the required estimate.
-\end{proof}
-
-To conclude, we remark that
-Corollary~\ref{cor:only_one_distance_1_on_straight_line}
-and
-Proposition~\ref{obs:estimate_points_on_straight_line}
-are the only results of this section
-that are used in the proofs below;
-all the other are auxiliary.
-
-\section{Planar integral point set in a square}
-
-In this section, we develop some methods that employ the fact that every planar integral point set is situated in a square.
-Then we establish a connection between distances in integral point sets and the Point Packing in the Square problem.
-
-\subsection{General results}
-
-Let us introduce the problem, basic notions and results.
-
-\begin{problem}[Point Packing in a Square (PPS)~\cite{locatelli2002packing,costa2013valid}]
-	\label{problem:PPS}
-	Given an integer $k > 1$, place $k$ points in the unit square $U = [ 0 , 1 ]^2$ such that their
-	minimum pairwise distance $m$ is maximal.
-\end{problem}
-
-\begin{definition}
-	For each $k > 1$, the corresponding maximal distance $m$ from Problem~\ref{problem:PPS}
-	is called the $k$-th PPS coefficient and denoted by $\varphi_k$.
-\end{definition}
-So, it's impossible to place $k$ points in a unit square in such a way that each pairwise distance of the points is greater than $\varphi_k$.
-
-\begin{theorem}
-	\label{thm:varphi_k_bounds}
-	\cite{costa2013valid}
-	For every $k\geq 2$ the following inequality holds:
-	\begin{equation*}
-		\sqrt{\frac{2}{k\sqrt{3}}}
-		\leq
-		\varphi_k
-		\leq
-		\frac{1}{k-1} +
-		\sqrt{
-			\frac{1}{(k-1)^2}
-			+
-			\frac{2}{(k-1)\sqrt{3}}
-		}
-	\end{equation*}
-\end{theorem}
-
-Now we will estimate the side length of a square which contains a given planar integral points set.
-
-\begin{lemma}
-	\cite[Lemma 4]{our-vmmsh-2018-translit}
-	\label{lem:square_container}
-	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
-	Then $M$ is situated in a square of side length $d$.
-\end{lemma}
-
-\begin{proof}
-	Consider $M_1, M_2 \in M$ such that $|M_1 - M_2| = d$
-	and a Cartesian coordinate system such that
-	$M_1 = (-d/2; 0)$, $M_2 = (d/2; 0)$.
-
-	Then for $M_i =(x_i, y_i)\in M$, $i > 2$,
-	the inequality $|x_i| < d/2$ holds
-	(otherwise we get $|M_i - M_1| > d$ or $|M_i - M_2| > d$).
-	Moreover, $|y_i| < d$ (otherwise we get $|M_i - M_1| > d$ or $|M_i - M_2| > d$).
-
-	Set $y_{\max} = \max_{i} y_i$, $y_{\min} = \min_{i} y_i$, $M_{\max}=(x_{\max}, y_{\max})$, $M_{\min}=(x_{\min}, y_{\min})$
-	(in case of multiple points providing the maximum or the minimum, the choice is arbitrary).
-	Then
-	\begin{multline}
-		d \geq |M_{\max} - M_{\min}| = \sqrt{(x_{\max} - x_{\min})^2 + (y_{\max} - y_{\min})^2}
-		\geq
-		\\\geq
-		\sqrt{(y_{\max} - y_{\min})^2} =
-		y_{\max} - y_{\min}
-	\end{multline}
-	Thus, the sides of the square that contains $M$ are $x=\pm d/2$,
-	$y=y_{\max}$ and $y=y_{\max} - d$.
-\end{proof}
-
-\subsection{Special results}
-
-It is clear that for $n>123$ we have $d(2,n) \geq d(2,123)$.
-Based on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
-\cite{kurz2008bounds}, we derive the following proposition.
-\begin{proposition}
-	\label{obs:4_leq_n_leq_21491}
-	The inequality
-	\begin{equation}
-		d(2,n) \geq 3^{1/4}\cdot2^{-3/2} \cdot n
-	\end{equation}
-	holds for $4 \leq n \leq 21491$.
-\end{proposition}
-
-Therefore, we will focus on planar integral points sets of more than 21491 points.
-
-Now we will use the upper bound of Theorem~\ref{thm:varphi_k_bounds}
-to estimate $\varphi_n$ with $\beta/\sqrt{n-2}$.
-
-\begin{proposition}
-	\label{varphi_n_where_n_geq_21492}
-	For $n \geq 21492$ we have
-	\begin{equation}
-		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{n-2}}
-		,
-	\end{equation}
-	where
-	\begin{equation}
-		\beta = \frac{1}{\sqrt{21490}} + \sqrt{ \frac{2}{\sqrt{3}} + \frac{1}{21490} } < 1.07464
-		.
-	\end{equation}
-\end{proposition}
-
-\begin{proof}
-	Indeed, for $n \geq 21492$ we have
-	\begin{multline*}
-		\varphi_n
-		\leq
-		\varphi_{n-1}
-		\leq
-		\frac{1}{n-2} +
-		\sqrt{
-			\frac{1}{(n-2)^2}
-			+
-			\frac{2}{(n-2)\sqrt{3}}
-		}
-		=
-		\\=
-		\frac{1}{\sqrt{n-2}}
-		\cdot
-		\frac{1}{\sqrt{n-2}}
-		 +
-		\sqrt{
-			\frac{1}{(n-2)^2}
-			+
-			\frac{2}{(n-2)\sqrt{3}}
-		}
-		\leq
-		\\ \leq
-		\frac{1}{\sqrt{n-2}}
-		\cdot
-		\frac{1}{\sqrt{21490}}
-		 +
-		\sqrt{
-			\frac{1}{(n-2)^2}
-			+
-			\frac{2}{(n-2)\sqrt{3}}
-		}
-		=
-		\\=
-		\frac{1}{\sqrt{n-2}}
-		\left(
-		\frac{1}{\sqrt{21490}}
-		 +
-		\sqrt{
-			\frac{1}{n-2}
-			+
-			\frac{2}{\sqrt{3}}
-		}
-		\right)
-		\leq
-		\\ \leq
-		\frac{1}{\sqrt{n-2}}
-		\left(
-		\frac{1}{\sqrt{21490}}
-		 +
-		\sqrt{
-			\frac{1}{21490}
-			+
-			\frac{2}{\sqrt{3}}
-		}
-		\;
-		\right)
-		.
-	\end{multline*}
-\end{proof}
-
-\section{Lower bound for the diameter}
-
-In this section,
-we improve the lower bound for the minimum diameter of planar integral point sets
-employing the results of two previous sections.
-
-\subsection{General results}
-
-\begin{definition}
-	A \textit{cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is the union of two straight lines:
-	the line through $M_1$ and $M_2$,
-	and the perpendicular bisector of line segment $M_1 M_2$.
-\end{definition}
-
-\begin{lemma}
-	\label{lem:intervals_cross}
-	If open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect,
-	%$M_1 \neq M_2$, $M_3 \neq M_4$,
-	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains 2 or 4 points.
-\end{lemma}
-
-\begin{proof}
-	Consider straight lines $l^1_1$, $l^1_2$, $l^2_1$, $l^2_2$
-	such that $cr(M_1,M_2) = l^1_1 \cup l^1_2$ and $cr(M_3,M_4) = l^2_1 \cup l^2_2$.
-
-	The cases $l^1_1 = l^2_1$, $l^1_2 = l^2_2$ and $l^1_1 = l^2_2$, $l^1_2 = l^2_1$ are impossible,
-	because the centers of open line segments $M_1 M_2$ and $M_3 M_4$ do not coincide as far as
-	the open line segments do not intersect.
-	So, all the possible cases (up to enumeration of the straight lines) are:
-	\begin{itemize}
-		\item
-			$l^1_1 = l^2_1$, $l^1_2 \neq l^2_2$.
-			The straight lines $l^1_2$ and $l^2_2$ are both orthogonal to $l^1_1$ and thus do not intersect;
-			thus, $cr(M_1,M_2) \cap cr(M_3,M_4) = l^1_1 = l^2_1$.
-		\item
-			$l^1_1$ is parallel to $l^2_1$ and $l^1_2$ is parallel to $l^2_2$.
-			Then
-			\begin{equation*}
-				cr(M_1,M_2) \cap cr(M_3,M_4) = (l^1_1 \cup l^1_2)\cap(l^2_1 \cup l^2_2) =
-				\\=
-				(l^1_1 \cap l^2_2) \cup (l^1_2 \cap l^2_1)
-				,
-			\end{equation*}
-			that is two points.
-		\item
-			$l^1_1$ is neither parallel nor orthogonal to $l^2_1$.
-			Then the intersection contains 4 points.
-	\end{itemize}
-\end{proof}
-
-\begin{remark}
-	It is important that we consider the intersection of open line segments in the claim of Lemma~\ref{lem:intervals_cross},
-	so e.g. the cases $M_1 = M_3$ and $M_3 \in M_1 M_2$ satisfy the conditions of Lemma~\ref{lem:intervals_cross}.
-\end{remark}
-
-\begin{lemma}
-	\label{lemma:quadr_diag_edges}
-	Let $ABCD$ be a convex quadrilateral on the plane.
-	Then $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
-	that is at least one diagonal is greater than at least one side.
-\end{lemma}
-
-\begin{proof}
-	Suppose the contrary.
-	Then $AC$ in the minimal side of the triangle $ABC$ and, thus, the angle $ABC$ is acute.
-	So are also angles $BCD$, $CDA$ and $DAB$.
-	We conclude that all angles of a convex quadrilateral $ABCD$ are acute, which is impossible.
-	The contradiction concludes the proof.
-\end{proof}
-
-\subsection{Special results}
-
-Now we are ready to prove the main theorem of the section.
-
-\begin{theorem}
-	\label{thm:main_estimate}
-	If $n\geq 4$, then $d(2,n) \geq \gamma (n - 2)$,
-	where
-	\begin{multline}
-		0.46530... =
-		3^{1/4} \cdot 2^{-3/2} >
-		\\ >
-		\gamma = \frac{\sqrt{16 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}+\frac{14791716}{{{\left( 2593 \sqrt{647}-5823\right) }^{2}}}}-\frac{3846}{2593 \sqrt{647}-5823}}{8 {{\left( \sqrt{\frac{2}{\sqrt{3}}+\frac{1}{21490}}+\frac{1}{\sqrt{21490}}\right) }^{2}}}
-		>\\
-		> 0.45557
-		> \frac{5}{11}
-		.
-	\end{multline}
-\end{theorem}
-
-\begin{proof}
-	For $4 \leq n \leq 21491$, the assertion of the theorem follows immediately from Proposition~\ref{obs:4_leq_n_leq_21491}.
-	Let us consider $M\in \mathfrak{M}(2,n)$, $n \geq 21492$, $\operatorname{diam} M = b$.
-	Lemma~\ref{lem:square_container} yields that $M$ is situated in a square of side length $b$.
-	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1 M_2|$ and $|M_3 M_4|$
-	are minimal in $M$ ($M_2$ and $M_3$ may coincide).
-	Then $|M_1 M_2| \leq b\varphi_{n-1}$ and $|M_3 M_4| \leq b\varphi_{n-1}$.
-	Due to Lemma~\ref{lemma:quadr_diag_edges}, open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect
-	(otherwise they are not minimal).
-
-	Let $C = cr(M_1 M_2) \cap cr(M_3 M_4)$.
-	Each point $N\in M$ satisfies one of the following conditions:
-
-	a) $N$ belongs to $C$; overall at most $\gamma_2 b + 6$ points by Lemma~\ref{lem:intervals_cross} and Proposition~\ref{obs:estimate_points_on_straight_line};
-
-	b) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas
-	with one of $|M_3 M_4| - 1$ hyperbolas; overall at most $4 (|M_1 M_2| - 1)(|M_3 M_4| - 1)$ points;
-
-	c) $N$ belongs to the intersection of one of $|M_1 M_2| - 1$ hyperbolas with $cr(M_3 M_4)$;
-	overall at most $4 (|M_1 M_2| - 1)$ points;
-
-	d) $N$ belongs to the intersection of one of $|M_3 M_4| - 1$ hyperbolas with $cr(M_1 M_2)$;
-	overall at most $4 (|M_3 M_4| - 1)$ points;
-	\\
-	(see~\cite{erdos1945integral} for details).
-	Summing up the above cases, we obtain the following estimate:
-	\begin{equation}
-		\label{eq:estimate_n_varphi_n_squared}
-		n \leq 4 b^2 \varphi_{n-1}^2 - 4 + \gamma_2 b + 6
-		.
-	\end{equation}
-	Proposition~\ref{varphi_n_where_n_geq_21492} turns estimate~\eqref{eq:estimate_n_varphi_n_squared}
-	into the following:
-	\begin{equation}
-		\label{eq:estimate_n_sqrt_n-2}
-		n-2 \leq 4 b^2 \frac{\beta^2}{n-2} + \gamma_2 b
-		,
-	\end{equation}
-	which obviously leads to inequality
-	\begin{equation}
-		1 \leq 4\beta^2 \left(\frac{b}{n-2}\right)^2  + \gamma_2 \frac{b}{n-2}
-		.
-	\end{equation}
-	Let us denote $\lambda = b/(n-2)$ and solve the following quadratic inequality for $\lambda$:
-	\begin{equation}
-		\label{eq:square_inequality_lambda}
-		4\beta^2 \lambda^2  + \gamma_2 \lambda - 1 \geq 0
-		.
-	\end{equation}
-	The discriminant is $\gamma_2^2 + 16 \cdot \beta^2$,
-	so we obtain the following estimate:
-	\begin{equation}
-		\frac{b}{n-2} = \lambda \geq \frac{-\gamma_2 + \sqrt{\gamma_2^2 + 16 \cdot \beta^2}}{8\beta^2}
-		.
-	\end{equation}
-	Calculating the expression above for our $\beta$ and $\gamma_2$, we conclude the proof.
-\end{proof}
-
-\begin{corollary}
-	If $n\geq 4$, then $d(2,n) > \frac{5}{11} n$.
-\end{corollary}
-
-\begin{proof}
-	For $4 \leq n \leq 21491$, the claim follows from Proposition~\ref{obs:4_leq_n_leq_21491} immediately.
-	For $n > 21491$, the inequality $0.45557(n-2) > \frac{5}{11}n$ holds.
-\end{proof}
-
-We can improve the constant in Theorem~\ref{thm:main_estimate},
-if some new exact values of $d(2,n)$ are found;
-however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$ from above.
-More precisely, we have the following theorem.
-
-\begin{theorem}
-	For every $\varepsilon > 0$ there exists a number $n_0$ such that the inequality
-	\begin{equation}
-		d(2,n) \geq n\cdot(3^{1/4}\cdot 2^{-3/2} - \varepsilon)
-	\end{equation}
-	holds for every $n>n_0$,
-	that is
-	\begin{equation}
-		\lim_{n\to\infty} \frac{d(2,n)}{n} \geq 3^{1/4}\cdot 2^{-3/2}
-		.
-	\end{equation}
-\end{theorem}
-
-\section{Integral point sets with distance 1}
-
-\subsection{Constructing planar integral point sets with distance 1}
-
-\begin{definition}
-	A set $M\in\mathfrak{M}(2,n)$ is called \textit{facher}
-	if $M$ consists of $n-1$ points on a straight line
-	and one point out of the line.
-\end{definition}
-
-\begin{definition}
-	A set $M\in\mathfrak{M}(m,n)$ is called \textit{optimal}
-	if $\operatorname{diam}M=d(m,n)$.
-\end{definition}
-
-\begin{definition}
-	\cite{kurz2005characteristic}
-	A squarefree number $q$ is called the \textit{characteristic} of $M\in\mathfrak{M}(2,\mathbb{N})$,
-	if for any points $M_1, M_2, M_3 \in M$ the area of triangle $M_1 M_2 M_3$
-	is $p_{1,2,3}\sqrt{q}$ for some rational $p_{1,2,3}$.
-\end{definition}
-For a given $M\in\mathfrak{M}(2,\mathbb{N})$, the characteristic is determined uniquely.
-
-Facher sets are the simplest planar integral point sets.
-It is known that for $9\leq n \leq 122$ all the optimal sets are facher~\cite{kurz2008minimum}.
-For every cardinality $n$ and every squarefree number $q$
-there exists a facher set $M\in\mathfrak{M}(2,n)$ with characteristic $q$~\cite[Theorem 5]{our-vmmsh-2018-translit}.
-In~\cite{antonov2008maximal}, the facher sets of characteristic 1 were investigated; they were called \textit{semi-crabs}.
-
-For every integer $n\geq 3$ Solymosi presented~\cite{solymosi2003note} a construction of a facher integral point set
-$M\in\mathfrak{M}(2,n)$
-such that equality $|M_1 M_2| = 2$ holds for some $M_1, M_2 \in M$.
-The constructed set has both odd and even distances.
-
-Now we improve Solymosi's result.
-
-\begin{construction}
-	\label{con:planar_set_with_minimeter_1}
-	Let us choose a positive integer $k > 1$ and set
-	\begin{equation}
-		a = 2^{2^k} - 1
-		.
-	\end{equation}
-	Then
-	\begin{equation}
-		\label{eq:a_equiv_3_mod_4}
-		a \equiv 3 \mod 4
-	\end{equation}
-	and, moreover,
-	\begin{multline}
-		a = \left(2^{2^{k-1}}\right)^2 - 1
-		=
-		\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-1}} - 1\right)
-		=
-		\\=
-		\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-2}} + 1\right) \cdot ... \cdot \left(2^{2^1} + 1\right) \left(2^2 - 1\right)
-		.
-	\end{multline}
-	We set $d_j = 2^{2^j} + 1$ for $1 \leq j \leq k-1$.
-	Then $d_j \equiv 1 \mod 4$.
-
-	Let $c_J = \prod_{j\in J} d_j$ for every subset of indices $J\subset I = \{1,2,...,k-1\}$
-	(and  $c_\varnothing = 1$).
-	We obtain
-	\begin{equation}
-		\label{eq:c_J_equiv_1_mod_4}
-		c_J\equiv 1 \mod{4}
-	\end{equation}
-	and, moreover, $a$ is divisible by $c_J$.
-
-	Statements~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4} yield $a/c_J \equiv 3 \mod 4$.
-	Let $b_J = (c_J - a/c_J)/2$, then $b_J \equiv 1 \mod 2$.
-	Let us further take $g_J = (c_J + a/c_J)/2$, then $g_J \equiv 0 \mod 2$.
-
-	Next we define the coordinates of the points as following:
-	\begin{equation}
-		M_{J\pm} =\left(\pm\frac{b_J}{2}, 0\right)
-		,
-		~~
-		N   =\left(0, \frac{\sqrt{a}}{2}\right)
-		.
-	\end{equation}
-	Then the distances are:
-	\begin{multline}
-		|N M_{J\pm}|
-		=
-		\left(\frac{b_J^2}{4} + \frac{a}{4}\right)^{1/2}
-		=
-		\frac{1}{2}\left(\left(\frac{c_J - a/c_J}{2}\right)^2 + a\right)^{1/2}
-		=
-		\\=
-		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 - \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2 + a\right)^{1/2}
-		=
-		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 + \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2    \right)^{1/2}
-		=
-		\\=
-		\frac{1}{2}\left(\left(\frac{c_J + a/c_J}{2}\right)^2\right)^{1/2}
-		=
-		\frac{1}{2}\left(\frac{c_J + a/c_J}{2}\right)
-		=
-		\frac{g_J}{2}
-		\in\mathbb{N}
-		,
-	\end{multline}
-	\begin{equation}
-		|M_{J_1 \pm}  M_{J_2 \pm}|
-		=
-		\left|\frac{b_{J_1}}{2} \pm \frac{b_{J_2}}{2}\right|
-		=
-		\left|\frac{b_{J_1} \pm b_{J_2}}{2}\right|
-		\in\mathbb{N}
-		.
-	\end{equation}
-	In particular, for $H = \{k-1\}$ we obtain $C_H = 2^{2^{k-1}}+1$ and
-	\begin{equation}
-		b_H =
-		\left( 2^{2^{k-1}}+1 - \frac{a}{2^{2^{k-1}}+1} \right)/2
-		=
-		\left(2^{2^{k-1}}+1 - \left(  2^{2^{k-1}}-1 \right) \right)/2
-		=
-		1
-		.
-	\end{equation}
-	Thus, one of the distances is
-	\begin{equation}
-		|M_{H+}  M_{H-}|
-		=
-		\left|\frac{b_{H}}{2} - \frac{-b_{H}}{2}\right|
-		=
-		\left|\frac{1}{2} - \frac{-1}{2}\right|
-		= 1
-		.
-	\end{equation}
-
-	Note that all the points $M_{J\pm}$ are distinct:
-	the equality $b_J =  b_K$ implies $J=K$;
-	the equality $b_J = -b_K$ implies $c_J = -c_K$ or $c_J = a / c_K$.
-	The first case is impossible because both $c_J$ and $c_K$ are positive;
-	the second case contradicts~\eqref{eq:a_equiv_3_mod_4} and~\eqref{eq:c_J_equiv_1_mod_4}.
-
-	So, $M = \{ M_{J\pm}, N\}$ is indeed a planar integral point set of $2^k+1$ points,
-	and distance 1 occurs in $M$.
-	Since $k$ can be taken arbitrary large, it follows that we can construct
-	a planar integral point set of arbitrary large cardinality so that the distance 1 occurs in that set.
-\end{construction}
-
-\begin{remark}
-	\label{rem:planar_set_with_minimeter_1_of_3_points}
-	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ ad litteram,
-	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then
-	obtain an equilateral triangle of side length 1.
-	This triangle is obviously the optimal set in $\mathfrak{M}(2,3)$.
-\end{remark}
-
-\begin{remark}
-	\label{rem:planar_set_with_minimeter_1_of_4_and_5_points}
-	For $k=2$ in Construction~\ref{con:planar_set_with_minimeter_1},
-	we obtain the optimal set in $\mathfrak{M}(2,5)$ presented in~\cite[Fig. 1]{harborth1993upper}.
-	If we remove one point from it,
-	then we get one of the two optimal sets in $\mathfrak{M}(2,4)$.
-\end{remark}
-
-\subsection{Description of planar integral point sets with distance 1}
-
-%Now we need two more definitions.
-\begin{definition}
-	\cite{antonov2008maximal}
-	A set $M\in \mathfrak{M}(m,n)$ is \textit{maximal},
-	if there is no set $M'\in \mathfrak{M}(m,n+1)$
-	such that $M \subsetneq M'$.
-\end{definition}
-
-%\begin{definition}
-%	A set $M\in \mathfrak{M}(m,n)$ is \textit{submaximal},
-%	if there is only one maximal integral point set $M'\in \mathfrak{M}(m,n+1)$,
-%	such that $M \subset M'$.
-%\end{definition}
-
-In order to describe all the sets $M \in \mathfrak{M}(2,n)$ with distance 1,
-we need~\cite[Proposition 6]{our-vmmsh-2018-translit}.
-For the reader's convenience, we will state a slightly rephrased version of it and provide the proof.
-\begin{lemma}
-	\label{lem:no_4_points_in_semigeneral_position_with_distance_1}
-	Let $M=\{M_1,M_2,M_3,M_4\}\in \mathfrak{M}(2,4)$ such that $|M_1 M_2|=1$.
-	If $l$ is the straight line through points $M_1$ and $M_2$,
-	then one of the points $M_3$ or $M_4$ belongs to $l$.
-\end{lemma}
-
-\begin{proof}
-	Let $m$ denote the perpendicular bisector of line segment $M_1 M_2$.
-	Due to the triangle inequality, there are two possibilities to place each of points $M_i$, $i=3,4$:
-
-	a) $M_i$ belongs to $l$ and $|M_i M_1|-|M_i M_2| = \pm 1$;
-
-	b) $M_i$ belongs to $m$ and $|M_i M_1| = |M_i M_2|$.
-
-	If (a) is true for both points, then $M\subset l$ and thus $M$ is not an integral point set.
-	If (a) is true for one point and (b) is true for another, then the claim of the lemma follows.
-	So, suppose to the contrary, that both points $M_3$ and $M_4$ belong to $m$.
-
-	The area of the triangle $M_1 M_3 M_4$ is rational because $|M_3 - M_4| \in \mathbb{Z}$.
-	Thus, the characteristic of $M$ is 1,
-	and there is a Cartesian coordinate system such that $M_1=(-1/2,0)$, $M_2=(1/2,0)$, $M_3=(0, a/2)$
-	(see~\cite[Theorem 4]{our-vmmsh-2018-translit}).
-	It is clear that $a\neq 0$.
-	We set $b = |M_1 - M_3|$ and $O=(0,0)$.
-	Applying the Pythagorean theorem to triangle $OM_1M_3$, we obtain the following Diophantine equation:
-	\begin{equation}
-		\frac{1}{4} + \frac{a^2}{4} = b^2
-		,
-	\end{equation}
-	or, equivalently,
-	\begin{equation}
-		1 + a^2 = (2b)^2
-		.
-	\end{equation}
-	This equation has no integral solutions.
-	This contradiction concludes the proof.
-\end{proof}
-
-Using Construction~\ref{con:planar_set_with_minimeter_1},
-Lemma~\ref{lem:no_4_points_in_semigeneral_position_with_distance_1} and Corollary~\ref{cor:only_one_distance_1_on_straight_line}
-together with results of~\cite[Section 6]{antonov2008maximal},
-we obtain the following theorem.
-
-\begin{theorem}
-	\label{thm:minimeter_1_planar}
-	For every $n\geq 3$ there is a planar integral set $M$ of $n$ points
-	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds.
-	This set consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.
-
-	And vice versa, if $M$ is a planar integral point set of $n$ points
-	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
-	then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
-	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
-	There is only one maximal integral point set $M' \supseteq M$,
-	and the perpendicular bisector is the axis of symmetry for $M'$.
-	Moreover, if $n > 3$, then the distance 1 occurs in $M$ (and $M'$) only once.
-\end{theorem}
-
-\subsection{Higher dimensions}
-
-To build integral points sets with distance 1 in higher dimensions,
-we need the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds}.
-
-\begin{construction}
-	\label{con:blowing_up}
-	Let $M\in\mathfrak{M}(2,n)$ have many points on a straight line $l_1$
-	and $M\setminus l_1 \subset l_2$ for a straight line $l_2$.
-	Moreover, let $l_1$ be parallel to $l_2$.
-
-	Construct $M' \subset \mathbb{R}^m$ by replacing all the points $M_1,...,M_n\in M \cap l_2$
-	by equal $(m-2)$--dimensional simplices $S_1,...,S_n$
-	in such a way that
-	\begin{equation}
-		\forall(j=1,...,n) \forall(N_i \in M \cap l_1 )\forall\left(M^j_k, M^j_h \in S_j\right)\left[|N_i M^j_k| = |N_i M^j_h|\right].
-	\end{equation}
-	Then under certain conditions (in particular, if $\#(M \cap l_2)=1$ that we actually need)
-	we get $M'\in\mathfrak{M}(m,\mathbb{N})$.
-
-	For further examples we also refer the reader to~\cite{avdeev2019particular}.
-\end{construction}
-
-Now we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,
-with distance 1 occurring in it.
-If we have an integral point set with distance 1 occurring,
-then we can easily dilate it to turn 1 into the desired distance.
-These facts, together with Theorem~\ref{thm:minimeter_1_planar}, give the following theorem.
-
-\begin{theorem}
-	For arbitrary integers $m \geq 2$, $n \geq m+1$, $d \geq 1$
-	there exists $M\in\mathfrak{M}(m,n)$,
-	such that for some $M_1, M_2\in M$ equality $|M_1 M_2| = d$ holds.
-\end{theorem}
-
-\begin{definition}
-	\cite{noll1989nclusters}
-	We will call an integral point set $M$ \textit{prime}, if the greatest common divisor
-	of all the distances occurring in $M$ is 1.
-\end{definition}
-
-If an integral point set is prime,
-then it cannot be squashed to an integral point set of the same power and structure but smaller diameter.
-
-\begin{theorem}
-	\label{thm:prime_with_desired_distance}
-	For every $m \geq 3$, $n \geq m+1$, $d \geq 1$ there exists a \textbf{prime}
-	integral point set $M\in\mathfrak{M}(m,n)$ which
-	contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
-\end{theorem}
-
-\begin{proof}
-	Take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
-	with sufficiently large height and then apply the ``blowing up'' procedure described in Construction~\ref{con:blowing_up},
-	replacing the point that is not situated on the line, with an $(m-2)$-dimensional simplex of proper side length.
-	That simplex will consist of $m-1$ points;
-	we can exclude practically all $M_J$ points, except the two points with distance 1.
-	So, the set consists of $m+1$ or more points and has distance 1 occurring in it.
-	As distance 1 occurs in the obtained set, it is prime.
-\end{proof}
-
-A slightly stronger result can also be claimed.
-
-\begin{theorem}
-	\label{thm:prime_with_unique_desired_distance}
-	For arbitrary integers $m > 2$, $n \geq m+1$, $d \geq 1$
-	there exists a prime set $M\in\mathfrak{M}(m,n)$,
-	such that for some $M_1, M_2\in M$ equality $|M_1 M_2| = d$ holds,
-	the distance $d$ is minimal and occurs in $M$ only once.
-\end{theorem}
-
-\begin{proof}
-	We apply Construction~\ref{con:planar_set_with_minimeter_1} to obtain $M'\in \mathfrak{M}(2,n-m+2)$
-	with $M'_1, M'_2 \in M'$ and $|M'_1 M'_2| = 1$.
-	Dilate $M'$ to $M''$ in such a way that $M'_1$ and $M'_2$ turn into $M''_1$ and $M''_2$ resp.
-	and $|M''_1 M''_2| = d$.
-	Then ``blow up'' $M''$ to $M\in \mathfrak{M}(m,n)$ using a simplex of side length $d+1$.
-	The greatest common divisor of $d$ and $d+1$ is $1$, so $M$ is prime.
-\end{proof}
-
-\section{Final remarks and open problems}
-
-\begin{remark}
-	The bound of Theorem~\ref{thm:main_estimate} is not tight.
-	The derivation of tight bounds for the minimum diameter $d(2, n)$
-	is still a challenging task for the forthcoming investigation~\cite[Section 7]{kurz2008minimum}.
-\end{remark}
-The known upper bound is $d(2,n)\leq 2^{c \log n \log \log n}$~\cite{harborth1993upper}.
-However, we hope that the present article can provide a framework for possible better estimates.
-
-\begin{conjecture}
-	The approach of Theorem~\ref{thm:main_estimate} can be generalized to higher dimensions.
-\end{conjecture}
-For the overview of current lower bounds for higher dimensions, we refer the reader to~\cite{nozaki2013lower}.
-
-There are also several ways to improve the current bound,
-based on Lemma~\ref{lem:square_container}.
-
-\begin{problem}
-	\label{prb:square_side_length_and_diameter_power}
-	Is there a number $\delta < 1$ such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	is situated in a square of side length $d^\delta$?
-\end{problem}
-If the answer for Problem~\ref{prb:square_side_length_and_diameter_power} is affirmative,
-then we can obtain a bound better than linear.
-
-\begin{problem}
-	\label{prb:square_side_length_and_diameter_linear}
-	What are the minimal numbers $\delta \leq 1$ and $h \leq 0$
-	such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	is situated in a square of side length $\delta d + h$?
-\end{problem}
-
-If the answer for Problem~\ref{prb:square_side_length_and_diameter_linear} is not $\delta = 1$ and $h = 0$,
-then the linear bound can be slightly improved.
-
-However, Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}
-do not seem to show the way for improving the bound
-due to the following well-known theorem,
-which can be proved by an application of circle inversion~\cite{solymosi2010question}.
-\begin{theorem}
-	\label{thm:rational_set_on_circle}
-	There exists a dense subset $P$ of the unit circle such that for any $P_1, P_2 \in P$
-	the distance $|P_1 P_2|$ is rational.
-\end{theorem}
-We conjecture that thorough investigation of existing examples of circular integral point sets
-~\cite{bat2018number,harborth1993upper,piepmeyer1996maximum}
-will lead to the negative answers for the
-Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}.
-
-Another possibility to increase the constant in our lower bound is introduced by the following problem.
-
-\begin{problem}
-	\label{prb:minimal_shape}
-	What is the minimal shape $\mathcal{S}$ such that any $M\in\mathfrak{M}(2,\mathbb{N})$ with $\operatorname{diam} M = d$
-	is situated in  $\mathcal{S}$?
-\end{problem}
-This problem is another generalization of Lemma~\ref{lem:square_container}.
-We know that $\mathcal{S}$ is not a circle of diameter $d$
-(the counterexample is an equilateral triangle of side length $d$)
-and we conjecture that $\mathcal{S}$ is not a Reuleaux triangle of width $d$
-(the possible counterexample belongs to a circle of diameter $d$, see Theorem~\ref{thm:rational_set_on_circle}).
-
-The solution to Problem~\ref{prb:minimal_shape} (if ever found) will lead to a sophisticated packing problem,
-something like ``How to put $n$ points in a convex hull of concentric circle and Reuleaux triangle?''
-Obviously, Problem~\ref{prb:minimal_shape} can be generalized to higher dimensions.
-
-For $n=3,4,5$, there are optimal sets in $\mathfrak{M}(2,n)$ that contain distance 1
-(see Remarks~\ref{rem:planar_set_with_minimeter_1_of_3_points} and~\ref{rem:planar_set_with_minimeter_1_of_4_and_5_points}).
-However, we can suggest the following conjecture.
-\begin{conjecture}
-	\label{con:no_optimal_with_edge_1}
-	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,
-	is not optimal.
-\end{conjecture}
-
-The motivation for Conjecture~\ref{con:no_optimal_with_edge_1} is based on the results of~\cite[Section 5]{kurz2008minimum},
-especially the following theorem.
-
-\begin{theorem}
-	For every $12 \leq n \leq 122$, there exist a facher optimal sets $M_n\in\mathfrak{M}(2,n)$
-	that consist of $n-1$ points $\{ (b_1,0), ..., (b_{n-1},0)\}$
-	and the point $(0,\sqrt{A_n})$,
-	where $b_1,...,b_{n-1}, A_n$ are integer and $\sqrt{A_n}\notin\mathbb{N}$.
-\end{theorem}
-Such a set cannot contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar},
-the points with distance 1 should be $(\pm 1/2, 0)$.
-
-For planar integral point sets containing distance 1, we can emphasize the following problem.
-\begin{problem}
-	For given $n\geq 3$ and squarefree number $q \neq 1$,
-	is there $M\in\mathfrak{M}(2,n)$, such that characteristic of $M$ is $q$
-	and distance 1 occurs in $M$?
-\end{problem}
-
-The consideration of integral point sets containing distance 1 in higher dimensions gives a rise to another problem.
-
-\begin{problem}
-	How many times can the distance 1 occur in a set $M\in\mathfrak{M}(m,n)$?
-\end{problem}
-The answer for $m=2$ is given by Theorem~\ref{thm:minimeter_1_planar}.
-For greater $m$, we will be bold enough to suggest the following conjecture.
-\begin{conjecture}
-	Consider $M\in\mathfrak{M}(m,n)$, $M_1,M_2\in M$, $|M_1 M_2| =1$.
-	Then either $n = m+1$ or $M$ is a ``blowup'' of a facher set and distance 1 occurs in $M$
-	no more than $1+\frac{(m-1)^2 - (m-1)}{2}$ times.
-\end{conjecture}
-Such an estimate is based on the number of edges in an $(m-2)$-dimensional simplex.
-
-We want to conclude the list of conjectures with the one which appears the easiest to deal with.
-
-\begin{conjecture}
-	\label{hyp:prime_planar}
-	For every $n \geq 3$, $d \geq 1$ there exists a \textbf{prime}
-	set $M\in\mathfrak{M}(2,n)$ which
-	contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
-\end{conjecture}
-This conjecture gives the same claim as Theorem~\ref{thm:prime_with_desired_distance} does, but for $m=2$.
-Obviously, Theorem~\ref{thm:prime_with_unique_desired_distance} can be generalized (for $m=2$) into a similar conjecture.
-
-\section{Acknowledgements}
-The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
-A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
-Dr.~A.S.~Usachev for proofreading,
-and E.A. Momot for the ``SciLexic'' project that helped the author with some word usage issues.
-Also the author would like to owe a special thanks to the anonymous referees
-for their helpful comments and suggestions on the paper.
-
-%\printbibliography
-
-%\printbibitembibliography
-
 \begin{thebibliography}{99}
@@
 %
-\bibitem{our-mz-rus-translit}
-N. N. Avdeev and E. M. Semenov, On point sets in the plane with integral distances (О множествах точек на плоскости с
-целочисленными расстояниями), \emph{Matematicheskie zametki} 100.\textbf{5} (2016), 757–761.
-%
 \bibitem{our-vmmsh-2018-translit}
@@
 %
-\bibitem{bat2018number}
-G. Bat-Ochir, On the number of points with pairwise integral distances on a circle, \emph{Discrete Appl. Math.} 254 (2018),
-17–32, \textsc{doi}: \href {https://doi.org/10.1016/j.dam.2018.07.004} {\nolinkurl {10.1016/j.dam.2018.07.004}}.
+\bibitem{my-semi-general-5-4-bound-2019}
+N. Avdeev, On diameter bounds for planar integral point sets in semi-general position (2019), arXiv: \href
+{http://arxiv.org/abs/1907.09331} {\nolinkurl {1907.09331}}, \textsc{url}: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190709331A/abstract}.%
+\bibitem{my-pps-linear-bound-2019}
+N. Avdeev, On existence of integral point sets and their diameter bounds, \emph{Australas. J. Combin.} 77.\textbf{1} (2020),
+100–116, arXiv: \href {http://arxiv.org/abs/1906.11926} {\nolinkurl {1906.11926}}, \textsc{url}: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190611926A/abstract}.
 %
@@
 %
-\bibitem{garibaldi2005erdos}
-J. Garibaldi and A. Iosevich, The Erdős distance problem: lecture notes (2005).%
 \bibitem{harborth1993upper}
@@
 {10.1007/bf02189331}}.
+%
+\bibitem{huff1948diophantine}
+G. B. Huff, Diophantine problems in geometry and elliptic ternary forms, \emph{Duke Math. J.} 15.\textbf{2} (1948), 443–453,
+\textsc{doi}: \href {https://doi.org/10.1215/S0012-7094-48-01543-9} {\nolinkurl {10.1215/S0012-7094-48-01543-9}}.
+%
+\bibitem{kemnitz1988punktmengen}
+A. Kemnitz, Punktmengen mit ganzzahligen Abständen, 1988.
+%
+\bibitem{kreisel2008heptagon}
+T. Kreisel and S. Kurz, There are integral heptagons, no three points on a line, no four on a circle, \emph{Discrete Comput.
+Geom.} 39.\textbf{4} (2008), 786–790, \textsc{doi}: \href {https://doi.org/10.1007/s00454-007-9038-6} {\nolinkurl
+{10.1007/s00454-007-9038-6}}.
 %
@@
 %
-\bibitem{locatelli2002packing}
-M. Locatelli and U. Raber, Packing equal circles in a square: a deterministic global optimization approach, \emph{Discrete Appl.
-Math.} 122.\textbf{1-3} (2002), 139–166.
+\bibitem{kurz2013constructing}
+S. Kurz et al., Constructing $7$-clusters, \emph{Serdica J. Comput.} 8.\textbf{1} (2014), 47–70, arXiv: \href
+{http://arxiv.org/abs/1312.2318} {\nolinkurl {1312.2318}}.
 %
-\bibitem{noll1989nclusters}
-L. C. Noll and D. I. Bell, {$n$}-clusters for {$1<n<7$}, \emph{Math. Comp.} 53.\textbf{187} (1989), 439–444.
+\bibitem{markot2005newverified}
+M. C. Markót and T. Csendes, A new verified optimization technique for the "packing circles in a unit square" problems,
+\emph{SIAM J. Optim.} 16.\textbf{1} (2005), 193–219.
+%
+\bibitem{momot2022example}
+E. A. Momot, R. E. Zvolinskiy, and A. E. Zvolinskiy, An example of rails integral point set with cardinality 106, Прикладная
+математика и фундаментальная информатика, 2022, 114–115.
 %
@@
 H. Nozaki, Lower bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 56 (2013), 139–143.
-%
-\bibitem{piepmeyer1996maximum}
-L. Piepmeyer, The maximum number of odd integral distances between points in the plane, \emph{Discrete Comput. Geom.}
-16.\textbf{1} (1996), 113–115, \textsc{doi}: \href {https://doi.org/10.1007/bf02711135} {\nolinkurl {10.1007/bf02711135}}.
 %
@@
 arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl {0806.3095}}.
+%
+\bibitem{our-ped-2018}
+Н. Н. Авдеев, On integral point sets in special position, \emph{Некоторые вопросы анализа, алгебры, геометрии и
+математического образования: материалы международной молодежной научной школы «Актуальные направления
+математического анализа и смежные вопросы»} 8 (2018), 5–6.%
+%
+\bibitem{our-pmm-2018}
+Н. Н. Авдеев, Об отыскании целоудалённых множеств специального вида, Актуальные проблемы прикладной математики,
+информатики и механики - сборник трудов Международной научной конференции. Научно-исследовательские публикации,
+2018, 492–498.
+%
+\bibitem{smurov1998stripcoverings}
+М. Смуров and А. Спивак, Покрытия полосками, \emph{Квант} \textbf{5} (1998), 6–12.
 \end{thebibliography}
-
-\end{document}

### 2024-10-09 char-4k-bound/char-4k-bound.tex
Switch to pregenerated references
[4k3] Last minor fixes

@@
 
-\usepackage[backend=biber,style=ajc2020unofficial]{biblatex}
-\addbibresource{../common/notmy.bib}
-\addbibresource{../common/my.bib}
+%\usepackage[backend=biber,style=ajc2020unofficial]{biblatex}
+%\addbibresource{../common/notmy.bib}
+%\addbibresource{../common/my.bib}
 
-%\usepackage{cite}
+\usepackage{cite}
 
@@
 	if no three points of $M$ are collinear.
-	Notation: $M\in\dot{\mathfrak{M}}$, and also $M\in\dot{\mathfrak{M}}_n$ for $n=\#M$.
+	Notation: $M\in\overline{\mathfrak{M}}$, and also $M\in\overline{\mathfrak{M}}_n$ for $n=\#M$.
 \end{definition}
@@
 	and no four points of $M$ are concircular.
-	Notation: $M\in\overline{\mathfrak{M}}$, and also $M\in\overline{\mathfrak{M}}_n$ for $n=\#M$.
+	Notation: $M\in\dot{\mathfrak{M}}$, and also $M\in\dot{\mathfrak{M}}_n$ for $n=\#M$.
 \end{definition}
@@
 		\pm\sqrt{k^2 - \left(\frac{m}{2}+\frac{a_i}{2m}\right)^2} =
-		\frac{\pm b_i \sqrt{q}}{2m},
+		\frac{\pm b_i \sqrt{p}}{2m},
 	\end{gather}
@@
 \begin{definition}
-	A planar integral point set situated on two parallel straight lines
+	A non-facher planar integral point set situated on two parallel straight lines
 	is called a \emph{rails} set.
@@
 
-\printbibliography
+%\printbibliography
+
+%\printbibitembibliography{}
+
+\input{references.tex}
 

### 2025-02-21 char-4k-bound/char-4k-bound.tex
[4k3] Unify the spelling of Erdős

@@
 
-	In 1945, Erdös gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
+	In 1945, Erdős gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
 	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
@@
 	That was the first lower bound for the diameter ever;
-	however, Erdös did not only estimate the diameter,
+	however, Erdős did not only estimate the diameter,
 	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolas.
@@
 	(in assumption that the set has at east 4 points).
-	All these lower bounds are based on the Erdos's framework --- cofocal hyperbolas.
+	All these lower bounds are based on the Erdős's framework --- cofocal hyperbolas.
 
@@
 
-	In the present paper, we put the power of Erdös's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
+	In the present paper, we put the power of Erdős's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
 	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
@@
 	\end{equation}
-	is called the \emph{$n$-th Erdos curve}.
+	is called the \emph{$n$-th Erdős curve}.
 \end{definition}
 
-Obviously, the 0-th Erdos curve is the perpendicular bisector of $M_1 M_2$,
-and all the other Erdos curves are branches of cofocal hyperbolas.
-On Figure~\ref{fig:Erdos_curves_numbered}, the Erdos curves are shown for an edge of length 3.
+Obviously, the 0-th Erdős curve is the perpendicular bisector of $M_1 M_2$,
+and all the other Erdős curves are branches of cofocal hyperbolas.
+On Figure~\ref{fig:Erdos_curves_numbered}, the Erdős curves are shown for an edge of length 3.
 
@@
 \center{\includegraphics[width=0.95\linewidth]{Erdos_curves_numbered.png}}
-\caption{Erdos curves}
+\caption{Erdős curves}
 \label{fig:Erdos_curves_numbered}
@@
 
-Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erdos curves.
-For the sake of brevity, Erdos curves with odd numbers are named odd Erdos curves,
-and the ones with even numbers are named even Erdos curves.
+Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erdős curves.
+For the sake of brevity, Erdős curves with odd numbers are named odd Erdős curves,
+and the ones with even numbers are named even Erdős curves.
 
@@
 %4
-\section{Erdös Curves and Characteristic}
+\section{Erdős Curves and Characteristic}
 
@@
 
-The following lemma is a classical Erdos-style intersection-enumeration one,
+The following lemma is a classical Erdős-style intersection-enumeration one,
 empowered by our Weeding Theorem I.
@@
 \section{Acknowledgements}
-The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
+The authors thank Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
 Dr. A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
 Dr.~A.S.~Usachev for proofreading,
-and E.A. Momot for the ``SciLexic'' project that helped the author with some word usage issues.
-%Also the author would like to owe a special thanks to the anonymous referees
-%for their helpful comments and suggestions on the paper.
+and E.A. Momot for the ``SciLexic'' project that helped the authors with some word usage issues.
+Also the authors would like to owe a special thanks to the anonymous referees
+for their helpful comments and suggestions on the paper.
 

### 2025-02-21 char-4k-bound/char-4k-bound.tex
[4k3] Implement the Referee's comments

@@
 In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological.
-Section 4 is devoted to the connection between Erdös curves and characteristic.
+Section 4 is devoted to the connection between Erdős curves and the characteristic.
 In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.
@@
 
-In this Section, we provide rigorous definitions and list some known results.
+In this section, we provide rigorous definitions and list some known results.
 

### 2025-02-22 char-4k-bound/char-4k-bound.tex
[4k3] Add a note about the asterisk notation
[4k3] Remove a misleading comma in 10 655
[4k3] Fix two remarks from the Referee

@@
 \end{definition}
+The characteristic of an IPS always exists and is unique.
 
@@
 	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
-	For $N\in\mathbb N_\pm$, $|n| < |M_1 M_2|$
-	we will say that the set of points
+	For an integer $n$, $|n| < |M_1 M_2|$,
+	we will call the set of points
 	\begin{equation}
@@
 	\end{equation}
-	is called the \emph{$n$-th Erdős curve}.
+	the \emph{$n$-th Erdős curve}.
 \end{definition}
@@
 Such notation is made possible by the Grid Theorem.
+Note that we use the asterisk instead of the centered dot to emphasize
+that we do not mean the usual coordinatewise multiplication;
+instead, the denominator is appended to both coordinates,
+but the square root multiplier is appended only to the second one.
 
@@
 	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
-	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10,655$.
+	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10 655$.
 	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.

### 2025-02-23 AvdMomZvol_rails/Article.tex
[rails] Typo fixes

@@
 	with $M_0$ being the center.
-	by the Law of sines, in the triangle $M_1 M_2 M_3$
+	By the Law of sines, in the triangle $M_1 M_2 M_3$
 	\begin{equation}
@@
 We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
-is weaker than the one given in~\eqref{eq:d_m_2m+2}.
+is weaker than the one given in~\eqref{eq:d_m_2m+2}).
 

### 2025-02-23 char-4k-bound/char-4k-bound.tex
[4k3] Restore the equation number (12)
[4k3] Fix a theorem number (thanks to the Referee)
[4k3] Omit the redundant condition of semi-general condition (thanks to the Referee)

@@
 \begin{lemma}
-	\cite[Theorem 3.10]{my-pps-linear-bound-2019}
+	\cite[Theorem 5.9]{my-pps-linear-bound-2019}
 	\label{lem:no_distance_one}
@@
 	\label{thm:weeding_I}
-	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$ be an IPS of semi-general position with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
+	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$ be an IPS with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
 Then, the length of the edge $M_1 M_2$ and its $n$-th Erdős curve, which contains the point $M_3$, have the same parity.
@@
 	\label{thm:weeding_II}
-	Let an IPS in semi-general position $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erdős curve of edge $M_1 M_2$.
-	Then the set $M$ has characteristic $p=8k+7$, $k\in \mathbb{N}_{0}$.
+	Let an IPS $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erdős curve of edge $M_1 M_2$.
+	Then $\operatorname{char} M=8k+7$, $k\in \mathbb{N}_{0}$.
 \end{theorem}
@@
 
-\begin{eqnarray}
+\begin{equation}
+	\begin{array}{rcl}
 		\label{d}
-\overline{d}(2,3)&=& 1,{}\nonumber\\
-\overline{d}(2,4)&=& 4,{}\nonumber\\
-\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{}\nonumber\\
-\overline{d}(2,7)&=& 33,{}\nonumber\\
-\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{}\nonumber\\
-\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{}\nonumber\\
-\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\nonumber\\
-\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{}\nonumber\\
-\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{}\nonumber\\
-\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{}\nonumber\\
-\overline{d}(2,28)&=& 10672,{}\nonumber\\
-\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{}\nonumber\\
+		\overline{d}(2,3)&=& 1,{}\\
+		\overline{d}(2,4)&=& 4,{} \\
+		\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{} \\
+		\overline{d}(2,7)&=& 33,{} \\
+		\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{} \\
+		\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{} \\
+		\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\\
+		\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{} \\
+		\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{} \\
+		\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{} \\
+		\overline{d}(2,28)&=& 10672,{} \\
+		\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{} \\
 		\overline{d}(2,37) &>& 20\ 000.
-\end{eqnarray}
+	\end{array}
+\end{equation}
 
-Now we are finally ready to proof our main result.
+Now we are finally ready to proove our main result.
 It improves the bound~\eqref{eq:chebsb_lower_bound}

### 2025-02-23 char-4k-bound/char-4k-bound.tex
[4k3] Restore the lost prime (thanks to the Referee)

@@
 	\label{lem:intersection_enumeration}
-	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}_n}$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$.
+	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}_n}'$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$.
 Then $\# M \le |M_1 M_2| \cdot |M_3 M_4| - 2$.

### 2025-02-24 AvdMomZvol_rails/Article.tex
[rails] Shortification, round one: P_7, P_8
[rails] Edits from Zvolinsky
[rails] Re-enable biblatex, the Long GOST style

--- 2021-11-22 AvdMomZvol_rails/Article.tex
@@
 \addbibresource{../common/my.bib}
+
+\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
+
+%\usepackage{cite}
 
@@
 \begin{figure}[htbp]
-	\includegraphics[width=.48\linewidth]{picture_d17.png}
-	\hfill
-	\includegraphics[width=.48\linewidth]{picture_d21.png}
-	\\
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 7 and diameter 17}
-	\label{picture_d17.png}}
-	\hfill
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 8 and diameter 21}
-	\label{picture_d21.png}}
+	\centering
+	\includegraphics[width=.5\linewidth]{picture_d21.png}
+	\\
+	\caption{PIPS $\mathcal{P}_8$ of cardinality 8 and diameter 21}
+	\label{picture_d21.png}
 \end{figure}
@@
 \item
-$\mathcal{P}=\sqrt{315}/{2} * \{ (13, 1),
+$\mathcal{P}_7 = \sqrt{315}/{2} * \{ (13, 1),
 (29, 1),
@@
 $
-(Figure~\ref{picture_d17.png})
-
-$f = 8$, $v = 6$, $w = 10$, $\operatorname{diam(\mathcal{P})} = 17$,
-
+,
+
+$f = 8$, $v = 6$, $w = 10$, $\operatorname{diam(\mathcal{P}_7)} = 17$,
 which gives $d(m, 2m + 3) \leq 17$.
@@
 \item
-$\mathcal{P}=\sqrt{315}/{2} * \{ (13, 1),
-(29, 1),
-(0, 0),
-(10, 0),
-(16, 0),
-(26, 0),
-(32, 0),
-(42, 0)\}
-$
-(Figure~\ref{picture_d21.png})
-
-$f = 8$, $v = 6$, $w = 10$, $\operatorname{diam(\mathcal{P})} = 21$,
-
+$\mathcal{P}_8 = \mathcal{P}_7 \cup \sqrt{315}/2 * \{(42, 0)\}
+$
+(Figure~\ref{picture_d21.png}),
+
+$f = 8$, $v = 6$, $w = 10$, $\operatorname{diam(\mathcal{P}_8)} = 21$,
 which gives $d(m, 2m + 4) \leq 21$.
@@
 \item
-$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
+$\mathcal{P}_9 = \sqrt{70}/{1} * \{ (\pm 44, 12),
 (-78 , 0),
@@
 $
-(Figure~\ref{picture_2.png})
-
-$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,
+,
+
+$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P}_9)} = 158$,
 
@@
 \item
-$\mathcal{P}=\sqrt{70}/{1} * \{ (\pm 44, 12),
-(\pm 78 , 0),
-(\pm 62, 0),
-(\pm 55 , 0),
-(\pm 10 , 0)\}
-$
-(Figure~\ref{picture_3.png}).
-
-$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P})} = 158$,
+$\mathcal{P}_{10} = \mathcal{P}_9 \cup \sqrt{70}/{1} * \{(78, 0)\}
+$
+(Figure~\ref{picture_3.png}),
+
+$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P}_{10})} = 158$,
 
@@
 \begin{figure}[h!]
-	\includegraphics[width=.48\linewidth]{picture_2.png}
-	\hfill
-	\includegraphics[width=.48\linewidth]{picture_3.png}
-	\\
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 9 and diameter 158}
-	\label{picture_2.png}}
-	\hfill
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 10 and diameter 158}
-	\label{picture_3.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
+	\centering
+	\includegraphics[width=.45\linewidth]{picture_3.png}
+	\\
+	\caption{PIPS $\mathcal{P}_{10}$ of cardinality 10 and diameter 158}
+	\label{picture_3.png}
+\end{figure}
+
+\item
+$\mathcal{P}_{11} = \sqrt{70}/{2} * \{ (\pm 99, 24),
 (\pm 207 , 0),
@@
 $
-(Figure~\ref{picture_4.png}).
-
-$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P})} = 252$,
+,
+
+$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P}_{11})} = 252$,
 
@@
 
+\item
+$\mathcal{P}_{12} = \mathcal{P}_{11} \cup \sqrt{70}/{2} * \{(297, 0)\}
+$
+(Figure~\ref{picture_5.png}),
+
+$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P}_{12})} = 297$,
+
+which gives $d(m, 2m + 8) \leq 297$.
+
 \begin{figure}[htbp]
-	\includegraphics[width=.48\linewidth]{picture_4.png}
-	\hfill
-	\includegraphics[width=.48\linewidth]{picture_5.png}
-	\\
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 11 and diameter 252}
-	\label{picture_4.png}}
-	\hfill
-	\parbox{.48\linewidth}{\caption{PIPS of cardinality 12 and diameter 297}
-	\label{picture_5.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{70}/{2} * \{ (\pm 99, 24),
-(\pm 297 , 0),
-(\pm 207 , 0),
-(\pm 145 , 0),
-(\pm 63 , 0),
-(\pm 25 , 0)\}
-$
-(Figure~\ref{picture_5.png}).
-
-$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P})} = 297$,
-
-which gives $d(m, 2m + 8) \leq 297$.
-
-\item
-$\mathcal{P}=\sqrt{19019}/{2} * \{ (\pm 200, 3),
+	\centering
+	\includegraphics[width=.5\linewidth]{picture_5.png}
+	\\
+	\caption{PIPS $\mathcal{P}_{12}$ of cardinality 12 and diameter 297}
+	\label{picture_5.png}
+\end{figure}
+
+\item
+$\mathcal{P}_{13} = \sqrt{19019}/{2} * \{ (\pm 200, 3),
 (\pm 873 , 0),
@@
 $
-(Figure~\ref{picture_6.png}).
-
-$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P})} = 1073$,
+,
+
+$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P}_{13})} = 1073$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{picture_6.png}}
-\parbox{0.7\linewidth}{\caption{PIPS of cardinality 13 and diameter 1073}
-\label{picture_6.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{19019}/{2} * \{ (\pm 200, 3),
-(\pm 1273 , 0),
-(\pm 873 , 0),
-(\pm 615 , 0),
-(\pm 377 , 0),
-(\pm 215 , 0),
-$
-
-$
-(\pm 23 , 0)\}
-$
-(Figure~\ref{picture_7.png}).
-
-$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P})} = 1273$,
+\item
+$\mathcal{P}_{14} = \mathcal{P}_{13} \cup \sqrt{19019}/{2} * \{(1273, 0)\}
+$
+(Figure~\ref{picture_7.png}),
+
+$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P}_{14})} = 1273$,
 
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{picture_7.png}}
-\parbox{0.7\linewidth}{\caption{PIPS of cardinality 14 and diameter 1273}
+\center{\includegraphics[width=0.6\linewidth]{picture_7.png}}
+\parbox{0.7\linewidth}{\caption{PIPS $\mathcal{P}_{14}$ of cardinality 14 and diameter 1273}
 \label{picture_7.png}}
@@
 \item
-$\mathcal{P}=\sqrt{385}/{2} * \{ (\pm 1105, 48),
+$\mathcal{P}_{15} = \sqrt{385}/{2} * \{ (\pm 1105, 48),
 (\pm 1587 , 0),
@@
 $
-(Figure~\ref{picture_8.png}).
-
-$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P})} = 1888$,
+,
+
+$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P}_{15})} = 1888$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.75\linewidth]{picture_8.png}}
-\parbox{0.75\linewidth}{\caption{PIPS of cardinality 15 and diameter 1888}
-\label{picture_8.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{385}/{2} * \{ (\pm 1105, 48),
-(\pm 2189 , 0),
-(\pm 1587 , 0),
-(\pm 1269 , 0),
-(\pm 763 , 0),
-(\pm 623 , 0),
-$
-
-$
-(\pm 529 , 0),
-(\pm 339 , 0)\}
-$
-(Figure~\ref{picture_9.png}).
-
-$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P})} = 2189$,
+\item
+$\mathcal{P}_{16} = \mathcal{P}_{15} \cup \sqrt{385}/{2} * \{(2189, 0)\}
+$
+(Figure~\ref{picture_9.png}),
+
+$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P}_{16})} = 2189$,
 
@@
 \center{\includegraphics[width=0.75\linewidth]{picture_9.png}}
-\parbox{0.75\linewidth}{\caption{PIPS of cardinality 16 and diameter 2189}
+\parbox{0.75\linewidth}{\caption{PIPS $\mathcal{P}_{16}$ of cardinality 16 and diameter 2189}
 \label{picture_9.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
+$\mathcal{P}_{17} = \sqrt{154}/{1} * \{ (\pm 874, 60),
 (\pm 1376 , 0),
@@
 $
-(Figure~\ref{picture_13.png}).
-
-$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 3224$,
+,
+
+$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P}_{17})} = 3224$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.85\linewidth]{picture_13.png}}
-\parbox{0.85\linewidth}{\caption{PIPS of cardinality 17 and diameter 3224}
-\label{picture_13.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
-(\pm 1848 , 0),
-(\pm 1376 , 0),
-(\pm 1036 , 0),
-(\pm 899 , 0),
-(\pm 849 , 0),
-$
-
-$
-(\pm 613 , 0),
-(\pm 576 , 0),
-(\pm 100 , 0)\}
-$
-(Figure~\ref{picture_14.png}).
-
-$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 3696$,
+\item
+$\mathcal{P}_{18} = \mathcal{P}_{17} \cup \sqrt{154}/{1} * \{(1848, 0)\}
+$
+(Figure~\ref{picture_14.png}),
+
+$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P}_{18})} = 3696$,
 
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.85\linewidth]{picture_14.png}}
-\parbox{0.85\linewidth}{\caption{PIPS of cardinality 18 and diameter 3696}
+\center{\includegraphics[width=0.8\linewidth]{picture_14.png}}
+\parbox{0.85\linewidth}{\caption{PIPS $\mathcal{P}_{18}$ of cardinality 18 and diameter 3696}
 \label{picture_14.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
-(\pm 1848 , 0),
-(\pm 1376 , 0),
-(\pm 1036 , 0),
-(\pm 899 , 0),
-(\pm 849 , 0),
-$
-
-$
-(\pm 613 , 0),
-(\pm 576 , 0),
-(\pm 100 , 0),
-(-3293 , 0)\}
-$
-(Figure~\ref{picture_15.png}).
-
-$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 5141$,
+$\mathcal{P}_{19} = \mathcal{P}_{18} \cup \sqrt{154}/{1} * \{(-3293, 0)\}
+$
+,
+
+$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P}_{19})} = 5141$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_15.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 19 and diameter 5141}
-\label{picture_15.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 874, 60),
-(\pm 3293 , 0),
-(\pm 1848 , 0),
-(\pm 1376 , 0),
-(\pm 1036 , 0),
-(\pm 899 , 0),
-$
-
-$
-(\pm 849 , 0),
-(\pm 613 , 0),
-(\pm 576 , 0),
-(\pm 100 , 0)\}
-$
-(Figure~\ref{picture_16.png}).
-
-$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P})} = 6586$,
+\item
+$\mathcal{P}_{20} = \mathcal{P}_{19} \cup \sqrt{154}/{1} * \{(3293, 0)\}
+$
+(Figure~\ref{picture_16.png}),
+
+$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P}_{20})} = 6586$,
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_16.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 20 and diameter 6586}
+\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{20}$ of cardinality 20 and diameter 6586}
 \label{picture_16.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
+$\mathcal{P}_{21} = \sqrt{154}/{1} * \{ (\pm 2622, 180),
 (\pm 5544 , 0),
@@
 $
-(Figure~\ref{picture_17.png}).
-
-$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 11692$,
+,
+
+$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P}_{21})} = 11692$,
 
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
-(\pm 6148 , 0),
-(\pm 5544 , 0),
-(\pm 4128 , 0),
-(\pm 3108 , 0),
-$
-
-$
-(\pm 2697 , 0),
-(\pm 2547 , 0),
-(\pm 1839 , 0),
-(\pm 1728 , 0),
-(\pm 904 , 0),
-(\pm 300 , 0)\}
-$
-(Figure~\ref{picture_18.png}).
-
-$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 12296$,
+$\mathcal{P}_{22} = \mathcal{P}_{21} \cup \sqrt{154}/{1} * \{(6148, 0)\}
+$
+(Figure~\ref{picture_18.png}),
+
+$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P}_{22})} = 12296$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.75\linewidth]{picture_17.png}}
-\parbox{0.75\linewidth}{\caption{PIPS of cardinality 21 and diameter 11692}
-\label{picture_17.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
-(\pm 6148 , 0),
-(\pm 5544 , 0),
-(\pm 4128 , 0),
-(\pm 3108 , 0),
-$
-
-$
-(\pm 2697  , 0),
-(\pm 2547 , 0),
-(\pm 1839 , 0),
-(\pm 1728 , 0),
-(\pm 904 , 0),
-(\pm 300 , 0),
-(-9879 , 0)\}
-$
-(Figure~\ref{picture_19.png}).
-
-$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 16027$,
+\item
+$\mathcal{P}_{23} = \mathcal{P}_{22} \cup \sqrt{154}/{1} * \{(-9879 , 0)\}
+$
+,
+
+$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P}_{23})} = 16027$,
 
@@
 \center{\includegraphics[width=0.75\linewidth]{picture_18.png}}
-\parbox{0.75\linewidth}{\caption{PIPS of cardinality 22 and diameter 12296}
+\parbox{0.75\linewidth}{\caption{PIPS $\mathcal{P}_{22}$ of cardinality 22 and diameter 12296}
 \label{picture_18.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 2622, 180),
-(\pm 9879 , 0),
-(\pm 6148 , 0),
-(\pm 5544 , 0),
-(\pm 4128 , 0),
-$
-
-$
-(\pm 3108 , 0),
-(\pm 2697 , 0),
-(\pm 2547 , 0),
-(\pm 1839 , 0),
-(\pm 1728 , 0),
-(\pm 904 , 0),
-(\pm 300 , 0)\}
-$
-(Figure~\ref{picture_20.png}).
-
-$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P})} = 19758$,
+$\mathcal{P}_{24} = \mathcal{P}_{23} \cup \sqrt{154}/{1} * \{ (9879, 0)\}
+$
+(Figure~\ref{picture_20.png}),
+
+$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P}_{24})} = 19758$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.85\linewidth]{picture_19.png}}
-\parbox{0.85\linewidth}{\caption{PIPS of cardinality 23 and diameter 16027}
-\label{picture_19.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 5244, 360),
+\item
+$\mathcal{P}_{25} = \sqrt{154}/{1} * \{ (\pm 5244, 360),
 (\pm 12296 , 0),
@@
 $
-(Figure~\ref{picture_21.png}).
-
-$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P})} = 32054$,
+,
+
+$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P}_{25})} = 32054$,
 
@@
 \center{\includegraphics[width=0.85\linewidth]{picture_20.png}}
-\parbox{0.85\linewidth}{\caption{PIPS of cardinality 24 and diameter 19758}
+\parbox{0.85\linewidth}{\caption{PIPS $\mathcal{P}_{24}$ of cardinality 24 and diameter 19758}
 \label{picture_20.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 5244, 360),
-(\pm 19758 , 0),
-(\pm 12296 , 0),
-(\pm 11088 , 0),
-(\pm 8579 , 0),
-$
-
-$
-(\pm 8256 , 0),
-(\pm 6216 , 0),
-(\pm 5394 , 0),
-(\pm 5094 , 0),
-(\pm 3678 , 0),
-(\pm 3456 , 0),
-(\pm 1808 , 0),
-$
-
-$
-(\pm 600 , 0)\}
-$
-(Figure~\ref{picture_22.png}).
-
-$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P})} = 39516$,
+$\mathcal{P}_{26} = \mathcal{P}_{25} \cup \sqrt{154}/{1} * \{(19758 , 0)\}
+$
+(Figure~\ref{picture_22.png}),
+
+$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P}_{26})} = 39516$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_21.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 25 and diameter 32054}
-\label{picture_21.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
+\item
+$\mathcal{P}_{27} = \sqrt{154}/{1} * \{ (\pm 36708, 2520),
 (\pm 116058 , 0),
@@
 $
-(Figure~\ref{picture_23.png}).
-
-$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 254364$,
+,
+
+$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P}_{27})} = 254364$,
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_22.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 26 and diameter 39516}
+\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{26}$ of cardinality 26 and diameter 39516}
 \label{picture_22.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
+$\mathcal{P}_{28} = \mathcal{P}_{27} \cup \sqrt{154}/{1} * \{(138306, 0)\}
+$
+(Figure~\ref{picture_24.png}),
+
+$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P}_{28})} = 276612$,
+
+which gives $d(m, 2m + 24) \leq 276612$.
+
+\item
+$\mathcal{P}_{29} = \sqrt{154}/{1} * \{ (\pm 36708, 2520),
 (\pm 138306 , 0),
@@
 (\pm 12656 , 0),
-(\pm 4200 , 0)\}
-$
-(Figure~\ref{picture_24.png}).
-
-$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 276612$,
-
-which gives $d(m, 2m + 24) \leq 276612$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_23.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 27 and diameter 254364}
-\label{picture_23.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
-(\pm 138306 , 0),
-(\pm 116058 , 0),
-(\pm 86072 , 0),
-(\pm 77616 , 0),
-$
-
-$
-(\pm 60053 , 0),
-(\pm 57792 , 0),
-(\pm 43512 , 0),
-(\pm 37758 , 0),
-(\pm 35658 , 0),
-(\pm 25746 , 0),
-$
-
-$
-(\pm 24192 , 0),
-(\pm 12656 , 0),
 (\pm 4200 , 0),
@@
 $
-(Figure~\ref{picture_25.png}).
-
-$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 338169$,
+,
+
+$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P}_{29})} = 338169$,
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_24.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 28 and diameter 276612}
+\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{28}$ of cardinality 28 and diameter 276612}
 \label{picture_24.png}}
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_25.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 29 and diameter 338169}
-\label{picture_25.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 36708, 2520),
-(\pm 199863 , 0),
-(\pm 138306 , 0),
-(\pm 116058 , 0),
-(\pm 86072 , 0),
-$
-
-$
-(\pm 77616 , 0),
-(\pm 60053 , 0),
-(\pm 57792 , 0),
-(\pm 43512 , 0),
-(\pm 37758 , 0),
-(\pm 35658 , 0),
-$
-
-$
-(\pm 25746 , 0),
-(\pm 24192 , 0),
-(\pm 12656 , 0),
-(\pm 4200 , 0)\}
-$
-(Figure~\ref{picture_26.png}).
-
-$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P})} = 399726$,
+\item
+$\mathcal{P}_{30} = \mathcal{P}_{29} \cup \sqrt{154}/{1} * \{(199863 , 0)\}
+$
+(Figure~\ref{picture_26.png}),
+
+$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P}_{30})} = 399726$,
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_26.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 30 and diameter 399726}
+\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{30}$ of cardinality 30 and diameter 399726}
 \label{picture_26.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 580290, 0),
+$\mathcal{P}_{31} = \sqrt{154}/{1} * \{ (\pm 580290, 0),
 (\pm 430360, 0),
@@
 $
-(Figure~\ref{picture_31.png}).
-
-$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P})} = 1614202$,
+,
+
+$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P}_{31})} = 1614202$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_31.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 31 and diameter 1614202}
-\label{picture_31.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 691530, 0),
-(\pm 580290, 0),
-(\pm 430360, 0),
-(\pm 388080, 0),
-$
-
-$
-(\pm 300265, 0),
-(\pm 288960, 0),
-(\pm 217560, 0),
-(\pm 188790, 0),
-(\pm 183540, 12600),
-$
-
-$
-(\pm 178290, 0),
-(\pm 128730, 0),
-(\pm 120960, 0),
-(\pm 63280, 0),
-(\pm 21000, 0),
-$
-
-$
-(\pm 11224, 0),
-(-1033912, 0),
-(-999315, 0)\}
-$
-(Figure~\ref{picture_32.png}).
-
-$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P})} = 1725442$,
+\item
+$\mathcal{P}_{32} = \mathcal{P}_{31} \cup \sqrt{154}/{1} * \{(691530  , 0)\}
+$
+(Figure~\ref{picture_32.png}),
+
+$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P}_{32})} = 1725442$,
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_32.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 32 and diameter 1725442}
+\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{32}$ of cardinality 32 and diameter 1725442}
 \label{picture_32.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 999315, 0),
-(\pm 691530, 0),
-(\pm 580290, 0),
-(\pm 430360, 0),
-$
-
-$
-(\pm 388080, 0),
-(\pm 300265, 0),
-(\pm 288960, 0),
-(\pm 217560, 0),
-(\pm 188790, 0),
-$
-
-$
-(\pm 183540, 12600),
-(\pm 178290, 0),
-(\pm 128730, 0),
-(\pm 120960, 0),
-(\pm 63280, 0),
-$
-
-$
-(\pm 21000, 0),
-(\pm 11224, 0),
-(-1033912, 0)\}
-$
-(Figure~\ref{picture_33.png}).
-
-$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P})} = 2033227$,
+$\mathcal{P}_{33} = \mathcal{P}_{32} \cup \sqrt{154}/{1} * \{(999315, 0)\}
+$
+,
+
+$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P}_{33})} = 2033227$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_33.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 33 and diameter 2033227}
-\label{picture_33.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 1033912, 0),
-(\pm 999315, 0),
-(\pm 691530, 0),
-(\pm 580290, 0),
-$
-
-$
-(\pm 430360, 0),
-(\pm 388080, 0),
-(\pm 300265, 0),
-(\pm 288960, 0),
-(\pm 217560, 0),
-$
-
-$
-(\pm 188790, 0),
-(\pm 183540, 12600),
-(\pm 178290, 0),
-(\pm 128730, 0),
-(\pm 120960, 0),
-$
-
-$
-(\pm 63280, 0),
-(\pm 21000, 0),
-(\pm 11224, 0)\}
-$
-(Figure~\ref{picture_34.png}).
-
-$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P})} = 2067824$,
+\item
+$\mathcal{P}_{34} = \mathcal{P}_{33} \cup \sqrt{154}/{1} * \{(1033912, 0)\}
+$
+(Figure~\ref{picture_34.png}),
+
+$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P}_{34})} = 2067824$,
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_34.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 34 and diameter 2067824}
+\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{34}$ of cardinality 34 and diameter 2067824}
 \label{picture_34.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 3997260, 0),
+$\mathcal{P}_{35} = \sqrt{154}/{1} * \{ (\pm 3997260, 0),
 (\pm 2766120, 0),
@@
 $
-(Figure~\ref{picture_35.png}).
-
-$f = 1468320$, $v = 12824$, $w = 1468376$, $\operatorname{diam(\mathcal{P})} = 8132908$,
+,
+
+$f = 1468320$, $v = 12824$, $w = 1468376$, $\operatorname{diam(\mathcal{P}_{35})} = 8132908$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_35.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 35 and diameter 8132908}
-\label{picture_35.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 4135648, 0),
-(\pm 3997260, 0),
-(\pm 2766120, 0),
-(\pm 2321160, 0),
-$
-
-$
-(\pm 1721440, 0),
-(\pm 1552320, 0),
-(\pm 1201060, 0),
-(\pm 1155840, 0),
-(\pm 870240, 0),
-$
-
-$
-(\pm 755160, 0),
-(\pm 734160, 50400),
-(\pm 713160, 0),
-(\pm 514920, 0),
-(\pm 483840, 0),
-$
-
-$
-(\pm 308175, 0),
-(\pm 253120, 0),
-(\pm 84000, 0),
-(\pm 44896, 0)\}
-$
-(Figure~\ref{picture_36.png}).
-
-$f = 1468320$, $v = 12824$, $w = 1468376$, $\operatorname{diam(\mathcal{P})} = 8271296$,
+\item
+$\mathcal{P}_{36} = \mathcal{P}_{35} \cup \sqrt{154}/{1} * \{(4135648, 0)\}
+$
+(Figure~\ref{picture_36.png}),
+
+$f = 1468320$, $v = 12824$, $w = 1468376$, $\operatorname{diam(\mathcal{P}_{36})} = 8271296$,
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_36.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 36 and diameter 8271296}
+\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{36}$ of cardinality 36 and diameter 8271296}
 \label{picture_36.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 12991095, 0),
+$\mathcal{P}_{37} = \sqrt{154}/{1} * \{ (\pm 12991095, 0),
 (\pm 8989890, 0),
@@
 (\pm 145912, 0),
-$
-
-$
 (-13440856, 0)\}
 $
-(Figure~\ref{picture_37.png}).
-
-$f = 4772040$, $v = 41678$, $w = 4772222$, $\operatorname{diam(\mathcal{P})} = 26431951$,
+,
+
+$f = 4772040$, $v = 41678$, $w = 4772222$, $\operatorname{diam(\mathcal{P}_{37})} = 26431951$,
 
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_37.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 37 and diameter 26431951}
-\label{picture_37.png}}
-\end{figure}
-
-\item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 13440856, 0),
-(\pm 12991095, 0),
-(\pm 8989890, 0),
-(\pm 7543770, 0),
-$
-
-$
-(\pm 5594680, 0),
-(\pm 5045040, 0),
-(\pm 3903445, 0),
-(\pm 3756480, 0),
-(\pm 3694845, 0),
-$
-
-$
-(\pm 2828280, 0),
-(\pm 2454270, 0),
-(\pm 2386020, 163800),
-(\pm 2317770, 0),
-(\pm 1673490, 0),
-$
-
-$
-(\pm 1572480, 0),
-(\pm 822640, 0),
-(\pm 484680, 0),
-(\pm 273000, 0),
-$
-
-$
-(\pm 145912, 0)\}
-$
-(Figure~\ref{picture_38.png}).
-
-$f = 4772040$, $v = 41678$, $w = 4772222$, $\operatorname{diam(\mathcal{P})} = 26881712$,
+\item
+$\mathcal{P}_{38} = \mathcal{P}_{37} \cup \sqrt{154}/{1} * \{(13440856, 0)\}
+$
+(Figure~\ref{picture_38.png}),
+
+$f = 4772040$, $v = 41678$, $w = 4772222$, $\operatorname{diam(\mathcal{P}_{38})} = 26881712$,
 
@@
 \center{\includegraphics[width=1\linewidth]{picture_38.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 38 and diameter 26881712}
+\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{38}$ of cardinality 38 and diameter 26881712}
 \label{picture_38.png}}
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 51964380, 0),
+$\mathcal{P}_{39} = \sqrt{154}/{1} * \{ (\pm 51964380, 0),
 (\pm 35959560, 0),
@@
 $
-(Figure~\ref{picture_39.png}).
-
-$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam(\mathcal{P})} = 105727804$,
+,
+
+$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam(\mathcal{P}_{39})} = 105727804$,
 
 which gives $d(m, 2m + 35) \leq 105727804$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_39.png}}
-\parbox{1\linewidth}{\caption{PIPS of cardinality 39 and diameter 105727804}
-\label{picture_39.png}}
-\end{figure}
 
@@
 \item
-$\mathcal{P}=\sqrt{154}/{1} * \{ (\pm 53763424, 0),
-(\pm 51964380 , 0),
-(\pm 35959560 , 0),
-(\pm 30175080 , 0),
-$
-
-$
-(\pm 22378720 , 0),
-(\pm 20180160 , 0),
-(\pm 15613780 , 0),
-(\pm 15025920 , 0),
-(\pm 14779380 , 0),
-$
-
-$
-(\pm 11313120 , 0),
-(\pm 9817080 , 0),
-(\pm 9544080 , 655200),
-(\pm 9271080 , 0),
-(\pm 6693960 , 0),
-$
-
-$
-(\pm 6289920 , 0),
-(\pm 4006275 , 0),
-(\pm 3290560 , 0),
-(\pm 1938720 , 0),
-(\pm 1092000 , 0),
-$
-
-$
-(\pm 583648 , 0)\}
-$
-(Figure~\ref{picture_40.png}).
-
-$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam}(\mathcal{P})
-= 107526848$,
+$\mathcal{P}_{40} = \mathcal{P}_{39} \cup \sqrt{154}/{1} * \{(53763424 , 0)\}
+$
+(Figure~\ref{picture_40.png}),
+
+$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam}(\mathcal{P}_{40}) = 107526848$,
 
@@
 $
-(Figure~\ref{picture_10.png}).
+(Figure~\ref{picture_10.png}),
 
@@
 
+\makeatletter
+\ltx@iffilelater{biblatex-gost.def}{2017/02/01}%
+{\toggletrue{bbx:gostbibliography}%
+\renewcommand*{\revsdnamepunct}{\addcomma}}{}
+\makeatother
+
 \printbibliography
+\printbibitembibliography
 %\bibliography{literature}
 
+%\begin{thebibliography}{99}
+%\end{thebibliography}
+
 \end{document}

### 2025-02-24 AvdMomZvol_rails/Article.tex
[rails] Adjust the authors metadata

@@
 	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
-	, Zvolinsky R.E., Momot E.A.
+	, Zvolinskii R.E., Momot E.A.
 	\\
@@
 
+Авдеев Николай Николаевич
+
+Воронежский государственный университет, кафедра теории функций и геометрии, аспирант
+
+Россия, 394018, г. Воронеж, Университетская пл., 1
+
+E-mail: nickkolok@mail.ru, avdeev@math.vsu.ru
+
+Avdeev Nikolai Nikolaevich
+
+Voronezh State University, Department of Function Theory and Geometry, postgraduate student
+
+Russia, 394018, Voronezh, University sq., 1
+
+Зволинский Роман Евгеньевич
+
+Воронежский государственный университет, кафедра теории функций и геометрии, аспирант
+
+Россия, 394018, г. Воронеж, Университетская пл., 1
+
+E-mail: roman.zvolinskiy@gmail.com , roman-zvolinskiy@mail.ru
+
+Zvolinskii Roman Evgen'evich
+
+Voronezh State University, Department of Function Theory and Geometry, postgraduate student
+
+Russia, 394018, Voronezh, University sq., 1
+
+Момот Екатерина Александровна
+
+Воронежский государственный университет, кафедра функционального анализа и операторных уравнений, выпускница
+
+Россия, 394018, г. Воронеж, Университетская пл., 1
+
+Momot Ekaterina Alexandrovna
+
+Voronezh State University, Department of Functional Analysis and Operator Equations (graduated from)
+
+Russia, 394018, Voronezh, University sq., 1
+
 \makeatletter

### 2025-02-25 AvdMomZvol_rails/Article.tex
[rails] More references

@@
 For $9 \leq n \leq 122$, the diameter $d(2,n)$ is attained on a facher point set~\cite{kurz2008minimum}.
+Facher sets have relatively simple structure and thus
+they are rather well-investigated~\cite{zvolinsky2021facher,lushina2021vestnikpmm}.
 
@@
 is attained on a circular set.
+Moreover, the first occurrence of integral point sets in modern literature (or at least the first one that the authors are aware of) is Anning's article from 1915 which focuses exactly on circular integral point sets~\cite{anning1915discussions}.
+For the most recent advances in circular IPS, we refer the reader to~\cite{undrakh2022refinement}.
 
@@
 is presented in~\cite{my-semi-general-5-4-bound-2019}.
+It is important to note that the cardinality of an IPS in semi-general position is bounded
+not only by its diameter (i.e. the maximal distance between its points),
+but also by the minimal distance.
+Solymosi established such a bound~\cite{solymosi2003note};
+in particular, he proved that if an IPS in semi-general position has distance 2,
+its cardinality cannot exceed 7.
+Recently, this result has been refined to 5 points~\cite{lushina2024cardinality}.
 
@@
 and all the rest on the other;
-on the other hand, it is still unknown whether there are any rails PIPSs with 4 and 5 points on the lines.
+for an example of cardinality 106, we refer the reader to~\cite{momot2022example}.
+On the other hand, it is still unknown whether there are any rails PIPSs with 4 and 5 points on the lines.
 

### 2025-02-25 AvdMomZvol_rails/Article.tex
[rails] Improve the final remarks
[rails] Edits from Zvolinsky: group the pics
[rails] Edits from Zvolinsky: group the sets
[rails] Remove two more redundant pics
[rails] Cite the most recent Lushina's work

@@
 \end{figure}
+
+If a rails IPS $\mathcal P$ is contained in two parallel lines $m_1$ amd $m_2$
+and $M_1,M_2 \in \mathcal P \cap m_1$, then $|M_1 M_2|\geq 3$
+(the second condition is essential)~\cite{lushina2025norailswith2}.
 
@@
 
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=.5\linewidth]{picture_d21.png}
-	\\
-	\caption{PIPS $\mathcal{P}_8$ of cardinality 8 and diameter 21}
-	\label{picture_d21.png}
-\end{figure}
-
 Below $f$ and $w$ are those from Theorem~\ref{thm:Kurz_blowup}.
@@
 \begin{figure}[h!]
-	\centering
-	\includegraphics[width=.45\linewidth]{picture_3.png}
+	\includegraphics[width=.48\linewidth]{picture_d21.png}
+	\hfill
+	\includegraphics[width=.48\linewidth]{picture_3.png}
 	\\
-	\caption{PIPS $\mathcal{P}_{10}$ of cardinality 10 and diameter 158}
-	\label{picture_3.png}
+	\parbox{.48\linewidth}{\caption{PIPS $\mathcal{P}_8$ of cardinality 8 and diameter 21}
+		\label{picture_d21.png}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{PIPS $\mathcal{P}_{10}$ of cardinality 10 and diameter 158}
+		\label{picture_3.png}}
 \end{figure}
@@
 
-\begin{figure}[htbp]
-	\centering
-	\includegraphics[width=.5\linewidth]{picture_5.png}
+\begin{figure}[h!]
+	\includegraphics[width=.48\linewidth]{picture_5.png}
+	\hfill
+	\includegraphics[width=.48\linewidth]{picture_7.png}
 	\\
-	\caption{PIPS $\mathcal{P}_{12}$ of cardinality 12 and diameter 297}
-	\label{picture_5.png}
+	\parbox{.48\linewidth}{\caption{PIPS $\mathcal{P}_{12}$ of cardinality 12 and diameter 297}
+		\label{picture_5.png}}
+	\hfill
+	\parbox{.48\linewidth}{\caption{PIPS $\mathcal{P}_{14}$ of cardinality 14 and diameter 1273}
+		\label{picture_7.png}}
 \end{figure}
@@
 
-\begin{figure}[h!]
-\center{\includegraphics[width=0.6\linewidth]{picture_7.png}}
-\parbox{0.7\linewidth}{\caption{PIPS $\mathcal{P}_{14}$ of cardinality 14 and diameter 1273}
-\label{picture_7.png}}
-\end{figure}
-
 \item
@@
 which gives $d(m, 2m + 19) \leq 16027$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.75\linewidth]{picture_18.png}}
-\parbox{0.75\linewidth}{\caption{PIPS $\mathcal{P}_{22}$ of cardinality 22 and diameter 12296}
-\label{picture_18.png}}
-\end{figure}
 
@@
 \item
-$\mathcal{P}_{29} = \sqrt{154}/{1} * \{ (\pm 36708, 2520),
-(\pm 138306 , 0),
-(\pm 116058 , 0),
-(\pm 86072 , 0),
-(\pm 77616 , 0),
-$
-
-$
-(\pm 60053 , 0),
-(\pm 57792 , 0),
-(\pm 43512 , 0),
-(\pm 37758 , 0),
-(\pm 35658 , 0),
-(\pm 25746 , 0),
-$
-
-$
-(\pm 24192 , 0),
-(\pm 12656 , 0),
-(\pm 4200 , 0),
-(-199863 , 0)\}
+$\mathcal{P}_{29} = \mathcal{P}_{28} \cup \sqrt{154}/{1} * \{(-199863 , 0)\}
 $
@@
 which gives $d(m, 2m + 25) \leq 338169$.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{picture_24.png}}
-\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{28}$ of cardinality 28 and diameter 276612}
-\label{picture_24.png}}
-\end{figure}
 
@@
 
-The source code can be obtained at https://gitlab.com/Nickkolok/ips-algo
-
-\section{Acknowledgements}
+The source code can be obtained at https://gitlab.com/Nickkolok/ips-algo .
+The preprint of this article is arXiv:1909.10386 [math.CO]
+
 The authors thank A.S. Usachev for his help in polishing the text of the paper.

### 2025-03-01 char-4k-bound/char-4k-bound.tex
[4k3] Verify the case p<10 explicitly - fix the Referee's comment
[4k3] Refactor Corollary 5.5. a bit - fix the Referee's comment

@@
 	\label{cor:sqrt_5_3}
-	For the height of a triangle $ABC$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}_4}'$, with sides $3 \leq a \leq b \leq c$, the following estimate holds:
+	In a triangle $ABC$ with sides $3 \leq a \leq b \leq c$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}_4}'$, the smallest height is at least $(2a-1)^{1/2}$.
+\end{corollary}
+
+\begin{proof}
+	Indeed, for $a\geq 3$ the inequality
 	$$
-		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2},
+		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2}
 	$$
-	where the difference between the left and right-hand sides increases with increasing $a$.
-\end{corollary}
+	holds, and
+	the difference between the left and right-hand sides increases with increasing $a$.
+\end{proof}
 
@@
 
-	Now we have to consider $m > \frac{6}{5}p^{2/5}$.
+	If $n\geq 4$ and $p < 10$, then the claim of the theorem follows immediately from~\eqref{d}.
+	So now we have to consider $m > \frac{6}{5}p^{2/5} > 3$.
 	Then for any points $A, B \in M \setminus {M_1}$, we have $|AB| > \frac{6}{5}p^{2/5}$.

### 2025-03-02 char-4k-bound/char-4k-bound.tex
[4k3] Refactor Theorem 4.2 (The Weeding Theorem I) - fix the Referee's comment

@@
 	\label{thm:weeding_I}
-	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$ be an IPS with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
-Then, the length of the edge $M_1 M_2$ and its $n$-th Erdős curve, which contains the point $M_3$, have the same parity.
+	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$
+	be an IPS with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$,
+	ant let $M_3$ belong to the $n$-th Erdős curve of the edge $M_1 M_2$.
+	Then, $n$ and the length of $M_1M_2$ have the same parity.
 \end{theorem}
@@
 \begin{proof}
-	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and let the $n$-th Erdős curve satisfy $|M_1 M_3| - |M_2 M_3| = n$.
+	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and $n = |M_1 M_3| - |M_2 M_3|$.
 	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is an even number, is also an odd number.

### 2025-03-02 char-4k-bound/char-4k-bound.tex
[4k3] Unification: \overline{\mathfrak{M}_n}' -> \overline{\mathfrak{M}}'_n

@@
 	\label{lem:no_distance_2}
-	Every set $M \in \overline{\mathfrak{M}_n}'$ with an edge $|M_1M_2|=2$ has a cardinality $n=3$.
+	Every set $M \in \overline{\mathfrak{M}}'_n$ with an edge $|M_1M_2|=2$ has the cardinality $n=3$.
 \end{lemma}
@@
 \begin{proof}
-	Let $M \in \overline{\mathfrak{M}_n}'$ and $M_1, M_2 \in M$.
+	Let $M \in \overline{\mathfrak{M}}'_n$ and $M_1, M_2 \in M$.
 	Then all points in the set $M$ lie on $cr(M_1, M_2)$.
@@
 
-	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}_4}'$.
+	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}}'_4$.
 	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
@@
 	\label{cor:sqrt_5_3}
-	In a triangle $ABC$ with sides $3 \leq a \leq b \leq c$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}_4}'$, the smallest height is at least $(2a-1)^{1/2}$.
+	In a triangle $ABC$ with sides $3 \leq a \leq b \leq c$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}}'_4$, the smallest height is at least $(2a-1)^{1/2}$.
 \end{corollary}
@@
 	\label{lem:intersection_enumeration}
-	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}_n}'$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$.
+	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}}'_n$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$.
 Then $\# M \le |M_1 M_2| \cdot |M_3 M_4| - 2$.
@@
 	\label{thm:main_result}
-	Let $M \in \overline{\mathfrak{M}}_n'$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
+	Let $M \in \overline{\mathfrak{M}}'_n$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
 	Then, for every integer $n \geq 3$, the inequality
@@
 
-	Let us consider $M \in \overline{\mathfrak{M}_n}'$, $n \geq 4$, $\operatorname{diam} M = p$.
+	Let us consider $M \in \overline{\mathfrak{M}}'_n$, $n \geq 4$, $\operatorname{diam} M = p$.
 

### 2025-03-05 AvdMomZvol_rails/Article.tex
[rails] Inject generated bibliography

@@
 
-\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
-\addbibresource{../common/notmy.bib}
-\addbibresource{../common/my.bib}
-
-\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
-
-%\usepackage{cite}
+%\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+%\addbibresource{../common/notmy.bib}
+%\addbibresource{../common/my.bib}
+
+%\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
+
+\usepackage{cite}
 
@@
 
-\makeatletter
-\ltx@iffilelater{biblatex-gost.def}{2017/02/01}%
-{\toggletrue{bbx:gostbibliography}%
-\renewcommand*{\revsdnamepunct}{\addcomma}}{}
-\makeatother
-
-\printbibliography
-\printbibitembibliography
-%\bibliography{literature}
-
-%\begin{thebibliography}{99}
-%\end{thebibliography}
+% The "long" GOST
+
+%\makeatletter
+%\ltx@iffilelater{biblatex-gost.def}{2017/02/01}%
+%{\toggletrue{bbx:gostbibliography}%
+%\renewcommand*{\revsdnamepunct}{\addcomma}}{}
+%\makeatother
+
+%\printbibliography
+%\printbibitembibliography
+
+% Generated from .bib files with biblatex-gost and biblatex2bibitem
+
+\begin{thebibliography}{99}
+
+{}
+\bibitem{ascher2020erdos}
+\emph{Ascher}, \emph{K.}, \emph{Braune}, \emph{L.},
+\emph{Turchet}, \emph{A.} The Erdős–Ulam problem, Lang’s conjecture and
+uniformity / K. Ascher, L. Braune, A. Turchet // Bulletin of the London Mathematical
+Society. — 2020. — Vol. 52, no. 6. — Pp. 1053–1063. — arXiv: \href
+{http://arxiv.org/abs/1809.02037} {\nolinkurl {1809.02037}}.
+{}
+\bibitem{solymosi2010question}
+\emph{Solymosi}, \emph{J.}, \emph{De Zeeuw}, \emph{F.} On a question of Erdős
+and Ulam / J. Solymosi, F. De Zeeuw // Discrete \& Computational Geometry. —
+2010. — Vol. 43, no. 2. — Pp. 393–401. — arXiv: \href {http://arxiv.org/abs/0806.3095}
+{\nolinkurl {0806.3095}}.
+{}
+\bibitem{kurz2007enumeration}
+\emph{Kurz}, \emph{S.} Enumeration of Integral Tetrahedra / S. Kurz // Journal of
+Integer Sequences. — 2007. — Vol. 10, no. 2. — P. 3.{}
+\bibitem{anning1945integral}
+\emph{Anning}, \emph{N. H.}, \emph{Erdős}, \emph{P.} Integral distances /
+N. H. Anning, P. Erdős // Bulletin of the American Mathematical Society. — 1945. —
+Vol. 51, no. 8. — Pp. 598–600. — DOI: \href
+{https://doi.org/10.1090/S0002-9904-1945-08407-9} {\nolinkurl
+{10.1090/S0002-9904-1945-08407-9}}.
+{}
+\bibitem{erdos1945integral}
+\emph{Erdős}, \emph{P.} Integral distances / P. Erdős // Bulletin of the American
+Mathematical Society. — 1945. — Vol. 51, no. 12. — P. 996. — DOI: \href
+{https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl
+{10.1090/S0002-9904-1945-08490-0}}.
+{}
+\bibitem{kemnitz1988punktmengen}
+\emph{Kemnitz}, \emph{A.} Punktmengen mit ganzzahligen Abständen /
+A. Kemnitz. — 1988.
+{}
+\bibitem{kurz2005characteristic}
+\emph{Kurz}, \emph{S.} On the characteristic of integral point sets in {$\mathbb
+{E}^m$} / S. Kurz // Australasian Journal of Combinatorics. — 2006. — Vol. 36. —
+Pp. 241–248. — arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl
+{math/0511704}}.
+{}
+\bibitem{kurz2008bounds}
+\emph{Kurz}, \emph{S.}, \emph{Laue}, \emph{R.} Bounds for the minimum
+diameter of integral point sets / S. Kurz, R. Laue // Australasian Journal of
+Combinatorics. — 2007. — Vol. 39. — Pp. 233–240. — arXiv: \href
+{http://arxiv.org/abs/0804.1296} {\nolinkurl {0804.1296}}.
+{}
+\bibitem{solymosi2003note}
+\emph{Solymosi}, \emph{J.} Note on integral distances / J. Solymosi // Discrete \&
+Computational Geometry. — 2003. — Vol. 30, no. 2. — Pp. 337–342. — DOI: \href
+{https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.
+{}
+\bibitem{my-pps-linear-bound-2019}
+\emph{Avdeev}, \emph{N.} On existence of integral point sets and their diameter
+bounds / N. Avdeev // Australasian Journal of Combinatorics. — 2020. — Vol. 77,
+no. 1. — Pp. 100–116. — arXiv: \href {http://arxiv.org/abs/1906.11926} {\nolinkurl
+{1906.11926}}. — URL: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190611926A/abstract}.
+{}\bibitem{nozaki2013lower}
+\emph{Nozaki}, \emph{H.} Lower bounds for the minimum diameter of integral point
+sets / H. Nozaki // Australasian Journal of Combinatorics. — 2013. — Vol. 56. —
+Pp. 139–143.
+{}
+\bibitem{our-vmmsh-2018}
+\emph{Авдеев}, \emph{Н. Н.}, \emph{Семёнов}, \emph{Е. М.} Множества точек с
+целочисленными расстояниями на плоскости и в евклидовом пространстве /
+Н. Н. Авдеев, Е. М. Семёнов // Математический форум (Итоги науки. Юг
+России). — 2018. — Pp. 217–236.
+{}
+\bibitem{antonov2008maximal}
+\emph{Antonov}, \emph{A. R.}, \emph{Kurz}, \emph{S.} Maximal integral point
+sets over {$\mathbb {Z}^2$} / A. R. Antonov, S. Kurz // International Journal of
+Computer Mathematics. — 2008. — Vol. 87, no. 12. — Pp. 2653–2676. — DOI: \href
+{https://doi.org/10.1080/00207160902993636} {\nolinkurl
+{10.1080/00207160902993636}}. — arXiv: \href {http://arxiv.org/abs/0804.1280}
+{\nolinkurl {0804.1280}}.
+{}
+\bibitem{kurz2008minimum}
+\emph{Kurz}, \emph{S.}, \emph{Wassermann}, \emph{A.} On the minimum
+diameter of plane integral point sets / S. Kurz, A. Wassermann // Ars Combinatoria. —
+2011. — Vol. 101. — Pp. 265–287. — arXiv: \href {http://arxiv.org/abs/0804.1307}
+{\nolinkurl {0804.1307}}.
+{}
+\bibitem{zvolinsky2021facher}
+\emph{Zvolinsky}, \emph{R. E.} Facher integral point sets with particular distances of
+arbitrary cardinality / R. E. Zvolinsky // Актуальные проблемы прикладной
+математики, информатики и механики. — 2021. — Pp. 668–674.
+{}
+\bibitem{lushina2021vestnikpmm}
+\emph{Лушина}, \emph{Е. А.}, \emph{Авдеев}, \emph{Н. Н.} О характеристике
+плоских множеств с целочисленными расстояниями с рёбрами 1 и 2 /
+Е. А. Лушина, Н. Н. Авдеев // Вестник факультета прикладной математики,
+информатики и механики. — 2021. — Vol. 15. — Pp. 117–123.
+{}
+\bibitem{our-ped-2018}
+\emph{Авдеев}, \emph{Н. Н.} On integral point sets in special position /
+Н. Н. Авдеев // Некоторые вопросы анализа, алгебры, геометрии и
+математического образования: материалы международной молодежной научной
+школы «Актуальные направления математического анализа и смежные
+вопросы». — 2018. — Vol. 8. — Pp. 5–6.{}
+\bibitem{our-pmm-2018}
+\emph{Авдеев}, \emph{Н. Н.} Об отыскании целоудалённых множеств
+специального вида / Н. Н. Авдеев // Актуальные проблемы прикладной
+математики, информатики и механики - сборник трудов Международной научной
+конференции. — Научно-исследовательские публикации, 2018. — Pp. 492–498.
+{}
+\bibitem{lushina2025norailswith2}
+\emph{Лушина}, \emph{Е. А.} О длине основания рельсового целоудалённого
+множества / Е. А. Лушина // Современные проблемы математики и её
+приложений. — Институт математики и механики УрО РАН им. Н.Н. Красовского,
+2025. — в печати.
+{}
+\bibitem{harborth1993upper}
+\emph{Harborth}, \emph{H.}, \emph{Kemnitz}, \emph{A.},
+\emph{Möller}, \emph{M.} An upper bound for the minimum diameter of integral
+point sets / H. Harborth, A. Kemnitz, M. Möller // Discrete \& Computational
+Geometry. — 1993. — Vol. 9, no. 4. — Pp. 427–432. — DOI: \href
+{https://doi.org/10.1007/bf02189331} {\nolinkurl {10.1007/bf02189331}}.
+{}
+\bibitem{piepmeyer1996maximum}
+\emph{Piepmeyer}, \emph{L.} The maximum number of odd integral distances
+between points in the plane / L. Piepmeyer // Discrete \& Computational Geometry. —
+1996. — Vol. 16, no. 1. — Pp. 113–115. — DOI: \href
+{https://doi.org/10.1007/bf02711135} {\nolinkurl {10.1007/bf02711135}}.
+{}
+\bibitem{bat2018number}
+\emph{Bat-Ochir}, \emph{G.} On the number of points with pairwise integral
+distances on a circle / G. Bat-Ochir // Discrete Applied Mathematics. — 2018. —
+Vol. 254. — Pp. 17–32. — DOI: \href {https://doi.org/10.1016/j.dam.2018.07.004}
+{\nolinkurl {10.1016/j.dam.2018.07.004}}.
+{}
+\bibitem{anning1915discussions}
+\emph{Anning}, \emph{N.} Discussions: Relating to a Geometric Representation of
+Integral Solutions of Certain Quadratic Equations / N. Anning // The American
+Mathematical Monthly. — 1915. — Vol. 22, no. 9. — Pp. 321–321.
+{}
+\bibitem{undrakh2022refinement}
+\emph{Undrakh}, \emph{B.}, \emph{Bat-Ochir}, \emph{G.} Refinement of the upper
+bound of the radius of a circular integral points set / B. Undrakh, G. Bat-Ochir // Gulf
+Journal of Mathematics. — 2022. — Vol. 12, no. 1. — Pp. 28–40.
+{}\bibitem{kreisel2008heptagon}
+\emph{Kreisel}, \emph{T.}, \emph{Kurz}, \emph{S.} There are integral heptagons,
+no three points on a line, no four on a circle / T. Kreisel, S. Kurz // Discrete \&
+Computational Geometry. — 2008. — Vol. 39, no. 4. — Pp. 786–790. — DOI: \href
+{https://doi.org/10.1007/s00454-007-9038-6} {\nolinkurl {10.1007/s00454-007-9038-6}}.
+{}
+\bibitem{kurz2013constructing}
+Constructing $7$-clusters / S. Kurz [et al.] // Serdica Journal of Computing. — 2014. —
+Vol. 8, no. 1. — Pp. 47–70. — arXiv: \href {http://arxiv.org/abs/1312.2318} {\nolinkurl
+{1312.2318}}.
+{}
+\bibitem{avdeev2021particular}
+\emph{Avdeev}, \emph{N.}, \emph{Momot}, \emph{E.},
+\emph{Zvolinsky}, \emph{A.} On particular examples of planar integral point sets and
+their classification / N. Avdeev, E. Momot, A. Zvolinsky. —. — arXiv: \href
+{http://arxiv.org/abs/2102.12462 [math.CO]} {\nolinkurl {2102.12462 [math.CO]}}.
+{}
+\bibitem{my-semi-general-5-4-bound-2019}
+\emph{Avdeev}, \emph{N.} On diameter bounds for planar integral point sets in
+semi-general position / N. Avdeev // Chebyshevskii Sbornik. —. — Vol. 22, no. 4. —
+Pp. 343–350. — DOI: \href {https://doi.org/10.22405/2226-8383-2021-22-4-343-350}
+{\nolinkurl {10.22405/2226-8383-2021-22-4-343-350}}. — arXiv: \href
+{http://arxiv.org/abs/1907.09331} {\nolinkurl {1907.09331}}. — URL: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190709331A/abstract}.
+{}
+\bibitem{lushina2024cardinality}
+\emph{Лушина}, \emph{Е. А.}, \emph{Бриндюк}, \emph{В. С.} О мощности
+плоского целоудаленного множества с ребром 2 / Е. А. Лушина, В. С. Бриндюк //
+~74-я Международная студенческая научно-техническая конференция. Материалы.
+Астрахань. — 2024. — Pp. 741–743.
+{}
+\bibitem{guy2013unsolved}
+\emph{Guy}, \emph{R.} Unsolved problems in number theory. Vol. 1 / R. Guy. —
+Springer Science \& Business Media, 2013. — DOI: \href
+{https://doi.org/10.1007/978-1-4757-1738-9} {\nolinkurl {10.1007/978-1-4757-1738-9}}.
+{}
+\bibitem{momot2022example}
+\emph{Momot}, \emph{E. A.}, \emph{Zvolinskiy}, \emph{R. E.},
+\emph{Zvolinskiy}, \emph{A. E.} An example of rails integral point set with
+cardinality 106 / E. A. Momot, R. E. Zvolinskiy, A. E. Zvolinskiy // Прикладная
+математика и фундаментальная информатика. — 2022. — Pp. 114–115.
+
+\end{thebibliography}
 

### 2025-03-12 char-4k-bound/char-4k-bound.tex
[4k3][typo] Just a result of another proofreading round
[4k3] Minor edits after external proofreading

@@
 elements of $M$ are integers, and $M$ does not contain collinear triples.
-We improve the lower bound for diameter of such sets in the particular case
+In this paper we improve the lower bound for diameter of such sets in the particular case
 when the characteristic of the set is of the form $4k+1$ or $4k+2$.
@@
 	A planar integral point set (IPS) is a set of points in the plane, such that the distance between any pair of its points is an integer, and at least one triple of its points is non-collinear.
-The latter condition is essential to avoid subsets of a straight line; those are de-facto equivalent to subsets of integer numbers and form a completely different combinatorial object.
+The latter condition is essential to avoid subsets of a straight line; those are de-facto equivalent to subsets of integers and form a completely different combinatorial object.
 
-	In 1945, Erdős gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
+	In 1945, Erdős gave an elegant proof that every IPS is finite~\cite{anning1945integral,erdos1945integral} .
 	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
@@
 	however, Erdős did not only estimate the diameter,
-	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolas.
+	but also established the toolset for further investigation.
+	The core of this approach is a system of cofocal hyperbolas.
 
@@
 	$$
-	Although Solymosi did not give the constant $c$ explicitly,
+	Although Solymosi did not state the value of $c$ explicitly,
 	it can be inferred~\cite{our-vmmsh-2018-translit} from his proof that $c=\frac{1}{24}$.
@@
 	\end{equation}
-	(in assumption that the set has at east 4 points).
+	(assuming that the set has at least 4 points).
 	All these lower bounds are based on the Erdős's framework --- cofocal hyperbolas.
@@
 	In 2000s, Kurz introduced~\cite{kurz2005characteristic} the function $d(2,n)$ that evaluates
-	to the minimal possible diameter of planar IPS of cardinality $n$.
-	Then Kurz employed Kemnitz's results and found~\cite[Subsection 4.2]{kurz2008minimum} exact values of $d(2,n)$
-	up to $n=122$ by exhaustive computer search.
+	the minimal possible diameter of planar IPS of cardinality $n$.
+	Then Kurz employed Kemnitz's results and found the exact values of $d(2,n)$
+	up to $n=122$ by exhaustive computer search~\cite[Subsection 4.2]{kurz2008minimum}.
 	(Taking the characteristic into consideration allows to boost such search significantly ---
@@
 
-	In the present paper, we put the power of Erdős's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
+	In the present paper, we combine the power of Erdős's and Kemnitz's approaches together.
+Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
 	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
@@
 In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological.
-Section 4 is devoted to the connection between Erdős curves and the characteristic.
+Section 4 describes a connection between Erdős curves and the characteristic.
 In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.
@@
 
-For the sake of brevity, the following notations will be used for sets
+For the sake of brevity, the following notations are used for sets
 of positive integers, non-negative integers and all the integers resp.:
@@
 
-For a finite set $M$, we will denote its cardinality by $\#M$.
+For a finite set $M$, we denote its cardinality by $\#M$.
 
@@
 
-When we say that a set in non-collinear,
+When we say that a set is non-collinear,
 we mean that it has at least one non-collinear triple.
 If we tighten the condition and require all the triples to be non-collinear,
-we get the next definition.
+we obtain the next definition.
 
@@
 \begin{definition}
-	If points $M_1, M_2 \in M \in \mathfrak{M}$,
-	then the line segment $M_1 M_2$ is said to be an \emph{edge} of $M$.
+	For any two points $M_1, M_2 \in M \in \mathfrak{M}$,
+	the line segment $M_1 M_2$ is said to be an \emph{edge} of $M$.
 \end{definition}
@@
 	For an integer $n$, $|n| < |M_1 M_2|$,
-	we will call the set of points
+	we call the set of points
 	\begin{equation}
@@
 Thus, if an IPS $M$ satisfies these conditions,
-we will write that
+we write that
 $M\in\overline{\mathfrak{M}}'$, and also $M\in\overline{\mathfrak{M}}'_n$ for $n=\#M$.
@@
 are neither too exotic nor pathological.
-In order to do this, we will provide important examples of IPS for each class.
+In order to do this, we provide important examples of IPS for each class.
 
@@
 are investigated in~\cite{antonov2008maximal}.
-There are also much more complex IPS with characteristic 1;
+There are also significantly more complex IPS with characteristic 1;
 for those from $\dot{\mathfrak{M}}_7$, we refer the reader to~\cite{kurz2013constructing}.
@@
 		\begin{cases}
-		\frac{\sqrt{(a+m^2)^2+pb^2}}{2m}&=m+n-s.\\
-		\frac{\sqrt{(a-m^2)^2+pb^2}}{2m}&=n
+		\frac{\sqrt{(a+m^2)^2+pb^2}}{2m}&=~m+n-s,\\
+		\frac{\sqrt{(a-m^2)^2+pb^2}}{2m}&=~n
+		.
 		\end{cases}
 	\end{equation}
-	Multiply each equation by $2m$ and square both sides of the equations in system~\eqref{eq:29}
+	Multiplying each equation by $2m$ and squaring both sides, we obtain
 	\begin{equation}
@@
 		\begin{cases}
-		(a+m^2)^2+pb^2&=4m^2(m+n-s)^2,\\
-		(a-m^2)^2+pb^2&=4m^2n^2
+		(a+m^2)^2+pb^2&=~4m^2(m+n-s)^2,\\
+		(a-m^2)^2+pb^2&=~4m^2n^2
+		.
 		\end{cases}
 	\end{equation}
-	and then subtract the second equation from the first one:
+	Subtract the second equation from the first one:
 	\begin{equation}
@@
 	\end{equation}
-	Substitute the obtained expression~\eqref{eq:32_pre_weeding_I}
-	into the second equation of the system~\eqref{eq:30_pre_weeding_I}:
+	Substitute the expression~\eqref{eq:32_pre_weeding_I}
+	into the second equation of~\eqref{eq:30_pre_weeding_I}:
 	\begin{equation}
-		((m-s)(m+2n-s)-m^2)^2+pb^2=4m^2n^2,
+		((m-s)(m+2n-s)-m^2)^2+pb^2=4m^2n^2.
 	\end{equation}
-	that immediately gives
+	This immediately gives
 	\begin{equation}
@@
 
-The following theorem follows immediately.
+The next theorem follows immediately.
 
@@
 	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and $n = |M_1 M_3| - |M_2 M_3|$.
-	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is an even number, is also an odd number.
-	Let $m$ be an even number; then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is an even number, as a difference of two even numbers.
+	If $m$ is odd, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is even. Thus, $n$ is odd.
+	If $m$ is even, then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is even, as a difference of two even numbers.
 \end{proof}
@@
 	\label{lem:pre_weeding_II}
-	Every set $M = \{A, B, C\} \in \mathfrak{M}_3$ with an even edge $|AB| = 2q$, where $|AC| - |BC| = 2q - s$, $q, s \in \mathbb{N}$, $s$ is an odd number and $s \leq 2q$, has a characteristic of the form $p = 8k + 7$, $k \in \mathbb{N}_0$.
+	Every set $M = \{A, B, C\} \in \mathfrak{M}_3$ with an even edge $|AB| = 2q$, where $|AC| - |BC| = 2q - s$, $q, s \in \mathbb{N}$, $s$ is odd and $s \leq 2q$, has a characteristic of the form $p = 8k + 7$, $k \in \mathbb{N}_0$.
 \end{lemma}
@@
 	According to Lemma~\ref{lem:pre_weeding_II}, the difference $n=|M_1 M_3|-|M_2 M_3|=2q-s$,
-	where $s$ is an odd number,
-	is also an odd number and $\operatorname{char} M = 8k + 7$, $k\in \mathbb{N}_{0}$.
+	where $s$ is odd,
+	is also odd and $\operatorname{char} M = 8k + 7$, $k\in \mathbb{N}_{0}$.
 \end{proof}
@@
 	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}}'_4$.
-	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
+	We employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
 	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
@@
 	By Definition~\ref{def:char}, this means that $\operatorname{char}{M}={1}$.
-	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
+	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we obtain
 	$$
@@
 	\end{equation}
-	Now, we find the solutions to equation~\eqref{eq: 42} in positive integers.
-	To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
+	To find the solutions to equation~\eqref{eq: 42} in positive integers,
+	we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
 	Then equation~\eqref{eq: 42} takes the form:
@@
 	\end{equation}
-	Break down the left-hand side of equation~\eqref{eq: 43} into factors:
+	Factorising the left-hand side of equation~\eqref{eq: 43} yields:
 	\begin{equation}
@@
 	Since $s,k\in\mathbb{N}$, the number 1 can be represented as the product of two integers in two ways: $1\cdot1$ and $-1\cdot(-1)$.
-In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$, which contradicts that $k\in\mathbb{N}$.
-	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$, which also contradicts that $s,k\in\mathbb{N}$.
+In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$.
+This contradicts that $k\in\mathbb{N}$.
+	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$. This also contradicts that $s,k\in\mathbb{N}$.
 
@@
 \begin{proof}
-	Indeed, consider the triangle $ABC$ and let $|BC|=a$, $|AC|=b$.
+	Consider the triangle $ABC$ and let $|BC|=a$, $|AC|=b$.
 	Using the triangle inequality $|AB| < |BC|+|AC|$,
 	we can represent the length of side $AB$ as $|AB|=a+b-s$, $s \in \mathbb{N}$ and $s \leq a$.
-	For $s=1$, the conditions of Lemma~\ref{lem:pre_weeding_I} are satisfied: $s$ is an odd number and $s \leq a$.
+	For $s=1$, the conditions of Lemma~\ref{lem:pre_weeding_I} are satisfied: $s$ is odd and $s \leq a$.
 	Therefore, the triangle with sides $a$, $b$, and $c=a+b-1$
@@
 \begin{proof}
-	Indeed, for $a\geq 3$ the inequality
+	For $a\geq 3$ the inequality
 	$$
@@
 
-The following lemma is a classical Erdős-style intersection-enumeration one,
+The following result is a classical Erdős-style intersection-enumeration lemma,
 empowered by our Weeding Theorem I.
@@
 \begin{proof}
-	We will distinguish three cases: two even edges, two odd edges and two edges with different parity.
+	Consider three cases: two even edges, two odd edges and two edges with different parity.
 
-	Let's consider the first case. Suppose both edges $M_1M_2$ and $M_3M_4$ have even lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:
+	\paragraph{Case 1.} Suppose that both edges $M_1M_2$ and $M_3M_4$ have even lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:
 
@@
 
-	Let's consider the second case. Suppose both edges $M_1M_2$ and $M_3M_4$ have odd lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:
+	\paragraph{Case 2.} Suppose that both edges $M_1M_2$ and $M_3M_4$ have odd lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:
 
@@
 
-	Let's consider the third case.
-Suppose the lengths of edges $M_1M_2$ and $M_3M_4$ are different.
-Without loss of generality suppose the edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
+	\paragraph{Case 3.} Suppose that the lengths of edges $M_1M_2$ and $M_3M_4$ are different. Without loss of generality suppose the edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
 	Then, for each point $N\in M$, one of the following conditions is satisfied:
@@
 
-Now we are finally ready to proove our main result.
+Now we are finally ready to prove our main result.
 It improves the bound~\eqref{eq:chebsb_lower_bound}
@@
 	\end{equation}
-	which is exactly the claim of the theorem.
+	which is exactly the assertion of the theorem.
 
@@
 
-	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
-	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
+	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the assertion follows.
+	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter also holds.
 	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10 655$.
@@
 	Another research area that is able to utilize our results fruitfully is the maximization of IPS.
-	The Weeding Theorems I and II can sometimes make the exhaustive search up to 4 times faster.
+	The Weeding Theorems I and II can sometimes make the exhaustive search up to four times faster.
 

### 2025-03-12 char-4k-bound/char-4k-bound.tex
[4k3] Use a more clear designation
[4k3] Restore a runaway equation number

@@
 	\begin{equation}
-		\operatorname{diam} M \ge \left( \frac{25}{36} n \right)^{5/4}
+		\operatorname{diam} M \ge \left( \frac{25}{36} \cdot \# M \right)^{5/4}
 		.
@@
 	or
-	\begin{eqnarray}
-		pb^2+1 &=&4(-4t^4+8t^3m+8t^3n-8t^3-4t^2m^2-12t^2mn+12t^2m- {}\nonumber\\
-			 &     &-4t^2n^2+12t^2n-6t^2+4tm^2n-4tmn^2-12tmn+6tm- {}\nonumber\\
+	\begin{equation}\begin{array}{rcl}
+		pb^2+1 &=&4(-4t^4+8t^3m+8t^3n-8t^3-4t^2m^2-12t^2mn+12t^2m- {}\\
+			 &     &-4t^2n^2+12t^2n-6t^2+4tm^2n-4tmn^2-12tmn+6tm- {}\\
 			 &     &-4tn^2+6tn-2t+2m^2n-m^2+2mn^2-3mn+m-n^2+n).%{}\nonumber\\
 	\label{eq:35}
-	\end{eqnarray}
+	\end{array}\end{equation}
 	If equation~\eqref{eq:35} has integer solutions, then both $p$ and $b$ are odd.

### 2025-08-24 AvdMomZvol_rails/Article.tex
[rails] Remove refs to non-existent figures
[rails] The non-English annotation

@@
 \maketitle
+
+УДК 519.146 + 514.11
+
+О некоторых оценках на диаметр целоудалённых множеств в многомерных пространствах
+
+Ключевые слова: целоудалённое множество, множество с целочисленными расстояниями, дискретная комбинаторная геометрия, диаметр
+
+Keywords: integral point set, discrete combinatorial geometry, diameter
+
+Аннотация:
+
+	Целоудалённое множество $\mathcal{P}$ есть такое множество из $n$ точек
+	в $m$-мерном евклидовом пространстве $\mathbb{R}^{m}$,
+	что все попарные расстояния между точками есть целые числа
+	и $\mathcal{P}$ не содержится ни в какой $(m-1)$-мерной гиперплоскости.
+	В данной статье обсуждаются некоторые классы плоских целоудалённых множеств ($m=2$);
+	преимущественно рассматриваются подмножества объединений двух прямых:
+	веерные, рельсовые, ножницеподобные и пирамидальные множества.
+	Рельсовые и пирамидальные множества используются для оценок
+	минимального диаметра целоудалённых множества в многомерных пространствах.
 
@@
 $\mathcal{P}_{22} = \mathcal{P}_{21} \cup \sqrt{154}/{1} * \{(6148, 0)\}
-$
-(Figure~\ref{picture_18.png}),
+$,
 
@@
 $\mathcal{P}_{28} = \mathcal{P}_{27} \cup \sqrt{154}/{1} * \{(138306, 0)\}
-$
-(Figure~\ref{picture_24.png}),
+$,
 

### 2025-08-24 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Some old polishing

@@
-\documentclass[11pt,A4paper]{article}
+\documentclass[10pt,A4paper]{article}
 \usepackage[utf8]{inputenc}
@@
 \title{
-	On particular examples of planar integral point sets and their classification
+	Particular examples of planar integral point sets and their classification
 	\footnote{
@@
 	The diameter of the integral point set $\mathcal{P}$ is defined as
-	\begin{equation}
+	$
 		\operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
@@
 		.
-	\end{equation}
+	$
 \end{dfn}
 
-Every planar IPS aso has a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.
-
-\begin{dfn}
+\begin{dfn}(\cite{kemnitz1988punktmengen,kurz2005characteristic})
 	The characteristic of a planar integral point set $M$ is the least positive integer $q$
@@
 \begin{dfn}
-	A planar integral point set $M$ is said to be in \emph{semi-general position}
+	A PIPS $M$ is said to be in \emph{semi-general position}
 	if no three points of $M$ are located in a straight line.
@@
 		(\pm328080480; 0);
-		\\
 		(\pm268861320; 0);
 		(\pm194124840; 0);
+		\\
 		(\pm182407680; 0);
@@
 		(\pm284695320; 0);
-		\\
 		(\pm1506967020; 0);
@@
 		(\pm875077320; 0);
+		\\
 		(\pm648982880; 0);
 		(\pm585224640; 0);
-		\\
 		(\pm452799620; 0);
@@
 		(\pm116181975; 0);
+		\\
 		(\pm428602020; 0);
-		\\
 		(\pm56222880; 0);
@@
 		( \pm23985026520 ; 0);
-		\\
 		( \pm389293216 ; 0);
 		( \pm6183810360 ; 0);
+		\\
 		( \pm4464871320 ; 0);
@@
 		( \pm728364000 ; 0);
-		\\
 		( \pm35860203808 ; 0);
 		( \pm34660241460 ; 0);
+		\\
 		( \pm7545851040 ; 0);
 		( \pm20126778360 ; 0);
-		\\
 		( \pm14926606240 ; 0);
@@
 		( \pm10414391260 ; 0);
+		\\
 		( \pm2194803520 ; 0);
-		\\
 		( \pm2672185425 ; 0);
@@
 The distance between the non-adoptable points is
-\begin{equation}
+$
 	\sqrt{\left(\frac{1767}{2} - \frac{340}{2}\right)^2 + \left(\frac{147}{2}\right)^2\cdot143} = 2\sqrt{320401}
-	.
-\end{equation}
-It's notable that 320401 is a prime number.
+$.
+Notably, 320401 is a prime.
 
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{./img/9_2890_1_02e6afb1c19846c6c4debb44e52cc468_other.png}}
+\center{\includegraphics[width=0.6\linewidth]{./img/9_2890_1_02e6afb1c19846c6c4debb44e52cc468_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 9 and diameter 2890}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{./img/19_20663808074_1_0070b97f26bb194a7de2ddc37ee38b08_symm.png}}
+\center{\includegraphics[width=0.85\linewidth]{./img/19_20663808074_1_0070b97f26bb194a7de2ddc37ee38b08_symm.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 20663808074}
@@
 		( 987843675 ; \pm1146332880);
-		\\
+		%\\
 		( \pm729918777 ; 0);
 		( \pm972103809 ; 0);
+		\\
 		( \pm1113030324 ; 0);
 		( \pm1400170149 ; 0);
-		\\
+		%\\
 		( \pm3137217825 ; 0);
@@
 On Figure~\ref{17_1024296_1_639b}, the following IPS is shown:
-\begin{multline}
+$
 	\mathcal{P}_{17}=
@@
 		( \pm98943 ; 0);
-		\\
+		%\\
 		( -694668 ; 0);
+		$\\$
 		( -71487 ; 0);
 		( -50367 ; 0);
+		%\\
 		( -14943 ; 0);
 		( 15457 ; 0);
-		\\
+		%\\
 		( 23582 ; 0);
@@
 	\}
-\end{multline}
+$
 Note that the axis of symmetry for $\mathcal{P}_{17}$ is the $x$ axis;
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{./img/17_1024296_1_639b7a0452227fde1aceabf02e8778e4_other.png}}
+\center{\includegraphics[width=0.7\linewidth]{./img/17_1024296_1_639b7a0452227fde1aceabf02e8778e4_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 17 and diameter 1024296}
@@
 \begin{figure}[h!]
-	\center{\includegraphics[width=1\linewidth]{./img/10_2431_1_0606659e825917be0c3c431c1375bf21_other.png}}
+	\center{\includegraphics[width=0.7\linewidth]{./img/10_2431_1_0606659e825917be0c3c431c1375bf21_other.png}}
 	\parbox{1\linewidth}{\caption{IPS of cardinality 10 and diameter 2431}
@@
 \begin{figure}[h!]
-	\center{\includegraphics[width=1\linewidth]{./img/15_19203744_1_80dbf8e05b25f69b8366e50d1d7463f6_other.png}}
+	\center{\includegraphics[width=0.7\linewidth]{./img/15_19203744_1_80dbf8e05b25f69b8366e50d1d7463f6_other.png}}
 	\parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 19203744}
@@
 Figure~\ref{10_2431_1_0606}:
-\begin{multline*}
+$
 	\mathcal{P}_{10} =
@@
 	\}
-\end{multline*}
+$
 
 Figure~\ref{15_19203744_1_80db}:
-\begin{multline*}
+$
 	\mathcal{P}_{15} =
@@
 		( 5385600 ; \pm2625480);
-		\\
+		%\\
 		( -6283200 ; 0);
@@
 		( -635040 ; 0);
+		$\\$
 		( 1884960 ; 0);
-		\\
+		%\\
 		( 3147210 ; 0);
@@
 	\}
-\end{multline*}
+$
 
@@
 %
-\begin{multline}
+\begin{equation}
 	\mathcal{P}_8=
@@
 	( 29575 ; 0);
-	\\
+	%\\
 	( -2366 ; -8112);
@@
 	\}
-\end{multline}
+\end{equation}
 %
 \begin{figure}[h!]
-\center{\includegraphics[width=1\linewidth]{./img/8_2535_1_other.png}}
+\center{\includegraphics[width=0.6\linewidth]{./img/8_2535_1_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 8 and diameter 2535}
@@
 the set cannot be rotated by $90^\circ$ but still be on lattice~\eqref{eq:char_lattice}:
-\begin{equation}
+$
 	\mathcal{P}_{8y}=
@@
 	\}
-\end{equation}
+$
 
 \begin{figure}[h!]
-	\center{\includegraphics[width=1\linewidth]{./img/8_2400_42_symm.png}}
+	\center{\includegraphics[width=0.6\linewidth]{./img/8_2400_42_symm.png}}
 	\parbox{1\linewidth}{\caption{IPS of cardinality 8 and diameter 2400}

### 2025-08-24 char-4k-bound/char-4k-bound.tex
[4k3] Thank the Referees

--- 2025-03-02 char-4k-bound/char-4k-bound.tex
@@
 
-	In 1945, Erdős gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
+	In 1945, Erdös gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
 	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
@@
 	That was the first lower bound for the diameter ever;
-	however, Erdős did not only estimate the diameter,
+	however, Erdös did not only estimate the diameter,
 	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolas.
@@
 	(in assumption that the set has at east 4 points).
-	All these lower bounds are based on the Erdős's framework --- cofocal hyperbolas.
+	All these lower bounds are based on the Erdos's framework --- cofocal hyperbolas.
 
@@
 
-	In the present paper, we put the power of Erdős's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
+	In the present paper, we put the power of Erdös's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
 	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
@@
 In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological.
-Section 4 is devoted to the connection between Erdős curves and the characteristic.
+Section 4 is devoted to the connection between Erdös curves and characteristic.
 In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.
@@
 
-In this section, we provide rigorous definitions and list some known results.
+In this Section, we provide rigorous definitions and list some known results.
 
@@
 \end{definition}
-The characteristic of an IPS always exists and is unique.
 
@@
 	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
-	For an integer $n$, $|n| < |M_1 M_2|$,
-	we will call the set of points
+	For $N\in\mathbb N_\pm$, $|n| < |M_1 M_2|$
+	we will say that the set of points
 	\begin{equation}
@@
 	\end{equation}
-	the \emph{$n$-th Erdős curve}.
+	is called the \emph{$n$-th Erdos curve}.
 \end{definition}
 
-Obviously, the 0-th Erdős curve is the perpendicular bisector of $M_1 M_2$,
-and all the other Erdős curves are branches of cofocal hyperbolas.
-On Figure~\ref{fig:Erdos_curves_numbered}, the Erdős curves are shown for an edge of length 3.
+Obviously, the 0-th Erdos curve is the perpendicular bisector of $M_1 M_2$,
+and all the other Erdos curves are branches of cofocal hyperbolas.
+On Figure~\ref{fig:Erdos_curves_numbered}, the Erdos curves are shown for an edge of length 3.
 
@@
 \center{\includegraphics[width=0.95\linewidth]{Erdos_curves_numbered.png}}
-\caption{Erdős curves}
+\caption{Erdos curves}
 \label{fig:Erdos_curves_numbered}
@@
 
-Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erdős curves.
-For the sake of brevity, Erdős curves with odd numbers are named odd Erdős curves,
-and the ones with even numbers are named even Erdős curves.
+Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erdos curves.
+For the sake of brevity, Erdos curves with odd numbers are named odd Erdos curves,
+and the ones with even numbers are named even Erdos curves.
 
@@
 \begin{lemma}
-	\cite[Theorem 5.9]{my-pps-linear-bound-2019}
+	\cite[Theorem 3.10]{my-pps-linear-bound-2019}
 	\label{lem:no_distance_one}
@@
 Such notation is made possible by the Grid Theorem.
-Note that we use the asterisk instead of the centered dot to emphasize
-that we do not mean the usual coordinatewise multiplication;
-instead, the denominator is appended to both coordinates,
-but the square root multiplier is appended only to the second one.
 
@@
 %4
-\section{Erdős Curves and Characteristic}
+\section{Erdös Curves and Characteristic}
 
@@
 	\label{thm:weeding_I}
-	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$
-	be an IPS with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$,
-	ant let $M_3$ belong to the $n$-th Erdős curve of the edge $M_1 M_2$.
-	Then, $n$ and the length of $M_1M_2$ have the same parity.
+	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$ be an IPS of semi-general position with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
+Then, the length of the edge $M_1 M_2$ and its $n$-th Erdős curve, which contains the point $M_3$, have the same parity.
 \end{theorem}
@@
 \begin{proof}
-	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and $n = |M_1 M_3| - |M_2 M_3|$.
+	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and let the $n$-th Erdős curve satisfy $|M_1 M_3| - |M_2 M_3| = n$.
 	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is an even number, is also an odd number.
@@
 	\label{thm:weeding_II}
-	Let an IPS $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erdős curve of edge $M_1 M_2$.
-	Then $\operatorname{char} M=8k+7$, $k\in \mathbb{N}_{0}$.
+	Let an IPS in semi-general position $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erdős curve of edge $M_1 M_2$.
+	Then the set $M$ has characteristic $p=8k+7$, $k\in \mathbb{N}_{0}$.
 \end{theorem}
@@
 	\label{lem:no_distance_2}
-	Every set $M \in \overline{\mathfrak{M}}'_n$ with an edge $|M_1M_2|=2$ has the cardinality $n=3$.
+	Every set $M \in \overline{\mathfrak{M}_n}'$ with an edge $|M_1M_2|=2$ has a cardinality $n=3$.
 \end{lemma}
@@
 \begin{proof}
-	Let $M \in \overline{\mathfrak{M}}'_n$ and $M_1, M_2 \in M$.
+	Let $M \in \overline{\mathfrak{M}_n}'$ and $M_1, M_2 \in M$.
 	Then all points in the set $M$ lie on $cr(M_1, M_2)$.
@@
 
-	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}}'_4$.
+	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}_4}'$.
 	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
@@
 	\label{cor:sqrt_5_3}
-	In a triangle $ABC$ with sides $3 \leq a \leq b \leq c$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}}'_4$, the smallest height is at least $(2a-1)^{1/2}$.
+	For the height of a triangle $ABC$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}_4}'$, with sides $3 \leq a \leq b \leq c$, the following estimate holds:
+	$$
+		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2},
+	$$
+	where the difference between the left and right-hand sides increases with increasing $a$.
 \end{corollary}
-
-\begin{proof}
-	Indeed, for $a\geq 3$ the inequality
-	$$
-		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2}
-	$$
-	holds, and
-	the difference between the left and right-hand sides increases with increasing $a$.
-\end{proof}
 
@@
 
-The following lemma is a classical Erdős-style intersection-enumeration one,
+The following lemma is a classical Erdos-style intersection-enumeration one,
 empowered by our Weeding Theorem I.
@@
 	\label{lem:intersection_enumeration}
-	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}}'_n$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$.
+	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}_n}$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$.
 Then $\# M \le |M_1 M_2| \cdot |M_3 M_4| - 2$.
@@
 
-\begin{equation}
-	\begin{array}{rcl}
+\begin{eqnarray}
 \label{d}
-		\overline{d}(2,3)&=& 1,{}\\
-		\overline{d}(2,4)&=& 4,{} \\
-		\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{} \\
-		\overline{d}(2,7)&=& 33,{} \\
-		\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{} \\
-		\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{} \\
-		\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\\
-		\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{} \\
-		\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{} \\
-		\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{} \\
-		\overline{d}(2,28)&=& 10672,{} \\
-		\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{} \\
+\overline{d}(2,3)&=& 1,{}\nonumber\\
+\overline{d}(2,4)&=& 4,{}\nonumber\\
+\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{}\nonumber\\
+\overline{d}(2,7)&=& 33,{}\nonumber\\
+\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{}\nonumber\\
+\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{}\nonumber\\
+\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\nonumber\\
+\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{}\nonumber\\
+\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{}\nonumber\\
+\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{}\nonumber\\
+\overline{d}(2,28)&=& 10672,{}\nonumber\\
+\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{}\nonumber\\
 \overline{d}(2,37) &>& 20\ 000.
-	\end{array}
-\end{equation}
+\end{eqnarray}
 
-Now we are finally ready to proove our main result.
+Now we are finally ready to proof our main result.
 It improves the bound~\eqref{eq:chebsb_lower_bound}
@@
 	\label{thm:main_result}
-	Let $M \in \overline{\mathfrak{M}}'_n$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
+	Let $M \in \overline{\mathfrak{M}}_n'$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
 	Then, for every integer $n \geq 3$, the inequality
@@
 
-	Let us consider $M \in \overline{\mathfrak{M}}'_n$, $n \geq 4$, $\operatorname{diam} M = p$.
+	Let us consider $M \in \overline{\mathfrak{M}_n}'$, $n \geq 4$, $\operatorname{diam} M = p$.
 
@@
 
-	If $n\geq 4$ and $p < 10$, then the claim of the theorem follows immediately from~\eqref{d}.
-	So now we have to consider $m > \frac{6}{5}p^{2/5} > 3$.
+	Now we have to consider $m > \frac{6}{5}p^{2/5}$.
 	Then for any points $A, B \in M \setminus {M_1}$, we have $|AB| > \frac{6}{5}p^{2/5}$.
@@
 	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
-	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10 655$.
+	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10,655$.
 	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.
@@
 \section{Acknowledgements}
-The authors thank Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
+The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
 Dr. A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
 Dr.~A.S.~Usachev for proofreading,
-and E.A. Momot for the ``SciLexic'' project that helped the authors with some word usage issues.
-Also the authors would like to owe a special thanks to the anonymous referees
+and E.A. Momot for the ``SciLexic'' project that helped the author with some word usage issues.
+Also the author would like to owe a special thanks to the anonymous referees
 for their helpful comments and suggestions on the paper.

### 2025-08-24 char-4k-bound/char-4k-bound.tex
[4k3][typos]

@@
 \section{Acknowledgements}
-The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
+The authors thank Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
 Dr. A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
 Dr.~A.S.~Usachev for proofreading,
-and E.A. Momot for the ``SciLexic'' project that helped the author with some word usage issues.
-Also the author would like to owe a special thanks to the anonymous referees
+and E.A. Momot for the ``SciLexic'' project that helped the authors with some word usage issues.
+Also the authors would like to owe a special thanks to the anonymous referees
 for their helpful comments and suggestions on the paper.

### 2025-08-25 char-4k-bound/char-4k-bound.tex
[4k3] Fixed one more runaway equation number

--- 2025-03-12 char-4k-bound/char-4k-bound.tex
@@
 	or, after expansion,
-	\begin{eqnarray}
-		pb^2 + 1&=& 8(-2t^4+8t^3q+4t^3n-4t^3-8t^2q^2-12t^2qn+12t^2q- {}\nonumber\\
-			 &     &-2t^2n^2+6t^2n-3t^2+8tq^2n-4tqn^2-12tqn+6tq- {}\nonumber\\
-			 &     &-2tn^2+3tn-t+4q^2n-2q^2+2qn^2-3qn+q)-  {}\nonumber\\
+	\begin{equation}
+		\label{eq:35.1}
+		\begin{array}{rcl}
+		pb^2 + 1&=& 8(-2t^4+8t^3q+4t^3n-4t^3-8t^2q^2-12t^2qn+12t^2q- {}\\
+			 &     &-2t^2n^2+6t^2n-3t^2+8tq^2n-4tqn^2-12tqn+6tq- {}\\
+			 &     &-2tn^2+3tn-t+4q^2n-2q^2+2qn^2-3qn+q)-  {}\\
 			 &     &-4n(n-1).
-		\label{eq:35.1}
-	\end{eqnarray}
+		\end{array}
+	\end{equation}
 	Among two consecutive natural numbers,

### 2025-08-25 char-4k-bound/char-4k-bound.tex
[4k3] Fix a preposition - hopefully the last proofreading fix

@@
 
-For the sake of completeness, on Figure~\ref{fig:rails_3_8.png}
+For the sake of completeness, in Figure~\ref{fig:rails_3_8.png}
 we give an example of rails set

### 2025-11-16 AvdMomZvolA_examples/ips_examples.tex
Switch back to biblatex2bibitem
[AMZA] Cite 2407.08121 as in print

@@
 
-%\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
-%\addbibresource{../common/notmy.bib}
-%\addbibresource{../common/my.bib}
-%\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
+\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+\addbibresource{../common/notmy.bib}
+\addbibresource{../common/my.bib}
+\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
 
 \usepackage{hyperref}
-\usepackage{cite}
+%\usepackage{cite}
 
@@
 it was noticed~\cite{kurz2008minimum} that such diameter is attained at sets with many points on a straight line;
-for some estimations on this tendency, we also refer the reader to~\cite{solymosi2003note}.
+for some estimations on this tendency, we also refer the reader to~\cite{solymosi2003note};
+(For the recently established connection between the characteristic and minimal possible diameter,
+see~\cite{avdeev-lushina2025australas}.)
 Thus, the following classification was introduced:
@@
 
-%\printbibliography
-%\printbibitembibliography
+\printbibliography
+\printbibitembibliography
+
+\end{document}
 

### 2025-11-16 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Abstract (Rus and Eng)

@@
 \paragraph{Abstract.}
-	A planar integral point set is a set of non-collinear points in plane
-	such that for any pair of the  points the Euclidean distance
-	between them is integral.
-	We discuss the classification of planar integral point sets
-	and provide examples of sets that are not covered by the existent classification.
+
+A planar integral point set (PIPS) is a finite set of non-collinear points in the Euclidean plane
+such that the Euclidean distance between any pair of points is an integer.
+These sets are characterized by their cardinality (finite number of points),
+diameter (maximum pairwise distance), and characteristic (the smallest positive integer \(q\)
+such that all triangular areas are commensurable with \(\sqrt{q}\)).
+The characteristic remains invariant under translations, dilations, reflections, and even the addition or removal of points.
+
+Existing classifications include sets in semi-general position (no three points collinear)
+and general position (no three collinear and no four concyclic).
+Circular sets and facher sets (all but one points on a line) are prominent examples,
+but finding sets in general position is challenging; for instance, the largest known has seven points,
+and no eight-point example exists yet.
+
+This work introduces new examples to advance classification, including rails sets (points on two parallel lines),
+sets with multiple symmetries.
+We also present sets with many shared points that cannot be merged.
+These constructions highlight potential of sequential extensions and limitations in merging sets,
+offering insights into the structure and properties of planar integral point sets.
+
+\paragraph{Аннотация.}
+
+Плоское целоудалённое множество есть конечное множество точек на евклидовой плоскости,
+не содержащееся ни в какой прямой,
+такое что евклидово расстояние между любой парой точек является целым числом.
+Эти множества характеризуются своей мощностью (конечным числом точек),
+диаметром (максимальным попарным расстоянием) и характеристикой (наименьшим положительным целым числом \(q\),
+таким что площади всех треугольников, образованных точками множества, соизмеримы с \(\sqrt{q}\)).
+Характеристика инвариантна относительно при сдвига, растяжения, отражения, а также добавления или удаления точек.
+
+Существующие классификации включают множества в полуобщем положении (никакие три точки не лежат на одной прямой)
+и в общем положении (никакие три точки не лежат на одной прямой и никакие четыре не лежат на одной окружности).
+Классическими примерами являются круговые множества и веерные множества (все точки, кроме одной, лежат на одной прямой),
+однако нахождение множеств общего положения представляет значительные трудности;
+например, наибольшее известное множество имеет семь точек, и пока не найдено множество из восьми точек общего положения.
+
+В данной работе представлены новые примеры для развития классификации,
+включая рельсовые множества (точки на двух параллельных прямых),
+множества с несколькими симметриями и стреловидные конфигурации.
+Мы также рассматриваем множества с большим количеством общих точек, которые нельзя объединить.
+Эти конструкции подчёркивают потенциал последовательных растяжений и ограничения на объединение множеств,
+демонсnрируя новые особенности структуры и свойств плоских целоудалённых множеств.
 

### 2025-11-17 AvdMomZvolA_examples/ips_examples.tex
UDC and MSC
[AMZA] Remove TODOs
[AMZA] Remove unnecessary theorem styles
[AMZA] Memorial note

@@
 
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}
 \theoremstyle{dfn}
 \newtheorem{dfn}{Definition}
-\theoremstyle{remark}
-\newtheorem{remark}{Remark}
 
@@
 \maketitle
+
+To the blessed memory of Professor Semën Samsonovich Kutateladze
+% Я взял такое написание с
+% https://mathscinet.ams.org/mathscinet/MRAuthorID/197072
+% и никоим образом на нём не настаиваю
 
@@
 
+UDC 519.146 + 514.11
+
+MSC 52C10
+
 \section{Introduction}
@@
 and is said to be the \emph{cardinality} of the IPS.
-Furthermore, we can naturally % TODO: naturally/essentially? cf. arxiv/1
+Furthermore, we can naturally
 define the diameter of a finite point set.
@@
 Moreover, we failed to extend it by dilation and looking for extra points on the $x$ axis.
-%
-%TODO: write about points which prevent the set from reflection
-%
+
 \begin{equation}
@@
 The authors thank Dr. Prof. E.M. Semenov and Dr.~A.S.~Usachev for proofreading.
+
+Special remark from N. Avdeev:
+I would like to express my deepest gratitude to the late Professor Semën Samsonovich Kutateladze,
+who played a pivotal role in my scientific career.
+He recommended the journal for my first significant scholarly article~\cite{our-mz-rus-translit}
+and facilitated its submission, marking the start of a pioneering series of works on this topic at our university.
+This contribution is dedicated to his memory.
 

### 2025-11-17 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Remove technical remark
[AMZA] The information about authors

@@
 
-%% Прекрасно понимаю, что следующая команда - дичь и вордописчество, но время не ждёт, время жмёт
 \author{
@@
 
+Авдеев Николай Николаевич
+
+Воронежский государственный университет, кафедра теории функций и геометрии, аспирант
+
+Россия, 394006, г. Воронеж, Университетская пл., 1
+
+E-mail: nickkolok@mail.ru, avdeev@math.vsu.ru
+
+Avdeev Nikolai Nikolaevich
+
+Voronezh State University, Department of Function Theory and Geometry, postgraduate student
+
+Russia, 394006, Voronezh, University sq., 1
+
+Момот Екатерина Александровна
+
+Воронежский государственный университет, кафедра функционального анализа и операторных уравнений, выпускница
+
+Россия, 394006, г. Воронеж, Университетская пл., 1
+
+Momot Ekaterina Alexandrovna
+
+Voronezh State University, Department of Functional Analysis and Operator Equations, graduated from
+
+Russia, 394006, Voronezh, University sq., 1
+
+Зволинский Александр Евгеньевич
+
+Воронежский государственный университет, кафедра теории функций и геометрии, выпускник
+
+Россия, 394006, г. Воронеж, Университетская пл., 1
+
+Zvolinsky Alexandr Evgenievich
+
+Voronezh State University, Department of Function Theory and Geometry, graduated from
+
+Russia, 394006, Voronezh, University sq., 1
+
 \end{document}

### 2025-11-17 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Page geometry - still WIP
[AMZA] Keywords in Russian

@@
 \newtheorem{dfn}{Definition}
+
+\sloppy \textheight = 230mm \textwidth = 155mm \topmargin = -3mm
+\oddsidemargin=3mm \evensidemargin=4pt
 
@@
 	diameter of an integral point set
+
+\paragraph{Ключевые слова.}
+	целоудалённое множество,
+	плоское целоудалённое множество,
+	классификация плоских целоудалённых множеств,
+	диаметр целоудалённого множества
 
@@
 
-\begin{equation}
+\begin{multline*}
 	\mathcal{P}_8=
@@
 	( 29575 ; 0);
-	%\\
+	\\
 	( -2366 ; -8112);
@@
 	\}
-\end{equation}
+\end{multline*}
 %
@@
 the set cannot be rotated by $90^\circ$ but still be on lattice~\eqref{eq:char_lattice}:
-$
+\begin{equation*}
 	\mathcal{P}_{8y}=
@@
 	\}
-$
+\end{equation*}
 

### 2025-11-17 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Remove the old bibliography

@@
 \end{document}
-
-\begin{thebibliography}{99}
-%
-\bibitem{anning1945integral}
-\emph{Anning} \emph{N. H.}, \emph{Erdős} \emph{P.}. Integral distances //
-Bulletin of the American Mathematical Society. — 1945. — Vol. 51, no. 8. —
-Pp. 598–600. — DOI: \href {https://doi.org/10.1090/S0002-9904-1945-08407-9}
-{\nolinkurl {10.1090/S0002-9904-1945-08407-9}}.
-%
-\bibitem{erdos1945integral}
-\emph{Erdős} \emph{P.}. Integral distances // Bulletin of the American Mathematical
-Society. — 1945. — Vol. 51, no. 12. — P. 996. — DOI: \href
-{https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl
-{10.1090/S0002-9904-1945-08490-0}}.%
-\bibitem{kemnitz1988punktmengen}
-\emph{Kemnitz} \emph{A.}. Punktmengen mit ganzzahligen Abständen. — 1988.
-%
-\bibitem{kurz2005characteristic}
-\emph{Kurz} \emph{S.}. On the characteristic of integral point sets in {$\mathbb
-{E}^m$} // Australasian Journal of Combinatorics. — 2006. — Vol. 36. —
-Pp. 241–248. — arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl
-{math/0511704}}.
-%
-\bibitem{kurz2008minimum}
-\emph{Kurz} \emph{S.}, \emph{Wassermann} \emph{A.}. On the minimum diameter
-of plane integral point sets // Ars Combinatoria. — 2011. — Vol. 101. — Pp. 265–287. —
-arXiv: \href {http://arxiv.org/abs/0804.1307} {\nolinkurl {0804.1307}}.
-%
-\bibitem{solymosi2003note}
-\emph{Solymosi} \emph{J.}. Note on integral distances // Discrete \& Computational
-Geometry. — 2003. — Vol. 30, no. 2. — Pp. 337–342. — DOI: \href
-{https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.
-%
-\bibitem{kreisel2008heptagon}
-\emph{Kreisel} \emph{T.}, \emph{Kurz} \emph{S.}. There are integral heptagons, no
-three points on a line, no four on a circle // Discrete \& Computational Geometry. —
-2008. — Vol. 39, no. 4. — Pp. 786–790. — DOI: \href
-{https://doi.org/10.1007/s00454-007-9038-6} {\nolinkurl {10.1007/s00454-007-9038-6}}.
-%
-\bibitem{our-ped-2018}
-\emph{Авдеев} \emph{Н. Н.}. On integral point sets in special position // Некоторые
-вопросы анализа, алгебры, геометрии и математического образования: материалы
-международной молодежной научной школы «Актуальные направления
-математического анализа и смежные вопросы». — 2018. — Vol. 8. — Pp. 5–6.
-%
-\bibitem{our-pmm-2018}
-\emph{Авдеев} \emph{Н. Н.}. Об отыскании целоудалённых множеств
-специального вида // Актуальные проблемы прикладной математики,
-информатики и механики - сборник трудов Международной научной
-конференции. — Научно-исследовательские публикации, 2018. — Pp. 492–498.
-%
-\bibitem{our-vmmsh-2018}
-\emph{Авдеев} \emph{Н. Н.}, \emph{Семёнов} \emph{Е. М.}. Множества точек с
-целочисленными расстояниями на плоскости и в евклидовом пространстве //
-Математический форум (Итоги науки. Юг России). — 2018. — Pp. 217–236.%
-\bibitem{our-vmmsh-2018-translit}
-\emph{Avdeev} \emph{N. N.}, \emph{Semenov} \emph{E. M.}. Integral point sets in
-the plane and Euclidean space (Множества точек с целочисленными расстояниями
-на плоскости и в евклидовом пространстве) // Matematicheskij forum (Itogi nauki.
-Yug Rossii). — 2018. — Pp. 217–236.
-%
-\bibitem{avdeev2019particular}
-\emph{Avdeev} \emph{N. N.}, \emph{Zvolinsky} \emph{R. E.},
-\emph{Momot} \emph{E. A.}. On particular diameter bounds for integral point sets in
-higher dimensions. — 2019. — arXiv: \href {http://arxiv.org/abs/1909.10386}
-{\nolinkurl {1909.10386}}. — URL: \url
-{https://ui.adsabs.harvard.edu/abs/2019arXiv190910386A/abstract}.
-%
-\bibitem{kurz2013constructing}
-Constructing $7$-clusters / S. Kurz [et al.] // Serdica Journal of Computing. — 2014. —
-Vol. 8, no. 1. — Pp. 47–70. — arXiv: \href {http://arxiv.org/abs/1312.2318} {\nolinkurl
-{1312.2318}}.
-%
-\bibitem{antonov2008maximal}
-\emph{Antonov} \emph{A. R.}, \emph{Kurz} \emph{S.}. Maximal integral point sets
-over {$\mathbb {Z}^2$} // International Journal of Computer Mathematics. —
-2008. — Vol. 87, no. 12. — Pp. 2653–2676. — DOI: \href
-{https://doi.org/10.1080/00207160902993636} {\nolinkurl
-{10.1080/00207160902993636}}. — arXiv: \href {http://arxiv.org/abs/0804.1280}
-{\nolinkurl {0804.1280}}.
-%
-\bibitem{solymosi2010question}
-\emph{Solymosi} \emph{J.}, \emph{De Zeeuw} \emph{F.}. On a question of Erdős
-and Ulam // Discrete \& Computational Geometry. — 2010. — Vol. 43, no. 2. —
-Pp. 393–401. — arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl
-{0806.3095}}.
-\end{thebibliography}
-
-The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
-
-\end{document}

### 2025-11-17 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Adjust definitions

@@
 
-\theoremstyle{dfn}
-\newtheorem{dfn}{Definition}
+\newcommand{\Definition}[1]{\smallskip{\sc Definition#1.}}
 
@@
 
-\begin{dfn}\label{dfn1}
+\Definition{ 1}\label{dfn1}
 	A planar integral point set (PIPS) is a set $\mathcal{P}$
@@
 	between points $P_{1}$ and $P_{2}$ is integral.
-\end{dfn}
 
@@
 
-\begin{dfn}
+\Definition{ 2}
 	The diameter of the integral point set $\mathcal{P}$ is defined as
@@
 	$
-\end{dfn}
 
-\begin{dfn}(\cite{kemnitz1988punktmengen,kurz2005characteristic})
+\Definition{ 3}(\cite{kemnitz1988punktmengen,kurz2005characteristic})
 	The characteristic of a planar integral point set $M$ is the least positive integer $q$
@@
 	is commensurable with $\sqrt{q}$.
-\end{dfn}
+
 Characteristic of an IPS does not change if the set is moved, dilated ot flipped;
@@
 Thus, the following classification was introduced:
-\begin{dfn}
+
+\Definition{ 4}
 	A PIPS $M$ is said to be in \emph{semi-general position}
 	if no three points of $M$ are located in a straight line.
-\end{dfn}
 
 The most dominating examples of PIPS in semi-general position are circular sets.
-\begin{dfn}
+
+\Definition{ 5}
 	A planar integral point sets that is situated on a circle is said to be a \textit{circular}
 	point set.
-\end{dfn}
 
 So, the following constraint appeared.
-\begin{dfn}
+
+\Definition{ 6}
 	A planar integral point set $M$ is said to be in \emph{general position}
@@
 	and no four points of $M$ are located in a circle.
-\end{dfn}
 
@@
 
-\begin{dfn}
+\Definition{ 7}
 	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
 	a \textit{facher} set.
-\end{dfn}
+
 Facher sets are predominant examples of planar integral point sets.
@@
 
-\begin{dfn}[\cite{avdeev2019particular}]
+\Definition{ 8}[\cite{avdeev2019particular}]
 	A non-facher planar integral point sets situated in two parallel straight lines
 	is called a \textit{rails} set.
-\end{dfn}
 

### 2025-11-17 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Russian article caption
[AMZA] Page geometry

@@
 	диаметр целоудалённого множества
+
+Некоторые примеры плоских целоудалённых множеств и их классификация
 
@@
 Figure~\ref{42_symm.png}:
-\begin{multline}
+\begin{multline*}
 	\mathcal{P}_{42}=\sqrt{154}/{1} * \{
@@
 	\}
-\end{multline}
+\end{multline*}
 
@@
 Figure~\ref{46_symm.png}:
-\begin{multline}
+\begin{multline*}
 	\mathcal{P}_{46}=
@@
 		(\pm328080480; 0);
+		\\
 		(\pm268861320; 0);
 		(\pm194124840; 0);
-		\\
 		(\pm182407680; 0);
@@
 		(\pm284695320; 0);
+		\\
 		(\pm1506967020; 0);
@@
 		(\pm875077320; 0);
-		\\
 		(\pm648982880; 0);
 		(\pm585224640; 0);
+		\\
 		(\pm452799620; 0);
@@
 		(\pm116181975; 0);
+		(\pm428602020; 0);
 		\\
-		(\pm428602020; 0);
 		(\pm56222880; 0);
@@
 Figure~\ref{48_symm.png}:
-\begin{multline}
+\begin{multline*}
 	\mathcal{P}_{48}=
@@
 		( \pm6365901360 ; -437018400);
+		\\
 		( \pm10022288640; 0);
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.6\linewidth]{./img/9_2890_1_02e6afb1c19846c6c4debb44e52cc468_other.png}}
+\center{\includegraphics[width=1\linewidth]{./img/9_2890_1_02e6afb1c19846c6c4debb44e52cc468_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 9 and diameter 2890}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.85\linewidth]{./img/19_20663808074_1_0070b97f26bb194a7de2ddc37ee38b08_symm.png}}
+\center{\includegraphics[width=1\linewidth]{./img/19_20663808074_1_0070b97f26bb194a7de2ddc37ee38b08_symm.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 19 and diameter 20663808074}
@@
 		( 987843675 ; \pm1146332880);
-		%\\
+		\\
 		( \pm729918777 ; 0);
 		( \pm972103809 ; 0);
-		\\
+		%\\
 		( \pm1113030324 ; 0);
 		( \pm1400170149 ; 0);
-		%\\
+		\\
 		( \pm3137217825 ; 0);
@@
 On Figure~\ref{17_1024296_1_639b}, the following IPS is shown:
-$
+\begin{multline*}
 	\mathcal{P}_{17}=
@@
 		( -694668 ; 0);
-		$\\$
+		( 15457 ; 0);
+		\\
 		( -71487 ; 0);
@@
 		( -14943 ; 0);
-		( 15457 ; 0);
 		%\\
@@
 		( 125307 ; 0);
-		( 172207 ; 0)
-		( 329628 ; 0);
+		( 172207 ; 0);
+		( 329628 ; 0)
 	\}
-$
+\end{multline*}
 Note that the axis of symmetry for $\mathcal{P}_{17}$ is the $x$ axis;
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.7\linewidth]{./img/17_1024296_1_639b7a0452227fde1aceabf02e8778e4_other.png}}
+\center{\includegraphics[width=1\linewidth]{./img/17_1024296_1_639b7a0452227fde1aceabf02e8778e4_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 17 and diameter 1024296}
@@
 \begin{figure}[h!]
-	\center{\includegraphics[width=0.7\linewidth]{./img/10_2431_1_0606659e825917be0c3c431c1375bf21_other.png}}
+	\center{\includegraphics[width=1\linewidth]{./img/10_2431_1_0606659e825917be0c3c431c1375bf21_other.png}}
 	\parbox{1\linewidth}{\caption{IPS of cardinality 10 and diameter 2431}
@@
 \begin{figure}[h!]
-	\center{\includegraphics[width=0.7\linewidth]{./img/15_19203744_1_80dbf8e05b25f69b8366e50d1d7463f6_other.png}}
+	\center{\includegraphics[width=1\linewidth]{./img/15_19203744_1_80dbf8e05b25f69b8366e50d1d7463f6_other.png}}
 	\parbox{1\linewidth}{\caption{IPS of cardinality 15 and diameter 19203744}
@@
 Figure~\ref{10_2431_1_0606}:
-$
+\begin{equation*}
 	\mathcal{P}_{10} =
@@
 	\}
-$
+\end{equation*}
 
 Figure~\ref{15_19203744_1_80db}:
-$
+\begin{multline*}
 	\mathcal{P}_{15} =
@@
 		( 5385600 ; \pm2625480);
-		%\\
+		\\
 		( -6283200 ; 0);
@@
 		( -635040 ; 0);
-		$\\$
 		( 1884960 ; 0);
-		%\\
+		\\
 		( 3147210 ; 0);
@@
 	\}
-$
+\end{multline*}
 
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.6\linewidth]{./img/8_2535_1_other.png}}
+\center{\includegraphics[width=0.9\linewidth]{./img/8_2535_1_other.png}}
 \parbox{1\linewidth}{\caption{IPS of cardinality 8 and diameter 2535}
@@
 \begin{figure}[h!]
-	\center{\includegraphics[width=0.6\linewidth]{./img/8_2400_42_symm.png}}
+	\center{\includegraphics[width=0.9\linewidth]{./img/8_2400_42_symm.png}}
 	\parbox{1\linewidth}{\caption{IPS of cardinality 8 and diameter 2400}

### 2025-11-17 AvdMomZvolA_examples/ips_examples.tex
[AMZA] Reincorporate the bibliography

@@
 
-\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
-\addbibresource{../common/notmy.bib}
-\addbibresource{../common/my.bib}
-\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
+%\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
+%\addbibresource{../common/notmy.bib}
+%\addbibresource{../common/my.bib}
+%\usepackage{../../../biblatex2bibitem/biblatex2bibitem}
 
 \usepackage{hyperref}
-%\usepackage{cite}
+\usepackage{cite}
 
@@
 
-\printbibliography
-\printbibitembibliography
-
 Авдеев Николай Николаевич
@@
 
+\begin{thebibliography}{99}
+{}
+\bibitem{anning1945integral}
+\emph{Anning} \emph{N. H.}, \emph{Erdős} \emph{P.} Integral distances // Bulletin of
+the American Mathematical Society. — 1945. — Vol. 51, no. 8. — Pp. 598–600. — DOI: \href
+{https://doi.org/10.1090/S0002-9904-1945-08407-9} {\nolinkurl
+{10.1090/S0002-9904-1945-08407-9}}.
+{}
+\bibitem{erdos1945integral}
+\emph{Erdős} \emph{P.} Integral distances // Bulletin of the American Mathematical
+Society. — 1945. — Vol. 51, no. 12. — P. 996. — DOI: \href
+{https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl
+{10.1090/S0002-9904-1945-08490-0}}.
+{}
+\bibitem{kemnitz1988punktmengen}
+\emph{Kemnitz} \emph{A.} Punktmengen mit ganzzahligen Abständen. — 1988.
+{}
+\bibitem{kurz2005characteristic}
+\emph{Kurz} \emph{S.} On the characteristic of integral point sets in {$\mathbb
+{E}^m$} // Australasian Journal of Combinatorics. — 2006. — Vol. 36. — Pp. 241–248. —
+arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl {math/0511704}}.
+{}
+\bibitem{kurz2008minimum}
+\emph{Kurz} \emph{S.}, \emph{Wassermann} \emph{A.} On the minimum diameter of
+plane integral point sets // Ars Combinatoria. — 2011. — Vol. 101. — Pp. 265–287. — arXiv:
+\href {http://arxiv.org/abs/0804.1307} {\nolinkurl {0804.1307}}.
+{}
+\bibitem{solymosi2003note}
+\emph{Solymosi} \emph{J.} Note on integral distances // Discrete \& Computational
+Geometry. — 2003. — Vol. 30, no. 2. — Pp. 337–342. — DOI: \href
+{https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.{}
+\bibitem{avdeev-lushina2025australas}
+\emph{Avdeev} \emph{N.}, \emph{Lushina} \emph{E.} On the characteristic and diameter
+of planar integral point sets // Australasian Journal of Combinatorics. — (in press). — DOI:
+\href {https://doi.org/10.48550/arXiv.2407.08121} {\nolinkurl
+{10.48550/arXiv.2407.08121}}. — arXiv: \href {http://arxiv.org/abs/2407.08121} {\nolinkurl
+{2407.08121}}. — URL: \url
+{https://ui.adsabs.harvard.edu/abs/2024arXiv240708121A/abstract}.
+{}
+\bibitem{kreisel2008heptagon}
+\emph{Kreisel} \emph{T.}, \emph{Kurz} \emph{S.} There are integral heptagons, no three
+points on a line, no four on a circle // Discrete \& Computational Geometry. — 2008. —
+Vol. 39, no. 4. — Pp. 786–790. — DOI: \href {https://doi.org/10.1007/s00454-007-9038-6}
+{\nolinkurl {10.1007/s00454-007-9038-6}}.
+{}
+\bibitem{our-ped-2018}
+\emph{Авдеев} \emph{Н. Н.} On integral point sets in special position // Некоторые
+вопросы анализа, алгебры, геометрии и математического образования: материалы
+международной молодежной научной школы «Актуальные направления математического
+анализа и смежные вопросы». — 2018. — Vol. 8. — Pp. 5–6.
+{}
+\bibitem{our-pmm-2018}
+\emph{Авдеев} \emph{Н. Н.} Об отыскании целоудалённых множеств специального
+вида // Актуальные проблемы прикладной математики, информатики и механики -
+сборник трудов Международной научной конференции. — Научно-исследовательские
+публикации, 2018. — Pp. 492–498.
+{}
+\bibitem{our-vmmsh-2018-translit}
+\emph{Avdeev} \emph{N. N.}, \emph{Semenov} \emph{E. M.} Integral point sets in the
+plane and Euclidean space (Множества точек с целочисленными расстояниями на
+плоскости и в евклидовом пространстве) // Matematicheskij forum (Itogi nauki. Yug
+Rossii). — 2018. — Pp. 217–236.
+{}
+\bibitem{avdeev2019particular}
+\emph{Avdeev} \emph{N. N.}, \emph{Zvolinsky} \emph{R. E.},
+\emph{Momot} \emph{E. A.} On particular diameter bounds for integral point sets in higher
+dimensions // Вестн. ВГУ. Серия: Физика. Математика. — 2025. — Vol. 1. — Pp. 62–77. —
+arXiv: \href {http://arxiv.org/abs/1909.10386} {\nolinkurl {1909.10386}}. — URL: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190910386A/abstract}.
+{}
+\bibitem{kurz2013constructing}
+Constructing $7$-clusters / S. Kurz [et al.] // Serdica Journal of Computing. — 2014. — Vol. 8,
+no. 1. — Pp. 47–70. — arXiv: \href {http://arxiv.org/abs/1312.2318} {\nolinkurl {1312.2318}}.
+{}
+\bibitem{antonov2008maximal}
+\emph{Antonov} \emph{A. R.}, \emph{Kurz} \emph{S.} Maximal integral point sets over
+{$\mathbb {Z}^2$} // International Journal of Computer Mathematics. — 2008. — Vol. 87,
+no. 12. — Pp. 2653–2676. — DOI: \href {https://doi.org/10.1080/00207160902993636}
+{\nolinkurl {10.1080/00207160902993636}}. — arXiv: \href {http://arxiv.org/abs/0804.1280}
+{\nolinkurl {0804.1280}}.{}
+\bibitem{solymosi2010question}
+\emph{Solymosi} \emph{J.}, \emph{De Zeeuw} \emph{F.} On a question of Erdős and
+Ulam // Discrete \& Computational Geometry. — 2010. — Vol. 43, no. 2. — Pp. 393–401. —
+arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl {0806.3095}}.
+{}
+\bibitem{our-mz-rus-translit}
+\emph{Avdeev} \emph{N. N.}, \emph{Semenov} \emph{E. M.} On point sets in the plane
+with integral distances (О множествах точек на плоскости с целочисленными
+расстояниями) // Matematicheskie zametki. — 2016. — Vol. 100, no. 5. — Pp. 757–761.
+\end{thebibliography}
+
 \end{document}

### 2025-11-19 char-4k-bound/char-4k-bound.tex
[4k3] First bunch of edits

@@
 \paragraph{Abstract.}
-A point set $M$ in Euclidean plane is called an integral point set in semi-general position if all the distances between the
+A point set $M$ in a Euclidean plane is called an integral point set in semi-general position if all the distances between the
 elements of $M$ are integers, and $M$ does not contain collinear triples.
-In this paper we improve the lower bound for diameter of such sets in the particular case
+In this paper we improve the lower bound for the diameter of such sets in the particular case
 when the characteristic of the set is of the form $4k+1$ or $4k+2$.

### 2025-11-27 char-4k-bound/char-4k-bound.tex
[4k3] Fuller address
[4k3] Erd{\H o}s

@@
 	\\
-	\textit{Voronezh State University}
+	\textit{Voronezh State University, Voronezh, Russia}
 }
@@
 
-	In 1945, Erdős gave an elegant proof that every IPS is finite~\cite{anning1945integral,erdos1945integral} .
-	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.
-Applying the same argument to the line segment $M_1M_3$, Erdös concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.
+	In 1945, Erd{\H o}s gave an elegant proof that every IPS is finite~\cite{anning1945integral,erdos1945integral} .
+	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for the length of the line segment $M_1M_2$.
+Applying the same argument to the line segment $M_1M_3$, Erd{\H o}s concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.
 
@@
 	That was the first lower bound for the diameter ever;
-	however, Erdős did not only estimate the diameter,
+	however, Erd{\H o}s did not only estimate the diameter,
 	but also established the toolset for further investigation.
@@
 
-	In the present paper, we combine the power of Erdős's and Kemnitz's approaches together.
-Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
+	In the present paper, we combine the power of Erd{\H o}s's and Kemnitz's approaches together.
+Due to the squarefree nature of the characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
 	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
@@
 In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological.
-Section 4 describes a connection between Erdős curves and the characteristic.
+Section 4 describes a connection between Erd{\H o}s curves and the characteristic.
 In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.
@@
 	\end{equation}
-	the \emph{$n$-th Erdős curve}.
+	the \emph{$n$-th Erd{\H o}s curve}.
 \end{definition}
 
-Obviously, the 0-th Erdős curve is the perpendicular bisector of $M_1 M_2$,
-and all the other Erdős curves are branches of cofocal hyperbolas.
-On Figure~\ref{fig:Erdos_curves_numbered}, the Erdős curves are shown for an edge of length 3.
+Obviously, the 0-th Erd{\H o}s curve is the perpendicular bisector of $M_1 M_2$,
+and all the other Erd{\H o}s curves are branches of cofocal hyperbolas.
+In Figure~\ref{fig:Erdos_curves_numbered}, the Erd{\H o}s curves are shown for an edge of length 3.
 
@@
 \center{\includegraphics[width=0.95\linewidth]{Erdos_curves_numbered.png}}
-\caption{Erdős curves}
+\caption{Erd{\H o}s curves}
 \label{fig:Erdos_curves_numbered}
@@
 
-Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erdős curves.
-For the sake of brevity, Erdős curves with odd numbers are named odd Erdős curves,
-and the ones with even numbers are named even Erdős curves.
+Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erd{\H o}s curves.
+For the sake of brevity, Erd{\H o}s curves with odd numbers are named odd Erd{\H o}s curves,
+and the ones with even numbers are named even Erd{\H o}s curves.
 
@@
 %4
-\section{Erdős Curves and Characteristic}
+\section{Erd{\H o}s Curves and Characteristic}
 
@@
 	be an IPS with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$,
-	ant let $M_3$ belong to the $n$-th Erdős curve of the edge $M_1 M_2$.
+	ant let $M_3$ belong to the $n$-th Erd{\H o}s curve of the edge $M_1 M_2$.
 	Then, $n$ and the length of $M_1M_2$ have the same parity.
@@
 	\label{thm:weeding_II}
-	Let an IPS $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erdős curve of edge $M_1 M_2$.
+	Let an IPS $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erd{\H o}s curve of edge $M_1 M_2$.
 	Then $\operatorname{char} M=8k+7$, $k\in \mathbb{N}_{0}$.
@@
 	Indeed, let $|M_1 M_2|=2q$, $q \in \mathbb{N}$.
-	For the $n$-th Erdős curve we have $|M_1 M_3|-|M_2 M_3|=n$.
+	For the $n$-th Erd{\H o}s curve we have $|M_1 M_3|-|M_2 M_3|=n$.
 	According to Lemma~\ref{lem:pre_weeding_II}, the difference $n=|M_1 M_3|-|M_2 M_3|=2q-s$,
@@
 
-The following result is a classical Erdős-style intersection-enumeration lemma,
+The following result is a classical Erd{\H o}s-style intersection-enumeration lemma,
 empowered by our Weeding Theorem I.

### 2025-11-28 char-4k-bound/char-4k-bound.tex
[4k3] Fix Acknowledgements
[4k3] Incorporate the bibliography

@@
 	(assuming that the set has at least 4 points).
-	All these lower bounds are based on the Erdős's framework --- cofocal hyperbolas.
-
-	Meanwhile in 1988 Kemnitz introduced~\cite{kemnitz1988punktmengen}
-	a characteristic of an IPS, that is a squarefree integer $q$ such as the area
-	of a triangle formed by any triple of points from the IPS is comparable with $\sqrt{q}$.
+	All these lower bounds are based on Erd{\H o}s's framework: cofocal hyperbolas.
+
+	Meanwhile in 1988, Kemnitz~\cite{kemnitz1988punktmengen} introduced a characteristic for an IPS.
+	The charateristic of an IPS is the one and only squarefree integer $q$ which is comparable
+	with the area of the triangle formed by any triple of points from the IPS.
 	(Indeed, Kemnitz proved this fact for any point set with rational distances ---
@@
 	up to $n=122$ by exhaustive computer search~\cite[Subsection 4.2]{kurz2008minimum}.
-	(Taking the characteristic into consideration allows to boost such search significantly ---
+	(Taking the characteristic into consideration allows one to boost such search significantly; this is
 	basically because all triples of an IPS form triangles with equal characteristic.)
 
-	For generalization in higher dimensions, we refer the reader to~\cite{nozaki2013lower}.
+	For generalization to higher dimensions, we refer the reader to~\cite{nozaki2013lower}.
 
@@
 
-For the sake of brevity, the following notations are used for sets
-of positive integers, non-negative integers and all the integers resp.:
+For the sake of brevity, the following notation is used for sets
+of positive integers, non-negative integers and all the integers respectively:
 \begin{equation}
@@
 In the present paper, we mostly focus on IPS in semi-general position.
-However, the next restriction step can be done as the following.
+However, the next restriction step can be done as follows.
 
@@
 When we think about IPS with characteristic $4k+1$,
-the very first example that comes to our mind is Egyptian triangle,
-that is the triangle with sides $(3;4;5)$.
+the very first example that comes to our mind is the Egyptian triangle,
+that is the triangle with sides $(3;4;5)$. The
 Egyptian triangle is obviously an IPS with characteristic $1 = 4 \cdot 0 +1$.
@@
 	( 4558 ; 0)
-\}
+\}.
 \end{multline}
@@
 In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$.
-This contradicts that $k\in\mathbb{N}$.
-	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$. This also contradicts that $s,k\in\mathbb{N}$.
+This contradicts the fact that $k\in\mathbb{N}$.
+	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$. This also contradicts $s,k\in\mathbb{N}$.
 
 	Thus, equation~\eqref{eq: 42} has no positive integer solutions.
-	The obtained contradiction completes the proof.
+	This contradiction completes the proof.
 \end{proof}
@@
 		a^2- \left(\frac{c^2+a^2-b^2}{2c} \right)^2
+		.
 	\end{multline}
@@
 		=
-		\\=
+		\\
 		 4\left(\frac{|M_1 M_2| \cdot |M_3 M_4|}{4} - \frac{|M_1 M_2|}{2} - \frac{|M_3 M_4|}{2}+1\right) + 4 + 4
 		=
-		\\=
+		\\
 		|M_1 M_2| \cdot |M_3 M_4| - 2 |M_1 M_2| - 2 |M_3 M_4| + 4 + 4 + 4
@@
 		\leq \\
-		\leq |M_1 M_2| \cdot |M_3 M_4| - 2
+		|M_1 M_2| \cdot |M_3 M_4| - 2
 		.
@@
 		=
-		\\=
+		\\
 		(|M_1 M_2| - 2)(|M_3 M_4| - 1) + 6
 		=
-		\\=
+		\\
 		|M_1 M_2| \cdot |M_3 M_4| - |M_1 M_2| - 2 |M_3 M_4| + 2 + 6
@@
 		\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{} \\
-		\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{} \\
+		\overline{d}(2,10)&=& \dots =\overline{d}(2,12)= 105,{} \\
 		\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\\
-		\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{} \\
-		\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{} \\
-		\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{} \\
+		\overline{d}(2,15)&=& \dots = \overline{d}(2,18) = 735,{} \\
+		\overline{d}(2,19)&=& \dots = \overline{d}(2,24) = 1995,{} \\
+		\overline{d}(2,25)&=&  \dots = \overline{d}(2,27)= 9555,{} \\
 		\overline{d}(2,28)&=& 10672,{} \\
-		\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{} \\
+		\overline{d}(2,29)&=&  \dots = \overline{d}(2,36) = 13975,{} \\
 		\overline{d}(2,37) &>& 20\ 000.
@@
 		\min_{A,B \in M} |AB| = |M_1 M_2|
+		,
 	\end{equation*}
@@
 		\min_{A,B \in M \setminus {M_1}} |AB| = |M_3 M_4| = m
+		.
 	\end{equation*}
@@
 \section{Conclusion}
-	The bound proved above (as well as~\eqref{eq:chebsb_lower_bound}) may appear to be far from precise values of $\overline{d}(2,n)$.
-	However, it's easy to see that values of $d(2;n)$ tend to repeat often;
+	The bound proved above (as well as~\eqref{eq:chebsb_lower_bound}) may appear to be far from the precise values of $\overline{d}(2,n)$.
+	However, it is easy to see that values of $d(2;n)$ tend to repeat often;
 	thus, it is rather not unrealistic that the bounds may converge to the precise values.
@@
 
-\section{Acknowledgements}
+\section*{Acknowledgements}
 The authors thank Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
@@
 and E.A. Momot for the ``SciLexic'' project that helped the authors with some word usage issues.
-Also the authors would like to owe a special thanks to the anonymous referees
+Also the authors owe special thanks to the anonymous referees
 for their helpful comments and suggestions on the paper.
 
-%\printbibliography
-
-%\printbibitembibliography{}
-
-\input{references.tex}
+\begin{thebibliography}{99}
+%
+\bibitem{anning1945integral}
+N. H. Anning and P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{8} (1945), 598–600, \textsc{doi}: \href
+{https://doi.org/10.1090/S0002-9904-1945-08407-9} {\nolinkurl {10.1090/S0002-9904-1945-08407-9}}.
+%
+\bibitem{antonov2008maximal}
+A. R. Antonov and S. Kurz, Maximal integral point sets over {$\mathbb {Z}^2$}, \emph{Int. J. Comput. Math.} 87.\textbf{12}
+(2008), 2653–2676, \textsc{doi}: \href {https://doi.org/10.1080/00207160902993636} {\nolinkurl {10.1080/00207160902993636}},
+arXiv: \href {http://arxiv.org/abs/0804.1280} {\nolinkurl {0804.1280}}.
+%
+\bibitem{our-vmmsh-2018-translit}
+N. N. Avdeev and E. M. Semenov, Integral point sets in the plane and Euclidean space (Множества точек с целочисленными
+расстояниями на плоскости и в евклидовом пространстве), \emph{Matematicheskij forum (Itogi nauki. Yug Rossii)} (2018),
+217–236.
+%
+\bibitem{avdeev2019particular}
+N. N. Avdeev, R. E. Zvolinsky, and E. A. Momot, On particular diameter bounds for integral point sets in higher dimensions
+Proceedings of Voronezh State University. Series: Physics. Mathematics. 2025. No 1. Pp. 62-77.
+arXiv: \href {http://arxiv.org/abs/1909.10386} {\nolinkurl {1909.10386}}.
+%
+\bibitem{my-semi-general-5-4-bound-2019}
+N. Avdeev, On diameter bounds for planar integral point sets in semi-general position (2019), arXiv: \href
+{http://arxiv.org/abs/1907.09331} {\nolinkurl {1907.09331}}, \textsc{url}: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190709331A/abstract}.%
+\bibitem{my-pps-linear-bound-2019}
+N. Avdeev, On existence of integral point sets and their diameter bounds, \emph{Australas. J. Combin.} 77.\textbf{1} (2020),
+100–116, arXiv: \href {http://arxiv.org/abs/1906.11926} {\nolinkurl {1906.11926}}, \textsc{url}: \url
+{https://ui.adsabs.harvard.edu/abs/2019arXiv190611926A/abstract}.
+%
+\bibitem{costa2013valid}
+A. Costa, Valid constraints for the point packing in a square problem, \emph{Discrete Appl. Math.} 161.\textbf{18} (2013),
+2901–2909.
+%
+\bibitem{erdos1945integral}
+P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{12} (1945), 996, \textsc{doi}: \href
+{https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl {10.1090/S0002-9904-1945-08490-0}}.
+%
+\bibitem{harborth1993upper}
+H. Harborth, A. Kemnitz, and M. Möller, An upper bound for the minimum diameter of integral point sets, \emph{Discrete
+Comput. Geom.} 9.\textbf{4} (1993), 427–432, \textsc{doi}: \href {https://doi.org/10.1007/bf02189331} {\nolinkurl
+{10.1007/bf02189331}}.
+%
+\bibitem{huff1948diophantine}
+G. B. Huff, Diophantine problems in geometry and elliptic ternary forms, \emph{Duke Math. J.} 15.\textbf{2} (1948), 443–453,
+\textsc{doi}: \href {https://doi.org/10.1215/S0012-7094-48-01543-9} {\nolinkurl {10.1215/S0012-7094-48-01543-9}}.
+%
+\bibitem{kemnitz1988punktmengen}
+A. Kemnitz, Punktmengen mit ganzzahligen Abständen, 1988.
+%
+\bibitem{kreisel2008heptagon}
+T. Kreisel and S. Kurz, There are integral heptagons, no three points on a line, no four on a circle, \emph{Discrete Comput.
+Geom.} 39.\textbf{4} (2008), 786–790, \textsc{doi}: \href {https://doi.org/10.1007/s00454-007-9038-6} {\nolinkurl
+{10.1007/s00454-007-9038-6}}.
+%
+\bibitem{kurz2005characteristic}
+S. Kurz, On the characteristic of integral point sets in {$\mathbb {E}^m$}, \emph{Australas. J. Combin.} 36 (2006), 241–248,
+arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl {math/0511704}}.
+%
+\bibitem{kurz2008bounds}
+S. Kurz and R. Laue, Bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 39 (2007), 233–240,
+arXiv: \href {http://arxiv.org/abs/0804.1296} {\nolinkurl {0804.1296}}.
+%
+\bibitem{kurz2008minimum}
+S. Kurz and A. Wassermann, On the minimum diameter of plane integral point sets, \emph{Ars Combin.} 101 (2011), 265–287,
+arXiv: \href {http://arxiv.org/abs/0804.1307} {\nolinkurl {0804.1307}}.
+%
+\bibitem{kurz2013constructing}
+S. Kurz et al., Constructing $7$-clusters, \emph{Serdica J. Comput.} 8.\textbf{1} (2014), 47–70, arXiv: \href
+{http://arxiv.org/abs/1312.2318} {\nolinkurl {1312.2318}}.
+%
+\bibitem{markot2005newverified}
+M. C. Markót and T. Csendes, A new verified optimization technique for the "packing circles in a unit square" problems,
+\emph{SIAM J. Optim.} 16.\textbf{1} (2005), 193–219.
+%
+\bibitem{momot2022example}
+E. A. Momot, R. E. Zvolinskiy, and A. E. Zvolinskiy, An example of rails integral point set with cardinality 106, Прикладная
+математика и фундаментальная информатика, 2022, 114–115.
+%
+\bibitem{nozaki2013lower}
+H. Nozaki, Lower bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 56 (2013), 139–143.
+%
+\bibitem{solymosi2003note}
+J. Solymosi, Note on integral distances, \emph{Discrete Comput. Geom.} 30.\textbf{2} (2003), 337–342, \textsc{doi}: \href
+{https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.
+%
+\bibitem{solymosi2010question}
+J. Solymosi and F. De Zeeuw, On a question of Erdős and Ulam, \emph{Discrete Comput. Geom.} 43.\textbf{2} (2010), 393–401,
+arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl {0806.3095}}.
+%
+\bibitem{our-ped-2018}
+Н. Н. Авдеев, On integral point sets in special position, \emph{Некоторые вопросы анализа, алгебры, геометрии и
+математического образования: материалы международной молодежной научной школы «Актуальные направления
+математического анализа и смежные вопросы»} 8 (2018), 5–6.%
+%
+\bibitem{our-pmm-2018}
+Н. Н. Авдеев, Об отыскании целоудалённых множеств специального вида, Актуальные проблемы прикладной математики,
+информатики и механики - сборник трудов Международной научной конференции. Научно-исследовательские публикации,
+2018, 492–498.
+%
+\bibitem{smurov1998stripcoverings}
+М. Смуров and А. Спивак, Покрытия полосками, \emph{Квант} \textbf{5} (1998), 6–12.
+\end{thebibliography}
 

### 2025-11-28 char-4k-bound/char-4k-bound.tex
[4k3] Bibliography - n-dashes

@@
 \bibitem{anning1945integral}
-N. H. Anning and P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{8} (1945), 598–600, \textsc{doi}: \href
+N. H. Anning and P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{8} (1945), 598--600, \textsc{doi}: \href
 {https://doi.org/10.1090/S0002-9904-1945-08407-9} {\nolinkurl {10.1090/S0002-9904-1945-08407-9}}.
@@
 A. R. Antonov and S. Kurz, Maximal integral point sets over {$\mathbb {Z}^2$}, \emph{Int. J. Comput. Math.} 87.\textbf{12}
-(2008), 2653–2676, \textsc{doi}: \href {https://doi.org/10.1080/00207160902993636} {\nolinkurl {10.1080/00207160902993636}},
+(2008), 2653--2676, \textsc{doi}: \href {https://doi.org/10.1080/00207160902993636} {\nolinkurl {10.1080/00207160902993636}},
 arXiv: \href {http://arxiv.org/abs/0804.1280} {\nolinkurl {0804.1280}}.
@@
 расстояниями на плоскости и в евклидовом пространстве), \emph{Matematicheskij forum (Itogi nauki. Yug Rossii)} (2018),
-217–236.
+217--236.
 %
@@
 N. Avdeev, On existence of integral point sets and their diameter bounds, \emph{Australas. J. Combin.} 77.\textbf{1} (2020),
-100–116, arXiv: \href {http://arxiv.org/abs/1906.11926} {\nolinkurl {1906.11926}}, \textsc{url}: \url
+100--116, arXiv: \href {http://arxiv.org/abs/1906.11926} {\nolinkurl {1906.11926}}, \textsc{url}: \url
 {https://ui.adsabs.harvard.edu/abs/2019arXiv190611926A/abstract}.
@@
 A. Costa, Valid constraints for the point packing in a square problem, \emph{Discrete Appl. Math.} 161.\textbf{18} (2013),
-2901–2909.
+2901--2909.
 %
@@
 H. Harborth, A. Kemnitz, and M. Möller, An upper bound for the minimum diameter of integral point sets, \emph{Discrete
-Comput. Geom.} 9.\textbf{4} (1993), 427–432, \textsc{doi}: \href {https://doi.org/10.1007/bf02189331} {\nolinkurl
+Comput. Geom.} 9.\textbf{4} (1993), 427--432, \textsc{doi}: \href {https://doi.org/10.1007/bf02189331} {\nolinkurl
 {10.1007/bf02189331}}.
@@
 \bibitem{huff1948diophantine}
-G. B. Huff, Diophantine problems in geometry and elliptic ternary forms, \emph{Duke Math. J.} 15.\textbf{2} (1948), 443–453,
+G. B. Huff, Diophantine problems in geometry and elliptic ternary forms, \emph{Duke Math. J.} 15.\textbf{2} (1948), 443--453,
 \textsc{doi}: \href {https://doi.org/10.1215/S0012-7094-48-01543-9} {\nolinkurl {10.1215/S0012-7094-48-01543-9}}.
@@
 T. Kreisel and S. Kurz, There are integral heptagons, no three points on a line, no four on a circle, \emph{Discrete Comput.
-Geom.} 39.\textbf{4} (2008), 786–790, \textsc{doi}: \href {https://doi.org/10.1007/s00454-007-9038-6} {\nolinkurl
+Geom.} 39.\textbf{4} (2008), 786--790, \textsc{doi}: \href {https://doi.org/10.1007/s00454-007-9038-6} {\nolinkurl
 {10.1007/s00454-007-9038-6}}.
@@
 \bibitem{kurz2005characteristic}
-S. Kurz, On the characteristic of integral point sets in {$\mathbb {E}^m$}, \emph{Australas. J. Combin.} 36 (2006), 241–248,
+S. Kurz, On the characteristic of integral point sets in {$\mathbb {E}^m$}, \emph{Australas. J. Combin.} 36 (2006), 241--248,
 arXiv: \href {http://arxiv.org/abs/math/0511704} {\nolinkurl {math/0511704}}.
@@
 \bibitem{kurz2008bounds}
-S. Kurz and R. Laue, Bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 39 (2007), 233–240,
+S. Kurz and R. Laue, Bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 39 (2007), 233--240,
 arXiv: \href {http://arxiv.org/abs/0804.1296} {\nolinkurl {0804.1296}}.
@@
 \bibitem{kurz2008minimum}
-S. Kurz and A. Wassermann, On the minimum diameter of plane integral point sets, \emph{Ars Combin.} 101 (2011), 265–287,
+S. Kurz and A. Wassermann, On the minimum diameter of plane integral point sets, \emph{Ars Combin.} 101 (2011), 265--287,
 arXiv: \href {http://arxiv.org/abs/0804.1307} {\nolinkurl {0804.1307}}.
@@
 \bibitem{kurz2013constructing}
-S. Kurz et al., Constructing $7$-clusters, \emph{Serdica J. Comput.} 8.\textbf{1} (2014), 47–70, arXiv: \href
+S. Kurz et al., Constructing $7$-clusters, \emph{Serdica J. Comput.} 8.\textbf{1} (2014), 47--70, arXiv: \href
 {http://arxiv.org/abs/1312.2318} {\nolinkurl {1312.2318}}.
@@
 M. C. Markót and T. Csendes, A new verified optimization technique for the "packing circles in a unit square" problems,
-\emph{SIAM J. Optim.} 16.\textbf{1} (2005), 193–219.
+\emph{SIAM J. Optim.} 16.\textbf{1} (2005), 193--219.
 %
@@
 E. A. Momot, R. E. Zvolinskiy, and A. E. Zvolinskiy, An example of rails integral point set with cardinality 106, Прикладная
-математика и фундаментальная информатика, 2022, 114–115.
+математика и фундаментальная информатика, 2022, 114--115.
 %
 \bibitem{nozaki2013lower}
-H. Nozaki, Lower bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 56 (2013), 139–143.
+H. Nozaki, Lower bounds for the minimum diameter of integral point sets, \emph{Australas. J. Combin.} 56 (2013), 139--143.
 %
 \bibitem{solymosi2003note}
-J. Solymosi, Note on integral distances, \emph{Discrete Comput. Geom.} 30.\textbf{2} (2003), 337–342, \textsc{doi}: \href
+J. Solymosi, Note on integral distances, \emph{Discrete Comput. Geom.} 30.\textbf{2} (2003), 337--342, \textsc{doi}: \href
 {https://doi.org/10.1007/s00454-003-0014-7} {\nolinkurl {10.1007/s00454-003-0014-7}}.
@@
 \bibitem{solymosi2010question}
-J. Solymosi and F. De Zeeuw, On a question of Erdős and Ulam, \emph{Discrete Comput. Geom.} 43.\textbf{2} (2010), 393–401,
+J. Solymosi and F. De Zeeuw, On a question of Erdős and Ulam, \emph{Discrete Comput. Geom.} 43.\textbf{2} (2010), 393--401,
 arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl {0806.3095}}.
@@
 математического образования: материалы международной молодежной научной школы «Актуальные направления
-математического анализа и смежные вопросы»} 8 (2018), 5–6.%
+математического анализа и смежные вопросы»} 8 (2018), 5--6.%
 %
@@
 информатики и механики - сборник трудов Международной научной конференции. Научно-исследовательские публикации,
-2018, 492–498.
+2018, 492--498.
 %
 \bibitem{smurov1998stripcoverings}
-М. Смуров and А. Спивак, Покрытия полосками, \emph{Квант} \textbf{5} (1998), 6–12.
+М. Смуров and А. Спивак, Покрытия полосками, \emph{Квант} \textbf{5} (1998), 6--12.
 \end{thebibliography}

### 2025-11-28 char-4k-bound/char-4k-bound.tex
[4k3] Bibliography - Erdos
[4k3] Bibliography - VSU paper format
[4k3] Bibliography - translate the refs titles as required

@@
 \bibitem{anning1945integral}
-N. H. Anning and P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{8} (1945), 598--600, \textsc{doi}: \href
+N. H. Anning and P. Erd{\H o}s, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{8} (1945), 598--600, \textsc{doi}: \href
 {https://doi.org/10.1090/S0002-9904-1945-08407-9} {\nolinkurl {10.1090/S0002-9904-1945-08407-9}}.
@@
 \bibitem{avdeev2019particular}
-N. N. Avdeev, R. E. Zvolinsky, and E. A. Momot, On particular diameter bounds for integral point sets in higher dimensions
-Proceedings of Voronezh State University. Series: Physics. Mathematics. 2025. No 1. Pp. 62-77.
+N. N. Avdeev, R. E. Zvolinsky, and E. A. Momot, On particular diameter bounds for integral point sets in higher dimensions,
+\it{Proceedings of Voronezh State University. Series: Physics. Mathematics}. \textbf{1} (2025), 62--77.
 arXiv: \href {http://arxiv.org/abs/1909.10386} {\nolinkurl {1909.10386}}.
@@
 \bibitem{erdos1945integral}
-P. Erdős, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{12} (1945), 996, \textsc{doi}: \href
+P. Erd{\H o}s, Integral distances, \emph{Bull. Amer. Math. Soc.} 51.\textbf{12} (1945), 996, \textsc{doi}: \href
 {https://doi.org/10.1090/S0002-9904-1945-08490-0} {\nolinkurl {10.1090/S0002-9904-1945-08490-0}}.
@@
 \bibitem{kemnitz1988punktmengen}
-A. Kemnitz, Punktmengen mit ganzzahligen Abständen, 1988.
+A. Kemnitz, Point sets with integer distances (Punktmengen mit ganzzahligen Abständen), 1988.
 %
@@
 \bibitem{solymosi2010question}
-J. Solymosi and F. De Zeeuw, On a question of Erdős and Ulam, \emph{Discrete Comput. Geom.} 43.\textbf{2} (2010), 393--401,
+J. Solymosi and F. De Zeeuw, On a question of Erd{\H o}s and Ulam, \emph{Discrete Comput. Geom.} 43.\textbf{2} (2010), 393--401,
 arXiv: \href {http://arxiv.org/abs/0806.3095} {\nolinkurl {0806.3095}}.
@@
 \bibitem{our-ped-2018}
-Н. Н. Авдеев, On integral point sets in special position, \emph{Некоторые вопросы анализа, алгебры, геометрии и
-математического образования: материалы международной молодежной научной школы «Актуальные направления
-математического анализа и смежные вопросы»} 8 (2018), 5--6.%
+N. N. Avdeev, On integral point sets in special position,
+\emph{Nekotorye voprosy analiza, algebry, geometrii i matematicheskogo obrazovaniya:
+materialy mezhdunarodnoi molodezhnoi nauchnoi shkoly «Aktual'nye napravleniya matematicheskogo analiza i smezhnye voprosy»}
+8 (2018), 5--6.%
 %
 \bibitem{our-pmm-2018}
-Н. Н. Авдеев, Об отыскании целоудалённых множеств специального вида, Актуальные проблемы прикладной математики,
-информатики и механики - сборник трудов Международной научной конференции. Научно-исследовательские публикации,
+N. N. Avdeev,
+On the search of special integral point sets
+(Об отыскании целоудалённых множеств специального вида),
+Aktual'nye problemy prikladnoj matematiki, informatiki i mekhaniki --- sbornik trudov Mezhdunarodnoj nauchnoj konferentsii,
 2018, 492--498.
@@
 \bibitem{smurov1998stripcoverings}
-М. Смуров and А. Спивак, Покрытия полосками, \emph{Квант} \textbf{5} (1998), 6--12.
+M. Smurov and A. Spivak, Covering by strips (Покрытия полосками), \emph{Kvant} \textbf{5} (1998), 6--12.
 \end{thebibliography}
